### 2020-01-14 ising/ising.tex
added ising data.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}
\usepackage{braket}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\begin{document}

\title{Flat histogram method comparison on 2D Ising Model
}

\author{Jordan K. Pommerenck} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  We examine the convergence of the Stochastic Approximation with a Dynamic
  update factor (SAD) algorithm applied to the 2D ising model. Comparison with
  SAMC and WL methods show that SAD performs robustly and without user input
  knowledge of an energy range. We confirm that SAD is more powerful in the
  common case in which the range of energies is not known in advance.
\end{abstract}

\maketitle

\section{Introduction}
Flat histogram Monte-Carlo simulation algorithms calculate the thermodynamic
properties of various systems over a range of temperatures.  The first histogram
method used a single canonical Monte Carlo simulation to predict properties for
nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
samples a narrow energy range given limited by a narrow temperature range, it
proves computationally inefficient at sampling large energy ranges.

Due to a canonical Monte Carlo simulation's inefficient sampling of large
energies, a variety of ``flat'' (or ``broad'') histogram
methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
wang2001determining, wang2001efficient} were developed to explore a wider range
of energies.  In addition to obtaining thermodynamic information for the entire
energy range for a single simulation, these approaches cannot be easily trapped
in a local energy minimum like a canonical simulation.

Wang and Landau introduced one of the most widely used flat-histogram
Monte-Carlo algorithms that accurately determined the density of states (DOS) for
a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
its power, the method unfortunately requires a priori knowledge of several
user-defined parameters. Thus, for any given system under study, the user needs
to determine the ideal parameters in order to apply the method. The Wang-Landau
algorithm is also known to violate detailed balance (although only for brief
time intervals)~\cite{yan2003fast, shell2002generalization}. With the violation
of detailed balance, convergence of the algorithm is not guaranteed.

Because of the uncertainty of convergence for WL, many studies were undertaken
to understand how the modification (or update) factor $\gamma$ impacts the
convergence~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. The ideal choice of decreasing the modification factor
helps to prevent CPU time from being wasted by continuing to perform
calculations after the error in the density of states is
saturated~\cite{belardinelli2008analysis}. Modification factors that decrease
faster than $1/t$ were found to lead to
non-convergence~\cite{belardinelli2007fast}.  These findings led to the
formation of the $1/t$-WL algorithm.

Liang independently considered whether WL could be treated as a special case of
Stochastic Approximation whose convergence could be mathematically
proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
al.~\cite{liang2007stochastic} argued that WL can be considered a form of
Stochastic Approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
convergence (if certain conditions are met). Despite the added benefit of
guaranteed convergence, the method still has a system specific user-defined
variable. Such variables often create difficulty when applying Monte-Carlo
methods across arbitrary systems.

Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
related Statistical Temperature Molecular Dynamics (STMD), an adaption
of the WL method that approximates the entropy as a piecewise linear function,
which improves convergence for systems with a continuously varying
energy~\cite{kim2006statistical, kim2007statistical}. STMC applied to WL
requires a temperature range be specified rather than an energy range.  Kim
\emph{et al.} extended this work as Replica Exchange Statistical Temperature
Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping STMC
simulations to improve convergence~\cite{kim2009replica}. Recently,
Junghans \emph{et al.} demonstrated a close connection between
metadynamics, which was introduced by Laio and
Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods, with STMD
forging the connection~\cite{junghans2014molecular}.

The SAD (Stochastic Approximation with
Dynamic $\gamma$) method outlined by Pommerenck
et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
that dynamically chooses the modification factor rather than relying on system
dependent parameters. SAD shares the same convergence properties with SAMC while
replacing un-physical user-defined parameters with the algorithms dynamic
choice.

In this work, we apply the family of weight-based flat histogram Monte
Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.

\section{Ising Model}
The 2D Ising spin-lattice system is widely used as a testbed when
benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The 2nd order
phase transition behavior and the ability to directly calculate the
exact solution for finite lattices~\cite{beale1996exact} make the
system sufficiently interesting for such theoretical comparisons. It is
also important to note that direct comparison of the other methods can
be made with WL as its original implementation was done on this
system~\cite{wang2001determining,wang2001efficient}.
We test the convergence of several flat histogram methods
on the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new}.
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the absence of a magnetic field ($h =
0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over nearest neighbor spin sites. Beale showed that for finite
lattices the Density of States could directly be calculated from the partition
function~\cite{beale1996exact}
\begin{align}
Z = \sum_E g(E) e^{-{\beta E}}
\end{align}
where $g(E)$ is the multiplicity of the system which is proportional to
$D(E)$. We can estimate the average deviation from the exact solution
using the relation~\cite{schneider2017convergence, shakirov2018convergence,
barash2017control}:
\begin{align}
\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
\end{align}
As per the implementation by Wang and
Landau~\cite{wang2001determining}, the total number of available energy
states are $N-1$.

\section{Flat histogram methods}\label{sec:histogram}
Flat histogram methods determine the density of states $D(E)$ over a broad range
of energies by simulating each energy with equivalent accuracy. Flat histogram
Monte Carlo methods propose randomly chosen ``moves'' which change the state of
the system and must satisfy detailed balance.  Each algorithm differs in how it
determines the probability of accepting a move and in what additional statistics
must be collected in order decide on that probability.

We describe four closely related flat histogram methods which each rely on a
weight function $w(E)$ to determine $D(E)$.  For these algorithms, the
probability of accepting a move is given by
\begin{equation}
	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
\end{equation}
which biases the simulation in favor of energies with low weights. The result of
weights $w(E)$ that are proportional to $D(E)$ is an entirely flat histogram. We
can relate the entropy to the weights in the microcanonical ensemble, since the
entropy is defined as $S(E) \equiv k_B\ln(D(E)) \approx \ln w(E)$.

Each approach uses a random walk in energy space to estimate $D(E)$.  At the
core of each of these approaches is to continuously update the weights at each
step of the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma_t
\end{equation}
where $t$ is number of the current move, $\gamma(t)$ is a move-dependent update
factor, and $E$ is the current energy.  This update causes the random walk to
avoid energies that have been frequently sampled, leading to a rapid exploration
of energy space. The four methods differ primarily in how they schedule the
decreasing of $\gamma_t$.

The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
landau2014guide} explores energy space by initially setting $\gamma=1$, and then
decreases $\gamma$ in discrete stages. An energy range of interest must be
specified~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
which can require multiple simulations if unknown.

The number (``counts'') of moves ending at each energy are stored in a
histogram.  For a sufficiently flat energy histogram (specified by the user to
typically be 0.8), $\gamma$ is decreased by a specified factor of $\frac12$ and
the histogram is reset to zero. The entire process is repeated until $\gamma$
reaches a desired cutoff.

Zhou et. al also showed that the WL algorithm never converges
exponentially~\cite{zhou2008optimal} and succesfully bounded the statistical
error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}. He also suggests
that a modification factor of $1/t$ is optimal for WL.

The $1/t$-WL method is a refinement of the original implementation that
addresses the error saturation of WL. Belardinelli and Pereyra
implemented a schedule~\cite{belardinelli2007fast} that enforced that
if $\gamma < \nicefrac{1}{t}$ then the update factor is set to
$\nicefrac{1}{t}$ and the histogram is no longer tracked. Using this
metric, the error saturation is avoided since the correct density of
states is approached asymptotically as
$t^{-\frac12}$~\cite{belardinelli2008analysis}.
Schneider $\text{et. al}$ outline minor refinements to the $1/t$-WL algorithm
including $\nicefrac{N_S}{t}$ scaling and switching from standard WL to
$1/t$-WL when the update factor $\gamma <
\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the original
$1/t$ implementation~\cite{belardinelli2007fast}, the update factor is
decreased once all $N_S$ energy states are visited at least once (i.e.
$H(E) \neq 0$) effectively avoiding the concept of `flatness'.

Another weight-based flat histogram method is
the Stochastic Approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
schedule by which the update factor $\gamma$ is continuously
decreased~\cite{liang2007stochastic, werlich2015stochastic,
schneider2017convergence}.  The update factor is defined in the
original implementation~\cite{liang2007stochastic} in terms of an
arbitrary tunable parameter $t_0$,
\begin{align}
\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
\end{align}
where as above $t$ is the number of moves that have been attempted.

SAMC offers extreme simplicity and proven convergence.
Provided the update factor satisfies
\begin{align}
\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
\end{align}
where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
to converge to the true density of states~\cite{liang2006theory,
liang2007stochastic}.  Unlike WL methods, the energy range need not be
known \emph{a priori.}  The convergenc time depends only on the choice of
parameter $t_0$.
The parameter $t_0$ can unfortunately be difficult to chose in advance
for arbitrary systems.
Liang \emph{et al.} give a rule of thumb in
which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
is the number of energy bins~\cite{liang2007stochastic}.  Schneider
\emph{et al.} found that for the Ising model this heuristic is helpful
for small spin systems, but that larger systems require an even higher
$t_0$ value~\cite{schneider2017convergence}. Our work also confirms this.

Pommerenck et al. propose a refinement~\cite{pommerenck2020stochastic} to SAMC
where the update factor is determined dynamically rather than by the user.
Stochastic Approximation with a dynamic $\gamma$ (SAD) requires the user to
provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
requiring a priori an energy range of interest; however, this may easier to
identify and is more physical than the SAMC parameter $t_0$. The update factor
can be written in terms of the current estimates for the highest $E_H$ and
lowest $E_L$ energies of interest and the last time that an energy in the range
of interest is encountered $t_L$.
\begin{align}
  \gamma_{t}^{\text{SAD}} =
     \frac{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
     }{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
     }
\end{align}
SAD only explores the energy range of interest as specified by the minimum
temperature of interest $T_{\min} < T < \infty$. During the simulation the two
energies $E_H$ and $E_L$, are refined such that the range of energies are conservatively
estimated. The weights are calculated for each energy region according to the original
prescription.
\begin{enumerate}
\item {$E < E_L$:} $w(E>E_H) = w(E_H)$
\item {$E_L < E < E_H$:} moves are handled the same as other weight-based
methods that are mentioned
\item {$E > E_H$:} $w(E<E_L) = w(E_L)e^{-\frac{E_L-E}{T_{\min}}}$
\end{enumerate}
Each time the simulation changes the value of $E_H$ or $E_L$, the weights
within the new portion of the interesting energy range are updated.

\section{Results}
We tested the algorithms on two 2D Ising model systems.  The first is a smaller
simulation with a lattic size of 32 and the second has a lattice size of 128.
For each system we use a reasonable root-mean-square displacement distance
$\delta_0 = 0.05\sigma$ for proposed moves. The simulations explore the energy
space of the systems with minimum reduced temperatures of $T_{\text{min}} = 1$
for simulations All simulations lead to the minimum important energy $E_{\min}$
and maximum entropy energy $E_{\max}$ being calculated (with the exception of
the WL methods where both of these parameters are needed a priori).

\begin{figure}
\includegraphics[width=\columnwidth]{figs/ising/gamma-n32.pdf}
  \caption{
  The update factor $\gamma_t$ versus the iteration number for the $N=32 \times 32$
  system.}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{figs/ising/N32-entropy-error-default.pdf}
  \caption{
  The average entropy error for each MC method for $N=32 \times 32$,
               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
               as a function of number of iterations run.  The error is
               averaged over 8 independent simulations, and the best
               and worst simulations for each method are shown as a
               semi-transparent shaded area.}\label{fig:n32}
\end{figure}

\subsection{The 32 $\times$ 32 Ising model}
For this simulation, we chose a minimum reduced temperature of $1$, which
corresponds to an interesting energy range from $-2048$ to $48$.  The number of
important energy states for this system is $N_S = 529$.

Fig.~\ref{fig:n32} shows the average error in the entropy as a function of time for
this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed lines
represent the average of the absolute value of the error in the entropy averaged
over eight simulations using different random number seeds. The range of average
errors for each simulation is shown as a shaded region around its mean error. By
the time $10^8$ moves have been made all but the WL simulation have begun to
converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^9$ moves.

\begin{figure}
\includegraphics[width=\columnwidth]{figs/ising/gamma-n128.pdf}
  \caption{
  The update factor $\gamma_t$ versus the iteration number for the $N=128 \times 128$
  system.}
\end{figure}

\begin{figure}
\includegraphics[width=\columnwidth]{figs/ising/N128-entropy-error-default.pdf}
  \caption{
  The average entropy error for each MC method for $N=128 \times 128$,
               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
               as a function of number of iterations run.  The error is
               averaged over 8 independent simulations, and the best
               and worst simulations for each method are shown as a
               semi-transparent shaded ar.}\label{fig:n128}
\end{figure}

\subsection{128 $\times$ 128 Ising system}
For this simulation, we chose a minimum reduced temperature of $1$, which
corresponds to an interesting energy range from $-32768$ to $100$.  The number of
important energy states for this system is $N_S = 8214$.
Fig.~\ref{fig:n128} shows the average error in the entropy as a function of time
for this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed
lines represent the average of the absolute value of the error in the entropy
averaged over eight simulations using different random number seeds. The range
of average errors for each simulation is shown as a shaded region around its
mean error. By the time $10^{12}$ moves have been made all but the WL and SAD
simulations have begun to converge as $1/\sqrt{t}$. We then see the WL error
saturate around $10^{12}$ moves. The error in the SAD method will continue to
converge faster than $1/\sqrt{t}$ until the update factor approaches that of
1/t-WL at which point both methods will converge equivalently.

\section{Conclusion}
We have examined the convergence of SAD versus other widely used weight-based
histogram methods.  We have shown that Sad samples the energy space
corresponding to a desired range of temperatures for a variety of system sizes.
We find that both SAD and 1/t-WL demonstrate excellent and robust convergence.
They both converge more rapidly than SAMC, and unlike WL do not suffer from
error saturation. We find that for larger systems SAD reduces the updated factor
more slowly (and conservatively) than 1/t-WL. This means that SAD will take
proportionately more moves to converge to the same value as 1/t-WL as system
size is increased. SAD requires the user to specify a temperature range of
interest rather than an energy range of interest as 1/t-WL does. For use cases
in which a range of desired temperatures is known, this will make the SAD method
considerably more convenient.

\bibliography{ising} % Produces the bibliography via BibTeX.

\end{document}

### 2020-01-16 figs/colors.py
added the data and figure scripts.

@@ -1,0 +1,281 @@
+import matplotlib.pyplot as plt
+
+_colors = { 'SAD': 'tab:orange',
+            r'SAD $\Delta E=0.001\epsilon$': 'tab:orange',
+            r'SAD $\Delta E=0.01\epsilon$': 'tab:orange',
+            r'SAD $\Delta E=0.1\epsilon$': 'tab:orange',
+
+            'WL': 'tab:green',
+            r'WL $E_{\min}=-133.53\epsilon$': 'tab:green',
+
+            '$1/t$-WL': 'xkcd:purple',# 'tab:blue',
+
+            r'$1/t$-WL $E_{\min}=-133.58\epsilon$': 'xkcd:purple',
+            r'$1/t$-WL $E_{\min}=-133.54\epsilon$': 'xkcd:purple',
+            r'$1/t$-WL $E_{\min}=-133.53\epsilon$': 'xkcd:purple',
+            r'$1/t$-WL $E_{\min}=-133.52\epsilon$': 'xkcd:purple',
+
+            'SAMC ($t_0 =10^{3}$)': 'r',
+            'SAMC ($t_0 =10^{4}$)': 'xkcd:orange',
+            'SAMC ($t_0 =10^{5}$)': 'xkcd:gold',
+            'SAMC ($t_0 =10^{6}$)': 'g',
+            'SAMC ($t_0 =10^{7}$)': 'tab:blue', # 'midnightblue',
+            'SAMC ($t_0 =10^{8}$)': 'darkmagenta',
+            'SAMC ($t_0 =10^{9}$)': 'xkcd:maroon',
+            '1/sqrt(t)': 'xkcd:dark gray',
+            'converged result': 'xkcd:black',
+            r'$\frac{1}{\sqrt{t}}$': '#eeeeee',
+}
+
+_linestyles = {
+    'SAMC ($t_0 =10^{3}$)': ':',
+    'SAMC ($t_0 =10^{4}$)': ':',
+    'SAMC ($t_0 =10^{5}$)': ':',
+    'SAMC ($t_0 =10^{6}$)': ':',
+    'SAMC ($t_0 =10^{7}$)': ':',
+    'SAMC ($t_0 =10^{8}$)': ':',
+    'SAMC ($t_0 =10^{9}$)': ':',
+    'WL': '--',
+    r'WL $E_{\min}=-133.53\epsilon$': '--',
+
+    '$1/t$-WL': '-.',
[…]

### 2020-01-16 figs/comparison-plot.py

@@ -1,0 +1,249 @@
+from __future__ import division, print_function
+import sys, os, matplotlib
+import numpy as np
+from collections import OrderedDict
+
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+matplotlib.rcParams['figure.figsize'] = (5, 4)
+
+if 'noshow' in sys.argv:
+        matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+from glob import glob
+import colors
+
+if os.path.exists('../data'):
+    os.chdir('..')
+
+energy = int(sys.argv[1])
+filebase = sys.argv[2]
+transcale = sys.argv[3]
+save_dir = sys.argv[4]
+
+tex_filebase = filebase.replace('.','_') # latex objects to extra "." characters
+
+methods = ['-sad3', '-sad3-s1', '-sad3-s2',
+            '-tmmc', '-tmi', '-tmi2', '-tmi3', '-toe', '-toe2', '-toe3',
+            '-vanilla_wang_landau']
+if 'allmethods' not in sys.argv:
+    methods = ['-sad3','-wl-50','-wl-inv-t-50','-wl-inv-t-256','-sad-50',
+               '-sad-256','-wl-256','-ising-sad-32','-ising-wl-32','-ising-wl-inv-t-32',
+               '-ising-sad-128','-ising-wl-128','-ising-wl-inv-t-128'] #'-ising-sad-128-T15' This file has a lot of data
+    # methods = [m+s for m in methods for s in '', '-s1', '-s2', '-s3', '-s4']
+    if transcale == 'slow':
+        methods = ['-sad3-slow','-sad-slow-T13','-wl-50-slow','-wl-inv-t-50-slow','-sad-50-slow',
+                   '-tmmc-slow', '-vanilla_wang_landau-slow','-vanilla_wang_landau-T13-slow','-sad3-T13-slow']
+    if transcale == 'fast':
+        methods = ['-sad3-fast','-tmmc-fast', '-vanilla_wang_landau-fast']
+
+# For WLTMMC compatibility with LVMC
+lvextra = glob('data/comparison/%s-wltmmc*' % filebase)
+split1 = [i.split('%s-'%filebase, 1)[-1] for i in lvextra]
+split2 = [i.split('-m', 1)[0] for i in split1]
+for meth in split2:
+    if "default" in transcale:
+        if meth[-3:] != '-tm' and "slow" not in meth and "fast" not in meth:
[…]

### 2020-01-16 figs/gamma-plot.py

@@ -1,0 +1,93 @@
+from __future__ import division
+import numpy as np
+import matplotlib.pyplot as plt
+import sys, glob, matplotlib
+import colors
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+
+filename = sys.argv[1]
+Tmin = float(sys.argv[2])
+
+plt.figure(figsize=(5, 4))
+
+try:
+    for wl in glob.glob("data/gamma/%s/wl*.txt" % filename):
+        print("in wl")
+        print(('vanilla_wang_landau'+ wl[len("data/gamma/%s/wl" % filename):-4]))
+        wlmoves, wlfactor = np.loadtxt(wl, dtype = float, unpack = True)
+        data = np.loadtxt(wl)
+        moves = data[:, 0]
+        factor = data[:, 1]
+        if (data[0, 0] == 'wl_factor'): # using c++ data!
+            moves = np.zeros(len(wlmoves)*2+2)
+            factor = np.zeros_like(moves)
+            factor[0] = 1
+            moves[0] = 1
+            for i in range(len(wlmoves)):
+                moves[2*i+1] = wlmoves[i]
+                moves[2*i+2] = wlmoves[i]
+                factor[2*i+1] = wlfactor[i]*2
+                factor[2*i+2] = wlfactor[i]
+        colors.loglog(moves, factor,
+                      'wl'
+                         + wl[len("data/gamma/%s/wl" % filename):-4])
+        plt.ylim(ymin=1e-10, ymax=1e1)
+        plt.xlim(xmin=1e3, xmax=1e12)
+
+except:
+    pass
+
+
+try:
+    for sad in glob.glob("data/gamma/%s/sad*.dat" % filename):
+        data = np.loadtxt(sad)
+        print((data.shape))
+        if data.shape[1] > 2: # i.e. we are not using parse-yaml-out.py
+            num_sad_states = data[:, 0]
+            time = data[:, 1]
+            ehi = data[:, 2]
+            elo = data[:, 3]
+            ts = np.exp(np.linspace(0, np.log(max(time)*1e4), 2000))
+            gamma = np.zeros_like(ts)
+            print((sad, time))
+            for j in range(len(time)):
+                for i in range(len(gamma)):
+                    if ts[i] > time[j]:
+                        t = ts[i]
[…]

### 2020-01-16 figs/ising-gamma-plot.py

@@ -1,0 +1,98 @@
+from __future__ import division
+import numpy as np
+import matplotlib.pyplot as plt
+import sys, glob, matplotlib
+import colors
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+
+filename = sys.argv[1]
+Tmin = float(sys.argv[2])
+
+plt.figure(figsize=(5,4))
+
+try:
+    for wl in glob.glob("data/gamma/ising/%s/*wl*.txt" % filename):
+        print("in wl")
+        print('vanilla_wang_landau'+ '-is' + wl[len("data/gamma/ising/%s/wl" % filename):-4])
+        wlmoves, wlfactor = np.loadtxt(wl, dtype = float, unpack = True)
+        data = np.loadtxt(wl)
+        moves = data[:,0]
+        factor = data[:,1]
+        if (data[0,0] == 'wl_factor'): # using c++ data!
+            moves = np.zeros(len(wlmoves)*2+2)
+            factor = np.zeros_like(moves)
+            factor[0] = 1
+            moves[0] = 1
+            for i in range(len(wlmoves)):
+                moves[2*i+1] = wlmoves[i]
+                moves[2*i+2] = wlmoves[i]
+                factor[2*i+1] = wlfactor[i]*2
+                factor[2*i+2] = wlfactor[i]
+        colors.loglog(moves, factor,
+                      'is'
+                         + wl[len("data/gamma/ising/%s/wl" % filename):-4])
+        #plt.ylim(1e-10, 1e1)
+        #plt.xlim(1e3, 1e13)
+
+except:
+    pass
+
+try:
+    for sad in glob.glob("data/gamma/ising/%s/*sad*.dat" % filename):
+        data = np.loadtxt(sad)
+        print(data.shape)
+        if data.shape[1] > 2: # i.e. we are not using parse-yaml-out.py
+            num_sad_states = data[:,0]
+            time = data[:,1]
+            ehi = data[:,2]
+            elo = data[:,3]
+            ts = np.exp(np.linspace(0, np.log(max(time)*1e4), 2000))
+            gamma = np.zeros_like(ts)
+            print(sad, time)
+            for j in range(len(time)):
+                for i in range(len(gamma)):
+                    if ts[i] > time[j]:
[…]

### 2020-01-16 figs/comparison-plot.py
fixed paper ising where to look for figs.

@@ -115,13 +115,6 @@ for method in methods:
                 iterations = np.array(temp_iter)
                 errorinentropy = np.array(temp_ee)
                 maxerror = np.array(temp_me)
-        if not os.path.exists('figs/lv'):
-                os.makedirs('figs/lv')
-
-        if not os.path.exists('figs/s000'):
-                os.makedirs('figs/s000')
-
-        os.makedirs('figs/ising', exist_ok=True)
 
         if filebase.startswith('lv'):
                 NxNsplit = filebase.split('-')

### 2020-01-16 ising/ising.tex
fixed paper ising where to look for figs.
fixed where fac was looking for bib file.

@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/gamma-n32.pdf}
+\includegraphics[width=\columnwidth]{ising/gamma-n32.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/N32-entropy-error-default.pdf}
+\includegraphics[width=\columnwidth]{ising/N32-entropy-error-default.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/gamma-n128.pdf}
+\includegraphics[width=\columnwidth]{ising/gamma-n128.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/N128-entropy-error-default.pdf}
+\includegraphics[width=\columnwidth]{ising/N128-entropy-error-default.pdf}
   \caption{
@@
 
-\bibliography{ising} % Produces the bibliography via BibTeX.
+\bibliography{ising/ising} % Produces the bibliography via BibTeX.
 

### 2020-01-20 figs/ising-gamma-plot.py
fixing up ising directory.

@@ -12,9 +12,9 @@ Tmin = float(sys.argv[2])
 plt.figure(figsize=(5,4))
 
 try:
-    for wl in glob.glob("data/gamma/ising/%s/*wl*.txt" % filename):
+    for wl in glob.glob("../data/gamma/ising/%s/*wl*.txt" % filename):
         print("in wl")
-        print('vanilla_wang_landau'+ '-is' + wl[len("data/gamma/ising/%s/wl" % filename):-4])
+        print('vanilla_wang_landau'+ '-is' + wl[len("../data/gamma/ising/%s/wl" % filename):-4])
         wlmoves, wlfactor = np.loadtxt(wl, dtype = float, unpack = True)
         data = np.loadtxt(wl)
         moves = data[:,0]
@@ -31,7 +31,7 @@ try:
                 factor[2*i+2] = wlfactor[i]
         colors.loglog(moves, factor,
                       'is'
-                         + wl[len("data/gamma/ising/%s/wl" % filename):-4])
+                         + wl[len("../data/gamma/ising/%s/wl" % filename):-4])
         #plt.ylim(1e-10, 1e1)
         #plt.xlim(1e3, 1e13)
 
@@ -39,7 +39,7 @@ except:
     pass
 
 try:
-    for sad in glob.glob("data/gamma/ising/%s/*sad*.dat" % filename):
+    for sad in glob.glob("../data/gamma/ising/%s/*sad*.dat" % filename):
         data = np.loadtxt(sad)
         print(data.shape)
         if data.shape[1] > 2: # i.e. we are not using parse-yaml-out.py
@@ -93,6 +93,6 @@ plt.tight_layout()
 plt.ylim(1e-10, 1e1)
 plt.xlim(1e3, 1e13)
 
-plt.savefig('ising/gamma-%s.pdf' % filename.replace('.','_'))
+plt.savefig('gamma-%s.pdf' % filename.replace('.','_'))
 if 'noshow' not in sys.argv:
     plt.show()

### 2020-01-20 scripts-python/ising-gamma-plot.py
fix ising-gamma to look in right directory.

@@ -12,9 +12,9 @@ Tmin = float(sys.argv[2])
 plt.figure(figsize=(5,4))
 
 try:
-    for wl in glob.glob("../data/gamma/ising/%s/*wl*.txt" % filename):
+    for wl in glob.glob("../ising/data/gamma/%s/*wl*.txt" % filename):
         print("in wl")
-        print('vanilla_wang_landau'+ '-is' + wl[len("../data/gamma/ising/%s/wl" % filename):-4])
+        print('vanilla_wang_landau'+ '-is' + wl[len("../ising/data/gamma/%s/wl" % filename):-4])
         wlmoves, wlfactor = np.loadtxt(wl, dtype = float, unpack = True)
         data = np.loadtxt(wl)
         moves = data[:,0]
@@ -31,7 +31,7 @@ try:
                 factor[2*i+2] = wlfactor[i]
         colors.loglog(moves, factor,
                       'is'
-                         + wl[len("../data/gamma/ising/%s/wl" % filename):-4])
+                         + wl[len("../ising/data/gamma/%s/wl" % filename):-4])
         #plt.ylim(1e-10, 1e1)
         #plt.xlim(1e3, 1e13)
 
@@ -39,7 +39,7 @@ except:
     pass
 
 try:
-    for sad in glob.glob("../data/gamma/ising/%s/*sad*.dat" % filename):
+    for sad in glob.glob("../ising/data/gamma/%s/*sad*.dat" % filename):
         data = np.loadtxt(sad)
         print(data.shape)
         if data.shape[1] > 2: # i.e. we are not using parse-yaml-out.py

### 2020-01-20 scripts-python/ising-gamma-plot.py
the python script now searches the right directory when built with fac.

@@ -11,6 +11,7 @@ Tmin = float(sys.argv[2])
 
 plt.figure(figsize=(5,4))
 
+
 try:
     for wl in glob.glob("../ising/data/gamma/%s/*wl*.txt" % filename):
         print("in wl")

### 2020-01-20 scripts-python/comparison-plot.py
moved ising data into ising directory.

@@ -13,7 +13,7 @@ import matplotlib.pyplot as plt
 from glob import glob
 import colors
 
-if os.path.exists('../data'):
+if os.path.exists('../ising/data'):
     os.chdir('..')
 
 energy = int(sys.argv[1])

### 2020-01-20 scripts-python/comparison-plot.py
comparison-plot now is able to handle the ising data.

@@ -13,13 +13,13 @@ import matplotlib.pyplot as plt
 from glob import glob
 import colors
 
-if os.path.exists('../ising/data'):
+if os.path.exists('ising/data'):
     os.chdir('..')
 
 energy = int(sys.argv[1])
 filebase = sys.argv[2]
 transcale = sys.argv[3]
-save_dir = sys.argv[4]
+system = sys.argv[4]
 
 tex_filebase = filebase.replace('.','_') # latex objects to extra "." characters
 
@@ -116,23 +116,16 @@ for method in methods:
                 errorinentropy = np.array(temp_ee)
                 maxerror = np.array(temp_me)
 
-        if filebase.startswith('lv'):
-                NxNsplit = filebase.split('-')
-                NxN = NxNsplit[-1].split('x')
-                # Formula to calculate N from title i.e. 100x10
-                # and use floor to always round up.
-                N = np.floor(0.25*0.20*float(NxN[0])*float(NxN[-1])*float(NxN[-1]))
-                ff = 1.0 # FIXME
-                moves = iterations * N
-        if filebase.startswith('s000'):
+        if system == 's000':
                 N = filebase.split('-N')[-1]
                 ff = filebase.split('-ff')[-1].split('-N')[0]
                 ff = float(ff)
                 # Get N directly from title.
                 moves = iterations * float(N)
-        if filebase.startswith('ising'):
+        if system == 'ising':
                 N = filebase.split('N')[-1]
                 moves = iterations * float(N)
+
         max_time = max(max_time, moves.max())
 
         if energy > 0:
@@ -204,10 +197,10 @@ for method in methods:
 plt.figure('maxerror')
 colors.loglog([1e6,max_time], [best_ever_max*np.sqrt(max_time/1e6), best_ever_max], method = r'$\frac{1}{\sqrt{t}}$')
 colors.legend()
-plt.savefig('%s/%s-max-entropy-error-%s.pdf' % (save_dir,tex_filebase,transcale))
+plt.savefig('%s-max-entropy-error-%s.pdf' % (tex_filebase,transcale))
 
 plt.figure('errorinentropy')
-if filebase == 'ising/N128':
[…]

### 2020-01-20 square-well/sad.tex
added square-well files and now working on inegrating with data.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{blue} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\begin{document}

\title{Stochastic Approximation Monte Carlo with a Dynamic Update
Factor
}

\author{Jordan K. Pommerenck} \author{Tanner T. Simpson}
\author{Michael A. Perlin} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  We present a new Monte Carlo algorithm based on the Stochastic
  Approximation Monte Carlo (SAMC) algorithm for directly calculating
  the density of states. The proposed method is Stochastic
  Approximation with a Dynamic update factor (SAD)
  which dynamically adjusts the update factor $\gamma_t$ during the course of
  the simulation. We test this method on the square-well fluid and
  the 31-atom Lennard-Jones cluster and
  compare the convergence behavior of several
  related
  Monte Carlo methods. We find that both the SAD and $1/t$-Wang-Landau ($1/t$-WL)
  methods rapidly converge to the
  correct density of states without the need for the user to specify an
  arbitrary tunable parameter $t_0$ as in the case of SAMC.  SAD requires
  as input the temperature range of interest, in contrast to
  $1/t$-WL, which requires that the user identify the interesting range
  of energies.
  %
  The convergence of the $1/t$-WL method is very sensitive to the energy
  range chosen for the low-temperature heat capacity of the
  Lennard-Jones cluster.
  %
  Thus, SAD is more powerful in the common case in which the range
  of energies is not known in advance.
\end{abstract}

\maketitle

\section{Introduction}
Over the past several decades, a number of flat histogram Monte Carlo simulation
algorithms have been developed which calculate the thermodynamic properties of
various systems over a range of temperatures.  This development began with the
original histogram method, which used a single canonical Monte Carlo simulation
to predict properties for nearby temperatures~\cite{ferrenberg1988new}.  For
large systems, this approach is limited to a narrow temperature range because a
single canonical simulation explores only a small range of energies. Berg and
Neuhaus developed multicanonical methods that introduced a weight function to
enable flat histogram sampling which improved the exploration of configuration
space and allowed the system to overcome free energy
barriers~\cite{berg1991multicanonical, berg1992multicanonical}. These works led
to an explosion in the interest of ``flat'' (or ``broad'') histogram
methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
wang2001determining, wang2001efficient, landau2004new, schulz2003avoiding,
yan2003fast, trebst2004optimizing, belardinelli2007wang, belardinelli2007fast,
belardinelli2008analysis, belardinelli2014intrinsic, singh2012density,
zhou2008optimal, schneider2017convergence, liang2006theory, liang2007stochastic,
liang2009improving, werlich2015stochastic, kim2006statistical,
kim2007statistical, kim2009replica, junghans2014molecular}, which explore a
wider range of energies. In addition to simulating a range of temperatures, in
contrast with low-temperature canonical Monte Carlo, these approaches
avoid being trapped in a local energy minimum.

Wang and Landau introduced the most widely used flat histogram algorithm (WL)
that uses an update factor and a statistical histogram to compute the density of
states of a given system~\cite{wang2001determining, wang2001efficient}.  While
the method is incredibly powerful, it has a few disadvantages. The most
significant disadvantage is that the method requires the user to select the
range of energies to be studied~\cite{landau2004new, wang2001efficient,
schulz2003avoiding, yan2003fast}. This requirement adds an additional
hurdle to its application to systems for which the interesting range of energies
is not known \emph{a priori}.  The simulation violates detailed balance, albeit
briefly as the size of the violation decreases with time, which complicates
convergence analysis.  In fact, the error in a WL computation has been
demonstrated to saturate at a non-zero value~\cite{yan2003fast}, i.e. the method
does not converge to the true density of states~\cite{belardinelli2007wang,
belardinelli2007fast, belardinelli2008analysis, belardinelli2014intrinsic,
singh2012density, zhou2008optimal}.

Belardinelli and Pereyra demonstrated that allowing the update factor to
decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang},
where $t$ corresponds to the number of moves.
This leads to their $1/t$-WL algorithm which ensures that the error continues to
decrease asymptotically as $1/\sqrt{t}$, which they demonstrated avoids
error saturation and asymptotically approaches the true density of
states~\cite{belardinelli2008analysis}. Zhou \emph{et al.} further confirmed that the
original WL algorithm never converges exponentially and successfully bounded the
statistical error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}.
Schneider \emph{et al.} outline minor refinements to the $1/t$-WL algorithm
including scaling the update factor with the number of energy
bins~\cite{schneider2017convergence}.

Liang began to consider Stochastic Approximation as a mathematical
generalization of the WL approach, with convergence that could be mathematically
proven~\cite{liang2006theory}. In 2007, Liang \emph{et al.}
developed Stochastic Approximation Monte Carlo (SAMC)~\cite{liang2007stochastic,
liang2009improving}, and proved its convergence, although the method still has a
system specific user-defined parameter which must be tuned when applying this
algorithm to a new system.  In contrast to the WL-based methods, the SAMC method
does not require users to determine the energy range of interest \emph{a
priori}. Werlich \emph{et al.} proposed the introduction of an additional tuning
parameter into SAMC~\cite{werlich2015stochastic}.

Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
related Statistical Temperature Molecular Dynamics (STMD) which is an
adaption of the WL method that approximates the entropy (or natural logarithm of the density of
states) as a piecewise linear function, which improves convergence for systems
with a continuously varying energy~\cite{kim2006statistical,
kim2007statistical}. STMC applied to WL requires a temperature range be
specified rather than an energy range.  Kim \emph{et al.} extended this work as
Replica Exchange Statistical Temperature Monte Carlo (RESTMC), which uses
replica exchange of multiple overlapping STMC simulations to improve
convergence~\cite{kim2009replica}. More recently, Junghans \emph{et al.} have
demonstrated a close connection between metadynamics, which was introduced by
Laio and Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods,
with STMD forging the connection~\cite{junghans2014molecular}.

In this work, we have developed an improved algorithm based on SAMC that does
not require an array of non-physical, user-defined inputs and therefore should
be easily applicable to any system. The method (like STMC above) does require
the user to define a temperature range of interest ($T_\text{min}$ to
$T_\infty$) which we explain in Section~\ref{sec:sad}.  We call this method SAD (Stochastic Approximation with a Dynamic
update factor), and will discuss it in detail in the methods section. We compare
its convergence properties with three existing flat histogram methods: WL,
$1/t$-WL, and SAMC.

In this work, we compare four flat histogram methods.  We outline the
general workings of each algorithm that we developed in detail while
summarizing algorithms that were developed in other works.  The
following methods are discussed and simulated for the square-well
fluid and the 31-atom Lennard-Jones cluster: Wang-Landau (WL),
$1/t$-Wang-Landau ($1/t$-WL), Stochastic Approximation Monte Carlo
(SAMC), and SAD.

\section{Flat histogram methods}\label{sec:histogram}
The goal of flat histogram methods (also called \emph{broad histogram}
or \emph{multicanonical} methods) is to simulate each energy with
similar accuracy so as to accurately determine the density of states
over a broad range of energies---and thus to determine the thermodynamic
quantities such as heat capacity or internal energy
over a broad range of temperatures.
Properties that require more information---such as a spatial
correlation function or a response function---can still be computed
for any temperature, provided statistics are collected for each
individual energy, which can then be reweighted for any
temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
errington2003direct}.

All the flat histogram Monte Carlo methods begin with randomly chosen
``moves'' which change the state of the system and must satisfy
detailed balance.  Each algorithm differs in how it determines the
probability of accepting a move and in what additional statistics must
be collected in order to decide on that probability.

Flat histogram methods calculate the density of states $D(E)$ for a discrete set
of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
trebst2004optimizing}. Therefore, energy binning becomes an important
consideration for systems with a continuum of possible energies.  Energy bins
are typically of uniform size for the entire energy
continuum~\cite{fasnacht2004adaptive}. Some methods such as
AdaWL~\cite{koh2013dynamically} employ a tunable mechanism for controlling the
binning for low entropic states in order to ensure the exploration of all
energies.  The method introduced in this paper is designed to scale appropriately
as bin size is changed.

In this section we will introduce four closely related flat histogram
methods each of which rely on a weight function $w(E)$.  In these
algorithms, the probability of accepting a move is given by
\begin{equation}
	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
\end{equation}
which biases the simulation in favor of energies with low weights.  A set of
weights that are proportional to the density of states $D(E)$ of the system will
result in an entirely flat histogram, which means that the weights should
converge to being proportional to the density of states.  The natural logarithm
of the weight is typically stored, since the density of states will often vary
over a few hundred orders of magnitude. In the microcanonical ensemble, the
entropy is defined as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the
logarithm of the weight is an approximation of the entropy.

Each approach uses a random walk in energy space to estimate the density of
states.  The core of these approaches is to update the weights at each step of
the simulation
\begin{equation}
	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
	+\gamma_t
\end{equation}
where $t$ is the number of the current move, $\gamma_t$ is an update factor
that varies over the simulation, and $E$ is the current energy.  This update
causes the random walk to avoid energies that have been frequently sampled,
enabling a rapid exploration of energy space. This approach violates detailed
balance, due to the acceptance probabilities changing with each move, but the
severity of this violation decreases as we decrease $\gamma_t$.  The four
methods differ primarily in how they schedule the decrease of $\gamma_t$.

\subsection{Wang-Landau}

The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
landau2014guide} begins with $\gamma^{\text{WL}}_{t=0}=1$, and then decreases
$\gamma^{\text{WL}}$ in discrete stages.  When the energy histogram is
sufficiently flat, $\gamma^{\text{WL}}$ is decreased by a specified factor of
$\frac12$.  The flatness is defined by the ratio between the minimum value of
the histogram and its average value.  When this flatness reaches a specified
threshold (typically 0.8), the $\gamma^{\text{WL}}$ value is decreased and the
histogram is reset to zero.  This approach requires that the energy range of
interest be known in advance, and difficulties can occur with this flatness
criteria due to the fact that some energies in this energy range might never be
sampled~\cite{troster2005wang}.  The entire process is repeated until
$\gamma^{\text{WL}}$ reaches a desired cutoff.

The Wang-Landau approach thus has three parameters that need be
specified: the factor by which to decrease $\gamma^{\text{WL}}$ when flatness is
achieved, the flatness criterion, and the cutoff that determines when
the computation is complete.  In addition, an energy range (or in
general, a \emph{set} of energies) must be supplied, so that the
flatness criterion can be defined.

\subsection{$1/t$-Wang-Landau}

The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
from dropping below $N_S/t$~\cite{belardinelli2008analysis,
schneider2017convergence}. The method follows the standard WL algorithm with two
modifications.  Firstly, the histogram is considered flat, and $\gamma_t$ is
decreased by a factor of two, when every energy state has been visited once,
i.e. when the WL ``flatness'' becomes nonzero. Secondly, when
$\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the algorithm switches to use
$\gamma_t = N_S/t$ for the remainder of the simulation:
\begin{align}
  \gamma_t^{1/t\text{-WL}} = \begin{cases}
     \gamma^{\text{WL}}_t & \gamma^{\text{WL}}_t > \frac{N_S}{t} \\
     \frac{N_S}{t} & t \ge t_0
 \end{cases}
\end{align}
where $t$ is the number of moves, $\gamma^{\text{WL}}_t$ is the Wang-Landau update factor
at move $t$, and $N_S$ is the number of energy bins.

\subsection{SAMC}
The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses the lack of
convergence of Wang-Landau's approach with a simple schedule by which the update
factor $\gamma_t$ is continuously decreased~\cite{liang2007stochastic,
werlich2015stochastic, schneider2017convergence}.  The update factor is defined
in the original implementation~\cite{liang2007stochastic} in terms of an tunable
parameter $t_0$,
\begin{align}
\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}\label{eq:1}
\end{align}
where as above $t$ is the number of moves that have been attempted.
SAMC offers extreme simplicity, combined with is proven convergence.
Provided the update factor satisfies
\begin{align}
\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
\end{align}
where $\zeta > 1$, Liang has proven that the weights converge to the true
density of states~\cite{liang2006theory, liang2007stochastic,
liang2009improving}.  In addition, the energy range need not be known \emph{a priori}.
The time to converge depends only on the choice of parameter $t_0$.
Unfortunately, there is no prescription for finding an acceptable value for
$t_0$, and while the algorithm formally converges, for a poor choice of $t_0$
that convergence can be far too slow to be practical. Liang \emph{et al.} give a
rule of thumb in which $t_0$ is chosen in the range from $2N_S$ to $100N_S$
where $N_S$ is the number of energy bins~\cite{liang2007stochastic}. Schneider
\emph{et al.} found that for the Ising model this heuristic is helpful for small
spin systems, but that larger systems require an even higher $t_0$
value~\cite{schneider2017convergence}.  We will describe below one case we
examined, in which $t_0$ needs to be as much as two orders of magnitude higher
than the rule of thumb of $100N_S$ in order to converge in $10^{12}$ moves.

Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t^{\text{SA}}$
by a factor $\gamma_0$~\cite{werlich2015stochastic}.  While this may result in
an improved rate of convergence, it adds yet another parameter that must be
empirically determined, and we have not explored this additional degree of
freedom.

\subsection{SAMC convergence time}\label{sec:samc-convergence}
A primary difficulty in using the SAMC method lies in identifying an
appropriate value for $t_0$.  Although SAMC is proven to formally
converge regardless of the $t_0$ value, a choice that is either too
high or too low will result in prohibitively slow convergence to the true
entropy of the system.  It is instructive to consider separately
values of $t_0$ that are too low or too high.

We can place a rigorous \emph{lower} bound $t_{\min}$ on the number of
moves required to find the true entropy by considering the total change
that needs to be made to the entropy:
\begin{align}
  \Delta S_{\text{tot}} \equiv \sum_E S(E) - S_{\min}.
\end{align}
This connects with $\gamma_t$ because the $\ln w(E)$ (our approximation for entropy)
has $\gamma_t$ added to it on each move.  Thus a minimum number of moves
that could possibly result in the true entropy $S(E)$ starting with a flat set of
weights is determined by the total entropy change required.  We can estimate
this number of moves required by summing the $\gamma_t$,
which we can approximate using an integral:
\begin{align}
   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t^{\text{SA}} \\
  % &= \int_0^{t_{\min}} \gamma(t)dt \\
  &\approx t_0 + \int_{t_0}^{t_{\min}} \frac{t_0}{t}dt
  \\
  &= t_0\left(1 + \ln\left(\frac{t_{\min}}{t_0}\right)\right)
\end{align}
Solving for ${t_{\min}}$ we find that
\begin{align}
  {t_{\min}} &= t_0 e^{\frac{\Delta S_{\text{tot}}}{t_0} - 1}
\end{align}
which means that the minimum time to converge grows exponentially
as $t_0$ is made smaller.  You \emph{seriously} don't want to underestimate
$t_0$!

One might reasonably choose to err by selecting a large $t_0$. The rate of
convergence is harder to estimate when $t_0$ is large, but in general
$\gamma_t^{\text{SA}}$ itself forms a lower bound on the accuracy with which the
entropy may be known, with an unknown prefactor which is related to the
coherence time ($t = t_0$) of the Monte Carlo simulation.  Since
$\gamma_t^{\text{SA}}$ is given by $t_0/t$, the time to converge to a given
accuracy is increased in proportion to the ratio by which we overestimate $t_0$.
Thus, while it is exponentially painful to underestimate $t_0$, overestimating
by several orders of magnitude is also not acceptable.  We should note that
these extreme limiting cases do not preclude the possibility that there is a
wide range of $t_0$ values that lead to an acceptable convergence rate.

\section{SAD Algorithm}\label{sec:sad}
The Stochastic Approximation with Dynamic update factor (SAD) method
is a variant of the SAMC Algorithm that attempts to dynamically choose
the modification factor rather than relying on system dependent
parameters such as $t_0$ or $\gamma_0$.  There is an immediate
advantage of such an algorithm where parameters are chosen independent
of system size or type. Each flat-histogram method has unique
advantages and disadvantages.  Wang-Landau and $1/t$-WL require an
energy range for initialization.  SAMC removes this energy range
requirement but requires simulating every possible energy. Our
proposed method SAD requires the user to input $T_\text{min}$, the
lowest temperature of interest, which is an immediate disadvantage of
the method. However, identifying a minimum temperature of interest
$T_\text{min}$ may be easier for a user than determining in advance
an energy range of interest or the unphysical parameter $t_0$.

We set the maximum temperature of interest as $T=\infty$ for a couple
of reasons.  First, when this temperature is reached, the system is
maximally randomized, so allowing infinite temperature further ensures
that the system can surmount an energy barrier of any size.  Second,
the high temperatures tend to be ``easy'' to simulate, and tend to
converge very quickly.  We note, however, that this is under the
assumption that infinite temperature corresponds to a finite energy,
as is the case for configurational sampling.  Were we to include
kinetic energy in a simulation (as in molecular dynamics), or to
examine a system such as a high density Lennard-Jones fluid with an
entropy maximum at an energy that is far from the energy range of
physical interest, then we would need to apply either a maximum
temperature or a maximum energy, either of which could slow down the
rate at which state space is explored.

\begin{figure}
  \includegraphics[width=\columnwidth]{N50-lndos-comparison}
  \caption{The entropy of a square-well fluid with 50 atoms and filling fraction
        0.3 as the green line.  The green hatched area reflects the
        minimum entropy change needed to converge to the true value.
        The light blue area is the quadratic approximation
        for the change in entropy.  The vertical dotted lines represent
        the energy corresponding to $T=1/3$ and $T=\infty$.}
  \label{fig:entropy-cartoon}
\end{figure}

While for SAMC, the update factor is defined in the original
implementation (see equation~\eqref{eq:1}), for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
the SAMC parameter
$t_0$ should have dimensions of entropy.
We begin with an estimate of the average value of the entropy (relative
to the lowest entropy at $T_{\min}$).  If we assume a quadratic
dependence on energy (see Fig.~\ref{fig:entropy-cartoon}), this is given by
\begin{align}
\langle S\rangle \approx \frac13 \frac{E({T=\infty}) - E(T_{\min})}{T_{\min}}
\end{align}
We approximate this energy difference by $E_H -E_L$ where $E_H$ and
$E_L$ are defined below. The entropy numerator of the update factor in
general should scale with the total number of interesting energy states
$N_S$, since updates to the weights are distributed between that many
energy states.  The product $N_S\langle S\rangle$ is the total change
of entropy required (starting from constant weights) to find the true
entropy, and puts a lower bound on the convergence time. After
\emph{long} times, when all the energies have been explored a long time ago, we
wish for a lower update factor in order to more rapidly refine the
remaining error in entropy.  We track the time at which we first visited
each possible energy.  We define $t_L$ to be the last time that we
encountered an energy that we currently believe is in the energy range
of interest, so a $t\gg t_L$ we feel confident that we have established
the true energy range of interest. We gradually transition to a lower
update factor (but still asymptotically scaling as $\gamma_t \propto
1/t$ to ensure eventual convergence).  Finally, we wish for an update
factor that is \emph{never} greater than 1, because a very large update
factor could introduce very large errors in entropy that may take many iterations
to remove.  The SAD expression for $\gamma_t$ which incorporates these
ideas is:
\begin{align}
  \gamma_{t}^{\text{SAD}} =
     \frac{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
     }{
       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
     }
\end{align}
where $E_H$ and $E_L$ are the current estimates for the highest and
lowest energies of interest as defined below.  This factor
asymptotically has the same $1/t$ behavior as the original SAMC
algorithm and with the same $N_S$ prefactor used by the $1/t$-WL
method; however for earlier values of $t$, the update factor drops as
$1/t^2$ and jumps every time a new energy is determined to be of
interest.  This behavior allows SAD to dynamically prevent the update
factor from decreasing too rapidly.

\begin{figure}
  \includegraphics[width=\columnwidth]{gamma-n256}
  \caption{The update factor $\gamma_t$ versus iteration for the
    square-well fluid with 256 atoms, for four different methods: WL, $1/t$-WL, SAMC,
    and SAD.}
    \label{fig:gamma-vs-t}
\end{figure}
Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for the related
methods SAD, WL, $1/t$-WL, and SAMC.  For SAMC, $\gamma_t$ remains
constant before dropping as $1/t$.  WL $\gamma_t$ remains at 1 for
many iterations, and then decreases very rapidly, with $1/t$-WL
behaving similarly but decreasing more aggressively before
transitioning to a more conservative $1/t$ behavior.  The update
factor for SAD fluctuates dynamically around a value less than 1 for
early MC moves, and then decreases as approximately $1/t$ while
continuing to fluctuate as new energies are found to be important.  At
intermediate times, the SAD $\gamma_t$ decreases as $1/t^2$ before
asymptoting to $N_S/t$, which is the same as $1/t$-WL.

Since SAD does not explore \emph{all} energy states, it needs to
determine what energy range corresponds to the temperature range of
interest defined by $T_{\min}<T<\infty$. The simulation is responsible
for determining and updating this energy range. Given the true entropy
$S(E)$, we can define the interesting energy range as
%\begin{align}
  $E(T_{\min}) <E< E(T=\infty)$
%\end{align}
where $E(T)$ is the energy that maximizes $S-E/T$.  During the course of the
simulation, this precise energy is challenging to evaluate accurately.
In order to ensure that we sample this entire energy range adequately,
we define two energy limits:  a high energy $E_H$ and a low
energy $E_L$, which define the range over which the energy histogram
is made flat. At move $t$, $E_H$ and $E_L$ are the greatest and lowest
energy that prior to that move
had the highest histogram value (i.e. been visited the most times) at some point
during the course of the simulation.
This definition results in a ``ratcheting'' effect, in which $E_H$
may only increase, while $E_L$ may only decrease over the course of the
simulation, which results in a conservative estimate of the range of
energies that need be sampled.

During the simulation when considering a move inside the energy
range of interest $E_L \le E \le E_H$, the weights are used as in the
three methods already described.  If $E\ge E_H$, the weight is taken
to be
\begin{align}\label{eq:highw}
  w(E>E_H) &= w(E_H),
\end{align}
which corresponds to an infinite temperature.  This choice ensures
that if the maximum in entropy is at an energy $E_{\max}>E_H$, then
the energy $E_{\max}$ will eventually have the highest number of
counts and the ratcheting will result in $E_H\ge E_{\max}$ .  At lower
energies, Boltzmann weights corresponding to the minimum temperature
are used:
\begin{align}\label{eq:loww}
  w(E<E_L) &= w(E_L)e^{-\frac{E_L-E}{T_{\min}}}.
\end{align}
This choice has the result that if the energy $E_{\min}$ at which the free
energy at $T_{\min}$ is minimized is less than $E_L$, the lower energy
limit will ratchet down to include $E_{\min}$.
Each time we change the value of $E_H$ or $E_L$, the weights outside the
new portion of the interesting energy range are set to the expressions
in Equations~\ref{eq:highw} and~\ref{eq:loww}.

%~ \paragraph{Energy binning with SAD}

A significant advantage of SAD over SAMC---which the $1/t$-WL and WL
methods share \emph{after} they have discovered all the energies---is that
the schedule for $\gamma_t$ automatically responds to the choice of bin
size.
SAD should perform similarly over a reasonable range of bin sizes
because $\gamma_t \propto \nicefrac{N_S}{t}$.  As the number of energy
states $N_S$ found increases (fine binning), the time spent $t$ in
each bin will decrease with the effect that the convergence should be
roughly independent of the bin size chosen.  SAMC could be used with a
prefactor $\gamma_0$ to aid in a similar
way~\cite{werlich2015stochastic} but this adds yet another parameter
for the user to choose.

%~ \paragraph{Scaling, step size, and SAD}
%~ Methods such as WL can fail to converge for complex systems in part
%~ because the same random MC moves are performed for the entire energy
%~ range.  The main problem lies in trying to determine how to optimize
%~ the step size such that the time to explore all energies is reasonable.
%~ Small step sizes lead to improved sampling at the expense of slow
%~ convergence. Koh et al. proposed multiplying $\gamma$ by an energy
%~ dependent factor in order to improve the converge of WL in regards to
%~ convergence and scaling~\cite{koh2013dynamically}.  Because the update
%~ factor for SAD has an energy dependent factor already present (and
%~ dynamically adjusted), we expect that SAD will perform consitently
%~ scaling free for a reasonable choice of step size.  SAD should perform
%~ extremely well compared with other flat histogram methods when applied
%~ to complex continuous systems such as polymers/proteins and spin models.

%~ \subsection{Convergence}
%~ Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
%~ convergence criteria.  However, we also want to ensure that SAD avoids
%~ the exponentially poor convergence time that SAMC suffers when $t_0$ is
%~ small as discussed in Section~\ref{sec:samc-convergence}.
%~ To place a lower bound on the convergence of SAD, we will again
%~ approximate the total entropy change using an integral.  To sidestep
%~ the dynamical nature of SAD, we ensure that \emph{after discovering the
%~ last important energy} at time $t_L$ the entropy change
%~ $\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
%~ Thus we integrate the update factor from $t_L$ to some minimum time
%~ $t_{\min}$ necessary for convergence.  Technically this is not a lower
%~ bound as in the case of SAMC because the algorithm will already have a
%~ reasonable approximation for the entropy.
%~ \begin{align}
%~ \Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
     %~ \frac{
       %~ 3\langle S\rangle + \frac{t}{t_L}
     %~ }{
       %~ 3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
     %~ } dt \\
%~ &\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
%~ \end{align}
%~ where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
%~ that it takes a long time to find all the interesting energy states.
%~ Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
%~ of order unity.  This omission only makes our bound more conservative,
%~ and enables us to solve to find that
%~ \begin{align}
  %~ t_{\min} &= \frac32 t_L
%~ \end{align}
%~ which means that we only need to run fifty percent longer after we
%~ discover the final interesting energy state in order to have made
%~ sufficient updates to \emph{possibly} converge to the true entropy.

\section{Results}\label{sec:results}

\subsection{Square-well fluid}
As our first test case, we consider the square-well fluid i.e. a
system of particles whose interactions are governed by a square-well
potential~\cite{singh2003surface, barker2004perturbationSW}.  The
square-well potential is an ideal test-bed as it is a simple model for
a liquid, which includes both attractive and repulsive
interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
To date, there have not been any published direct convergence comparison tests
for flat histogram methods applied to the square-well fluid.
The potential $U(\textbf{r})$ for such a system is given by
\begin{equation}
 U(\textbf{r})=\begin{cases} \infty &
 \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
 \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
 \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
\end{equation}
where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
reduced range of the potential well, and $\epsilon$ is its depth. This model has
the further advantage that binning is not required because the energy is
discrete.

\begin{figure*}
  \includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
\includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
  \caption{(a) The average entropy error for each MC method for $N=50$,
               $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
               as a function of number of iterations run. The error is
               averaged over 8 independent simulations, and the best
               and worst simulations for each method are shown as a
               semi-transparent shaded area, and
           (b) the update factor $\gamma_t$ versus iteration number
               for the same simulations.
           (c) The average entropy error for each MC method for the
               same physical system with a smaller displacement distance
               $\delta_0 = 0.005\sigma$,
               as a function of number of iterations run, and
           (d) the update factor $\gamma_t$ versus iteration number
               for the same simulations.
  }\label{fig:fast-slow-gamma}
\end{figure*}

We tested the algorithms on two square-well fluid systems.  The first
is a smaller simulation with a particle number of 50, a well-width of
$\lambda = 1.3$, and a volume corresponding to a filling fraction
(defined as the fraction of volume filled by atoms) of $\eta =
0.3$. The second system is larger, with a particle number of 256, a
well-width of $\lambda = 1.5$, and a volume corresponding to a filling
fraction of $\eta = 0.17$.  For each system we use a reasonable
root-mean-square displacement distance $\delta_0 = 0.05\sigma$ for
proposed moves, and for the smaller system we also use an unreasonably
small displacement distance of $0.005\sigma$. The simulations explore
the energy space of the systems with minimum reduced temperatures of
$T_{\text{min}} = 1/3$ for simulations of the smaller system, and
$T_{\min}=1$ for the larger system.  All simulations lead to the
minimum important energy $E_{\min}$ and maximum entropy energy
$E_{\max}$ being calculated (with the exception of the WL methods
where both of these parameters are needed \emph{a priori}).

The SAMC simulations
computed the density of states for the entire range of possible
energies.  The SAD simulations determined the energy range of interest
dynamically as described above, based on a specified $T_{\min}$.  For
the WL and $1/t$-WL simulations, we constrained the simulation to
remain in the energy range corresponding to $T_{\min} < T < \infty$,
as determined by a previous SAMC simulation.  Thus the WL and $1/t$-WL
simulations were given extra information that in practice would not be
available without additional computational effort, and the SAMC simulations
computed the entropy over the entire range of possible energies, which
required more effort.

We use the average entropy error versus moves as a metric to compare
simulation runtimes and overall convergence. The overall accuracy
is determined by examining the fractional error of a particular method to
a precise reference system. For each simulation, the reference system
is chosen to be the final output of a SAMC simulation with a fixed energy range
corresponding to the temperature range of interest.
Although SAMC does not require an energy range as an input parameter,
we find that by limiting the simulation to this energy range, we can
achieve much faster convergence with a smaller $t_0$.  We compute an average
of the error by averaging the error in the entropy over the
interesting energy range, and then averaging \emph{this} error over several
simulations run with different random number seeds.
%~ \begin{align}
%~ \epsilon_\text{avg} =
  %~ \frac{1}{N_S \text{num seeds}}\sum_s^{\text{num seeds}}\sum_{i}^{N_S} \left|S_\text{ref}(E_i) - S_\text{method}(E_i)\right|
%~ \end{align}

\subsubsection{50 square-well atoms}
For this 50 atom simulation, we chose a minimum reduced
temperature of $1/3$, which corresponds to an interesting energy range
from $-248$ to $-120$.  The number of important energy states
for this system is therefore $N_S = 129$.  The entropy of this system is shown in
Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
energy range the entropy differs by 198, corresponding to a ratio of
$10^{86}$ between the highest and lowest density of states.

In order to explore the effect of simulation details on convergence,
we consider two values for the displacement distance by which atoms
are moved during a Monte Carlo step.  We began with a reasonable
displacement distance of $\delta_0 = 0.05\sigma$, which corresponds to
an acceptance rate of proposed moves of 38\%.  We further ran
simulations with a much smaller displacement distance of $\delta_0 =
0.005\sigma$, which resulted in an acceptance rate of 86\%, which
converged more slowly.

Figure~\ref{fig:fast-slow-gamma}a shows the average error in the
entropy as a function of time for this system with the reasonable
displacement distance of $\delta_0 = 0.05\sigma$.  The solid/dashed lines
represent the average of the absolute value of the error in the
entropy averaged over eight simulations using different random number
seeds.  The range of average errors for each simulation is shown as a
shaded region around its mean error.  By the time $10^8$ moves have
been made all but the SAMC simulation with the shortest $t_0$ have
begun to converge as $1/\sqrt{t}$.  We then see the WL error
saturate around $10^{10}$ moves.

Figure~\ref{fig:fast-slow-gamma}c shows the average error in
the entropy as a function of time for this system with the
unreasonably small displacement distance of $\delta_0 = 0.005\sigma$.
%
The smaller translation scale causes all methods to take additional time to
explore all energies. Based on random walk scaling, the convergence time of an
ideal method should scale roughly as $\delta_0^{-2}$ in the limit of small
$\delta_0$, that is, one order of magnitude decrease in the displacement
distance should result in two order of magnitude increase in convergence time.
SAMC simulations with a $t_0$ value that rapidly converged for $\delta =
0.05\sigma$ do not converge at all in $10^{12}$ moves for a translation scale of
$\delta = 0.005\sigma$. It is also worth noting that for the smaller
displacement distance, the SAMC rule of thumb of choosing $t_0$ to be
approximately $100N_S$ is no longer valid. SAD, WL, and $1/t$-WL handle the
shift in displacement distance and converge roughly as expected.

The methods SAD, WL, and $1/t$-WL compensate for the smaller
displacement distance by reducing $\gamma_t$ more slowly, as can be seen from
Figure~\ref{fig:fast-slow-gamma}b and ~\ref{fig:fast-slow-gamma}d.
The update factors take approximately 10$\times$ longer to reach steady-state
for the smaller displacement distance. Because of this update behavior,
these methods are less sensitive to the choice of displacement
distance than SAMC is.

%~ SAMC methods that converge
%~ in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not
%~ necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.
%
%~ Sad performs similarly regardless of the displacement distance chosen although
%~ it transitions to $1/\sqrt{t}$ convergence after

\begin{figure}
\includegraphics[width=\columnwidth]{periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
  \caption{
  The average entropy error for each MC method for $N=256$,
               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
               as a function of number of iterations run.  The error is
               averaged over 8 independent simulations, and the best
               and worst simulations for each method are shown as a
               semi-transparent shaded area.  The update factor for this
               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
\end{figure}

\subsubsection{256 square-well atoms}
Next we introduce a considerably larger simulation containing 256
atoms which has a maximum entropy about 1500 greater than its minimum.
This makes exploring the entire range of energies extremely expensive,
and strongly favors the methods that restrict the energy (or
temperature) range of interest.  For this simulation, we chose a much
higher minimum reduced temperature of $1.0$, which corresponds to an
interesting energy range from $-915$ to $-509$.  The number of
important energy states for this system is then $N_S = 407$.  The
minimum entropy over this energy range is just 395 less than the
maximum, corresponding to a ratio of only $10^{118}$ between the
highest and lowest density of states.

%$395.2833013702813 - 4.754731035740479e^{-8}$ = $395.2833$ so $2^{396.2833}$
%gives us about $10^{118}$ from 'samc-1e4-256-s8.yaml'

Figure~\ref{fig:n256} shows the average error in the entropy as a
function of moves for this system with the reasonable displacement
distance of $\delta_0 = 0.05\sigma$.  The solid lines represent the
average of the absolute value of the error in the entropy averaged
over eight simulations using different random number seeds.  The range
of average errors for each simulation is shown as a shaded region
around its mean error.  By the time $10^8$ moves have been made all
but the SAMC simulation with the shortest $t_0$ shown have begun to
converge as $1/\sqrt{t}$.  We then see the WL error saturates around
$10^{10}$ moves.  Once again, the convergence of SAD is essentially
the same as that of the $1/t$-WL method.

\begin{figure}
  \includegraphics[width=\columnwidth]{lj-cv}
  \caption{The heat capacity of the LJ31 cluster after $10^{12}$ moves
    as calculated with each method.  Each simulation used a bin size
    of $\Delta E=0.01\epsilon$.}
  \label{fig:lj-cv}
\end{figure}

\subsection{31-atom Lennard-Jones cluster}

We also examined convergence of each algorithm on the 31-atom cluster
(LJ31), which is the smallest Lennard-Jones cluster that exhibits
broken ergodicity and a solid-solid phase transition at a low
temperature~\cite{doye1998thermodynamics,
  martiniani2014superposition}.  Clusters of Lennard-Jones atoms have
been frequently used for testing Monte Carlo
algorithms~\cite{wales1997global, neirotti2000phase,
  frantsuzov2005size, mandelshtam2006multiple}.  The LJ31 cluster
is commonly used for testing convergence of algorithms
because it features a two-funnel energy landscape with a significant
barrier between the two low-energy states~\cite{calvo2000entropic,
  calvo2000phase, poulain2006performances,
  martiniani2014superposition}.  In particular, the LJ31 cluster
features a low-temperature solid-solid phase transition that is challenging to
converge~\cite{mandelshtam2006multiple}.  Poulain \emph{et al.}
tested a number of variants of the WL algorithm on LJ31 and failed to
converge the spike in heat capacity corresponding to the solid-solid
phase transition~\cite{poulain2006performances}.

%% \jpcom{This \P\ should be commented once summarized.
%% Poulain \emph{et al.} found very poor convergence for a variety of
%% modified WL methods on the LJ31
%% cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
%% energy range or bin size that was used in their simulation, so their
%% results cannot be reproduced.  They do mention that they use 1000
%% energy bins, which if they used an energy range of
%% $[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
%% E\approx 0.13\epsilon$, which we find to be too large to accurately
%% predict the heat capacity peak at $0.03\epsilon/k_B$.  Looks like they
%% only ran $\sim 10^{10}$ moves. We can't converge in that few moves either.}

Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster
over the temperature range from $k_BT=0.01\epsilon$ to
$k_BT=0.40\epsilon$.  There are two peaks, a large peak around $k_BT
\approx 0.3\epsilon$ corresponding to melting temperature, and a
small peak around $k_BT \approx 0.027\epsilon$ for the solid-solid
transition from a Mackay-to-anti-Mackay (M $\rightarrow$ aM)
transition~\cite{doye1997thermally, hendy2001molecular,
  mandelshtam2006multiple}.  We constrain our atoms within a spherical
box of radius 2.5$\sigma$, as is common in the
literature~\cite{poulain2006performances, mandelshtam2006multiple}.
We note that using a larger box (with radius $5\sigma$) has a large
effect on the melting peak, and a smaller but still significant effect
on the solid-solid transition.  Since our focus is on the convergence
rate, we will restrict ourselves to the smaller box size.  In addition, we
further restricted all our simulations to negative total potential energies.

\begin{figure}
  \includegraphics[width=\columnwidth]{lj-cv-error}
  \caption{The maximum error in the heat capacity of the LJ31 cluster
    in the temperature range 0.01~$\epsilon\le k_BT \le 0.05\epsilon$.
    Note that the value of $C_V$ ranges from 90$k_B$ to 120$k_B$, so
    the smallest (maximum) errors seen are around 1\%.  All WL methods
    shown were run with a histogram bin width of $\Delta
    E=0.01\epsilon$.  The converged reference heat capacity was
    computed with a $1/t$-WL simulation with the energy constrained to
    a range $-133.53\epsilon\le E\le -110\epsilon$.
  }
  \label{fig:lj-cv-error}
\end{figure}

In order to test the low-temperature convergence, we focused on the
heat capacity in the temperature range $0.01\epsilon \le k_BT \le
0.05\epsilon$ (see inset in Fig.~\ref{fig:lj-cv}), which contains
the solid-solid phase transition.
%
We compute the heat capacity from the entropy using a standard
canonical analysis, which gives
\begin{align}
  C_V &= 31\times\frac32k_B + \frac1{k_BT^2}\left< (E- \langle
  E\rangle)^2\right>
  \label{eq:cv}
\end{align}
where the first term represents the kinetic energy contribution, and
the second term is the potential energy contribution, with canonical
averages using the microcanonical entropy to compute multiplicities.
%
For our reference heat capacity, we ran a $1/t$-WL simulation with the
energy constrained to a range $-133.53\epsilon\le E\le -110\epsilon$.
We chose this energy range through experimentation, as we found the
convergence rate to be very sensitive to the minimum energy chosen.

%% \jpcom{this is a basically a repeat of above...
%% The heat capacity changes significantly even at these low temperatures
%% when the size of the cavity is increased.  We use a 2.5 cavity because
%% ``everyone else'' does~\cite{poulain2006performances}.  For purposes
%% of testing the convergence of the algorithm, we don't need to solve
%% for the true heat capacity.  Furthermore, the heat capacity at fixed
%% volume is not the most meaningful quantity to predict for a cluster.
%% }
We ran simulations with several bin widths using each algorithm we studied
and show results for $\Delta E=0.1\epsilon$ and
$0.01\epsilon$.
When examining the converged results, we saw that each bin size yielded a
measurably different low temperature heat capacity. The largest
bin size of $0.1\epsilon$ introduced a maximum error in heat capacity
of more than $20k_B$ compared with the smaller bin size.  We will focus on
the smaller bin size, and
used a bin size of $0.01\epsilon$ for our converged reference heat capacity.

Figure~\ref{fig:lj-cv-error} shows the maximum error in heat capacity
over the temperature range $0.01\epsilon\le k_BT\le 0.05\epsilon$ as a
function of the number of Monte Carlo moves performed.
%
We explored the convergence behavior of each algorithm at two energy
bin widths ($\Delta E=0.01\epsilon$ and $\Delta E=0.1\epsilon$).  We
were unable to converge the WL or $1/t$-WL algorithms with the larger
bin size.
%
The rate of convergence of SAD for the heat capacity seems to be roughly
independent of the bin width.  In contrast (but not shown) the entropy
itself shows much larger statistical fluctuations with smaller bin
sizes.  The canonical averaging process in Eq.~\ref{eq:cv} averages
out those statistical fluctuations, which reduces the penalty of using
a smaller bin size.

%% Dependence on bin width.  Figure~\ref{fig:lj-cv-error} shows that the
%% heat capacity converges just as well for bin widths varying over two
%% orders of magnitude, although for bins as large as $\Delta
%% E=0.1\epsilon$ the error in the heat capacity is significant, as is
%% apparent in Fig~\ref{fig:lj-cv}.  The entropy itself converges much
%% more slowly for a smaller $\Delta E$, but when summing over energies
%% the statistical fluctuations in the entropy cancel. WL and $1/t$-WL
%% show similar behavior, so this is not new, but is encouraging.

The convergence behavior of the WL and $1/t$-WL is incredibly
sensitive to the minimum energy chosen, which complicates the process of
obtaining a workable result.  The known ground state energy of this
cluster is $-133.586422\epsilon$~\cite{northby1987structure,
  wales1997global}.  With a minimum energy of $-133.58\epsilon$, none
of our WL or $1/t$-WL simulations converged in 10$^{12}$ moves.  With minimum
energies of $-133.54\epsilon$, $-133.53\epsilon$, and
$-133.52\epsilon$, the simulations do seem to converge, although the
lowest energy case was significantly slower.
This result highlights the advantage of being able to
specify a temperature range rather than an energy range when computing
properties as a function of temperature.  It is not only more
convenient, but also far more efficient due the the fact that only a
single simulation need be performed.

%% Specifying temperature range rather than energy range is beneficial.
%% Both WL and $1/t$-WL are incredibly sensitive to the minimum energy
%% chosen, as seen in Fig.~\ref{fig:lj-cv-error}.  Cite previous work
%% that showed them not converging because it chose too low a minimum
%% energy, proving that it is \emph{not} a trivial procedure to determine
%% an appropriate energy range.  Specifying the temperature range
%% \emph{does} matter! We had to do a whole bunch of $1/t$-WL
%% computations in order to determine what would be a sufficient
%% low-energy cutoff.

Finally, we want to mention our results for SAMC, which are not shown
in our plots in order to reduce clutter.  With a well-chosen power of ten value for
$t_0$, SAMC can converge to the correct result, but less rapidly than
either $1/t$-WL or SAD.  However, with a poor $t_0$
parameter, the method cannot converge in practice.  In addition, the
optimal $t_0$ parameter depends on the bin width as well as the energy
range chosen, further complicating matters.

%% SAMC is a major pain.  The $t_0$ value needs to be adjusted when
%% $\Delta E$ is changed or when the energy range is changed.  The energy
%% range has the same downsides as the WL methods, although it manifests
%% differently, with failure to converge as the problem, rather than
%% failure for $\gamma_t$ to decrease.

\section{Conclusions}
We have introduced a new algorithm, a variant of the SAMC method, which
effectively samples the
energy space corresponding to a desired range of temperatures for a
few systems.
%
We find that both SAD and $1/t$-WL converge more rapidly than SAMC, and
unlike WL consistently converge to the correct density of states. SAD requires
the user to specify a temperature range of interest rather than an
energy range of interest as $1/t$-WL does.  For use cases in which
a range of desired temperatures is known, this will make the SAD method
considerably more convenient.

We find that SAMC converges for a reasonable choice of
$t_0$ but this parameter can be difficult to tune especially across
significantly differing systems.  We find that even simple
changes to the Monte Carlo moves can have a dramatic effect on the
range of practical $t_0$ values.  Additionally, SAMC does not
converge as rapidly as either SAD or $1/t$-WL even for the best choice of $t_0$,
when a relatively small range of energies is required, because it always
simulates all possible energies.

Finally, we find that for examining a Lennard-Jones cluster the WL and
$1/t$-WL algorithms---versions of which had previously been found to
be unsuccessful on this problem~\cite{poulain2006performances}---have
convergence properties that are highly dependent on the choice of
energy range to be examined.  This is particularly problematic when
examining the low-temperature heat capacity, where it is difficult to
determine the lowest energy that will have a significant impact at the
temperatures of interest. This is an example of how SAD is more
powerful when the range of energies is not known in advance.

%\section{Acknowledgement}

%We gratefully acknowledge D.W. Siderius and
%J.R. Errington for helpful communications regarding the implementation
%of WL-TMMC.

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}

### 2020-01-21 scripts-python/gamma-plot.py
gamma-plot is now correctly building the figures.

@@ -12,9 +12,9 @@ Tmin = float(sys.argv[2])
 plt.figure(figsize=(5, 4))
 
 try:
-    for wl in glob.glob("data/gamma/%s/wl*.txt" % filename):
+    for wl in glob.glob("../square-well/data/gamma/%s/wl*.txt" % filename):
         print("in wl")
-        print(('vanilla_wang_landau'+ wl[len("data/gamma/%s/wl" % filename):-4]))
+        print(('vanilla_wang_landau'+ wl[len("../square-well/data/gamma/%s/wl" % filename):-4]))
         wlmoves, wlfactor = np.loadtxt(wl, dtype = float, unpack = True)
         data = np.loadtxt(wl)
         moves = data[:, 0]
@@ -31,7 +31,7 @@ try:
                 factor[2*i+2] = wlfactor[i]
         colors.loglog(moves, factor,
                       'wl'
-                         + wl[len("data/gamma/%s/wl" % filename):-4])
+                         + wl[len("../square-well/data/gamma/%s/wl" % filename):-4])
         plt.ylim(ymin=1e-10, ymax=1e1)
         plt.xlim(xmin=1e3, xmax=1e12)
 
@@ -40,7 +40,7 @@ except:
 
 
 try:
-    for sad in glob.glob("data/gamma/%s/sad*.dat" % filename):
+    for sad in glob.glob("../square-well/data/gamma//%s/sad*.dat" % filename):
         data = np.loadtxt(sad)
         print((data.shape))
         if data.shape[1] > 2: # i.e. we are not using parse-yaml-out.py
@@ -87,7 +87,7 @@ for t0 in t0s:
     plt.ylabel(r'$\gamma_{t}$')
     colors.legend()
 plt.tight_layout()
-plt.savefig('figs/gamma-%s.pdf' % filename.replace('.', '_'))
+plt.savefig('gamma-%s.pdf' % filename.replace('.', '_'))
 if 'noshow' not in sys.argv:
     plt.show()
 

### 2020-01-21 scripts-python/create-N50-lndos-comparison.py
N50 figure is now building correctly.

@@ -1,0 +1,61 @@
+from __future__ import division, print_function
+import sys, os, matplotlib
+import numpy as np
+
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+if 'noshow' in sys.argv:
+        matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+import colors
+import readnew
+
+plt.figure(figsize=(5, 4))
+
+fname = '../square-well/data/samc-1e4-50-s1-reference-lndos.dat'
+e, lndos = readnew.e_lndos(fname)
+
+emax = -8
+emin = -276
+Tmin = 0.333333
+eminimportant = -265
+eminimportant = -248
+eSmax = -120
+Smin = -377.174
+Sminimportant = -255
+Sminimportant = -197.9025045
+#plt.plot(e, (e - eminimportant)/.2 + Sminimportant - Smin, 'r:')
+
+
+ei = np.arange(eminimportant, eSmax, 0.1)
+B = 1/(2*Tmin*(eSmax - eminimportant))
+Stop = Sminimportant + B*(eSmax - eminimportant)**2
+plt.fill_between(ei,
+                 Stop - B*(ei - eSmax)**2 - Smin,
+                 ei*0 + Sminimportant - Smin,
+                 color='tab:blue', alpha=0.25,
+                label=r'quadratic $\Delta S_{\textrm{tot}}$')
+
+interesting = (e > eminimportant)*(e < eSmax)
+plt.rcParams['hatch.color'] = 'tab:green'
+plt.rcParams['hatch.linewidth'] = 3
+plt.fill_between(e[interesting],
+                 lndos[interesting] - Smin,
+                 e[interesting]*0 + Sminimportant - Smin,
+                 #color='tab:green',
+                hatch='\\\\\\', label=r'actual $\Delta S_{\textrm{tot}}$',
+                facecolor='none', edgecolor='tab:green', linewidth=0)
+
+
+plt.plot(e, lndos - Smin, color='tab:green', label='$S$')
+plt.axvline(eminimportant, linestyle=':', color='tab:gray')
+plt.axvline(eSmax, linestyle=':', color='tab:gray')
+plt.ylabel('$S/k_B$')
+plt.ylim(0, (Stop-Smin)*1.05)
+plt.xlim(emin, emax)
+plt.xlabel('$E$')
+plt.legend(loc='lower right')
+plt.tight_layout()
+
+plt.savefig('N50-lndos-comparison.pdf')
+#plt.show()

### 2020-01-21 scripts-python/gamma-and-entropy-plot.py
nearly fixed the gamma-entropy plot in square well paper.

@@ -19,9 +19,10 @@ if os.path.exists('../data'):
 # Arguments for entropy portion
 energy = int(sys.argv[1])
 filebase = sys.argv[2]
-transcale = sys.argv[3] 
+transcale = sys.argv[3]
 
-tex_filebase = filebase.replace('.','_') # latex objects to extra "." characters
+save_filebase = filebase.split('/')[-1]
+tex_filebase = save_filebase.replace('.','_') # latex objects to extra "." characters
 
 methods = ['-sad3','-vanilla_wang_landau']
 if 'allmethods' not in sys.argv:
@@ -34,7 +35,7 @@ if 'allmethods' not in sys.argv:
         methods = ['-sad3-fast','-vanilla_wang_landau-fast']
 
 # For SAMC compatibility with LVMC
-lvextra1 = glob('data/comparison/%s-samc*' % filebase)
+lvextra1 = glob('../square-well/data/comparison/%s-samc*' % filebase)
 split3 = [i.split('%s-'%filebase, 1)[-1] for i in lvextra1]
 split4 = [i for i in split3 if i[-3:-1] != '-s']
 
@@ -56,7 +57,8 @@ print('methods are', methods)
 for method in methods:
     print('trying method', method)
     try:
-        dirname = 'data/comparison/%s%s/' % (filebase,method)
+        dirname = '../square-well/data/comparison/%s%s/' % (filebase,method)
+        print(dirname)
         if not os.path.exists(dirname) or os.listdir(dirname) == []:
                 continue
 
@@ -67,9 +69,6 @@ for method in methods:
         errorinentropy = data[1]
         maxerror = data[2]
         best_ever_max = min(best_ever_max, maxerror.min())
-
-        if not os.path.exists('figs/s000'):
-                os.makedirs('figs/s000')
 
         if filebase.startswith('s000'):
                 N = filebase.split('-N')[-1]
@@ -116,13 +115,13 @@ moves = np.array([1e3, 1e12])
 for i in np.arange(-8, 19, 1.0):
     colors.loglog(moves, 10**i/np.sqrt(0.1*moves), method = r'1/sqrt(t)')
 plt.xlim(moves[0], moves[1])
-if filebase == 's000/periodic-ww1.30-ff0.30-N50':
[…]

### 2020-01-21 scripts-python/comparison-plot.py
all but the lj-cluster data and scripts are in.

@@ -15,6 +15,9 @@ import colors
 
 if os.path.exists('ising/data'):
     os.chdir('..')
+    
+if os.path.exists('square-well/data'):
+    os.chdir('..')
 
 energy = int(sys.argv[1])
 filebase = sys.argv[2]
@@ -53,8 +56,8 @@ for meth in split2:
             methods.append('-%s' % meth)
 
 # For SAMC compatibility with LVMC
-if filebase.startswith('s000'):
-    lvextra1 = glob('data/comparison/%s-samc*' % filebase)
+if system == 's000':
+    lvextra1 = glob('data/comparison/%s/%s-samc*' % (system,filebase))
 else:
     lvextra1 = glob('data/comparison/%s-*samc*' % filebase)
 split3 = [i.split('%s-'%filebase, 1)[-1] for i in lvextra1]
@@ -79,7 +82,11 @@ print('methods are', methods)
 for method in methods:
     print('trying method', method)
     try:
-        dirname = 'data/comparison/%s%s/' % (filebase,method)
+        if system == 's000':
+                dirname = 'data/comparison/%s/%s%s/' % (system,filebase,method)
+        else:
+                dirname = 'data/comparison/%s%s/' % (filebase,method)
+
         if not os.path.exists(dirname) or os.listdir(dirname) == []:
                 continue
 
@@ -197,7 +204,7 @@ for method in methods:
 plt.figure('maxerror')
 colors.loglog([1e6,max_time], [best_ever_max*np.sqrt(max_time/1e6), best_ever_max], method = r'$\frac{1}{\sqrt{t}}$')
 colors.legend()
-plt.savefig('%s-max-entropy-error-%s.pdf' % (tex_filebase,transcale))
+#plt.savefig('%s-max-entropy-error-%s.pdf' % (tex_filebase,transcale))
 
 plt.figure('errorinentropy')
 if filebase == 'N128':
@@ -223,7 +230,7 @@ elif filebase == 'N128':
 
 colors.legend()
 plt.tight_layout()
-print('filename', 'figs/%s-entropy-error-%s.pdf' % (tex_filebase,transcale))
+print('filename', '%s-entropy-error-%s.pdf' % (tex_filebase,transcale))
 plt.savefig('%s-entropy-error-%s.pdf' % (tex_filebase,transcale))
 
 #plt.figure('error-vs-energy')

### 2020-01-21 scripts-python/gamma-and-entropy-plot.py

@@ -13,7 +13,7 @@ import matplotlib.pyplot as plt
 from glob import glob
 import colors
 
-if os.path.exists('../data'):
+if os.path.exists('square-well/data'):
     os.chdir('..')
 
 # Arguments for entropy portion
@@ -115,13 +115,13 @@ moves = np.array([1e3, 1e12])
 for i in np.arange(-8, 19, 1.0):
     colors.loglog(moves, 10**i/np.sqrt(0.1*moves), method = r'1/sqrt(t)')
 plt.xlim(moves[0], moves[1])
-if filebase == 'periodic-ww1.30-ff0.30-N50':
+if filebase == 's000/periodic-ww1.30-ff0.30-N50':
     plt.ylim(1e-3, 1e4)
     if "slow" in transcale:
         plt.ylim(1e-2, 1e5)
-elif filebase == 'periodic-ww1.30-ff0.30-N500':
+elif filebase == 's000/periodic-ww1.30-ff0.30-N500':
     plt.ylim(1e-1, 1e3)
-elif filebase == 'periodic-ww1.50-ff0.17-N256':
+elif filebase == 's000/periodic-ww1.50-ff0.17-N256':
     plt.ylim(1e-3, 1e3)
 #colors.legend()
 #plt.tight_layout()
@@ -145,7 +145,7 @@ elif "slow" in transcale:
 try:
     for wl in glob("../square-well/data/gamma/n%s/%s" % (N,wl_globstring)):
         print('in wl')
-        print('vanilla_wang_landau'+ wl[len("data/gamma/n%s/wl" % N):-4])
+        print('vanilla_wang_landau'+ wl[len("../square-well/data/gamma/n%s/wl" % N):-4])
         wlmoves, wlfactor = np.loadtxt(wl, dtype = float, unpack = True)
         data = np.loadtxt(wl)
         moves = data[:,0]
@@ -162,7 +162,7 @@ try:
                 factor[2*i+2] = wlfactor[i]
         colors.loglog(moves, factor,
                       'wl'
-                         + wl[len("data/gamma/n%s/wl" % N):-4])
+                         + wl[len("../square-well/data/gamma/n%s/wl" % N):-4])
         plt.ylim(ymin=1e-10, ymax=1e1)
         plt.xlim(xmin=1e2, xmax=1e12)
 

### 2020-01-21 thesis/chapter-1.tex
added thesis to the repository.

\chapter{Introduction}

% \begin{enumerate}
%     \item Chapter 1 – General Introduction Links all manuscripts
%  \item Chapter 2 – First Manuscript
%  \item Chapter 3 – Second Manuscript
%  \item Chapter 4 – Third Manuscript
%  \item Chapter 6 – General Conclusion (common conclusion linking all manuscripts
% thematically)
%  \item Common bibliography covering all manuscripts, although each manuscript may have its own reference section
% \end{enumerate}

\section{Micro-Canonical Ensemble}\label{micro}
A statistical ensemble is defined as a probability distribution $\mathcal{P}\left(\Omega\right)$ for the state of the system.
The micro-canonical\index{ensemble!micro-canonical} is a representation for a system $\Gamma$ that is determined by a fixed number\index{extensive!number $N$} of particles $N$ (and fixed components $\sum N_i$ for a multi-component system), a fixed volume\index{extensive!volume $V$} $V$, and a fixed energy\index{extensive!energy $E$} $E$. All possible states for the system $\Gamma\left(N,V,E\right)$ have the same number of particles, energy, and volume due to the system being in isolation from any external environment.
The micro-canonical ensemble\index{ensemble!micro-canonical} is for this reason also known as the $NVE$ ensemble as the micro-states depend on the \ind{extensive} macroscopic details.

The probability $\mathcal{P}$ of selecting a micro-state at random for a given range of energies centered about $E$ is given by
\begin{align}
    \mathcal{P} = \frac{1}{\Omega(E)}
\end{align}
where $\Omega(E)$ is the number of micro-states in the system $\Gamma(E)$.

\subsection{Definition of Entropy $S(E)$}\label{entropy}
The entropy $S(E)$ can be defined for any given system $\Gamma$ with discrete energy states using one of the most important formulae in physics:
\begin{align}\label{sdos}
    S(E,V,N) \equiv k_B \ln\Omega(E,V,N)
\end{align}
where the entropy is defined as the logarithm of the density of states and $k_B$ is the Boltzmann constant. An important aspect of this definition in terms of logarithms is that the entropy for multiple isolated systems is additive.
\begin{align}
    S(E_1,E_2) &= k_B \ln(\Omega(E_1)\Omega_2(E_2)) \\
    &= k_B \left( \Omega_1(E_1) + \Omega_2(E_2)\right) \\
    &= k_B \left( S_1(E_1) + S_2(E_2)\right)
\end{align}
From the \nth{1} law of thermodynamics, the system temperature can be written in terms of the density of states by refering to Eq.~\HR{sdos}.
\begin{align}
    T = \left(\diff{S(E)}{E}\right)_V = \frac{k_B}{\Omega(E)}\left(\diff{\Omega(E)}{E}\right)_V
\end{align}
The relationship between entropy $S(E)$, $\Omega(E)$, and $T$ will be discussed in further depth in sections \HR{laws} and \HR{weights}.

\subsection{The Laws of Thermodynamics and $S(E)$}\label{laws}
The \nth{1} law of thermodynamics states that for an isolated system $\Gamma$, the total internal energy is equal to the sum of the heat $Q$ and work $W$ done.
\begin{align}
    \D{U} = \D{Q} + \D{W}
\end{align}
The \nth{1} law follows directly from conservation of energy. The thermodynamic identity states that for any infinitesimal change in the system and fixed number of particles $N$, the internal energy can be written in terms of the entropy.
\begin{align}
    \D{U} =  T\D{S} - p\D{V} \\
    T = \left(\diff{U}{S}\right)_V
\end{align}

The \nth{2} law of thermodynamics states that for any isolated system $\Gamma$ the entropy $S(E) \geq 0$. The heat of the system can also be related to the entropy using the relation:
\begin{align}
    \delta Q = T\delta S
\end{align}
The relationship between heat and entropy for a system is invaluable for the \nth{1} law to relate the conjugate variables T, $S(E)$, and $U(E)$.

\section{Canonical Ensemble}
The Canonical ensemble is a representation for a system $\Gamma$ that is determined by a fixed number of particles $N$ (and fixed components $\sum N_i$ for a multi-component system), a fixed volume $V$, and a fixed temperature $T$. All possible states for the system $\Gamma\left(N,V,T\right)$ have the same number of particles, energy, and temperature due to the system being in thermal equilibrium with a heat bath.
The canonical ensemble is for this reason also known as the $NVT$ ensemble.

The probability $\mathcal{P}$ of selecting a given micro-state at random is
\begin{align}
    \mathcal{P} = e^{\nicefrac{F-E}{k_BT}}
\end{align}
where $F\left(N,V,T\right)$ is the free energy of the system $\Gamma$ and will be discussed in detail in section~\HR{potentials}.

\subsection{Boltzmann Factor and Partition Function}
In statistical thermodynamics $\beta$ is defined as the coldness or the reciprocal of the temperature.
\begin{align}
    \beta = \frac{1}{k_BT} = \frac{1}{k_B}
\end{align}
where $k_B$ is the Boltzmann constant. The partition function for a canonical ensemble can then be written in terms of the Boltzmann factor where $E$ is the total energy of the system in the respective microstate at index $i$.
\begin{align}
    Z = \sum_i e^{-\beta E_i}
\end{align}

\subsection{Thermodynamic Potentials}\label{potentials}
Thermodynamic potentials describe a given thermodynamic state of a given system and
have units of energy $k_BT$. In this section, we will discuss five common thermodynamic potentials and how they are mathematically related through Maxwell relations.
While the internal energy $U(E)$ has been briefly mentioned in section~\HR{laws}, we outline in detail the relationship between the state variables.
\begin{align}
    \D{U(S,V,N)} = T\D{S}-p\D{V}+\mu\D{N}
\end{align}
We can write the temperature $T$, pressure $p$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxU}
    T = \left( \pdv{U}{S}\right )_{V,N}  \quad\quad 
    p = -\left( \pdv{U}{V}\right )_{S,N}  \quad\quad 
    \mu = \left( \pdv{U}{N}\right )_{S,V}  \quad
\end{align}
The Helmholtz free energy $F(T,V,N)$ can be written as follows:
\begin{align}
    \D{F(T,V,N)} = \D{(U-TS)} = -S\D{T}-p\D{V}+\mu\D{N}
\end{align}
We can write the entropy $S$, pressure $p$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxF}
    S = \left( \pdv{F}{T}\right )_{V,N}  \quad\quad 
    p = -\left( \pdv{F}{V}\right )_{T,N}  \quad\quad 
    \mu = \left( \pdv{F}{N}\right )_{T,V}  \quad
\end{align}
The Enthalpy $H(S,p,N)$ can be written as follows:
\begin{align}
    \D{H(S,p,N)} = \D{(U+pV)} = -T\D{S}+V\D{p}+\mu\D{N}
\end{align}
We can write the temperature $T$, volume $V$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxH}
    T = \left( \pdv{H}{S}\right )_{p,N}  \quad\quad 
    V = -\left( \pdv{H}{p}\right )_{S,N}  \quad\quad 
    \mu = \left( \pdv{H}{N}\right )_{S,p}  \quad
\end{align}
The Gibbs free energy $G(T,p,N)$ can be written as follows:
\begin{align}
    \D{G(T,p,N)} = \D{(U+pV-TS)} = -S\D{T}+V\D{p}+\mu\D{N}
\end{align}
We can write the entropy $S$, volume $V$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxG}
    S = \left( \pdv{G}{T}\right )_{p,N}  \quad\quad 
    V = -\left( \pdv{G}{p}\right )_{T,N}  \quad\quad 
    \mu = \left( \pdv{G}{N}\right )_{T,p}  \quad
\end{align}

\subsection{Maxwell Relations}\label{maxwell}
Maxwell's relations are second derivative equalities that relate important thermodynamic variables. They are presented here to facilitate a critical derivation in Section~\HR{gas-ads}. Maxwell relations are an extension of the Schwarz theorem which states that the order of differentiation is irrelevant.
\begin{align}
    \pdv{}{\theta}\left(\pdv{\Theta}{\phi}\right)=\pdv{}{\phi}\left(\pdv{\Theta}{\theta}\right)
\end{align}
From this theorem and Section~\HR{potentials}, we can derive the most common of Maxwell's relations. The relation involving internal energy $U\left(S,V,N\right)$ relates the natural variables of pressure $p$ and temperature $T$ and can be derived using Equation~\ref{maxU}.
\begin{align}
    \left( {\pdv{U}{S}{V}} \right )_{N} = \left( \pdv{T}{V}\right )_{S,N} = - \left( \pdv{p}{S}\right )_{V,N}
\end{align}
The relation involving the Helmholtz free energy $F\left(T,V,N\right)$ relates the natural variables of entropy $S$ and pressure $p$ and can be derived using Equation~\ref{maxF}.
\begin{align}
    \left( {\pdv{F}{T}{V}} \right )_{N} = \left( \pdv{S}{V}\right )_{T,N} = \left( \pdv{p}{T}\right )_{V,N}
\end{align}
The relation involving the Enthalpy $H\left(S,p,N\right)$ relates the natural variables of temperature $T$ and volume $V$ and can be derived using Equation~\ref{maxH}.
\begin{align}
    \left( {\pdv{H}{S}{p}} \right )_{N} = \left( \pdv{T}{p}\right )_{S,N} =  \left( \pdv{V}{S}\right )_{p,N}
\end{align}
The relation involving the Gibbs free energy $G\left(T,p,N\right)$ relates the natural variables of entropy $S$ and volume $V$ and can be derived using Equation~\ref{maxG}.
\begin{align}
    \left( {\pdv{G}{T}{p}} \right )_{N} = -\left( \pdv{S}{p}\right )_{T,N} = - \left( \pdv{V}{T}\right )_{p,N}
\end{align}

\subsection{Thermodynamic Properties}\label{thermoprop}
The Maxwell relations from Section~\HR{maxwell} can be written in terms of thermodynamic properties of the material.  The isochoric heat capacity $C_V$ of a given material is defined as the amount of heat necessary to raise the temperature by a given amount.
\begin{align}
    C_V = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{V,N} = \frac{1}{T}\left(\pdv{S}{T}\right )_{V,N}
\end{align}
The isobaric heat capacity $C_p$, which is always greater than or equal to $C_V$ according to the Mayer relation, can also be written in terms of the entropy $S$ and temperature $T$ of the system.
\begin{align}
    C_p = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{p,N} = T\left(\pdv{S}{T}\right )_{p,N}
\end{align}
The Mayer relation gives the difference of the isobaric and isochoric heat capacities in terms of the thermal expansion coefficient and the isothermal compressibility.
\begin{align}
    C_p - C_V = VT\frac{\alpha^2}{\kappa_T}
\end{align}
The isobaric coefficient of thermal expansion determines to what extent a given material will deform as a result of a change in temperature.
\begin{align}
    \alpha_p = \frac{1}{V}\left( \pdv{V}{T} \right )_{p,N}
\end{align}
Similarly, the isochoric coefficient of thermal expansion can be written as the following:
\begin{align}
    \alpha_V = \frac{1}{p}\left( \pdv{p}{T} \right )_{V,N}
\end{align}
The compressibility determines to what extent a given material will change volume based on the pressure applied to the system.  The isothermal compressibility can be written in terms of the following:
\begin{align}
    \kappa_T = -\frac{1}{V}\left( \pdv{V}{p} \right )_{T,N}
\end{align}
Similarly, the isentropic compressibility can be written as the following:
\begin{align}
    \kappa_S = -\frac{1}{V}\left( \pdv{V}{p} \right )_{S,N}
\end{align}
A major driving force behind introducing the concepts of thermal expansion and compressibility is to right complex derivatives in terms of these parameters.  This will be invaluable as will be shown in Section~\HR{gas-ads}. The internal energy derivative and entropy derivative can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{U}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} - \alpha TV \quad\quad
    \left( \pdv{S}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha T} + \alpha V
\end{align}
Additionally, the enthalpy derivatives can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{H}{T} \right )_{V,N} = C_p + \frac{\alpha V}{\kappa_T}\left(1-\alpha T\right) \quad\quad
    \left( \pdv{H}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} + V\left(1-\alpha T\right)
\end{align}
Finally, the Gibbs free energy derivatives can be written in terms of the isothermal compressibility, the thermal coefficient of expansion, and the entropy.
\begin{align}
    \left( \pdv{G}{T} \right )_{V,N} = \frac{\alpha V}{\kappa_T}-S \quad\quad
    \left( \pdv{G}{p} \right )_{V,N} = V-\frac{\kappa_T S}{\alpha}
\end{align}

\section{Grand Canonical Ensemble}
The grand canonical ensemble is a representation for a system $\Gamma$ that is determined by a fixed chemical potential $\mu$, a fixed volume $V$, and a fixed temperature $T$. All possible states for the system $\Gamma\left(\mu,V,T\right)$ have the same chemical potential, volume, and temperature due to the system being in thermal and chemical equilibrium with a given resevoir.
The grand canonical ensemble is for this reason also known as the $\mu VT$ ensemble.

The probability $\mathcal{P}$ of selecting a given micro-state at random is
\begin{align}
    \mathcal{P} = e^{\nicefrac{\Phi + \mu N-E}{k_BT}}
\end{align}
where $\Phi\left(\mu,V,T\right)$ is the grand potential of the system $\Gamma$. The grand potential is often referred to as the Landau potential and written:
\begin{align}
    \Phi \equiv F -\mu N = U - TS -\mu N
\end{align}

\subsection{Chemical Potential $\mu$}
The chemical potential $\mu$ of a given system $\Gamma$ is representative of the energy that can be gained or lost due (often because of a phase transition) to a change in the conjugate variable particle number $N$. Analogous to potential energy, higher chemical potential drives particles to lower potential (thereby usually changing the particle number). The chemical potential $\mu_i$ of species $i$ can be represented using a variety of free energies (all of which are equivalent but not necessarily useful for any given experiment).
\begin{align}
    \mu_i = \left( \pdv{G}{N_i}\right )_{T,P,N_{j \neq i}}  \quad
    \mu_i = \left( \pdv{H}{N_i}\right )_{S,P,N_{j \neq i}}  \quad
    \mu_i = \left( \pdv{F}{N_i}\right )_{T,V,N_{j \neq i}}  
\end{align}

\subsection{Grand Potential and Thermodynamic Quantities}
A variety of important thermodynamic quantities can be written in terms of the grand potential. In differential form, the grand potential can be written as the following (for a single component system $\Gamma$):
\begin{align}
    \D{\Phi(\mu,V,T)} = -S\D{T}-p\D{V}-N\D{\mu}
\end{align}
The number $N$, pressure $p$, and Gibbs entropy $S$ can be written in terms of the grand potential.
\begin{align}
    N = -\left( \pdv{\Phi}{\mu}\right )_{T,V}  \quad
    p = -\left( \pdv{\Phi}{V}\right )_{T,\mu}  \quad
    S = -\left( \pdv{\Phi}{T}\right )_{V, \mu} 
\end{align}

\section{Monte Carlo Simulation}\label{monte}

% talk about Canonical here
\subsection{Broad Histogram Methods}\label{bhm}
\subsection{Weights, multiplicity, detailed balance}\label{weights}

\subsection{System Details}
% Ising, Square well, Lennard-Jones clusters, MOF?

\section{Real System Simulation}
% brief outline for broad histogram methods and how the motivate study of MOFs

\subsection{Metal-organic Frameworks (MOFs)}
\subsection{Gas Adsorption}\label{gas-ads}
% qst

\section{Summary}

% Tie everything in together here!

### 2020-01-21 thesis/chapter-2.tex
added thesis to the repository.

--- 2020-01-20 square-well/sad.tex
@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
-\usepackage{graphicx} % Include figure files
-\usepackage{color}
-\usepackage{nicefrac} % Include for inline fractions
+\chapter{Stochastic Approximation Monte Carlo with a Dynamic Update
+Factor}
 
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
-\usepackage{mdframed}
+\section{Contribution of Authors}
+Jordan K. Pommerenck, Tanner T. Simpson, Michael A. Perlin, David J. Roundy
 
-% define colors for comments
-\definecolor{dark-gray}{gray}{0.10}
-\definecolor{light-gray}{gray}{0.70}
-
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
-
-\newcommand{\davidsays}[1]{{\color{blue} [\green{David:} \emph{#1}]}}
-\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
-\begin{document}
-
-\title{Stochastic Approximation Monte Carlo with a Dynamic Update
-Factor
-}
-
-\author{Jordan K. Pommerenck} \author{Tanner T. Simpson}
-\author{Michael A. Perlin} \author{David Roundy}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-
-\begin{abstract}
-  We present a new Monte Carlo algorithm based on the Stochastic
-  Approximation Monte Carlo (SAMC) algorithm for directly calculating
-  the density of states. The proposed method is Stochastic
-  Approximation with a Dynamic update factor (SAD)
-  which dynamically adjusts the update factor $\gamma_t$ during the course of
-  the simulation. We test this method on the square-well fluid and
-  the 31-atom Lennard-Jones cluster and
-  compare the convergence behavior of several
-  related
-  Monte Carlo methods. We find that both the SAD and $1/t$-Wang-Landau ($1/t$-WL)
-  methods rapidly converge to the
-  correct density of states without the need for the user to specify an
-  arbitrary tunable parameter $t_0$ as in the case of SAMC.  SAD requires
-  as input the temperature range of interest, in contrast to
-  $1/t$-WL, which requires that the user identify the interesting range
-  of energies.
-  %
-  The convergence of the $1/t$-WL method is very sensitive to the energy
-  range chosen for the low-temperature heat capacity of the
-  Lennard-Jones cluster.
-  %
-  Thus, SAD is more powerful in the common case in which the range
-  of energies is not known in advance.
-\end{abstract}
-
-\maketitle
+% \affiliation{Department of Physics, Oregon State University,
+%   Corvallis, OR 97331}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{N50-lndos-comparison}
-  \caption{The entropy of a square-well fluid with 50 atoms and filling fraction
+  \includegraphics[width=\textwidth]{../square-well/N50-lndos-comparison.pdf}
+  \caption[The entropy of a square-well fluid with 50 atoms and filling fraction
+        0.3 as the green line.]{The entropy of a square-well fluid with 50 atoms and filling fraction
         0.3 as the green line.  The green hatched area reflects the
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{gamma-n256}
+  \includegraphics[width=\textwidth]{../square-well/gamma-n256.pdf}
   \caption{The update factor $\gamma_t$ versus iteration for the
@@
 
-%~ \paragraph{Scaling, step size, and SAD}
-%~ Methods such as WL can fail to converge for complex systems in part
-%~ because the same random MC moves are performed for the entire energy
-%~ range.  The main problem lies in trying to determine how to optimize
-%~ the step size such that the time to explore all energies is reasonable.
-%~ Small step sizes lead to improved sampling at the expense of slow
-%~ convergence. Koh et al. proposed multiplying $\gamma$ by an energy
-%~ dependent factor in order to improve the converge of WL in regards to
-%~ convergence and scaling~\cite{koh2013dynamically}.  Because the update
-%~ factor for SAD has an energy dependent factor already present (and
-%~ dynamically adjusted), we expect that SAD will perform consitently
-%~ scaling free for a reasonable choice of step size.  SAD should perform
-%~ extremely well compared with other flat histogram methods when applied
-%~ to complex continuous systems such as polymers/proteins and spin models.
-
-%~ \subsection{Convergence}
-%~ Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
-%~ convergence criteria.  However, we also want to ensure that SAD avoids
-%~ the exponentially poor convergence time that SAMC suffers when $t_0$ is
-%~ small as discussed in Section~\ref{sec:samc-convergence}.
-%~ To place a lower bound on the convergence of SAD, we will again
-%~ approximate the total entropy change using an integral.  To sidestep
-%~ the dynamical nature of SAD, we ensure that \emph{after discovering the
-%~ last important energy} at time $t_L$ the entropy change
-%~ $\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
-%~ Thus we integrate the update factor from $t_L$ to some minimum time
-%~ $t_{\min}$ necessary for convergence.  Technically this is not a lower
-%~ bound as in the case of SAMC because the algorithm will already have a
-%~ reasonable approximation for the entropy.
-%~ \begin{align}
-%~ \Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
-     %~ \frac{
-       %~ 3\langle S\rangle + \frac{t}{t_L}
-     %~ }{
-       %~ 3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
-     %~ } dt \\
-%~ &\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
-%~ \end{align}
-%~ where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
-%~ that it takes a long time to find all the interesting energy states.
-%~ Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
-%~ of order unity.  This omission only makes our bound more conservative,
-%~ and enables us to solve to find that
-%~ \begin{align}
-  %~ t_{\min} &= \frac32 t_L
-%~ \end{align}
-%~ which means that we only need to run fifty percent longer after we
-%~ discover the final interesting energy state in order to have made
-%~ sufficient updates to \emph{possibly} converge to the true entropy.
-
 \section{Results}\label{sec:results}
@@
 \begin{figure*}
-  \includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
-\includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
-  \caption{(a) The average entropy error for each MC method for $N=50$,
+  \includegraphics[width=0.5\textwidth]{../square-well/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
+\includegraphics[width=0.5\textwidth]{../square-well/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
+  \caption[(a) The average entropy error for each MC method for $N=50$,
+               $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
+               as a function of number of iterations run.]{(a) The average entropy error for each MC method for $N=50$,
                $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
@@
 simulations run with different random number seeds.
-%~ \begin{align}
-%~ \epsilon_\text{avg} =
-  %~ \frac{1}{N_S \text{num seeds}}\sum_s^{\text{num seeds}}\sum_{i}^{N_S} \left|S_\text{ref}(E_i) - S_\text{method}(E_i)\right|
-%~ \end{align}
 
@@
 
-%~ SAMC methods that converge
-%~ in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not
-%~ necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.
-%
-%~ Sad performs similarly regardless of the displacement distance chosen although
-%~ it transitions to $1/\sqrt{t}$ convergence after
-
 \begin{figure}
-\includegraphics[width=\columnwidth]{periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
-  \caption{
+\includegraphics[width=\textwidth]{../square-well/periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
+  \caption[The average entropy error for each MC method for $N=256$,
+               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
+               as a function of number of iterations run.]{
   The average entropy error for each MC method for $N=256$,
@@
 
-%$395.2833013702813 - 4.754731035740479e^{-8}$ = $395.2833$ so $2^{396.2833}$
-%gives us about $10^{118}$ from 'samc-1e4-256-s8.yaml'
-
 Figure~\ref{fig:n256} shows the average error in the entropy as a
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{lj-cv}
-  \caption{The heat capacity of the LJ31 cluster after $10^{12}$ moves
+  \includegraphics[width=\textwidth]{../square-well/lj-cv.pdf}
+  \caption[The heat capacity of the LJ31 cluster after $10^{12}$ moves
+    as calculated with each method.]{The heat capacity of the LJ31 cluster after $10^{12}$ moves
     as calculated with each method.  Each simulation used a bin size
@@
 
-%% \jpcom{This \P\ should be commented once summarized.
-%% Poulain \emph{et al.} found very poor convergence for a variety of
-%% modified WL methods on the LJ31
-%% cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
-%% energy range or bin size that was used in their simulation, so their
-%% results cannot be reproduced.  They do mention that they use 1000
-%% energy bins, which if they used an energy range of
-%% $[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
-%% E\approx 0.13\epsilon$, which we find to be too large to accurately
-%% predict the heat capacity peak at $0.03\epsilon/k_B$.  Looks like they
-%% only ran $\sim 10^{10}$ moves. We can't converge in that few moves either.}
-
 Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{lj-cv-error}
-  \caption{The maximum error in the heat capacity of the LJ31 cluster
+  \includegraphics[width=\textwidth]{../square-well/lj-cv-error.pdf}
+  \caption[The maximum error in the heat capacity of the LJ31 cluster
+    in the temperature range 0.01~$\epsilon\le k_BT \le 0.05\epsilon$]{The maximum error in the heat capacity of the LJ31 cluster
     in the temperature range 0.01~$\epsilon\le k_BT \le 0.05\epsilon$.
@@
 powerful when the range of energies is not known in advance.
-
-%\section{Acknowledgement}
-
-%We gratefully acknowledge D.W. Siderius and
-%J.R. Errington for helpful communications regarding the implementation
-%of WL-TMMC.
-
-\bibliography{paper}% Produces the bibliography via BibTeX.
-
-\end{document}

### 2020-01-21 thesis/chapter-3.tex
added thesis to the repository.

--- 2020-01-16 ising/ising.tex
@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
-\usepackage{graphicx} % Include figure files
-\usepackage{color}
-\usepackage{nicefrac} % Include for inline fractions
-
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
-\usepackage{mdframed}
-\usepackage{braket}
-
-% define colors for comments
-\definecolor{dark-gray}{gray}{0.10}
-\definecolor{light-gray}{gray}{0.70}
-
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
-
-\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
-\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
-\begin{document}
-
-\title{Flat histogram method comparison on 2D Ising Model
+\chapter{Flat histogram method comparison on 2D Ising Model
 }
 
-\author{Jordan K. Pommerenck} \author{David Roundy}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-
-\begin{abstract}
-  We examine the convergence of the Stochastic Approximation with a Dynamic
-  update factor (SAD) algorithm applied to the 2D ising model. Comparison with
-  SAMC and WL methods show that SAD performs robustly and without user input
-  knowledge of an energy range. We confirm that SAD is more powerful in the
-  common case in which the range of energies is not known in advance.
-\end{abstract}
-
-\maketitle
+\section{Contribution of Authors}
+Jordan K. Pommerenck, Tanner T. Simpson, Michael A. Perlin, David J. Roundy
+% \affiliation{Department of Physics, Oregon State University,
+%   Corvallis, OR 97331}
 
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{ising/gamma-n32.pdf}
+\includegraphics[width=\textwidth]{../ising/gamma-n32.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{ising/N32-entropy-error-default.pdf}
-  \caption{
+\includegraphics[width=\textwidth]{../ising/N32-entropy-error-default.pdf}
+  \caption[The average entropy error for each MC method for $N=32 \times 32$,
+               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+               as a function of number of iterations run.]{
   The average entropy error for each MC method for $N=32 \times 32$,
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{ising/gamma-n128.pdf}
+\includegraphics[width=\textwidth]{../ising/gamma-n128.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{ising/N128-entropy-error-default.pdf}
-  \caption{
+\includegraphics[width=\textwidth]{../ising/N128-entropy-error-default.pdf}
+  \caption[The average entropy error for each MC method for $N=128 \times 128$,
+               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+               as a function of number of iterations run.]{
   The average entropy error for each MC method for $N=128 \times 128$,
@@
 considerably more convenient.
-
-\bibliography{ising/ising} % Produces the bibliography via BibTeX.
-
-\end{document}

### 2020-01-21 thesis/pretext.tex
added thesis to the repository.

%\thispagestyle{plain}
\thispagestyle{empty}
\begin{center}
	\large
	AN ABSTRACT OF THE THESIS OF
\end{center}

\justify{}
\underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
presented on \underline{\jpsays{include the correct date here!}}.
\vspace{1.0cm}

\justify{}
Title: \underline{\jpsays{Title of the thesis}}.
\vspace{2.0cm}

\justify{}
Abstract approved:
\underline{\hspace{11cm}}
\vspace{0.5cm}

%\centering{}
\hspace{7.0cm} David J. Roundy
\vspace{1.0cm}

\justify{}
\doublespacing
"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip
ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore 
eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia
deserunt mollit anim id est laborum."

\iffalse
\singlespacing
"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt..."
\fi

\newpage{}
\thispagestyle{empty}
\singlespacing

\vspace*{4.0cm}
\begin{center}
$\copyright$ Copyright by Jordan K. Pommerenck \\
\jpsays{include the correct date here!} \\
(All Rights Reserved)
\end{center}

\newpage{}
\thispagestyle{empty}
\singlespacing

\begin{center}
\jpsays{Title of the thesis}

\vspace{1.0cm}
by \\
Jordan K. Pommerenck \\
\vspace{3.0cm}
A DISSERTATION \\
\vspace{0.5cm}
submitted to \\
\vspace{0.5cm}
Oregon State University \\
\vspace{3.0cm}
in partial fulfillment of \\
the requirements for the \\
degree of \\
\vspace{1.0cm}
Doctor of Philosophy \\
\vspace{3.0cm}
Presented October 26, 2019  \\
Commencement December 2019 \jpsays{include the correct date here!}
\end{center}

\newpage{}
\thispagestyle{empty}
\singlespacing
\justify{}
\underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
\jpsays{get the date right here? August 26, 2015}. \\

\justify{}
\vspace{0.5cm}
APPROVED: \\

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Major Professor, representing Physics
\vspace{1.0cm}
\end{flushleft}

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Chair of the Department of Physics
\vspace{1.0cm}
\end{flushleft}

\justify{}
\underline{\hspace{15cm}}
\begin{flushleft}
Dean of the Graduate School
\vspace{1.0cm}
\end{flushleft}

\vspace{3.0cm}
\justify{}
I understand that my dissertation will become part of the permanent collection
of Oregon State University libraries. My signature below authorizes release of my
dissertation to any reader upon request. \\
\vspace{0.5cm}
\begin{flushleft}
\underline{\hspace{15cm}}
\end{flushleft}
\centering{}
Jordan K. Pommerenck, Author

\newpage{}
\thispagestyle{empty}
\begin{center}
	\large
	ACKNOWLEDGEMENTS
\end{center}
\justify{}
\doublespacing
"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
dolore magna aliqua. Ut

### 2020-01-21 thesis/thesis.tex
added thesis to the repository.

\documentclass[12pt,a4paper]{report}
\usepackage{fancyhdr}

% for creating an index at end of document
\usepackage{imakeidx}
\makeindex[columns=1, title=Alphabetical Index]

% indexing commands
\newcommand{\bind}[1]{\textbf{#1}\index{#1}}
\newcommand{\iind}[1]{\textit{#1}\index{#1}}
\newcommand{\ind}[1]{#1\index{#1}}

% link to section command
\newcommand{\HR}[1]{\textbf{\hyperref[#1]{\ref{#1}}}}

\usepackage[hypertexnames=false]{hyperref}
\hypersetup{
    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=true,        % show Acrobat’s toolbar?
    pdfmenubar=true,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={Pommerenck Thesis},    % title
    pdfauthor={Jordan K. Pommerenck},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1, key2, key3}, % list of keywords
    pdfnewwindow=true,      % links in new PDF window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
    citecolor=black,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}

\pagestyle{fancy}
\fancyhf{}
\fancyheadoffset{0cm}
\renewcommand{\headrulewidth}{0pt} 
\renewcommand{\footrulewidth}{0pt}
\fancyhead[R]{\thepage}
\fancypagestyle{plain}{%
   \fancyhf{}%
   \fancyhead[R]{\thepage}%
}

\usepackage{blindtext}
\usepackage{parskip}

\usepackage[left=1.5in,right=1in,top=1in,bottom=1in,includefoot,heightrounded]{geometry}
\usepackage[pagestyles]{titlesec}

\sethead{\itshape\chaptername\thechapter}{}{\itshape\chaptertitle}
\titleformat{\chapter}[display]{\normalfont\bfseries\Large\centering}{}{18pt}{\Large}
\titlespacing{\chapter}{0pt}{0pt}{1cm}

\usepackage{blindtext}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[document]{ragged2e}
% differential and partials package
\usepackage[thinc]{esdiff}
\usepackage{physics}
\newcommand{\D}[1]{\text{d}#1}
% the nth superscript package
\usepackage[super]{nth}

\usepackage{indentfirst}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\usepackage{amsmath,amssymb}
\usepackage{nicefrac}

\usepackage[backend=bibtex,maxnames=200,sorting=none]{biblatex}
%\usepackage[square,numbers]{natbib}

\addbibresource{thesis.bib}

% All the packages I have added...
\usepackage{setspace} % Allows double spacing with the \doublespacing command
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{\thepage}

% \usepackage{color}
% \newcommand{\red}[1]{{\bf \color{red} #1}}
% \newcommand{\green}[1]{{\bf \color{green} #1}}

% \newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
% \newcommand{\jpsays}[1]{{\color{red} [\red{Jordan:} \emph{#1}]}}

\author{Jordan K. Pommerenck}

% rename Contents to ToC and center
\usepackage{tocloft}
\renewcommand{\cftchapfont}{\normalfont}
\renewcommand{\cftchappagefont}{\normalfont}
\renewcommand{\contentsname}{\hfill\Large\normalfont TABLE OF CONTENTS}
\renewcommand{\cftaftertoctitle}{\hfill}
\renewcommand{\listfigurename}{\hfill\Large\normalfont LIST OF FIGURES\hfill}
\renewcommand{\cftaftertoctitle}{\hfill}

\begin{document}
\input{pretext.tex}

\clearpage{}
\tableofcontents
\thispagestyle{empty}

\clearpage{}
\listoffigures
\thispagestyle{empty}

\clearpage{}

\setcounter{page}{1} % needed if pages are to be continuously numbered

\input{chapter-1}
%\input{Chapters/Chapter2_Files/sad.tex}
%\input{Chapters/Chapter3_Files/ising.tex}

%\chapter{Conclusion}
%\input{Chapters/chapter06}

\printbibliography[title={Bibliography}]

\clearpage{}
%\printindex
\end{document}

### 2020-01-28 thesis/chapter-1.tex
update thesis bib and introduction.

@@
 
-% \begin{enumerate}
-%     \item Chapter 1 – General Introduction Links all manuscripts
-%  \item Chapter 2 – First Manuscript
-%  \item Chapter 3 – Second Manuscript
-%  \item Chapter 4 – Third Manuscript
-%  \item Chapter 6 – General Conclusion (common conclusion linking all manuscripts
-% thematically)
-%  \item Common bibliography covering all manuscripts, although each manuscript may have its own reference section
-% \end{enumerate}
+Monte Carlo methods are a powerful simulation tool that allow scientists to explore properties of systems that may otherwise by inaccessible via experimental methods.
+Monte Carlo methods can be used to solve a variety of physical problems yielding insight into statistical averages due to random processes.
+During the last two decades considerable research has been done on developing and refining flat-histogram Monte Carlo methods.
+These Monte Carlo (MC) methods are capable of exploring all of the energy phase space for all given temperatures in a \emph{single} simulation.
+In this work, we first outline the necessary statistical mechanics background.
+We provide a description of MC methods and flat-histogram methods, as well as, our newly developed method SAD (stochastic approximation Monte Carlo with a dynamic update factor).
+Finally, we discuss physics rich test systems that can be used to explore the convergence and statistical properties of our newly developed method SAD.
+In the following chapters, we follow the manuscript outline with each chapter corresponding to our published or submitted research on the SAD method applied to various physical systems.
 
@@
 \end{align}
-From the \nth{1} law of thermodynamics, the system temperature can be written in terms of the density of states by refering to Eq.~\HR{sdos}.
+From the \nth{1} law of thermodynamics, the system temperature can be written in terms of the density of states by referring to Eq.~\HR{sdos}.
 \begin{align}
@@
 \end{align}
-The relationship between entropy $S(E)$, $\Omega(E)$, and $T$ will be discussed in further depth in sections \HR{laws} and \HR{weights}.
+The relationship between entropy $S(E)$, $\Omega(E)$, and $T$ will be discussed in further depth in Section~\HR{laws}.
 
@@
 Maxwell's relations are second derivative equalities that relate important thermodynamic variables.
-They are presented here to facilitate a critical derivation in Section~\HR{gas-ads}.
 Maxwell relations are an extension of the Schwarz theorem which states that the order of differentiation is irrelevant.
@@
 A major driving force behind introducing the concepts of thermal expansion and compressibility is to right complex derivatives in terms of these parameters.
-This will be invaluable as will be shown in Section~\HR{gas-ads}.
 The internal energy derivative and entropy derivative can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
@@
 \section{Monte Carlo Simulation}\label{monte}
+Experiment and simulation blend together to help form a more complete understanding of our knowledge of how the universe works.
+Computational `experiments' such as Monte Carlo methods aid in this understanding by allowing scientists to explore system configurations that are not easily accessible to traditional experiment.
 
-% talk about Canonical here
-\subsection{Broad Histogram Methods}\label{bhm}
-\subsection{Weights, multiplicity, detailed balance}\label{weights}
+In the field of statistical physics, Monte Carlo methods are used to determine equilibrium 
+properties (thermal averages) of multi-particle systems.  A number of Monte Carlo simulations 
+have been developed over the last century. Some of these methods are completely unique while 
+others are more sophisticated variants of an original method. All of the methods share a number 
+of similarities including picking random moves (we will discuss this process in Section~\HR{rng}) and determining the acceptance or rejection of a 
+move based on a probability distribution. They differ primarily in how they explore phase space 
+and thus compute thermodynamic properties such as the density of states $\Omega(E)$. The accuracy of any MC simulation depends on the details of how the system is `explored'. To reduce the error 
+in simulation one can in principle simply continue to allow the simulation more time to explore 
+phase space.
 
-\subsection{System Details}
-% Ising, Square well, Lennard-Jones clusters, MOF?
+The motivation for new research into developing MC methods stems from the desire to solve ever-increasingly complex physical problems in a faster simulation time.
+Many of the originally developed MC methods are unable to effectively explore the complex phase space of real (and even many test) systems.
+Flat-histogram methods (introduced in Section~\HR{fhm}) are able to explore all energies and temperatures in a `single' simulation.
+In addition, they are able to explore complex energy landscapes without getting `stuck' (either never converging or converging at such a slow rate as to not practically converge).
+Many of these flat-histogram methods also have proven convergence on the statistical error thus guaranteeing convergence.
+In the following sections, we will discuss a number of MC methods including importance sampling canonical methods, and flat-histogram methods leading up to our method SAD introduced in this work discussed in Section~\HR{fhm}.
 
-\section{Real System Simulation}
-% brief outline for broad histogram methods and how the motivate study of MOFs
+\subsection{Random Number Generation}\label{rng}
 
-\subsection{Metal-organic Frameworks (MOFs)}
-\subsection{Gas Adsorption}\label{gas-ads}
-% qst
+Since Monte Carlo moves are inherently stochastic in nature, it is imperative that any useful random number generator be fast and ideally non-deterministic.
+In practice, pseudo-random number generators (PRNGs) are used in order to generate the fast random streams although they are deterministic.
+In order for a PRNG to be acceptable for use in Monte Carlo simulation it must meet certain criteria~\cite{sawilowsky2003you}.
+First, the PRNG must have a long `period' before repeating values.
+Poor PRNGs (that is ones with short periods) can lead to systematic error in Monte Carlo simulations~\cite{ferrenberg1992monte, plascak2002cluster, ossola2004systematic}.
+Second, the PRNG must actually be able to pass certain statistical tests such as implemented in the TestU01 library~\cite{l2007testu01}.
+
+There are a variety of PRNGs each with benefits and few without some form of drawback.
+We detail some of the well known PRNGs here.
+Linear congruential generators (LCGs) have a very fast time performance and are straightforward to program (in terms of size and complexity).
+Unfortunately, LCGs have multiple issues regarding the generated random number statistical quality~\cite{coddington1994analysis, barry1996recommendations}.
+While the period of the PRNG can be made acceptable by using 64-bit integers, the statistical correlations remain.
+Other PRNGs are the belong to the shift-register family (Xorshift RNGs).
+Like the LCG family, the shift-register RNGs fail many statistical tests~\cite{panneton2005xorshift} and are therefore often considered unacceptable for use in Monte Carlo simulations.
+
+In our works, we use the \textbf{Xoroshiro128+} RNG which is named after its logical operations (XOR $\xrightarrow{}$ rotate $\xrightarrow{}$ shift $\xrightarrow{}$ rotate).
+The RNG has ultra-fast speed 0.72 $\nicefrac{\text{ns}}{\text{64-bits}}$ and incredible statistical quality~\cite{blackman2018scrambled}.
+
+\subsection{Canonical Monte Carlo}\label{cmc}
+In this section, we introduce canonical Monte Carlo and how it relates to flat-histogram methods introduced in Section~\HR{fhm}.
+The first importance sampling method developed was the famous Metropolis method~\cite{metropolis1953equation}.
+The method works by choosing an initial state for the system $\Gamma$.
+Next, a random site $\Gamma_i$ is chosen and a move is proposed that will result in an energy change $\Delta E$ if accepted.
+The acceptance of the move is determined by whether a random number chosen on a range from zero to one is less than a Boltzmann factor.
+The process of moves and random number generation is repeated until some termination condition is met.
+This canonical method performs calculations at a fixed temperature $T$ only.
+Re-weighting can extrapolate in the vicinity of the simulated temperature, but this approach is limited.
+Another limitation of canonical Monte Carlo is that for complex energy landscapes and/or large systems the simulation can undergo a phenomena known as `critical slowing down' as the simulation will never or almost never visit the ground state energy $E_{\min}$~\cite{berg2000introduction}.
+These limitations are what drive the need for methods that can avoid getting `stuck' and that can simulate all temperatures in a single simulation.
+
+\subsection{Flat-Histogram Methods}\label{fhm}
+The goal of flat-histogram methods (also called \emph{broad-histogram}
+or \emph{multicanonical} methods) is to simulate each energy with
+similar accuracy.  In this way, they can accurately determine the density of states
+over a broad range of energies and thus determine the thermodynamic
+quantities such as heat capacity or internal energy
+over a broad range of temperatures.
+Properties that require more information, such as a spatial
+correlation function or a response function, can still be computed
+for any temperature, provided statistics are collected for each
+individual energy, which can then be reweighted for any
+temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
+errington2003direct}.
+All the flat-histogram Monte Carlo methods begin with randomly chosen
+`moves' which change the state of the system and must satisfy
+detailed balance.  All of the tested algorithms differ in how they determine the
+probability of accepting a move and in what additional statistics must
+be collected in order to decide on that probability.
+Flat-histogram methods calculate the density of states $D(E)$ for a discrete set
+of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
+trebst2004optimizing}. Therefore, energy binning becomes an important
+consideration for systems with a continuum of possible energies.  Energy bins
+are typically of uniform size for the entire energy
+continuum~\cite{fasnacht2004adaptive}.
+
+All of the closely related flat-histogram
+methods rely on a weight function $w(E)$.  In these
+algorithms, the probability of accepting a move is given by
+\begin{equation}
+	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
+	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
+\end{equation}
+which biases the simulation in favor of energies with low weights.  A set of
+weights that are proportional to the density of states $D(E)$ of the system will
+result in an entirely flat histogram, which means that the weights should
+converge to being proportional to the density of states.  The natural logarithm
+of the weight is typically stored, since the density of states will often vary
+over a few hundred orders of magnitude. In the microcanonical ensemble, the
+entropy is defined as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the
+logarithm of the weight is an approximation of the entropy.
+
+Each approach uses a random walk in energy space to estimate the density of
+states.  The core of these approaches is to update the weights at each step of
+the simulation
+\begin{equation}
+	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
+	+\gamma_t
+\end{equation}
+where $t$ is the number of the current move, $\gamma_t$ is an update factor
+that varies over the simulation, and $E$ is the current energy.  This update
+causes the random walk to avoid energies that have been frequently sampled,
+enabling a rapid exploration of energy space. This approach violates detailed
+balance, due to the acceptance probabilities changing with each move, but the
+severity of this violation decreases as we decrease $\gamma_t$.  The four
+methods differ primarily in how they schedule the decrease of $\gamma_t$.
+
+In this work, we develop a new method called Stochastic Approximation with Dynamic update factor (SAD) method which is a variant of the SAMC Algorithm that attempts to dynamically choose
+the modification factor rather than relying on system dependent
+parameters such as $t_0$ or $\gamma_0$.  There is an immediate
+advantage of such an algorithm where parameters are chosen independent
+of system size or type. Our
+proposed method SAD requires the user to input $T_\text{min}$, the
+lowest temperature of interest, which is an immediate disadvantage of
+the method. However, identifying a minimum temperature of interest
+$T_\text{min}$ may be easier for a user than determining in advance
+an energy range of interest or the unphysical parameter $t_0$ such as other
+flat-histogram methods must do.
+
+\section{Test System Details}
+We introduce a number of systems that are suitable for benchmarking the convergence properties of various flat-histogram MC methods.
+Testing on a variety of systems is necessary for ensuring the robustness of a given MC method.
+The square-well fluid is chosen as a test-bed due to the fact that convergence properties of flat-histogram weight-based MC methods have never been studied on this system.
+Secondly, the MC method will have to simulate the potentially complex liquid-vapor phase coexistence of the system.
+The 2D Ising model is a simple system with a second order phase transition.
+In addition, the majority of MC methods have been tested on the Ising model so the speed and accuracy with which a given MC method can converge can often be directly compared with the literature.
+A Lennard-Jones cluster is a heavily studied system that can be simple or incredibly complex for a MC method to accurately solve.
+In order to solve for solid-solid transitions for certain cluster configurations, it is necessary to explore incredibly low reduced temperatures.
+This can make finding the global minima configuration extremely difficult for MC methods in general.
+
+\subsection{Square-Well Fluid}
+The square-well fluid is a system of particles whose interactions are governed by a square-well potential~\cite{singh2003surface, barker2004perturbationSW}.
+The square-well potential is an ideal test-bed for Monte Carlo methods as it is a simple model for a liquid, which includes both attractive and repulsive interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
+The potential $U(\textbf{r})$ for such a system is given by
+\begin{equation}
+ U(\textbf{r})=\begin{cases} \infty &
+ \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
+ \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
+ \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
+\end{equation}
+where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the reduced range of the potential well, and $\epsilon$ is its depth.
+
+\subsection{2D Ising Model}
+The 2D Ising spin-lattice system is widely used as a test-bed when benchmarking or comparing Monte-Carlo methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}.
+The \nth{2} order phase transition behavior and the ability to directly calculate the exact solution for finite lattices~\cite{beale1996exact} make the system sufficiently interesting for such theoretical comparisons.
+The Hamiltonian for the periodic 2D square lattice ferromagnetic Ising model with nearest
+neighbor interactions~\cite{landau2004new} cab be written as the following:
+\begin{align}
+\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
+\end{align}
+The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
+for up or down spins respectively. In the absence of a magnetic field ($h =
+0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
+kaufman1949crystal}:
+\begin{align}
+\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
+\end{align}
+where the sum is over nearest neighbor spin sites.
+
+\subsection{Lennard-Jones Clusters}
+Lennard-Jones cluster systems are widely used as a challenging test-bed when comparing Monte-Carlo methods~\cite{doye1998thermodynamics, poulain2006performances, martiniani2014superposition}.
+The potential energy for interacting atomic particles can be written as follows:
+\begin{align}
+U = 4\epsilon \left[\left( \sigma/r \right)^{12} - \left( \sigma/r \right)^{6}\right]
+\end{align}
+A number of challenges exit when trying to accurately simulate the density of states of a cluster system.
+Narrow potential energy funnels~\cite{mandelshtam2006multiple, wales1997global} and low temperature solid-solid transitions~\cite{neirotti2000phase, calvo2000phase, calvo2000entropic} make predicting thermodynamic properties such as heat capacity an extremely difficult process.
+In addition, since flat-histogram methods explore all energies and all temperatures it is possible that for high temperatures a large cluster volume is possible (which could lead to boil-off of the particles).
+A `constraining' sphere with radius $R_c$ is used to prevent this but it is usually empirically determined further complicating the simulation process.
 
@@
 
-% Tie everything in together here!
+We have provided the background necessary for understanding the various flat-histogram methods and the various applied test systems.
+In the following chapters, we detail our manuscripts and the method SAD applied to the various physical test systems.
+We test SAD against a variety of other weight-based flat-histogram methods on the square-well fluid, Lennard-Jones clusters, and the 2D Ising model.
+Each of these systems provides unique challenges to our newly developed method SAD and convergence robustness is determined best by the application to multiple systems.

### 2020-02-07 thesis/chapter-1.tex
working on getting thesis to compile. Still some encoding character error.
added new ising scripts and chapter section in thesis.

@@
 We provide a description of MC methods and flat-histogram methods, as well as, our newly developed method SAD (stochastic approximation Monte Carlo with a dynamic update factor).
-Finally, we discuss physics rich test systems that can be used to explore the convergence and statistical properties of our newly developed method SAD.
-In the following chapters, we follow the manuscript outline with each chapter corresponding to our published or submitted research on the SAD method applied to various physical systems.
+We discuss physics rich test systems that can be used to explore the convergence and statistical properties of our newly developed method SAD.
+
+A driving motivation for developing novel Monte Carlo methods that have physically based tunable parameters is to provide simulation methods for exploring metal-organic frameworks (MOFs).
+In this work, we examine MOFs and
+their application to gas storage and delivery. We formulate a theoretical upper
+bound on the maximum amount of gas that can be stored and released at different
+pressures.
+We propose SAD as a powerful method for examining thermodynamic properties for MOFs.
+We lay a groundwork for 2D SAD which would be able to compute isotherms for any given MOF thereby providing a powerful way for researchers to determine a MOF's suitability for gas storage applications.
+
+In the following chapters, we follow the manuscript outline with each chapter corresponding to our published or submitted research on the SAD method applied to various physical systems and our work on establishing a theoretical upper bound on gas storage and delivery for MOFs.
 
@@
 
-In this work, we develop a new method called Stochastic Approximation with Dynamic update factor (SAD) method which is a variant of the SAMC Algorithm that attempts to dynamically choose
+In this work, we develop a new method called Stochastic Approximation with Dynamic update factor (SAD) method
+which is a variant of the SAMC Algorithm that attempts to dynamically choose
 the modification factor rather than relying on system dependent
@@
 \section{Test System Details}
-We introduce a number of systems that are suitable for benchmarking the convergence properties of various flat-histogram MC methods.
-Testing on a variety of systems is necessary for ensuring the robustness of a given MC method.
-The square-well fluid is chosen as a test-bed due to the fact that convergence properties of flat-histogram weight-based MC methods have never been studied on this system.
-Secondly, the MC method will have to simulate the potentially complex liquid-vapor phase coexistence of the system.
-The 2D Ising model is a simple system with a second order phase transition.
-In addition, the majority of MC methods have been tested on the Ising model so the speed and accuracy with which a given MC method can converge can often be directly compared with the literature.
-A Lennard-Jones cluster is a heavily studied system that can be simple or incredibly complex for a MC method to accurately solve.
-In order to solve for solid-solid transitions for certain cluster configurations, it is necessary to explore incredibly low reduced temperatures.
-This can make finding the global minima configuration extremely difficult for MC methods in general.
+We introduce a number of systems that are suitable for benchmarking the convergence properties of various
+flat-histogram MC methods. Testing on a variety of systems is necessary for ensuring the robustness of a
+given MC method. The square-well fluid is chosen as a test-bed due to the fact that convergence properties
+of flat-histogram weight-based MC methods have never been studied on this system. Secondly, the MC method
+will have to simulate the potentially complex liquid-vapor phase coexistence of the system. The 2D Ising
+model is a simple system with a second order phase transition. In addition, the majority of MC methods have
+been tested on the Ising model so the speed and accuracy with which a given MC method can converge can often
+be directly compared with the literature. A Lennard-Jones cluster is a heavily studied system that can be
+simple or incredibly complex for a MC method to accurately solve. In order to solve for solid-solid transitions
+for certain cluster configurations, it is necessary to explore incredibly low reduced temperatures. This can
+make finding the global minima configuration extremely difficult for MC methods in general.
 
 \subsection{Square-Well Fluid}
-The square-well fluid is a system of particles whose interactions are governed by a square-well potential~\cite{singh2003surface, barker2004perturbationSW}.
-The square-well potential is an ideal test-bed for Monte Carlo methods as it is a simple model for a liquid, which includes both attractive and repulsive interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
-The potential $U(\textbf{r})$ for such a system is given by
+The square-well fluid is a system of particles whose interactions are governed by a square-well
+potential~\cite{singh2003surface, barker2004perturbationSW}.  The square-well potential is an ideal test-bed
+for Monte Carlo methods as it is a simple model for a liquid, which includes both attractive and repulsive
+interactions~\cite{barker1967-SW-perturbation, vega1992phase}. The potential $U(\textbf{r})$ for such a system
+is given by
 \begin{equation}
@@
 \subsection{2D Ising Model}
-The 2D Ising spin-lattice system is widely used as a test-bed when benchmarking or comparing Monte-Carlo methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}.
-The \nth{2} order phase transition behavior and the ability to directly calculate the exact solution for finite lattices~\cite{beale1996exact} make the system sufficiently interesting for such theoretical comparisons.
+The 2D Ising spin-lattice system is widely used as a test-bed when benchmarking or comparing Monte-Carlo
+methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The \nth{2} order phase
+transition behavior and the ability to directly calculate the exact solution for finite
+lattices~\cite{beale1996exact} make the system sufficiently interesting for such theoretical comparisons.
 The Hamiltonian for the periodic 2D square lattice ferromagnetic Ising model with nearest
@@
 \subsection{Lennard-Jones Clusters}
-Lennard-Jones cluster systems are widely used as a challenging test-bed when comparing Monte-Carlo methods~\cite{doye1998thermodynamics, poulain2006performances, martiniani2014superposition}.
-The potential energy for interacting atomic particles can be written as follows:
+Lennard-Jones cluster systems are widely used as a challenging test-bed when comparing Monte-Carlo
+methods~\cite{doye1998thermodynamics, poulain2006performances, martiniani2014superposition}. The potential
+energy for interacting atomic particles can be written as follows:
 \begin{align}
@@
 A number of challenges exit when trying to accurately simulate the density of states of a cluster system.
-Narrow potential energy funnels~\cite{mandelshtam2006multiple, wales1997global} and low temperature solid-solid transitions~\cite{neirotti2000phase, calvo2000phase, calvo2000entropic} make predicting thermodynamic properties such as heat capacity an extremely difficult process.
-In addition, since flat-histogram methods explore all energies and all temperatures it is possible that for high temperatures a large cluster volume is possible (which could lead to boil-off of the particles).
-A `constraining' sphere with radius $R_c$ is used to prevent this but it is usually empirically determined further complicating the simulation process.
+Narrow potential energy funnels~\cite{mandelshtam2006multiple, wales1997global} and low temperature solid-solid
+transitions~\cite{neirotti2000phase, calvo2000phase, calvo2000entropic} make predicting thermodynamic properties
+such as heat capacity an extremely difficult process. In addition, since flat-histogram methods explore all
+energies and all temperatures it is possible that for high temperatures a large cluster volume is possible
+(which could lead to boil-off of the particles). A `constraining' sphere with radius $R_c$ is used to prevent
+this but it is usually empirically determined further complicating the simulation process.
+
+\section{Gas Storage and Delivery in Metal-Organic Frameworks}
+
+Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles~\cite{?}.
+One strategy to densify these gases is to pack the fuel tank with a porous material~\cite{?}.
+Metal-organic frameworks (MOFs) are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases~\cite{?}.
+The US Department of Energy (DOE) has set volumetric deliverable capacity targets~\cite{?} which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels.
+Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction~\cite{?}; none have established a rigorous upper bound on the deliverable capacity.
+
+Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing. 
+To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
+Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
+We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
+However, limitations to the scope of applicability of our upper bound guide future material development efforts.
+Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes~\cite{?}.
+Secondly, the fundamental physics of our upper bound do not pertain to
+any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target~\cite{?}.
+
+Consider a pressure vessel onboard a vehicle (i.e.
+fuel tank) packed with porous material.
+At the filling stage, the tank is connected and opened to a (pure) gaseous reservoir at pressure $p_\text{full}$ and allowed to equilibrate.
+At this point, the adsorbed gas tank is considered full.
+While driving, gas desorbs from the adsorbent to the engine/fuel cell, driven by a pressure differential.
+The tank is considered depleted/empty when the pressure has dropped to $p_\text{empty}$, the pressure at which the flow rate of gas from the tank to the engine is insufficient.
+
+However, given $p_\text{empty} \neq 0$ (pulling vacuum), residual gas will remain trapped in the adsorbent.
+Therefore, the driving range of the vehicle is primarily determined by the \emph{deliverable capacity}, defined as the density of gas in the tank at $p_\text{full}$ minus the residual gas remaining at $p_\text{empty}$.
+The [isothermal, volumetric] deliverable capacity is an intrinsic property of the nanoporous material and its interaction with the gas.
+
+\jpsays{explain how 2D SAD is needed for this and how we are working toward it}
 
 \section{Summary}
+We have provided the background necessary for understanding the various flat-histogram methods and the
+various applied test systems. In the following chapters, we detail our manuscripts and the method SAD
+applied to the various physical test systems. We test SAD against a variety of other weight-based flat-histogram
+methods on the square-well fluid, Lennard-Jones clusters, and the 2D Ising model. Each of these systems provides
+unique challenges to our newly developed method SAD and convergence robustness is determined best by the application
+to multiple systems.
 
-We have provided the background necessary for understanding the various flat-histogram methods and the various applied test systems.
-In the following chapters, we detail our manuscripts and the method SAD applied to the various physical test systems.
-We test SAD against a variety of other weight-based flat-histogram methods on the square-well fluid, Lennard-Jones clusters, and the 2D Ising model.
-Each of these systems provides unique challenges to our newly developed method SAD and convergence robustness is determined best by the application to multiple systems.
+\jpsays{Tie this in with gas adsorption!}

### 2020-02-07 scripts-python-make-data/ising-multi-heat-capacity.py

@@ -1,0 +1,190 @@
+#!/usr/bin/env python
+
+from __future__ import division
+import sys, os
+import numpy as np
+import readnew
+from glob import glob
+import matplotlib.pyplot as plt
+import yaml
+import os.path
+import time # Need to wait some time if file is being written
+import argparse
+
+parser = argparse.ArgumentParser(description =
+    """
+        This script functions by reading in yaml files from a given directory
+        and comparing with a given reference .DAT file. The heat capacity
+        is computed and compared to the reference.
+    """)
+
+parser.add_argument('--file_dir', required=True,
+                    help='The directory where the files are located. \
+                          Example:/home/jordan/ising-cp-data/')
+parser.add_argument('--reference', required=True,
+                    help='The directory where the reference file is located. \
+                          Exmaple:data/ising-32-reference-lndos.dat')
+parser.add_argument('--save_dir', required=True,
+                    help='The directory where the data will be saved. \
+                          Exmaple:data/comparison/N32')
+parser.add_argument('--N', type=int,
+                    help='The length of the ising model. Used to divide \
+                          the number of moves.')
+parser.add_argument('--Emin', type=int,
+                    help='The energy at the entropy minimum.')
+parser.add_argument('--Emax', type=int,
+                    help='The energy at the entropy maximum.')
+
+
+args = parser.parse_args()
+
+file_dir = args.file_dir                # change name from filename_location
+ref = args.reference                    # reference to ref
+save_dir = args.save_dir                # filebase
+
+
+
+# seed_avg = int(sys.argv[8])
+
+# filename = sys.argv[9:]
+# print(('filenames are ', filename))
+
+def heat_capacity(T, E, S):
+    """
+        This function calculates the heat capacity by taking in
+        the temperature 'T', energy 'E', and entropy 'S' as NumPy
+        arrays.  The number 'Num' is also necessary to complete
+        the heat capacity calculation.
+    """
+    C = np.zeros_like(T)
+    for i in range(len(T)):
+        boltz_arg = S - E/T[i]
+        P = np.exp(boltz_arg - boltz_arg.max())
+        P = P/P.sum()
+        U = (E*P).sum()
+        C[i] = ((E-U)**2*P).sum()/T[i]**2
+    return C
+
[…]

### 2020-02-07 scripts-python-make-data/readnew.py

@@ -1,0 +1,352 @@
+#!/usr/bin/python2
+
+import numpy
+
+def e_hist(fbase):
+    try:
+        trans = numpy.loadtxt(fbase +"-transitions.dat", dtype=numpy.float)
+        energy = -trans[:, 0]
+        trans = trans[:, 1:]
+        hist = numpy.sum(trans, axis=1)
+    except:
+        # energy histogram file; indexed by [-energy,counts]
+        e_hist = numpy.loadtxt(fbase+"-E.dat", ndmin=2, dtype=numpy.float)
+        energy = -e_hist[:, 0] # array of energies
+        hist = e_hist[:, 1]
+    return energy, hist
+
+def e_lndos(f):
+    if '.dat' not in f:
+        f = f+"-lndos.dat"
+    e_lndos = numpy.loadtxt(f, ndmin=2, dtype=numpy.float)
+    energy = -e_lndos[:, 0] # array of energies
+    lndos = e_lndos[:, 1]
+    return energy, lndos
+
+def e_lndos_ps(fbase):
+    if '.dat' not in fbase:
+        fbase = fbase + "-lndos.dat"
+    e_lndos_ps = numpy.loadtxt(fbase, ndmin=2, dtype=numpy.float)
+    energy = -e_lndos_ps[:, 0]
+    lndos = e_lndos_ps[:, 1]
+    ps = e_lndos_ps[:, 2] # pessimistic samples
+
+    return energy, lndos, ps
+
+def e_lndos_ps_lndostm(fbase):
+    if '.dat' not in fbase:
+        fbase = fbase + "-lndos.dat"
+    e_lndos_ps = numpy.loadtxt(fbase, ndmin=2, dtype=numpy.float)
+    energy = -e_lndos_ps[:, 0]
+    lndos = e_lndos_ps[:, 1]
+    ps = e_lndos_ps[:, 2] # pessimistic samples
+    lndostm = None
+    if len(e_lndos_ps[0,:]) >= 4:
+        lndostm = e_lndos_ps[:, 3]
+
+    return energy, lndos, ps, lndostm
+
+def e_lnw(fbase):
+    e_lnw = numpy.loadtxt(fbase+"-lnw.dat", ndmin=2, dtype=numpy.float)
+
+    energy = -e_lnw[:, 0] # array of energies
+    lnw = e_lnw[:, 1]
+    return energy, lnw
+
+def T_u_cv_s_minT(fbase):
+    max_T = 20.0
[…]

### 2020-02-07 scripts-python-make-data/ising-multi-heat-capacity.py
script now adds heat capacity for plotting.

@@ -16,6 +16,17 @@ parser = argparse.ArgumentParser(description =
         This script functions by reading in yaml files from a given directory
         and comparing with a given reference .DAT file. The heat capacity
         is computed and compared to the reference.
+        
+        EXAMPLE USAGE:
+        python3 scripts-python-make-data/ising-multi-heat-capacity.py
+        --file_dir=../ising-cp-data/
+        --reference=../deft/papers/histogram/data/ising-32-reference-lndos.dat
+        --save_dir=ising/data/comparison/N32
+        --filename ising-sad-32
+        --N=32
+        --Emin=2048
+        --Emax=0
+        --seed_avg=8
     """)
 
 parser.add_argument('--file_dir', required=True,
@@ -27,27 +38,36 @@ parser.add_argument('--reference', required=True,
 parser.add_argument('--save_dir', required=True,
                     help='The directory where the data will be saved. \
                           Exmaple:data/comparison/N32')
-parser.add_argument('--N', type=int,
+parser.add_argument('--N', type=int, required=True,
                     help='The length of the ising model. Used to divide \
                           the number of moves.')
-parser.add_argument('--Emin', type=int,
+parser.add_argument('--Emin', type=int, required=True,
                     help='The energy at the entropy minimum.')
-parser.add_argument('--Emax', type=int,
+parser.add_argument('--Emax', type=int, required=True,
                     help='The energy at the entropy maximum.')
-
+parser.add_argument('--seed_avg', type=int, required=True,
+                    help='The number of seed files to average over.')
+parser.add_argument('--filename', required=True, nargs='+',
+                    help='The filename to perform analysis on. \
+                          Exmaple:ising-sad-32')
 
 args = parser.parse_args()
 
-file_dir = args.file_dir                # change name from filename_location
-ref = args.reference                    # reference to ref
-save_dir = args.save_dir                # filebase
+# Rename all argparse parameters.
+file_dir = args.file_dir
+ref = args.reference
+save_dir = args.save_dir
+N = args.N
+Emin = args.Emin
+Emax = args.Emax
+seed_avg = args.seed_avg
 
-
-
-# seed_avg = int(sys.argv[8])
-
-# filename = sys.argv[9:]
-# print(('filenames are ', filename))
+# The correct way to handle accepting multiple arguments using nargs
+# to create a list.
[…]

### 2020-02-17 thesis/pretext.tex
added abstract and some data.

@@
 \underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
-presented on \underline{\jpsays{include the correct date here!}}.
+presented on \underline{\jpsays{include the correct date here FORMAT = April 10, 2020!}}.
 \vspace{1.0cm}
@@
 \doublespacing
-"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
-dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip
-ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore 
-eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia
-deserunt mollit anim id est laborum."
-
-\iffalse
-\singlespacing
-"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt..."
-\fi
+This work introduces the development of the novel flat-histogram Monte 
+Carlo (MC) method SAD and explores the convergence properties of a 
+variety of related weight-based MC methods. The new method is applied 
+to a variety of physical `test' systems including the 2D Ising model, 
+square-well fluid, and Lennard-Jones clusters. A driving motivation for 
+developing novel Monte Carlo methods that have physically based tunable 
+parameters is to provide simulation methods for exploring metal-organic 
+frameworks (MOFs). A theoretical framework for gas adsorption and 
+delivery in metal-organic frameworks is developed and compared with 
+experimental data. SAD is a introduced as a powerful method for 
+examining thermodynamic properties for MOFs.  A \emph{preliminary} 
+groundwork is laid for the future development of 2D SAD which would be 
+able to compute isotherms for any given MOF thereby providing a 
+powerful way for researchers to determine a MOF's suitability for gas 
+storage applications.
 
@@
 \vspace{3.0cm}
-Presented October 26, 2019  \\
-Commencement December 2019 \jpsays{include the correct date here!}
+Presented April 10, 2020  \\
+Commencement June 2020 \jpsays{include the correct date here!}
 \end{center}
@@
 \underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
-\jpsays{get the date right here? August 26, 2015}. \\
+\jpsays{get the date right here? April 10, 2015}. \\
 
@@
 \doublespacing
-"Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
-dolore magna aliqua. Ut
+\jpsays{FIXME: "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
+dolore magna aliqua. Ut}

### 2020-02-17 scripts-python-make-data/ising-multi-heat-capacity.py

@@ -151,7 +151,7 @@ for f in filename:
             maxerror = np.zeros(N_save_times)
             cverror = np.zeros(N_save_times)
 
-            for i in range(0, N_save_times):
+            for i in range(60, N_save_times):
                 # below just set average S equal between lndos and lndosref
                 if 'ising' in save_dir:
 
@@ -181,6 +181,19 @@ for f in filename:
                 errorinentropy[i] = np.sum(abs(doserror))/len(doserror) #- np.mean(doserror)
                 maxerror[i] = np.amax(doserror) - np.amin(doserror)
                 cverror[i] = np.sum(abs(my_cv_error))/len(my_cv_error)
+                plt.figure('foo')
+                plt.clf()
+                plt.plot(T, heat_capacity(T, flip_ising_E, ising_lndos), label='myself')
+                plt.plot(T, cvref, label='cvref')
+                plt.legend(loc='best')
+                plt.title(f'save {i}')
+                plt.figure('foo diff')
+                plt.clf()
+                plt.plot(T, heat_capacity(T, flip_ising_E, ising_lndos) - cvref)
+                plt.title(f'error {i}')
+                plt.pause(0.000001)
+                # DEBUG by looking into DOS and DOS ERROR!
+                
             
             # remove N from moves in yaml file because N is added back in the
             # comparison-plot script

### 2020-02-17 scripts-python-make-data/run-ising-comparisons.sh

@@ -19,3 +19,5 @@ python figs/yaml-multi-comparison.py /home/jordan/ising-cp-data/ /home/jordan/de
 
 # ising-cp-data not weniger-cp-data
 #32408
+
+python3 scripts-python-make-data/ising-multi-heat-capacity.py --file_dir=../ising-cp-data/ --reference=../deft/papers/histogram/data/ising-32-reference-lndos.dat --save_dir=ising/data/comparison/N32 --filename ising-sad-32 ising-samc-1e5-32 ising-samc-1e6-32 ising-samc-1e7-32 ising-wl-32 ising-wl-inv-t-32 --N=32 --Emin=2048 --Emax=0 --seed_avg=8

### 2020-02-17 scripts-python/comparison-plot.py

@@ -198,6 +198,13 @@ for method in methods:
                     elif "fast" in transcale:
                         plt.title(r'$N=%d$, $\delta_0 = 0.5$' % (int(N)))
                 colors.legend()
+                if data.shape[0] > 5:
+                    heat_capacity = data[5]
+                    plt.figure('heat_capacity')
+                    colors.loglog(moves, heat_capacity, method = method[1:])
+                    plt.xlabel(r'Moves')
+                    plt.ylabel(r'Cv')
+                    colors.legend()
 
     except:
         raise

### 2020-03-17 thesis/pretext.tex
completed the pretext and acknowledgements.

@@
 \underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
-presented on \underline{\jpsays{include the correct date here FORMAT = April 10, 2020!}}.
+presented on April 30, 2020
 \vspace{1.0cm}
@@
 \justify{}
-Title: \underline{\jpsays{Title of the thesis}}.
+Title: \underline{Advancing renewable gas storage using flat-histogram methods}.
 \vspace{2.0cm}
@@
 variety of related weight-based MC methods. The new method is applied
-to a variety of physical `test' systems including the 2D Ising model, 
+to a variety of physical ‘test’ systems including the 2D Ising model,
 square-well fluid, and Lennard-Jones clusters. A driving motivation for
@@
 delivery in metal-organic frameworks is developed and compared with
-experimental data. SAD is a introduced as a powerful method for 
-examining thermodynamic properties for MOFs.  A \emph{preliminary} 
-groundwork is laid for the future development of 2D SAD which would be 
-able to compute isotherms for any given MOF thereby providing a 
-powerful way for researchers to determine a MOF's suitability for gas 
-storage applications.
+experimental data. SAD is introduced as a powerful method for examining
+thermodynamic properties for MOFs. A preliminary groundwork is laid for
+the future development of a multi-dimensional SAD which would be able
+to compute all isotherms for any given MOF thereby providing a powerful
+way for researchers to determine a MOF’s suitability for gas storage
+applications.
 
@@
 $\copyright$ Copyright by Jordan K. Pommerenck \\
-\jpsays{include the correct date here!} \\
+April 30, 2020 \\
 (All Rights Reserved)
@@
 \begin{center}
-\jpsays{Title of the thesis}
+Advancing renewable gas storage using flat-histogram methods
 
@@
 \vspace{3.0cm}
-Presented April 10, 2020  \\
-Commencement June 2020 \jpsays{include the correct date here!}
+Presented April 30, 2020  \\
+Commencement June 2020
 \end{center}
@@
 \underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
-\jpsays{get the date right here? April 10, 2015}. \\
+April 30, 2020. \\
 
@@
 I understand that my dissertation will become part of the permanent collection
-of Oregon State University libraries. My signature below authorizes release of my
-dissertation to any reader upon request. \\
+of Oregon State University libraries. My signature below authorizes release of my dissertation to any reader upon request. \\
 \vspace{0.5cm}
@@
 \doublespacing
-\jpsays{FIXME: "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et
-dolore magna aliqua. Ut}
+I gratefully acknowledge my graduate advisor David Roundy for his
+constant support and guidance throughout my research. His breadth of knowledge
+and hands-off approach have tremendously helped me to develop as a scientific
+researcher.
+I express thanks to my graduate committee members: David McIntyre, Davide Lazzati, Yun-Shik Lee, and GCR Chih-hung (Alex) Chang.
+I would like to acknowledge the many outstanding graduate students that have positively impacted me (too many to name here but you know who you are).
+I would like to acknowledge some of the outstanding undergraduate students
+from our research lab that have worked on related research:
+Tanner Simpson, Cade Trotter, and Christopher May.
+Thank you all for your helpful discussions related to research and life in
+general.
+Finally, I would like to thank my parents for their unconditional support
+throughout my PhD research.

### 2020-03-18 thesis/pretext.tex
went through the pretext again.

@@
 \doublespacing
-This work introduces the development of the novel flat-histogram Monte
+This work introduces the novel flat-histogram Monte
 Carlo (MC) method SAD and explores the convergence properties of a
@@
 to a variety of physical ‘test’ systems including the 2D Ising model,
-square-well fluid, and Lennard-Jones clusters. A driving motivation for
+square-well fluid, and a Lennard-Jones cluster. A driving motivation for
 developing novel Monte Carlo methods that have physically based tunable
@@
 the future development of a multi-dimensional SAD which would be able
-to compute all isotherms for any given MOF thereby providing a powerful
-way for researchers to determine a MOF’s suitability for gas storage
+to compute all properties of interest for any given MOF thereby providing a powerful way for researchers to determine a MOF’s suitability for gas storage
 applications.
@@
 constant support and guidance throughout my research. His breadth of knowledge
-and hands-off approach have tremendously helped me to develop as a scientific
-researcher.
+and hands-off advising have tremendously helped me to develop as an independent scientific researcher.
 I express thanks to my graduate committee members: David McIntyre, Davide Lazzati, Yun-Shik Lee, and GCR Chih-hung (Alex) Chang.
-I would like to acknowledge the many outstanding graduate students that have positively impacted me (too many to name here but you know who you are).
-I would like to acknowledge some of the outstanding undergraduate students
-from our research lab that have worked on related research:
-Tanner Simpson, Cade Trotter, and Christopher May.
-Thank you all for your helpful discussions related to research and life in
-general.
-Finally, I would like to thank my parents for their unconditional support
-throughout my PhD research.
+I would like to acknowledge the many outstanding graduate students that have positively impacted me personally and professionally (too many to name here but you know who you are).
+I would like to acknowledge some of the outstanding undergraduate students from our research lab that have worked on related research: Tanner Simpson, Cade Trotter, and Christopher May.
+Thank you all for your helpful discussions related to research and life in general.
+Finally, I would like to thank my parents for their unconditional support throughout my PhD research.

### 2020-04-15 ising/ising.tex
added new gamma data for plotting and now working on paper.

--- 2020-01-14 ising/ising.tex
@@
 phase transition behavior and the ability to directly calculate the
-exact solution for finite lattices~\cite{beale1996exact} make the
+exact solution for finite lattices~\cite{beale1996exact, haggkvist2004computation} make the
 system sufficiently interesting for such theoretical comparisons. It is
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/gamma-n32.pdf}
+\includegraphics[width=\columnwidth]{gamma-n32.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/N32-entropy-error-default.pdf}
+\includegraphics[width=\columnwidth]{N32-entropy-error-default.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/gamma-n128.pdf}
+\includegraphics[width=\columnwidth]{gamma-n128.pdf}
   \caption{
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/N128-entropy-error-default.pdf}
+\includegraphics[width=\columnwidth]{N128-entropy-error-default.pdf}
   \caption{
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{N128-Cv-error-default.pdf}
+  \caption{The error in the heat capacity of the $N=128 \times 128$ after $10^{9}$ moves as calculated with each method.}
+  \label{fig:N128-cv-error}
+\end{figure}
+
 \section{Conclusion}

### 2020-04-15 scripts-python-make-data/run-ising-comparisons.sh

@@ -6,10 +6,15 @@ python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-32-s1 32 isin
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-sad-32-s1 32 ising-sad-32
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-inv-t-32-s1 32 ising-wl-inv-t-32
 
+python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-32-minGamma-s1 32 ising-wl-32-minGamma
+
 # This is the reference for 128 X 128 system
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-128-s1 128 ising-wl-128
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-sad-128-s1 128 ising-sad-128
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-inv-t-128-s1 128 ising-wl-inv-t-128
+
+python figs/parse-ising-out.py /home/jordan/sad-monte-carlo/ising-wl-128-minGamma-s1 128 ising-wl-128-minGamma
+#cp data/gamma/ising/n128/ising-wl-128-minGamma.txt ../../../thesis-pommerenck/ising/data/gamma/n128/.
 
 # This is the reference for 32 X 32 system
 #python yaml-multi-comparison.py /home/jordan/ising-cp-data/ /home/jordan/deft/papers/histogram/data/ising-32-reference-lndos.dat ising/N32 32 2048 0 true 8 ising-sad-32 ising-samc-1e5-32 ising-samc-1e6-32 ising-samc-1e7-32 ising-wl-32 ising-wl-inv-t-32

### 2020-04-15 scripts-python/colors.py

@@ -180,6 +180,7 @@ _legend_label = {
     'ising-wl-128': 'WL',
     'ising-wl-32': 'WL',
     'ising-wl-32-minGamma': 'WL-$\gamma_{t}$-min',
+    'ising-wl-128-minGamma': 'WL-$\gamma_{t}$-min',
     'ising-wl-inv-t-128': '$1/t$-WL',
     'ising-wl-inv-t-32': '$1/t$-WL',
     'ising-samc-1e3-32': 'SAMC ($t_0 =10^{3}$)',

### 2020-04-15 scripts-python/comparison-plot.py

@@ -219,8 +219,8 @@ for method in methods:
                         plt.xlim(1e5,1e12)
                         plt.ylim(1,1e4)
                     else:
-                        plt.xlim(1e8,10**(13.8))
-                        plt.ylim(1e1,1e6)
+                        plt.xlim(1e9,10**(13.8))
+                        plt.ylim(5e1,1e6)
                     plt.xlabel(r'Moves')
                     plt.ylabel(r'Cv')
                     colors.legend()

### 2020-04-15 scripts-python-make-data/run-ising-comparisons.sh
added final data just waiting on minGamma.

@@ -6,7 +6,8 @@ python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-32-s1 32 isin
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-sad-32-s1 32 ising-sad-32
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-inv-t-32-s1 32 ising-wl-inv-t-32
 
-python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-32-minGamma-s1 32 ising-wl-32-minGamma
+python figs/parse-ising-out.py /home/jordan/sad-monte-carlo/ising-wl-128-minGamma-s1 32 ising-wl-32-minGamma
+#cp data/gamma/ising/n32/ising-wl-32-minGamma.txt ../../../thesis-pommerenck/ising/data/gamma/n32/.
 
 # This is the reference for 128 X 128 system
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-128-s1 128 ising-wl-128

### 2020-04-16 ising/ising.tex
made some comments on the ising paper.

@@
 
+\jpsays{consider adding in minGamma WL here or before 1/t-WL method}
+
 Another weight-based flat histogram method is
@@
 corresponds to an interesting energy range from $-32768$ to $100$.  The number of
-important energy states for this system is $N_S = 8214$.
+important energy states for this system is $N_S = 8214$~
+\jpsays{This is wrong! This is simulation range. This section needs rewritten.  Why is displacement distance even mentioned?}.
 Fig.~\ref{fig:n128} shows the average error in the entropy as a function of time
@@
 \begin{figure}
+  \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
+  \caption{The error in the heat capacity of the $N=32 \times 32$ as calculated with each method.}
+  \label{fig:N32-cv-error}
+\end{figure}
+
+\begin{figure}
   \includegraphics[width=\columnwidth]{N128-Cv-error-default.pdf}
-  \caption{The error in the heat capacity of the $N=128 \times 128$ after $10^{9}$ moves as calculated with each method.}
+  \caption{The error in the heat capacity of the $N=128 \times 128$ as calculated with each method.}
   \label{fig:N128-cv-error}
 \end{figure}
+
+\subsection{Separate Heat Capacity Section?}
+Why does the heat capacity error saturate (i.e. the error is around 1 out 
+but the heat capacity value is much as 1600!)
+The smaller system SAD does great and minGamma does great but minGamma needs
+the user to pick the right minGamma for the production run otherwise you get
+the 128 by 128 results.
+
+We can either break this up or mix in?
+
+Why are we looking at Cv?
+
+(1) the errors in the derivative will be more pronounced. (numerical differentiation of noisy data)
+
+(2) interested in computing thermodynamic quantities
+
+(3) compare with published literature?
 

### 2020-04-17 ising/ising.tex
added some text to the paper.

@@
 
-Due to a canonical Monte Carlo simulation's inefficient sampling of large
-energies, a variety of ``flat'' (or ``broad'') histogram
+Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram sampling which improved the exploration of configuration space and allowed the system to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
+These works led to increase in the development of a variety of ``flat'' (or ``broad'') histogram
 methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-wang2001determining, wang2001efficient} were developed to explore a wider range
+wang2001determining, wang2001efficient} which could explore a wider range
 of energies.  In addition to obtaining thermodynamic information for the entire
@@
 \section{Conclusion}
+\jpsays{add something here about production run WL-minGamma}
 We have examined the convergence of SAD versus other widely used weight-based
@@
 
+\section{Acknowledgments}
+We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model.
+
+Johannes Zierenberg for production run discussions
+
+Martin Weigel for talking about EPA could cite earlier.
 \bibliography{ising} % Produces the bibliography via BibTeX.

### 2020-04-17 scripts-python-make-data/run-ising-comparisons.sh
fixed the data script so the right data is used.

@@ -6,7 +6,7 @@ python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-32-s1 32 isin
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-sad-32-s1 32 ising-sad-32
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-inv-t-32-s1 32 ising-wl-inv-t-32
 
-python figs/parse-ising-out.py /home/jordan/sad-monte-carlo/ising-wl-128-minGamma-s1 32 ising-wl-32-minGamma
+python figs/parse-ising-out.py /home/jordan/sad-monte-carlo/ising-wl-32-minGamma-s1 32 ising-wl-32-minGamma
 #cp data/gamma/ising/n32/ising-wl-32-minGamma.txt ../../../thesis-pommerenck/ising/data/gamma/n32/.
 
 # This is the reference for 128 X 128 system

### 2020-04-17 ising/ising.tex
fixed the data script so the right data is used.
made notes for self on editing the ising paper.

@@
   common case in which the range of energies is not known in advance.
+  \jpsays{production run WL is also tested and does well} 
 \end{abstract}
@@
 \section{Introduction}
+
+\jpsays{comment out the graphs we are not going to use i.e. the average entropy error.  Change Cv to cv in plots i.e. divide by 32 X 32 etc...}
+
 Flat histogram Monte-Carlo simulation algorithms calculate the thermodynamic
@@
 of detailed balance, convergence of the algorithm is not guaranteed.
+
+\jpsays{here might be a good place to mention production run WL}
 
@@
 
+\jpsays{add in the heat capacity formula here and comment out the average deviation form the exact solution for the entropy}
+
 \section{Flat histogram methods}\label{sec:histogram}
@@
 
-\jpsays{consider adding in minGamma WL here or before 1/t-WL method}
+\jpsays{consider adding in minGamma WL here or before 1/t-WL method.
+An alternative method of implementing the WL algorithm is to first implement a “production” run~\cite{}.
+The WL method updates the weights w(E) until some Gamma min is reached.
+At this point, the weights are fixed and the method proceeds.}
 
@@
 \section{Results}
+
+\jpsays{this all needs rewritten in terms of heat capacity and WL min gamma.}
 We tested the algorithms on two 2D Ising model systems.  The first is a smaller
@@
 \section{Acknowledgments}
+
+\jpsays{edit here!}
 We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model.

### 2020-04-20 ising/ising.tex
made a few paper edits.

@@
 \section{Introduction}
-
-\jpsays{comment out the graphs we are not going to use i.e. the average entropy error.  Change Cv to cv in plots i.e. divide by 32 X 32 etc...}
-
 Flat histogram Monte-Carlo simulation algorithms calculate the thermodynamic
@@
 benchmarking or comparing Monte-Carlo
-methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The 2nd order
+methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing, barash2019estimating}. The 2nd order
 phase transition behavior and the ability to directly calculate the
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{N32-entropy-error-default.pdf}
-  \caption{
-  The average entropy error for each MC method for $N=32 \times 32$,
-               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-               as a function of number of iterations run.  The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded area.}\label{fig:n32}
+  \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
+  \caption{The maximum error in the heat capacity for each method for the $N=32 \times 32$ and $T_{\min} = 1$ as a function of number of iterations run.
+The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
+  \label{fig:N32-cv-error}
 \end{figure}
+
+% \begin{figure}
+% \includegraphics[width=\columnwidth]{N32-entropy-error-default.pdf}
+%   \caption{
+%   The average entropy error for each MC method for $N=32 \times 32$,
+%                $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+%                as a function of number of iterations run.  The error is
+%                averaged over 8 independent simulations, and the best
+%                and worst simulations for each method are shown as a
+%                semi-transparent shaded area.}\label{fig:n32}
+% \end{figure}
 
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{N128-entropy-error-default.pdf}
-  \caption{
-  The average entropy error for each MC method for $N=128 \times 128$,
-               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-               as a function of number of iterations run.  The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded ar.}\label{fig:n128}
+  \includegraphics[width=\columnwidth]{N128-Cv-error-default.pdf}
+  \caption{The maximum error in the heat capacity for each method for the $N=128 \times 128$ and $T_{\min} = 1$ as a function of number of iterations run.
+The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
+  \label{fig:N128-cv-error}
 \end{figure}
+
+% \begin{figure}
+% \includegraphics[width=\columnwidth]{N128-entropy-error-default.pdf}
+%   \caption{
+%   The average entropy error for each MC method for $N=128 \times 128$,
+%                $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+%                as a function of number of iterations run.  The error is
+%                averaged over 8 independent simulations, and the best
+%                and worst simulations for each method are shown as a
+%                semi-transparent shaded ar.}\label{fig:n128}
+% \end{figure}
 
@@
 1/t-WL at which point both methods will converge equivalently.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
-  \caption{The error in the heat capacity of the $N=32 \times 32$ as calculated with each method.}
-  \label{fig:N32-cv-error}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N128-Cv-error-default.pdf}
-  \caption{The error in the heat capacity of the $N=128 \times 128$ as calculated with each method.}
-  \label{fig:N128-cv-error}
-\end{figure}
 
@@
 \section{Conclusion}
-\jpsays{add something here about production run WL-minGamma}
+\jpsays{add something here about production run WL-minGamma
+The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter Gamma min makes the method somewhat more difficult to implement in practice.}
 We have examined the convergence of SAD versus other widely used weight-based
@@
 
-\jpsays{edit here!}
 We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model.
+We also with to thank Johannes Zierenberg for helpful discussions regarding WL production simulations and
+Martin Weigel for discussions regarding entropic population annealing.
 
-Johannes Zierenberg for production run discussions
-
-Martin Weigel for talking about EPA could cite earlier.
 \bibliography{ising} % Produces the bibliography via BibTeX.

### 2020-04-20 scripts-python-make-data/run-ising-comparisons.sh

@@ -37,6 +37,8 @@ python3 scripts-python-make-data/ising-multi-heat-capacity.py --file_dir=../sad-
 
 python3 scripts-python-make-data/ising-multi-heat-capacity.py --file_dir=../sad-monte-carlo/ --reference=/home/jordan/deft/papers/histogram/data/ising-128-T15-reference-lndos.dat --save_dir=ising/data/comparison/N128 --filename ising-wl-128-minGamma --N=128 --Emin=31972 --Emax=0 --seed_avg=8
 
+python3 scripts-python-make-data/ising-multi-heat-capacity.py --file_dir=../sad-monte-carlo/ --reference=/home/jordan/deft/papers/histogram/data/ising-128-T15-reference-lndos.dat --save_dir=ising/data/comparison/N128 --filename ising-wl-128-minGamma-1e6 --N=128 --Emin=31972 --Emax=0 --seed_avg=8
+
 # Heat Capacity Plotting
 python3 scripts-python/ising-cv-save-csv.py --file_dir=../ising-cp-data/ --reference=../deft/papers/histogram/data/ising-32-reference-lndos.dat --save_dir=ising/data/heat-capacity --filename ising-sad-32 ising-samc-1e5-32 ising-samc-1e6-32 ising-samc-1e7-32 ising-wl-32 ising-wl-inv-t-32 ising-wl-32-minGamma --N=32 --Emin=2048 --Emax=0 --seed_avg=1
 

### 2020-04-20 scripts-python/comparison-plot.py

@@ -201,11 +201,18 @@ for method in methods:
                 colors.legend()
                 if data.shape[0] > 5:
                     plt.figure('heat_capacity')
-                    heat_capacity = data[5]
+                    if filebase == 'N32':
+                        plt.xlim(1e5,1e12)
+                        plt.ylim(1e-3,1e1)
+                        cv_factor = (32*32)
+                    else:
+                        plt.xlim(1e9,10**(13.8))
+                        plt.ylim(1e-3,1e1)
+                        cv_factor = 128*128
                     try:
-                        minmean_cv = data[6]
-                        maxmean_cv = data[7]
-
+                        heat_capacity = data[5] / cv_factor
+                        minmean_cv = data[6] / cv_factor
+                        maxmean_cv = data[7] / cv_factor
                         plt.fill_between(moves, minmean_cv, maxmean_cv,
                                         edgecolor='none', linewidth=0,
                                         color=colors.color(method[1:]),
@@ -215,14 +222,8 @@ for method in methods:
                     colors.loglog(moves, heat_capacity, method = method[1:])
                     for i in np.arange(-8, 19, 1.0):
                         colors.loglog(moves, 10**i/np.sqrt(0.1*moves), method = r'1/sqrt(t)')
-                    if filebase == 'N32':
-                        plt.xlim(1e5,1e12)
-                        plt.ylim(1,1e4)
-                    else:
-                        plt.xlim(1e9,10**(13.8))
-                        plt.ylim(5e1,1e6)
-                    plt.xlabel(r'Moves')
-                    plt.ylabel(r'Cv')
+                    plt.xlabel(r'$\textrm{Moves}$')
+                    plt.ylabel(r'Maximum error in $c_V$')
                     colors.legend()
                     plt.savefig('%s-Cv-error-%s.pdf' % (tex_filebase,transcale))
 

### 2020-04-21 ising/ising.tex
made some paper changes.

@@
   We examine the convergence of the Stochastic Approximation with a Dynamic
-  update factor (SAD) algorithm applied to the 2D ising model. Comparison with
+  update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
   SAMC and WL methods show that SAD performs robustly and without user input
-  knowledge of an energy range. We confirm that SAD is more powerful in the
-  common case in which the range of energies is not known in advance.
-  \jpsays{production run WL is also tested and does well} 
+  knowledge of an energy range. We also compare SAD with a final production run WL simulation with fixed weights. While SAD is more powerful in the
+  common case in which the range of energies is not known in advance, the 2D Ising model energy range is readily known and easily discovered by WL methods.
 \end{abstract}
@@
 
-Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram sampling which improved the exploration of configuration space and allowed the system to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
+Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram
+sampling which improved the exploration of configuration space and allowed the
+system to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
 These works led to increase in the development of a variety of ``flat'' (or ``broad'') histogram
@@
 
-\jpsays{here might be a good place to mention production run WL}
+The WL method can also be implemented as part of a production run which aids in
+equilibration~\cite{gross2018massively}. WL is used to generate the weights
+resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
+minimum $\gamma$ is reached, the simulation switches to a final production run
+with fixed weights. WL implemented in this way can ensure ergodicity and
+detailed-balance. The converge of the production run is still impacted by the
+choice of the minimum $\gamma$.
 
@@
 \end{align}
+% where $g(E)$ is the multiplicity of the system which is proportional to
+% $D(E)$. We can estimate the average deviation from the exact solution
+% using the relation~\cite{schneider2017convergence, shakirov2018convergence,
+% barash2017control}:
+% \begin{align}
+% \braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
+% \end{align}
 where $g(E)$ is the multiplicity of the system which is proportional to
-$D(E)$. We can estimate the average deviation from the exact solution
-using the relation~\cite{schneider2017convergence, shakirov2018convergence,
-barash2017control}:
+$D(E)$. We can estimate the maximum deviation in the canonical specific heat capacity $c_V$ from the exact solution~\cite{schneider2017convergence, shakirov2018convergence,
+barash2017control,barash2017gpu}:
 \begin{align}
-\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
+\braket{\epsilon(t)}_E = \max_E \left( c_V(E,S,T,t) - c_V(E,S,T,t)_{\text{Beale}}\right)
 \end{align}
-As per the implementation by Wang and
-Landau~\cite{wang2001determining}, the total number of available energy
-states are $N-1$.
-
-\jpsays{add in the heat capacity formula here and comment out the average deviation form the exact solution for the entropy}
+% As per the implementation by Wang and
+% Landau~\cite{wang2001determining}, the total number of available energy
+% states are $N-1$.
+Computing the specific heat capacity $c_V$ presents a difficult challenge for
+any Monte-Carlo method due to fluctuations in the derivative of the internal energy.  Methods that accurately compute $c_V$ also by extension accurately compute the internal energy.
 
@@
 We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model.
-We also with to thank Johannes Zierenberg for helpful discussions regarding WL production simulations and
+We also with to thank Johannes Zierenberg for helpful discussions regarding WL production run simulations and
 Martin Weigel for discussions regarding entropic population annealing.

### 2020-04-21 scripts-python-make-data/run-ising-comparisons.sh

@@ -17,6 +17,9 @@ python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-inv-t-128-s1 
 python figs/parse-ising-out.py /home/jordan/sad-monte-carlo/ising-wl-128-minGamma-s1 128 ising-wl-128-minGamma
 #cp data/gamma/ising/n128/ising-wl-128-minGamma.txt ../../../thesis-pommerenck/ising/data/gamma/n128/.
 
+python figs/parse-ising-out.py /home/jordan/sad-monte-carlo/ising-wl-128-minGamma-1e6-s1 128 ising-wl-128-minGamma-1e6
+#cp data/gamma/ising/n128/ising-wl-128-minGamma-1e6.txt ../../../thesis-pommerenck/ising/data/gamma/n128/.
+
 # This is the reference for 32 X 32 system
 #python yaml-multi-comparison.py /home/jordan/ising-cp-data/ /home/jordan/deft/papers/histogram/data/ising-32-reference-lndos.dat ising/N32 32 2048 0 true 8 ising-sad-32 ising-samc-1e5-32 ising-samc-1e6-32 ising-samc-1e7-32 ising-wl-32 ising-wl-inv-t-32
 # python3 scripts-python-make-data/ising-multi-heat-capacity.py --file_dir=../ising-cp-data/ --reference=../deft/papers/histogram/data/ising-32-reference-lndos.dat --save_dir=ising/data/comparison/N32 --N=32 --Emin=2048 --Emax=0 --seed_avg=8 --filename ising-sad-32 ising-samc-1e5-32 ising-samc-1e6-32 ising-samc-1e7-32 ising-wl-32 ising-wl-inv-t-32 

### 2020-04-22 ising/ising.tex
making major ising paper edits.

@@
 
-\title{Flat histogram method comparison on 2D Ising Model
+\title{Flat-histogram method comparison on 2D Ising Model
 }
@@
 \section{Introduction}
-Flat histogram Monte-Carlo simulation algorithms calculate the thermodynamic
+Flat-histogram Monte-Carlo simulation algorithms calculate the thermodynamic
 properties of various systems over a range of temperatures.  The first histogram
@@
 
-The WL method can also be implemented as part of a production run which aids in
-equilibration~\cite{gross2018massively}. WL is used to generate the weights
-resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
-minimum $\gamma$ is reached, the simulation switches to a final production run
-with fixed weights. WL implemented in this way can ensure ergodicity and
-detailed-balance. The converge of the production run is still impacted by the
-choice of the minimum $\gamma$.
-
 Because of the uncertainty of convergence for WL, many studies were undertaken
@@
 non-convergence~\cite{belardinelli2007fast}.  These findings led to the
-formation of the $1/t$-WL algorithm.
+formation of the $1/t$-WL algorithm and also sparked some debate as to whether
+the simulation should be split into multiple parts, called ``production run" WL, in order to preserve ergodicity and
+detailed balance~\cite{jayasri2005wang, mukhopadhyay2008monte}.
 
@@
 
-In this work, we apply the family of weight-based flat histogram Monte
+In this work, we apply the family of weight-based flat-histogram Monte
 Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
@@
 system~\cite{wang2001determining,wang2001efficient}.
-We test the convergence of several flat histogram methods
+We test the convergence of several flat-histogram methods
 on the periodic 2D square lattice ferromagnetic Ising model with nearest
@@
 
-\section{Flat histogram methods}\label{sec:histogram}
-Flat histogram methods determine the density of states $D(E)$ over a broad range
-of energies by simulating each energy with equivalent accuracy. Flat histogram
+\section{Flat-histogram methods}\label{sec:histogram}
+Flat-histogram methods determine the density of states $D(E)$ over a broad range
+of energies by simulating each energy with equivalent accuracy. Flat-histogram
 Monte Carlo methods propose randomly chosen ``moves'' which change the state of
@@
 
-We describe four closely related flat histogram methods which each rely on a
+We describe several closely related flat-histogram methods which each rely on a
 weight function $w(E)$ to determine $D(E)$.  For these algorithms, the
@@
 which biases the simulation in favor of energies with low weights. The result of
-weights $w(E)$ that are proportional to $D(E)$ is an entirely flat histogram. We
+weights $w(E)$ that are proportional to $D(E)$ is an entirely flat-histogram. We
 can relate the entropy to the weights in the microcanonical ensemble, since the
@@
 
-Each approach uses a random walk in energy space to estimate $D(E)$.  At the
-core of each of these approaches is to continuously update the weights at each
+Flat-histogram methods employ a random walk in energy space to estimate $D(E)$.  Each method operates by continuously updating the weights at each
 step of the simulation
@@
 factor, and $E$ is the current energy.  This update causes the random walk to
-avoid energies that have been frequently sampled, leading to a rapid exploration
-of energy space. The four methods differ primarily in how they schedule the
-decreasing of $\gamma_t$.
+avoid frequently sampling the same energies, leading to a rapid exploration
+of energy space. Flat-histogram methods differ primarily in how they schedule
+the decrease of $\gamma_t$.
 
 The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
-landau2014guide} explores energy space by initially setting $\gamma=1$, and then
-decreases $\gamma$ in discrete stages. An energy range of interest must be
-specified~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
-which can require multiple simulations if unknown.
+landau2014guide} explores energy space by setting $\gamma_{t=0}^{\text{WL}}=1$,and then decreases $\gamma^{\text{WL}}$ in prescribed stages. An energy range of
+interest must be specified~\cite{wang2001efficient, schulz2003avoiding,yan2003fast}, which often requires multiple simulations if unknown.
 
 The number (``counts'') of moves ending at each energy are stored in a
-histogram.  For a sufficiently flat energy histogram (specified by the user to
-typically be 0.8), $\gamma$ is decreased by a specified factor of $\frac12$ and
-the histogram is reset to zero. The entire process is repeated until $\gamma$
-reaches a desired cutoff.
+histogram.
+For a sufficiently flat energy histogram (typically user-specified to be 0.8), $\gamma^{\text{WL}}$ is decreased by a specified factor of $\frac12$ and the histogram is reset to zero.
+The entire process is repeateduntil $\gamma^{\text{WL}}$ reaches a desired cutoff.
 
@@
 
-\jpsays{consider adding in minGamma WL here or before 1/t-WL method.
-An alternative method of implementing the WL algorithm is to first implement a “production” run~\cite{}.
-The WL method updates the weights w(E) until some Gamma min is reached.
-At this point, the weights are fixed and the method proceeds.}
+The WL method can also be implemented as part of a production run which aids in
+equilibration~\cite{gross2018massively}. WL is used to generate the weights
+resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
+minimum $\gamma$ is reached, the simulation switches to a final production run
+with fixed weights. WL implemented in this way can ensure ergodicity and
+detailed-balance; however, the convergence of simulation is still impacted by
+the choice of the minimum $\gamma$.
 
-Another weight-based flat histogram method is
+Another weight-based flat-histogram method is
 the Stochastic Approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
@@
 is the number of energy bins~\cite{liang2007stochastic}.  Schneider
-\emph{et al.} found that for the Ising model this heuristic is helpful
-for small spin systems, but that larger systems require an even higher
-$t_0$ value~\cite{schneider2017convergence}. Our work also confirms this.
+\emph{et al.} found and we confirm that for the Ising model this heuristic is helpful for small spin systems, but that larger systems require an even higher
+$t_0$ value~\cite{schneider2017convergence}.
 

### 2020-04-22 ising/ising.tex
update the 1/t-WL method paragraphs.

@@
 convergence~\cite{zhou2005understanding,lee2006convergence,
-belardinelli2007wang}. The ideal choice of decreasing the modification factor
-helps to prevent CPU time from being wasted by continuing to perform
-calculations after the error in the density of states is
-saturated~\cite{belardinelli2008analysis}. Modification factors that decrease
-faster than $1/t$ were found to lead to
-non-convergence~\cite{belardinelli2007fast}.  These findings led to the
+belardinelli2007wang}.
+Belardinelli and Pereyra showed that an update factor that decreases faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang, belardinelli2008analysis, zhou2008optimal}, where $t$ corresponds to the number of moves.
+Schneider \emph{et al.} outline minor refinements algorithm including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.
+These studies led to the
 formation of the $1/t$-WL algorithm and also sparked some debate as to whether
-the simulation should be split into multiple parts, called ``production run" WL, in order to preserve ergodicity and
+the entire simulation should be split into multiple parts, called ``production run" WL, in order to preserve ergodicity and
 detailed balance~\cite{jayasri2005wang, mukhopadhyay2008monte}.
@@
 For a sufficiently flat energy histogram (typically user-specified to be 0.8), $\gamma^{\text{WL}}$ is decreased by a specified factor of $\frac12$ and the histogram is reset to zero.
-The entire process is repeateduntil $\gamma^{\text{WL}}$ reaches a desired cutoff.
+The entire process is repeated until $\gamma^{\text{WL}}$ reaches a desired cutoff.
 
-Zhou et. al also showed that the WL algorithm never converges
-exponentially~\cite{zhou2008optimal} and succesfully bounded the statistical
-error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}. He also suggests
-that a modification factor of $1/t$ is optimal for WL.
+The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
+from dropping below $N_S/t$~\cite{belardinelli2008analysis,
+schneider2017convergence}. The method follows the standard WL algorithm with two
+modifications.  Firstly, when each energy state has been visited once, the histogram is considered flat and $\gamma_t$ is
+decreased by a factor of two. Secondly, when
+$\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the update factor becomes
+$\gamma_t = N_S/t$ for the remainder of the simulation:
+\begin{align}
+  \gamma_t^{1/t\text{-WL}} = \begin{cases}
+     \gamma^{\text{WL}}_t & \gamma^{\text{WL}}_t > \frac{N_S}{t} \\
+     \frac{N_S}{t} & t \ge t_0
+ \end{cases}
+\end{align}
+where $t$ is the number of moves, $\gamma^{\text{WL}}_t$ is the Wang-Landau update factor
+at move $t$, and $N_S$ is the number of energy bins.
 
-The $1/t$-WL method is a refinement of the original implementation that
-addresses the error saturation of WL. Belardinelli and Pereyra
-implemented a schedule~\cite{belardinelli2007fast} that enforced that
-if $\gamma < \nicefrac{1}{t}$ then the update factor is set to
-$\nicefrac{1}{t}$ and the histogram is no longer tracked. Using this
-metric, the error saturation is avoided since the correct density of
-states is approached asymptotically as
-$t^{-\frac12}$~\cite{belardinelli2008analysis}.
-Schneider $\text{et. al}$ outline minor refinements to the $1/t$-WL algorithm
-including $\nicefrac{N_S}{t}$ scaling and switching from standard WL to
-$1/t$-WL when the update factor $\gamma <
-\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the original
-$1/t$ implementation~\cite{belardinelli2007fast}, the update factor is
-decreased once all $N_S$ energy states are visited at least once (i.e.
-$H(E) \neq 0$) effectively avoiding the concept of `flatness'.
-
-The WL method can also be implemented as part of a production run which aids in
+The WL method can alternatively be implemented as part of a production run which aids in
 equilibration~\cite{gross2018massively}. WL is used to generate the weights

### 2020-04-22 ising/ising.tex
update to the paper; now on conclusion.

@@
 a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
-its power, the method unfortunately requires a priori knowledge of several
+its power, the method unfortunately requires \emph{a priori.} knowledge of several
 user-defined parameters. Thus, for any given system under study, the user needs
@@
 \begin{align}
-\braket{\epsilon(t)}_E = \max_E \left( c_V(E,S,T,t) - c_V(E,S,T,t)_{\text{Beale}}\right)
+\braket{\epsilon(t)}_E = \left|\max_E \left(c_V(E,S,T,t) - c_V(E,S,T,t)_{\text{Beale}}\right)\right|
 \end{align}
@@
 resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
-minimum $\gamma$ is reached, the simulation switches to a final production run
+minimum $\gamma^{\text{WL}}_t$ is reached, the simulation switches to a final production run
 with fixed weights. WL implemented in this way can ensure ergodicity and
 detailed-balance; however, the convergence of simulation is still impacted by
-the choice of the minimum $\gamma$.
+the choice of the minimum $\gamma^{\text{WL}}_t$.
 
@@
 the Stochastic Approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
-schedule by which the update factor $\gamma$ is continuously
+schedule by which the update factor $\gamma^{\text{SA}}_t$ is continuously
 decreased~\cite{liang2007stochastic, werlich2015stochastic,
@@
 
-SAMC offers extreme simplicity and proven convergence.
-Provided the update factor satisfies
+The implementation of SAMC is extremely simplistic.
+In addition, Liang has proven that the weights converge to the true
+density of states~\cite{liang2006theory, liang2007stochastic,
+liang2009improving} provided the update factor satisfies
 \begin{align}
@@
 \end{align}
-where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
-to converge to the true density of states~\cite{liang2006theory,
-liang2007stochastic}.  Unlike WL methods, the energy range need not be
-known \emph{a priori.}  The convergenc time depends only on the choice of
+where $\zeta > 1$.  Unlike WL methods, the energy range need not be
+known \emph{a priori.} and the convergence time depends only on the choice of
 parameter $t_0$.
-The parameter $t_0$ can unfortunately be difficult to chose in advance
+Unfortunately, $t_0$ can be difficult to chose in advance
 for arbitrary systems.
@@
 provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
-requiring a priori an energy range of interest; however, this may easier to
+requiring \emph{a priori.} an energy range of interest; however, this is almost always easier to
 identify and is more physical than the SAMC parameter $t_0$. The update factor
@@
 \section{Results}
-
-\jpsays{this all needs rewritten in terms of heat capacity and WL min gamma.}
-We tested the algorithms on two 2D Ising model systems.  The first is a smaller
-simulation with a lattic size of 32 and the second has a lattice size of 128.
-For each system we use a reasonable root-mean-square displacement distance
-$\delta_0 = 0.05\sigma$ for proposed moves. The simulations explore the energy
-space of the systems with minimum reduced temperatures of $T_{\text{min}} = 1$
-for simulations All simulations lead to the minimum important energy $E_{\min}$
+We tested the algorithms on two different system sizes of the 2D Ising model.  The first is a smaller
+simulation with a lattice size of $N = 32 \times 32$ and the second has a lattice size of $N = 128 \times 128$. The SAD method explores the energy
+space of the each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations lead to the minimum important energy $E_{\min}$
 and maximum entropy energy $E_{\max}$ being calculated (with the exception of
-the WL methods where both of these parameters are needed a priori).
+the WL methods where both of these parameters are needed \emph{a priori.}).
 
@@
   system.}
+  \label{fig:N32-gamma}
 \end{figure}
@@
 \subsection{The 32 $\times$ 32 Ising model}
-For this simulation, we chose a minimum reduced temperature of $1$, which
-corresponds to an interesting energy range from $-2048$ to $48$.  The number of
-important energy states for this system is $N_S = 529$.
+Fig.~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
+All of the update factors initially start at 1 $\gamma_{t=0} = 1$.
+SAD dynamically updates $\gamma_t$ throughout the simulation.
+After about $10^{10}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
+Both WL and ``production'' WL update factors $\gamma_t$ are shown.
+The production WL begins after $\gamma_t$ has reached $10^{-4}$.
+The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
+All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point the decrease as $1/t$.
 
-Fig.~\ref{fig:n32} shows the average error in the entropy as a function of time for
-this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed lines
-represent the average of the absolute value of the error in the entropy averaged
-over eight simulations using different random number seeds. The range of average
-errors for each simulation is shown as a shaded region around its mean error. By
+Fig.~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
+this system. The solid/dashed lines
+represent the average of the maximum value of the error in the specific heat $c_V$ averaged
+over eight simulations using different random number seeds. The range of maximum
+errors for each simulation is shown as a shaded region. By
 the time $10^8$ moves have been made all but the WL simulation have begun to
@@
   system.}
+  \label{fig:N128-gamma}
 \end{figure}
@@
 \subsection{128 $\times$ 128 Ising system}
-For this simulation, we chose a minimum reduced temperature of $1$, which
-corresponds to an interesting energy range from $-32768$ to $100$.  The number of
-important energy states for this system is $N_S = 8214$~
-\jpsays{This is wrong! This is simulation range. This section needs rewritten.  Why is displacement distance even mentioned?}.
-Fig.~\ref{fig:n128} shows the average error in the entropy as a function of time
-for this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed
-lines represent the average of the absolute value of the error in the entropy
-averaged over eight simulations using different random number seeds. The range
-of average errors for each simulation is shown as a shaded region around its
-mean error. By the time $10^{12}$ moves have been made all but the WL and SAD
-simulations have begun to converge as $1/\sqrt{t}$. We then see the WL error
-saturate around $10^{12}$ moves. The error in the SAD method will continue to
-converge faster than $1/\sqrt{t}$ until the update factor approaches that of
-1/t-WL at which point both methods will converge equivalently.
+Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
+All of the update factors initially start at 1 $\gamma_{t=0} = 1$.
+SAD dynamically updates $\gamma_t$ throughout the simulation.
+After about $10^{13}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
+Both WL and ``production'' WL update factors $\gamma_t$ are shown.
+The production WL begins after $\gamma_t$ has reached $10^{-4}$ and $10^{-6}$.
+The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
+All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point the decrease as $1/t$.
 
-\subsection{Separate Heat Capacity Section?}
-Why does the heat capacity error saturate (i.e. the error is around 1 out 
-but the heat capacity value is much as 1600!)
-The smaller system SAD does great and minGamma does great but minGamma needs
-the user to pick the right minGamma for the production run otherwise you get
-the 128 by 128 results.
-
-We can either break this up or mix in?
-
-Why are we looking at Cv?
-
-(1) the errors in the derivative will be more pronounced. (numerical differentiation of noisy data)
-
-(2) interested in computing thermodynamic quantities
-
-(3) compare with published literature?
+Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a function of time for
+this system. The solid/dashed lines
+represent the average of the maximum value of the error in the specific heat $c_V$ averaged
+over eight simulations using different random number seeds. The range of maximum
+errors for each simulation is shown as a shaded region. By
+the time $10^13$ moves have been made all but the WL simulation have begun to
+converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^12$ moves.
 

### 2020-04-22 scripts-python/comparison-plot.py

@@ -206,7 +206,7 @@ for method in methods:
                         plt.ylim(1e-3,1e1)
                         cv_factor = (32*32)
                     else:
-                        plt.xlim(1e9,10**(13.8))
+                        plt.xlim(1e8,10**(13.8))
                         plt.ylim(1e-3,1e1)
                         cv_factor = 128*128
                     try:

### 2020-04-22 scripts-python/ising-gamma-plot.py

@@ -48,7 +48,7 @@ try:
             time = data[:,1]
             ehi = data[:,2]
             elo = data[:,3]
-            ts = np.exp(np.linspace(0, np.log(max(time)*1e4), 2000))
+            ts = np.exp(np.linspace(0, np.log(max(time)*1e6), 2000))
             gamma = np.zeros_like(ts)
             print(sad, time)
             for j in range(len(time)):
@@ -80,8 +80,12 @@ def gamma_sa(t,t0):
 
 if '128' in filename:
     t0s = ['1e5','1e6','1e7','1e8','1e9']
+    plt.ylim(1e-10, 1e1)
+    plt.xlim(1e3, 1e13)
 elif '32' in filename:
     t0s = ['1e3','1e4','1e5','1e6','1e7']
+    plt.ylim(1e-10, 1e1)
+    plt.xlim(1e3, 1e13)
 
 
 
@@ -91,8 +95,7 @@ for t0 in t0s:
     plt.ylabel(r'$\gamma_{t}$')
     colors.legend()
 plt.tight_layout()
-plt.ylim(1e-10, 1e1)
-plt.xlim(1e3, 1e13)
+
 
 plt.savefig('gamma-%s.pdf' % filename.replace('.','_'))
 if 'noshow' not in sys.argv:

### 2020-04-23 ising/ising.tex
finished the paper. Now proofing a printed copy.

@@
   SAMC and WL methods show that SAD performs robustly and without user input
-  knowledge of an energy range. We also compare SAD with a final production run WL simulation with fixed weights. While SAD is more powerful in the
-  common case in which the range of energies is not known in advance, the 2D Ising model energy range is readily known and easily discovered by WL methods.
+  knowledge of an energy range.
+We also compare SAD with a final production run WL simulation with fixed weights which is the first of its kind recorded.
+Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
 \end{abstract}
@@
 Fig.~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
-All of the update factors initially start at 1 $\gamma_{t=0} = 1$.
+All of the update factors initially start at $\gamma_{t=0} = 1$.
 SAD dynamically updates $\gamma_t$ throughout the simulation.
 After about $10^{10}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
-Both WL and ``production'' WL update factors $\gamma_t$ are shown.
-The production WL begins after $\gamma_t$ has reached $10^{-4}$.
+Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown.
+The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$.
 The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
@@
 Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
-All of the update factors initially start at 1 $\gamma_{t=0} = 1$.
+All of the update factors initially start at $\gamma_{t=0} = 1$.
 SAD dynamically updates $\gamma_t$ throughout the simulation.
 After about $10^{13}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
-Both WL and ``production'' WL update factors $\gamma_t$ are shown.
-The production WL begins after $\gamma_t$ has reached $10^{-4}$ and $10^{-6}$.
+Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown.
+The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$.
 The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
@@
 \section{Conclusion}
-\jpsays{add something here about production run WL-minGamma
-The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter Gamma min makes the method somewhat more difficult to implement in practice.}
-We have examined the convergence of SAD versus other widely used weight-based
-histogram methods.  We have shown that Sad samples the energy space
-corresponding to a desired range of temperatures for a variety of system sizes.
-We find that both SAD and 1/t-WL demonstrate excellent and robust convergence.
-They both converge more rapidly than SAMC, and unlike WL do not suffer from
+The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter Gamma min makes the method somewhat more difficult to implement in practice.
+
+We find that both SAD, 1/t-WL, and production WL (provided an ideal minimum update factor before beginning the production run) demonstrate excellent and robust convergence.
+They all converge more rapidly than SAMC, and unlike WL do not suffer from
 error saturation. We find that for larger systems SAD reduces the updated factor
-more slowly (and conservatively) than 1/t-WL. This means that SAD will take
+more slowly (and conservatively) than 1/t-WL and production WL. This means that SAD will take
 proportionately more moves to converge to the same value as 1/t-WL as system
 size is increased. SAD requires the user to specify a temperature range of
-interest rather than an energy range of interest as 1/t-WL does. For use cases
-in which a range of desired temperatures is known, this will make the SAD method
-considerably more convenient.
+interest rather than an energy range of interest as the WL methods do.
+The 2D Ising model is a system where the energy range is known \emph{a priori.}, which makes a comparison between WL methods and SAD (which requires a minimum temperature) more biased towards WL methods.
+For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
+considerably more convenient. This work also visually demonstrates that production WL performs extremely well and is preferable to WL for ensuring both ergodicity and detailed-balance.
 

### 2020-04-24 ising/ising.tex
went over paper with a fine comb and edited.

@@
 \begin{abstract}
-  We examine the convergence of the Stochastic Approximation with a Dynamic
+We examine the convergence of the stochastic approximation with a dynamic
 update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
-  SAMC and WL methods show that SAD performs robustly and without user input
-  knowledge of an energy range.
-We also compare SAD with a final production run WL simulation with fixed weights which is the first of its kind recorded.
+stochastic approximation Monte Carlo (SAMC) and Wang-Landau (WL) methods show that SAD performs robustly and without user input knowledge of an energy range.
+We also compare SAD with ``production'' run WL with fixed weights which is the first of its kind recorded.
 Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
@@
 \section{Introduction}
-Flat-histogram Monte-Carlo simulation algorithms calculate the thermodynamic
+Flat-histogram Monte Carlo simulation algorithms calculate the thermodynamic
 properties of various systems over a range of temperatures.  The first histogram
@@
 nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
-samples a narrow energy range given limited by a narrow temperature range, it
-proves computationally inefficient at sampling large energy ranges.
+samples a narrow energy range, it proves computationally inefficient at sampling large energy ranges.
 
 Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram
-sampling which improved the exploration of configuration space and allowed the
-system to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
+sampling which improved the exploration of configurational space and allowed the
+simulation to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
 These works led to increase in the development of a variety of ``flat'' (or ``broad'') histogram
@@
 Wang and Landau introduced one of the most widely used flat-histogram
-Monte-Carlo algorithms that accurately determined the density of states (DOS) for
+Monte Carlo algorithms that determined the density of states (DOS) for
 a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
@@
 
-Because of the uncertainty of convergence for WL, many studies were undertaken
+Because of the uncertainty of convergence for WL, many studies have been undertaken
 to understand how the modification (or update) factor $\gamma$ impacts the
@@
 Liang independently considered whether WL could be treated as a special case of
-Stochastic Approximation whose convergence could be mathematically
+stochastic approximation whose convergence could be mathematically
 proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
 al.~\cite{liang2007stochastic} argued that WL can be considered a form of
-Stochastic Approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
+stochastic approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
 convergence (if certain conditions are met). Despite the added benefit of
 guaranteed convergence, the method still has a system specific user-defined
-variable. Such variables often create difficulty when applying Monte-Carlo
+variable. Such variables often create difficulty when applying Monte Carlo
 methods across arbitrary systems.
@@
 
-The SAD (Stochastic Approximation with
-Dynamic $\gamma$) method outlined by Pommerenck
+The SAD (stochastic approximation with a
+dynamic $\gamma$) method outlined by Pommerenck
 et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
@@
 The 2D Ising spin-lattice system is widely used as a testbed when
-benchmarking or comparing Monte-Carlo
+benchmarking or comparing Monte Carlo
 methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing, barash2019estimating}. The 2nd order
@@
 Computing the specific heat capacity $c_V$ presents a difficult challenge for
-any Monte-Carlo method due to fluctuations in the derivative of the internal energy.  Methods that accurately compute $c_V$ also by extension accurately compute the internal energy.
+any Monte Carlo method due to fluctuations in the derivative of the internal energy.  Methods that accurately compute $c_V$ also by extension accurately compute the internal energy.
 
@@
 factor, and $E$ is the current energy.  This update causes the random walk to
-avoid frequently sampling the same energies, leading to a rapid exploration
+avoid frequent sampling of the same energies, leading to a rapid exploration
 of energy space. Flat-histogram methods differ primarily in how they schedule
@@
 with fixed weights. WL implemented in this way can ensure ergodicity and
-detailed-balance; however, the convergence of simulation is still impacted by
+detailed balance; however, the convergence of simulation is still impacted by
 the choice of the minimum $\gamma^{\text{WL}}_t$.
@@
 Another weight-based flat-histogram method is
-the Stochastic Approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
+the stochastic approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
 schedule by which the update factor $\gamma^{\text{SA}}_t$ is continuously
@@
 where the update factor is determined dynamically rather than by the user.
-Stochastic Approximation with a dynamic $\gamma$ (SAD) requires the user to
+stochastic approximation with a dynamic $\gamma$ (SAD) requires the user to
 provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
@@
 simulation with a lattice size of $N = 32 \times 32$ and the second has a lattice size of $N = 128 \times 128$. The SAD method explores the energy
-space of the each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations lead to the minimum important energy $E_{\min}$
+space of each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations lead to the minimum important energy $E_{\min}$
 and maximum entropy energy $E_{\max}$ being calculated (with the exception of
@@
 The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
-All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point the decrease as $1/t$.
+All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.
 
@@
 errors for each simulation is shown as a shaded region. By
-the time $10^13$ moves have been made all but the WL simulation have begun to
-converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^12$ moves.
+the time $10^{13}$ moves have been made all but the WL simulation have begun to
+converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$ moves.
 
@@
 For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
-considerably more convenient. This work also visually demonstrates that production WL performs extremely well and is preferable to WL for ensuring both ergodicity and detailed-balance.
+considerably more convenient. This work also visually demonstrates that ``production'' WL performs extremely well and is preferable to WL for ensuring both ergodicity and detailed balance.
 
@@
 We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model.
-We also with to thank Johannes Zierenberg for helpful discussions regarding WL production run simulations and
+We also wish to thank Johannes Zierenberg for helpful discussions regarding WL production run simulations and
 Martin Weigel for discussions regarding entropic population annealing.

### 2020-04-24 scripts-python-make-data/ising-multi-heat-capacity.py
add debug print

@@ -170,6 +170,8 @@ for f in filename:
                     # invoke np.flip since ising_E is indexed backward!
                     # this is critical for my_cv_error or you get wrong answer.
                     flip_ising_E = np.flip(np.copy(ising_E))
+                    print('max discrepancy in energy:',
+                          abs(eref[0:Emin-Emax+1] - flip_ising_E).max())
 
                     norm_factor = np.mean(ising_norm) - np.mean(lndosref[0:Emin-Emax+1])
                     doserror = ising_lndos - lndosref[0:Emin-Emax+1] - norm_factor

### 2020-04-24 scripts-python-make-data/ising-multi-heat-capacity.py
fix error in computing ising heat capacities

@@ -156,27 +156,22 @@ for f in filename:
                 # below just set average S equal between lndos and lndosref
                 if 'ising' in save_dir:
 
-                    ising_norm = lndos[i][maxyaml:minyaml+1] # remove impossible state
-
-                    ising_lndos = lndos[i][maxyaml:minyaml+1][::-1] # remove impossible state
+                    ising_lndos = np.flip(np.copy(lndos[i][maxyaml:minyaml+1])) # remove impossible state
 
                     # the states are counted backward hence the second to last state would be at index = 1
-                    ising_norm = np.delete(ising_norm, [1])
                     ising_lndos = np.delete(ising_lndos, [len(ising_lndos)-2])
 
-                    ising_E = np.array(energies[maxyaml:minyaml+1])
-                    ising_E = np.delete(ising_E, [len(ising_E)-2])
-                    
-                    # invoke np.flip since ising_E is indexed backward!
+                    # invoke np.flip since ising_E and ising_lndos are indexed backward!
                     # this is critical for my_cv_error or you get wrong answer.
-                    flip_ising_E = np.flip(np.copy(ising_E))
-                    print('max discrepancy in energy:',
-                          abs(eref[0:Emin-Emax+1] - flip_ising_E).max())
+                    ising_E = np.flip(np.copy(energies[maxyaml:minyaml+1]))
+                    ising_E = np.delete(ising_E, [len(ising_E)-2])
 
-                    norm_factor = np.mean(ising_norm) - np.mean(lndosref[0:Emin-Emax+1])
+                    norm_factor = np.mean(ising_lndos) - np.mean(lndosref[0:Emin-Emax+1])
                     doserror = ising_lndos - lndosref[0:Emin-Emax+1] - norm_factor
+                    # print(np.array([ising_E, eref[0:Emin-Emax+1],
+                    #                 ising_lndos-norm_factor, lndosref[0:Emin-Emax+1]]).T)
 
-                    my_cv_error = heat_capacity(T, flip_ising_E, ising_lndos) - cvref
+                    my_cv_error = heat_capacity(T, ising_E, ising_lndos) - cvref
 
                     # if i == arb:
                     

### 2020-04-27 ising/ising.tex
updating the data for the paper.

@@
 this system. The solid/dashed lines
-represent the average of the maximum value of the error in the specific heat $c_V$ averaged
+represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
 over eight simulations using different random number seeds. The range of maximum

### 2020-04-28 scripts-python/colors.py
updated necessary data and made few corrections in paper.

@@ -26,8 +26,8 @@ _colors = { 'SAD': 'tab:orange',
             'converged result': 'xkcd:black',
             r'$\frac{1}{\sqrt{t}}$': '#eeeeee',
             'Exact': 'xkcd:black',
-            'WL $(\gamma_{t} = 10^{-4})$': 'r',
-            'WL $(\gamma_{t} = 10^{-6})$': 'midnightblue',
+            'WL $(\gamma_{t}^{\min} = 10^{-4})$': 'r',
+            'WL $(\gamma_{t}^{\min} = 10^{-6})$': 'midnightblue',
             
 }
 
@@ -93,8 +93,8 @@ _legend_order = [
     'ising-sad-32',
     'ising-wl-128',
     'ising-wl-32',
-    'WL $(\gamma_{t} = 10^{-4})$',
-    'WL $(\gamma_{t} = 10^{-6})$',
+    'WL $(\gamma_{t}^{\min} = 10^{-4})$',
+    'WL $(\gamma_{t}^{\min} = 10^{-6})$',
     'ising-wl-inv-t-128',
     'ising-wl-inv-t-32',
     'ising-samc-1e3-32',
@@ -181,9 +181,9 @@ _legend_label = {
     'ising-sad-32': 'SAD',
     'ising-wl-128': 'WL',
     'ising-wl-32': 'WL',
-    'ising-wl-32-minGamma': 'WL $(\gamma_{t} = 10^{-4})$',
-    'ising-wl-128-minGamma':  'WL $(\gamma_{t} = 10^{-4})$',
-    'ising-wl-128-minGamma-1e6':  'WL $(\gamma_{t} = 10^{-6})$',
+    'ising-wl-32-minGamma': 'WL $(\gamma_{t}^{\min} = 10^{-4})$',
+    'ising-wl-128-minGamma':  'WL $(\gamma_{t}^{\min} = 10^{-4})$',
+    'ising-wl-128-minGamma-1e6':  'WL $(\gamma_{t}^{\min} = 10^{-6})$',
     'ising-wl-inv-t-128': '$1/t$-WL',
     'ising-wl-inv-t-32': '$1/t$-WL',
     'ising-samc-1e3-32': 'SAMC ($t_0 =10^{3}$)',

### 2020-04-28 scripts-python/comparison-plot.py

@@ -224,7 +224,10 @@ for method in methods:
                         colors.loglog(moves, 10**i/np.sqrt(0.1*moves), method = r'1/sqrt(t)')
                     plt.xlabel(r'$\textrm{Moves}$')
                     plt.ylabel(r'Maximum error in $c_V$')
-                    colors.legend()
+                    if filebase == 'N128':
+                      colors.legend(loc='lower left')
+                    else:
+                      colors.legend()
                     plt.savefig('%s-Cv-error-%s.pdf' % (tex_filebase,transcale))
 
     except:
@@ -255,6 +258,7 @@ elif filebase == 'N32':
     plt.ylim(1e-3,1e3)
 elif filebase == 'N128':
     plt.ylim(1e-3,1e4)
+
 
 colors.legend()
 plt.tight_layout()

### 2020-04-29 ising/ising.tex
updated necessary data and made few corrections in paper.
added figure to paper and improved figure visually.

@@
 detailed balance; however, the convergence of simulation is still impacted by
-the choice of the minimum $\gamma^{\text{WL}}_t$.
+the choice of the minimum $\gamma^{\min}_t$.
 
@@
 \begin{figure}
+  \includegraphics[width=\columnwidth]{N32-Cv.pdf}
+    \caption{
+    The specific heat capacity versus the reciprocal temperature $\beta$ for the $N=32 \times 32$ system for each histogram method.}
+    \label{fig:N32-cv}
+  \end{figure}
+
+\begin{figure}
   \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
-  \caption{The maximum error in the heat capacity for each method for the $N=32 \times 32$ and $T_{\min} = 1$ as a function of number of iterations run.
+  \caption{The maximum error in the specific heat capacity for each method for the $N=32 \times 32$ and $T_{\min} = 1$ as a function of number of iterations run.
 The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
@@
   \includegraphics[width=\columnwidth]{N128-Cv-error-default.pdf}
-  \caption{The maximum error in the heat capacity for each method for the $N=128 \times 128$ and $T_{\min} = 1$ as a function of number of iterations run.
+  \caption{The maximum error in the specific heat capacity for each method for the $N=128 \times 128$ and $T_{\min} = 1$ as a function of number of iterations run.
 The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
@@
 \section{Conclusion}
-The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter Gamma min makes the method somewhat more difficult to implement in practice.
+The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter $\gamma_{t}^{\min}$ makes the method somewhat more difficult to implement in practice.
 

### 2020-04-29 scripts-python-make-data/run-ising-comparisons.sh

@@ -46,3 +46,5 @@ python3 scripts-python-make-data/ising-multi-heat-capacity.py --file_dir=../sad-
 python3 scripts-python/ising-cv-save-csv.py --file_dir=../ising-cp-data/ --reference=../deft/papers/histogram/data/ising-32-reference-lndos.dat --save_dir=ising/data/heat-capacity --filename ising-sad-32 ising-samc-1e5-32 ising-samc-1e6-32 ising-samc-1e7-32 ising-wl-32 ising-wl-inv-t-32 ising-wl-32-minGamma --N=32 --Emin=2048 --Emax=0 --seed_avg=1
 
 python3 scripts-python/ising-cv-plot.py --file_dir=ising/data/heat-capacity --N=32
+
+python3 scripts-python/ising-cv-save-csv.py --file_dir=../ising-cp-data/ --reference=../deft/papers/histogram/data/ising-32-reference-lndos.dat --save_dir=ising/data/heat-capacity --filename ising-sad-32 ising-wl-32 ising-wl-inv-t-32 ising-wl-32-minGamma --N=32 --Emin=2048 --Emax=0 --seed_avg=1

### 2020-04-29 scripts-python/colors.py

@@ -50,6 +50,7 @@ _linestyles = {
 
     r'$\frac{1}{\sqrt{t}}$': '-',
     'converged result': ':',
+    'Exact': '--',
 }
 
 _legend_order = [
@@ -272,6 +273,9 @@ def style_args(method):
     if method == '1/sqrt(t)':
         args['zorder'] = -59
         args['linewidth'] = 1.1
+    if 'Exact' in fix_legend(method):
+        args['zorder'] = 60
+        args['linewidth'] = 0.75
     return args
 
 def plot(x, y, method=None, axes=None):

### 2020-04-29 scripts-python/ising-cv-save-csv.py

@@ -150,27 +150,21 @@ for f in filename:
                 minyaml = lndos.shape[1]-1
 
             
-            times = [100000000, 1333521432, 1778279410,2371373716,3162277660, 4216965034, 5623413252, 7498942093]
-            index = np.argwhere(my_time == times[1])[0][0]
+            times = [1000000000, 1333521432, 1778279410,2371373716,3162277660, 4216965034, 5623413252, 7498942093]
+            index = np.argwhere(my_time == times[0])[0][0]
             # below just set average S equal between lndos and lndosref
             if 'ising' in save_dir:
-
-                ising_norm = lndos[index][maxyaml:minyaml+1] # remove impossible state
-
-                ising_lndos = lndos[index][maxyaml:minyaml+1][::-1] # remove impossible state
-
+                
+                ising_lndos = np.flip(np.copy(lndos[index][maxyaml:minyaml+1])) # remove impossible state
                 # the states are counted backward hence the second to last state would be at index = 1
-                ising_norm = np.delete(ising_norm, [1])
                 ising_lndos = np.delete(ising_lndos, [len(ising_lndos)-2])
-
-                ising_E = np.array(energies[maxyaml:minyaml+1])
-                ising_E = np.delete(ising_E, [len(ising_E)-2])
-                    
-                # invoke np.flip since ising_E is indexed backward!
+                
+                # invoke np.flip since ising_E and ising_lndos are indexed backward!
                 # this is critical for my_cv_error or you get wrong answer.
-                flip_ising_E = np.flip(np.copy(ising_E))
+                ising_E = np.flip(np.copy(energies[maxyaml:minyaml+1]))
+                ising_E = np.delete(ising_E, [len(ising_E)-2])
 
-                my_cv = heat_capacity(T, flip_ising_E, ising_lndos)
+                my_cv = heat_capacity(T, ising_E, ising_lndos)
                         
                 # add column
                 df['%s' % f] = my_cv

### 2020-04-30 thesis/chapter-5.tex
added in final chapters including author contributions.

\chapter{General Conclusion}

Hello this is the conclusion.

### 2020-04-30 thesis/thesis.tex
added thesis files but not building yet.
working on getting thesis to compile. Still some encoding character error.
needed to turn a unicode symbol into unicode.
added in final chapters including author contributions.

@@
     pdfproducer={Producer}, % producer of the document
-    pdfkeywords={keyword1, key2, key3}, % list of keywords
+    pdfkeywords={flat-histogram, Monte-Carlo, Thesis}, % list of keywords
     pdfnewwindow=true,      % links in new PDF window
@@
 \usepackage[left=1.5in,right=1in,top=1in,bottom=1in,includefoot,heightrounded]{geometry}
-\usepackage[pagestyles]{titlesec}
+%\usepackage[pagestyles]{titlesec}
 
-\sethead{\itshape\chaptername\thechapter}{}{\itshape\chaptertitle}
-\titleformat{\chapter}[display]{\normalfont\bfseries\Large\centering}{}{18pt}{\Large}
-\titlespacing{\chapter}{0pt}{0pt}{1cm}
+%\sethead{\itshape\chaptername\thechapter}{}{\itshape\chaptertitle}
+%\titleformat{\chapter}[display]{\normalfont\bfseries\Large\centering}{}{18pt}{\Large}
+%\titlespacing{\chapter}{0pt}{0pt}{1cm}
 
-\usepackage{blindtext}
+%\usepackage{blindtext}
 \usepackage[T1]{fontenc}
@@
 
-\addbibresource{thesis.bib}
+%\addbibresource{thesis.bib}
 
@@
 
-% \usepackage{color}
-% \newcommand{\red}[1]{{\bf \color{red} #1}}
-% \newcommand{\green}[1]{{\bf \color{green} #1}}
+\usepackage{color}
+\newcommand{\red}[1]{{\bf \color{red} #1}}
+\newcommand{\green}[1]{{\bf \color{green} #1}}
 
-% \newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
-% \newcommand{\jpsays}[1]{{\color{red} [\red{Jordan:} \emph{#1}]}}
+\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
+\newcommand{\jpsays}[1]{{\color{red} [\red{Jordan:} \emph{#1}]}}
 
@@
 \renewcommand{\cftaftertoctitle}{\hfill}
+
+% THE BIBLIOGRAPHY NEEDS TO COME BEFORE ALL OF THE CHAPTERS
+\bibliography{thesis}
 
@@
 
+% introduction
 \input{chapter-1}
-%\input{Chapters/Chapter2_Files/sad.tex}
-%\input{Chapters/Chapter3_Files/ising.tex}
-
-%\chapter{Conclusion}
-%\input{Chapters/chapter06}
+% sq-well paper
+\input{chapter-2}
+% ising paper
+\input{chapter-3}
+% gas-adsorption paper
+\input{chapter-4}
+% general conclusion linking all manuscripts
+\input{chapter-5}
 

### 2020-04-30 thesis/pretext.tex
added in final chapters including author contributions.
wrote author contribution section in thesis.

--- 2020-03-17 thesis/pretext.tex
@@
 \doublespacing
-This work introduces the development of the novel flat-histogram Monte
-Carlo (MC) method SAD and explores the convergence properties of a
+This work introduces the novel flat-histogram Monte
+Carlo (MC) method stochastic approximation with a dynamic update factor (SAD) and explores the convergence properties of a
 variety of related weight-based MC methods. The new method is applied
-to a variety of physical ‘test’ systems including the 2D Ising model,
-square-well fluid, and Lennard-Jones clusters. A driving motivation for
+to a number of physical `test’ systems including the 2D Ising model, a
+square-well fluid, and a Lennard-Jones cluster. A driving motivation for
 developing novel Monte Carlo methods that have physically based tunable
@@
 the future development of a multi-dimensional SAD which would be able
-to compute all isotherms for any given MOF thereby providing a powerful
-way for researchers to determine a MOF’s suitability for gas storage
-applications.
+to compute all properties of interest for any given MOF thereby providing a powerful way for researchers to determine any porous media's suitability for gas storage applications.
 
@@
 constant support and guidance throughout my research. His breadth of knowledge
-and hands-off approach have tremendously helped me to develop as a scientific
-researcher.
-I express thanks to my graduate committee members: David McIntyre, Davide Lazzati, Yun-Shik Lee, and GCR Chih-hung (Alex) Chang.
-I would like to acknowledge the many outstanding graduate students that have positively impacted me (too many to name here but you know who you are).
-I would like to acknowledge some of the outstanding undergraduate students
-from our research lab that have worked on related research:
-Tanner Simpson, Cade Trotter, and Christopher May.
-Thank you all for your helpful discussions related to research and life in
-general.
-Finally, I would like to thank my parents for their unconditional support
-throughout my PhD research.
+and hands-off approach to advising have tremendously helped my development as an independent scientific researcher.
+I express thanks to my entire graduate committee: David McIntyre, Davide Lazzati, Yun-Shik Lee, and graduate council representative Chih-hung (Alex) Chang.
+I would like to acknowledge the many outstanding undergraduate and graduate students that have positively impacted me personally and professionally while at Oregon State University.
+Finally, I would like to thank my parents for their unconditional support throughout my Ph.D.
+research.
+
+\newpage{}
+\thispagestyle{empty}
+\begin{center}
+	\large
+	CONTRIBUTION OF AUTHORS
+\end{center}
+\justify{}
+\doublespacing
+Michael A.
+Perlin worked on much of the early C{}\verb!++!
+codebase.
+This codebase was used primarily in the early study of many of the MC methods before switching to RUST.
+In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
+Tanner T.
+Simpson performed many of the early simulations for the flat-histogram methods using the C{}\verb!++!
+codebase.
+He also developed using Python some of the plotting scripts that allowed the saved C{}\verb!++!
+data to be visualized.
+In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
+Cory M.
+Simon collaborated together with us on the third manuscript \emph{An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}.
+He contributed significantly to the introduction section of the paper and helped in the editing phase throughout preparation.
+In addition, he designed the first figure in the paper and the first figure in the supplemental to the paper.

### 2020-04-30 thesis/thesis.tex
wrote author contribution section in thesis.

@@
     pdfproducer={Producer}, % producer of the document
-    pdfkeywords={flat-histogram, Monte-Carlo, Thesis}, % list of keywords
+    pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
     pdfnewwindow=true,      % links in new PDF window

### 2020-04-30 thesis/chapter-5.tex
added the conclusion to thesis.

@@
 
-Hello this is the conclusion.
+In the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}, we developed and refined the novel Monte Carlo method SAD.
+We compared the convergence properties of SAD to a variety of other weight-based flat-histogram methods.
+We also examined the convergence behavior applied to the square-well fluid and a Lennard-Jones cluster.
+
+In the second manuscript \emph{Flat-histogram method comparison on 2D Ising model}, we tested SAD on the 2D Ising model against a number of weight-based methods including a ``production'' run WL.
+We found that even for systems where it is convenient to choose an energy range, SAD performs incredibly well.
+
+In the third manuscript \emph{An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}, we develop an upper bound on gas delivery and compare the theory with experimental data.
+We found that while hydrogen may not be viable at room temperature, methane storage can potentially meet the Department of Energy (DOE) targets.
+The theory also suggests an ideal energy of attraction which will help when determining what MOFs to simulate.
+
+Each of these manuscripts laid the groundwork for the development of 2D SAD.
+SAD has been tested on a variety of physical systems and shows tremendous promise for simulating isotherms.
+In addition, an upper bound was developed that will aid in simulating MOFs using 2D Sad.
+
+There are two important steps that remain to complete the extension.
+The first is the development of SAD with a minimum particle number $N_{\min}$ instead of a $T_{\min}$.
+By varying the number of particles, the simulation can yield thermodynamic insights for all densities.
+The second step that remains is the merger of each SAD method such that a multidimensional 2D SAD is formed.
+Such a Monte Carlo method would be an invaluable tool for driving research in the gas storage field.

### 2020-05-01 gas-adsorption/main.tex
added gas storage to thesis pommerenck.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{amsmath}
\usepackage{subfig}

\usepackage{nicefrac} % Include for inline fractions
\usepackage{url}

% \usepackage{cite}
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage{color}  % Coloured text etc.
%
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

\newcommand{\addcite}{{\bf \color{red} []}}
\newcommand{\xvec}{\mathbf{x}}
\newcommand{\thetavec}{\boldsymbol \theta}

\newcommand\V{\Phi}
\newcommand\Vk{\tilde\Phi(\vec k)}
\newcommand\volume{V}

\newcommand{\pink}[1]{{\bf \color{pink} #1}}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\orange}[1]{{\bf \color{orange} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

%\newcommandx{\davidsays}[1]{\begin{mdframed}[backgroundcolor=blue!24]\textbf{David says:} #1\end{mdframed} }
%\newcommandx{\davidsays}[2][1=inline]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{David says:} #2} }
\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }

%% \renewcommand\davidsays[1]{}
%% \renewcommand{\corysays}[2][1=inline]{}
%TC:macro \corysays [ignore]
%TC:macro \davidsays [ignore]
%TC:macro \jordan [ignore]

\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }

\newtheorem{limitation}{Limitation}

\begin{document}
\title{
An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}
% other ideas:
%Gas delivery via pressure-swing adsorption in nanoporous materials: setting storage targets}
% A thermodynamic upper bound to gas delivery via pressure-swing adsorption in nanoporous materials:}

\author{Jordan K. Pommerenck}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}
\author{Cory M. Simon}
\affiliation{School of Chemical, Biological, and Environmental Engineering,
  Corvallis, OR 97331}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles []. One strategy to densify these gases is to pack the fuel tank with a porous material [].
  Metal-organic frameworks are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases []. The US Department of Energy (DOE) has set volumetric deliverable capacity targets [] which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels. Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction []; none have established a rigorous upper bound on the deliverable capacity.
  Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing. % "pressure-swing storage doesn't make sense to me. storage and delivery  maybe but awk.
  % I added rigid since that could be important...
  To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
  \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
  Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
  We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
  % However, this upper bound directly leads to important realizations which should guide future development.
  However, limitations to the scope of applicability of our upper bound guide future material development efforts.
 % Firstly, one could extract the gas at a higher temperature than that used while filling [].
  Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
  Secondly, the fundamental physics of our upper bound do not pertain to
  % rule out
  any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
  % firstly, secondly vs first, second?
\end{abstract}

\newcommand\pfull{\ensuremath{p_{\text{full}}}}
\newcommand\pempty{\ensuremath{p_{\text{empty}}}}

\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}

\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}

\maketitle

\newcommand\gst{\ensuremath{\Delta g_{st}}}

\paragraph*{to do}
\begin{enumerate}
    \item See david/corysays
    \item make the code presentable
    \item do the FIXME in code
\end{enumerate}

\section{Introduction}
The transportation sector is dominantly powered by petroleum-based fuels \cite{davis2009transportation}. In the US, it is responsible for 36\% of energy-related carbon dioxide emissions \cite{useia} and 50,000 premature deaths per year associated with particulate matter and ozone emissions \cite{caiazzo2013air}. 
Natural gas and hydrogen (H$_2$) are alternative transportation fuels that emit fewer/less pollutants 
than petroleum-based fuels, and thus their widespread adoption could mitigate climate change 
\cite{mcglade2015geographical} and improve air quality for human health. Further, as the finite global 
petroleum resources are declining rapidly \cite{sorrell2010global}, the development of technologies for the widespread
adoption of sustainable transportation fuels, such as hydrogen, is critical.

Natural gas, mostly methane, is considered a transition (to a renewable and clean) fuel because 
it emits 25\% less carbon dioxide \cite{eia2013much} and less toxic 
byproducts \cite{wang2000full} upon combustion per unit energy produced compared to gasoline. From an 
economic standpoint, the supply of natural gas in the US is increasing as a result of hydraulic 
fracturing and horizontal drilling techniques \cite{usnatgassupply}. A positive environmental outlook 
for natural gas, however, is predicated on mitigating fugitive emissions (methane is itself a potent 
greenhouse gas) \cite{alvarez2012greater} and groundwater contamination \cite{osborn2011methane}
from hydraulic fracturing.

Hydrogen (H$_2$) is the ultimate transportation fuel because it emits only water when it electrochemically reacts with oxygen in a fuel cell to produce electricity and power a vehicle.
% Moreover, elemental hydrogen is very abundant, though bonded with oxygen (in water) or carbon (in hydrocarbons); H$_2$ does not occur naturally in abundant quantities.
Currently, hydrogen (H$_2$) is primarily produced by steam reforming of natural gas followed by the
water-gas shift reaction, which emits carbon dioxide \cite{crabtree2004hydrogen}. Notably, the 
environmental allure of hydrogen is predicated on its production via a renewable means, e.g.\ splitting
water using wind-generated electricity.

% \jordan{I got the values from wikipedia but the citation that wikipedia gave didn't link to the correct paper? Article is Energy Density.
% THIS PAPER says H2 at 700 bar is 4.7 MJ/L: Review on processing of metal–organic framework (MOF) materials towards system integration for hydrogen storage
% }
At ambient conditions, both methane
%(0.036 MJ/L)
and hydrogen gas
%(0.01 MJ/L)
possess a very low volumetric energy density compared to (liquid) gasoline.
%(34.2 MJ/L)
Consequently, under storage space constraints in passenger vehicles, natural gas and hydrogen must be densified for onboard storage to achieve a reasonable driving range on a ``full'' tank of fuel. Traditional densification approaches are liquefaction (cryogenic storage near atmospheric pressure)
% (at 1 atm, liquid CH$_4$ at 111~K to achieve 22.2 MJ/L~\cite{makal2012methane}, liquid H$_2$ at 20 K to achieve 8~MJ/L \cite{suh2011hydrogen})
or compression to high pressures (room temperature storage).
% (200 bar for CH$_4$ to achieve 9.2 MJ/L \cite{makal2012methane} and 700 bar for H$_2$ to achieve 4.5-5.3 MJ/L~\cite{unknown}).
Both approaches require expensive infrastructure at refilling stations and significant energy input; e.g., liquefaction of hydrogen consumes at least 30\% of its energy content \cite{bossel2003energy}. Moreover, high-pressure storage tanks are heavy, thick-walled, and non-conformable, while cryogenic storage tanks are bulky, expensive, and afflicted by boil-off losses \cite{hasan2009minimizing}.
% \jordan{
% Both forms of densification suffer several drawbacks varying from safety issues,  excessively expensive, boil-off~\cite{hasan2009minimizing}, and extensive refilling infrastructure~\cite{bossel2003energy}. 
% }
% % Jordan: I commented this out! % Cory: I shorted it. I think it's important.

A promising approach to densify natural gas \cite{makal2012methane,mason2014evaluating} and hydrogen \cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is through physical adsorption in nanoporous materials at room temperature~\cite{schoedel2016role}. The internal surfaces of porous materials attract gas molecules through van der Waals, electrostatic, etc. interactions to achieve a higher adsorbed gas density than the bulk gas at the same temperature and pressure, allowing for room temperature and lower-pressure storage and alleviating many drawbacks of high-pressure storage.
% This would thereby reduce the cost of infrastructure at refilling stations, allow thinner-walled and therefore cheaper and lighter pressure vessels, reduce the energy requirements for gas densification, and alleviate safety concerns with high pressure storage.

%As most homes in the US are connected to a natural gas pipeline \addcite, adsorbed natural gas fuel tanks could permit at-home fueling.

% Jordan: I commented this out!
% Cory: I think I agree that this paragraph is not necessary, thanks.
% Current research is towards synthesizing a nanoporous material that can maximally densify natural gas/hydrogen whilst satisfying cost and stability constraints. So far, metal-organic frameworks (MOFs) \cite{furukawa2013chemistry} have demonstrated the highest internal surface areas ($>$7000 m$^2$/g \cite{farha2012metal}) and absorptive capacities for hydrogen \cite{suh2011hydrogen,garcia2018benchmark} and methane \cite{makal2012methane,mason2014evaluating}. Importantly, MOFs are highly adjustable materials synthesized by combining organic molecules (struts) and metal clusters (nodes) that
% % I cut [often] here, since I'm not sure what the exception is and I don't think it's relevant.
% coordinate to form a three-dimensional, porous framework \cite{furukawa2013chemistry}. This modular synthesis enables one to fine-tune the MOF chemistry to arrive at an optimal material for methane or hydrogen storage and delivery \cite{schoedel2016role}. See Fig.~\ref{fig:example_MOF} for the crystal structure of a canonical MOF, CuBTC \cite{chui1999chemically}.

For a vehicle employing a porous material to store natural gas or hydrogen, the (volumetric) \emph{deliverable capacity} of the gas is the primary thermodynamic property of the material that determines the driving range~\cite{mason2014evaluating}. The adsorbed gas storage tank delivers the gaseous fuel to the engine via an (assumed) isothermal pressure swing \cite{sircar2002pressure}. The deliverable capacity (see Fig.~\ref{fig:delcap}) is the density of the gas in the material at the storage pressure $\pfull$ minus the residual gas that remains adsorbed at the lowest pressure $\pempty$ such that sufficient flow is maintained to feed the engine.
For commercial feasibility, the US Department of Energy (DOE) has set deliverable capacity targets for both adsorbed methane and hydrogen storage and delivery for vehicles. For methane, the Advanced Research Projects Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$ per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure swing \cite{simon2015materials}. For hydrogen, the DOE set a series of progressive targets at five year intervals, with the ultimate target of 0.05 kg H$_2$/L \cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a minimum of -40 $^\circ$C \cite{allendorf2018assessment}. Thus far, despite the emergence of highly tunable materials with enormous surface areas, such as metal-organic frameworks \cite{furukawa2013chemistry}, no porous material has met these deliverable capacity targets.

\begin{figure}
    \centering
    \subfloat[][]{
        \includegraphics[width=0.6\columnwidth]{figs/hkust-1_fancy.png} \label{fig:example_MOF}
    }
    \qquad
    \subfloat[][]{
        \includegraphics[width=0.98\columnwidth]{figs/usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
    }
    \caption{Gas storage and delivery using metal-organic frameworks (MOFs). (a) the crystal structure of an archetype MOF, CuBTC \cite{chui1999chemically}. (b) the methane adsorption isotherm in CuBTC \cite{chui1999chemically} at 298 K (blue) (data from Ref.~\cite{mason2014evaluating}). The deliverable capacity $\rho_D$ is illustrated as the density of gas in the MOF at the storage pressure $\pfull$ minus the density at the discharge pressure $\pempty$.
    }
    \label{fig:fig1}
\end{figure}

To set realistic performance targets and optimally allocate resources for research efforts, in this work, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a rigid porous material and uses as input the experimentally measured properties of the bulk gas. Our extremum is provided by a substrate that offers a spatially uniform potential energy field felt by the gas. Applying our framework to methane and hydrogen gas, we find the US DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material. Optimistically, new paradigms outside the scope of applicability of our theoretical framework, such as gas-induced structural transitions of the material, hold promise for meeting these targets, as evidenced by flexible MOF Co(bdp) which currently boasts the largest methane deliverable capacity~\cite{mason2015methane}.

% Jordan: I commented this out!
% To optimally allocate resources for research efforts and set realistic performance targets, it is important to assess the thermodynamic limits of technologies. Here, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a nanoporous material and uses as input experimental measurements of the bulk gas. Applying our framework to methane and hydrogen gas, we find the DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.

\section{Gas storage \& delivery by isothermal, pressure-swing adsorption}
% Here, we pose the context and question that we address in this work.
% Consider pure gas storage and delivery by isothermal, pressure-swing adsorption \cite{sircar2002pressure}.
Consider a pressure vessel onboard a vehicle (i.e. fuel tank) packed with porous material. At the filling stage, the tank is connected to a (pure) gaseous reservoir at pressure $\pfull$ and allowed to equilibrate. At this point, the adsorbed gas tank is considered full. While driving, gas desorbs from the adsorbent to the engine/fuel cell, driven by a pressure differential.
The tank is considered depleted/empty when the pressure has dropped to $\pempty$, the pressure at which the flow rate of gas from the tank to the engine is insufficient.
%
However, given $\pempty \neq 0$ (pulling vacuum), residual gas will remain trapped in the adsorbent. Therefore, the driving range of the vehicle is primarily determined by the \emph{deliverable capacity}, defined as the density of gas in the tank at $\pfull$ minus density of the the residual gas remaining at $\pempty$. The [isothermal, volumetric] deliverable capacity is an intrinsic property of the nanoporous material and its interaction with the gas.
%Precisely, let $\rho(p; T)$ be the density of gas in the adsorbent when in equilibrium with a bath of pure gas at pressure $p$ and temperature $T$, the \emph{adsorption isotherm}; $\rho(p,T)$ is essentially the equation of state of the gas adsorbed in the porous material. The [intrinsic, isothermal] deliverable capacity $\rho_D$ of the gas in the adsorbent is:
%\begin{equation}
%    \rho_D(\pfull, \pempty; T) \equiv \rho(\pfull; T) - \rho(\pempty; T).
%    \label{eq:delcap}
%\end{equation}
% The pressures $\pfull$ and $\pempty$ correspond to chemical potentials $\mufull$ and $\muempty$ via an equation of state for the gas.
% Eqn.~\ref{eq:delcap}
% This assumes that the tank reaches equilibrium %, neglecting the kinetics of ad/desorption
% at the end of the filling and depletion stages.

%The assumption of isothermal operation neglects (i) heat released/consumed during ad/desorption, which can diminish the deliverable capacity by raising/lowering the temperature of the adsorbent during filling/discharge \cite{mota1997dynamics,chang1996behavior}, and (ii) an engineering strategy to use waste heat from the engine to heat the adsorbent and drive off residual gas when it nears $\pempty$ \cite{gomez2014exploring}.

%Second, an engineering strategy to increase the deliverable capacity beyond the isothermal deliverable capacity is to reroute waste heat from the engine to the adsorbed gas tank during delivery. By raising the temperature of the adsorbent during the discharge stage, we drive off a fraction of the residual gas that would otherwise remain when the tank reaches $\pempty$ . In such a case, the [non-isothermal] deliverable capacity of the tank is $\rho(\pfull, T)-\rho(\pempty; T_H)$ where $T_H$ is the temperature to which we heat the adsorbent at discharge. This strategy, however, requires a more complicated design.

% \begin{limitation}\label{limit:isothermal}
% We assume isothermal pressure-swing adsorption. It is possible to achieve a greater deliverable capacity by raising the temperature of the adsorbent as the gas is extracted to drive off the residual gas.
% \end{limitation}

\section{Review of previous work}
There has been considerable work attempting to establish an upper bound on the isothermal deliverable capacity in pressure-swing adsorption.
% We now review previous work towards placing an upper bound on the isothermal deliverable capacity of hydrogen or methane by pressure-swing adsorption.
% The first body of work rests on the simplified Langmuir model; the second body of work rests upon imposing energetic models for the gas-substrate and gas-gas interactions.

Early work showed that, in the simplified Langmuir model, there exists an optimal free energy of adsorption (which determines the Langmuir constant) that maximizes the deliverable capacity $\rho_D$~\cite{matranga1992storage,bhatia2006optimum,simon2014optimizing}.
If the gas-substrate interaction is too weak (strong), too little (much) gas adsorbs (is retained) at $\pfull$ ($\pempty$), diminishing $\rho_D$. An upper bound on the deliverable capacity of a Langmuir material follows if each adsorption site is endowed with the optimal free energy of adsorption. However, remaining is the question of how many adsorption sites per volume a porous material can practically offer, under the constraint that these adsorption sites offer the optimal free energy of adsorption. Moreover, gas-gas attractions, neglected in the Langmuir model, could recruit more gas in the material at $\pfull$ than at $\pempty$ and enhance the deliverable capacity~\cite{simon2014optimizing}.

G\'omez-Gualdr\'on \emph{et al.} \cite{gomez2017impact} introduced a model that accounted for gas-gas interactions via an intermolecular potential and idealized the substrate in two different ways: (1) discrete adsorption sites packed into an FCC lattice and (2) a volume endowed with a spatially uniform, background potential energy field. According to molecular simulation of methane adsorption, the ARPA-E deliverable capacity target of 315 L STP/L could be reached in both of these idealized substrates.
%The ARPA-E target of 315 L STP/L was reached when adsorption sites were endowed with an optimal heat of adsorption of 11.5 kJ/mol and adsorption sites were $5.3$ \AA~apart. In a second model, they endowed a volume with a background potential energy field, then simulated the adsorption of methane in the volume in the grand-canonical ensemble, using a Lennard-Jones potential for methane-methane interactions. The maximal deliverable capacity of 357 L STP/L was achieved with a background heat of adsorption of 8.6 kJ/mol. 
However, both models neglected the space occupied by atoms of the porous material that are needed to endow the adsorption sites/volume with the attractive potential energy. % To address this, in the third model, they constructed pseudo materials by placing Lennard-Jones spheres in space with tunable well depths. When constrained to form a connected network, none of the psuedo-materials reached the ARPA-E deliverable capacity target.

A third body of work attempted to account for both gas-gas interactions \emph{and} steric interactions of the gas with the atoms of the porous material.
% explores the upper limits of the deliverable capacity while accounting for gas-gas interactions \emph{and} the atoms of the porous material that are required to attract the guest; these material atoms exclude the adsorbate from regions of the control volume. 
%More works obtained an empirical upper bound on the deliverable capacity by explicitly accounting for the atoms of the material needed to create the potential energy field felt by the gas molecules.
Simulations of methane adsorption in hundreds of thousands of explicit nanoporous crystal structures---both real and hypothetical---suggested that the ARPA-E deliverable capacity target is infeasible \cite{simon2015materials}; the highest simulated methane deliverable capacity was 196 cm$^3$~STP/cm$^3$. Confidence in this conclusion rests upon (i) the accuracy of the intermolecular potentials describing the molecular interactions and (ii) the sufficient sampling of material space, i.e.\ that the structures considered are representative of the set of possible materials. To further explore material space and address sensitivity to the potentials used, scaling the Lennard-Jones potential well depths of material atoms to model enhanced interactions \cite{gomez2014exploring}, placing Lennard-Jones spheres in a unit cell randomly to form ``pseudo-materials'' \cite{kaija2018high}, and generating fictitious potential energy fields via a generative adversarial network trained on zeolite structures \cite{lee2019predicting} all generated model substrates that all failed to meet the ARPA-E methane deliverable capacity target.

% Brown and Freeman \cite{bhown2011analysis}

In this work, we place a rigorous upper bound on the deliverable capacity of a pure gas in a rigid substrate by endowing a control volume with a spatially uniform background energy field. Instead of using a molecular model for the gas~\cite{gomez2017impact}, we use the experimental equation of state to account for gas-gas interactions. In addition, we prove using the calculus of variations that this spatially uniform substrate yields a maximal deliverable capacity. We use our framework to place an upper bound on the deliverable capacity of methane and hydrogen.

\section{An upper bound on the deliverable capacity of a pure gas in a rigid porous material}\label{sec:upper-bound}
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/methane-298-rho-mu}
  \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several MOFs at 298\ K as a function of chemical potential (bottom axis) and pressure (top axis). The bulk methane density is from the National Institute of Standards and Technology (NIST)~\cite{nist}. The methane adsorption isotherms in the MOFs are experimental data from Ref.~\cite{mason2014evaluating, furukawa2009storage}.
  }
  \label{fig:density-vs-mu-ch4}
\end{figure}

We now develop a thermodynamic framework that places an upper bound on the deliverable capacity of any pure gas in a rigid porous material.

The thermodynamic properties of a bulk, pure gas are characterized by an equation of state. 
Of particular interest for gas storage and delivery is the density of the gas as a function of chemical potential $\mu$ and temperature $T$, $\rho_g(\mu,T)$, which is shown for methane gas in Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K. For comparison, we also show the density of methane adsorbed into several porous materials.

% \davidsays{This paragraph could move considerably later. Or cut entirely?}
% We model the interaction of substrate with gas as a simple Gibbs free energy of adsorption per mole (or atom) \gst.  We will assume that this Gibbs free energy is independent of the concentration of gas.  At low concentrations, this assumption is reasonable, because the gas molecules can be assumed to be widely separated and their interactions weak.  In the limit of high concentrations, we expect that the attraction of the gas to the substrate will be diminished, since the ``good spots'' will already be occupied.  The choice of a concentration-independent $\gst$ will thus have a tendency to overestimate the deliverable capacity, providing an upper bound (for more on this, see Sec.~\ref{sec:qualitative-max}).

To place an upper bound on the deliverable capacity, consider a substrate whose sole interaction with the gas is to introduce a spatially uniform potential energy $\V$
for gas molecules in the control volume (where $\V<0$ for an attractive potential). Because this interaction is uniform within the substrate, the gas-gas interactions and thus fluid structure in such a substrate are identical to those of the pure gas at the same density and temperature. This allows us to obtain the adsorption properties of this model substrate using the experimentally measured properties of the pure gas~\cite{nist}. 
Intuitively, the deliverable capacity of a gas in such a homogeneous substrate with the optimal potential energy is an upper bound because, in a realistic material, (i) spatial inhomogeneity of the potential energy results in some points in the control volume offering a suboptimal attraction for the gas and (ii) the atoms required to create the potential field will exclude gas from occupying a fraction of the control volume.
%\corysays{the suboptimal energy is hard to argue since that would affect entropy too...}

% \corysays{I propose to replace ``homogenous'' with ``spatially uniform'' everywhere instead of using both interchangeably.
% \davidsays{We replaced ``homogeneous potential'' with spatially uniform, but were less confident replacing ``homogeneous substrate'', which has a slightly different meaning, and might need a slightly different name.  e.g. maybe ``uniform substrate'' or ``model substrate''?}}

%\corysays{what do u think about callin this an ``ideal substrate'' or giving it some name so we can refer to it hereon? \davidsays{I think I'd prefer ``homogeneous substrate'' or ``idealized substrate''}}

%\subsection{Finding the density}\label{sec:finding-n}

The spatially uniform potential $\V$ representing gas-substrate interactions behaves as an external potential and effectively shifts the chemical potential (or, equivalently, molar Gibbs free energy) of the gas in the material, just as gravitational potential energy causes the density of air to vary with altitude.  Consequently, the density of gas in our homogeneous substrate is:
\begin{align}
    \rho(\mu,T) &= \rho_g(\mu - \V,T). \label{eq:mof-density}
\end{align}
See Sec.~\ref{sec:V_shifts_chem_pot} for a derivation.
% where $\rho_g(\mu,T)$ is the density of the pure gas at chemical potential $\mu$, shown for methane in Fig.~\ref{fig:density-vs-mu-ch4}. 
% cory says: ^ we already defined \rho_g in pargraph above.

The deliverable capacity of gas in our homogeneous substrate is thus:
\begin{align}
    \rho_D(\V,T) &= \rho(\mufull,T) - \rho(\muempty,T),
    \label{eq:DofPhi}
    \\
    &= \rho_g(\mufull-\V,T) - \rho_g(\muempty-\V,T),
\end{align}
where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the pressures $\pfull$ and $\pempty$, respectively.
Thus, $\rho_D$ for our homogeneous substrate as a function of $\V$ is the difference between two shifted versions of $\rho_g(\mu; T)$.
Figure~\ref{fig:methane-298-D} shows the two shifted bulk methane density curves and their difference.
The horizontal axis
is the difference in molar Gibbs free energy at fixed density and temperature between the gas within the substrate and the pure gas (\gst), which is equal to $\V$ in our model.
We see a potential $\V$ that maximizes the deliverable capacity of our ideal substrate, balancing the need to maximize the density at $\pfull$ against the need to minimize residual gas retained at $\pempty$. 
%Our theoretical framework thus recovers an upper bound on the deliverable capacity of methane (298\ K, $\pfull=$65\ bar, $\pempty=$5.8\ bar) in porous materials as 374\ L STP/L.
The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, \red{achieved for $\V_{opt} = ...$}.
%
%As a ballpark, we can expect the attractive potential to be optimal around the maximum slope of $\rho_g(\mu)$.  This is the point where the steric repulsion between the gas atoms starts leading to saturation of the density.

An essential question is whether our idealized substrate offering the spatially uniform potential $\V_{opt}$ for the gas places an upper bound on the deliverable capacity in all rigid porous materials. 
%yields the maximum deliverable capacity over all possible potential energy fields. 
In the SI (see Section~\ref{sec:fixme} \davidsays{Look up section label}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest. The potential energy field is \emph{static} if it is unaffected by the presence of the gas; consequently, our model does not apply to flexible materials that undergo gas-induced conformation changes \cite{schneemann2014flexible}. We next argue that this extremum is a maximum by addressing two alternative possibilities: the extremum could turn out to be (i) a saddle point or (ii) a local, rather than global, maximum.

Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum $\rho_D$ over all \emph{spatially uniform}, static potential energy fields. To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a maximum deliverable capacity over \emph{all} (including non-spatially uniform) static potential energy fields (a broader claim), consider a non-spatially-uniform variation on $\V_{opt}$.
We will show that the effect of this variation is to reduce the deliverable capacity.
At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, the additional gas molecules will be forced into higher energy locations. Thus, the mean attraction will be lowest at \pfull\  and highest at \pempty.  This will result in $\gst$ increasing monotonically with with pressure.  As shown in the Supplementary Information (see Sec.~\ref{sec:monotonic}), this results in the deliverable capacity of a non-uniform potential being lower than the deliverable capacity of a uniform potential with $\V=\gst$ for either $\pfull$ or $\pempty$.
Thus a spatially uniform potential will give the best deliverable capacity amongst all static potential energy fields that give the same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will give the a greater deliverable than any static non-uniform potential.

% Yo David, I changed ur paragraph below. tell me wut u think.
% To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a \emph{maximum} deliverable capacity over all static potential energy fields, consider the case of a material that offers a non-spatially-uniform potential energy felt by the gas molecules.We wish to show that this material will have a lower deliverable capacity than our ideal substrate. At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, some of the gas molecules will be forced into higher energy locations. Thus, the mean attraction will be lowest when the tank is ``full,'' and highest when the tank is ``empty.'' This is the opposite of what we would want to achieve a high deliverable capacity. Based on this qualitative reasoning, we expect the spatially uniform potential \red{$\V_{opt}$} to give the maximal deliverable capacity, since it will provide the most attraction for the molecules added at high densities.

% \corysays{``To qualitatively argue that the spatially uniform potential provides a \emph{maximum} deliverable capacity'' do we need to qualify that this spatially uniform potential must hv the optimal $\V$? obviously a spatially uniform potential can be better than a spatially uniform one with a poor $\V$. maybe we should start with our optimal homogenous potential, then say introduce a little blip in it from there, like in calculus of variations.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{figs/methane-298-n-vs-G}
    \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
    Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line (not visible at this scale).}
    \label{fig:methane-298-D}
\end{figure}

\section{Results}
Although our theoretical framework allows us to place an upper bound on the deliverable capacity of many different gases in a rigid porous material, we focus on the maximal deliverable capacity of methane and hydrogen gas in our homogeneous substrate due to their application as transportation fuels. To characterize the density of the gases, $\rho_g(p, T)$, we use experimental data from NIST~\cite{nist}, which naturally includes quantum effects that particularly affect hydrogen at low temperature~\cite{kumar2006quantum}. In the context of storage onboard passenger vehicles, we compare our upper bound with several prominent porous materials using experimental adsorption isotherms from the literature; we also compare with deliverable capacity targets set by the US DOE.

Figure~\ref{fig:methane-298-D} shows the upper bound for methane storage at room temperature~(374~cm$^3$~STP/cm$^3$). In addition to the predicted maximum deliverable capacity as a function of \gst,
\corysays{subtlety here: isnt ur proof restricted to saying $\V_{opt}$ provides the upper bound? does it really prove that, for fixed and nonoptimal $\gst$, a spatially uniform potential is optimal over a non-spatially uniform one? I don't think it does...}
the ARPA-E target of 315~cm$^3$~STP/cm$^3$~\cite{arpaemove} is shown for context. For an adsorbent with at least 84\% void fraction, the ARPA-E target is theoretically possible. The experimental deliverable capacities of several adsorbents~\cite{mason2014evaluating} are also shown over a range of $\gst$ (converted from the measured adsorption isotherm as explained in \davidsays{confirm this} Sec.~\ref{sec:phi-is-delta-g}). These can be compared with the highest observed deliverable capacities for methane at room temperature in rigid materials, up to 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}. 
% don't think the below is helpful or correct: Gst opt is reached by some... there are void fraction constraints too.
% To reach the ARPA-E target, an adsorbent material with a higher $|\gst|$ would need to be found.

\corysays{I know it doesn't matter, but should we choose L STP/L or cm3 STP/cm3? we use both, some might get confused/distracted? \davidsays{We chose to use the units used for each of the targets thinking it would be more familiar and/or easier to compare.  I don't like either set of units, and if we were to pick ``our'' choice, I'd lean towards something dimensionless with a sane comparison (e.g. relative to the gas density at our high pressure)}}

\corysays{need to explain why the MOFs are lines not points... that \gst has a range for MOFs.
\jpsays{due to a different deliverable capacity at full pressure and empty pressure. For hydrogen the SI (by Paula Garcia-Holley) has total uptake plots so we could make a curve even but would have a lot of error due to manual extraction?}}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-298-n-vs-G}
    \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.  Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
    \label{fig:hydrogen-298-D}
\end{figure}

\corysays{Figs.~\ref{fig:hydrogen-298-D} ~\ref{fig:methane-298-D} why not just let $\gst$ be negative and have an arrow pointing in direction of more attraction? as written the legend is not correct since those are not the functions being shown. in the main text we never discussed what \gst is for the MOF and why the MOFs are lines not points (\gst is a function of density now)}

\jpsays{I edited everything below this comment. The original is commented out}

Storage of hydrogen is considerably more challenging owing to its relatively weak interaction with adsorbents. The DOE ULTIMATE deliverable capacity target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C \red{check}. Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve for hydrogen storage along with a range of experimental measurements for known MOFs.  The DOE ULTIMATE deliverable capacity target is theoretically possible, however by such a small margin that we can safely rule out the possibility of reaching this target through storage and delivery of hydrogen in \emph{any} rigid substrate at room temperature. Such a material would require a void fraction of at least 94\% \red{check}. On top of this, the DOE ULTIMATE target requires an optimal $|\gst|$ of 10~kJ/mol which is far greater than what is found in observed porous materials. This reflects the known fact that hydrogen interacts with substrates far more weakly than methane does. One possibility that has been pursued is storage at cryogenic temperatures. See Section~\ref{sec:cryo-hydrogen} in the Supplementary Information for the upper bound for storage of hydrogen at 77 K.

Why do we not experimentally observe materials approaching the upper bound on the deliverable capacity? Foremost, any adsorbent substrate will be composed of atoms, which will exclude the gas from some volume. Due to strong short-range interactions, substrate atoms must be uniformly distributed to achieve strong attraction for gas atoms. Uniform atom distribution sets a limit on the pore volume.
% Because strong interactions are short-range, achieving a strong attraction requires that substrate atoms be distributed throughout the volume, which places a limit on the pore volume.
In contrast, real materials have a spatially non-uniform attraction for gas atoms. The result of this is that there are regions of space with non-optimal attraction. For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.
% Finally, real materials must have a spatially non-uniform attraction for gas atoms, resulting in some regions of space where the attraction is not optimal. For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.

\section{Conclusion}
We have established an upper bound on the deliverable capacity via pressure-swing adsorption in rigid porous solids based on experimentally measured properties of pure gases. While these upper bounds do not rule out the discovery of materials that reach current DOE targets, they cast strong doubt on the possibility of achieving these goals when we consider the additional constraints imposed due to steric hindrance between substrate atoms and the adsorbate. Our upper bound does indicate that those goals cannot be exceeded by more than 19\% for methane and 6\% for hydrogen \red{check}. Fortunately, there are some limitations to these upper bounds which suggest avenues for future developments which could allow the targets to be met.

The first limitation of our proof is that we restricted ourselves to \emph{isothermal} pressure swing storage. By raising the temperature of the adsorbent during gas discharge to drive off residual gas \cite{gomez2014exploring}, the deliverable capacity could be enhanced, albeit at the cost of a more complicated engineering design of the fuel tank and vehicle.
%While this approach poses engineering challenges, in principle it could allow an arbitrarily large deliverable capacity.

% \corysays{no, cant lower filling temp. when park it will heat up then go above pressure rating of tank...}

The second limitation of our proof arises in the assumption of a rigid substrate. The rigid substrate acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
% The second limitation of our proof arises in the assumption of a rigid substrate, which acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
For most porous materials this is a reasonable approximation, and this assumption is frequently made in both the simulation and theory of porous materials~\cite{witman2017influence}. However, there are cases where the substrate can in effect provide a very strongly gas-density-dependent interaction through structural flexibility~\cite{schneemann2014flexible}. A flagship example is MOF Co(bdp)~\cite{choi2008broadly}, which possesses a wine-rack-like topology capable of hinge motion. At low methane pressure, Co(bdp) adopts a collapsed, nonporous state, but expands to a porous state and fills with gas at higher pressures \cite{mason2015methane}. This allows Co(bdp) to fully expel its residual gas at low pressures.
Our upper bound does not consider such material effects and thus flexible materials could have significantly higher deliverable capacities.
% Such materials can indeed violate our upper bound and thus could lead to significantly improved deliverable capacities.

We note that our upper bound can readily be applied to the storage of other gasses of interest, with the proviso that the gas is far from crystallization. Our code is available at \davidsays{XXXXX}.

% \clearpage

% Here we will articulate the assumptions required for our upper bound to hold. Some of these assumptions serve as a guide for chemists who aim to synthesize a material that violates our upper bound.

% Our proof that the spatially uniform potential is an extremum assumes that the fluid is stable at the temperature and densities involved.
% \begin{limitation}\label{limit:crystallization}
% Our proof assumes the gas is cannot become spontaneously solid at the temperature of interest.
% \end{limitation}
% Our upper bound will \emph{not} be valid for gasses which may crystallize in their pure form. This is because a periodic potential can interact with a high-density crystal far more strongly than it interacts with a low-density gas.  We are not aware of a practical application of this principle, but theoretically it is possible.  For the application of gas storage for light gasses such as methane or hydrogen at moderately high temperatures (say liquid nitrogen or warmer), the pure gas is sufficiently far from its crystal form that we need not concern ourselves with crystallization.  We do not see this particular assumption as one that can be worked around.

% There remains the possibility that the spatially uniform potential is a local maximum, but there is another global maximum.  We do not see how to formally rule out this possibility for an arbitrary gas-gas interaction, but the qualitative arguments above seem sufficient to rule out this possibility for a real gas.

% However, there was a more fundamental assumption, which is that the porous material acts as an effective potential for the gas molecules, and that the gas-gas interaction is unmodified.
% \begin{limitation}\label{limit:field}
% Our proof assumes the interaction of the gas with the substrate does not modify the substrate.  Equivalently,  the interaction of substrate with gas operates as a simple potential energy field.
% \end{limitation}
% For \emph{most} porous materials this seems likely to be a pretty good approximation.  However, there are unusual cases where the substrate can in effect provide a very strongly density-dependent interaction.  This is perhaps easiest to see in the example of \davidsays{look up the collapsing MOFs description from the RMS-MOF paper}, in which the gas can introduce a phase change in the solid, which has the effect of drastically lowering the density of gas in the material.  Such materials can indeed violate our upper bound. \davidsays{Remember to discuss this in the conclusions or perhaps discussion, we can frame these limitations as guides.  The limitations of our upper bound are the design requirements for a material that will do any better.}

% [impurities?]

% \section{Conclusion}
% As many have suspected \davidsays{cite previous work}, pressure-swing adsorption in rigid porous solids is not going to enable reaching targets.  \davidsays{conclude explicitly about each target}  Unlike previous work, our upper bound on deliverable capacity is based on experimentally known properties of the gas of interest rather than simulations.

% The limitations on our proof of an upper bound on the deliverable capacity provide hints as to approaches that could in principle achieve the targets.  In this paper we have identified three limitations of our upper bound on the deliverable capacity.  Two of these three limitations suggest engineering strategies that do hold promise of achieving the targets for hydrogen and methane storage.  Limitation \ref{limit:crystallization}, which states that the gas must not crystallize does not means for improvement, since it is a limit on the gasses to which our bound may be applied, and neither hydrogen nor methane are in any danger of crystallizing.

% Crystal density vs. pelletiziation density.

% A more subtle limitation of our proof is that it assumes that the substrate itself is not modified by the presence of the gas.  While this assumption is never precisely true, since any adsorbed atom will change the electron distribution of the substrate to some degree, it tends to be sufficiently accurate that most atomistic simulations can get away with assuming that the substrate is unmodified.  A dramatic exception to that rule is the field of flexible MOFs, in which the presence of the gas can induce a phase change which can \emph{drastically} modify the gas-substrate interaction~\davidsays{cite experiments}\cite{witman2019}.

% wut about mixtures?

%TC:ignore

\bibliography{bibfile}

\appendix
\clearpage
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}%
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{SI \Alph{section}}%

\section{An external, spatially uniform potential $\V$ shifts the chemical potential $\mu$ of the gas} \label{sec:V_shifts_chem_pot}
We show that imposing an external, spatially uniform potential $\V$ to a gas has the effect of shifting the chemical potential $\mu$ of the gas, recovering eqn.~\ref{eq:mof-density}. Consider a control volume $\Omega$ with volume $V=|\Omega|$ (large enough to neglect boundary effects) that is endowed with the external, spatially uniform potential $\V$. Impose the grand-canonical ensemble, where this control volume can exchange energy and particles with a bath of gas at temperature $T$ and chemical potential $\mu$.

To denote a microstate of this system, let $N$ be the number of gas particles in the control volume and $\mathbf{x}_1,...,\mathbf{x}_N$ be their positions. Then, the potential energy $E$ of a microstate of the control volume is:
\begin{equation}
    E(\mathbf{x}_1,...,\mathbf{x}_N) = N\V + E_{gg}(\mathbf{x}_1,...,\mathbf{x}_N),
\end{equation} where $E_{gg}$ is the (unknown and complicated) interatomic potential for gas-gas interactions that governs the (real) gas properties. The first term arises from each gas molecule feeling the external potential $\V$, where $\V < 0$ corresponds to attraction.

\corysays{wut about orientations? shld we include those?}

The grand-canonical partition function of the control volume is:
\begin{multline}
    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\xvec_1, ..., \xvec_N)} e^{\beta (\mu - \V) N} d\xvec_1 \cdots d\xvec_N.
    \label{eq:gcpf}
\end{multline} We recognize this as the grand canonical partition function of the bulk gas in the control volume without the external potential, but shifted in chemical potential:
\begin{equation}
    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
    \label{eq:xi_vs_xi0}
\end{equation}
where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the bulk gas in the control volume in the absence of an external potential. Importantly, this equivalency is predicated on the potential $\V$ being spatially uniform. Therefore, the thermodynamic properties of the gas sitting in the spatially uniform external potential $\V$-- i.e., adsorbed in our idealized substrate-- are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$ ($T$ fixed here). 
As $\V$ becomes more negative, corresponding to a more attractive adsorbent, the thermodynamic properties of the adsorbed gas in our ideal substrate are equivalent to the gas at higher chemical potential.

\section{Proof of extremum}
To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.  In this proof we will make use of the Fourier transform of this potential:
\begin{align}
    \Vk \equiv \iiint \V(\vec r) e^{-i\vec k\cdot \vec r} d\vec r
\end{align}
The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$. Therefore, to show that a uniform potential extremizes the deliverable capacity, we must show that the functional derivative of the deliverable capacity with respect to $\Vk$ is zero for \emph{nonzero} values of $\vec k$, i.e.
\begin{align}
    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \vec k\ne 0.
\end{align}
We note that this functional derivative may be non-zero for $\vec k=0$ because we separately maximize with respect to the particular uniform potential $\V$.
This means that
\begin{align}
    \frac{\delta N_H}{\delta \V(\vec k)} &= -\frac{\delta N_L}{\delta \Vk}
\end{align}
where $N_H$ and $N_L$ are the number of particles at the low and high pressure.

Because the chemical potential $\mu$ varies monotonically with $N$ at fixed temperature, we can consider how the chemical potential varies as we change $\Vk$ with the number of molecules held fixed.  We demonstrate this using the cyclic chain rule, which shows us that
\begin{align}
    \left(\frac{\delta N}{\delta \Vk}\right)_{\mu} &=
    -\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}
    \left(\frac{\partial N}{\partial \mu}\right)_{\Vk}.
\end{align}
Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \V(\vec k)}\right)_{\mu}=0$.  Thus we consider
\begin{align}
    \left(\frac{\delta \mu}{\delta \tilde\V(\vec k)}\right)_N
    &= \left(\frac{\delta \left(\frac{\partial F}{\partial N}\right)_{\volume}}{\delta \Vk}\right)_N
    \\
    &= \left(\frac{\partial \left(\frac{\delta F}{\delta \Vk}\right)_{\volume}}{\partial N}\right)_{\volume}
    \label{eq:dmudpot}
\end{align}
where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reordered the functional and partial derivatives.
Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ just gives the number density:
\begin{align}
    \frac{\delta F}{\delta \Vk} &= \rho(\vec k)
\end{align}
The number density is itself spatially uniform for any system that is stable in a fluid state at this density (i.e. does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.  Thus the functional derivative $\frac{\delta F}{\delta \V(\vec r)}$ is actually a uniform function.
We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
\begin{align}
    \frac{\delta \mu}{\delta \Vk} &\propto \delta(\vec k) \\
    \frac{\delta N}{\delta \Vk} &\propto \delta(\vec k)
\end{align}
Thus the functional derivative of both the chemical potential and $N$ with regard to $\V(\vec r)$ are themselves spatially uniform.  Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e. $\vec k=0$), the derivative of $D$ with respect to any change of potential is zero.

This demonstrates that a spatially uniform potential leads to an extremum value of the deliverable capacity.  This proof is insufficient, however, to show that it must be a true maximum.

\section{The real-substrate analogy of $\V$ is $\gst$}
\label{sec:phi-is-delta-g}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/g_st_piston_2.png}
    \caption{Cartoon illustrating the definition of $\gst$ and a hypothetical experiment to measure it with a piston and removable barrier that separates two equal volumes, one with a porous material (bottom) and one with free space (top). The initial state (left) is a bulk gas phase at density $\rho$ and temperature $T$. An infinitely-thin, removable barrier separates the gas from the porous material (bottom). The barrier is then slowly removed, and the piston slowly pushes all gas molecules into the porous material in an isothermal process. The final state (right) is an adsorbed gas phase at density $\rho$ and temperature $T$. The quantity $\gst$ follows from {\color{red} the heat released?}.}
    \label{fig:delta-G-cartoon}
\end{figure}

The parameter describing our idealized substrate is $\V$, the spatially uniform potential felt by a gas molecule adsorbed in the idealized substrate. 
A natural question is how this potential relates to the properties of real substrates. 
The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the gas (see Sec.~\ref{sec:V_shifts_chem_pot}).
Because our ideal substrate shifts the chemical potential of the gas molecules by providing a spatially uniform potential energy field, the entropy of the gas in the ideal substrate is equal to the entropy of the gas in its bulk state at the same density and temperature.
In contrast, a real substrate provides a non-spatially uniform potential. Consequently, the entropy of the gas inside a real substrate is \emph{not} equal to the entropy of the bulk gas at the same density and temperature. 
Therefore, the parameter analogous to $\V$ in a real substrate will involve both energy and entropy.
The real-substrate analogy of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
\begin{equation}
   \gst(\rho, T) \equiv
   %\frac{G_{\text{ads}}(\rho, T) - G_{\text{gas}}(\rho, T)}{N}\\ 
    \mu_{\text{ads}}(\rho, T) - \mu_{\text{gas}}(\rho, T).
  % \gst(\rho, T) &\equiv g_{\text{gas}}(\rho, T) - g_{\text{ads}}(\rho, T)
  \label{eq:g_st}
\end{equation}
The isosteric Gibbs free energy difference $\gst$ is the difference in molar Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and the bulk gas \emph{with the same density of gas molecules}. The quantity $\gst$ does \emph{not} correspond to a change in the molar Gibbs free energy as a molecule is adsorbed, which is zero under conditions of coexistence. 
The quantity $\gst$ in a real substrate is a direct analogy to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to  measure $\gst$ via a piston with a removable partition that separates a volume of free space from the same volume of substrate. Note that $\gst$ is a property of both the substrate and the identity of the gas. Because real substrates offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho, T)=\V$. Consequently, throughout this article, we show $\gst(\rho, T)$ for real substrates at both conditions relevant to gas storage and delivery, $\pfull$ and $\pempty$. 
% In the initial state, the partition isolates the bulk gas exhibiting density $\rho$ and temperature $T$. Then, the partition is removed and the piston pushes down on the gas so that the gas slowly adsorbs into the MOF in an isothermal process. \red{isothermal rite?} In the final state, the piston has pushed all gas into the MOF, resulting in an adsorbed gas phase with the same density $\rho$ and temperature $T$ as the initial state. The isosteric Gibbs free energy of adsorption, $\gst$, then follows from \red{is it the amount of heat released?}.

\corysays{Fig.~\ref{fig:qst-vs-delta-G} looks too jumbled. two separate figs? plus $G_{st}$ is capitalized here, should be lowercase. where is all this data from}

In practice, we can readily compute $\gst(\rho, T)$ of a real gas/substrate system from (i) the (experimental or simulated) equilibrium adsorption isotherm of the gas in the substrate and (ii) the (experimental or simulated) chemical potential of the bulk gas. 
Consider the real substrate in thermodynamic equilibrium with a bulk gas at fixed temperature $T$ and pressure $p$, and let $\rho=\rho(p, T)$ be the density of gas in the substrate. 
At coexistence, the chemical potential of the bulk gas is equal to the chemical potential of the adsorbed gas in the substrate. Thus, we can use the experimentally known molar Gibbs free energy of the pure, bulk gas system at temperature $T$ and pressure $p$ to determine the molar Gibbs free energy of the adsorbed system: $\mu_{\text{ads}}(\rho, T)=\mu_{\text{gas}}(p, T)$. We can then also look up the known chemical potential of the bulk gas at the same density and temperature as in the substrate, $\mu_{\text{gas}}(\rho, T)$. Via eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.

An interesting question is how $\gst$ relates to the commonly measured and reported isosteric heat of adsorption $q_{st}$, which is roughly the energy change when a gas molecule is adsorbed~\cite{sircar1999isosteric, tian2017differential}. Figure~\ref{fig:qst-vs-delta-G} shows how $q_{st}$ compares to $\gst$ for several prominent adsorbents. In every case, $|q_{st}|>|\gst|$ because the gas in the adsorbent always has less entropy than the gas in the bulk at the same density and temperature. That is, while adsorption is energetically favored, it is entropically disfavored due to the restrictions imposed on the configuration of the gas molecules via steric interactions with the substrate itself; this counters the energetic attraction.

\corysays{can u mathematically justify why $\gst$ is the relevant parameter through the grand canonical partition function in eqn.~\ref{eq:gcpf}? it is not true that $\Xi(\mu, V, T)$ for the real substrate is equal to $\Xi(\mu+\gst, V, T)$}

\corysays{Fig.~\ref{fig:qst-vs-delta-G} should be split in two. one for methane one for h2.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{figs/qst-vs-delta-G}
    \caption{Relationship between $\gst$ and the isosteric heat $q_{st}$ for several prominent adsorbents. The dots represent the properties of methane adsorption at 298\ K and at 5\ bar and 100\ bar. The $\times$'s represent the properties of hydrogen adsorption at 77\ K and at 5~bar and 100~bar.}
    \label{fig:qst-vs-delta-G}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{figs/methane-298-gst}
    \caption{The density-dependence of $\gst$ of methane in several adsorbents (298\ K). Note that $\gst$ is monotonic in $\rho$. Data from Refs.~\ref{mason2014evaluating,furukawa2009storage}.
    }
    \label{fig:methane-gst}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{figs/four-cases}
    \caption{Four cases we consider.}
    \label{fig:delta-gst-maximum}
\end{figure}

\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding to a full and empty density. The deliverable capacity of the MOF is equal to the difference between the full and empty density.
Examining Fig.~\ref{fig:methane-gst}, we see that a significant variety of known experimental $\Delta g_\text{st}$ curves are monotonic.  Furthermore, our qualitative argument in Section~\ref{sec:upper-bound} suggests that this function \emph{should} be monotonically increasing for rigid substrates, as increasing the density of gas forces some of the gas to reside in higher-energy sites.  If this is always the case, the deliverable capacity will always be bounded by our theoretical upper bound, even for a value of $\V$ that is not optimal.

To show this, we perform a thought experiment illustrated in Fig.~\ref{fig:delta-gst-maximum}.  We begin with two pairs of boxes containing gas.  Two of the boxes are filled with a MOF, and the other two have a uniform potential energy $\V$.  We could imagine the two boxes at $\V$ as being empty but at very low altitude where gravity provides a potential energy difference (relative to the box with the MOF in it) of $\V$ to each gas molecule (perhaps we are on Jupiter?).  One box of each sort has a density of gas $\rhofull$ corresponding to $\pfull$ in the MOF system, while the other two boxes has gas density $\rhoempty$.

We now consider what happens if we introduce a diffusive connection between two boxes that are initially at the same density, for instance by connecting them with a tube.  If the chemical potential of the two boxes are equal, their density will be unchanged, otherwise gas will flow into the box with lower chemical potential.  The difference in chemical potential is equal to the difference
\begin{align}
   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
   \\
   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
\end{align}
Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at their initial density after they are connected.  Thus the deliverable capacity of the MOF will be greater than the deliverable capacity of the model with uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$ does \emph{not} monotonically increase.  By the same token, if we consider the case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable capacity than our model, the MOF must have less residual gas, which means that gas must spontaneously flow from the low-density MOF to the box with potential $\V$, which means that $\gst(\rhoempty)>\V$.  So once again exceeding our upper bound requires a material with a $\gst(\rho)$ that does not increase monotonically.

Taken together, this indicates that not only is our absolute upper bound an upper bound for rigid MOFs, but the green curve labeled $\rho_D(\gst)$ in Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and \ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal $\gst$.

\section{Cryogenic hydrogen storage}\label{sec:cryo-hydrogen}
\begin{figure}
    \centering
    FIGURE HERE A B C D
    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-77-n-vs-G}
    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$. Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
    \label{fig:hydrogen-77-D}
\end{figure}

One approach to increase the deliverable capacity is to reduce the storage temperature. This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling point of nitrogen. The DOE ULTIMATE target in this case looks far more achievable, and with a much lower $|\gst|$. In fact, an empty tank at this temperature can satisfy the DOE 2020 target. The DOE ULTIMATE target is 14\% below the upper bound. Actual MOFs fall far short of the theoretical maximum. \red{the target is not just a storage density, it specifies temps and pressures too... is 77 K allowed?} \davidsays{My recollection of my reading of the DOE target was that anything was allowed, but that the target included ``everything'', which would include both the tank and any cryogenic equipment required.  So the answer is not so simple.}

%TC:endignore

\end{document}

### 2020-05-01 gas-adsorption/old_storage/APS-March-2020-abstract.tex
added gas storage to thesis pommerenck.

\documentclass[letterpaper,onecolumn,amsmath,amssymb,jcp,aps,12pt]{revtex4-1}
\usepackage{xargs}      % Use more than one optional parameter in a new commands
\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}
\usepackage{color}
%\usepackage[utf8]{inputenc}
%\usepackage[english]{babel}
\usepackage[document]{ragged2e}

\begin{document}

\pagenumbering{gobble}
\begin{center}
    \textbf{Abstract}: Limited to 1300 characters
\end{center}

\center{}
\textbf{Authors:} Jordan K. Pommerenck, Cory M. Simon, David J. Roundy

\justify{}
The transportation sector accounts for 38\% of US energy-related carbon dioxide emissions \cite{useia} and generates toxic air pollution (particulate matter, ozone, NO$_x$, SO$_x$, carbon monoxide, volatile organic compounds) \cite{caiazzo2013air}. Alternative transportation fuels, such as natural gas and hydrogen, are therefore critical to mitigate climate change \cite{mcglade2015geographical} and improve air quality.
%
Since they possess a low volumetric energy density compared to (liquid) gasoline, both natural gas and hydrogen gas must be densified such as through physical adsorption on nanoporous materials \cite{schoedel2016role} in order for reasonable light vehicle operating range.  
%
The US Department of Energy (DOE) has set on-board light vehicle storage targets for adsorbed natural gas and hydrogen. We provide a theoretical upper bound for the deliverable capacity of natural gas and hydrogen that applies to any uniform potential nanoporous material via pressure swing adsorption. We compare experimental data for MOFs to the upper bound with $q_\text{st}$ converted to $G_\text{st}$. We conclude that the DOE storage targets are theoretically impossible for room temperature hydrogen storage but could certainly be met by increasing the release temperature above the storage temperature. We also conclude that while the DOE storage targets are theoretically possible for natural gas, any applicable adsorbent would need a prohibitively high void fraction.

\justify{}
\textbf{Keywords:} MOF, natural gas, hydrogen, upper bound, vehicular storage

\bibliography{bibfile}
\end{document}

### 2020-05-01 gas-adsorption/old_storage/APS-abstract.tex
added gas storage to thesis pommerenck.

\documentclass[letterpaper,onecolumn,amsmath,amssymb,jcp,aps,12pt]{revtex4-1}
\usepackage{xargs}      % Use more than one optional parameter in a new commands
\usepackage{titlesec}
\newcommand{\sectionbreak}{\clearpage}
\usepackage{color}
%\usepackage[utf8]{inputenc}
%\usepackage[english]{babel}
\usepackage[document]{ragged2e}

\begin{document}

\pagenumbering{gobble}
\begin{center}
    \textbf{Abstract}: Limited to 1300 characters
\end{center}

\center{}
\textbf{Authors:} Jordan K. Pommerenck, Cory M. Simon, David J. Roundy

\justify{}
The transportation sector accounts for 38\% of US energy-related carbon dioxide emissions \cite{useia} and generates toxic air pollution (particulate matter, ozone, NO$_x$, SO$_x$, carbon monoxide, volatile organic compounds) \cite{caiazzo2013air}. Alternative transportation fuels, such as natural gas or hydrogen, are therefore critical to mitigate climate change \cite{mcglade2015geographical} and improve air quality.
%
Both natural gas and hydrogen (gas) possess a very low volumetric energy density compared to (liquid) gasoline. Consequently, under storage space constraints in passenger vehicles, they must be densified such as through physical adsorption on nanoporous materials \cite{schoedel2016role} in order to drive a reasonable distance on a full tank of fuel.  
%
The US Department of Energy (DOE) set storage targets for adsorbed natural gas and hydrogen onboard light vehicles.
To assess the feasibility of these targets, we provide a theoretical upper bound on the density of natural gas and hydrogen that can be stored in and delivered by nanoporous materials via a pressure swing.
We conclude that, while the DOE storage targets are theoretically possible, the material would require a void fraction that is outside the range of void fractions in known materials exhibiting sufficient interactions with the gas.

\justify{}
\textbf{Keywords:} MOF, natural gas, hydrogen, void fraction, vehicular storage

% used if I wanted a new page with new section and numbering
%\section{NextSection}
%\pagenumbering{arabic}

\bibliography{bibfile}
\end{document}

### 2020-05-01 gas-adsorption/old_storage/isosteric-heat.tex
added gas storage to thesis pommerenck.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{amsmath}

\usepackage{nicefrac} % Include for inline fractions
\usepackage{url}

% \usepackage{cite}
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
%
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2} }
\usepackage{mdframed}

\newcommand{\addcite}{{\bf \color{red} []}}
\newcommand{\xvec}{\mathbf{x}}
\newcommand{\thetavec}{\boldsymbol \theta}

\newcommand\V{\Phi}
\newcommand\Vk{\tilde\Phi(\vec k)}
\newcommand\volume{V}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommandx{\davidsays}[2][1=inline]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{David says:} #2} }
\newcommandx{\jpsays}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }
\newcommand{\jpinline}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }

\begin{document}
\title{Adsorbed, absorbed or sorbed gasses?}
% other ideas:
%Gas delivery via pressure-swing adsorption in nanoporous materials: setting storage targets}
% A thermodynamic upper bound to gas delivery via pressure-swing adsorption in nanoporous materials:}

\author{Jordan K. Pommerenck}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}
\author{Cory M. Simon}
\affiliation{School of Chemical, Biological, and Environmental Engineering,
  Corvallis, OR 97331}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  We discuss the isosteric heat of adsorption, and how it is not the energy of adsorption and how it relates to
  $\Phi$.
\end{abstract}

\maketitle

\subsection{New ideas/things/points to make}
\begin{enumerate}
    \item ($Q_{st}$) Net vs. absolute for Clausius Clapeyron?
    \item ($Q_{st}$) Can't use ideal gas for Qst Clausius Clapeyron things at 100 bar (try it!) Quantify error introduced.
    \item ($Q_{st}$) Relate Qst to the actual adsorption energy, the potential field created by the MOF. (Also to other differences)
    \item ($Q_{st}$) Relate calorimetry to T dependence of isotherms
    \item how does $Q_{st}$ relate to the actual potential energy that an adsorbate experiences while in the MOF?
    \item Point out that by measuring nabs vs p they measure the Gibbs free energy
      (thus entropy comes from the T dependence)...
      and does this lead to other ways of analyzing experimental data?
      Question: How to find $Q_{st}$ from the Gibbs free energy?
    \item Address mechanical pressure theoretically?
    \item Calculating $Q_{st}$: method of interpolating the data: linear interpolation vs model fitting vs splines. An advantage of model: allows you to extrapolate to both low and high pressures.
\end{enumerate}
\subsection{Ideas/things/points that others have made which we should emphasize}
\begin{enumerate}
 \item Explain carefully excess/net/absolute distinction.
    \item Argue carefully that surface excess is a red herring \textbf{This has already been done!}
\end{enumerate}

\section*{todo Tuesday June 25}

\begin{description}
    \item[Cory] does intro and abs/net/exc figure.
    \item[Jordan] does n vs T isothermal and p vs T isosteric plot and maybe qst (S37 from supplemental could be interesting) from Long's data.  If there is no maximum excess then look for other data?
    \item[David] visits relatives, thinks about what we can get from the Gibbs free energy (available volume)
\end{description}

\newcommand\Nabs{N_{\text{abs}}}
\newcommand\Sabs{S_{\text{abs}}}
\newcommand\Gabs{G_{\text{abs}}}
\newcommand\sabs{s_{\text{abs}}}
\newcommand\vabs{v_{\text{abs}}}
\newcommand\muads{\mu_{\text{ads}}}

\newcommand\Nmof{N_{\text{MOF}}}
\newcommand\Vmof{V_{\text{MOF}}}
\newcommand\mumof{\mu_{\text{MOF}}}

\newcommand\Ngas{N_{\text{gas}}}
\newcommand\Sgas{S_{\text{gas}}}
\newcommand\sgas{s_{\text{gas}}}
\newcommand\Vgas{V_{\text{gas}}}
\newcommand\vgas{v_{\text{gas}}}
\newcommand\mugas{\mu_{\text{gas}}}

\newcommand\pisostere{p_{\text{iosteric}}} % this is the pressure isotherm with the density also held fixed

\section{Introduction}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{MOF5-excess-methane-isotherm.pdf}
    \includegraphics[width=0.95\columnwidth]{MOF5-absolute-methane-isotherm.pdf}
    \caption{The absolute, net, and excess adsorption isotherms for methane in MOF-5
    taken from Mason and Long paper 2014.}
    \label{fig:mof-5-isotherms}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{MOF5-excess-methane-isostere.pdf}
    \includegraphics[width=0.95\columnwidth]{MOF5-absolute-methane-isostere.pdf}
    \caption{The absolute, net, and excess adsorption isosteres for methane in MOF-5.  Note that the pressure is not a function of temperature for an excess adsorption isostere.}
    \label{fig:mof-5-isosteres}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{MOF5-excess-methane-Qst.pdf}
    \includegraphics[width=0.95\columnwidth]{MOF5-absolute-methane-Qst.pdf}
    \caption{The absolute, net, and excess adsorption Qst for methane in MOF-5.}
    \label{fig:mof-5-isosteres}
\end{figure}

\subsection{Why gas storage matters (Cory TODO)}

\section{Definitions (Cory TODO)}
\emph{Densities excess/absolute/net.  Which matter for what.  While explaining, refer to Fig.~\ref{fig:mof-5-isotherms} and Fig.~\ref{fig:mof-5-isosteres}.}

\section{Density measurement}
\davidsays{I don't think we can write this paper without \emph{some} discussion of what is actually being measured.
  Hopefully with less detail than a few other papers we like, which we will cite.}
\subsection{Gravimetric measurement?}
\subsection{Volumetric measurement?}
\subsection{``Breakthrough'' measurement?}

\section{What can we extract from isotherms?}

\subsection{Isosteric heat}
Aside from the density isotherms, the most widely reported adsorption property is the isosteric heat.  In this section, we will define this quantity, explain how it is measured experimentally, and discuss its properties~\footnote{Add citation for a few papers about it}.  The most widely used approach for measuring the isosteric heat uses the Clausius-Clapeyron relation together with an assumption that the gas phase is ideal~\footnote{Who gets cited for Clausius-Clapeyron equation?}.  We will begin by presenting a derivation of this result that starts with the Gibbs-Duhem relation.
\subsubsection{Gibbs-Duhem}
We will begin by reminding the reader of the origin of the Gibbs-Duhem relation.  It begins with Euler's homogeneous function theorem, which shows that an extensive function of extensive variables must be given by a product of its inputs with the partial derivative with respect to those inputs.  When applied to a system with adsorbate as well as an adsorbent, we must consider the number of each species as a separate extensive quantity.  Thus
\begin{align}
    U &= TS - p\Vmof + \mumof \Nmof + \muads \Nabs
\end{align}
where we $\Nabs$ refers to the \emph{absolute} number of molecules adsorbed, which is necessary because the pressure $p$ in this expression is the derivative of the internal energy with respect to the \textbf{total volume FIXME make this align with definitions above} of the system.
Taking a differential of this expression and comparing with the thermodynamic identity gives us the Gibbs-Duhem relationship
\begin{align}
     0 &= S dT - \Vmof dp + \mumof d\Nmof + \Nabs d\muads
\end{align}
Once we agree that the quantity of MOF will not change, we can eliminate $d\Nmof$ from the equation, obtaining the more familiar expression.
\begin{align}
     \Nabs d\muads &= -S dT + \Vmof dp
\end{align}
We note that this expression is true for the pure gas as well as the adsorbed gas.  We can now compare equal quantities of gas in equilibrium with the adsorbent.  Since the two are in diffusive equilibrium, both systems will have equal chemical potentials.  Thus
\begin{align}
    0 &= Nd\mugas - Nd\muads \\
    &= (-\Sgas dT + \Vgas dp) - (-\Sabs dT + \Vmof d\pisostere) \\
    &= (\Sabs - \Sgas)dT - (\Vmof - \Vgas)d\pisostere \\
    \left(\frac{\partial p}{\partial T}\right)_{\Nabs} &= \frac{\sabs - \sgas}{\vabs - \vgas}
\end{align}
where $\sabs$ is the molar entropy of the adsorbate, $\sgas$ is the molar entropy of the coexisting gas, $\vabs$ is the molar volume of the adsorbate, and $\vgas$ is the molar volume of the coexisting gas.  At this point, we can identify the isosteric heat as
\begin{align}
    Q_{st} &\equiv T(\sgas - \sabs) \\
    &= T(\vgas - \vabs)\left(\frac{\partial p}{\partial T}\right)_{\Nabs}
\end{align}
At this point, two approximations are commonly made.  One is to assume that the density of the gas is much lower than the density in the MOF.  This assumption is false at sufficiently high pressures---i.e. the density of the gas is \emph{greater} than the adsorbed density---due to steric repulsion with the MOF atoms.  The second assumption is that the gas is ideal, which is also a poor approximation at high pressures.  In the low-pressure limit, we find the commonly used expression~\footnote{Cite numerous papers here that use this approximation}
\begin{align}
    Q_{st} &= \frac{RT^2}{p}\left(\frac{\partial p}{\partial T}\right)_{\Nabs}
\end{align}
which is derived by~\footnote{Who is cited for this?}.
\davidsays{We should demonstrate how and when this breaks down.}

\subsubsection{Clausius-Clapeyron}
\subsubsection{Different expressions for Qst}
\subsubsection{ideal gas simplifications}
\paragraph{How big an error is it?}
\subsection{Gibbs free energy}
\emph{We know the density as a function of Gibbs free energy, which is readily determined from the bulk pressure in equilibrium.  Once $G(\rho,T)$ (or equivalently $\mu(\rho,T)$) is known we can work with that more easily?}

\begin{align}
    dG &= -SdT + Vdp + \mu dN
\end{align}
which reminds us that
\begin{align}
    S &= -\left(\frac{\partial G}{\partial T}\right)_{p,N}
\end{align}
Our definition of $Q_{st}$ was
\begin{align}
    Q_{st} &\equiv T\left(\frac{\Sgas}{\Ngas} - \frac{\Sabs}{\Nabs}\right)
    \\
    &= T\left(\frac{\Sgas}{\Ngas} + \frac{1}{\Nabs}\left(\frac{\partial \Gabs}{\partial T}\right)_{\Nabs}\right)
\end{align}

\subsection{Entropy}

\emph{What entropy should we talk about?}

\bibliography{bibfile}

\clearpage
\appendix

\section{Literature Review}
\jpsays{I will attempt to break up papers of interest by topic using forward/backward search.}
\subsection{Derivation of $q_\text{st}$}
\begin{enumerate}
\item {\textbf{1985}} We read together Sircar et al.~\cite{sircar1985excess} which derives $q_\text{st}$ in a long and notationally difficult way (Cory noted the Maxwell relation not specifically stated in this paper).

\item {\textbf{1998}} Huanhua Pan et al. greatly improves on this derivation of $q_\text{st}$~\cite{pan1998examination}. He has a much cleaner derivation, notes Maxwell relation, and discusses how assumtions (1) neglecting the adsorbed phase molar volume and (2) assuming an ideal gas are made to get to Classius-Clapeyron equation. These assumptions can be avoided by using 3 different methods which they discuss in detail.

\item {\textbf{1999}} Sircar et al. derives $q_\text{st}$ using both GSE and conventional model. He is a strong proponent for using surface excess. He derives the heat of adsorption and notes assumptions necessary to use it in conjuction with isostere data~\cite{sircar1999isosteric}.

\item {\textbf{2002}} Myers and Monson use a simple thermodynamic model, consisting of the Langmuir equation for the adsorption isotherm and the ideal gas equation of state for bulk properties, to calculate excess and absolute thermodynamic variables~\cite{myers2002adsorption}.

\item {\textbf{2017}} Tian et al. cites~\cite{pan1998examination} stating that the Gibsean formalism [surface excess] predicts a nonphysical isosteric heat when the adsorption isotherm exhibits a maximum He cites~\cite{myers2014physical} for this idea. The paper introduces a rigorous thermodynamic procedure for heat analysis that is free of inconsistency yet convenient for practical applications. The controversial issues with conventional isostere methods have also been clarified in the context of the exact results~\cite{tian2017differential}.
\textcolor{red}{This and Myers and Monson work both hint at or directly argue that surface excess is `red herring'}
% New Subsection
\subsection{Calorimetry experiments}
\item {\textbf{1996} and \textbf{2001}} Dunne et al. (Myers and Sircar both coauthors) describes calorimetry experiments in detail~\cite{dunne1996calorimetric}. Cao and Sircar also further describe the calorimetry measurement process in this work~\cite{cao2001heats}.

\item {\textbf{2005}} Sircar et al. references his derivation $q_\text{st}$ using GSE (He is a strong proponent of this). He references experiments using Tian–Calvet type micro-calorimeters to obtain isostere data~\cite{sircar2005heat}.

% New Subsection
\subsection{Net, Excess, and Absolute Adsorption}
\item {\textbf{2013}} Talu et al. has an excellent discussion of the measurable properties of microporous adsorption (what things do experimentalists measure). Also, Figure 1 illustrates accessible regions for storage using the Gibbs `hyper-surface' (GSE) for each type of adsorption (N, E, or A)~\cite{talu2013net}.

\item {\textbf{2014}} Myers and Monson state the GSE is inappropriate for porous solids. Absolute sorption should be reported particularly for high pressures 100 bar or more~\cite{myers2014physical}.

\corysays{how exactly do they get the pore volume of the MOF?}
\textbf{ANSWER} The pore size distribution from the gas adsorption method is commonly analyzed from the nitrogen or Ar adsorption (87.3 K) isotherm at their boiling temperature. From the isotherm data, DFT is used to calculate the pore size (with the user specifying the pore geometry~\cite{groen2003pore}.

\corysays{why is cryogenic Ar useful for determining micropore volume in particular? and why is volumetric technique more accurate at low pressures?}
\textbf{QUANTACHROME INSTRUMENTS}
It is now well established (see the latest IUPAC recommendations in Pure. Appl. Chem. 87 (2015) 1051)) that argon molecules provide distinct advantages over nitrogen molecules for gas sorption analyses, including the following:
Unlike nitrogen, argon has no quadrupole moment. Thus, using argon as adsorbate eliminates specific chemical interactions with polar/ionic surface sites;
As a result, argon physisorption isotherms provide much more reliable fingerprints of the interactions modeled by today’s most advanced techniques (e.g., QSDFT) for pore size characterization; and
Argon sorption analyses at its boiling point (87K) can be significantly faster than conventional N2/77K experiments, because the filling of similar pores can occur much more readily at much higher relative pressures;

\item {\textbf{2016}} Brandani and Lev Sarkisov thermodyncamic models need the absolute adsorbed amount (opposed to net or excess). The authors recommend experimentalists including the non-accessible volume if they continue to report excess adsorbed so that net adsorption can be easily calculated~\cite{brandani2016net}.

\item {\textbf{2017}} Brandani et al. The analysis presented is a further indication that absolute adsorption is the thermodynamic variable to use in describing adsorption~\cite{brandani2017net}. Notes that Talu~\cite{talu2013net} IAST (Ideal Adsorption Solution Theory) equations obtained are
inconsistent.  Absolute adsorption is what should be used as Myers~\cite{myers2014physical} derived correctly.
\end{enumerate}

\section{Introduction}
The isosteric heat of adsorption $q_\text{st}$ represents a cornerstone thermodynamic variable for
designing
experimental gas storage systems~\cite{rudzinski2012adsorption, sturluson2019role, patil2016noria, banerjee2016metal}. Adsorption isotherm measurements have been used to
determine the amount of adsorbate that can viably transition to the adsorbent phase for storage~\cite{matranga1992storage, mulfort2007chemical}. For gas storage systems, $q_\text{st}$
is useful in helping to determine which materials would work as an adsorbent. High $q_\text{st}$ leads to strong
attraction (leading to storage), while a weak $q_\text{st}$ leads to easy extraction of the
adsorbate. Finding the optimal balance yields the maximum deliverable capacity that can be extracted from
the gas storage system~\cite{song2015nbo, bae2010optimal}.

Experimentalists have long used the Clausius-Clapeyron equation in order to extract $q_\text{st}$ from
isotherm data~\cite{pan1998examination, lee2005gas, geier2013selective}. The conventional method employed by researchers is to extract data from an isostere
i.e. constant loading of a P vs. T dataset.  Next, a derivative is taken of the log of the pressure vs the temperature. $q_\text{st}$ can then be defined as~\cite{yang1997adsorption}:
\begin{equation}
    q_\text{st} \equiv R T^2\left( \frac{\delta \ln f}{\delta T}
    \right)_{n}
    \label{eq:qst-common}
\end{equation}
If the gas is ideal, then the fugacity is equal to the adsorption pressure $f = P$. The $q_\text{st}$ can directly be extracted from the slope of the experimental isostere.
The Clausius-Clapeyron equation however assumes that the adsorbate behaves as an ideal gas which means at
high pressures there will be error~\cite{pan1998examination} although for real gasses one could derive
$q_\text{st}$~\cite{askalany2015derivation}.
\jpsays{People also use the Van't Hoff equation or is that an intermediate step to get CC equation?}

Describe $\Phi$ in terms of isosteric heat and what exactly it is in the system how it relates to the enthalpy of adsorption.
\begin{equation}
-q_\text{st} = \Phi + \Delta H
\end{equation}
\subsection{What is the problem}

The differential heat of adsorption $q_\text{st}^d$ derived through the Gibbs Surface Excess (GSE) Model is actually (assuming an ideal gas mixture):
\begin{equation}
    q_\text{st}^d = R T^2\left( \frac{\delta \ln \left( P y_i \right)}{\delta T} \right)_{n_i^m} - RT
\end{equation}
For a single-component adsorbate $y_i = 1$. Thus the GSE model has an additional term $RT$ versus the conventional model
in literature~\cite{sircar1999heat,sircar1999isosteric}.

\subsection{How does this paper solve the problem}
Because of the apparent misunderstanding in literature about $q_\text{st}$ and its relation to the
experimentally measured heat of adsorption, we show via derivation and insightful cartoon how to relate
the heat of adsorption $q_\text{st}$ to the potential field of the adsorbent $\Phi$.

\section{David derives $q_\text{st}$ using Gibbs-Duhem}
We can start with the Gibbs-Duhem equation
\begin{align}
  Nd\mu &= -SdT + Vdp
\end{align}
\davidsays{Note that this is only valid for the entire volume (absolute absorption)
based on how Euler's Homogeneous Function Theorem is derived.}
which is true for both the pure gas and the adsorbed gas, and tells us
how the chemical potential changes when you change the temperature and
pressure.

So let's ask what the difference is between equal amounts of gas and
adsorbate.  If we assume that the MOF is incompressible, then this
means the density in the MOF is constant, thus isosteric.
\begin{align}
  Nd\mu_\text{gas} - Nd\mu_\text{ads} &= -S_\text{gas}dT +
  V_\text{gas}dp
  - \left(-S_\text{ads}dT + V_\text{ads}dp\right)
  \\
  0 &= -(s_\text{gas} - s_\text{ads})dT + (v_\text{gas}-v_\text{ads})dp
\end{align}
where we used the fact that the gas and adsorbate are taken to be in
equilibrium.  This tells us that
\begin{align}
  \left(\frac{\partial p}{\partial T}\right)_{n_\text{ads}}
  &= \frac{s_\text{gas} - s_\text{ads}}{v_\text{gas}-v_\text{ads}}
\end{align}
Once again we can solve for $\Delta s$ and thus $q_\text{st}$.  This
heat, with think, is the heat that happens (leaves?) if you start with
an empty mof and a full gas, and then gradually decrease the volume of
the gas and fill up the MOF so you end up with all the atoms in the
mof.
\begin{align}
  q_\text{st} &= T\Delta s \label{eq:david-qst}
  \\
  &= T(v_\text{gas}-v_\text{ads})
  \left(\frac{\partial p}{\partial T}\right)_{n_\text{ads}}
\end{align}
and this lets us know the isosteric heat.  Again, we could assume that
$v_\text{gas}\gg v_\text{ads}$ and that the gas is ideal.  This would
tell us that
\begin{align}
  q_\text{st} &\approx
  Tv_\text{gas}\left(\frac{\partial p}{\partial
    T}\right)_{n_\text{ads}}
  \\
  &\approx T\frac{RT}{p}\left(\frac{\partial p}{\partial
    T}\right)_{n_\text{ads}}
\end{align}

\subsection{What $\Delta s$?}
The $\Delta s$ in Eq.~\ref{eq:david-qst} is the difference in entropy between the adsorbate in the MOF and the pure gas that coexists with it.  Since these two systems have the same molar Gibbs free energy (and chemical potential), we can see that
\begin{align}
    q_{st} &\equiv T(s_\text{gas} - s_\text{ads}) \\
    g_\text{gas} &= g_\text{ads} \\
    h_\text{gas} - Ts_\text{gas} &= h_\text{ads} - Ts_\text{ads} \\
    h_\text{gas} - h_\text{ads} &= T(s_\text{gas} - s_\text{ads})
    \\
    &= q_{st}
\end{align}
Thus the isosteric heat is an enthalpy difference between these two coexisting states.  It is not (yet seen to be?) the differential change of enthalpy when a single atom moves from one state to the other, as some claim.

\davidsays{If this is true, we need to show it is true, otherwise the opposite.}
\jpsays{to self perhaps see what is done in here~\cite{pan1998examination}}
\section{Measuring $q_{st}$}
\corysays{
Given that we are writing a paper to bring clarity Qst, I think it would strengthen the paper if we provided two experimental protocols to accurately measure it:
(1) temperature dependence of adsorption isotherms [check mark]
(2) calorimetry [getting there]
}
\subsection{Clausius-Clapeyron}
\davidsays{This is where you measure a couple of isotherms and take the right derivative.}
\subsection{Calorimetry} In the work by Sircar et al.~\cite{sircar1999isosteric}, he gives a description of experimental protocol for calorimetry. They measure the helium voids of the sample $v^0$ and dosage $v^d$ sides after placing in a calorimeter cell.  Thermopiles measure the heat as a voltage vs time. For adsorption of a single-component \emph{pure, ideal gas}:
\begin{align}
    q_i^0 = \frac{\frac{\Delta Q_i}{\Delta N_i} - \bar h_i - v^0\frac{\Delta P_i}{\Delta N_i}}{1-\frac{v^0}{RT}\frac{\Delta P_i}{\Delta N_i}}
\end{align}
The change in pressure in the sample side during the expt is $\Delta P_i$ and $q_i^0$ is the isosteric
heat of adsorption of pure gas $i$ at $P^0$ and $T^0$.
The isosteric heat of adsorption $q_\text{st}$ is calculated by~\cite{dunne1996calorimetric, parrillo1993characterization}:
\begin{align}
    q_\text{st} = q_d + RT
\end{align}
where $\Delta \bar u \equiv q_d$ is the differential energy of desorption.  If the gas enters the calorimeter at the temperature of the sample cell and if the same number of
moles of gas enters the sample cell and a reference cell wired in reverse polarity, the calorimeter
measures $q_\text{st}$ directly.

\davidsays{I think this just comes from the definition of enthalpy.  $H=U+PV$  For an ideal gas, this means that $H=U+NkT$ by the ideal gas law.  Thus
\begin{align}
    Q_{st} &= \Delta H \\
    &= \Delta U + \Delta(PV) \\
    &\approx \Delta U + \Delta(NkT) \\
   q_{st} &= \frac{Q_{st}}{N} \\
     &\approx \Delta u + k_BT
\end{align}
}
\davidsays{
Remaining and related question:  when we define $\Delta u$ etc what are the two states being subtracted?
}
\section{Friday 4/26 Discussion on $q_\text{st}$}
In the literature, the isosteric heat of adsorption is defined to be the following:
\begin{align}
    q_\text{st} = \left(\frac{\partial H_\text{Tot}}{\partial N_\text{Mof}}\right)_T
\end{align}
The total enthalpy can be written in terms of the number of molecules of adsorbate in the corresponding phase.
\begin{align}
    H_\text{Tot} = N_\text{Mof} h_\text{Mof}\left(p,T\right) + N_\text{Gas} h_\text{Gas}\left(p,T\right)
\end{align}
We derive the isosteric heat of adsorption using a thermodynamic approach.
\begin{align}
    q_\text{st} = \frac{H_\text{all Mof} - H_\text{all Gas}}{N}
\end{align}
This yields us that the single particle difference $q_\text{st}$.
\begin{align}
    q_\text{st} = \frac{\left(N_\text{Mof}+1\right) h_\text{Mof}\left(p,T\right) - \left(N_\text{Gas}-1\right) h_\text{Gas}\left(p,T\right)}{1}
\end{align}

\section{Discussion}
\jpsays{Break this up into subsections in order to determine the ordering of thoughts.  (1) Discuss
problem in terms of a potential $\Phi$ uniform or not i.e. we need a setup. (2) Show a cartoon or sketch
that shows how to relate $\Phi$ to $q_\text{st}$.
}

\subsection{Relate $\Phi$ to $q_\text{st}$}
\jpsays{took this from the other paper}
We consider the density of an adsorbate given a spatially uniform ``external'' potential $\Phi$ created by
interaction with an adsorbent.  This potential will have the effect of shifting the chemical potential of
the material, the same way that the gravitational potential energy causes the density of the atmosphere to
vary with altitude.  The density at a chemical potential $\mu$ is thus given by
\begin{align}
    \rho(\mu) &= \rho_0(\mu + \V)\label{eq:mof-density}
\end{align}
where $\rho_0(\mu)$ is the density of the pure gas at chemical potential $\mu$.

Formally, we recover eqn.~\ref{eq:mof-density} by imposing that the energy of a system of $N$ particles at positions $\mathbf{x}_1,...\mathbf{x}_N \in \Omega$ in a space $\Omega$ with volume $V=|\Omega|$ endowed with uniform external potential $\Phi$ is:
\begin{equation}
    E(\mathbf{x}_1,...\mathbf{x}_N) = N\Phi + E_{gg}(\mathbf{x}_1,...\mathbf{x}_N),
\end{equation} where $E_{gg}$ is the (unknown and complicated) interatomic potential for gas-gas interactions that governs the (real) gas properties. Then, the grand-canonical partition function of this control volume in equilibrium with a bulk gas exhibiting chemical potential $\mu$ is:
\begin{multline}
    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{\Lambda^{-3N}}{N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\xvec_1, ..., \xvec_N)} e^{\beta (\mu - \V) N} d\xvec_1 \cdots d\xvec_N \\ =
    \Xi_0(\mu-\V, V, T),
    \label{eq:xi_realgas}
\end{multline} where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the real gas. Therefore, the thermodynamic properties of the gas sitting in this external potential $\Phi$, in equilibrium with a bath of gas of chemical potential $\mu$, are equivalent to the properties of the gas at chemical potential $\mu-\V$.

\section{ORDERING THOUGHTS}
{\color{red}
\begin{enumerate}
    \item Intro: An explanation for where $-q_\text{st}$ comes from. I presume from fitting adsorption gas isotherms? This would help to motivate the importance of understanding the subtleties between isosteric heat of adsorption and energy of adsorption.
    \item Intro: Discussion of how $-q_\text{st}$ is often conflated in the literature with the energy of adsorption
\end{enumerate}
}

The differential energy of adsorption is defined as the following equation~\cite{myers2002adsorption}:
\begin{equation}
    \Delta \bar u_i = \left( \frac{\delta U}{\delta n_i} \right)_{T,n_j} - u_i^o
\end{equation}
The reference state $u_i^o$ is a gas with infinite separation among its constituent molecules. The notation $e$ denotes excess. $\Delta \bar u_i$ could
also be written in terms of the fugacity.
\begin{equation}
    \Delta \bar u_i = -R T^2\left( \frac{\delta \ln f_i}{\delta T} \right)_{n_i,n_j} + RT
\end{equation}
In this formulation, it is easier to see that the differential enthalpy of adsorption is the first term of the energy of adsorption if a suitable conversion between absolute $n_i$ to excess $n_i^e$ exists.
\begin{equation}
    \Delta \bar h_i^e = -R T^2\left( \frac{\delta \ln f_i}{\delta T} \right)_{n_i^e,n_j^e}
\end{equation}
The isosteric heat of adsorption is denoted as $\Delta \bar h_i^e$~\cite{myers2002adsorption}.
\jpsays{Discuss the conversion between absolute and excess variables 6/13 in paper by Myers and Monson.  Read below and synthesize...}

\subsection{Conversion of absolute to excess variables}
\jpsays{translate this into something understandable...}
4. Conversion of Absolute to Excess VariablesConversion of absolute (ni) to excess (nie) adsorption byeq 28 requires the volume of the gas phase (Vg). If thetheoretical  model  ignores  adsorption  on  the  externalsurface of the adsorbent particles, thenVgis the porevolume.  If  the  model  accounts  for  adsorption  on  theexternal surface, thenVgis the void volume of the system,which includes the pore volume and the space external tothe porous material. This conversion of absolute to excessadsorption is, in principle, precise. However, it does requirea definition of pore volume, which as we discuss belowdepends on the size of the molecule probing the porousmaterial.The reverse process of calculating absolute adsorption by  experimental  methods  is  impossible  in  principle.Absolute  adsorption  depends  on  the  model  selected  to divide the helium void space into two parts:  the spacewithin the potential field of the solid and the space outside the potential field of the solid. For example, a reasonablemodel of a microporous adsorbent might define the space lying within the potential field of the solid as its porevolume, so that absolute adsorption is the total amountof gas contained in the pores. The experimental heliumvoid volume is the sum of the pore volume plus the externalvolume of the gas phase. There is no way to determine experimentally the space lying within the potential fieldof the solid.

\subsection{Gibbsian Surface Excess Model}
Sircar et al. shows that the GSE (Gibbsian surface excess model)~\cite{sircar1999heat} model for an ideal gas mixture (for a pure gas $y_i = 1$) can be used to write the isosteric heat of adsorption~\cite{sircar1999isosteric} related to the total internal energy $U^o$:
\begin{equation}
    \left( \frac{\delta U^o}{\delta n_i^m}\right)_{T,n_{j\neq i}^m} = -R T^2\left( \frac{\delta \ln \bar p_i}{\delta T} \right)_{n_i^m} + RT
\end{equation}
where $\bar p_i$ is the partial pressure.
\subsection{Conventional Model}
The conventional model~\cite{sircar1999isosteric} assumes that a differential amount of pure
adsorbate $dn$ is transferred from gas to the adsorbed phase denoted $a$ for a closed system with
P and T held constant.  We can write (where $U^o=U+U^a$):
\begin{equation}
    dU^o = -dQ - Pdv^o
\end{equation}
The total volume in the gas phase is $V=v^o-V^a$ giving the pressure and volume in the gas phase.
\begin{align}
    V=\frac{nRT}{P}; \quad dV = -\frac{RT}{P}dn^a \qquad \textrm{constant} \quad P,T \\
    Pdv^o=PdV^a - RTdn^a \qquad \textrm{constant} \quad P,T
\end{align}
These equations can be combined to produce an equation for the heat of adsorption.
\begin{equation}
    \left( \frac{\delta Q}{\delta n^a}\right)_{P,T} = h^* - \left( \frac{\delta H^a}{\delta n^a}\right)_{P,T}
\end{equation}
where $h^*$ is the molar enthalpy of a pure gas at temperature T and pressure P of 1 atm. Much of
published literature simplifies (the above) by making a number of assumptions (which don't seem likely to
be valid very often...). $\bar q$ is defined to be the isosteric heat of adsorption under these assumptions.
\begin{equation}
    \left( \frac{\delta Q}{\delta n^a}\right)_{P,T} = \bar q = R T^2\left( \frac{\delta \ln P}{\delta T} \right)_{n^a}
\end{equation}
The following assumptions:
\begin{enumerate}
    \item the adsorbed phase should be liquid like~\cite{young1962physical}.
    \item the adsorbent should be energetically homogeneous or at the least a weakly heterogeneous adsorbent.
\end{enumerate}
In contrast, the GSE model requires no such assumptions.

\subsection{Figures and Ideas}
Figures and Ideas are below
% \begin{figure}
%   \begin{flushleft}
%     a)\\\vspace{-3em}
%     \includegraphics[width=\columnwidth]{methane-298-rho-mu.pdf}\\
%     b)\\\vspace{-3em}
%     \includegraphics[width=\columnwidth]{methane-298-rho-pressure.pdf}
%   \end{flushleft}
%   \vspace{-1em}
%   \caption{The density as a function of (a) chemical potential and (b) pressure for pure methane.}
%   \label{fig:density-vs-mu-ch4}
% \end{figure}

% \begin{figure}
%   \begin{flushleft}
%     \includegraphics[width=\columnwidth]{methane-298-experimental-gibbs}
%   \end{flushleft}
%   \vspace{-1em}
%   \caption{The Gibbs free energy difference.}
%   \label{fig:gibbs}
% \end{figure}

\jpsays{Ideas to think about: Gibbs adsorption equation (GAE), Gibbs dividing surface, differential heat of adsorption, a diagram
showing how $\Phi$ and $\Delta H$ relate to $-q_\text{st}$}

% cited papers!
%https://pubs.acs.org/doi/full/10.1021/jp9903817
%https://pubs.acs.org/doi/10.1021/la026399h

% uncited yet...
%Gibbs surface excess:
%https://www.sciencedirect.com/science/article/pii/S0001868614000049

\jpsays{We should define our $\Phi$ in terms of $-q_\text{st}$ (so that we could write our plots in terms of $q_\text{st}$ rather than $\Phi$) such that we would have:}
\begin{equation}
-q_\text{st} = \Phi + \Delta H
\end{equation}
In some papers I find that the Van't Hoff equation is used to fit for the isosteric heat of adsorption:
\begin{equation}
\left( \frac{\partial \ln K}{\partial \frac{1}{T}} \right)_\theta = -\frac{\Delta H}{R}
\end{equation}
Since the adsorption constants are at equilibrium they obey the equation and $K$ must be isosteric.

In this work, the authors claim to fit CO2 isotherms using this equation:~\cite{sim2014gas}
\begin{equation}
\left( \frac{\partial \ln P}{\partial \frac{1}{T}} \right) = -\frac{\Delta q_\text{st}}{R}
\end{equation}

\end{document}

### 2020-05-01 gas-adsorption/old_storage/main_Nature_Format.tex
added gas storage to thesis pommerenck.

%% Template for a preprint Letter or Article for submission
%% to the journal Nature.
%% Written by Peter Czoschke, 26 February 2004
%%

\documentclass{nature}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{amsmath}

\usepackage{nicefrac} % Include for inline fractions
\usepackage{url}

%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them

\bibliographystyle{naturemag}

% \usepackage{cite}
\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
%
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2} }
\usepackage{mdframed}

\newcommand{\addcite}{{\bf \color{red} []}}
\newcommand{\xvec}{\mathbf{x}}
\newcommand{\thetavec}{\boldsymbol \theta}

\newcommand\V{\Phi}
\newcommand\Vk{\tilde\Phi(\vec k)}
\newcommand\volume{V}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }

% The title should be less than 90 characters! Ours is currently 84.
\title{An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}

%% Notice placement of commas and superscripts and use of &
%% in the author list
\author{Jordan K. Pommerenck$^1$, Cory M. Simon$^2$ \& David Roundy$^1$}

\begin{document}

\maketitle

\begin{affiliations}
 \item Department of Physics, Oregon State University, Corvallis, OR 97331
 \item School of Chemical, Biological, and Environmental Engineering,
       Corvallis, OR 97331.
\end{affiliations}

% For Nature, the abstract is really an introductory paragraph set
% in bold type.  This paragraph must be ``fully referenced'' and
% less than 180 words for Letters.  This is the thing that is
% supposed to be aimed at people from other disciplines and is
% arguably the most important part to getting your paper past the
% editors.  End this paragraph with a sentence like ``Here we
% show...'' or something similar.
\begin{abstract}
  We present a framework that places an intrinsic upper limit on the amount of
  gas that can be stored in a given volume of material and then delivered at a
  lower pressure.
\end{abstract}

\section{Background}

\subsection{Review of previous work}

\section{Introduction}
The transportation sector accounts for 38\% of US energy-related carbon dioxide emissions \cite{useia} and generates toxic air pollution (particulate matter, ozone, NO$_x$, SO$_x$, carbon monoxide, volatile organic compounds) \cite{caiazzo2013air}. Alternative transportation fuels cleaner than petroleum, which currently dominates the transportation fuel market \cite{davis2009transportation}, are therefore critical to mitigate climate change \cite{mcglade2015geographical} and improve air quality for human health. Further motivating the development of technologies to produce sustainable transportation fuels at low cost, the finite global petroleum resources are declining rapidly \cite{sorrell2010global}.

Natural gas, mostly methane, is considered a transition [to a renewable and clean] fuel because, on an energy basis, it emits 25\% less carbon dioxide upon combustion than gasoline \cite{eia2013much} as well as less other toxic byproducts \cite{wang2000full}. Moreover, owing to advances in hydraulic fracturing and horizontal drilling, natural gas is abundant \cite{usnatgassupply} and cheap in the US. A positive environmental outlook for natural gas, however, is predicated on mitigating fugitive emissions (methane is a potent greenhouse gas itself) \cite{alvarez2012greater} and groundwater contamination \cite{osborn2011methane} from hydraulic fracturing.

Hydrogen (H$_2$) is the ultimate transportation fuel because it emits only water when it electrochemically reacts with oxygen in a fuel cell to produce electricity and power a vehicle. Moreover, elemental hydrogen is very abundant, though bonded with oxygen (in water) or carbon (in hydrocarbons); H$_2$ does not occur naturally. Currently, hydrogen (H$_2$) is primarily produced by the steam reforming of natural gas followed by the water-gas shift reaction, which emits carbon dioxide \cite{crabtree2004hydrogen}. Notably, the environmental allure of hydrogen is predicated on its production via a renewable means, e.g.\ splitting water using wind-generated electricity.

Gaseous at ambient conditions, both natural gas ({\color{red} x} MJ/L) and hydrogen ({\color{red} x} MJ/L) possess a very low volumetric energy density compared to (liquid) gasoline (34.2 MJ/L). Consequently, under storage space constraints in passenger vehicles, natural gas and hydrogen must be densified for onboard vehicle storage to drive a reasonable distance on a full tank of fuel. Traditional densification approaches are liquefaction (1 atm, CH$_4$: 111 K, H$_2$: 20 K to achieve 22.2 ML/L \cite{makal2012methane} and 8 MJ/L \cite{suh2011hydrogen}, respectively) or compression to high pressures (room temperature, CH$_4$: up to 200 bar to achieve 9.2 MJ/L \cite{makal2012methane}, H$_2$: up to 700 bar to achieve {\color{red} x} MJ/L). Both suffer several drawbacks. High-pressure storage tanks pose a safety hazard and require heavy, thick-walled, non-conforable vessels. Cryogenic storage vessels are bulky and expensive. Due to imperfect insulation, boil-off losses are a concern for cryogenic storage at ambient pressure, especially for methane since it is a potent greenhouse gas \cite{hasan2009minimizing}. Both compression to high pressures and liquefaction require expensive infrastructure at refilling stations and significant energy input; e.g., liquefaction of hydrogen consumes at least 30\% of its energy content \cite{bossel2003energy}.

A different approach to densify natural gas and hydrogen for vehicular storage is through physical adsorption on nanoporous materials \cite{schoedel2016role}. Consequent to van der Waals interactions between gas molecules and their internal surfaces, porous materials can achieve adsorbed gas densities competitive to compressed gas vessels but at significantly lower pressures. This would thereby reduce the cost of infrastructure at refilling stations, allow thinner-walled and therefore cheaper and lighter pressure vessels, reduce the energy requirements for densification, and alleviate safety concerns with high pressure storage. As most homes in the US are connected to natural gas lines \addcite, adsorbed natural gas fuel tanks could permit at-home fueling. Current research is towards synthesizing a nanoporous material that can maximally densify natural gas and hydrogen whilst satisfying cost and stability constraints. So far, metal-organic frameworks (MOFs) \cite{furukawa2013chemistry} have demonstrated the highest internal surface areas ($>$7000 m$^2$/g \cite{farha2012metal}) and absorptive capacities for hydrogen \cite{suh2011hydrogen,garcia2018benchmark} and methane \cite{makal2012methane,mason2014evaluating}. Importantly, MOFs are highly tunable materials synthesized by combining organic molecules (struts) and metal clusters (nodes) that [often] coordinate to form a three-dimensional, porous framework \cite{furukawa2013chemistry}. This tunability offers the intriguing possibility to synthesize the optimal material for methane or hydrogen storage and delivery \cite{schoedel2016role}.

\unsure[inline]{I'm reading they want to get a H2 density higher than liquid with a porous material. is that possible?!}

Employing nanoporous materials for gas storage and delivery \cite{schoedel2016role}. Pressure swing \cite{sircar2002pressure}.

Targets.
For methane storage, the Advanced \cite{arpaemove}

Generally, knowledge on the thermodynamic limits of a technology is important for allocating research funds to technologies with the most potential for development and setting realistic performance targets.

\begin{figure}
    \centering
    picture of MOF + example methane adsorption isotherm illustrating deliverable capacity or too basic?
    \caption{FIXME}
    \label{fig:example-experiment}
\end{figure}

\section{Gas storage \& delivery by pressure swing adsorption}
% Here, we pose the context and question that we address in this work.
Consider an empty porous material that is packed into a volume $V$ comprising the space within a pressure vessel onboard a vehicle. The pressure swing adsorption process \cite{sircar2002pressure} for vehicular gas storage and delivery (to the engine) is idealized as follows. First, at the filling stage, the pressure vessel is hooked up to a (pure) gaseous reservoir exhibiting a pressure $P_{H}$ at the fuel station. The adsorbed phase is allowed to reach thermodynamic equilibrium with the bulk gas phase. At this point, the adsorbed gas fuel tank is considered full, and the vehicle is ready for a road trip along the Oregon coast. While driving, gas is driven off of the adsorbent and into the engine by a pressure differential. A pressure of $P_{L}$ must be sustained in the fuel tank for sufficient flow to the engine\unsure{or does the natural gas engine need this pressure?}; therefore, the fuel tank is fully used when it exhibits a pressure of $P_{L}$. However, given $P_{L} \neq 0$ (pulling vacuum), so-called cushion or stranded gas will remain in the adsorbent within a fully used fuel tank. Therefore, the amount of gas adsorbed on the porous material at $P_H$ does not determine the driving range of the vehicle. Instead, the \emph{deliverable capacity}, defined as the amount of gas adsorbed in the volume at pressure $P_H$ minus that retained in the adsorbent at pressure $P_L$, determines the driving range. The deliverable capacity of an adsorbent is predicted on the assumption of (i) isothermal operation and (ii) maintenance of thermal equilibrium between the adsorbed and gas phase within the fuel tank.

Mathematically, let $n(P; T)$ be the amount of gas adsorbed per volume of adsorbent at thermodynamic equilibrium when immersed in a bath of gas at pressure $P$ and temperature $T$, the \emph{adsorption isotherm}. The [intrinsic, isothermal] deliverable capacity $\rho_D$ of the adsorbent is then defined as:
\begin{equation}
    \rho_D(P_H, P_L; T) = \rho(P_H; T) - \rho(P_L; T).
    \label{eq:delcap}
\end{equation}
Note that the pressures $P_H$ and $P_L$ correspond to chemical potentials $\mu_H$ and $\mu_L$ via an equation of state for the gas.

The assumption of isothermal operation in eqn.~\ref{eq:delcap} neglects the heat that is (i) released to the surroundings upon adsorption during the filling process and (ii) taken from the surroundings upon desorption during the delivery process. If the heat upon ad/desorption is not properly managed, significant losses in the deliverable capacity result as a consequence of (i) less gas adsorbed in a hot adsorbent during filling and (ii) more gas retained in a cold adsorbent during delivery \cite{mota1997dynamics,chang1996behavior}. A pessimistic performance metric for a material that accounts for heat effects is the adiabatic deliverable capacity, i.e. the gas delivered under adiabatic conditions; while independent of tank design, this depends on the heat capacity of the material.

An engineering strategy to increase the usable capacity of a tank beyond the isothermal deliverable capacity is to use waste heat from the engine to increase the temperature of the adsorbent and thereby drive off cushion gas when the tank approaches pressure $P_L$ \cite{gomez2014exploring}. In such a case, the usable capacity of the tank is $n(P_H, T)-n(P_L; T_H)$ where $T_H$ is the temperature to which we heat the adsorbent. This strategy, however, requires a complicated design.

\begin{limitation}\label{limit:isothermal}
This work assumes isothermal pressure-swing adsorption.  It is possible to achieve greater deliverable capacity by raising the temperature at which the gas is extracted.
\end{limitation}

Finally, we remark that eqn.~\ref{eq:delcap} assumes thermodynamic equilibrium, neglecting the kinetics of ad/desorption during filling and delivery.

\section{Review of previous work}
Several authors have sought to determine an upper bound on the deliverable capacity in a pressure swing adsorption process, particularly for methane storage \& delivery.

Matranga et al. \cite{matranga1992storage} and, later, Bhatia and Myers \cite{bhatia2006optimum} initiated research on the thermodynamics of the optimal adsorption site for gas storage and delivery via pressure swing adsorption and determined the maximal utility of an adsorption site. Under a Langmuir adsorption model, they reasoned that there exists an optimal affinity of the gas for the material (Langmuir constant) that maximizes the deliverable capacity $\rho_D$. If the gas adsorbs too weakly, too little gas will adsorb at $P_H$; too strongly, too much gas will be retained at $P_L$; both diminish $\rho_D$. Following is the optimal utility of each adsorption site if it is endowed with the optimal free energy of adsorption. Assuming a typical entropy of adsorption, Bhatia and Myers \cite{bhatia2006optimum} set widely used rules of thumb for the optimal enthalpy of adsorption of hydrogen and methane to maximize $\rho_D$. Simon et al. \cite{simon2014optimizing}, through molecular simulations of methane adsorption in zeolites, later showed that there exists significant variation in the entropy of adsorption and therefore an optimal heat of adsorption is a flawed target to pursue. Following from these studies is a theoretical upper bound of the fraction of adsorption sites that can deliver a molecule in a pressure swing adsorption process, but these studies neglected (i) guest-guest interactions which can enhance the utility of an adsorption site \cite{simon2014optimizing} and (ii) the question of how many adsorption sites one can pack into a nanoporous material that harbor the optimal free energy of adsorption.

Gomez-Gualdron et al.\ \cite{gomez2017impact} obtained an upper bound on the methane deliverable capacity via a few different methods. First, they packed adsorption sites into an FCC lattice (the most efficient packing of spheres into a volume), endowed each adsorption site with the optimal energy of adsorption, and accounted for methane-methane interactions via a Lennard-Jones potential. The ARPA-E target of 315 v STP/v was reached when adsorption sites were endowed with an optimal heat of adsorption of 11.5 kJ/mol and adsorption sites were $5.3$ \AA~apart. Second, similar to here, they endowed a volume with a background energy field, then simulated the adsorption of methane in the volume using a Lennard-Jones potential for methane-methane interactions. The maximal deliverable capacity of 357 v STP/v was achieved with a background heat of adsorption of 8.6 kJ/mol. Both of these methods did not explicitly consider the space taken up by the material atoms that create the background energy field. To address this, Gomez-Gualdron also constructed pseudo materials by placing Lennard-Jones spheres in space with tunable well depths. When constrained to form a connected network, none of the psuedo-materials reached the ARPA-E target.

Another strategy to assess the feasibility of deliverable capacity targets is by conducting molecular simulations of adsorption in tens of thousands of existing and (realistic) hypothetical materials and observing the tail of the distribution of simulated deliverable capacities \cite{gomez2014exploring,simon2015materials}. For example, Simon et al.\ \cite{simon2015materials} simulated methane adsorption in hundreds of thousands of existing and hypothetical nanoporous crystal structures and found the highest simulated methane deliverable capacity to be 196 v STP/v, suggesting the ARPA-E target could be impossible to reach. Such a conclusion rest upon the accuracy of the molecular model describing molecular interactions as well as the sufficient sampling of achievable ``chemical space'' by the sets of materials considered. However, Gomez-Gualdron et al.\ \cite{gomez2014exploring} scaled the Lennard-Jones potential well depths in the simulations and still found no materials meeting the ARPA-E deliverable capacity target.

Finally, Lee et al.\ \cite{lee2019predicting} trained a generative adversarial network to generate fictitious but realistic potential energy surfaces hosted by zeolites for methane adsorption, then conducted molecular simulations of methane adsorption in 100 000 generated potential energy surfaces to assess performance limitations. No generated potential energy surface met the ARPA-E target. Conclusions under this framework are also predicated on the accuracy of the molecular model as well as sufficient sampling of the space of possible potential energy surfaces fed to the generative adversarial network during training.

Brown and Freeman \cite{bhown2011analysis}

\section{An upper bound on gas adsorption in control volume}
We consider a host (which could be a MOF, another porous solid, or even a liquid solvent) interacting with a gas.  Although the process of adding guest to host could be anything from adsorption to solvation, we will refer to it as adsorption throughout this paper, since that is the common case.  The goal of this process is to load the guest into the host at some high pressure, and then extract the guest until it reaches some lower pressure.

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{methane-298-rho-mu}
  \caption{The density as a function of chemical potential for bulk
    methane.  The density is written as a ratio of the volume of
    methane at STP to the volume of the gas. The pressure axis is labelled
    along the top.  We plot with dashed lines the density of adsorbed
    methane as a function of chemical potential for several
    adsorbents~\cite{long} \davidsays{we should also mention the MOFs
      here}}
  \label{fig:density-vs-mu-ch4}
\end{figure}

Each guest is characterized by a set of equations of state for the pure guest system.  We will focus on the number density as a function of chemical potential $\mu$ (or equivalently molar Gibbs free energy) and temperature $\rho(\mu,T)$.  We illustrate in Fig.~\ref{fig:density-vs-mu-ch4} this equation of state for methane. We also show a somewhat more familiar equation of state illustrating the density as a function of pressure $\rho(p,T)$.  For comparison, we also show the number density as a function of chemical potential for methane in a variety of hosts.

% \davidsays{This paragraph could move considerably later. Or cut entirely?}
% We model the interaction of host with guest as a simple Gibbs free energy of adsorption per mole (or atom) $\Delta G_{st}$.  We will assume that this Gibbs free energy is independent of the concentration of guest.  At low concentrations, this assumption is reasonable, because the guests can be assumed to be widely separated and their interactions weak.  In the limit of high concentrations, we expect that the attraction of the guest to the host will be diminished, since the ``good spots'' will already be occupied.  The choice of a concentration-independent $\Delta G_{st}$ will thus have a tendency to overestimate the deliverable capacity, providing an upper bound (for more on this, see Sec.~\ref{sec:qualitative-max}).

To place an upper bound on the deliverable capacity $\rho_D$ in Eq.~\ref{eq:delcap}, we consider a host whose sole interaction with the guest is to introduce an attractive potential energy $\V$ for gas molecules in the control volume.  Because this interaction is uniform within the host, the interactions between gas molecules are identical to that of a pure gas, which allows us to predict the properties of this model host using the experimentally known properties of the pure gas~\cite{nist}.

%\subsection{Finding the density}\label{sec:finding-n}

The spatially uniform ``external'' potential $\V$ representing guest-host interactions has the effect of shifting the chemical potential of the material, in the same way that the gravitational potential energy causes the density of the atmosphere to vary with altitude.  The density at a chemical potential $\mu$ is thus given by
\begin{align}
    \rho(\mu) &= \rho_0(\mu + \V)\label{eq:mof-density}
\end{align}
where $\rho_0(\mu)$ is the density of the pure gas at chemical potential $\mu$ \davidsays{we could maybe find a better name than $\rho_0$?}.  Figure~\ref{fig:density-vs-mu-ch4}a shows $\rho_0$ for methane using data available from NIST~\cite{nist}.

% Formally, we recover Eq.~\ref{eq:mof-density} by imposing that the energy of a system of $N$ particles at positions $\mathbf{x}_1,...\mathbf{x}_N \in \Omega$ in a space $\Omega$ with volume $V=|\Omega|$ endowed with uniform external potential $\V$ is:
% \begin{equation}
%     E(\mathbf{x}_1,...\mathbf{x}_N) = N\V + E_{gg}(\mathbf{x}_1,...\mathbf{x}_N),
% \end{equation} where $E_{gg}$ is the (unknown and complicated) interatomic potential for guest-guest interactions that governs the (real) gas properties. Then, the grand-canonical partition function of this control volume in equilibrium with a bulk gas exhibiting chemical potential $\mu$ is:
% \begin{multline}
%     \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{\Lambda^{-3N}}{N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\xvec_1, ..., \xvec_N)} e^{\beta (\mu - \V) N} d\xvec_1 \cdots d\xvec_N \\ =
%     \Xi_0(\mu-\V, V, T),
% \end{multline} where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the real gas. Therefore, the thermodynamic properties of the gas sitting in this external potential $\V$, in equilibrium with a bath of gas of chemical potential $\mu$, are equivalent to the properties of the gas at chemical potential $\mu-\V$.

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{methane-298-methane-298-n-vs-G.pdf}
    \caption{The optimal deliverable capacity for methane as a function of the attractive Gibbs free energy $\Delta G_{st}$.  Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\Delta G_{st}$ at the empty and full pressures shown as dots connected by a line.}
    \label{fig:methane-298-D}
\end{figure}

The deliverable capacity is given by
\begin{align}
    \rho_D &= \rho(\mu_H) - \rho(\mu_L)
\end{align}
where $\mu_H$ and $\mu_L$ are the chemical potentials when the tank is filled and when it is considered empty, respectively.  For a uniform potential $\V$ we can use Eq.~\ref{eq:mof-density} to relate this to the properties of the bulk fluid
\begin{equation}
    \rho_D(\V) = \rho_0(\mu_H+\V) - \rho_0(\mu_L+\V).
    \label{eq:DofPhi}
\end{equation}
Thus the deliverable capacity $\rho_D(\V)$ is the difference between the curve shown in Fig.~\ref{fig:density-vs-mu-ch4}a shifted by the upper and lower chemical potentials and offset by $\V$.  Figure~\ref{fig:methane-298-D} shows the two shifted density curves as well as their difference.  There exists an optimal adsorption free energy, which balances the need to maximize the full-tank density against the need to minimize the empty-tank density.
%
%As a ballpark, we can expect the attractive potential to be optimal around the maximum slope of $\rho_0(\mu)$.  This is the point where the steric repulsion between the gas atoms starts leading to saturation of the density.

An essential question is whether this approach of solving for the potential shift $\V$ that yields the maximum deliverable capacity.  In the supplementary material we show that this approach yields an \emph{extremum} of the deliverable capacity, provided the guest does not crystallize in the density range and temperature of interest.  We will argue in the remainder of this section that this extremum is actually a maximum.  To show this, we must further address two possibilities:  the extremum may turn out to actually be a saddle point, or it could turn out to be a local rather than global maximum.

% \subsection{Why is the uniform potential optimal?}\label{sec:qualitative-max}

To qualitatively argue that the homogeneous potential provides a \emph{maximum} deliverable capacity, let us now consider the case of a more realistic material which causes the guest molecules to feel a non-uniform potential energy.  This means that there are some lower-energy sites (and orientations) and some higher-energy sites (and orientations).  We wish to show that this material will have a lower deliverable capacity.  At low densities of guests, the low-energy sites will be preferentially occupied, while at higher densities the some of the guest molecules will be forced by steric repulsion into higher energy locations.  Thus the mean attraction will be minimum when the tank is ``full,'' and maximum when the tank is ``empty.''  This is the opposite of what we would want to acheive a high deliverable capacity.  Based on this qualitative reasoning, we expect a homogeneous potential to give optimal deliverable capacity, since it will provide the most attraction for the molecules added at high densities.

% \davidsays{Is this paragraph out of place, or duplicated?
% %
% Before going on to show that a uniform potential gives an optimal deliverable capacity, we will first point out a limitation of this maximum capacity.  In general, one can increase the deliverable capacity by extracting the gas at a higher temperature than the temperature at which the tank is filled.  A higher temperature will always drive off more of the gas from the MOF, since the gas state has maximal entropy.  When the temperature of extraction differs in this way, a uniform potential does \emph{not} give an optimum deliverable capacity, because it maximizes the entropy of the adsorbed gas, which thus tends to reduce the effect of higher temperatures driving off more gas.}

\subsection*{Limitations}
Here we will articulate the assumptions required for our upper bound.  Some of these assumptions may serve as a guide for those who wish to violate our upper bound, while others are more solid.

\begin{limitation}\label{limit:crystallization}
Our proof assumes the gas is stable with regard to crystallization.
\end{limitation}

Our proof that the homogeneous potential is an extremum makes a very important assumption that the fluid is stable at the temperature and densities involved.  Our upper bound will \emph{not} be valid for guests which may crystallize in their pure form.  This is because a periodic potential can interact with a high-density crystal far more strongly than it interacts with a low-density gas.  We are not aware of a practical application of this principle, but theoretically it is possible.  For the application of gas storage for light gasses such as methane or hydrogen at moderately high temperatures (say liquid nitrogen or warmer), the pure guest is sufficiently far from its crystal form that we need not concern ourselves with crystallization.  We do not see this particular assumption as one that can be worked around.

There remains the possibility that the homogeneous potential is a local maximum, but there is another global maximum.  We do not see how to formally rule out this possibility for an arbitrary gas-gas interaction, but the qualitative arguments above seem sufficient to rule out this possibility for a real gas.

\begin{limitation}\label{limit:field}
Our proof assumes the interaction of the gas with the substrate does not modify the substrate.  This is equivalent to the assumption that the interaction of substrate with gas operates as a simple potential energy field.
\end{limitation}

However, there was a more fundamental assumption, which is that the porous material acts as an effective potential for the guest molecules, and that the guest-guest interaction is unmodified.  For \emph{most} porous materials this seems likely to be a pretty good approximation.  However, there are unusual cases where the host can in effect provide a very strongly density-dependent interaction.  This is perhaps easiest to see in the example of \davidsays{look up the collapsing MOFs description from the RMS-MOF paper}, in which the guest can introduce a phase change in the solid, which has the effect of drastically lowering the density of guest in the material.  Such materials can indeed violate our upper bound. \davidsays{Remember to discuss this in the conclusions or perhaps discussion, we can frame these limitations as guides.  The limitations of our upper bound are the design requirements for a material that will do any better.}

\section{Results}
We explore the theoretical deliverable capacity for two gases stored in a homogeneous potential. We consider methane and hydrogen gas due to the significant environmental and transportation implications. Experimental data is obtained from NIST and naturally includes quantum effects. We are able to place a theoretical upper bound for \emph{any} adsorbent material with a uniform potential. In the context of light vehicle storage, we compare the experimental data obtained from the literature for porous material MOFs and compare with important goals and research targets set by DOE.

\subsection{Methane}
Fig.~\ref{fig:methane-298-D} shows the theoretical upper bound for methane storage. In addition to the predicted maximum deliverable capacity as a function of $\Delta G_{st}$, the ARPA-E target of 315~cm$^3$~STP/cm$^3$~\cite{arpaemove} is shown for context.  For any adsorbent with a 100\% void fraction, the target appears to be theoretically possible. The experimental deliverable capacity for several adsorbents~\cite{mason2014evaluating} are also shown over a range of $\Delta G_{st}$ (converted from $q_\text{st}$ as explained above). These can be compared with the highest observed deliverable capacity for methane at room temperature of 206~cm$^3$~STP/cm$^3$~\cite{gomez2014exploring}. In order, to
reach the ARPA-E target, an adsorbent material would have to be identified with a higher $\Delta G_{st}$.

\subsection{Hydrogen}

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{hydrogen-298-hydrogen-298-n-vs-G}
    \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $\Delta G_{st}$.  Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\Delta G_{st}$ at the empty and full pressures shown as x's connected by a line.}
    \label{fig:hydrogen-298-D}
\end{figure}

Storage of hydrogen is considerably more challenging, and the DOE ULTIMATE deliverable capacity target~\cite{DOE} is close to thermodynamically impossible at 25$^\circ$C.  Figure~\ref{fig:hydrogen-298-D} shows the theoretical maximum curve for hydrogen storage along with a range of experimental measurements for known MOFs.  The DOE ULTIMATE deliverable capacity target is theoretically possible, however by such a small margin that we can safely rule out the possibility of reaching this target through storage and removal of hydrogen in \emph{any} adsorbent material at room temperature and with this pressure range.  We also note that actual MOFs have a $\Delta G_{st}$ that is far below that which is required to reach this target.  This reflects the known fact that hydrogen interacts much more weakly with substrates, and has a much lower isosteric heat.

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{hydrogen-77-hydrogen-77-n-vs-G}
    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\Delta G_{st}$.  Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\Delta G_{st}$ at the empty and full pressures shown as x's connected by a line.}
    \label{fig:hydrogen-77-D}
\end{figure}

One approach to increase the deliverable capacity is to reduce the temperature.  This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows the upper bound to the deliverable capacity for hydrogen at liquid nitrogen temperatures.  The DOE ULTIMATE target in this case looks far more achievable, and with a much lower $\Delta G_{st}$.  In fact, an empty tank at this temperature can satisfy the DOE 2020 target.  Actual MOFs fall far short of the theoretical maximum.

\section{Conclusion}
As many have suspected \davidsays{cite previous work}, pressure swing adsorption in rigid porous solid is not going to enable reaching targets.  \davidsays{conclude explicitly about each target}  Unlike previous work, our upper bound on deliverable capacity is based on experimentally known properties of the gas of interest rather than simulations.

The limitations on our proof of an upper bound on the deliverable capacity provide hints as to approaches that could in principle achieve the targets.  In this paper we have identified three limitations of our upper bound on the deliverable capacity.  Two of these three limitations suggest engineering strategies that do hold promise of achieving the targets for hydrogen and methane storage.  Limitation \ref{limit:crystallization}, which states that the gas must not crystallize does not means for improvement, since it is a limit on the gasses to which our bound may be applied, and neither hydrogen nor methane are in any danger of crystallizing.

Limitation~\ref{limit:isothermal}

Limitation~\ref{limit:field}

One such approach is to use temperature swings along with pressure swings.

A more subtle limitation of our proof is that it assumes that the substrate itself is not modified by the presence of the gas.  While this assumption is never precisely true, since any adsorbed atom will change the electron distribution of the substrate to some degree, it tends to be sufficiently accurate that most atomistic simulations can get away with assuming that the substrate is unmodified.  A dramatic exception to that rule is the field of flexible MOFs, in which the presence of the gas can induce a phase change which \emph{drastically} modifies the gas-substrate interaction~\davidsays{cite experiments}.

\clearpage 
% Then the body of the main text appears after the intro paragraph.
% Figure environments can be left in place in the document.
% \verb|\includegraphics| commands are ignored since Nature wants
% the figures sent as separate files and the captions are
% automatically moved to the end of the document (they are printed
% out with the \verb|\end{document}| command. However, tables must
% be manually moved to the end of the document, after the addendum.

% \begin{figure}
% %%%\includegraphics{something} % this command will be ignored
% \caption{Each figure legend should begin with a brief title for
% the whole figure and continue with a short description of each
% panel and the symbols used. For contributions with methods
% sections, legends should not contain any details of methods, or
% exceed 100 words (fewer than 500 words in total for the whole
% paper). In contributions without methods sections, legends should
% be fewer than 300 words (800 words or fewer in total for the whole
% paper).}
% \end{figure}

% \section*{Another Section}

% Sections can only be used in Articles.  Contributions should be
% organized in the sequence: title, text, methods, references,
% Supplementary Information line (if any), acknowledgements,
% interest declaration, corresponding author line, tables, figure
% legends.

% Spelling must be British English (Oxford English Dictionary)

% In addition, a cover letter needs to be written with the
% following:
% \begin{enumerate}
%  \item A 100 word or less summary indicating on scientific grounds
% why the paper should be considered for a wide-ranging journal like
% \textsl{Nature} instead of a more narrowly focussed journal.
%  \item A 100 word or less summary aimed at a non-scientific audience,
% written at the level of a national newspaper.  It may be used for
% \textsl{Nature}'s press release or other general publicity.
%  \item The cover letter should state clearly what is included as the
% submission, including number of figures, supporting manuscripts
% and any Supplementary Information (specifying number of items and
% format).
%  \item The cover letter should also state the number of
% words of text in the paper; the number of figures and parts of
% figures (for example, 4 figures, comprising 16 separate panels in
% total); a rough estimate of the desired final size of figures in
% terms of number of pages; and a full current postal address,
% telephone and fax numbers, and current e-mail address.
% \end{enumerate}

% See \textsl{Nature}'s website
% (\texttt{http://www.nature.com/nature/submit/gta/index.html}) for
% complete submission guidelines.

% \begin{methods}
% Put methods in here.  If you are going to subsection it, use
% \verb|\subsection| commands.  Methods section should be less than
% 800 words and if it is less than 200 words, it can be incorporated
% into the main text.

% \subsection{Method subsection.}

% Here is a description of a specific method used.  Note that the
% subsection heading ends with a full stop (period) and that the
% command is \verb|\subsection{}| not \verb|\subsection*{}|.

% \end{methods}

\bibliographystyle{naturemag}
\bibliography{bibfile}

%% Here is the endmatter stuff: Supplementary Info, etc.
%% Use \item's to separate, default label is "Acknowledgements"

\begin{addendum}
 \item Put acknowledgements here.
 \item[Competing Interests] The authors declare that they have no
competing financial interests.
 \item[Correspondence] Correspondence and requests for materials
should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
\end{addendum}

\end{document}

### 2020-05-01 gas-adsorption/supplemental-proof.tex
added gas storage to thesis pommerenck.

--- 2020-05-01 gas-adsorption/main.tex
@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp]{revtex4-1}
-\usepackage{graphicx} % Include figure files
-\usepackage{color}
+\documentclass[twocolumn]{article}
 \usepackage{amsmath}
-\usepackage{subfig}
-
-\usepackage{nicefrac} % Include for inline fractions
-\usepackage{url}
-
-% \usepackage{cite}
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage{color}  % Coloured text etc.
-%
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
-\usepackage{mdframed}
-
-\newcommand{\addcite}{{\bf \color{red} []}}
-\newcommand{\xvec}{\mathbf{x}}
-\newcommand{\thetavec}{\boldsymbol \theta}
-
-\newcommand\V{\Phi}
-\newcommand\Vk{\tilde\Phi(\vec k)}
-\newcommand\volume{V}
-
-\newcommand{\pink}[1]{{\bf \color{pink} #1}}
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\orange}[1]{{\bf \color{orange} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
-
-%\newcommandx{\davidsays}[1]{\begin{mdframed}[backgroundcolor=blue!24]\textbf{David says:} #1\end{mdframed} }
-%\newcommandx{\davidsays}[2][1=inline]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{David says:} #2} }
-\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
-\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
-\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }
-
-%% \renewcommand\davidsays[1]{}
-%% \renewcommand{\corysays}[2][1=inline]{}
-%TC:macro \corysays [ignore]
-%TC:macro \davidsays [ignore]
-%TC:macro \jordan [ignore]
-
-\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }
-
-\newtheorem{limitation}{Limitation}
 
 \begin{document}
-\title{
-An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}
-% other ideas:
-%Gas delivery via pressure-swing adsorption in nanoporous materials: setting storage targets}
-% A thermodynamic upper bound to gas delivery via pressure-swing adsorption in nanoporous materials:}
 
-\author{Jordan K. Pommerenck}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-\author{Cory M. Simon}
-\affiliation{School of Chemical, Biological, and Environmental Engineering,
-  Corvallis, OR 97331}
-\author{David Roundy}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
+\section*{Proof of extremum}
+\davidsays{I think this section should probably move to supplementary information, if we're going to send to Nature.
+It suddenly gets way more mathematical here, and we start introducing Fourier space, etc.}
 
-\begin{abstract}
-  Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles [].
-One strategy to densify these gases is to pack the fuel tank with a porous material [].
-  Metal-organic frameworks are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases [].
-The US Department of Energy (DOE) has set volumetric deliverable capacity targets [] which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels.
-Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction []; none have established a rigorous upper bound on the deliverable capacity.
-  Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing.
-% "pressure-swing storage doesn't make sense to me.
-storage and delivery  maybe but awk.
-  % I added rigid since that could be important...
-  To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
-  \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
-  Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
-  We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-  % However, this upper bound directly leads to important realizations which should guide future development.
-  However, limitations to the scope of applicability of our upper bound guide future material development efforts.
- % Firstly, one could extract the gas at a higher temperature than that used while filling [].
-  Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
-  Secondly, the fundamental physics of our upper bound do not pertain to
-  % rule out
-  any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
-  % firstly, secondly vs first, second?
-\end{abstract}
-
-\newcommand\pfull{\ensuremath{p_{\text{full}}}}
-\newcommand\pempty{\ensuremath{p_{\text{empty}}}}
-
-\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
-\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}
-
-\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
-\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}
-
-\maketitle
-
-\newcommand\gst{\ensuremath{\Delta g_{st}}}
-
-\paragraph*{to do}
-\begin{enumerate}
-    \item See david/corysays
-    \item make the code presentable
-    \item do the FIXME in code
-\end{enumerate}
-
-\section{Introduction}
-The transportation sector is dominantly powered by petroleum-based fuels \cite{davis2009transportation}.
-In the US, it is responsible for 36\% of energy-related carbon dioxide emissions \cite{useia} and 50,000 premature deaths per year associated with particulate matter and ozone emissions \cite{caiazzo2013air}.
-Natural gas and hydrogen (H$_2$) are alternative transportation fuels that emit fewer/less pollutants 
-than petroleum-based fuels, and thus their widespread adoption could mitigate climate change 
-\cite{mcglade2015geographical} and improve air quality for human health. Further, as the finite global 
-petroleum resources are declining rapidly \cite{sorrell2010global}, the development of technologies for the widespread
-adoption of sustainable transportation fuels, such as hydrogen, is critical.
-
-Natural gas, mostly methane, is considered a transition (to a renewable and clean) fuel because 
-it emits 25\% less carbon dioxide \cite{eia2013much} and less toxic 
-byproducts \cite{wang2000full} upon combustion per unit energy produced compared to gasoline. From an 
-economic standpoint, the supply of natural gas in the US is increasing as a result of hydraulic 
-fracturing and horizontal drilling techniques \cite{usnatgassupply}. A positive environmental outlook 
-for natural gas, however, is predicated on mitigating fugitive emissions (methane is itself a potent 
-greenhouse gas) \cite{alvarez2012greater} and groundwater contamination \cite{osborn2011methane}
-from hydraulic fracturing.
-
-Hydrogen (H$_2$) is the ultimate transportation fuel because it emits only water when it electrochemically reacts with oxygen in a fuel cell to produce electricity and power a vehicle.
-% Moreover, elemental hydrogen is very abundant, though bonded with oxygen (in water) or carbon (in hydrocarbons); H$_2$ does not occur naturally in abundant quantities.
-Currently, hydrogen (H$_2$) is primarily produced by steam reforming of natural gas followed by the
-water-gas shift reaction, which emits carbon dioxide \cite{crabtree2004hydrogen}. Notably, the 
-environmental allure of hydrogen is predicated on its production via a renewable means, e.g.\ splitting
-water using wind-generated electricity.
-
-% \jordan{I got the values from wikipedia but the citation that wikipedia gave didn't link to the correct paper? Article is Energy Density.
-% THIS PAPER says H2 at 700 bar is 4.7 MJ/L: Review on processing of metal–organic framework (MOF) materials towards system integration for hydrogen storage
-% }
-At ambient conditions, both methane
-%(0.036 MJ/L)
-and hydrogen gas
-%(0.01 MJ/L)
-possess a very low volumetric energy density compared to (liquid) gasoline.
-%(34.2 MJ/L)
-Consequently, under storage space constraints in passenger vehicles, natural gas and hydrogen must be densified for onboard storage to achieve a reasonable driving range on a ``full'' tank of fuel.
-Traditional densification approaches are liquefaction (cryogenic storage near atmospheric pressure)
-% (at 1 atm, liquid CH$_4$ at 111~K to achieve 22.2 MJ/L~\cite{makal2012methane}, liquid H$_2$ at 20 K to achieve 8~MJ/L \cite{suh2011hydrogen})
-or compression to high pressures (room temperature storage).
-% (200 bar for CH$_4$ to achieve 9.2 MJ/L \cite{makal2012methane} and 700 bar for H$_2$ to achieve 4.5-5.3 MJ/L~\cite{unknown}).
-Both approaches require expensive infrastructure at refilling stations and significant energy input; e.g., liquefaction of hydrogen consumes at least 30\% of its energy content \cite{bossel2003energy}.
-Moreover, high-pressure storage tanks are heavy, thick-walled, and non-conformable, while cryogenic storage tanks are bulky, expensive, and afflicted by boil-off losses \cite{hasan2009minimizing}.
-% \jordan{
-% Both forms of densification suffer several drawbacks varying from safety issues,  excessively expensive, boil-off~\cite{hasan2009minimizing}, and extensive refilling infrastructure~\cite{bossel2003energy}.
-% }
-% % Jordan: I commented this out! % Cory: I shorted it. I think it's important.
-
-A promising approach to densify natural gas \cite{makal2012methane,mason2014evaluating} and hydrogen \cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is through physical adsorption in nanoporous materials at room temperature~\cite{schoedel2016role}.
-The internal surfaces of porous materials attract gas molecules through van der Waals, electrostatic, etc.
-interactions to achieve a higher adsorbed gas density than the bulk gas at the same temperature and pressure, allowing for room temperature and lower-pressure storage and alleviating many drawbacks of high-pressure storage.
-% This would thereby reduce the cost of infrastructure at refilling stations, allow thinner-walled and therefore cheaper and lighter pressure vessels, reduce the energy requirements for gas densification, and alleviate safety concerns with high pressure storage.
-
-%As most homes in the US are connected to a natural gas pipeline \addcite, adsorbed natural gas fuel tanks could permit at-home fueling.
-
-% Jordan: I commented this out!
-% Cory: I think I agree that this paragraph is not necessary, thanks.
-% Current research is towards synthesizing a nanoporous material that can maximally densify natural gas/hydrogen whilst satisfying cost and stability constraints.
-So far, metal-organic frameworks (MOFs) \cite{furukawa2013chemistry} have demonstrated the highest internal surface areas ($>$7000 m$^2$/g \cite{farha2012metal}) and absorptive capacities for hydrogen \cite{suh2011hydrogen,garcia2018benchmark} and methane \cite{makal2012methane,mason2014evaluating}.
-Importantly, MOFs are highly adjustable materials synthesized by combining organic molecules (struts) and metal clusters (nodes) that
-% % I cut [often] here, since I'm not sure what the exception is and I don't think it's relevant.
-% coordinate to form a three-dimensional, porous framework \cite{furukawa2013chemistry}.
-This modular synthesis enables one to fine-tune the MOF chemistry to arrive at an optimal material for methane or hydrogen storage and delivery \cite{schoedel2016role}.
-See Fig.~\ref{fig:example_MOF} for the crystal structure of a canonical MOF, CuBTC \cite{chui1999chemically}.
-
-For a vehicle employing a porous material to store natural gas or hydrogen, the (volumetric) \emph{deliverable capacity} of the gas is the primary thermodynamic property of the material that determines the driving range~\cite{mason2014evaluating}.
-The adsorbed gas storage tank delivers the gaseous fuel to the engine via an (assumed) isothermal pressure swing \cite{sircar2002pressure}.
-The deliverable capacity (see Fig.~\ref{fig:delcap}) is the density of the gas in the material at the storage pressure $\pfull$ minus the residual gas that remains adsorbed at the lowest pressure $\pempty$ such that sufficient flow is maintained to feed the engine.
-For commercial feasibility, the US Department of Energy (DOE) has set deliverable capacity targets for both adsorbed methane and hydrogen storage and delivery for vehicles.
-For methane, the Advanced Research Projects Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$ per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure swing \cite{simon2015materials}.
-For hydrogen, the DOE set a series of progressive targets at five year intervals, with the ultimate target of 0.05 kg H$_2$/L \cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a minimum of -40 $^\circ$C \cite{allendorf2018assessment}.
-Thus far, despite the emergence of highly tunable materials with enormous surface areas, such as metal-organic frameworks \cite{furukawa2013chemistry}, no porous material has met these deliverable capacity targets.
-
-\begin{figure}
-    \centering
-    \subfloat[][]{
-        \includegraphics[width=0.6\columnwidth]{figs/hkust-1_fancy.png} \label{fig:example_MOF}
-    }
-    \qquad
-    \subfloat[][]{
-        \includegraphics[width=0.98\columnwidth]{figs/usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
-    }
-    \caption{Gas storage and delivery using metal-organic frameworks (MOFs).
-(a) the crystal structure of an archetype MOF, CuBTC \cite{chui1999chemically}.
-(b) the methane adsorption isotherm in CuBTC \cite{chui1999chemically} at 298 K (blue) (data from Ref.~\cite{mason2014evaluating}).
-The deliverable capacity $\rho_D$ is illustrated as the density of gas in the MOF at the storage pressure $\pfull$ minus the density at the discharge pressure $\pempty$.
-    }
-    \label{fig:fig1}
-\end{figure}
-
-To set realistic performance targets and optimally allocate resources for research efforts, in this work, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a rigid porous material and uses as input the experimentally measured properties of the bulk gas.
-Our extremum is provided by a substrate that offers a spatially uniform potential energy field felt by the gas.
-Applying our framework to methane and hydrogen gas, we find the US DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-Optimistically, new paradigms outside the scope of applicability of our theoretical framework, such as gas-induced structural transitions of the material, hold promise for meeting these targets, as evidenced by flexible MOF Co(bdp) which currently boasts the largest methane deliverable capacity~\cite{mason2015methane}.
-
-% Jordan: I commented this out!
-% To optimally allocate resources for research efforts and set realistic performance targets, it is important to assess the thermodynamic limits of technologies.
-Here, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a nanoporous material and uses as input experimental measurements of the bulk gas.
-Applying our framework to methane and hydrogen gas, we find the DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-
-\section{Gas storage \& delivery by isothermal, pressure-swing adsorption}
-% Here, we pose the context and question that we address in this work.
-% Consider pure gas storage and delivery by isothermal, pressure-swing adsorption \cite{sircar2002pressure}.
-Consider a pressure vessel onboard a vehicle (i.e.
-fuel tank) packed with porous material.
-At the filling stage, the tank is connected to a (pure) gaseous reservoir at pressure $\pfull$ and allowed to equilibrate.
-At this point, the adsorbed gas tank is considered full.
-While driving, gas desorbs from the adsorbent to the engine/fuel cell, driven by a pressure differential.
-The tank is considered depleted/empty when the pressure has dropped to $\pempty$, the pressure at which the flow rate of gas from the tank to the engine is insufficient.
-%
-However, given $\pempty \neq 0$ (pulling vacuum), residual gas will remain trapped in the adsorbent.
-Therefore, the driving range of the vehicle is primarily determined by the \emph{deliverable capacity}, defined as the density of gas in the tank at $\pfull$ minus density of the the residual gas remaining at $\pempty$.
-The [isothermal, volumetric] deliverable capacity is an intrinsic property of the nanoporous material and its interaction with the gas.
-%Precisely, let $\rho(p; T)$ be the density of gas in the adsorbent when in equilibrium with a bath of pure gas at pressure $p$ and temperature $T$, the \emph{adsorption isotherm}; $\rho(p,T)$ is essentially the equation of state of the gas adsorbed in the porous material.
-The [intrinsic, isothermal] deliverable capacity $\rho_D$ of the gas in the adsorbent is:
-%\begin{equation}
-%    \rho_D(\pfull, \pempty; T) \equiv \rho(\pfull; T) - \rho(\pempty; T).
-%    \label{eq:delcap}
-%\end{equation}
-% The pressures $\pfull$ and $\pempty$ correspond to chemical potentials $\mufull$ and $\muempty$ via an equation of state for the gas.
-% Eqn.~\ref{eq:delcap}
-% This assumes that the tank reaches equilibrium %, neglecting the kinetics of ad/desorption
-% at the end of the filling and depletion stages.
-
-%The assumption of isothermal operation neglects (i) heat released/consumed during ad/desorption, which can diminish the deliverable capacity by raising/lowering the temperature of the adsorbent during filling/discharge \cite{mota1997dynamics,chang1996behavior}, and (ii) an engineering strategy to use waste heat from the engine to heat the adsorbent and drive off residual gas when it nears $\pempty$ \cite{gomez2014exploring}.
-
-%Second, an engineering strategy to increase the deliverable capacity beyond the isothermal deliverable capacity is to reroute waste heat from the engine to the adsorbed gas tank during delivery.
-By raising the temperature of the adsorbent during the discharge stage, we drive off a fraction of the residual gas that would otherwise remain when the tank reaches $\pempty$ .
-In such a case, the [non-isothermal] deliverable capacity of the tank is $\rho(\pfull, T)-\rho(\pempty; T_H)$ where $T_H$ is the temperature to which we heat the adsorbent at discharge.
-This strategy, however, requires a more complicated design.
-
-% \begin{limitation}\label{limit:isothermal}
-% We assume isothermal pressure-swing adsorption.
-It is possible to achieve a greater deliverable capacity by raising the temperature of the adsorbent as the gas is extracted to drive off the residual gas.
-% \end{limitation}
-
-\section{Review of previous work}
-There has been considerable work attempting to establish an upper bound on the isothermal deliverable capacity in pressure-swing adsorption.
-% We now review previous work towards placing an upper bound on the isothermal deliverable capacity of hydrogen or methane by pressure-swing adsorption.
-% The first body of work rests on the simplified Langmuir model; the second body of work rests upon imposing energetic models for the gas-substrate and gas-gas interactions.
-
-Early work showed that, in the simplified Langmuir model, there exists an optimal free energy of adsorption (which determines the Langmuir constant) that maximizes the deliverable capacity $\rho_D$~\cite{matranga1992storage,bhatia2006optimum,simon2014optimizing}.
-If the gas-substrate interaction is too weak (strong), too little (much) gas adsorbs (is retained) at $\pfull$ ($\pempty$), diminishing $\rho_D$.
-An upper bound on the deliverable capacity of a Langmuir material follows if each adsorption site is endowed with the optimal free energy of adsorption.
-However, remaining is the question of how many adsorption sites per volume a porous material can practically offer, under the constraint that these adsorption sites offer the optimal free energy of adsorption.
-Moreover, gas-gas attractions, neglected in the Langmuir model, could recruit more gas in the material at $\pfull$ than at $\pempty$ and enhance the deliverable capacity~\cite{simon2014optimizing}.
-
-G\'omez-Gualdr\'on \emph{et al.} \cite{gomez2017impact} introduced a model that accounted for gas-gas interactions via an intermolecular potential and idealized the substrate in two different ways: (1) discrete adsorption sites packed into an FCC lattice and (2) a volume endowed with a spatially uniform, background potential energy field.
-According to molecular simulation of methane adsorption, the ARPA-E deliverable capacity target of 315 L STP/L could be reached in both of these idealized substrates.
-%The ARPA-E target of 315 L STP/L was reached when adsorption sites were endowed with an optimal heat of adsorption of 11.5 kJ/mol and adsorption sites were $5.3$ \AA~apart.
-In a second model, they endowed a volume with a background potential energy field, then simulated the adsorption of methane in the volume in the grand-canonical ensemble, using a Lennard-Jones potential for methane-methane interactions.
-The maximal deliverable capacity of 357 L STP/L was achieved with a background heat of adsorption of 8.6 kJ/mol.
-However, both models neglected the space occupied by atoms of the porous material that are needed to endow the adsorption sites/volume with the attractive potential energy.
-% To address this, in the third model, they constructed pseudo materials by placing Lennard-Jones spheres in space with tunable well depths.
-When constrained to form a connected network, none of the psuedo-materials reached the ARPA-E deliverable capacity target.
-
-A third body of work attempted to account for both gas-gas interactions \emph{and} steric interactions of the gas with the atoms of the porous material.
-% explores the upper limits of the deliverable capacity while accounting for gas-gas interactions \emph{and} the atoms of the porous material that are required to attract the guest; these material atoms exclude the adsorbate from regions of the control volume.
-%More works obtained an empirical upper bound on the deliverable capacity by explicitly accounting for the atoms of the material needed to create the potential energy field felt by the gas molecules.
-Simulations of methane adsorption in hundreds of thousands of explicit nanoporous crystal structures---both real and hypothetical---suggested that the ARPA-E deliverable capacity target is infeasible \cite{simon2015materials}; the highest simulated methane deliverable capacity was 196 cm$^3$~STP/cm$^3$.
-Confidence in this conclusion rests upon (i) the accuracy of the intermolecular potentials describing the molecular interactions and (ii) the sufficient sampling of material space, i.e.\ that the structures considered are representative of the set of possible materials.
-To further explore material space and address sensitivity to the potentials used, scaling the Lennard-Jones potential well depths of material atoms to model enhanced interactions \cite{gomez2014exploring}, placing Lennard-Jones spheres in a unit cell randomly to form ``pseudo-materials'' \cite{kaija2018high}, and generating fictitious potential energy fields via a generative adversarial network trained on zeolite structures \cite{lee2019predicting} all generated model substrates that all failed to meet the ARPA-E methane deliverable capacity target.
-
-% Brown and Freeman \cite{bhown2011analysis}
-
-In this work, we place a rigorous upper bound on the deliverable capacity of a pure gas in a rigid substrate by endowing a control volume with a spatially uniform background energy field.
-Instead of using a molecular model for the gas~\cite{gomez2017impact}, we use the experimental equation of state to account for gas-gas interactions.
-In addition, we prove using the calculus of variations that this spatially uniform substrate yields a maximal deliverable capacity.
-We use our framework to place an upper bound on the deliverable capacity of methane and hydrogen.
-
-\section{An upper bound on the deliverable capacity of a pure gas in a rigid porous material}\label{sec:upper-bound}
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/methane-298-rho-mu}
-  \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several MOFs at 298\ K as a function of chemical potential (bottom axis) and pressure (top axis).
-The bulk methane density is from the National Institute of Standards and Technology (NIST)~\cite{nist}.
-The methane adsorption isotherms in the MOFs are experimental data from Ref.~\cite{mason2014evaluating, furukawa2009storage}.
-  }
-  \label{fig:density-vs-mu-ch4}
-\end{figure}
-
-We now develop a thermodynamic framework that places an upper bound on the deliverable capacity of any pure gas in a rigid porous material.
-
-The thermodynamic properties of a bulk, pure gas are characterized by an equation of state. 
-Of particular interest for gas storage and delivery is the density of the gas as a function of chemical potential $\mu$ and temperature $T$, $\rho_g(\mu,T)$, which is shown for methane gas in Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K.
-For comparison, we also show the density of methane adsorbed into several porous materials.
-
-% \davidsays{This paragraph could move considerably later. Or cut entirely?}
-% We model the interaction of substrate with gas as a simple Gibbs free energy of adsorption per mole (or atom) \gst.
-We will assume that this Gibbs free energy is independent of the concentration of gas.
-At low concentrations, this assumption is reasonable, because the gas molecules can be assumed to be widely separated and their interactions weak.
-In the limit of high concentrations, we expect that the attraction of the gas to the substrate will be diminished, since the ``good spots'' will already be occupied.
-The choice of a concentration-independent $\gst$ will thus have a tendency to overestimate the deliverable capacity, providing an upper bound (for more on this, see Sec.~\ref{sec:qualitative-max}).
-
-To place an upper bound on the deliverable capacity, consider a substrate whose sole interaction with the gas is to introduce a spatially uniform potential energy $\V$
-for gas molecules in the control volume (where $\V<0$ for an attractive potential).
-Because this interaction is uniform within the substrate, the gas-gas interactions and thus fluid structure in such a substrate are identical to those of the pure gas at the same density and temperature.
-This allows us to obtain the adsorption properties of this model substrate using the experimentally measured properties of the pure gas~\cite{nist}.
-Intuitively, the deliverable capacity of a gas in such a homogeneous substrate with the optimal potential energy is an upper bound because, in a realistic material, (i) spatial inhomogeneity of the potential energy results in some points in the control volume offering a suboptimal attraction for the gas and (ii) the atoms required to create the potential field will exclude gas from occupying a fraction of the control volume.
-%\corysays{the suboptimal energy is hard to argue since that would affect entropy too...}
-
-% \corysays{I propose to replace ``homogenous'' with ``spatially uniform'' everywhere instead of using both interchangeably.
-% \davidsays{We replaced ``homogeneous potential'' with spatially uniform, but were less confident replacing ``homogeneous substrate'', which has a slightly different meaning, and might need a slightly different name.
-e.g.
-maybe ``uniform substrate'' or ``model substrate''?}}
-
-%\corysays{what do u think about callin this an ``ideal substrate'' or giving it some name so we can refer to it hereon?
-\davidsays{I think I'd prefer ``homogeneous substrate'' or ``idealized substrate''}}
-
-%\subsection{Finding the density}\label{sec:finding-n}
-
-The spatially uniform potential $\V$ representing gas-substrate interactions behaves as an external potential and effectively shifts the chemical potential (or, equivalently, molar Gibbs free energy) of the gas in the material, just as gravitational potential energy causes the density of air to vary with altitude.
-Consequently, the density of gas in our homogeneous substrate is:
-\begin{align}
-    \rho(\mu,T) &= \rho_g(\mu - \V,T). \label{eq:mof-density}
-\end{align}
-See Sec.~\ref{sec:V_shifts_chem_pot} for a derivation.
-% where $\rho_g(\mu,T)$ is the density of the pure gas at chemical potential $\mu$, shown for methane in Fig.~\ref{fig:density-vs-mu-ch4}. 
-% cory says: ^ we already defined \rho_g in pargraph above.
-
-The deliverable capacity of gas in our homogeneous substrate is thus:
-\begin{align}
-    \rho_D(\V,T) &= \rho(\mufull,T) - \rho(\muempty,T),
-    \label{eq:DofPhi}
-    \\
-    &= \rho_g(\mufull-\V,T) - \rho_g(\muempty-\V,T),
-\end{align}
-where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the pressures $\pfull$ and $\pempty$, respectively.
-Thus, $\rho_D$ for our homogeneous substrate as a function of $\V$ is the difference between two shifted versions of $\rho_g(\mu; T)$.
-Figure~\ref{fig:methane-298-D} shows the two shifted bulk methane density curves and their difference.
-The horizontal axis
-is the difference in molar Gibbs free energy at fixed density and temperature between the gas within the substrate and the pure gas (\gst), which is equal to $\V$ in our model.
-We see a potential $\V$ that maximizes the deliverable capacity of our ideal substrate, balancing the need to maximize the density at $\pfull$ against the need to minimize residual gas retained at $\pempty$.
-%Our theoretical framework thus recovers an upper bound on the deliverable capacity of methane (298\ K, $\pfull=$65\ bar, $\pempty=$5.8\ bar) in porous materials as 374\ L STP/L.
-The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, \red{achieved for $\V_{opt} = ...$}.
-%
-%As a ballpark, we can expect the attractive potential to be optimal around the maximum slope of $\rho_g(\mu)$.
-This is the point where the steric repulsion between the gas atoms starts leading to saturation of the density.
-
-An essential question is whether our idealized substrate offering the spatially uniform potential $\V_{opt}$ for the gas places an upper bound on the deliverable capacity in all rigid porous materials.
-%yields the maximum deliverable capacity over all possible potential energy fields. 
-In the SI (see Section~\ref{sec:fixme} \davidsays{Look up section label}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest.
-The potential energy field is \emph{static} if it is unaffected by the presence of the gas; consequently, our model does not apply to flexible materials that undergo gas-induced conformation changes \cite{schneemann2014flexible}.
-We next argue that this extremum is a maximum by addressing two alternative possibilities: the extremum could turn out to be (i) a saddle point or (ii) a local, rather than global, maximum.
-
-Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum $\rho_D$ over all \emph{spatially uniform}, static potential energy fields.
-To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a maximum deliverable capacity over \emph{all} (including non-spatially uniform) static potential energy fields (a broader claim), consider a non-spatially-uniform variation on $\V_{opt}$.
-We will show that the effect of this variation is to reduce the deliverable capacity.
-At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, the additional gas molecules will be forced into higher energy locations.
-Thus, the mean attraction will be lowest at \pfull\  and highest at \pempty.
-This will result in $\gst$ increasing monotonically with with pressure.
-As shown in the Supplementary Information (see Sec.~\ref{sec:monotonic}), this results in the deliverable capacity of a non-uniform potential being lower than the deliverable capacity of a uniform potential with $\V=\gst$ for either $\pfull$ or $\pempty$.
-Thus a spatially uniform potential will give the best deliverable capacity amongst all static potential energy fields that give the same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will give the a greater deliverable than any static non-uniform potential.
-
-% Yo David, I changed ur paragraph below. tell me wut u think.
-% To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a \emph{maximum} deliverable capacity over all static potential energy fields, consider the case of a material that offers a non-spatially-uniform potential energy felt by the gas molecules.We wish to show that this material will have a lower deliverable capacity than our ideal substrate.
-At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, some of the gas molecules will be forced into higher energy locations.
-Thus, the mean attraction will be lowest when the tank is ``full,'' and highest when the tank is ``empty.'' This is the opposite of what we would want to achieve a high deliverable capacity.
-Based on this qualitative reasoning, we expect the spatially uniform potential \red{$\V_{opt}$} to give the maximal deliverable capacity, since it will provide the most attraction for the molecules added at high densities.
-
-% \corysays{``To qualitatively argue that the spatially uniform potential provides a \emph{maximum} deliverable capacity'' do we need to qualify that this spatially uniform potential must hv the optimal $\V$?
-obviously a spatially uniform potential can be better than a spatially uniform one with a poor $\V$.
-maybe we should start with our optimal homogenous potential, then say introduce a little blip in it from there, like in calculus of variations.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/methane-298-n-vs-G}
-    \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
-    Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line (not visible at this scale).}
-    \label{fig:methane-298-D}
-\end{figure}
-
-\section{Results}
-Although our theoretical framework allows us to place an upper bound on the deliverable capacity of many different gases in a rigid porous material, we focus on the maximal deliverable capacity of methane and hydrogen gas in our homogeneous substrate due to their application as transportation fuels.
-To characterize the density of the gases, $\rho_g(p, T)$, we use experimental data from NIST~\cite{nist}, which naturally includes quantum effects that particularly affect hydrogen at low temperature~\cite{kumar2006quantum}.
-In the context of storage onboard passenger vehicles, we compare our upper bound with several prominent porous materials using experimental adsorption isotherms from the literature; we also compare with deliverable capacity targets set by the US DOE.
-
-Figure~\ref{fig:methane-298-D} shows the upper bound for methane storage at room temperature~(374~cm$^3$~STP/cm$^3$). In addition to the predicted maximum deliverable capacity as a function of \gst,
-\corysays{subtlety here: isnt ur proof restricted to saying $\V_{opt}$ provides the upper bound?
-does it really prove that, for fixed and nonoptimal $\gst$, a spatially uniform potential is optimal over a non-spatially uniform one?
-I don't think it does...}
-the ARPA-E target of 315~cm$^3$~STP/cm$^3$~\cite{arpaemove} is shown for context.
-For an adsorbent with at least 84\% void fraction, the ARPA-E target is theoretically possible.
-The experimental deliverable capacities of several adsorbents~\cite{mason2014evaluating} are also shown over a range of $\gst$ (converted from the measured adsorption isotherm as explained in \davidsays{confirm this} Sec.~\ref{sec:phi-is-delta-g}).
-These can be compared with the highest observed deliverable capacities for methane at room temperature in rigid materials, up to 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}.
-% don't think the below is helpful or correct: Gst opt is reached by some... there are void fraction constraints too.
-% To reach the ARPA-E target, an adsorbent material with a higher $|\gst|$ would need to be found.
-
-\corysays{I know it doesn't matter, but should we choose L STP/L or cm3 STP/cm3?
-we use both, some might get confused/distracted?
-\davidsays{We chose to use the units used for each of the targets thinking it would be more familiar and/or easier to compare.
-I don't like either set of units, and if we were to pick ``our'' choice, I'd lean towards something dimensionless with a sane comparison (e.g.
-relative to the gas density at our high pressure)}}
-
-\corysays{need to explain why the MOFs are lines not points... that \gst has a range for MOFs.
-\jpsays{due to a different deliverable capacity at full pressure and empty pressure.
-For hydrogen the SI (by Paula Garcia-Holley) has total uptake plots so we could make a curve even but would have a lot of error due to manual extraction?}}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-298-n-vs-G}
-    \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
-    \label{fig:hydrogen-298-D}
-\end{figure}
-
-\corysays{Figs.~\ref{fig:hydrogen-298-D} ~\ref{fig:methane-298-D} why not just let $\gst$ be negative and have an arrow pointing in direction of more attraction?
-as written the legend is not correct since those are not the functions being shown.
-in the main text we never discussed what \gst is for the MOF and why the MOFs are lines not points (\gst is a function of density now)}
-
-\jpsays{I edited everything below this comment. The original is commented out}
-
-Storage of hydrogen is considerably more challenging owing to its relatively weak interaction with adsorbents.
-The DOE ULTIMATE deliverable capacity target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C \red{check}.
-Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve for hydrogen storage along with a range of experimental measurements for known MOFs.
-The DOE ULTIMATE deliverable capacity target is theoretically possible, however by such a small margin that we can safely rule out the possibility of reaching this target through storage and delivery of hydrogen in \emph{any} rigid substrate at room temperature.
-Such a material would require a void fraction of at least 94\% \red{check}.
-On top of this, the DOE ULTIMATE target requires an optimal $|\gst|$ of 10~kJ/mol which is far greater than what is found in observed porous materials.
-This reflects the known fact that hydrogen interacts with substrates far more weakly than methane does.
-One possibility that has been pursued is storage at cryogenic temperatures.
-See Section~\ref{sec:cryo-hydrogen} in the Supplementary Information for the upper bound for storage of hydrogen at 77 K.
-
-Why do we not experimentally observe materials approaching the upper bound on the deliverable capacity?
-Foremost, any adsorbent substrate will be composed of atoms, which will exclude the gas from some volume.
-Due to strong short-range interactions, substrate atoms must be uniformly distributed to achieve strong attraction for gas atoms.
-Uniform atom distribution sets a limit on the pore volume.
-% Because strong interactions are short-range, achieving a strong attraction requires that substrate atoms be distributed throughout the volume, which places a limit on the pore volume.
-In contrast, real materials have a spatially non-uniform attraction for gas atoms.
-The result of this is that there are regions of space with non-optimal attraction.
-For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.
-% Finally, real materials must have a spatially non-uniform attraction for gas atoms, resulting in some regions of space where the attraction is not optimal.
-For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.
-
-\section{Conclusion}
-We have established an upper bound on the deliverable capacity via pressure-swing adsorption in rigid porous solids based on experimentally measured properties of pure gases.
-While these upper bounds do not rule out the discovery of materials that reach current DOE targets, they cast strong doubt on the possibility of achieving these goals when we consider the additional constraints imposed due to steric hindrance between substrate atoms and the adsorbate.
-Our upper bound does indicate that those goals cannot be exceeded by more than 19\% for methane and 6\% for hydrogen \red{check}.
-Fortunately, there are some limitations to these upper bounds which suggest avenues for future developments which could allow the targets to be met.
-
-The first limitation of our proof is that we restricted ourselves to \emph{isothermal} pressure swing storage.
-By raising the temperature of the adsorbent during gas discharge to drive off residual gas \cite{gomez2014exploring}, the deliverable capacity could be enhanced, albeit at the cost of a more complicated engineering design of the fuel tank and vehicle.
-%While this approach poses engineering challenges, in principle it could allow an arbitrarily large deliverable capacity.
-
-% \corysays{no, cant lower filling temp. when park it will heat up then go above pressure rating of tank...}
-
-The second limitation of our proof arises in the assumption of a rigid substrate.
-The rigid substrate acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
-% The second limitation of our proof arises in the assumption of a rigid substrate, which acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
-For most porous materials this is a reasonable approximation, and this assumption is frequently made in both the simulation and theory of porous materials~\cite{witman2017influence}.
-However, there are cases where the substrate can in effect provide a very strongly gas-density-dependent interaction through structural flexibility~\cite{schneemann2014flexible}.
-A flagship example is MOF Co(bdp)~\cite{choi2008broadly}, which possesses a wine-rack-like topology capable of hinge motion.
-At low methane pressure, Co(bdp) adopts a collapsed, nonporous state, but expands to a porous state and fills with gas at higher pressures \cite{mason2015methane}.
-This allows Co(bdp) to fully expel its residual gas at low pressures.
-Our upper bound does not consider such material effects and thus flexible materials could have significantly higher deliverable capacities.
-% Such materials can indeed violate our upper bound and thus could lead to significantly improved deliverable capacities.
-
-We note that our upper bound can readily be applied to the storage of other gasses of interest, with the proviso that the gas is far from crystallization. Our code is available at \davidsays{XXXXX}.
-
-% \clearpage
-
-% Here we will articulate the assumptions required for our upper bound to hold. Some of these assumptions serve as a guide for chemists who aim to synthesize a material that violates our upper bound.
-
-% Our proof that the spatially uniform potential is an extremum assumes that the fluid is stable at the temperature and densities involved.
-% \begin{limitation}\label{limit:crystallization}
-% Our proof assumes the gas is cannot become spontaneously solid at the temperature of interest.
-% \end{limitation}
-% Our upper bound will \emph{not} be valid for gasses which may crystallize in their pure form.
-This is because a periodic potential can interact with a high-density crystal far more strongly than it interacts with a low-density gas.
-We are not aware of a practical application of this principle, but theoretically it is possible.
-For the application of gas storage for light gasses such as methane or hydrogen at moderately high temperatures (say liquid nitrogen or warmer), the pure gas is sufficiently far from its crystal form that we need not concern ourselves with crystallization.
-We do not see this particular assumption as one that can be worked around.
-
-% There remains the possibility that the spatially uniform potential is a local maximum, but there is another global maximum.
-We do not see how to formally rule out this possibility for an arbitrary gas-gas interaction, but the qualitative arguments above seem sufficient to rule out this possibility for a real gas.
-
-% However, there was a more fundamental assumption, which is that the porous material acts as an effective potential for the gas molecules, and that the gas-gas interaction is unmodified.
-% \begin{limitation}\label{limit:field}
-% Our proof assumes the interaction of the gas with the substrate does not modify the substrate.  Equivalently,  the interaction of substrate with gas operates as a simple potential energy field.
-% \end{limitation}
-% For \emph{most} porous materials this seems likely to be a pretty good approximation.
-However, there are unusual cases where the substrate can in effect provide a very strongly density-dependent interaction.
-This is perhaps easiest to see in the example of \davidsays{look up the collapsing MOFs description from the RMS-MOF paper}, in which the gas can introduce a phase change in the solid, which has the effect of drastically lowering the density of gas in the material.
-Such materials can indeed violate our upper bound.
-\davidsays{Remember to discuss this in the conclusions or perhaps discussion, we can frame these limitations as guides.
-The limitations of our upper bound are the design requirements for a material that will do any better.}
-
-% [impurities?]
-
-% \section{Conclusion}
-% As many have suspected \davidsays{cite previous work}, pressure-swing adsorption in rigid porous solids is not going to enable reaching targets.
-\davidsays{conclude explicitly about each target}  Unlike previous work, our upper bound on deliverable capacity is based on experimentally known properties of the gas of interest rather than simulations.
-
-% The limitations on our proof of an upper bound on the deliverable capacity provide hints as to approaches that could in principle achieve the targets.
-In this paper we have identified three limitations of our upper bound on the deliverable capacity.
-Two of these three limitations suggest engineering strategies that do hold promise of achieving the targets for hydrogen and methane storage.
-Limitation \ref{limit:crystallization}, which states that the gas must not crystallize does not means for improvement, since it is a limit on the gasses to which our bound may be applied, and neither hydrogen nor methane are in any danger of crystallizing.
-
-% Crystal density vs. pelletiziation density.
-
-% A more subtle limitation of our proof is that it assumes that the substrate itself is not modified by the presence of the gas.
-While this assumption is never precisely true, since any adsorbed atom will change the electron distribution of the substrate to some degree, it tends to be sufficiently accurate that most atomistic simulations can get away with assuming that the substrate is unmodified.
-A dramatic exception to that rule is the field of flexible MOFs, in which the presence of the gas can induce a phase change which can \emph{drastically} modify the gas-substrate interaction~\davidsays{cite experiments}\cite{witman2019}.
-
-% wut about mixtures?
-
-%TC:ignore
-
-\bibliography{bibfile}
-
-\appendix
-\clearpage
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}%
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}%
-\renewcommand{\thesection}{SI \Alph{section}}%
-
-\section{An external, spatially uniform potential $\V$ shifts the chemical potential $\mu$ of the gas} \label{sec:V_shifts_chem_pot}
-We show that imposing an external, spatially uniform potential $\V$ to a gas has the effect of shifting the chemical potential $\mu$ of the gas, recovering eqn.~\ref{eq:mof-density}.
-Consider a control volume $\Omega$ with volume $V=|\Omega|$ (large enough to neglect boundary effects) that is endowed with the external, spatially uniform potential $\V$.
-Impose the grand-canonical ensemble, where this control volume can exchange energy and particles with a bath of gas at temperature $T$ and chemical potential $\mu$.
-
-To denote a microstate of this system, let $N$ be the number of gas particles in the control volume and $\mathbf{x}_1,...,\mathbf{x}_N$ be their positions.
-Then, the potential energy $E$ of a microstate of the control volume is:
-\begin{equation}
-    E(\mathbf{x}_1,...,\mathbf{x}_N) = N\V + E_{gg}(\mathbf{x}_1,...,\mathbf{x}_N),
-\end{equation} where $E_{gg}$ is the (unknown and complicated) interatomic potential for gas-gas interactions that governs the (real) gas properties.
-The first term arises from each gas molecule feeling the external potential $\V$, where $\V < 0$ corresponds to attraction.
-
-\corysays{wut about orientations? shld we include those?}
-
-The grand-canonical partition function of the control volume is:
-\begin{multline}
-    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\xvec_1, ..., \xvec_N)} e^{\beta (\mu - \V) N} d\xvec_1 \cdots d\xvec_N.
-    \label{eq:gcpf}
-\end{multline} We recognize this as the grand canonical partition function of the bulk gas in the control volume without the external potential, but shifted in chemical potential:
-\begin{equation}
-    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
-    \label{eq:xi_vs_xi0}
-\end{equation}
-where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the bulk gas in the control volume in the absence of an external potential.
-Importantly, this equivalency is predicated on the potential $\V$ being spatially uniform.
-Therefore, the thermodynamic properties of the gas sitting in the spatially uniform external potential $\V$-- i.e., adsorbed in our idealized substrate-- are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$ ($T$ fixed here).
-As $\V$ becomes more negative, corresponding to a more attractive adsorbent, the thermodynamic properties of the adsorbed gas in our ideal substrate are equivalent to the gas at higher chemical potential.
-
-\section{Proof of extremum}
-To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.
-In this proof we will make use of the Fourier transform of this potential:
+To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between guest and substrate $\V(\vec r)$ that varies with space.
+In this proof we will actually make use of the Fourier transform of this potential:
 \begin{align}
@@
 \end{align}
-where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reordered the functional and partial derivatives.
+where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reorderd the functional and partial derivatives.
 Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ just gives the number density:
@@
 \end{align}
-The number density is itself spatially uniform for any system that is stable in a fluid state at this density (i.e.
+The number density is itself homogeneous for any system that is stable in a fluid state at this density (i.e.
 does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.
@@
 \end{align}
-Thus the functional derivative of both the chemical potential and $N$ with regard to $\V(\vec r)$ are themselves spatially uniform.
-Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e.
+Thus the functional derivative of both the chemical potential and $N$ with regard to $\V(\vec r)$ are itself homogeneous.
+Since we already maximize $D$ with respect to the homogeneous component of the potential (i.e.
 $\vec k=0$), the derivative of $D$ with respect to any change of potential is zero.
 
-This demonstrates that a spatially uniform potential leads to an extremum value of the deliverable capacity.  This proof is insufficient, however, to show that it must be a true maximum.
-
-\section{The real-substrate analogy of $\V$ is $\gst$}
-\label{sec:phi-is-delta-g}
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figs/g_st_piston_2.png}
-    \caption{Cartoon illustrating the definition of $\gst$ and a hypothetical experiment to measure it with a piston and removable barrier that separates two equal volumes, one with a porous material (bottom) and one with free space (top).
-The initial state (left) is a bulk gas phase at density $\rho$ and temperature $T$.
-An infinitely-thin, removable barrier separates the gas from the porous material (bottom).
-The barrier is then slowly removed, and the piston slowly pushes all gas molecules into the porous material in an isothermal process.
-The final state (right) is an adsorbed gas phase at density $\rho$ and temperature $T$.
-The quantity $\gst$ follows from {\color{red} the heat released?}.}
-    \label{fig:delta-G-cartoon}
-\end{figure}
-
-The parameter describing our idealized substrate is $\V$, the spatially uniform potential felt by a gas molecule adsorbed in the idealized substrate. 
-A natural question is how this potential relates to the properties of real substrates. 
-The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the gas (see Sec.~\ref{sec:V_shifts_chem_pot}).
-Because our ideal substrate shifts the chemical potential of the gas molecules by providing a spatially uniform potential energy field, the entropy of the gas in the ideal substrate is equal to the entropy of the gas in its bulk state at the same density and temperature.
-In contrast, a real substrate provides a non-spatially uniform potential.
-Consequently, the entropy of the gas inside a real substrate is \emph{not} equal to the entropy of the bulk gas at the same density and temperature.
-Therefore, the parameter analogous to $\V$ in a real substrate will involve both energy and entropy.
-The real-substrate analogy of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
-\begin{equation}
-   \gst(\rho, T) \equiv
-   %\frac{G_{\text{ads}}(\rho, T) - G_{\text{gas}}(\rho, T)}{N}\\ 
-    \mu_{\text{ads}}(\rho, T) - \mu_{\text{gas}}(\rho, T).
-  % \gst(\rho, T) &\equiv g_{\text{gas}}(\rho, T) - g_{\text{ads}}(\rho, T)
-  \label{eq:g_st}
-\end{equation}
-The isosteric Gibbs free energy difference $\gst$ is the difference in molar Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and the bulk gas \emph{with the same density of gas molecules}.
-The quantity $\gst$ does \emph{not} correspond to a change in the molar Gibbs free energy as a molecule is adsorbed, which is zero under conditions of coexistence.
-The quantity $\gst$ in a real substrate is a direct analogy to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
-Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to  measure $\gst$ via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
-Note that $\gst$ is a property of both the substrate and the identity of the gas.
-Because real substrates offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho, T)=\V$.
-Consequently, throughout this article, we show $\gst(\rho, T)$ for real substrates at both conditions relevant to gas storage and delivery, $\pfull$ and $\pempty$.
-% In the initial state, the partition isolates the bulk gas exhibiting density $\rho$ and temperature $T$.
-Then, the partition is removed and the piston pushes down on the gas so that the gas slowly adsorbs into the MOF in an isothermal process.
-\red{isothermal rite?} In the final state, the piston has pushed all gas into the MOF, resulting in an adsorbed gas phase with the same density $\rho$ and temperature $T$ as the initial state.
-The isosteric Gibbs free energy of adsorption, $\gst$, then follows from \red{is it the amount of heat released?}.
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} looks too jumbled. two separate figs? plus $G_{st}$ is capitalized here, should be lowercase. where is all this data from}
-
-In practice, we can readily compute $\gst(\rho, T)$ of a real gas/substrate system from (i) the (experimental or simulated) equilibrium adsorption isotherm of the gas in the substrate and (ii) the (experimental or simulated) chemical potential of the bulk gas.
-Consider the real substrate in thermodynamic equilibrium with a bulk gas at fixed temperature $T$ and pressure $p$, and let $\rho=\rho(p, T)$ be the density of gas in the substrate. 
-At coexistence, the chemical potential of the bulk gas is equal to the chemical potential of the adsorbed gas in the substrate.
-Thus, we can use the experimentally known molar Gibbs free energy of the pure, bulk gas system at temperature $T$ and pressure $p$ to determine the molar Gibbs free energy of the adsorbed system: $\mu_{\text{ads}}(\rho, T)=\mu_{\text{gas}}(p, T)$.
-We can then also look up the known chemical potential of the bulk gas at the same density and temperature as in the substrate, $\mu_{\text{gas}}(\rho, T)$.
-Via eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.
-
-An interesting question is how $\gst$ relates to the commonly measured and reported isosteric heat of adsorption $q_{st}$, which is roughly the energy change when a gas molecule is adsorbed~\cite{sircar1999isosteric, tian2017differential}.
-Figure~\ref{fig:qst-vs-delta-G} shows how $q_{st}$ compares to $\gst$ for several prominent adsorbents.
-In every case, $|q_{st}|>|\gst|$ because the gas in the adsorbent always has less entropy than the gas in the bulk at the same density and temperature.
-That is, while adsorption is energetically favored, it is entropically disfavored due to the restrictions imposed on the configuration of the gas molecules via steric interactions with the substrate itself; this counters the energetic attraction.
-
-\corysays{can u mathematically justify why $\gst$ is the relevant parameter through the grand canonical partition function in eqn.~\ref{eq:gcpf}?
-it is not true that $\Xi(\mu, V, T)$ for the real substrate is equal to $\Xi(\mu+\gst, V, T)$}
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} should be split in two. one for methane one for h2.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/qst-vs-delta-G}
-    \caption{Relationship between $\gst$ and the isosteric heat $q_{st}$ for several prominent adsorbents.
-The dots represent the properties of methane adsorption at 298\ K and at 5\ bar and 100\ bar.
-The $\times$'s represent the properties of hydrogen adsorption at 77\ K and at 5~bar and 100~bar.}
-    \label{fig:qst-vs-delta-G}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/methane-298-gst}
-    \caption{The density-dependence of $\gst$ of methane in several adsorbents (298\ K). Note that $\gst$ is monotonic in $\rho$. Data from Refs.~\ref{mason2014evaluating,furukawa2009storage}.
-    }
-    \label{fig:methane-gst}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/four-cases}
-    \caption{Four cases we consider.}
-    \label{fig:delta-gst-maximum}
-\end{figure}
-
-\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
-Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding to a full and empty density.
-The deliverable capacity of the MOF is equal to the difference between the full and empty density.
-Examining Fig.~\ref{fig:methane-gst}, we see that a significant variety of known experimental $\Delta g_\text{st}$ curves are monotonic.
-Furthermore, our qualitative argument in Section~\ref{sec:upper-bound} suggests that this function \emph{should} be monotonically increasing for rigid substrates, as increasing the density of gas forces some of the gas to reside in higher-energy sites.
-If this is always the case, the deliverable capacity will always be bounded by our theoretical upper bound, even for a value of $\V$ that is not optimal.
-
-To show this, we perform a thought experiment illustrated in Fig.~\ref{fig:delta-gst-maximum}.
-We begin with two pairs of boxes containing gas.
-Two of the boxes are filled with a MOF, and the other two have a uniform potential energy $\V$.
-We could imagine the two boxes at $\V$ as being empty but at very low altitude where gravity provides a potential energy difference (relative to the box with the MOF in it) of $\V$ to each gas molecule (perhaps we are on Jupiter?).
-One box of each sort has a density of gas $\rhofull$ corresponding to $\pfull$ in the MOF system, while the other two boxes has gas density $\rhoempty$.
-
-We now consider what happens if we introduce a diffusive connection between two boxes that are initially at the same density, for instance by connecting them with a tube.
-If the chemical potential of the two boxes are equal, their density will be unchanged, otherwise gas will flow into the box with lower chemical potential.
-The difference in chemical potential is equal to the difference
-\begin{align}
-   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
-   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
-   \\
-   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
-\end{align}
-Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at their initial density after they are connected.
-Thus the deliverable capacity of the MOF will be greater than the deliverable capacity of the model with uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e.
-if $\gst(\rho)$ does \emph{not} monotonically increase.
-By the same token, if we consider the case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable capacity than our model, the MOF must have less residual gas, which means that gas must spontaneously flow from the low-density MOF to the box with potential $\V$, which means that $\gst(\rhoempty)>\V$.
-So once again exceeding our upper bound requires a material with a $\gst(\rho)$ that does not increase monotonically.
-
-Taken together, this indicates that not only is our absolute upper bound an upper bound for rigid MOFs, but the green curve labeled $\rho_D(\gst)$ in Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and \ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal $\gst$.
-
-\section{Cryogenic hydrogen storage}\label{sec:cryo-hydrogen}
-\begin{figure}
-    \centering
-    FIGURE HERE A B C D
-    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-77-n-vs-G}
-    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
-    \label{fig:hydrogen-77-D}
-\end{figure}
-
-One approach to increase the deliverable capacity is to reduce the storage temperature.
-This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling point of nitrogen.
-The DOE ULTIMATE target in this case looks far more achievable, and with a much lower $|\gst|$.
-In fact, an empty tank at this temperature can satisfy the DOE 2020 target.
-The DOE ULTIMATE target is 14\% below the upper bound.
-Actual MOFs fall far short of the theoretical maximum.
-\red{the target is not just a storage density, it specifies temps and pressures too...
-is 77 K allowed?} \davidsays{My recollection of my reading of the DOE target was that anything was allowed, but that the target included ``everything'', which would include both the tank and any cryogenic equipment required.
-So the answer is not so simple.}
-
-%TC:endignore
+This demonstrates that a homogeneous potential leads to an extremum value of the deliverable capacity.  This proof is insufficient, however, to show that it must be a true maximum.
 

### 2020-05-01 ising/ising.tex
wrote a short section describing specific heat capacity plot.
added the conclusion to thesis.
added gas storage to thesis pommerenck.

@@
 
-\title{Flat-histogram method comparison on 2D Ising Model
+\title{Flat-histogram method comparison on 2D Ising model
 }
@@
     \caption{
-    The specific heat capacity versus the reciprocal temperature $\beta$ for the $N=32 \times 32$ system for each histogram method.}
+    The specific heat capacity versus the reciprocal temperature $\beta$ for the $N=32 \times 32$ system for each histogram method at $10^{9}$ moves.}
     \label{fig:N32-cv}
@@
 All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.
+
+Fig.~\ref{fig:N32-cv} shows the specific heat capacity vs.
+the reciprocal temperature $\beta$ at $10^{9}$ moves.
+Each method (except for the SAMC methods) is shown using the same single random number seed.
 

### 2020-05-01 gas-adsorption/colors.py

@@ -1,0 +1,65 @@
+import matplotlib.pyplot as plt
+import matplotlib, re
+import numpy as np
+
+from matplotlib.font_manager import FontProperties
+small_font = FontProperties()
+small_font.set_size('small')
+
+prop_cycle = plt.rcParams['axes.prop_cycle']
+colors = list(prop_cycle.by_key()['color'])
+
+the_order = [
+    'MOF5',
+    'HKUST1',
+    'PCN14',
+    'Ni2dobdc',
+    'Co2dobdc',
+    'Mg2dobdc',
+    'AX21',
+    'NU125',
+    'NOTT112',
+    'rhtMOF7',
+    'CuMOF74',
+    'PCN250',
+    'NU1000',
+    'UiO67',
+    'UiO68Ant',
+    'CYCU3Al',
+    'Zn2bdc2dabco',
+    'COF102',
+]
+
+# cm = plt.get_cmap('viridis')
+for x in np.linspace(0, 1, len(the_order) - len(colors)):
+    colors.append(matplotlib.colors.hsv_to_rgb((x,1,1-x*0.25)))
+# colors = [cm(1.*i/(len(the_order)-1)) for i in range(len(the_order))]
+
+def order(mofs):
+    return [x for x in the_order if x in mofs]+[x for x in mofs if x not in the_order]
+
+def color(mof):
+    if mof in the_order:
+        return colors[the_order.index(mof)]
+    return 'k'
+
+latex_mof_names = {
+    'Ni2dobdc': 'Ni$_2$(dodbc)',
+    'Co2dobdc': 'Co$_2$(dodbc)',
+    'Mg2dobdc':'Mg$_2$(dodbc)',
+    'Zn2bdc2dabco': 'Zn$_2$(bdc)$_2$(dabco)',
+    'CYCU3Al': 'CYCU$_3$Al',
+    'UiO68Ant': 'UiO68Ant',
+    'rhtMOF7': 'rht-MOF-7',
+}
+def latex_mof(mof):
+    if mof in latex_mof_names:
+        return latex_mof_names[mof]
+    print(mof, re.split('(\d+)', mof))
+    letters, numbers,empty = re.split('(\d+)', mof)
+    if empty == '':
+        return letters + '-' + numbers
+    return mof
+
+#print(latex_mof('AX231'))
+#print(latex_mof('COF102'))

### 2020-05-01 gas-adsorption/data_scripts/ammonia.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  17.03052*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Ammonia/L'
+Name = 'Ammonia'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/argon.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  39.948*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Argon/L'
+Name = 'Argon'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/benzene.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  78.11184*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Benzene/L'
+Name = 'Benzene'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/butane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  58.1222*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Butane/L'
+Name = 'Butane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/carbon_dioxide.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  44.01*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Carbon Dioxide/L'
+Name = 'Carbon Dioxide'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/carbon_monoxide.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  28.01*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Carbon Monoxide/L'
+Name = 'Carbon Monoxide'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/cyclopropane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  560.49158*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Cyclopropane/L'
+Name = 'Cyclopropane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/decane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  142.28*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Decane/L'
+Name = 'Decane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/deuterium.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  4.028*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Deuterium/L'
+Name = 'Deuterium'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/dodecane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  170.33*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Dodecane/L'
+Name = 'Dodecane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/ethane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  30.07*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Ethane/L'
+Name = 'Ethane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/ethene.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  28.05316*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Ethene/L'
+Name = 'Ethene'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/flourine.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  18.998403*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Flourine/L'
+Name = 'Flourine'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/helium.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  4.002602*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Helium/L'
+Name = 'Helium'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/heptane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  100.20194*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Heptane/L'
+Name = 'Heptane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/hydrogen.py

@@ -1,0 +1,168 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight = 2.016*nu.g/nu.mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Hydrogen/L'
+Name = 'Hydrogen'
+
+n_goals = []
+n_goals.append((0.05*kg/L/molecular_weight, 'DOE ULTIMATE','#999999', '-'))
+n_goals.append((0.04*kg/L/molecular_weight, 'DOE 2025',    '#888888', '-'))
+n_goals.append((0.03*kg/L/molecular_weight, 'DOE 2020',    '#777777', '-'))
+
+H2_298_Data = {
+    'HKUST1': {
[…]

### 2020-05-01 gas-adsorption/data_scripts/hydrogen_sulfide.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  34.08088*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Hydrogen Sulfide/L'
+Name = 'Hydrogen Sulfide'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/isobutane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  624.93638*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Isobutane/L'
+Name = 'Isobutane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/krypton.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  83.798*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Krypton/L'
+Name = 'Krypton'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/methane.py

@@ -1,0 +1,106 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight = 16.04*nu.g/mol
+density_units = 0.044135*mol/L # density at STP in mol/L
+density_unit_name = 'cm$^3$ CH$_4$ STP/cm$^3$'
+Name = 'Methane'
+
+n_goals = []
+n_arpa_methane_goal = 315*0.044135*mol/L # 315* density at STP
+n_goals.append((n_arpa_methane_goal, 'ARPA-E', '#999999', '-'))
+
+def linear_interpolate(x, x1, y1, x2, y2):
+    return ((x2-x)*y1 + (x-x1)*y2)/(x2-x1)
+
+# Methane experimental data is from Mason and Long 2014
+# The COF-102 data is from https://doi.org/10.1021/ja9015765
[…]

### 2020-05-01 gas-adsorption/data_scripts/methanol.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  32.04*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Methanol/L'
+Name = 'Methanol'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/neon.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  20.1797*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Neon/L'
+Name = 'Neon'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/nitrogen.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  28.014*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Nitrogen/L'
+Name = 'Nitrogen'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/nonane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  128.25*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Nonane/L'
+Name = 'Nonane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/octane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  114.26*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Octane/L'
+Name = 'Octane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/oxygen.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  15.9994*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Oxygen/L'
+Name = 'Oxygen'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/parahydrogen.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  2.016*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Parahydrogen/L'
+Name = 'Parahydrogen'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/pentane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  72.14878*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Pentane/L'
+Name = 'Pentane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/propane.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  44.09562*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Propane/L'
+Name = 'Propane'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/propene.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  42.0797*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Propene/L'
+Name = 'Propene'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/propyne.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  40.0639*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Propyne/L'
+Name = 'Propyne'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/sulfur_dioxide.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  64.066*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Sulfur Dioxide/L'
+Name = 'Sulfur Dioxide'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/toluene.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight =  92.13842*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Toluene/L'
+Name = 'Toluene'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/water.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight = 18.01528*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Water/L'
+Name = 'Water'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/data_scripts/xenon.py

@@ -1,0 +1,38 @@
+"""
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits as nu
+from numericalunits import kg, L, mol, mmol, cm, kJ, mg
+gram = nu.g
+
+molecular_weight = 131.29*gram/mol
+density_units = gram/L/molecular_weight
+density_unit_name = 'g Xenon/L'
+Name = 'Xenon'
+
+n_goals = []
+
+def isotherm_experiments(T, p_empty_bar, p_full_bar):
+    return {}
+
+    

### 2020-05-01 gas-adsorption/download-data.sh

@@ -1,0 +1,10 @@
+#!/bin/sh
+
+set -ev
+
+python3 isothermal-ads-gas-save-csv.py 1 10000 298 methane
+
+python3 isothermal-ads-gas-save-csv.py 1 10000 298 hydrogen
+python3 isothermal-ads-gas-save-csv.py 1 10000 77 hydrogen
+python3 isothermal-ads-gas-save-csv.py 1 10000 160 hydrogen
+

### 2020-05-01 gas-adsorption/gibbs-qst.py

@@ -1,0 +1,458 @@
+from __future__ import division, print_function
+import sys, os, re, matplotlib
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits
+
+import colors
+
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+
+# --- Build unit system --- #
+Kelvin = numericalunits.K
+atm = numericalunits.atm
+bar = numericalunits.bar
+gram = numericalunits.g
+kg = numericalunits.kg
+L = numericalunits.L
+Joule = numericalunits.J
+mol = numericalunits.mol
+psi = numericalunits.psi
+cm = numericalunits.cm
+mmol = numericalunits.mmol
+
+mg = numericalunits.mg    # milligram
+mL = numericalunits.mL    # milliliter
+kJ = numericalunits.kJ     # kiloJoule
+
+# HOW TO RUN:
+# python3 gibbs-qst.py
+
+# Define a Global Boltzmann constant
+kB = 8.314*Joule/mol/Kelvin
+
+# Define interpolation functions
+def my_interp(x0,x,y):
+    return np.interp(x0, x, y)
+
+def mu_from_n(particular_n):
+    return my_interp(particular_n, n, mu)
+
+def n_from_mu(particular_mu):
+    return my_interp(particular_mu, mu, n)
+
+def linear_interpolate(x, x1, y1, x2, y2):
+    return ((x2-x)*y1 + (x-x1)*y2)/(x2-x1)
+
+#A list of the five MOFs example AX21-298.txt
+#[AX21,Mg2dobdc,Co2dobdc,MOF5,Ni2dobdc,HKUST1,PCN14]
+
+# For the Hydrogen Data (I have to get Isotherms manually using
+# online plot digitizer although I can save the online project file
+# as data)
+
+#rho_MOF = 18.32*mmol/gram*molecular_weight/MOF_DENSITY
+
+# Looks like I need Qst at p_empty and p_full since most of the curves
+# are not constant!
+H2_crystal_density = {
[…]

### 2020-05-01 gas-adsorption/isothermal-ads-gas-plot.py

@@ -1,0 +1,354 @@
+from __future__ import division, print_function
+import sys, os, re, matplotlib
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits, colors
+
+matplotlib.rcParams['text.usetex'] = True
+
+from matplotlib.font_manager import FontProperties
+
+small_font = FontProperties()
+small_font.set_size('small')
+
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+
+"""
+Create deliverable capacity and density plots from NIST Thermophysical fluid
+data generated with isothermal-save-gas-csv.py using Python 3
+
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+
+
+# Typical usage:
+#        python3 isothermal-ads-gas-plot.py xenon 298 5.8 200
+if len(sys.argv) < 5:
+    print("""python3 isothermal-ads-gas-plot.py [GAS] [T in K] [Pempty in bar] [Pfull in bar]
+
[…]

### 2020-05-01 gas-adsorption/isothermal-save-gas-csv.py

@@ -1,0 +1,149 @@
+from __future__ import division, print_function
+import sys, os, re
+import numpy as np
+import matplotlib.pyplot as plt
+
+import pandas as pd # import data tables from NIST and store with pandas
+
+"""
+Generate isothermal data from NIST Thermophysical fluid database using Python 3
+
+Copyright (c) 2019 - 2020 Jordan K. Pommerenck
+
+Permission is hereby granted, free of charge, to any person obtaining a copy
+of this software and associated documentation files (the "Software"), to deal
+in the Software without restriction, including without limitation the rights
+to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+copies of the Software, and to permit persons to whom the Software is
+furnished to do so, subject to the following conditions:
+
+The above copyright notice and this permission notice shall be included in all
+copies or substantial portions of the Software.
+
+THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+SOFTWARE.
+"""
+
+usage = '''arguments: dp pmax T FLUID
+  pressures are in bar, T in Kelvin
+
+This program downloads data from the (NIST) Thermophysical properties of fluid
+system database. This data can then be read using python with pandas or the
+users program of choice.
+
+Typical usage:
+    python3 isothermal-save-gas-csv.py 0.1 2000 298 xenon
+    # The above only needs to be done once
+    python3 isothermal-ads-gas-plot.py xenon 298 5.8 200
+'''
+
+# --- Command line arguments --- #
+if len(sys.argv) != 5:
+    print(usage)
+    exit(1)
+pInc = float(sys.argv[1])
+pMax = float(sys.argv[2])
+Temperature = float(sys.argv[3])
+fluid = sys.argv[4]
+
[…]

### 2020-05-01 gas-adsorption/old_storage/IsothermCalculate.py

@@ -1,0 +1,82 @@
+from __future__ import division, print_function
+import sys, os, re, matplotlib
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits
+
+import colors
+
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+
+# All data comes from DOI: 10.1021/acsenergylett.8b00154
+v_coefs = {
+    'HKUST1': {
+        'a': [-784.62, 11.334, 0.64936, -0.066476, 0.003472, -0.00014164, 3.71e-6],
+        'b': [5.6102, -0.039763, 0.0071242, -0.00095632, 6.27e-5, -1.06e-6, -1.88e-8]
+    },
+    'NOTT112': {
+        'a': [-613.79, 13.013, -0.21045, -0.0071214, 0.00048134, -1.0052e-5, 2.4781e-7, -5.3734e-9],
+        'b': [4.431, -0.049712, 0.0040162, -0.00019197, 5.7258e-6, -6.8488e-8, -2.1531e-9, 7.1793e-11]
+    },
+    'NU125': {
+        'a': [-593.6, -0.029702, 0.50094, -0.0088905, -0.00056209, 1.9733e-5, -1.1469e-7, 1.5171e-9],
+        'b': [4.3146, 0.0030633, 0.0036414, -0.00044774, 2.0118e-5, -2.5e-7, -2.7021e-9, 2.3075e-11]
+    },
+    'rhtMOF7': {
+        'a': [-712.95, -1.2764, 1.8356, -0.054302, -0.0026518, 0.00013347],
+        'b': [5.1841, 0.055013, -0.00070978, -0.00092537, 8.6874e-5, -2.2699e-6]
+    },
+    'CuMOF74': {
+        'a': [-703.34, -4.9553, 2.4354, -0.01936, -0.0034717, -8.9968e-5],
+        'b': [5.513, 0.05651, 0.015549, -0.0036854, 0.00022566, -1.1167e-6]
+    },
+    'PCN250': {
+        'a': [-792.52, 27.244, 0.092688, -0.49499, 0.048062, -0.001344],
+        'b': [5.4822, -0.056556, -0.021315, 0.0074926, -0.00064637, 1.7667e-5]
+    },
+}
+
+
+
+
+IsothermT = [77,298]
+x = np.arange(0.001,50,0.001)
+
+for mof in sorted(v_coefs):
+    a_coef = 0
[…]

### 2020-05-01 gas-adsorption/old_storage/adsorbed-gas-isosteric-data.py

@@ -1,0 +1,233 @@
+from __future__ import division, print_function
+import sys, os, re, matplotlib
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits
+import math as mh
+import collections
+# Data from the work:  Benchmark Study of Hydrogen Storage in
+# Metal−Organic Frameworks under Temperature and Pressure Swing Conditions
+
+# HOW TO RUN:
+# python3 adsorbed-gas-isosteric-data.py MOF5 absolute methane 300
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+
+prop_cycle = plt.rcParams['axes.prop_cycle']
+colors = prop_cycle.by_key()['color']
+
+mof = sys.argv[1]
+ads_type = sys.argv[2]
+basename = sys.argv[3]
+my_isostere = int(sys.argv[4])
+
+def my_interp(x0,x,y):
+    return np.interp(x0, x, y)
+
+def mu_from_n(particular_n):
+    return my_interp(particular_n, n, mu)
+
+def n_from_mu(particular_mu):
+    return my_interp(particular_mu, mu, n)
+
+def mu_from_p(particular_p):
+    return my_interp(particular_p, p, mu)
+
+data = np.loadtxt(mof + ('-%s-%s.txt' % (ads_type,basename)), skiprows=3,unpack=True,delimiter=',')
+
+# define the temperature range for the papers data
+T_label = []
+Temp_str = np.loadtxt(mof + ('-%s-%s.txt' % (ads_type,basename)),dtype=str,unpack=True,delimiter=',')
+for i in range(len(Temp_str)):
+    if Temp_str[i][0]:
+        T_label.append(int(Temp_str[i][0]))
+
+# ---------------- Single-site Langmuir Fit (from reference) ------------------#
+Lang_rho = []
+Lang_p = []
+Lang_T = []
+
+Lang_T_2d, Lang_p_2d = np.meshgrid(T_label, np.arange(0, 100.1, 0.1))
+Lang_rho_2d = np.zeros_like(Lang_p_2d)
+Lang_rho_2d[:,:] = np.nan
+axis_value = 0
+# Table S3 Langmuir fitting parameters (MOF-5):
+nsat1 = 30.5*22.414 # mmol/g*22.414 mL/mol --> cm3 (STP)/g
[…]

### 2020-05-01 gas-adsorption/old_storage/adsorbed-gas-json.py

@@ -1,0 +1,159 @@
+from __future__ import division, print_function
+import sys, os, re, matplotlib, glob
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits
+import json
+
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+# --- Build unit system --- #
+
+Kelvin = numericalunits.K
+atm = numericalunits.atm
+bar = numericalunits.bar
+gram = numericalunits.g
+kg = numericalunits.kg
+L = numericalunits.L
+Joule = numericalunits.J
+mol = numericalunits.mol
+mmol = numericalunits.mmol
+psi = numericalunits.psi
+
+mg = numericalunits.mg    # milligram
+mL = numericalunits.mL    # milliliter
+kJ = numericalunits.kJ     # kiloJoule
+meter = numericalunits.m
+cm = numericalunits.cm
+wt = 1
+
+mol_per_L_STP = 22.414*mol/L
+
+# define H2 and CH4 constants
+molar_mass_H2 = 2.016*gram/mol
+molar_mass_CH4 = 16.04*gram/mol
+
+rho_H2 = 8.988e-5*gram/(cm**3)
+rho_CH4 = 0.000656*gram/(cm**3)
+
+def molar_mass(json):
+    masses = {
+    'Hydrogen': 2.016*gram/mol,
+    'H2': 2.016*gram/mol,
+    'CH4': 16.04*gram/mol,
+    'Methane': 16.04*gram/mol,
+    }
+    return masses[json['adsorbates'][0]['name']]
+
+def crystal_density(json):
+    densities = {
+    'IRMOF-1': 0.61*gram/cm**3, # Furuta, Kanoya, Sakai, and Hosoe
+    'CuPDC': 1.196*gram/cm**3, # Furuta, Kanoya, Sakai, and Hosoe
+
+    #DOI:10.1021/ic201376t
+    # Synonyms Basolite C300, C300, Cu-BTC, Cu3(BTC)2, HKUST-1, MOF-199
+    'HKUST-1': 0.88*gram/cm**3,
+    'PCN-6-Prime': 0.28*gram/cm**3, # a prime version of PCN-6?
+    'PCN-6': 0.56*gram/cm**3,
+    'PCN-20': 0.49*gram/cm**3,
+    'MOF-399': 0.13*gram/cm**3,
[…]

### 2020-05-01 gas-adsorption/old_storage/adsorbed-gas-read.py

@@ -1,0 +1,724 @@
+from __future__ import division, print_function
+import sys, os, re, matplotlib
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits
+
+matplotlib.rcParams['text.usetex'] = True
+
+import colors
+
+from matplotlib.font_manager import FontProperties
+
+small_font = FontProperties()
+small_font.set_size('small')
+
+# import python json data extract package
+ads_json = __import__('adsorbed-gas-json')
+
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+# --- Build unit system --- #
+
+Kelvin = numericalunits.K
+atm = numericalunits.atm
+bar = numericalunits.bar
+gram = numericalunits.g
+kg = numericalunits.kg
+L = numericalunits.L
+Joule = numericalunits.J
+mol = numericalunits.mol
+psi = numericalunits.psi
+cm = numericalunits.cm
+mmol = numericalunits.mmol
+
+mg = numericalunits.mg    # milligram
+mL = numericalunits.mL    # milliliter
+kJ = numericalunits.kJ     # kiloJoule
+
+basename = sys.argv[1]
+basename_at_dif_T = sys.argv[2]
+p_empty = float(sys.argv[3])*bar
+p_full = float(sys.argv[4])*bar
+
+
+molecular_weight = 2.016*gram/mol
+if 'methane' in basename:
+    molecular_weight = 16.04*gram/mol
+
+kB = 8.314*Joule/mol/Kelvin
+data = [np.loadtxt('%s.csv' % basename, skiprows=1),
+             np.loadtxt('%s.csv' % basename_at_dif_T, skiprows=1)]
+
+T = data[0][:,0]*Kelvin            # Temperature
+#print(np.asarray(T).shape)
+p = data[0][:,1]*bar               # Pressure
+rho = data[0][:,2]*mol/L           # Density
+V = data[0][:,3]*L/mol             # Volume
+U = data[0][:,4]*kJ/mol            # Internal Energy
[…]

### 2020-05-01 gas-adsorption/old_storage/adsorbed-gas.py

@@ -1,0 +1,184 @@
+from __future__ import division, print_function
+import sys, os, re
+import numpy as np
+import matplotlib.pyplot as plt
+
+import pint
+ureg = pint.UnitRegistry()
+import pandas as pd # import data tables from NIST and store with pandas
+
+# --- Function to find nearest value in array and return index and element --- #
+
+def find_nearest(array, value):
+    #array = np.asarray(array)
+    maxval = array.max()
+    #print('type', type(maxval), maxval)
+    #print(value, array.max(), value < array.max(), value < maxval)
+    #print('difference', array.max() - value)
+    assert(value < array.max())
+    assert(value >= array.min())
+    idx = (np.abs(array - value)).argmin()
+    return idx, array[idx]
+
+# --- Build unit system --- #
+
+Kelvin = ureg.kelvin
+atm = ureg.atm
+gram = ureg.g
+L = ureg.L
+Joule = ureg.J
+mol = ureg.mol
+psi = ureg.psi
+
+mg = ureg.mg    # milligram
+mL = ureg.mL    # milliliter
+kJ = ureg.kJ     # kiloJoule
+
+# --- Command line arguments --- #
+
+if len(sys.argv) != 7:
+    print('arguments: plo pFilled pmax pinc T FLUID\n  pressures are in atm (approximately bar), T in Kelvin')
+    exit(1)
+pLow = sys.argv[1]
+# Final Pressure and Fixed Volume
+pFilled = float(sys.argv[2])*atm
+pMax = sys.argv[3]
+pInc = sys.argv[4]
+Temperature = sys.argv[5]
+fluid = sys.argv[6]
+
+basename = '{}-{}-{}-{}'.format(fluid, pLow, sys.argv[2], Temperature)
+
+MM = 16.043 * gram/mol     # molar mass of Methane
+NA = 6.02214e23 * 1/mol   # Avogadro's constant
+kB = 1.38064852e-23 * Joule/Kelvin
+
+# --- Read NIST tables using Pandas --- #
+
+methane_id = 'C74828'
+hydrogen_id = 'C1333740'
+nist_ids = {
+'methane': 'C74828',
+'hydrogen': 'C1333740',
[…]

### 2020-05-01 gas-adsorption/old_storage/new-adsorbed-gas-isosteric.py

@@ -1,0 +1,145 @@
+from __future__ import division, print_function
+import sys, os, re, matplotlib
+import numpy as np
+import matplotlib.pyplot as plt
+import numericalunits
+import math as mh
+import collections
+# Data from the work:  Benchmark Study of Hydrogen Storage in
+# Metal−Organic Frameworks under Temperature and Pressure Swing Conditions
+
+# HOW TO RUN:
+# python3 new-adsorbed-gas-isosteric.py MOF5 absolute methane
+matplotlib.rcParams['text.usetex'] = True
+matplotlib.rc('font', family='serif')
+
+# TO DO:
+
+# 1. Add implementation that does not assume an ideal gas for methane.
+#    (Requires NIST data.)
+
+# 3. Attempt a Gibbs free energy analysis.
+
+prop_cycle = plt.rcParams['axes.prop_cycle']
+colors = prop_cycle.by_key()['color']
+
+mof = sys.argv[1]
+ads_type = sys.argv[2]
+basename = sys.argv[3]
+
+def my_interp(x0,x,y):
+    return np.interp(x0, x, y)
+
+data = np.loadtxt(mof + ('-%s-%s.txt' % (ads_type,basename)), skiprows=2,unpack=True,delimiter=',')
+
+# define the temperature range for the papers data
+T_label = []
+Temp_str = np.loadtxt(mof + ('-%s-%s.txt' % (ads_type,basename)),dtype=str,unpack=True,delimiter=',')
+print(mof + ('-%s-%s.txt' % (ads_type,basename)))
+for i in range(len(Temp_str)):
+    if Temp_str[i][0]:
+        T_label.append(int(Temp_str[i][0]))
+T_raw = np.array(T_label)
+
+#print(data)
+
+pressure = data[::2,:]
+rho = data[1::2,:]
+T = np.zeros_like(pressure)
+for i in range(len(T_raw)):
+    T[i,:] = T_raw[i]
+
+for i in range(T.shape[0]):
+    plt.plot(pressure[i,:], rho[i,:], '.-', label='{}$^\circ C$'.format(T[i,0]))
+plt.legend(loc='best')
+plt.xlabel('$p$')
+plt.ylabel(r'$\rho$')
+
+plt.figure('rho vs T at fixed p')
+
[…]

### 2020-05-01 gas-adsorption/old_storage/nist.py

@@ -1,0 +1,66 @@
+from __future__ import division, print_function
+import sys, os, re
+import numpy as np
+import matplotlib.pyplot as plt
+import json
+
+import pandas as pd # import data tables from NIST and store with pandas
+
+temperature_pressure_dict {
+    (50, 1.2): {
+    "density": 5.4,
+    "enthalpy": 4.4,
+    }
+}
+isothermal_dict = {
+    "50": { #store each temperature for an isotherm.
+        "Pressure":     [[1,2,3],'kg'],
+        "Density":      [[1,2,3],'kg'],
+        "Volume":       [[1,2,3],'kg'],
+        "Energy":       [[1,2,3],'kg'],
+        "Enthalpy":     [[1,2,3],'kg'],
+        "Entropy":      [[1,2,3],'kg'],
+        "Gibbs":        [[1,2,3],'kg'],
+
+    },
+    "50.1": { #store each temperature for an isotherm.
+        "Pressure":     [[1,2,3],'kg'],
+
+    },
+}
+
+# --- Command line arguments --- #
+
+with open('data.json', 'w', encoding='utf-8') as file:
+     json.dump(isothermal_dict, file, sort_keys=True, ensure_ascii=False, indent=4)
+
+with open('data.json') as file:
+    data = json.load(file)
+print(data)
+print(sys.getsizeof(data)/1000.0, 'kB of Ram')
+html = 'https://webbook.nist.gov/cgi/fluid.cgi?Action=Load&ID=C74828&Type=IsoTherm&Digits=5&PLow=5&PHigh=100&PInc=.1&T=300&RefState=DEF&TUnit=K&PUnit=bar&DUnit=mol%2Fl&HUnit=kJ%2Fmol&WUnit=m%2Fs&VisUnit=uPa*s&STUnit=N%2Fm'
+#
+tables = pd.read_html(html)[0]
+#
+Pressure = (pd.DataFrame(tables[1:],columns=[1])).values.astype(np.float)
+#print(np.core.defchararray.strip(Pressure),'[')
+print(Pressure.tolist())
+# print(type(tables))
+print(type(Pressure))
+
+
+def find_density(p, T, fluid):
+    ''' Cache data and only download if we don't have it already '''
+    pass
+
+def find_isotherm_internal_energy(pmin, pmax, T, fluid):
+    ''' Cache data and only download if we don't have it already '''
+    pass
+
+def find_isotherm_entropy(pmin, pmax, T, fluid):
[…]

### 2020-05-01 ising/ising.tex

@@
 stochastic approximation whose convergence could be mathematically
-proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
-al.~\cite{liang2007stochastic} argued that WL can be considered a form of
+proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang \emph{et
+al.}~\cite{liang2007stochastic} argued that WL can be considered a form of
 stochastic approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
@@
 
-Pommerenck et al. propose a refinement~\cite{pommerenck2020stochastic} to SAMC
+Pommerenck \emph{et al.} propose a refinement~\cite{pommerenck2020stochastic} to SAMC
 where the update factor is determined dynamically rather than by the user.
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N32-Cv.pdf}
-    \caption{
-    The specific heat capacity versus the reciprocal temperature $\beta$ for the $N=32 \times 32$ system for each histogram method at $10^{9}$ moves.}
-    \label{fig:N32-cv}
-  \end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
-  \caption{The maximum error in the specific heat capacity for each method for the $N=32 \times 32$ and $T_{\min} = 1$ as a function of number of iterations run.
-The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
-  \label{fig:N32-cv-error}
-\end{figure}
-
 % \begin{figure}
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{N32-Cv.pdf}
+  \caption{
+    The specific heat capacity versus the reciprocal temperature $\beta$ for the $N=32 \times 32$ system for each histogram method at $10^{9}$ moves.}
+  \label{fig:N32-cv}
+\end{figure}
+
 \subsection{The 32 $\times$ 32 Ising model}
-Fig.~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
+Figure~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
 All of the update factors initially start at $\gamma_{t=0} = 1$.
@@
 
-Fig.~\ref{fig:N32-cv} shows the specific heat capacity vs.
+All the methods except for SAMC gave a reasonable approximation for
+the heat capacity peak resulting from the phase transition within 10$^9$ moves.
+Figure~\ref{fig:N32-cv} shows the specific heat capacity vs.
 the reciprocal temperature $\beta$ at $10^{9}$ moves.
@@
 
-Fig.~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
+\begin{figure}
+  \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
+  \caption{The maximum error in the specific heat capacity for each method for the $N=32 \times 32$ and $T_{\min} = 1$ as a function of number of iterations run.
+The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
+  \label{fig:N32-cv-error}
+\end{figure}
+
+Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
 this system. The solid/dashed lines

### 2020-05-01 ising/ising.tex
a few rewordings, reorder a figure, and some davidsays

@@
 where $g(E)$ is the multiplicity of the system which is proportional to
-$D(E)$. We can estimate the maximum deviation in the canonical specific heat capacity $c_V$ from the exact solution~\cite{schneider2017convergence, shakirov2018convergence,
+$D(E)$. We can compute the maximum deviation in the canonical specific heat capacity $c_V$ from the exact solution~\cite{schneider2017convergence, shakirov2018convergence,
 barash2017control,barash2017gpu}:
 \begin{align}
-\braket{\epsilon(t)}_E = \left|\max_E \left(c_V(E,S,T,t) - c_V(E,S,T,t)_{\text{Beale}}\right)\right|
+\epsilon \equiv \max_T \left|\left(c_V(T)_{\text{MC}} - c_V(T)_{\text{Beale}}\right)\right|
 \end{align}
@@
 Computing the specific heat capacity $c_V$ presents a difficult challenge for
-any Monte Carlo method due to fluctuations in the derivative of the internal energy.  Methods that accurately compute $c_V$ also by extension accurately compute the internal energy.
+any Monte Carlo method due to fluctuations in the derivative of the
+internal energy.  Methods that accurately compute $c_V$ also by
+extension accurately compute the internal energy.
+
+\davidsays{talk about phase transition, e.g. at what temperature?}
 
@@
 of energy space. Flat-histogram methods differ primarily in how they schedule
-the decrease of $\gamma_t$.
+the decrease of $\gamma_t$.  Figure~\ref{fig:N32-gamma} shows ...
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{gamma-n32.pdf}
+  \caption{
+  The update factor $\gamma_t$ versus the iteration number for the $N=32 \times 32$
+  system.}
+  \label{fig:N32-gamma}
+\end{figure}
 
 The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
-landau2014guide} explores energy space by setting $\gamma_{t=0}^{\text{WL}}=1$,and then decreases $\gamma^{\text{WL}}$ in prescribed stages. An energy range of
+landau2014guide} explores energy space by setting
+$\gamma_{t=0}^{\text{WL}}=1$, and then decreases $\gamma^{\text{WL}}$ in prescribed stages. An energy range of
 interest must be specified~\cite{wang2001efficient, schulz2003avoiding,yan2003fast}, which often requires multiple simulations if unknown.
-
+%
 The number (``counts'') of moves ending at each energy are stored in a
@@
 
-The WL method can alternatively be implemented as part of a production run which aids in
-equilibration~\cite{gross2018massively}. WL is used to generate the weights
-resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
-minimum $\gamma^{\text{WL}}_t$ is reached, the simulation switches to a final production run
-with fixed weights. WL implemented in this way can ensure ergodicity and
-detailed balance; however, the convergence of simulation is still impacted by
-the choice of the minimum $\gamma^{\min}_t$.
+The WL method can be terminated after $\gamma$ reaches a specified
+minimum $\gamma^{\min}$ followed by a \emph{production run} in which the
+weights are held fixed~\cite{gross2018massively}. WL is used to
+generate the weights resulting in a
+flat-histogram~\cite{janke2017generalized}.  The entropy is then
+computed (up to a constant) by adding the logarithm of the production
+histogram to the logarithm of the weights.  The production run thus
+satisfies detailed balance, and will ideally be ergodic; however, the
+convergence of the simulation is still impacted by the choice of the
+minimum $\gamma^{\min}$.
 
@@
 
-\begin{figure}
-\includegraphics[width=\columnwidth]{gamma-n32.pdf}
-  \caption{
-  The update factor $\gamma_t$ versus the iteration number for the $N=32 \times 32$
-  system.}
-  \label{fig:N32-gamma}
-\end{figure}
-
 % \begin{figure}
@@
 \subsection{The 32 $\times$ 32 Ising model}
-Figure~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
-All of the update factors initially start at $\gamma_{t=0} = 1$.
-SAD dynamically updates $\gamma_t$ throughout the simulation.
-After about $10^{10}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
-Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown.
-The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$.
-The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
-All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.
+Figure~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each
+of the flat-histogram methods. All of the update factors initially
+start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$
+throughout the simulation. After about $10^{10}$ moves,
+$\gamma^{\text{SAD}}_t$ proceeds as $1/t$. We show
+$\gamma^{\text{WL}}_t$ for WL both with and without a production
+run. The WL production run begins after $\gamma^{\text{WL}}_t$ has reached
+$10^{-4}$. The update factor for $1/t$-WL decreases similarly to WL
+before finding all the energy states and switching to $1/t$. All of
+the SAMC update factors equal 1 until the number of moves is equal to
+$t_0$ at which point they decrease as $1/t$.
 
@@
 \subsection{128 $\times$ 128 Ising system}
-Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
-All of the update factors initially start at $\gamma_{t=0} = 1$.
-SAD dynamically updates $\gamma_t$ throughout the simulation.
-After about $10^{13}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
+Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each
+of the flat-histogram methods. All of the update factors initially
+start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$
+throughout the simulation. After about $10^{13}$ moves,
+$\gamma^{\text{SAD}}_t$ proceeds as $1/t$. \davidsays{fixme similarly}
 Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown.
@@
 
-We find that both SAD, 1/t-WL, and production WL (provided an ideal minimum update factor before beginning the production run) demonstrate excellent and robust convergence.
-They all converge more rapidly than SAMC, and unlike WL do not suffer from
+We find that both SAD, 1/t-WL, and WL followed by a production run (provided an ideal minimum update factor before beginning the production run) demonstrate excellent and robust convergence.
+They all converge more rapidly than SAMC, and unlike pure WL do not suffer from
 error saturation. We find that for larger systems SAD reduces the updated factor
-more slowly (and conservatively) than 1/t-WL and production WL. This means that SAD will take
+more slowly (and conservatively) than 1/t-WL and WL followed by a
+production run. This means that SAD will take
 proportionately more moves to converge to the same value as 1/t-WL as system
@@
 For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
-considerably more convenient. This work also visually demonstrates that ``production'' WL performs extremely well and is preferable to WL for ensuring both ergodicity and detailed balance.
+considerably more convenient. This work also visually demonstrates
+that WL followed by a production run performs extremely well and is preferable to WL for ensuring both ergodicity and detailed balance.
 

### 2020-05-02 ising/ising.tex

@@
 
-Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
+\davidsays{Explain the range of temperature compared, and why} Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
 this system. The solid/dashed lines
@@
 \section{Conclusion}
+\davidsays{remove this?}
 The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter $\gamma_{t}^{\min}$ makes the method somewhat more difficult to implement in practice.
 
-We find that both SAD, 1/t-WL, and WL followed by a production run (provided an ideal minimum update factor before beginning the production run) demonstrate excellent and robust convergence.
+We find that both SAD, 1/t-WL, and WL followed by a production run
+(with an adequately small $\gamma^{\min}$) demonstrate excellent and robust convergence.
 They all converge more rapidly than SAMC, and unlike pure WL do not suffer from

### 2020-05-02 gas-adsorption/gas-ads-body.tex
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

--- 2020-05-01 gas-adsorption/main.tex
@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp]{revtex4-1}
-\usepackage{graphicx} % Include figure files
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{subfig}
-
-\usepackage{nicefrac} % Include for inline fractions
-\usepackage{url}
-
-% \usepackage{cite}
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage{color}  % Coloured text etc.
-%
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
-\usepackage{mdframed}
-
-\newcommand{\addcite}{{\bf \color{red} []}}
-\newcommand{\xvec}{\mathbf{x}}
-\newcommand{\thetavec}{\boldsymbol \theta}
-
-\newcommand\V{\Phi}
-\newcommand\Vk{\tilde\Phi(\vec k)}
-\newcommand\volume{V}
-
-\newcommand{\pink}[1]{{\bf \color{pink} #1}}
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\orange}[1]{{\bf \color{orange} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
-
-%\newcommandx{\davidsays}[1]{\begin{mdframed}[backgroundcolor=blue!24]\textbf{David says:} #1\end{mdframed} }
-%\newcommandx{\davidsays}[2][1=inline]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{David says:} #2} }
-\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
-\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
-\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }
-
-%% \renewcommand\davidsays[1]{}
-%% \renewcommand{\corysays}[2][1=inline]{}
-%TC:macro \corysays [ignore]
-%TC:macro \davidsays [ignore]
-%TC:macro \jordan [ignore]
-
-\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }
-
-\newtheorem{limitation}{Limitation}
-
-\begin{document}
-\title{
-An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}
-% other ideas:
-%Gas delivery via pressure-swing adsorption in nanoporous materials: setting storage targets}
-% A thermodynamic upper bound to gas delivery via pressure-swing adsorption in nanoporous materials:}
-
-\author{Jordan K. Pommerenck}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-\author{Cory M. Simon}
-\affiliation{School of Chemical, Biological, and Environmental Engineering,
-  Corvallis, OR 97331}
-\author{David Roundy}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-
-\begin{abstract}
-  Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles [].
-One strategy to densify these gases is to pack the fuel tank with a porous material [].
-  Metal-organic frameworks are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases [].
-The US Department of Energy (DOE) has set volumetric deliverable capacity targets [] which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels.
-Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction []; none have established a rigorous upper bound on the deliverable capacity.
-  Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing.
-% "pressure-swing storage doesn't make sense to me.
-storage and delivery  maybe but awk.
-  % I added rigid since that could be important...
-  To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
-  \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
-  Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
-  We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-  % However, this upper bound directly leads to important realizations which should guide future development.
-  However, limitations to the scope of applicability of our upper bound guide future material development efforts.
- % Firstly, one could extract the gas at a higher temperature than that used while filling [].
-  Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
-  Secondly, the fundamental physics of our upper bound do not pertain to
-  % rule out
-  any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
-  % firstly, secondly vs first, second?
-\end{abstract}
-
-\newcommand\pfull{\ensuremath{p_{\text{full}}}}
-\newcommand\pempty{\ensuremath{p_{\text{empty}}}}
-
-\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
-\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}
-
-\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
-\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}
-
-\maketitle
-
-\newcommand\gst{\ensuremath{\Delta g_{st}}}
-
-\paragraph*{to do}
-\begin{enumerate}
-    \item See david/corysays
-    \item make the code presentable
-    \item do the FIXME in code
-\end{enumerate}
-
 \section{Introduction}
@@
     \subfloat[][]{
-        \includegraphics[width=0.6\columnwidth]{figs/hkust-1_fancy.png} \label{fig:example_MOF}
+        \includegraphics[width=0.6\columnwidth]{hkust-1_fancy.png} \label{fig:example_MOF}
     }
@@
     \subfloat[][]{
-        \includegraphics[width=0.98\columnwidth]{figs/usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
+        \includegraphics[width=0.98\columnwidth]{usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
     }
@@
   \centering
-  \includegraphics[width=\columnwidth]{figs/methane-298-rho-mu}
+  \includegraphics[width=\columnwidth]{methane-298-rho-mu}
   \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several MOFs at 298\ K as a function of chemical potential (bottom axis) and pressure (top axis).
@@
     \centering
-    \includegraphics[width=0.95\columnwidth]{figs/methane-298-n-vs-G}
+    \includegraphics[width=0.95\columnwidth]{methane-298-n-vs-G}
     \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
@@
     \centering
-    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-298-n-vs-G}
+    \includegraphics[width=0.95\columnwidth]{hydrogen-298-n-vs-G}
     \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
@@
 % wut about mixtures?
-
-%TC:ignore
-
-\bibliography{bibfile}
-
-\appendix
-\clearpage
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}%
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}%
-\renewcommand{\thesection}{SI \Alph{section}}%
-
-\section{An external, spatially uniform potential $\V$ shifts the chemical potential $\mu$ of the gas} \label{sec:V_shifts_chem_pot}
-We show that imposing an external, spatially uniform potential $\V$ to a gas has the effect of shifting the chemical potential $\mu$ of the gas, recovering eqn.~\ref{eq:mof-density}.
-Consider a control volume $\Omega$ with volume $V=|\Omega|$ (large enough to neglect boundary effects) that is endowed with the external, spatially uniform potential $\V$.
-Impose the grand-canonical ensemble, where this control volume can exchange energy and particles with a bath of gas at temperature $T$ and chemical potential $\mu$.
-
-To denote a microstate of this system, let $N$ be the number of gas particles in the control volume and $\mathbf{x}_1,...,\mathbf{x}_N$ be their positions.
-Then, the potential energy $E$ of a microstate of the control volume is:
-\begin{equation}
-    E(\mathbf{x}_1,...,\mathbf{x}_N) = N\V + E_{gg}(\mathbf{x}_1,...,\mathbf{x}_N),
-\end{equation} where $E_{gg}$ is the (unknown and complicated) interatomic potential for gas-gas interactions that governs the (real) gas properties.
-The first term arises from each gas molecule feeling the external potential $\V$, where $\V < 0$ corresponds to attraction.
-
-\corysays{wut about orientations? shld we include those?}
-
-The grand-canonical partition function of the control volume is:
-\begin{multline}
-    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\xvec_1, ..., \xvec_N)} e^{\beta (\mu - \V) N} d\xvec_1 \cdots d\xvec_N.
-    \label{eq:gcpf}
-\end{multline} We recognize this as the grand canonical partition function of the bulk gas in the control volume without the external potential, but shifted in chemical potential:
-\begin{equation}
-    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
-    \label{eq:xi_vs_xi0}
-\end{equation}
-where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the bulk gas in the control volume in the absence of an external potential.
-Importantly, this equivalency is predicated on the potential $\V$ being spatially uniform.
-Therefore, the thermodynamic properties of the gas sitting in the spatially uniform external potential $\V$-- i.e., adsorbed in our idealized substrate-- are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$ ($T$ fixed here).
-As $\V$ becomes more negative, corresponding to a more attractive adsorbent, the thermodynamic properties of the adsorbed gas in our ideal substrate are equivalent to the gas at higher chemical potential.
-
-\section{Proof of extremum}
-To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.
-In this proof we will make use of the Fourier transform of this potential:
-\begin{align}
-    \Vk \equiv \iiint \V(\vec r) e^{-i\vec k\cdot \vec r} d\vec r
-\end{align}
-The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$.
-Therefore, to show that a uniform potential extremizes the deliverable capacity, we must show that the functional derivative of the deliverable capacity with respect to $\Vk$ is zero for \emph{nonzero} values of $\vec k$, i.e.
-\begin{align}
-    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \vec k\ne 0.
-\end{align}
-We note that this functional derivative may be non-zero for $\vec k=0$ because we separately maximize with respect to the particular uniform potential $\V$.
-This means that
-\begin{align}
-    \frac{\delta N_H}{\delta \V(\vec k)} &= -\frac{\delta N_L}{\delta \Vk}
-\end{align}
-where $N_H$ and $N_L$ are the number of particles at the low and high pressure.
-
-Because the chemical potential $\mu$ varies monotonically with $N$ at fixed temperature, we can consider how the chemical potential varies as we change $\Vk$ with the number of molecules held fixed.
-We demonstrate this using the cyclic chain rule, which shows us that
-\begin{align}
-    \left(\frac{\delta N}{\delta \Vk}\right)_{\mu} &=
-    -\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}
-    \left(\frac{\partial N}{\partial \mu}\right)_{\Vk}.
-\end{align}
-Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \V(\vec k)}\right)_{\mu}=0$.
-Thus we consider
-\begin{align}
-    \left(\frac{\delta \mu}{\delta \tilde\V(\vec k)}\right)_N
-    &= \left(\frac{\delta \left(\frac{\partial F}{\partial N}\right)_{\volume}}{\delta \Vk}\right)_N
-    \\
-    &= \left(\frac{\partial \left(\frac{\delta F}{\delta \Vk}\right)_{\volume}}{\partial N}\right)_{\volume}
-    \label{eq:dmudpot}
-\end{align}
-where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reordered the functional and partial derivatives.
-Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ just gives the number density:
-\begin{align}
-    \frac{\delta F}{\delta \Vk} &= \rho(\vec k)
-\end{align}
-The number density is itself spatially uniform for any system that is stable in a fluid state at this density (i.e.
-does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.
-Thus the functional derivative $\frac{\delta F}{\delta \V(\vec r)}$ is actually a uniform function.
-We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
-\begin{align}
-    \frac{\delta \mu}{\delta \Vk} &\propto \delta(\vec k) \\
-    \frac{\delta N}{\delta \Vk} &\propto \delta(\vec k)
-\end{align}
-Thus the functional derivative of both the chemical potential and $N$ with regard to $\V(\vec r)$ are themselves spatially uniform.
-Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e.
-$\vec k=0$), the derivative of $D$ with respect to any change of potential is zero.
-
-This demonstrates that a spatially uniform potential leads to an extremum value of the deliverable capacity.  This proof is insufficient, however, to show that it must be a true maximum.
-
-\section{The real-substrate analogy of $\V$ is $\gst$}
-\label{sec:phi-is-delta-g}
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figs/g_st_piston_2.png}
-    \caption{Cartoon illustrating the definition of $\gst$ and a hypothetical experiment to measure it with a piston and removable barrier that separates two equal volumes, one with a porous material (bottom) and one with free space (top).
-The initial state (left) is a bulk gas phase at density $\rho$ and temperature $T$.
-An infinitely-thin, removable barrier separates the gas from the porous material (bottom).
-The barrier is then slowly removed, and the piston slowly pushes all gas molecules into the porous material in an isothermal process.
-The final state (right) is an adsorbed gas phase at density $\rho$ and temperature $T$.
-The quantity $\gst$ follows from {\color{red} the heat released?}.}
-    \label{fig:delta-G-cartoon}
-\end{figure}
-
-The parameter describing our idealized substrate is $\V$, the spatially uniform potential felt by a gas molecule adsorbed in the idealized substrate. 
-A natural question is how this potential relates to the properties of real substrates. 
-The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the gas (see Sec.~\ref{sec:V_shifts_chem_pot}).
-Because our ideal substrate shifts the chemical potential of the gas molecules by providing a spatially uniform potential energy field, the entropy of the gas in the ideal substrate is equal to the entropy of the gas in its bulk state at the same density and temperature.
-In contrast, a real substrate provides a non-spatially uniform potential.
-Consequently, the entropy of the gas inside a real substrate is \emph{not} equal to the entropy of the bulk gas at the same density and temperature.
-Therefore, the parameter analogous to $\V$ in a real substrate will involve both energy and entropy.
-The real-substrate analogy of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
-\begin{equation}
-   \gst(\rho, T) \equiv
-   %\frac{G_{\text{ads}}(\rho, T) - G_{\text{gas}}(\rho, T)}{N}\\ 
-    \mu_{\text{ads}}(\rho, T) - \mu_{\text{gas}}(\rho, T).
-  % \gst(\rho, T) &\equiv g_{\text{gas}}(\rho, T) - g_{\text{ads}}(\rho, T)
-  \label{eq:g_st}
-\end{equation}
-The isosteric Gibbs free energy difference $\gst$ is the difference in molar Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and the bulk gas \emph{with the same density of gas molecules}.
-The quantity $\gst$ does \emph{not} correspond to a change in the molar Gibbs free energy as a molecule is adsorbed, which is zero under conditions of coexistence.
-The quantity $\gst$ in a real substrate is a direct analogy to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
-Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to  measure $\gst$ via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
-Note that $\gst$ is a property of both the substrate and the identity of the gas.
-Because real substrates offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho, T)=\V$.
-Consequently, throughout this article, we show $\gst(\rho, T)$ for real substrates at both conditions relevant to gas storage and delivery, $\pfull$ and $\pempty$.
-% In the initial state, the partition isolates the bulk gas exhibiting density $\rho$ and temperature $T$.
-Then, the partition is removed and the piston pushes down on the gas so that the gas slowly adsorbs into the MOF in an isothermal process.
-\red{isothermal rite?} In the final state, the piston has pushed all gas into the MOF, resulting in an adsorbed gas phase with the same density $\rho$ and temperature $T$ as the initial state.
-The isosteric Gibbs free energy of adsorption, $\gst$, then follows from \red{is it the amount of heat released?}.
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} looks too jumbled. two separate figs? plus $G_{st}$ is capitalized here, should be lowercase. where is all this data from}
-
-In practice, we can readily compute $\gst(\rho, T)$ of a real gas/substrate system from (i) the (experimental or simulated) equilibrium adsorption isotherm of the gas in the substrate and (ii) the (experimental or simulated) chemical potential of the bulk gas.
-Consider the real substrate in thermodynamic equilibrium with a bulk gas at fixed temperature $T$ and pressure $p$, and let $\rho=\rho(p, T)$ be the density of gas in the substrate. 
-At coexistence, the chemical potential of the bulk gas is equal to the chemical potential of the adsorbed gas in the substrate.
-Thus, we can use the experimentally known molar Gibbs free energy of the pure, bulk gas system at temperature $T$ and pressure $p$ to determine the molar Gibbs free energy of the adsorbed system: $\mu_{\text{ads}}(\rho, T)=\mu_{\text{gas}}(p, T)$.
-We can then also look up the known chemical potential of the bulk gas at the same density and temperature as in the substrate, $\mu_{\text{gas}}(\rho, T)$.
-Via eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.
-
-An interesting question is how $\gst$ relates to the commonly measured and reported isosteric heat of adsorption $q_{st}$, which is roughly the energy change when a gas molecule is adsorbed~\cite{sircar1999isosteric, tian2017differential}.
-Figure~\ref{fig:qst-vs-delta-G} shows how $q_{st}$ compares to $\gst$ for several prominent adsorbents.
-In every case, $|q_{st}|>|\gst|$ because the gas in the adsorbent always has less entropy than the gas in the bulk at the same density and temperature.
-That is, while adsorption is energetically favored, it is entropically disfavored due to the restrictions imposed on the configuration of the gas molecules via steric interactions with the substrate itself; this counters the energetic attraction.
-
-\corysays{can u mathematically justify why $\gst$ is the relevant parameter through the grand canonical partition function in eqn.~\ref{eq:gcpf}?
-it is not true that $\Xi(\mu, V, T)$ for the real substrate is equal to $\Xi(\mu+\gst, V, T)$}
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} should be split in two. one for methane one for h2.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/qst-vs-delta-G}
-    \caption{Relationship between $\gst$ and the isosteric heat $q_{st}$ for several prominent adsorbents.
-The dots represent the properties of methane adsorption at 298\ K and at 5\ bar and 100\ bar.
-The $\times$'s represent the properties of hydrogen adsorption at 77\ K and at 5~bar and 100~bar.}
-    \label{fig:qst-vs-delta-G}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/methane-298-gst}
-    \caption{The density-dependence of $\gst$ of methane in several adsorbents (298\ K). Note that $\gst$ is monotonic in $\rho$. Data from Refs.~\ref{mason2014evaluating,furukawa2009storage}.
-    }
-    \label{fig:methane-gst}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/four-cases}
-    \caption{Four cases we consider.}
-    \label{fig:delta-gst-maximum}
-\end{figure}
-
-\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
-Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding to a full and empty density.
-The deliverable capacity of the MOF is equal to the difference between the full and empty density.
-Examining Fig.~\ref{fig:methane-gst}, we see that a significant variety of known experimental $\Delta g_\text{st}$ curves are monotonic.
-Furthermore, our qualitative argument in Section~\ref{sec:upper-bound} suggests that this function \emph{should} be monotonically increasing for rigid substrates, as increasing the density of gas forces some of the gas to reside in higher-energy sites.
-If this is always the case, the deliverable capacity will always be bounded by our theoretical upper bound, even for a value of $\V$ that is not optimal.
-
-To show this, we perform a thought experiment illustrated in Fig.~\ref{fig:delta-gst-maximum}.
-We begin with two pairs of boxes containing gas.
-Two of the boxes are filled with a MOF, and the other two have a uniform potential energy $\V$.
-We could imagine the two boxes at $\V$ as being empty but at very low altitude where gravity provides a potential energy difference (relative to the box with the MOF in it) of $\V$ to each gas molecule (perhaps we are on Jupiter?).
-One box of each sort has a density of gas $\rhofull$ corresponding to $\pfull$ in the MOF system, while the other two boxes has gas density $\rhoempty$.
-
-We now consider what happens if we introduce a diffusive connection between two boxes that are initially at the same density, for instance by connecting them with a tube.
-If the chemical potential of the two boxes are equal, their density will be unchanged, otherwise gas will flow into the box with lower chemical potential.
-The difference in chemical potential is equal to the difference
-\begin{align}
-   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
-   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
-   \\
-   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
-\end{align}
-Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at their initial density after they are connected.
-Thus the deliverable capacity of the MOF will be greater than the deliverable capacity of the model with uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e.
-if $\gst(\rho)$ does \emph{not} monotonically increase.
-By the same token, if we consider the case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable capacity than our model, the MOF must have less residual gas, which means that gas must spontaneously flow from the low-density MOF to the box with potential $\V$, which means that $\gst(\rhoempty)>\V$.
-So once again exceeding our upper bound requires a material with a $\gst(\rho)$ that does not increase monotonically.
-
-Taken together, this indicates that not only is our absolute upper bound an upper bound for rigid MOFs, but the green curve labeled $\rho_D(\gst)$ in Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and \ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal $\gst$.
-
-\section{Cryogenic hydrogen storage}\label{sec:cryo-hydrogen}
-\begin{figure}
-    \centering
-    FIGURE HERE A B C D
-    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-77-n-vs-G}
-    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
-    \label{fig:hydrogen-77-D}
-\end{figure}
-
-One approach to increase the deliverable capacity is to reduce the storage temperature.
-This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling point of nitrogen.
-The DOE ULTIMATE target in this case looks far more achievable, and with a much lower $|\gst|$.
-In fact, an empty tank at this temperature can satisfy the DOE 2020 target.
-The DOE ULTIMATE target is 14\% below the upper bound.
-Actual MOFs fall far short of the theoretical maximum.
-\red{the target is not just a storage density, it specifies temps and pressures too...
-is 77 K allowed?} \davidsays{My recollection of my reading of the DOE target was that anything was allowed, but that the target included ``everything'', which would include both the tank and any cryogenic equipment required.
-So the answer is not so simple.}
-
-%TC:endignore
-
-\end{document}

### 2020-05-02 gas-adsorption/gas-ads.tex
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

--- 2020-05-01 gas-adsorption/main.tex
@@
 \usepackage{graphicx} % Include figure files
+\graphicspath{{figs/}}
 \usepackage{color}
@@
 
-\section{Introduction}
-The transportation sector is dominantly powered by petroleum-based fuels \cite{davis2009transportation}.
-In the US, it is responsible for 36\% of energy-related carbon dioxide emissions \cite{useia} and 50,000 premature deaths per year associated with particulate matter and ozone emissions \cite{caiazzo2013air}.
-Natural gas and hydrogen (H$_2$) are alternative transportation fuels that emit fewer/less pollutants 
-than petroleum-based fuels, and thus their widespread adoption could mitigate climate change 
-\cite{mcglade2015geographical} and improve air quality for human health. Further, as the finite global 
-petroleum resources are declining rapidly \cite{sorrell2010global}, the development of technologies for the widespread
-adoption of sustainable transportation fuels, such as hydrogen, is critical.
-
-Natural gas, mostly methane, is considered a transition (to a renewable and clean) fuel because 
-it emits 25\% less carbon dioxide \cite{eia2013much} and less toxic 
-byproducts \cite{wang2000full} upon combustion per unit energy produced compared to gasoline. From an 
-economic standpoint, the supply of natural gas in the US is increasing as a result of hydraulic 
-fracturing and horizontal drilling techniques \cite{usnatgassupply}. A positive environmental outlook 
-for natural gas, however, is predicated on mitigating fugitive emissions (methane is itself a potent 
-greenhouse gas) \cite{alvarez2012greater} and groundwater contamination \cite{osborn2011methane}
-from hydraulic fracturing.
-
-Hydrogen (H$_2$) is the ultimate transportation fuel because it emits only water when it electrochemically reacts with oxygen in a fuel cell to produce electricity and power a vehicle.
-% Moreover, elemental hydrogen is very abundant, though bonded with oxygen (in water) or carbon (in hydrocarbons); H$_2$ does not occur naturally in abundant quantities.
-Currently, hydrogen (H$_2$) is primarily produced by steam reforming of natural gas followed by the
-water-gas shift reaction, which emits carbon dioxide \cite{crabtree2004hydrogen}. Notably, the 
-environmental allure of hydrogen is predicated on its production via a renewable means, e.g.\ splitting
-water using wind-generated electricity.
-
-% \jordan{I got the values from wikipedia but the citation that wikipedia gave didn't link to the correct paper? Article is Energy Density.
-% THIS PAPER says H2 at 700 bar is 4.7 MJ/L: Review on processing of metal–organic framework (MOF) materials towards system integration for hydrogen storage
-% }
-At ambient conditions, both methane
-%(0.036 MJ/L)
-and hydrogen gas
-%(0.01 MJ/L)
-possess a very low volumetric energy density compared to (liquid) gasoline.
-%(34.2 MJ/L)
-Consequently, under storage space constraints in passenger vehicles, natural gas and hydrogen must be densified for onboard storage to achieve a reasonable driving range on a ``full'' tank of fuel.
-Traditional densification approaches are liquefaction (cryogenic storage near atmospheric pressure)
-% (at 1 atm, liquid CH$_4$ at 111~K to achieve 22.2 MJ/L~\cite{makal2012methane}, liquid H$_2$ at 20 K to achieve 8~MJ/L \cite{suh2011hydrogen})
-or compression to high pressures (room temperature storage).
-% (200 bar for CH$_4$ to achieve 9.2 MJ/L \cite{makal2012methane} and 700 bar for H$_2$ to achieve 4.5-5.3 MJ/L~\cite{unknown}).
-Both approaches require expensive infrastructure at refilling stations and significant energy input; e.g., liquefaction of hydrogen consumes at least 30\% of its energy content \cite{bossel2003energy}.
-Moreover, high-pressure storage tanks are heavy, thick-walled, and non-conformable, while cryogenic storage tanks are bulky, expensive, and afflicted by boil-off losses \cite{hasan2009minimizing}.
-% \jordan{
-% Both forms of densification suffer several drawbacks varying from safety issues,  excessively expensive, boil-off~\cite{hasan2009minimizing}, and extensive refilling infrastructure~\cite{bossel2003energy}.
-% }
-% % Jordan: I commented this out! % Cory: I shorted it. I think it's important.
-
-A promising approach to densify natural gas \cite{makal2012methane,mason2014evaluating} and hydrogen \cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is through physical adsorption in nanoporous materials at room temperature~\cite{schoedel2016role}.
-The internal surfaces of porous materials attract gas molecules through van der Waals, electrostatic, etc.
-interactions to achieve a higher adsorbed gas density than the bulk gas at the same temperature and pressure, allowing for room temperature and lower-pressure storage and alleviating many drawbacks of high-pressure storage.
-% This would thereby reduce the cost of infrastructure at refilling stations, allow thinner-walled and therefore cheaper and lighter pressure vessels, reduce the energy requirements for gas densification, and alleviate safety concerns with high pressure storage.
-
-%As most homes in the US are connected to a natural gas pipeline \addcite, adsorbed natural gas fuel tanks could permit at-home fueling.
-
-% Jordan: I commented this out!
-% Cory: I think I agree that this paragraph is not necessary, thanks.
-% Current research is towards synthesizing a nanoporous material that can maximally densify natural gas/hydrogen whilst satisfying cost and stability constraints.
-So far, metal-organic frameworks (MOFs) \cite{furukawa2013chemistry} have demonstrated the highest internal surface areas ($>$7000 m$^2$/g \cite{farha2012metal}) and absorptive capacities for hydrogen \cite{suh2011hydrogen,garcia2018benchmark} and methane \cite{makal2012methane,mason2014evaluating}.
-Importantly, MOFs are highly adjustable materials synthesized by combining organic molecules (struts) and metal clusters (nodes) that
-% % I cut [often] here, since I'm not sure what the exception is and I don't think it's relevant.
-% coordinate to form a three-dimensional, porous framework \cite{furukawa2013chemistry}.
-This modular synthesis enables one to fine-tune the MOF chemistry to arrive at an optimal material for methane or hydrogen storage and delivery \cite{schoedel2016role}.
-See Fig.~\ref{fig:example_MOF} for the crystal structure of a canonical MOF, CuBTC \cite{chui1999chemically}.
-
-For a vehicle employing a porous material to store natural gas or hydrogen, the (volumetric) \emph{deliverable capacity} of the gas is the primary thermodynamic property of the material that determines the driving range~\cite{mason2014evaluating}.
-The adsorbed gas storage tank delivers the gaseous fuel to the engine via an (assumed) isothermal pressure swing \cite{sircar2002pressure}.
-The deliverable capacity (see Fig.~\ref{fig:delcap}) is the density of the gas in the material at the storage pressure $\pfull$ minus the residual gas that remains adsorbed at the lowest pressure $\pempty$ such that sufficient flow is maintained to feed the engine.
-For commercial feasibility, the US Department of Energy (DOE) has set deliverable capacity targets for both adsorbed methane and hydrogen storage and delivery for vehicles.
-For methane, the Advanced Research Projects Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$ per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure swing \cite{simon2015materials}.
-For hydrogen, the DOE set a series of progressive targets at five year intervals, with the ultimate target of 0.05 kg H$_2$/L \cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a minimum of -40 $^\circ$C \cite{allendorf2018assessment}.
-Thus far, despite the emergence of highly tunable materials with enormous surface areas, such as metal-organic frameworks \cite{furukawa2013chemistry}, no porous material has met these deliverable capacity targets.
-
-\begin{figure}
-    \centering
-    \subfloat[][]{
-        \includegraphics[width=0.6\columnwidth]{figs/hkust-1_fancy.png} \label{fig:example_MOF}
-    }
-    \qquad
-    \subfloat[][]{
-        \includegraphics[width=0.98\columnwidth]{figs/usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
-    }
-    \caption{Gas storage and delivery using metal-organic frameworks (MOFs).
-(a) the crystal structure of an archetype MOF, CuBTC \cite{chui1999chemically}.
-(b) the methane adsorption isotherm in CuBTC \cite{chui1999chemically} at 298 K (blue) (data from Ref.~\cite{mason2014evaluating}).
-The deliverable capacity $\rho_D$ is illustrated as the density of gas in the MOF at the storage pressure $\pfull$ minus the density at the discharge pressure $\pempty$.
-    }
-    \label{fig:fig1}
-\end{figure}
-
-To set realistic performance targets and optimally allocate resources for research efforts, in this work, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a rigid porous material and uses as input the experimentally measured properties of the bulk gas.
-Our extremum is provided by a substrate that offers a spatially uniform potential energy field felt by the gas.
-Applying our framework to methane and hydrogen gas, we find the US DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-Optimistically, new paradigms outside the scope of applicability of our theoretical framework, such as gas-induced structural transitions of the material, hold promise for meeting these targets, as evidenced by flexible MOF Co(bdp) which currently boasts the largest methane deliverable capacity~\cite{mason2015methane}.
-
-% Jordan: I commented this out!
-% To optimally allocate resources for research efforts and set realistic performance targets, it is important to assess the thermodynamic limits of technologies.
-Here, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a nanoporous material and uses as input experimental measurements of the bulk gas.
-Applying our framework to methane and hydrogen gas, we find the DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-
-\section{Gas storage \& delivery by isothermal, pressure-swing adsorption}
-% Here, we pose the context and question that we address in this work.
-% Consider pure gas storage and delivery by isothermal, pressure-swing adsorption \cite{sircar2002pressure}.
-Consider a pressure vessel onboard a vehicle (i.e.
-fuel tank) packed with porous material.
-At the filling stage, the tank is connected to a (pure) gaseous reservoir at pressure $\pfull$ and allowed to equilibrate.
-At this point, the adsorbed gas tank is considered full.
-While driving, gas desorbs from the adsorbent to the engine/fuel cell, driven by a pressure differential.
-The tank is considered depleted/empty when the pressure has dropped to $\pempty$, the pressure at which the flow rate of gas from the tank to the engine is insufficient.
-%
-However, given $\pempty \neq 0$ (pulling vacuum), residual gas will remain trapped in the adsorbent.
-Therefore, the driving range of the vehicle is primarily determined by the \emph{deliverable capacity}, defined as the density of gas in the tank at $\pfull$ minus density of the the residual gas remaining at $\pempty$.
-The [isothermal, volumetric] deliverable capacity is an intrinsic property of the nanoporous material and its interaction with the gas.
-%Precisely, let $\rho(p; T)$ be the density of gas in the adsorbent when in equilibrium with a bath of pure gas at pressure $p$ and temperature $T$, the \emph{adsorption isotherm}; $\rho(p,T)$ is essentially the equation of state of the gas adsorbed in the porous material.
-The [intrinsic, isothermal] deliverable capacity $\rho_D$ of the gas in the adsorbent is:
-%\begin{equation}
-%    \rho_D(\pfull, \pempty; T) \equiv \rho(\pfull; T) - \rho(\pempty; T).
-%    \label{eq:delcap}
-%\end{equation}
-% The pressures $\pfull$ and $\pempty$ correspond to chemical potentials $\mufull$ and $\muempty$ via an equation of state for the gas.
-% Eqn.~\ref{eq:delcap}
-% This assumes that the tank reaches equilibrium %, neglecting the kinetics of ad/desorption
-% at the end of the filling and depletion stages.
-
-%The assumption of isothermal operation neglects (i) heat released/consumed during ad/desorption, which can diminish the deliverable capacity by raising/lowering the temperature of the adsorbent during filling/discharge \cite{mota1997dynamics,chang1996behavior}, and (ii) an engineering strategy to use waste heat from the engine to heat the adsorbent and drive off residual gas when it nears $\pempty$ \cite{gomez2014exploring}.
-
-%Second, an engineering strategy to increase the deliverable capacity beyond the isothermal deliverable capacity is to reroute waste heat from the engine to the adsorbed gas tank during delivery.
-By raising the temperature of the adsorbent during the discharge stage, we drive off a fraction of the residual gas that would otherwise remain when the tank reaches $\pempty$ .
-In such a case, the [non-isothermal] deliverable capacity of the tank is $\rho(\pfull, T)-\rho(\pempty; T_H)$ where $T_H$ is the temperature to which we heat the adsorbent at discharge.
-This strategy, however, requires a more complicated design.
-
-% \begin{limitation}\label{limit:isothermal}
-% We assume isothermal pressure-swing adsorption.
-It is possible to achieve a greater deliverable capacity by raising the temperature of the adsorbent as the gas is extracted to drive off the residual gas.
-% \end{limitation}
-
-\section{Review of previous work}
-There has been considerable work attempting to establish an upper bound on the isothermal deliverable capacity in pressure-swing adsorption.
-% We now review previous work towards placing an upper bound on the isothermal deliverable capacity of hydrogen or methane by pressure-swing adsorption.
-% The first body of work rests on the simplified Langmuir model; the second body of work rests upon imposing energetic models for the gas-substrate and gas-gas interactions.
-
-Early work showed that, in the simplified Langmuir model, there exists an optimal free energy of adsorption (which determines the Langmuir constant) that maximizes the deliverable capacity $\rho_D$~\cite{matranga1992storage,bhatia2006optimum,simon2014optimizing}.
-If the gas-substrate interaction is too weak (strong), too little (much) gas adsorbs (is retained) at $\pfull$ ($\pempty$), diminishing $\rho_D$.
-An upper bound on the deliverable capacity of a Langmuir material follows if each adsorption site is endowed with the optimal free energy of adsorption.
-However, remaining is the question of how many adsorption sites per volume a porous material can practically offer, under the constraint that these adsorption sites offer the optimal free energy of adsorption.
-Moreover, gas-gas attractions, neglected in the Langmuir model, could recruit more gas in the material at $\pfull$ than at $\pempty$ and enhance the deliverable capacity~\cite{simon2014optimizing}.
-
-G\'omez-Gualdr\'on \emph{et al.} \cite{gomez2017impact} introduced a model that accounted for gas-gas interactions via an intermolecular potential and idealized the substrate in two different ways: (1) discrete adsorption sites packed into an FCC lattice and (2) a volume endowed with a spatially uniform, background potential energy field.
-According to molecular simulation of methane adsorption, the ARPA-E deliverable capacity target of 315 L STP/L could be reached in both of these idealized substrates.
-%The ARPA-E target of 315 L STP/L was reached when adsorption sites were endowed with an optimal heat of adsorption of 11.5 kJ/mol and adsorption sites were $5.3$ \AA~apart.
-In a second model, they endowed a volume with a background potential energy field, then simulated the adsorption of methane in the volume in the grand-canonical ensemble, using a Lennard-Jones potential for methane-methane interactions.
-The maximal deliverable capacity of 357 L STP/L was achieved with a background heat of adsorption of 8.6 kJ/mol.
-However, both models neglected the space occupied by atoms of the porous material that are needed to endow the adsorption sites/volume with the attractive potential energy.
-% To address this, in the third model, they constructed pseudo materials by placing Lennard-Jones spheres in space with tunable well depths.
-When constrained to form a connected network, none of the psuedo-materials reached the ARPA-E deliverable capacity target.
-
-A third body of work attempted to account for both gas-gas interactions \emph{and} steric interactions of the gas with the atoms of the porous material.
-% explores the upper limits of the deliverable capacity while accounting for gas-gas interactions \emph{and} the atoms of the porous material that are required to attract the guest; these material atoms exclude the adsorbate from regions of the control volume.
-%More works obtained an empirical upper bound on the deliverable capacity by explicitly accounting for the atoms of the material needed to create the potential energy field felt by the gas molecules.
-Simulations of methane adsorption in hundreds of thousands of explicit nanoporous crystal structures---both real and hypothetical---suggested that the ARPA-E deliverable capacity target is infeasible \cite{simon2015materials}; the highest simulated methane deliverable capacity was 196 cm$^3$~STP/cm$^3$.
-Confidence in this conclusion rests upon (i) the accuracy of the intermolecular potentials describing the molecular interactions and (ii) the sufficient sampling of material space, i.e.\ that the structures considered are representative of the set of possible materials.
-To further explore material space and address sensitivity to the potentials used, scaling the Lennard-Jones potential well depths of material atoms to model enhanced interactions \cite{gomez2014exploring}, placing Lennard-Jones spheres in a unit cell randomly to form ``pseudo-materials'' \cite{kaija2018high}, and generating fictitious potential energy fields via a generative adversarial network trained on zeolite structures \cite{lee2019predicting} all generated model substrates that all failed to meet the ARPA-E methane deliverable capacity target.
-
-% Brown and Freeman \cite{bhown2011analysis}
-
-In this work, we place a rigorous upper bound on the deliverable capacity of a pure gas in a rigid substrate by endowing a control volume with a spatially uniform background energy field.
-Instead of using a molecular model for the gas~\cite{gomez2017impact}, we use the experimental equation of state to account for gas-gas interactions.
-In addition, we prove using the calculus of variations that this spatially uniform substrate yields a maximal deliverable capacity.
-We use our framework to place an upper bound on the deliverable capacity of methane and hydrogen.
-
-\section{An upper bound on the deliverable capacity of a pure gas in a rigid porous material}\label{sec:upper-bound}
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/methane-298-rho-mu}
-  \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several MOFs at 298\ K as a function of chemical potential (bottom axis) and pressure (top axis).
-The bulk methane density is from the National Institute of Standards and Technology (NIST)~\cite{nist}.
-The methane adsorption isotherms in the MOFs are experimental data from Ref.~\cite{mason2014evaluating, furukawa2009storage}.
-  }
-  \label{fig:density-vs-mu-ch4}
-\end{figure}
-
-We now develop a thermodynamic framework that places an upper bound on the deliverable capacity of any pure gas in a rigid porous material.
-
-The thermodynamic properties of a bulk, pure gas are characterized by an equation of state. 
-Of particular interest for gas storage and delivery is the density of the gas as a function of chemical potential $\mu$ and temperature $T$, $\rho_g(\mu,T)$, which is shown for methane gas in Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K.
-For comparison, we also show the density of methane adsorbed into several porous materials.
-
-% \davidsays{This paragraph could move considerably later. Or cut entirely?}
-% We model the interaction of substrate with gas as a simple Gibbs free energy of adsorption per mole (or atom) \gst.
-We will assume that this Gibbs free energy is independent of the concentration of gas.
-At low concentrations, this assumption is reasonable, because the gas molecules can be assumed to be widely separated and their interactions weak.
-In the limit of high concentrations, we expect that the attraction of the gas to the substrate will be diminished, since the ``good spots'' will already be occupied.
-The choice of a concentration-independent $\gst$ will thus have a tendency to overestimate the deliverable capacity, providing an upper bound (for more on this, see Sec.~\ref{sec:qualitative-max}).
-
-To place an upper bound on the deliverable capacity, consider a substrate whose sole interaction with the gas is to introduce a spatially uniform potential energy $\V$
-for gas molecules in the control volume (where $\V<0$ for an attractive potential).
-Because this interaction is uniform within the substrate, the gas-gas interactions and thus fluid structure in such a substrate are identical to those of the pure gas at the same density and temperature.
-This allows us to obtain the adsorption properties of this model substrate using the experimentally measured properties of the pure gas~\cite{nist}.
-Intuitively, the deliverable capacity of a gas in such a homogeneous substrate with the optimal potential energy is an upper bound because, in a realistic material, (i) spatial inhomogeneity of the potential energy results in some points in the control volume offering a suboptimal attraction for the gas and (ii) the atoms required to create the potential field will exclude gas from occupying a fraction of the control volume.
-%\corysays{the suboptimal energy is hard to argue since that would affect entropy too...}
-
-% \corysays{I propose to replace ``homogenous'' with ``spatially uniform'' everywhere instead of using both interchangeably.
-% \davidsays{We replaced ``homogeneous potential'' with spatially uniform, but were less confident replacing ``homogeneous substrate'', which has a slightly different meaning, and might need a slightly different name.
-e.g.
-maybe ``uniform substrate'' or ``model substrate''?}}
-
-%\corysays{what do u think about callin this an ``ideal substrate'' or giving it some name so we can refer to it hereon?
-\davidsays{I think I'd prefer ``homogeneous substrate'' or ``idealized substrate''}}
-
-%\subsection{Finding the density}\label{sec:finding-n}
-
-The spatially uniform potential $\V$ representing gas-substrate interactions behaves as an external potential and effectively shifts the chemical potential (or, equivalently, molar Gibbs free energy) of the gas in the material, just as gravitational potential energy causes the density of air to vary with altitude.
-Consequently, the density of gas in our homogeneous substrate is:
-\begin{align}
-    \rho(\mu,T) &= \rho_g(\mu - \V,T). \label{eq:mof-density}
-\end{align}
-See Sec.~\ref{sec:V_shifts_chem_pot} for a derivation.
-% where $\rho_g(\mu,T)$ is the density of the pure gas at chemical potential $\mu$, shown for methane in Fig.~\ref{fig:density-vs-mu-ch4}. 
-% cory says: ^ we already defined \rho_g in pargraph above.
-
-The deliverable capacity of gas in our homogeneous substrate is thus:
-\begin{align}
-    \rho_D(\V,T) &= \rho(\mufull,T) - \rho(\muempty,T),
-    \label{eq:DofPhi}
-    \\
-    &= \rho_g(\mufull-\V,T) - \rho_g(\muempty-\V,T),
-\end{align}
-where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the pressures $\pfull$ and $\pempty$, respectively.
-Thus, $\rho_D$ for our homogeneous substrate as a function of $\V$ is the difference between two shifted versions of $\rho_g(\mu; T)$.
-Figure~\ref{fig:methane-298-D} shows the two shifted bulk methane density curves and their difference.
-The horizontal axis
-is the difference in molar Gibbs free energy at fixed density and temperature between the gas within the substrate and the pure gas (\gst), which is equal to $\V$ in our model.
-We see a potential $\V$ that maximizes the deliverable capacity of our ideal substrate, balancing the need to maximize the density at $\pfull$ against the need to minimize residual gas retained at $\pempty$.
-%Our theoretical framework thus recovers an upper bound on the deliverable capacity of methane (298\ K, $\pfull=$65\ bar, $\pempty=$5.8\ bar) in porous materials as 374\ L STP/L.
-The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, \red{achieved for $\V_{opt} = ...$}.
-%
-%As a ballpark, we can expect the attractive potential to be optimal around the maximum slope of $\rho_g(\mu)$.
-This is the point where the steric repulsion between the gas atoms starts leading to saturation of the density.
-
-An essential question is whether our idealized substrate offering the spatially uniform potential $\V_{opt}$ for the gas places an upper bound on the deliverable capacity in all rigid porous materials.
-%yields the maximum deliverable capacity over all possible potential energy fields. 
-In the SI (see Section~\ref{sec:fixme} \davidsays{Look up section label}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest.
-The potential energy field is \emph{static} if it is unaffected by the presence of the gas; consequently, our model does not apply to flexible materials that undergo gas-induced conformation changes \cite{schneemann2014flexible}.
-We next argue that this extremum is a maximum by addressing two alternative possibilities: the extremum could turn out to be (i) a saddle point or (ii) a local, rather than global, maximum.
-
-Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum $\rho_D$ over all \emph{spatially uniform}, static potential energy fields.
-To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a maximum deliverable capacity over \emph{all} (including non-spatially uniform) static potential energy fields (a broader claim), consider a non-spatially-uniform variation on $\V_{opt}$.
-We will show that the effect of this variation is to reduce the deliverable capacity.
-At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, the additional gas molecules will be forced into higher energy locations.
-Thus, the mean attraction will be lowest at \pfull\  and highest at \pempty.
-This will result in $\gst$ increasing monotonically with with pressure.
-As shown in the Supplementary Information (see Sec.~\ref{sec:monotonic}), this results in the deliverable capacity of a non-uniform potential being lower than the deliverable capacity of a uniform potential with $\V=\gst$ for either $\pfull$ or $\pempty$.
-Thus a spatially uniform potential will give the best deliverable capacity amongst all static potential energy fields that give the same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will give the a greater deliverable than any static non-uniform potential.
-
-% Yo David, I changed ur paragraph below. tell me wut u think.
-% To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a \emph{maximum} deliverable capacity over all static potential energy fields, consider the case of a material that offers a non-spatially-uniform potential energy felt by the gas molecules.We wish to show that this material will have a lower deliverable capacity than our ideal substrate.
-At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, some of the gas molecules will be forced into higher energy locations.
-Thus, the mean attraction will be lowest when the tank is ``full,'' and highest when the tank is ``empty.'' This is the opposite of what we would want to achieve a high deliverable capacity.
-Based on this qualitative reasoning, we expect the spatially uniform potential \red{$\V_{opt}$} to give the maximal deliverable capacity, since it will provide the most attraction for the molecules added at high densities.
-
-% \corysays{``To qualitatively argue that the spatially uniform potential provides a \emph{maximum} deliverable capacity'' do we need to qualify that this spatially uniform potential must hv the optimal $\V$?
-obviously a spatially uniform potential can be better than a spatially uniform one with a poor $\V$.
-maybe we should start with our optimal homogenous potential, then say introduce a little blip in it from there, like in calculus of variations.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/methane-298-n-vs-G}
-    \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
-    Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line (not visible at this scale).}
-    \label{fig:methane-298-D}
-\end{figure}
-
-\section{Results}
-Although our theoretical framework allows us to place an upper bound on the deliverable capacity of many different gases in a rigid porous material, we focus on the maximal deliverable capacity of methane and hydrogen gas in our homogeneous substrate due to their application as transportation fuels.
-To characterize the density of the gases, $\rho_g(p, T)$, we use experimental data from NIST~\cite{nist}, which naturally includes quantum effects that particularly affect hydrogen at low temperature~\cite{kumar2006quantum}.
-In the context of storage onboard passenger vehicles, we compare our upper bound with several prominent porous materials using experimental adsorption isotherms from the literature; we also compare with deliverable capacity targets set by the US DOE.
-
-Figure~\ref{fig:methane-298-D} shows the upper bound for methane storage at room temperature~(374~cm$^3$~STP/cm$^3$). In addition to the predicted maximum deliverable capacity as a function of \gst,
-\corysays{subtlety here: isnt ur proof restricted to saying $\V_{opt}$ provides the upper bound?
-does it really prove that, for fixed and nonoptimal $\gst$, a spatially uniform potential is optimal over a non-spatially uniform one?
-I don't think it does...}
-the ARPA-E target of 315~cm$^3$~STP/cm$^3$~\cite{arpaemove} is shown for context.
-For an adsorbent with at least 84\% void fraction, the ARPA-E target is theoretically possible.
-The experimental deliverable capacities of several adsorbents~\cite{mason2014evaluating} are also shown over a range of $\gst$ (converted from the measured adsorption isotherm as explained in \davidsays{confirm this} Sec.~\ref{sec:phi-is-delta-g}).
-These can be compared with the highest observed deliverable capacities for methane at room temperature in rigid materials, up to 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}.
-% don't think the below is helpful or correct: Gst opt is reached by some... there are void fraction constraints too.
-% To reach the ARPA-E target, an adsorbent material with a higher $|\gst|$ would need to be found.
-
-\corysays{I know it doesn't matter, but should we choose L STP/L or cm3 STP/cm3?
-we use both, some might get confused/distracted?
-\davidsays{We chose to use the units used for each of the targets thinking it would be more familiar and/or easier to compare.
-I don't like either set of units, and if we were to pick ``our'' choice, I'd lean towards something dimensionless with a sane comparison (e.g.
-relative to the gas density at our high pressure)}}
-
-\corysays{need to explain why the MOFs are lines not points... that \gst has a range for MOFs.
-\jpsays{due to a different deliverable capacity at full pressure and empty pressure.
-For hydrogen the SI (by Paula Garcia-Holley) has total uptake plots so we could make a curve even but would have a lot of error due to manual extraction?}}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-298-n-vs-G}
-    \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
-    \label{fig:hydrogen-298-D}
-\end{figure}
-
-\corysays{Figs.~\ref{fig:hydrogen-298-D} ~\ref{fig:methane-298-D} why not just let $\gst$ be negative and have an arrow pointing in direction of more attraction?
-as written the legend is not correct since those are not the functions being shown.
-in the main text we never discussed what \gst is for the MOF and why the MOFs are lines not points (\gst is a function of density now)}
-
-\jpsays{I edited everything below this comment. The original is commented out}
-
-Storage of hydrogen is considerably more challenging owing to its relatively weak interaction with adsorbents.
-The DOE ULTIMATE deliverable capacity target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C \red{check}.
-Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve for hydrogen storage along with a range of experimental measurements for known MOFs.
-The DOE ULTIMATE deliverable capacity target is theoretically possible, however by such a small margin that we can safely rule out the possibility of reaching this target through storage and delivery of hydrogen in \emph{any} rigid substrate at room temperature.
-Such a material would require a void fraction of at least 94\% \red{check}.
-On top of this, the DOE ULTIMATE target requires an optimal $|\gst|$ of 10~kJ/mol which is far greater than what is found in observed porous materials.
-This reflects the known fact that hydrogen interacts with substrates far more weakly than methane does.
-One possibility that has been pursued is storage at cryogenic temperatures.
-See Section~\ref{sec:cryo-hydrogen} in the Supplementary Information for the upper bound for storage of hydrogen at 77 K.
-
-Why do we not experimentally observe materials approaching the upper bound on the deliverable capacity?
-Foremost, any adsorbent substrate will be composed of atoms, which will exclude the gas from some volume.
-Due to strong short-range interactions, substrate atoms must be uniformly distributed to achieve strong attraction for gas atoms.
-Uniform atom distribution sets a limit on the pore volume.
-% Because strong interactions are short-range, achieving a strong attraction requires that substrate atoms be distributed throughout the volume, which places a limit on the pore volume.
-In contrast, real materials have a spatially non-uniform attraction for gas atoms.
-The result of this is that there are regions of space with non-optimal attraction.
-For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.
-% Finally, real materials must have a spatially non-uniform attraction for gas atoms, resulting in some regions of space where the attraction is not optimal.
-For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.
-
-\section{Conclusion}
-We have established an upper bound on the deliverable capacity via pressure-swing adsorption in rigid porous solids based on experimentally measured properties of pure gases.
-While these upper bounds do not rule out the discovery of materials that reach current DOE targets, they cast strong doubt on the possibility of achieving these goals when we consider the additional constraints imposed due to steric hindrance between substrate atoms and the adsorbate.
-Our upper bound does indicate that those goals cannot be exceeded by more than 19\% for methane and 6\% for hydrogen \red{check}.
-Fortunately, there are some limitations to these upper bounds which suggest avenues for future developments which could allow the targets to be met.
-
-The first limitation of our proof is that we restricted ourselves to \emph{isothermal} pressure swing storage.
-By raising the temperature of the adsorbent during gas discharge to drive off residual gas \cite{gomez2014exploring}, the deliverable capacity could be enhanced, albeit at the cost of a more complicated engineering design of the fuel tank and vehicle.
-%While this approach poses engineering challenges, in principle it could allow an arbitrarily large deliverable capacity.
-
-% \corysays{no, cant lower filling temp. when park it will heat up then go above pressure rating of tank...}
-
-The second limitation of our proof arises in the assumption of a rigid substrate.
-The rigid substrate acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
-% The second limitation of our proof arises in the assumption of a rigid substrate, which acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
-For most porous materials this is a reasonable approximation, and this assumption is frequently made in both the simulation and theory of porous materials~\cite{witman2017influence}.
-However, there are cases where the substrate can in effect provide a very strongly gas-density-dependent interaction through structural flexibility~\cite{schneemann2014flexible}.
-A flagship example is MOF Co(bdp)~\cite{choi2008broadly}, which possesses a wine-rack-like topology capable of hinge motion.
-At low methane pressure, Co(bdp) adopts a collapsed, nonporous state, but expands to a porous state and fills with gas at higher pressures \cite{mason2015methane}.
-This allows Co(bdp) to fully expel its residual gas at low pressures.
-Our upper bound does not consider such material effects and thus flexible materials could have significantly higher deliverable capacities.
-% Such materials can indeed violate our upper bound and thus could lead to significantly improved deliverable capacities.
-
-We note that our upper bound can readily be applied to the storage of other gasses of interest, with the proviso that the gas is far from crystallization. Our code is available at \davidsays{XXXXX}.
-
-% \clearpage
-
-% Here we will articulate the assumptions required for our upper bound to hold. Some of these assumptions serve as a guide for chemists who aim to synthesize a material that violates our upper bound.
-
-% Our proof that the spatially uniform potential is an extremum assumes that the fluid is stable at the temperature and densities involved.
-% \begin{limitation}\label{limit:crystallization}
-% Our proof assumes the gas is cannot become spontaneously solid at the temperature of interest.
-% \end{limitation}
-% Our upper bound will \emph{not} be valid for gasses which may crystallize in their pure form.
-This is because a periodic potential can interact with a high-density crystal far more strongly than it interacts with a low-density gas.
-We are not aware of a practical application of this principle, but theoretically it is possible.
-For the application of gas storage for light gasses such as methane or hydrogen at moderately high temperatures (say liquid nitrogen or warmer), the pure gas is sufficiently far from its crystal form that we need not concern ourselves with crystallization.
-We do not see this particular assumption as one that can be worked around.
-
-% There remains the possibility that the spatially uniform potential is a local maximum, but there is another global maximum.
-We do not see how to formally rule out this possibility for an arbitrary gas-gas interaction, but the qualitative arguments above seem sufficient to rule out this possibility for a real gas.
-
-% However, there was a more fundamental assumption, which is that the porous material acts as an effective potential for the gas molecules, and that the gas-gas interaction is unmodified.
-% \begin{limitation}\label{limit:field}
-% Our proof assumes the interaction of the gas with the substrate does not modify the substrate.  Equivalently,  the interaction of substrate with gas operates as a simple potential energy field.
-% \end{limitation}
-% For \emph{most} porous materials this seems likely to be a pretty good approximation.
-However, there are unusual cases where the substrate can in effect provide a very strongly density-dependent interaction.
-This is perhaps easiest to see in the example of \davidsays{look up the collapsing MOFs description from the RMS-MOF paper}, in which the gas can introduce a phase change in the solid, which has the effect of drastically lowering the density of gas in the material.
-Such materials can indeed violate our upper bound.
-\davidsays{Remember to discuss this in the conclusions or perhaps discussion, we can frame these limitations as guides.
-The limitations of our upper bound are the design requirements for a material that will do any better.}
-
-% [impurities?]
-
-% \section{Conclusion}
-% As many have suspected \davidsays{cite previous work}, pressure-swing adsorption in rigid porous solids is not going to enable reaching targets.
-\davidsays{conclude explicitly about each target}  Unlike previous work, our upper bound on deliverable capacity is based on experimentally known properties of the gas of interest rather than simulations.
-
-% The limitations on our proof of an upper bound on the deliverable capacity provide hints as to approaches that could in principle achieve the targets.
-In this paper we have identified three limitations of our upper bound on the deliverable capacity.
-Two of these three limitations suggest engineering strategies that do hold promise of achieving the targets for hydrogen and methane storage.
-Limitation \ref{limit:crystallization}, which states that the gas must not crystallize does not means for improvement, since it is a limit on the gasses to which our bound may be applied, and neither hydrogen nor methane are in any danger of crystallizing.
-
-% Crystal density vs. pelletiziation density.
-
-% A more subtle limitation of our proof is that it assumes that the substrate itself is not modified by the presence of the gas.
-While this assumption is never precisely true, since any adsorbed atom will change the electron distribution of the substrate to some degree, it tends to be sufficiently accurate that most atomistic simulations can get away with assuming that the substrate is unmodified.
-A dramatic exception to that rule is the field of flexible MOFs, in which the presence of the gas can induce a phase change which can \emph{drastically} modify the gas-substrate interaction~\davidsays{cite experiments}\cite{witman2019}.
-
-% wut about mixtures?
-
-%TC:ignore
+\input{gas-ads-body.tex}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figs/g_st_piston_2.png}
+    \includegraphics[width=\columnwidth]{g_st_piston_2.png}
     \caption{Cartoon illustrating the definition of $\gst$ and a hypothetical experiment to measure it with a piston and removable barrier that separates two equal volumes, one with a porous material (bottom) and one with free space (top).
@@
     \centering
-    \includegraphics[width=0.95\columnwidth]{figs/qst-vs-delta-G}
+    \includegraphics[width=0.95\columnwidth]{qst-vs-delta-G}
     \caption{Relationship between $\gst$ and the isosteric heat $q_{st}$ for several prominent adsorbents.
@@
     \centering
-    \includegraphics[width=0.95\columnwidth]{figs/methane-298-gst}
+    \includegraphics[width=0.95\columnwidth]{methane-298-gst}
     \caption{The density-dependence of $\gst$ of methane in several adsorbents (298\ K). Note that $\gst$ is monotonic in $\rho$. Data from Refs.~\ref{mason2014evaluating,furukawa2009storage}.
@@
     \centering
-    \includegraphics[width=0.95\columnwidth]{figs/four-cases}
+    \includegraphics[width=0.95\columnwidth]{four-cases}
     \caption{Four cases we consider.}
@@
     FIGURE HERE A B C D
-    \includegraphics[width=0.95\columnwidth]{figs/hydrogen-77-n-vs-G}
+    \includegraphics[width=0.95\columnwidth]{hydrogen-77-n-vs-G}
     \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$.

### 2020-05-02 ising/ising-body.tex
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

--- 2020-05-02 ising/ising.tex
@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
-\usepackage{graphicx} % Include figure files
-\usepackage{color}
-\usepackage{nicefrac} % Include for inline fractions
-
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
-\usepackage{mdframed}
-\usepackage{braket}
-
-% define colors for comments
-\definecolor{dark-gray}{gray}{0.10}
-\definecolor{light-gray}{gray}{0.70}
-
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
-
-\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
-\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
-\begin{document}
-
-\title{Flat-histogram method comparison on 2D Ising model
-}
-
-\author{Jordan K. Pommerenck} \author{David Roundy}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-
-\begin{abstract}
-We examine the convergence of the stochastic approximation with a dynamic
-update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
-stochastic approximation Monte Carlo (SAMC) and Wang-Landau (WL) methods show that SAD performs robustly and without user input knowledge of an energy range.
-We also compare SAD with ``production'' run WL with fixed weights which is the first of its kind recorded.
-Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
-\end{abstract}
-
-\maketitle
-
 \section{Introduction}
@@
 Martin Weigel for discussions regarding entropic population annealing.
-
-\bibliography{ising} % Produces the bibliography via BibTeX.
-
-\end{document}

### 2020-05-02 ising/ising.tex
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

--- 2020-05-01 ising/ising.tex
@@
 
-\section{Introduction}
-Flat-histogram Monte Carlo simulation algorithms calculate the thermodynamic
-properties of various systems over a range of temperatures.  The first histogram
-method used a single canonical Monte Carlo simulation to predict properties for
-nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
-samples a narrow energy range, it proves computationally inefficient at sampling large energy ranges.
-
-Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram
-sampling which improved the exploration of configurational space and allowed the
-simulation to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
-These works led to increase in the development of a variety of ``flat'' (or ``broad'') histogram
-methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-wang2001determining, wang2001efficient} which could explore a wider range
-of energies.  In addition to obtaining thermodynamic information for the entire
-energy range for a single simulation, these approaches cannot be easily trapped
-in a local energy minimum like a canonical simulation.
-
-Wang and Landau introduced one of the most widely used flat-histogram
-Monte Carlo algorithms that determined the density of states (DOS) for
-a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
-its power, the method unfortunately requires \emph{a priori.} knowledge of several
-user-defined parameters. Thus, for any given system under study, the user needs
-to determine the ideal parameters in order to apply the method. The Wang-Landau
-algorithm is also known to violate detailed balance (although only for brief
-time intervals)~\cite{yan2003fast, shell2002generalization}. With the violation
-of detailed balance, convergence of the algorithm is not guaranteed.
-
-Because of the uncertainty of convergence for WL, many studies have been undertaken
-to understand how the modification (or update) factor $\gamma$ impacts the
-convergence~\cite{zhou2005understanding,lee2006convergence,
-belardinelli2007wang}.
-Belardinelli and Pereyra showed that an update factor that decreases faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang, belardinelli2008analysis, zhou2008optimal}, where $t$ corresponds to the number of moves.
-Schneider \emph{et al.} outline minor refinements algorithm including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.
-These studies led to the
-formation of the $1/t$-WL algorithm and also sparked some debate as to whether
-the entire simulation should be split into multiple parts, called ``production run" WL, in order to preserve ergodicity and
-detailed balance~\cite{jayasri2005wang, mukhopadhyay2008monte}.
-
-Liang independently considered whether WL could be treated as a special case of
-stochastic approximation whose convergence could be mathematically
-proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang \emph{et
-al.}~\cite{liang2007stochastic} argued that WL can be considered a form of
-stochastic approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
-convergence (if certain conditions are met). Despite the added benefit of
-guaranteed convergence, the method still has a system specific user-defined
-variable. Such variables often create difficulty when applying Monte Carlo
-methods across arbitrary systems.
-
-Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
-related Statistical Temperature Molecular Dynamics (STMD), an adaption
-of the WL method that approximates the entropy as a piecewise linear function,
-which improves convergence for systems with a continuously varying
-energy~\cite{kim2006statistical, kim2007statistical}. STMC applied to WL
-requires a temperature range be specified rather than an energy range.  Kim
-\emph{et al.} extended this work as Replica Exchange Statistical Temperature
-Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping STMC
-simulations to improve convergence~\cite{kim2009replica}. Recently,
-Junghans \emph{et al.} demonstrated a close connection between
-metadynamics, which was introduced by Laio and
-Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods, with STMD
-forging the connection~\cite{junghans2014molecular}.
-
-The SAD (stochastic approximation with a
-dynamic $\gamma$) method outlined by Pommerenck
-et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
-that dynamically chooses the modification factor rather than relying on system
-dependent parameters. SAD shares the same convergence properties with SAMC while
-replacing un-physical user-defined parameters with the algorithms dynamic
-choice.
-
-In this work, we apply the family of weight-based flat-histogram Monte
-Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
-
-\section{Ising Model}
-The 2D Ising spin-lattice system is widely used as a testbed when
-benchmarking or comparing Monte Carlo
-methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing, barash2019estimating}. The 2nd order
-phase transition behavior and the ability to directly calculate the
-exact solution for finite lattices~\cite{beale1996exact, haggkvist2004computation} make the
-system sufficiently interesting for such theoretical comparisons. It is
-also important to note that direct comparison of the other methods can
-be made with WL as its original implementation was done on this
-system~\cite{wang2001determining,wang2001efficient}.
-We test the convergence of several flat-histogram methods
-on the periodic 2D square lattice ferromagnetic Ising model with nearest
-neighbor interactions~\cite{landau2004new}.
-\begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
-\end{align}
-The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
-for up or down spins respectively. In the absence of a magnetic field ($h =
-0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
-kaufman1949crystal}:
-\begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
-\end{align}
-where the sum is over nearest neighbor spin sites. Beale showed that for finite
-lattices the Density of States could directly be calculated from the partition
-function~\cite{beale1996exact}
-\begin{align}
-Z = \sum_E g(E) e^{-{\beta E}}
-\end{align}
-% where $g(E)$ is the multiplicity of the system which is proportional to
-% $D(E)$. We can estimate the average deviation from the exact solution
-% using the relation~\cite{schneider2017convergence, shakirov2018convergence,
-% barash2017control}:
-% \begin{align}
-% \braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
-% \end{align}
-where $g(E)$ is the multiplicity of the system which is proportional to
-$D(E)$. We can estimate the maximum deviation in the canonical specific heat capacity $c_V$ from the exact solution~\cite{schneider2017convergence, shakirov2018convergence,
-barash2017control,barash2017gpu}:
-\begin{align}
-\braket{\epsilon(t)}_E = \left|\max_E \left(c_V(E,S,T,t) - c_V(E,S,T,t)_{\text{Beale}}\right)\right|
-\end{align}
-% As per the implementation by Wang and
-% Landau~\cite{wang2001determining}, the total number of available energy
-% states are $N-1$.
-Computing the specific heat capacity $c_V$ presents a difficult challenge for
-any Monte Carlo method due to fluctuations in the derivative of the internal energy.  Methods that accurately compute $c_V$ also by extension accurately compute the internal energy.
-
-\section{Flat-histogram methods}\label{sec:histogram}
-Flat-histogram methods determine the density of states $D(E)$ over a broad range
-of energies by simulating each energy with equivalent accuracy. Flat-histogram
-Monte Carlo methods propose randomly chosen ``moves'' which change the state of
-the system and must satisfy detailed balance.  Each algorithm differs in how it
-determines the probability of accepting a move and in what additional statistics
-must be collected in order decide on that probability.
-
-We describe several closely related flat-histogram methods which each rely on a
-weight function $w(E)$ to determine $D(E)$.  For these algorithms, the
-probability of accepting a move is given by
-\begin{equation}
-	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
-	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
-\end{equation}
-which biases the simulation in favor of energies with low weights. The result of
-weights $w(E)$ that are proportional to $D(E)$ is an entirely flat-histogram. We
-can relate the entropy to the weights in the microcanonical ensemble, since the
-entropy is defined as $S(E) \equiv k_B\ln(D(E)) \approx \ln w(E)$.
-
-Flat-histogram methods employ a random walk in energy space to estimate $D(E)$.  Each method operates by continuously updating the weights at each
-step of the simulation
-\begin{equation}
-	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
-	+\gamma_t
-\end{equation}
-where $t$ is number of the current move, $\gamma(t)$ is a move-dependent update
-factor, and $E$ is the current energy.  This update causes the random walk to
-avoid frequent sampling of the same energies, leading to a rapid exploration
-of energy space. Flat-histogram methods differ primarily in how they schedule
-the decrease of $\gamma_t$.
-
-The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
-landau2014guide} explores energy space by setting $\gamma_{t=0}^{\text{WL}}=1$,and then decreases $\gamma^{\text{WL}}$ in prescribed stages. An energy range of
-interest must be specified~\cite{wang2001efficient, schulz2003avoiding,yan2003fast}, which often requires multiple simulations if unknown.
-
-The number (``counts'') of moves ending at each energy are stored in a
-histogram.
-For a sufficiently flat energy histogram (typically user-specified to be 0.8), $\gamma^{\text{WL}}$ is decreased by a specified factor of $\frac12$ and the histogram is reset to zero.
-The entire process is repeated until $\gamma^{\text{WL}}$ reaches a desired cutoff.
-
-The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
-from dropping below $N_S/t$~\cite{belardinelli2008analysis,
-schneider2017convergence}. The method follows the standard WL algorithm with two
-modifications.  Firstly, when each energy state has been visited once, the histogram is considered flat and $\gamma_t$ is
-decreased by a factor of two. Secondly, when
-$\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the update factor becomes
-$\gamma_t = N_S/t$ for the remainder of the simulation:
-\begin{align}
-  \gamma_t^{1/t\text{-WL}} = \begin{cases}
-     \gamma^{\text{WL}}_t & \gamma^{\text{WL}}_t > \frac{N_S}{t} \\
-     \frac{N_S}{t} & t \ge t_0
- \end{cases}
-\end{align}
-where $t$ is the number of moves, $\gamma^{\text{WL}}_t$ is the Wang-Landau update factor
-at move $t$, and $N_S$ is the number of energy bins.
-
-The WL method can alternatively be implemented as part of a production run which aids in
-equilibration~\cite{gross2018massively}. WL is used to generate the weights
-resulting in a flat-histogram~\cite{janke2017generalized}. Once a user-defined
-minimum $\gamma^{\text{WL}}_t$ is reached, the simulation switches to a final production run
-with fixed weights. WL implemented in this way can ensure ergodicity and
-detailed balance; however, the convergence of simulation is still impacted by
-the choice of the minimum $\gamma^{\min}_t$.
-
-Another weight-based flat-histogram method is
-the stochastic approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
-schedule by which the update factor $\gamma^{\text{SA}}_t$ is continuously
-decreased~\cite{liang2007stochastic, werlich2015stochastic,
-schneider2017convergence}.  The update factor is defined in the
-original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$,
-\begin{align}
-\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
-\end{align}
-where as above $t$ is the number of moves that have been attempted.
-
-The implementation of SAMC is extremely simplistic.
-In addition, Liang has proven that the weights converge to the true
-density of states~\cite{liang2006theory, liang2007stochastic,
-liang2009improving} provided the update factor satisfies
-\begin{align}
-\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
-\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
-\end{align}
-where $\zeta > 1$.  Unlike WL methods, the energy range need not be
-known \emph{a priori.} and the convergence time depends only on the choice of
-parameter $t_0$.
-Unfortunately, $t_0$ can be difficult to chose in advance
-for arbitrary systems.
-Liang \emph{et al.} give a rule of thumb in
-which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
-is the number of energy bins~\cite{liang2007stochastic}.  Schneider
-\emph{et al.} found and we confirm that for the Ising model this heuristic is helpful for small spin systems, but that larger systems require an even higher
-$t_0$ value~\cite{schneider2017convergence}.
-
-Pommerenck \emph{et al.} propose a refinement~\cite{pommerenck2020stochastic} to SAMC
-where the update factor is determined dynamically rather than by the user.
-stochastic approximation with a dynamic $\gamma$ (SAD) requires the user to
-provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
-requiring \emph{a priori.} an energy range of interest; however, this is almost always easier to
-identify and is more physical than the SAMC parameter $t_0$. The update factor
-can be written in terms of the current estimates for the highest $E_H$ and
-lowest $E_L$ energies of interest and the last time that an energy in the range
-of interest is encountered $t_L$.
-\begin{align}
-  \gamma_{t}^{\text{SAD}} =
-     \frac{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
-     }{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
-     }
-\end{align}
-SAD only explores the energy range of interest as specified by the minimum
-temperature of interest $T_{\min} < T < \infty$. During the simulation the two
-energies $E_H$ and $E_L$, are refined such that the range of energies are conservatively
-estimated. The weights are calculated for each energy region according to the original
-prescription.
-\begin{enumerate}
-\item {$E < E_L$:} $w(E>E_H) = w(E_H)$
-\item {$E_L < E < E_H$:} moves are handled the same as other weight-based
-methods that are mentioned
-\item {$E > E_H$:} $w(E<E_L) = w(E_L)e^{-\frac{E_L-E}{T_{\min}}}$
-\end{enumerate}
-Each time the simulation changes the value of $E_H$ or $E_L$, the weights
-within the new portion of the interesting energy range are updated.
-
-\section{Results}
-We tested the algorithms on two different system sizes of the 2D Ising model.  The first is a smaller
-simulation with a lattice size of $N = 32 \times 32$ and the second has a lattice size of $N = 128 \times 128$. The SAD method explores the energy
-space of each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations lead to the minimum important energy $E_{\min}$
-and maximum entropy energy $E_{\max}$ being calculated (with the exception of
-the WL methods where both of these parameters are needed \emph{a priori.}).
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{gamma-n32.pdf}
-  \caption{
-  The update factor $\gamma_t$ versus the iteration number for the $N=32 \times 32$
-  system.}
-  \label{fig:N32-gamma}
-\end{figure}
-
-% \begin{figure}
-% \includegraphics[width=\columnwidth]{N32-entropy-error-default.pdf}
-%   \caption{
-%   The average entropy error for each MC method for $N=32 \times 32$,
-%                $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-%                as a function of number of iterations run.  The error is
-%                averaged over 8 independent simulations, and the best
-%                and worst simulations for each method are shown as a
-%                semi-transparent shaded area.}\label{fig:n32}
-% \end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N32-Cv.pdf}
-  \caption{
-    The specific heat capacity versus the reciprocal temperature $\beta$ for the $N=32 \times 32$ system for each histogram method at $10^{9}$ moves.}
-  \label{fig:N32-cv}
-\end{figure}
-
-\subsection{The 32 $\times$ 32 Ising model}
-Figure~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
-All of the update factors initially start at $\gamma_{t=0} = 1$.
-SAD dynamically updates $\gamma_t$ throughout the simulation.
-After about $10^{10}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
-Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown.
-The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$.
-The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
-All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.
-
-All the methods except for SAMC gave a reasonable approximation for
-the heat capacity peak resulting from the phase transition within 10$^9$ moves.
-Figure~\ref{fig:N32-cv} shows the specific heat capacity vs.
-the reciprocal temperature $\beta$ at $10^{9}$ moves.
-Each method (except for the SAMC methods) is shown using the same single random number seed.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
-  \caption{The maximum error in the specific heat capacity for each method for the $N=32 \times 32$ and $T_{\min} = 1$ as a function of number of iterations run.
-The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
-  \label{fig:N32-cv-error}
-\end{figure}
-
-Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
-this system. The solid/dashed lines
-represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
-over eight simulations using different random number seeds. The range of maximum
-errors for each simulation is shown as a shaded region. By
-the time $10^8$ moves have been made all but the WL simulation have begun to
-converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^9$ moves.
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{gamma-n128.pdf}
-  \caption{
-  The update factor $\gamma_t$ versus the iteration number for the $N=128 \times 128$
-  system.}
-  \label{fig:N128-gamma}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N128-Cv-error-default.pdf}
-  \caption{The maximum error in the specific heat capacity for each method for the $N=128 \times 128$ and $T_{\min} = 1$ as a function of number of iterations run.
-The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
-  \label{fig:N128-cv-error}
-\end{figure}
-
-% \begin{figure}
-% \includegraphics[width=\columnwidth]{N128-entropy-error-default.pdf}
-%   \caption{
-%   The average entropy error for each MC method for $N=128 \times 128$,
-%                $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-%                as a function of number of iterations run.  The error is
-%                averaged over 8 independent simulations, and the best
-%                and worst simulations for each method are shown as a
-%                semi-transparent shaded ar.}\label{fig:n128}
-% \end{figure}
-
-\subsection{128 $\times$ 128 Ising system}
-Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the flat-histogram methods.
-All of the update factors initially start at $\gamma_{t=0} = 1$.
-SAD dynamically updates $\gamma_t$ throughout the simulation.
-After about $10^{13}$ moves, $\gamma^{\text{SAD}}_t$ proceeds as $1/t$.
-Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown.
-The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$.
-The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
-All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.
-
-Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a function of time for
-this system. The solid/dashed lines
-represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
-over eight simulations using different random number seeds. The range of maximum
-errors for each simulation is shown as a shaded region. By
-the time $10^{13}$ moves have been made all but the WL simulation have begun to
-converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$ moves.
-
-\section{Conclusion}
-The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter $\gamma_{t}^{\min}$ makes the method somewhat more difficult to implement in practice.
-
-We find that both SAD, 1/t-WL, and production WL (provided an ideal minimum update factor before beginning the production run) demonstrate excellent and robust convergence.
-They all converge more rapidly than SAMC, and unlike WL do not suffer from
-error saturation. We find that for larger systems SAD reduces the updated factor
-more slowly (and conservatively) than 1/t-WL and production WL. This means that SAD will take
-proportionately more moves to converge to the same value as 1/t-WL as system
-size is increased. SAD requires the user to specify a temperature range of
-interest rather than an energy range of interest as the WL methods do.
-The 2D Ising model is a system where the energy range is known \emph{a priori.}, which makes a comparison between WL methods and SAD (which requires a minimum temperature) more biased towards WL methods.
-For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
-considerably more convenient. This work also visually demonstrates that ``production'' WL performs extremely well and is preferable to WL for ensuring both ergodicity and detailed balance.
-
-\section{Acknowledgments}
-
-We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model.
-We also wish to thank Johannes Zierenberg for helpful discussions regarding WL production run simulations and
-Martin Weigel for discussions regarding entropic population annealing.
+\input{ising-body.tex}
 

### 2020-05-02 square-well/sq-well-body.tex
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

--- 2020-01-20 square-well/sad.tex
@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
-\usepackage{graphicx} % Include figure files
-\usepackage{color}
-\usepackage{nicefrac} % Include for inline fractions
-
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
-\usepackage{mdframed}
-
-% define colors for comments
-\definecolor{dark-gray}{gray}{0.10}
-\definecolor{light-gray}{gray}{0.70}
-
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
-
-\newcommand{\davidsays}[1]{{\color{blue} [\green{David:} \emph{#1}]}}
-\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
-\begin{document}
-
-\title{Stochastic Approximation Monte Carlo with a Dynamic Update
-Factor
-}
-
-\author{Jordan K. Pommerenck} \author{Tanner T. Simpson}
-\author{Michael A. Perlin} \author{David Roundy}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-
-\begin{abstract}
-  We present a new Monte Carlo algorithm based on the Stochastic
-  Approximation Monte Carlo (SAMC) algorithm for directly calculating
-  the density of states. The proposed method is Stochastic
-  Approximation with a Dynamic update factor (SAD)
-  which dynamically adjusts the update factor $\gamma_t$ during the course of
-  the simulation. We test this method on the square-well fluid and
-  the 31-atom Lennard-Jones cluster and
-  compare the convergence behavior of several
-  related
-  Monte Carlo methods. We find that both the SAD and $1/t$-Wang-Landau ($1/t$-WL)
-  methods rapidly converge to the
-  correct density of states without the need for the user to specify an
-  arbitrary tunable parameter $t_0$ as in the case of SAMC.  SAD requires
-  as input the temperature range of interest, in contrast to
-  $1/t$-WL, which requires that the user identify the interesting range
-  of energies.
-  %
-  The convergence of the $1/t$-WL method is very sensitive to the energy
-  range chosen for the low-temperature heat capacity of the
-  Lennard-Jones cluster.
-  %
-  Thus, SAD is more powerful in the common case in which the range
-  of energies is not known in advance.
-\end{abstract}
-
-\maketitle
-
 \section{Introduction}

### 2020-05-02 square-well/sq-well.tex
gamma-plot is now correctly building the figures.
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

--- 2020-01-20 square-well/sad.tex
@@
 
-\section{Introduction}
-Over the past several decades, a number of flat histogram Monte Carlo simulation
-algorithms have been developed which calculate the thermodynamic properties of
-various systems over a range of temperatures.  This development began with the
-original histogram method, which used a single canonical Monte Carlo simulation
-to predict properties for nearby temperatures~\cite{ferrenberg1988new}.  For
-large systems, this approach is limited to a narrow temperature range because a
-single canonical simulation explores only a small range of energies. Berg and
-Neuhaus developed multicanonical methods that introduced a weight function to
-enable flat histogram sampling which improved the exploration of configuration
-space and allowed the system to overcome free energy
-barriers~\cite{berg1991multicanonical, berg1992multicanonical}. These works led
-to an explosion in the interest of ``flat'' (or ``broad'') histogram
-methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-wang2001determining, wang2001efficient, landau2004new, schulz2003avoiding,
-yan2003fast, trebst2004optimizing, belardinelli2007wang, belardinelli2007fast,
-belardinelli2008analysis, belardinelli2014intrinsic, singh2012density,
-zhou2008optimal, schneider2017convergence, liang2006theory, liang2007stochastic,
-liang2009improving, werlich2015stochastic, kim2006statistical,
-kim2007statistical, kim2009replica, junghans2014molecular}, which explore a
-wider range of energies. In addition to simulating a range of temperatures, in
-contrast with low-temperature canonical Monte Carlo, these approaches
-avoid being trapped in a local energy minimum.
+\input{sq-well-body.tex}
 
-Wang and Landau introduced the most widely used flat histogram algorithm (WL)
-that uses an update factor and a statistical histogram to compute the density of
-states of a given system~\cite{wang2001determining, wang2001efficient}.  While
-the method is incredibly powerful, it has a few disadvantages. The most
-significant disadvantage is that the method requires the user to select the
-range of energies to be studied~\cite{landau2004new, wang2001efficient,
-schulz2003avoiding, yan2003fast}. This requirement adds an additional
-hurdle to its application to systems for which the interesting range of energies
-is not known \emph{a priori}.  The simulation violates detailed balance, albeit
-briefly as the size of the violation decreases with time, which complicates
-convergence analysis.  In fact, the error in a WL computation has been
-demonstrated to saturate at a non-zero value~\cite{yan2003fast}, i.e. the method
-does not converge to the true density of states~\cite{belardinelli2007wang,
-belardinelli2007fast, belardinelli2008analysis, belardinelli2014intrinsic,
-singh2012density, zhou2008optimal}.
-
-Belardinelli and Pereyra demonstrated that allowing the update factor to
-decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang},
-where $t$ corresponds to the number of moves.
-This leads to their $1/t$-WL algorithm which ensures that the error continues to
-decrease asymptotically as $1/\sqrt{t}$, which they demonstrated avoids
-error saturation and asymptotically approaches the true density of
-states~\cite{belardinelli2008analysis}. Zhou \emph{et al.} further confirmed that the
-original WL algorithm never converges exponentially and successfully bounded the
-statistical error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}.
-Schneider \emph{et al.} outline minor refinements to the $1/t$-WL algorithm
-including scaling the update factor with the number of energy
-bins~\cite{schneider2017convergence}.
-
-Liang began to consider Stochastic Approximation as a mathematical
-generalization of the WL approach, with convergence that could be mathematically
-proven~\cite{liang2006theory}. In 2007, Liang \emph{et al.}
-developed Stochastic Approximation Monte Carlo (SAMC)~\cite{liang2007stochastic,
-liang2009improving}, and proved its convergence, although the method still has a
-system specific user-defined parameter which must be tuned when applying this
-algorithm to a new system.  In contrast to the WL-based methods, the SAMC method
-does not require users to determine the energy range of interest \emph{a
-priori}. Werlich \emph{et al.} proposed the introduction of an additional tuning
-parameter into SAMC~\cite{werlich2015stochastic}.
-
-Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
-related Statistical Temperature Molecular Dynamics (STMD) which is an
-adaption of the WL method that approximates the entropy (or natural logarithm of the density of
-states) as a piecewise linear function, which improves convergence for systems
-with a continuously varying energy~\cite{kim2006statistical,
-kim2007statistical}. STMC applied to WL requires a temperature range be
-specified rather than an energy range.  Kim \emph{et al.} extended this work as
-Replica Exchange Statistical Temperature Monte Carlo (RESTMC), which uses
-replica exchange of multiple overlapping STMC simulations to improve
-convergence~\cite{kim2009replica}. More recently, Junghans \emph{et al.} have
-demonstrated a close connection between metadynamics, which was introduced by
-Laio and Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods,
-with STMD forging the connection~\cite{junghans2014molecular}.
-
-In this work, we have developed an improved algorithm based on SAMC that does
-not require an array of non-physical, user-defined inputs and therefore should
-be easily applicable to any system. The method (like STMC above) does require
-the user to define a temperature range of interest ($T_\text{min}$ to
-$T_\infty$) which we explain in Section~\ref{sec:sad}.  We call this method SAD (Stochastic Approximation with a Dynamic
-update factor), and will discuss it in detail in the methods section. We compare
-its convergence properties with three existing flat histogram methods: WL,
-$1/t$-WL, and SAMC.
-
-In this work, we compare four flat histogram methods.  We outline the
-general workings of each algorithm that we developed in detail while
-summarizing algorithms that were developed in other works.  The
-following methods are discussed and simulated for the square-well
-fluid and the 31-atom Lennard-Jones cluster: Wang-Landau (WL),
-$1/t$-Wang-Landau ($1/t$-WL), Stochastic Approximation Monte Carlo
-(SAMC), and SAD.
-
-\section{Flat histogram methods}\label{sec:histogram}
-The goal of flat histogram methods (also called \emph{broad histogram}
-or \emph{multicanonical} methods) is to simulate each energy with
-similar accuracy so as to accurately determine the density of states
-over a broad range of energies---and thus to determine the thermodynamic
-quantities such as heat capacity or internal energy
-over a broad range of temperatures.
-Properties that require more information---such as a spatial
-correlation function or a response function---can still be computed
-for any temperature, provided statistics are collected for each
-individual energy, which can then be reweighted for any
-temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
-errington2003direct}.
-
-All the flat histogram Monte Carlo methods begin with randomly chosen
-``moves'' which change the state of the system and must satisfy
-detailed balance.  Each algorithm differs in how it determines the
-probability of accepting a move and in what additional statistics must
-be collected in order to decide on that probability.
-
-Flat histogram methods calculate the density of states $D(E)$ for a discrete set
-of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
-trebst2004optimizing}. Therefore, energy binning becomes an important
-consideration for systems with a continuum of possible energies.  Energy bins
-are typically of uniform size for the entire energy
-continuum~\cite{fasnacht2004adaptive}. Some methods such as
-AdaWL~\cite{koh2013dynamically} employ a tunable mechanism for controlling the
-binning for low entropic states in order to ensure the exploration of all
-energies.  The method introduced in this paper is designed to scale appropriately
-as bin size is changed.
-
-In this section we will introduce four closely related flat histogram
-methods each of which rely on a weight function $w(E)$.  In these
-algorithms, the probability of accepting a move is given by
-\begin{equation}
-	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
-	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
-\end{equation}
-which biases the simulation in favor of energies with low weights.  A set of
-weights that are proportional to the density of states $D(E)$ of the system will
-result in an entirely flat histogram, which means that the weights should
-converge to being proportional to the density of states.  The natural logarithm
-of the weight is typically stored, since the density of states will often vary
-over a few hundred orders of magnitude. In the microcanonical ensemble, the
-entropy is defined as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the
-logarithm of the weight is an approximation of the entropy.
-
-Each approach uses a random walk in energy space to estimate the density of
-states.  The core of these approaches is to update the weights at each step of
-the simulation
-\begin{equation}
-	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
-	+\gamma_t
-\end{equation}
-where $t$ is the number of the current move, $\gamma_t$ is an update factor
-that varies over the simulation, and $E$ is the current energy.  This update
-causes the random walk to avoid energies that have been frequently sampled,
-enabling a rapid exploration of energy space. This approach violates detailed
-balance, due to the acceptance probabilities changing with each move, but the
-severity of this violation decreases as we decrease $\gamma_t$.  The four
-methods differ primarily in how they schedule the decrease of $\gamma_t$.
-
-\subsection{Wang-Landau}
-
-The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
-landau2014guide} begins with $\gamma^{\text{WL}}_{t=0}=1$, and then decreases
-$\gamma^{\text{WL}}$ in discrete stages.  When the energy histogram is
-sufficiently flat, $\gamma^{\text{WL}}$ is decreased by a specified factor of
-$\frac12$.  The flatness is defined by the ratio between the minimum value of
-the histogram and its average value.  When this flatness reaches a specified
-threshold (typically 0.8), the $\gamma^{\text{WL}}$ value is decreased and the
-histogram is reset to zero.  This approach requires that the energy range of
-interest be known in advance, and difficulties can occur with this flatness
-criteria due to the fact that some energies in this energy range might never be
-sampled~\cite{troster2005wang}.  The entire process is repeated until
-$\gamma^{\text{WL}}$ reaches a desired cutoff.
-
-The Wang-Landau approach thus has three parameters that need be
-specified: the factor by which to decrease $\gamma^{\text{WL}}$ when flatness is
-achieved, the flatness criterion, and the cutoff that determines when
-the computation is complete.  In addition, an energy range (or in
-general, a \emph{set} of energies) must be supplied, so that the
-flatness criterion can be defined.
-
-\subsection{$1/t$-Wang-Landau}
-
-The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
-from dropping below $N_S/t$~\cite{belardinelli2008analysis,
-schneider2017convergence}. The method follows the standard WL algorithm with two
-modifications.  Firstly, the histogram is considered flat, and $\gamma_t$ is
-decreased by a factor of two, when every energy state has been visited once,
-i.e. when the WL ``flatness'' becomes nonzero. Secondly, when
-$\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the algorithm switches to use
-$\gamma_t = N_S/t$ for the remainder of the simulation:
-\begin{align}
-  \gamma_t^{1/t\text{-WL}} = \begin{cases}
-     \gamma^{\text{WL}}_t & \gamma^{\text{WL}}_t > \frac{N_S}{t} \\
-     \frac{N_S}{t} & t \ge t_0
- \end{cases}
-\end{align}
-where $t$ is the number of moves, $\gamma^{\text{WL}}_t$ is the Wang-Landau update factor
-at move $t$, and $N_S$ is the number of energy bins.
-
-\subsection{SAMC}
-The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses the lack of
-convergence of Wang-Landau's approach with a simple schedule by which the update
-factor $\gamma_t$ is continuously decreased~\cite{liang2007stochastic,
-werlich2015stochastic, schneider2017convergence}.  The update factor is defined
-in the original implementation~\cite{liang2007stochastic} in terms of an tunable
-parameter $t_0$,
-\begin{align}
-\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}\label{eq:1}
-\end{align}
-where as above $t$ is the number of moves that have been attempted.
-SAMC offers extreme simplicity, combined with is proven convergence.
-Provided the update factor satisfies
-\begin{align}
-\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
-\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
-\end{align}
-where $\zeta > 1$, Liang has proven that the weights converge to the true
-density of states~\cite{liang2006theory, liang2007stochastic,
-liang2009improving}.  In addition, the energy range need not be known \emph{a priori}.
-The time to converge depends only on the choice of parameter $t_0$.
-Unfortunately, there is no prescription for finding an acceptable value for
-$t_0$, and while the algorithm formally converges, for a poor choice of $t_0$
-that convergence can be far too slow to be practical. Liang \emph{et al.} give a
-rule of thumb in which $t_0$ is chosen in the range from $2N_S$ to $100N_S$
-where $N_S$ is the number of energy bins~\cite{liang2007stochastic}. Schneider
-\emph{et al.} found that for the Ising model this heuristic is helpful for small
-spin systems, but that larger systems require an even higher $t_0$
-value~\cite{schneider2017convergence}.  We will describe below one case we
-examined, in which $t_0$ needs to be as much as two orders of magnitude higher
-than the rule of thumb of $100N_S$ in order to converge in $10^{12}$ moves.
-
-Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t^{\text{SA}}$
-by a factor $\gamma_0$~\cite{werlich2015stochastic}.  While this may result in
-an improved rate of convergence, it adds yet another parameter that must be
-empirically determined, and we have not explored this additional degree of
-freedom.
-
-\subsection{SAMC convergence time}\label{sec:samc-convergence}
-A primary difficulty in using the SAMC method lies in identifying an
-appropriate value for $t_0$.  Although SAMC is proven to formally
-converge regardless of the $t_0$ value, a choice that is either too
-high or too low will result in prohibitively slow convergence to the true
-entropy of the system.  It is instructive to consider separately
-values of $t_0$ that are too low or too high.
-
-We can place a rigorous \emph{lower} bound $t_{\min}$ on the number of
-moves required to find the true entropy by considering the total change
-that needs to be made to the entropy:
-\begin{align}
-  \Delta S_{\text{tot}} \equiv \sum_E S(E) - S_{\min}.
-\end{align}
-This connects with $\gamma_t$ because the $\ln w(E)$ (our approximation for entropy)
-has $\gamma_t$ added to it on each move.  Thus a minimum number of moves
-that could possibly result in the true entropy $S(E)$ starting with a flat set of
-weights is determined by the total entropy change required.  We can estimate
-this number of moves required by summing the $\gamma_t$,
-which we can approximate using an integral:
-\begin{align}
-   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t^{\text{SA}} \\
-  % &= \int_0^{t_{\min}} \gamma(t)dt \\
-  &\approx t_0 + \int_{t_0}^{t_{\min}} \frac{t_0}{t}dt
-  \\
-  &= t_0\left(1 + \ln\left(\frac{t_{\min}}{t_0}\right)\right)
-\end{align}
-Solving for ${t_{\min}}$ we find that
-\begin{align}
-  {t_{\min}} &= t_0 e^{\frac{\Delta S_{\text{tot}}}{t_0} - 1}
-\end{align}
-which means that the minimum time to converge grows exponentially
-as $t_0$ is made smaller.  You \emph{seriously} don't want to underestimate
-$t_0$!
-
-One might reasonably choose to err by selecting a large $t_0$. The rate of
-convergence is harder to estimate when $t_0$ is large, but in general
-$\gamma_t^{\text{SA}}$ itself forms a lower bound on the accuracy with which the
-entropy may be known, with an unknown prefactor which is related to the
-coherence time ($t = t_0$) of the Monte Carlo simulation.  Since
-$\gamma_t^{\text{SA}}$ is given by $t_0/t$, the time to converge to a given
-accuracy is increased in proportion to the ratio by which we overestimate $t_0$.
-Thus, while it is exponentially painful to underestimate $t_0$, overestimating
-by several orders of magnitude is also not acceptable.  We should note that
-these extreme limiting cases do not preclude the possibility that there is a
-wide range of $t_0$ values that lead to an acceptable convergence rate.
-
-\section{SAD Algorithm}\label{sec:sad}
-The Stochastic Approximation with Dynamic update factor (SAD) method
-is a variant of the SAMC Algorithm that attempts to dynamically choose
-the modification factor rather than relying on system dependent
-parameters such as $t_0$ or $\gamma_0$.  There is an immediate
-advantage of such an algorithm where parameters are chosen independent
-of system size or type. Each flat-histogram method has unique
-advantages and disadvantages.  Wang-Landau and $1/t$-WL require an
-energy range for initialization.  SAMC removes this energy range
-requirement but requires simulating every possible energy. Our
-proposed method SAD requires the user to input $T_\text{min}$, the
-lowest temperature of interest, which is an immediate disadvantage of
-the method. However, identifying a minimum temperature of interest
-$T_\text{min}$ may be easier for a user than determining in advance
-an energy range of interest or the unphysical parameter $t_0$.
-
-We set the maximum temperature of interest as $T=\infty$ for a couple
-of reasons.  First, when this temperature is reached, the system is
-maximally randomized, so allowing infinite temperature further ensures
-that the system can surmount an energy barrier of any size.  Second,
-the high temperatures tend to be ``easy'' to simulate, and tend to
-converge very quickly.  We note, however, that this is under the
-assumption that infinite temperature corresponds to a finite energy,
-as is the case for configurational sampling.  Were we to include
-kinetic energy in a simulation (as in molecular dynamics), or to
-examine a system such as a high density Lennard-Jones fluid with an
-entropy maximum at an energy that is far from the energy range of
-physical interest, then we would need to apply either a maximum
-temperature or a maximum energy, either of which could slow down the
-rate at which state space is explored.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{N50-lndos-comparison}
-  \caption{The entropy of a square-well fluid with 50 atoms and filling fraction
-        0.3 as the green line.  The green hatched area reflects the
-        minimum entropy change needed to converge to the true value.
-        The light blue area is the quadratic approximation
-        for the change in entropy.  The vertical dotted lines represent
-        the energy corresponding to $T=1/3$ and $T=\infty$.}
-  \label{fig:entropy-cartoon}
-\end{figure}
-
-While for SAMC, the update factor is defined in the original
-implementation (see equation~\eqref{eq:1}), for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
-thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
-the SAMC parameter
-$t_0$ should have dimensions of entropy.
-We begin with an estimate of the average value of the entropy (relative
-to the lowest entropy at $T_{\min}$).  If we assume a quadratic
-dependence on energy (see Fig.~\ref{fig:entropy-cartoon}), this is given by
-\begin{align}
-\langle S\rangle \approx \frac13 \frac{E({T=\infty}) - E(T_{\min})}{T_{\min}}
-\end{align}
-We approximate this energy difference by $E_H -E_L$ where $E_H$ and
-$E_L$ are defined below. The entropy numerator of the update factor in
-general should scale with the total number of interesting energy states
-$N_S$, since updates to the weights are distributed between that many
-energy states.  The product $N_S\langle S\rangle$ is the total change
-of entropy required (starting from constant weights) to find the true
-entropy, and puts a lower bound on the convergence time. After
-\emph{long} times, when all the energies have been explored a long time ago, we
-wish for a lower update factor in order to more rapidly refine the
-remaining error in entropy.  We track the time at which we first visited
-each possible energy.  We define $t_L$ to be the last time that we
-encountered an energy that we currently believe is in the energy range
-of interest, so a $t\gg t_L$ we feel confident that we have established
-the true energy range of interest. We gradually transition to a lower
-update factor (but still asymptotically scaling as $\gamma_t \propto
-1/t$ to ensure eventual convergence).  Finally, we wish for an update
-factor that is \emph{never} greater than 1, because a very large update
-factor could introduce very large errors in entropy that may take many iterations
-to remove.  The SAD expression for $\gamma_t$ which incorporates these
-ideas is:
-\begin{align}
-  \gamma_{t}^{\text{SAD}} =
-     \frac{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
-     }{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
-     }
-\end{align}
-where $E_H$ and $E_L$ are the current estimates for the highest and
-lowest energies of interest as defined below.  This factor
-asymptotically has the same $1/t$ behavior as the original SAMC
-algorithm and with the same $N_S$ prefactor used by the $1/t$-WL
-method; however for earlier values of $t$, the update factor drops as
-$1/t^2$ and jumps every time a new energy is determined to be of
-interest.  This behavior allows SAD to dynamically prevent the update
-factor from decreasing too rapidly.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{gamma-n256}
-  \caption{The update factor $\gamma_t$ versus iteration for the
-    square-well fluid with 256 atoms, for four different methods: WL, $1/t$-WL, SAMC,
-    and SAD.}
-    \label{fig:gamma-vs-t}
-\end{figure}
-Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for the related
-methods SAD, WL, $1/t$-WL, and SAMC.  For SAMC, $\gamma_t$ remains
-constant before dropping as $1/t$.  WL $\gamma_t$ remains at 1 for
-many iterations, and then decreases very rapidly, with $1/t$-WL
-behaving similarly but decreasing more aggressively before
-transitioning to a more conservative $1/t$ behavior.  The update
-factor for SAD fluctuates dynamically around a value less than 1 for
-early MC moves, and then decreases as approximately $1/t$ while
-continuing to fluctuate as new energies are found to be important.  At
-intermediate times, the SAD $\gamma_t$ decreases as $1/t^2$ before
-asymptoting to $N_S/t$, which is the same as $1/t$-WL.
-
-Since SAD does not explore \emph{all} energy states, it needs to
-determine what energy range corresponds to the temperature range of
-interest defined by $T_{\min}<T<\infty$. The simulation is responsible
-for determining and updating this energy range. Given the true entropy
-$S(E)$, we can define the interesting energy range as
-%\begin{align}
-  $E(T_{\min}) <E< E(T=\infty)$
-%\end{align}
-where $E(T)$ is the energy that maximizes $S-E/T$.  During the course of the
-simulation, this precise energy is challenging to evaluate accurately.
-In order to ensure that we sample this entire energy range adequately,
-we define two energy limits:  a high energy $E_H$ and a low
-energy $E_L$, which define the range over which the energy histogram
-is made flat. At move $t$, $E_H$ and $E_L$ are the greatest and lowest
-energy that prior to that move
-had the highest histogram value (i.e. been visited the most times) at some point
-during the course of the simulation.
-This definition results in a ``ratcheting'' effect, in which $E_H$
-may only increase, while $E_L$ may only decrease over the course of the
-simulation, which results in a conservative estimate of the range of
-energies that need be sampled.
-
-During the simulation when considering a move inside the energy
-range of interest $E_L \le E \le E_H$, the weights are used as in the
-three methods already described.  If $E\ge E_H$, the weight is taken
-to be
-\begin{align}\label{eq:highw}
-  w(E>E_H) &= w(E_H),
-\end{align}
-which corresponds to an infinite temperature.  This choice ensures
-that if the maximum in entropy is at an energy $E_{\max}>E_H$, then
-the energy $E_{\max}$ will eventually have the highest number of
-counts and the ratcheting will result in $E_H\ge E_{\max}$ .  At lower
-energies, Boltzmann weights corresponding to the minimum temperature
-are used:
-\begin{align}\label{eq:loww}
-  w(E<E_L) &= w(E_L)e^{-\frac{E_L-E}{T_{\min}}}.
-\end{align}
-This choice has the result that if the energy $E_{\min}$ at which the free
-energy at $T_{\min}$ is minimized is less than $E_L$, the lower energy
-limit will ratchet down to include $E_{\min}$.
-Each time we change the value of $E_H$ or $E_L$, the weights outside the
-new portion of the interesting energy range are set to the expressions
-in Equations~\ref{eq:highw} and~\ref{eq:loww}.
-
-%~ \paragraph{Energy binning with SAD}
-
-A significant advantage of SAD over SAMC---which the $1/t$-WL and WL
-methods share \emph{after} they have discovered all the energies---is that
-the schedule for $\gamma_t$ automatically responds to the choice of bin
-size.
-SAD should perform similarly over a reasonable range of bin sizes
-because $\gamma_t \propto \nicefrac{N_S}{t}$.  As the number of energy
-states $N_S$ found increases (fine binning), the time spent $t$ in
-each bin will decrease with the effect that the convergence should be
-roughly independent of the bin size chosen.  SAMC could be used with a
-prefactor $\gamma_0$ to aid in a similar
-way~\cite{werlich2015stochastic} but this adds yet another parameter
-for the user to choose.
-
-%~ \paragraph{Scaling, step size, and SAD}
-%~ Methods such as WL can fail to converge for complex systems in part
-%~ because the same random MC moves are performed for the entire energy
-%~ range.  The main problem lies in trying to determine how to optimize
-%~ the step size such that the time to explore all energies is reasonable.
-%~ Small step sizes lead to improved sampling at the expense of slow
-%~ convergence. Koh et al. proposed multiplying $\gamma$ by an energy
-%~ dependent factor in order to improve the converge of WL in regards to
-%~ convergence and scaling~\cite{koh2013dynamically}.  Because the update
-%~ factor for SAD has an energy dependent factor already present (and
-%~ dynamically adjusted), we expect that SAD will perform consitently
-%~ scaling free for a reasonable choice of step size.  SAD should perform
-%~ extremely well compared with other flat histogram methods when applied
-%~ to complex continuous systems such as polymers/proteins and spin models.
-
-%~ \subsection{Convergence}
-%~ Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
-%~ convergence criteria.  However, we also want to ensure that SAD avoids
-%~ the exponentially poor convergence time that SAMC suffers when $t_0$ is
-%~ small as discussed in Section~\ref{sec:samc-convergence}.
-%~ To place a lower bound on the convergence of SAD, we will again
-%~ approximate the total entropy change using an integral.  To sidestep
-%~ the dynamical nature of SAD, we ensure that \emph{after discovering the
-%~ last important energy} at time $t_L$ the entropy change
-%~ $\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
-%~ Thus we integrate the update factor from $t_L$ to some minimum time
-%~ $t_{\min}$ necessary for convergence.  Technically this is not a lower
-%~ bound as in the case of SAMC because the algorithm will already have a
-%~ reasonable approximation for the entropy.
-%~ \begin{align}
-%~ \Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
-     %~ \frac{
-       %~ 3\langle S\rangle + \frac{t}{t_L}
-     %~ }{
-       %~ 3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
-     %~ } dt \\
-%~ &\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
-%~ \end{align}
-%~ where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
-%~ that it takes a long time to find all the interesting energy states.
-%~ Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
-%~ of order unity.  This omission only makes our bound more conservative,
-%~ and enables us to solve to find that
-%~ \begin{align}
-  %~ t_{\min} &= \frac32 t_L
-%~ \end{align}
-%~ which means that we only need to run fifty percent longer after we
-%~ discover the final interesting energy state in order to have made
-%~ sufficient updates to \emph{possibly} converge to the true entropy.
-
-\section{Results}\label{sec:results}
-
-\subsection{Square-well fluid}
-As our first test case, we consider the square-well fluid i.e. a
-system of particles whose interactions are governed by a square-well
-potential~\cite{singh2003surface, barker2004perturbationSW}.  The
-square-well potential is an ideal test-bed as it is a simple model for
-a liquid, which includes both attractive and repulsive
-interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
-To date, there have not been any published direct convergence comparison tests
-for flat histogram methods applied to the square-well fluid.
-The potential $U(\textbf{r})$ for such a system is given by
-\begin{equation}
- U(\textbf{r})=\begin{cases} \infty &
- \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
- \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
- \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
-\end{equation}
-where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
-reduced range of the potential well, and $\epsilon$ is its depth. This model has
-the further advantage that binning is not required because the energy is
-discrete.
-
-\begin{figure*}
-  \includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
-\includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
-  \caption{(a) The average entropy error for each MC method for $N=50$,
-               $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run. The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded area, and
-           (b) the update factor $\gamma_t$ versus iteration number
-               for the same simulations.
-           (c) The average entropy error for each MC method for the
-               same physical system with a smaller displacement distance
-               $\delta_0 = 0.005\sigma$,
-               as a function of number of iterations run, and
-           (d) the update factor $\gamma_t$ versus iteration number
-               for the same simulations.
-  }\label{fig:fast-slow-gamma}
-\end{figure*}
-
-We tested the algorithms on two square-well fluid systems.  The first
-is a smaller simulation with a particle number of 50, a well-width of
-$\lambda = 1.3$, and a volume corresponding to a filling fraction
-(defined as the fraction of volume filled by atoms) of $\eta =
-0.3$. The second system is larger, with a particle number of 256, a
-well-width of $\lambda = 1.5$, and a volume corresponding to a filling
-fraction of $\eta = 0.17$.  For each system we use a reasonable
-root-mean-square displacement distance $\delta_0 = 0.05\sigma$ for
-proposed moves, and for the smaller system we also use an unreasonably
-small displacement distance of $0.005\sigma$. The simulations explore
-the energy space of the systems with minimum reduced temperatures of
-$T_{\text{min}} = 1/3$ for simulations of the smaller system, and
-$T_{\min}=1$ for the larger system.  All simulations lead to the
-minimum important energy $E_{\min}$ and maximum entropy energy
-$E_{\max}$ being calculated (with the exception of the WL methods
-where both of these parameters are needed \emph{a priori}).
-
-The SAMC simulations
-computed the density of states for the entire range of possible
-energies.  The SAD simulations determined the energy range of interest
-dynamically as described above, based on a specified $T_{\min}$.  For
-the WL and $1/t$-WL simulations, we constrained the simulation to
-remain in the energy range corresponding to $T_{\min} < T < \infty$,
-as determined by a previous SAMC simulation.  Thus the WL and $1/t$-WL
-simulations were given extra information that in practice would not be
-available without additional computational effort, and the SAMC simulations
-computed the entropy over the entire range of possible energies, which
-required more effort.
-
-We use the average entropy error versus moves as a metric to compare
-simulation runtimes and overall convergence. The overall accuracy
-is determined by examining the fractional error of a particular method to
-a precise reference system. For each simulation, the reference system
-is chosen to be the final output of a SAMC simulation with a fixed energy range
-corresponding to the temperature range of interest.
-Although SAMC does not require an energy range as an input parameter,
-we find that by limiting the simulation to this energy range, we can
-achieve much faster convergence with a smaller $t_0$.  We compute an average
-of the error by averaging the error in the entropy over the
-interesting energy range, and then averaging \emph{this} error over several
-simulations run with different random number seeds.
-%~ \begin{align}
-%~ \epsilon_\text{avg} =
-  %~ \frac{1}{N_S \text{num seeds}}\sum_s^{\text{num seeds}}\sum_{i}^{N_S} \left|S_\text{ref}(E_i) - S_\text{method}(E_i)\right|
-%~ \end{align}
-
-\subsubsection{50 square-well atoms}
-For this 50 atom simulation, we chose a minimum reduced
-temperature of $1/3$, which corresponds to an interesting energy range
-from $-248$ to $-120$.  The number of important energy states
-for this system is therefore $N_S = 129$.  The entropy of this system is shown in
-Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
-energy range the entropy differs by 198, corresponding to a ratio of
-$10^{86}$ between the highest and lowest density of states.
-
-In order to explore the effect of simulation details on convergence,
-we consider two values for the displacement distance by which atoms
-are moved during a Monte Carlo step.  We began with a reasonable
-displacement distance of $\delta_0 = 0.05\sigma$, which corresponds to
-an acceptance rate of proposed moves of 38\%.  We further ran
-simulations with a much smaller displacement distance of $\delta_0 =
-0.005\sigma$, which resulted in an acceptance rate of 86\%, which
-converged more slowly.
-
-Figure~\ref{fig:fast-slow-gamma}a shows the average error in the
-entropy as a function of time for this system with the reasonable
-displacement distance of $\delta_0 = 0.05\sigma$.  The solid/dashed lines
-represent the average of the absolute value of the error in the
-entropy averaged over eight simulations using different random number
-seeds.  The range of average errors for each simulation is shown as a
-shaded region around its mean error.  By the time $10^8$ moves have
-been made all but the SAMC simulation with the shortest $t_0$ have
-begun to converge as $1/\sqrt{t}$.  We then see the WL error
-saturate around $10^{10}$ moves.
-
-Figure~\ref{fig:fast-slow-gamma}c shows the average error in
-the entropy as a function of time for this system with the
-unreasonably small displacement distance of $\delta_0 = 0.005\sigma$.
-%
-The smaller translation scale causes all methods to take additional time to
-explore all energies. Based on random walk scaling, the convergence time of an
-ideal method should scale roughly as $\delta_0^{-2}$ in the limit of small
-$\delta_0$, that is, one order of magnitude decrease in the displacement
-distance should result in two order of magnitude increase in convergence time.
-SAMC simulations with a $t_0$ value that rapidly converged for $\delta =
-0.05\sigma$ do not converge at all in $10^{12}$ moves for a translation scale of
-$\delta = 0.005\sigma$. It is also worth noting that for the smaller
-displacement distance, the SAMC rule of thumb of choosing $t_0$ to be
-approximately $100N_S$ is no longer valid. SAD, WL, and $1/t$-WL handle the
-shift in displacement distance and converge roughly as expected.
-
-The methods SAD, WL, and $1/t$-WL compensate for the smaller
-displacement distance by reducing $\gamma_t$ more slowly, as can be seen from
-Figure~\ref{fig:fast-slow-gamma}b and ~\ref{fig:fast-slow-gamma}d.
-The update factors take approximately 10$\times$ longer to reach steady-state
-for the smaller displacement distance. Because of this update behavior,
-these methods are less sensitive to the choice of displacement
-distance than SAMC is.
-
-%~ SAMC methods that converge
-%~ in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not
-%~ necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.
-%
-%~ Sad performs similarly regardless of the displacement distance chosen although
-%~ it transitions to $1/\sqrt{t}$ convergence after
-
-\begin{figure}
-\includegraphics[width=\columnwidth]{periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
-  \caption{
-  The average entropy error for each MC method for $N=256$,
-               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
-               as a function of number of iterations run.  The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded area.  The update factor for this
-               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
-\end{figure}
-
-\subsubsection{256 square-well atoms}
-Next we introduce a considerably larger simulation containing 256
-atoms which has a maximum entropy about 1500 greater than its minimum.
-This makes exploring the entire range of energies extremely expensive,
-and strongly favors the methods that restrict the energy (or
-temperature) range of interest.  For this simulation, we chose a much
-higher minimum reduced temperature of $1.0$, which corresponds to an
-interesting energy range from $-915$ to $-509$.  The number of
-important energy states for this system is then $N_S = 407$.  The
-minimum entropy over this energy range is just 395 less than the
-maximum, corresponding to a ratio of only $10^{118}$ between the
-highest and lowest density of states.
-
-%$395.2833013702813 - 4.754731035740479e^{-8}$ = $395.2833$ so $2^{396.2833}$
-%gives us about $10^{118}$ from 'samc-1e4-256-s8.yaml'
-
-Figure~\ref{fig:n256} shows the average error in the entropy as a
-function of moves for this system with the reasonable displacement
-distance of $\delta_0 = 0.05\sigma$.  The solid lines represent the
-average of the absolute value of the error in the entropy averaged
-over eight simulations using different random number seeds.  The range
-of average errors for each simulation is shown as a shaded region
-around its mean error.  By the time $10^8$ moves have been made all
-but the SAMC simulation with the shortest $t_0$ shown have begun to
-converge as $1/\sqrt{t}$.  We then see the WL error saturates around
-$10^{10}$ moves.  Once again, the convergence of SAD is essentially
-the same as that of the $1/t$-WL method.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{lj-cv}
-  \caption{The heat capacity of the LJ31 cluster after $10^{12}$ moves
-    as calculated with each method.  Each simulation used a bin size
-    of $\Delta E=0.01\epsilon$.}
-  \label{fig:lj-cv}
-\end{figure}
-
-\subsection{31-atom Lennard-Jones cluster}
-
-We also examined convergence of each algorithm on the 31-atom cluster
-(LJ31), which is the smallest Lennard-Jones cluster that exhibits
-broken ergodicity and a solid-solid phase transition at a low
-temperature~\cite{doye1998thermodynamics,
-  martiniani2014superposition}.  Clusters of Lennard-Jones atoms have
-been frequently used for testing Monte Carlo
-algorithms~\cite{wales1997global, neirotti2000phase,
-  frantsuzov2005size, mandelshtam2006multiple}.  The LJ31 cluster
-is commonly used for testing convergence of algorithms
-because it features a two-funnel energy landscape with a significant
-barrier between the two low-energy states~\cite{calvo2000entropic,
-  calvo2000phase, poulain2006performances,
-  martiniani2014superposition}.  In particular, the LJ31 cluster
-features a low-temperature solid-solid phase transition that is challenging to
-converge~\cite{mandelshtam2006multiple}.  Poulain \emph{et al.}
-tested a number of variants of the WL algorithm on LJ31 and failed to
-converge the spike in heat capacity corresponding to the solid-solid
-phase transition~\cite{poulain2006performances}.
-
-%% \jpcom{This \P\ should be commented once summarized.
-%% Poulain \emph{et al.} found very poor convergence for a variety of
-%% modified WL methods on the LJ31
-%% cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
-%% energy range or bin size that was used in their simulation, so their
-%% results cannot be reproduced.  They do mention that they use 1000
-%% energy bins, which if they used an energy range of
-%% $[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
-%% E\approx 0.13\epsilon$, which we find to be too large to accurately
-%% predict the heat capacity peak at $0.03\epsilon/k_B$.  Looks like they
-%% only ran $\sim 10^{10}$ moves. We can't converge in that few moves either.}
-
-Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster
-over the temperature range from $k_BT=0.01\epsilon$ to
-$k_BT=0.40\epsilon$.  There are two peaks, a large peak around $k_BT
-\approx 0.3\epsilon$ corresponding to melting temperature, and a
-small peak around $k_BT \approx 0.027\epsilon$ for the solid-solid
-transition from a Mackay-to-anti-Mackay (M $\rightarrow$ aM)
-transition~\cite{doye1997thermally, hendy2001molecular,
-  mandelshtam2006multiple}.  We constrain our atoms within a spherical
-box of radius 2.5$\sigma$, as is common in the
-literature~\cite{poulain2006performances, mandelshtam2006multiple}.
-We note that using a larger box (with radius $5\sigma$) has a large
-effect on the melting peak, and a smaller but still significant effect
-on the solid-solid transition.  Since our focus is on the convergence
-rate, we will restrict ourselves to the smaller box size.  In addition, we
-further restricted all our simulations to negative total potential energies.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{lj-cv-error}
-  \caption{The maximum error in the heat capacity of the LJ31 cluster
-    in the temperature range 0.01~$\epsilon\le k_BT \le 0.05\epsilon$.
-    Note that the value of $C_V$ ranges from 90$k_B$ to 120$k_B$, so
-    the smallest (maximum) errors seen are around 1\%.  All WL methods
-    shown were run with a histogram bin width of $\Delta
-    E=0.01\epsilon$.  The converged reference heat capacity was
-    computed with a $1/t$-WL simulation with the energy constrained to
-    a range $-133.53\epsilon\le E\le -110\epsilon$.
-  }
-  \label{fig:lj-cv-error}
-\end{figure}
-
-In order to test the low-temperature convergence, we focused on the
-heat capacity in the temperature range $0.01\epsilon \le k_BT \le
-0.05\epsilon$ (see inset in Fig.~\ref{fig:lj-cv}), which contains
-the solid-solid phase transition.
-%
-We compute the heat capacity from the entropy using a standard
-canonical analysis, which gives
-\begin{align}
-  C_V &= 31\times\frac32k_B + \frac1{k_BT^2}\left< (E- \langle
-  E\rangle)^2\right>
-  \label{eq:cv}
-\end{align}
-where the first term represents the kinetic energy contribution, and
-the second term is the potential energy contribution, with canonical
-averages using the microcanonical entropy to compute multiplicities.
-%
-For our reference heat capacity, we ran a $1/t$-WL simulation with the
-energy constrained to a range $-133.53\epsilon\le E\le -110\epsilon$.
-We chose this energy range through experimentation, as we found the
-convergence rate to be very sensitive to the minimum energy chosen.
-
-%% \jpcom{this is a basically a repeat of above...
-%% The heat capacity changes significantly even at these low temperatures
-%% when the size of the cavity is increased.  We use a 2.5 cavity because
-%% ``everyone else'' does~\cite{poulain2006performances}.  For purposes
-%% of testing the convergence of the algorithm, we don't need to solve
-%% for the true heat capacity.  Furthermore, the heat capacity at fixed
-%% volume is not the most meaningful quantity to predict for a cluster.
-%% }
-We ran simulations with several bin widths using each algorithm we studied
-and show results for $\Delta E=0.1\epsilon$ and
-$0.01\epsilon$.
-When examining the converged results, we saw that each bin size yielded a
-measurably different low temperature heat capacity. The largest
-bin size of $0.1\epsilon$ introduced a maximum error in heat capacity
-of more than $20k_B$ compared with the smaller bin size.  We will focus on
-the smaller bin size, and
-used a bin size of $0.01\epsilon$ for our converged reference heat capacity.
-
-Figure~\ref{fig:lj-cv-error} shows the maximum error in heat capacity
-over the temperature range $0.01\epsilon\le k_BT\le 0.05\epsilon$ as a
-function of the number of Monte Carlo moves performed.
-%
-We explored the convergence behavior of each algorithm at two energy
-bin widths ($\Delta E=0.01\epsilon$ and $\Delta E=0.1\epsilon$).  We
-were unable to converge the WL or $1/t$-WL algorithms with the larger
-bin size.
-%
-The rate of convergence of SAD for the heat capacity seems to be roughly
-independent of the bin width.  In contrast (but not shown) the entropy
-itself shows much larger statistical fluctuations with smaller bin
-sizes.  The canonical averaging process in Eq.~\ref{eq:cv} averages
-out those statistical fluctuations, which reduces the penalty of using
-a smaller bin size.
-
-%% Dependence on bin width.  Figure~\ref{fig:lj-cv-error} shows that the
-%% heat capacity converges just as well for bin widths varying over two
-%% orders of magnitude, although for bins as large as $\Delta
-%% E=0.1\epsilon$ the error in the heat capacity is significant, as is
-%% apparent in Fig~\ref{fig:lj-cv}.  The entropy itself converges much
-%% more slowly for a smaller $\Delta E$, but when summing over energies
-%% the statistical fluctuations in the entropy cancel. WL and $1/t$-WL
-%% show similar behavior, so this is not new, but is encouraging.
-
-The convergence behavior of the WL and $1/t$-WL is incredibly
-sensitive to the minimum energy chosen, which complicates the process of
-obtaining a workable result.  The known ground state energy of this
-cluster is $-133.586422\epsilon$~\cite{northby1987structure,
-  wales1997global}.  With a minimum energy of $-133.58\epsilon$, none
-of our WL or $1/t$-WL simulations converged in 10$^{12}$ moves.  With minimum
-energies of $-133.54\epsilon$, $-133.53\epsilon$, and
-$-133.52\epsilon$, the simulations do seem to converge, although the
-lowest energy case was significantly slower.
-This result highlights the advantage of being able to
-specify a temperature range rather than an energy range when computing
-properties as a function of temperature.  It is not only more
-convenient, but also far more efficient due the the fact that only a
-single simulation need be performed.
-
-%% Specifying temperature range rather than energy range is beneficial.
-%% Both WL and $1/t$-WL are incredibly sensitive to the minimum energy
-%% chosen, as seen in Fig.~\ref{fig:lj-cv-error}.  Cite previous work
-%% that showed them not converging because it chose too low a minimum
-%% energy, proving that it is \emph{not} a trivial procedure to determine
-%% an appropriate energy range.  Specifying the temperature range
-%% \emph{does} matter! We had to do a whole bunch of $1/t$-WL
-%% computations in order to determine what would be a sufficient
-%% low-energy cutoff.
-
-Finally, we want to mention our results for SAMC, which are not shown
-in our plots in order to reduce clutter.  With a well-chosen power of ten value for
-$t_0$, SAMC can converge to the correct result, but less rapidly than
-either $1/t$-WL or SAD.  However, with a poor $t_0$
-parameter, the method cannot converge in practice.  In addition, the
-optimal $t_0$ parameter depends on the bin width as well as the energy
-range chosen, further complicating matters.
-
-%% SAMC is a major pain.  The $t_0$ value needs to be adjusted when
-%% $\Delta E$ is changed or when the energy range is changed.  The energy
-%% range has the same downsides as the WL methods, although it manifests
-%% differently, with failure to converge as the problem, rather than
-%% failure for $\gamma_t$ to decrease.
-
-\section{Conclusions}
-We have introduced a new algorithm, a variant of the SAMC method, which
-effectively samples the
-energy space corresponding to a desired range of temperatures for a
-few systems.
-%
-We find that both SAD and $1/t$-WL converge more rapidly than SAMC, and
-unlike WL consistently converge to the correct density of states. SAD requires
-the user to specify a temperature range of interest rather than an
-energy range of interest as $1/t$-WL does.  For use cases in which
-a range of desired temperatures is known, this will make the SAD method
-considerably more convenient.
-
-We find that SAMC converges for a reasonable choice of
-$t_0$ but this parameter can be difficult to tune especially across
-significantly differing systems.  We find that even simple
-changes to the Monte Carlo moves can have a dramatic effect on the
-range of practical $t_0$ values.  Additionally, SAMC does not
-converge as rapidly as either SAD or $1/t$-WL even for the best choice of $t_0$,
-when a relatively small range of energies is required, because it always
-simulates all possible energies.
-
-Finally, we find that for examining a Lennard-Jones cluster the WL and
-$1/t$-WL algorithms---versions of which had previously been found to
-be unsuccessful on this problem~\cite{poulain2006performances}---have
-convergence properties that are highly dependent on the choice of
-energy range to be examined.  This is particularly problematic when
-examining the low-temperature heat capacity, where it is difficult to
-determine the lowest energy that will have a significant impact at the
-temperatures of interest. This is an example of how SAD is more
-powerful when the range of energies is not known in advance.
-
-%\section{Acknowledgement}
-
-%We gratefully acknowledge D.W. Siderius and
-%J.R. Errington for helpful communications regarding the implementation
-%of WL-TMMC.
-
-\bibliography{paper}% Produces the bibliography via BibTeX.
+\bibliography{sq-well}% Produces the bibliography via BibTeX.
 

### 2020-05-02 thesis/chapter-2.tex
needed to turn a unicode symbol into unicode.
added in final chapters including author contributions.
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

@@
 
-\section{Contribution of Authors}
-Jordan K. Pommerenck, Tanner T. Simpson, Michael A. Perlin, David J. Roundy
-
-% \affiliation{Department of Physics, Oregon State University,
-%   Corvallis, OR 97331}
-
-\section{Introduction}
-Over the past several decades, a number of flat histogram Monte Carlo simulation
-algorithms have been developed which calculate the thermodynamic properties of
-various systems over a range of temperatures.  This development began with the
-original histogram method, which used a single canonical Monte Carlo simulation
-to predict properties for nearby temperatures~\cite{ferrenberg1988new}.  For
-large systems, this approach is limited to a narrow temperature range because a
-single canonical simulation explores only a small range of energies. Berg and
-Neuhaus developed multicanonical methods that introduced a weight function to
-enable flat histogram sampling which improved the exploration of configuration
-space and allowed the system to overcome free energy
-barriers~\cite{berg1991multicanonical, berg1992multicanonical}. These works led
-to an explosion in the interest of ``flat'' (or ``broad'') histogram
-methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-wang2001determining, wang2001efficient, landau2004new, schulz2003avoiding,
-yan2003fast, trebst2004optimizing, belardinelli2007wang, belardinelli2007fast,
-belardinelli2008analysis, belardinelli2014intrinsic, singh2012density,
-zhou2008optimal, schneider2017convergence, liang2006theory, liang2007stochastic,
-liang2009improving, werlich2015stochastic, kim2006statistical,
-kim2007statistical, kim2009replica, junghans2014molecular}, which explore a
-wider range of energies. In addition to simulating a range of temperatures, in
-contrast with low-temperature canonical Monte Carlo, these approaches
-avoid being trapped in a local energy minimum.
-
-Wang and Landau introduced the most widely used flat histogram algorithm (WL)
-that uses an update factor and a statistical histogram to compute the density of
-states of a given system~\cite{wang2001determining, wang2001efficient}.  While
-the method is incredibly powerful, it has a few disadvantages. The most
-significant disadvantage is that the method requires the user to select the
-range of energies to be studied~\cite{landau2004new, wang2001efficient,
-schulz2003avoiding, yan2003fast}. This requirement adds an additional
-hurdle to its application to systems for which the interesting range of energies
-is not known \emph{a priori}.  The simulation violates detailed balance, albeit
-briefly as the size of the violation decreases with time, which complicates
-convergence analysis.  In fact, the error in a WL computation has been
-demonstrated to saturate at a non-zero value~\cite{yan2003fast}, i.e. the method
-does not converge to the true density of states~\cite{belardinelli2007wang,
-belardinelli2007fast, belardinelli2008analysis, belardinelli2014intrinsic,
-singh2012density, zhou2008optimal}.
-
-Belardinelli and Pereyra demonstrated that allowing the update factor to
-decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang},
-where $t$ corresponds to the number of moves.
-This leads to their $1/t$-WL algorithm which ensures that the error continues to
-decrease asymptotically as $1/\sqrt{t}$, which they demonstrated avoids
-error saturation and asymptotically approaches the true density of
-states~\cite{belardinelli2008analysis}. Zhou \emph{et al.} further confirmed that the
-original WL algorithm never converges exponentially and successfully bounded the
-statistical error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}.
-Schneider \emph{et al.} outline minor refinements to the $1/t$-WL algorithm
-including scaling the update factor with the number of energy
-bins~\cite{schneider2017convergence}.
-
-Liang began to consider Stochastic Approximation as a mathematical
-generalization of the WL approach, with convergence that could be mathematically
-proven~\cite{liang2006theory}. In 2007, Liang \emph{et al.}
-developed Stochastic Approximation Monte Carlo (SAMC)~\cite{liang2007stochastic,
-liang2009improving}, and proved its convergence, although the method still has a
-system specific user-defined parameter which must be tuned when applying this
-algorithm to a new system.  In contrast to the WL-based methods, the SAMC method
-does not require users to determine the energy range of interest \emph{a
-priori}. Werlich \emph{et al.} proposed the introduction of an additional tuning
-parameter into SAMC~\cite{werlich2015stochastic}.
-
-Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
-related Statistical Temperature Molecular Dynamics (STMD) which is an
-adaption of the WL method that approximates the entropy (or natural logarithm of the density of
-states) as a piecewise linear function, which improves convergence for systems
-with a continuously varying energy~\cite{kim2006statistical,
-kim2007statistical}. STMC applied to WL requires a temperature range be
-specified rather than an energy range.  Kim \emph{et al.} extended this work as
-Replica Exchange Statistical Temperature Monte Carlo (RESTMC), which uses
-replica exchange of multiple overlapping STMC simulations to improve
-convergence~\cite{kim2009replica}. More recently, Junghans \emph{et al.} have
-demonstrated a close connection between metadynamics, which was introduced by
-Laio and Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods,
-with STMD forging the connection~\cite{junghans2014molecular}.
-
-In this work, we have developed an improved algorithm based on SAMC that does
-not require an array of non-physical, user-defined inputs and therefore should
-be easily applicable to any system. The method (like STMC above) does require
-the user to define a temperature range of interest ($T_\text{min}$ to
-$T_\infty$) which we explain in Section~\ref{sec:sad}.  We call this method SAD (Stochastic Approximation with a Dynamic
-update factor), and will discuss it in detail in the methods section. We compare
-its convergence properties with three existing flat histogram methods: WL,
-$1/t$-WL, and SAMC.
-
-In this work, we compare four flat histogram methods.  We outline the
-general workings of each algorithm that we developed in detail while
-summarizing algorithms that were developed in other works.  The
-following methods are discussed and simulated for the square-well
-fluid and the 31-atom Lennard-Jones cluster: Wang-Landau (WL),
-$1/t$-Wang-Landau ($1/t$-WL), Stochastic Approximation Monte Carlo
-(SAMC), and SAD.
-
-\section{Flat histogram methods}\label{sec:histogram}
-The goal of flat histogram methods (also called \emph{broad histogram}
-or \emph{multicanonical} methods) is to simulate each energy with
-similar accuracy so as to accurately determine the density of states
-over a broad range of energies---and thus to determine the thermodynamic
-quantities such as heat capacity or internal energy
-over a broad range of temperatures.
-Properties that require more information---such as a spatial
-correlation function or a response function---can still be computed
-for any temperature, provided statistics are collected for each
-individual energy, which can then be reweighted for any
-temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
-errington2003direct}.
-
-All the flat histogram Monte Carlo methods begin with randomly chosen
-``moves'' which change the state of the system and must satisfy
-detailed balance.  Each algorithm differs in how it determines the
-probability of accepting a move and in what additional statistics must
-be collected in order to decide on that probability.
-
-Flat histogram methods calculate the density of states $D(E)$ for a discrete set
-of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
-trebst2004optimizing}. Therefore, energy binning becomes an important
-consideration for systems with a continuum of possible energies.  Energy bins
-are typically of uniform size for the entire energy
-continuum~\cite{fasnacht2004adaptive}. Some methods such as
-AdaWL~\cite{koh2013dynamically} employ a tunable mechanism for controlling the
-binning for low entropic states in order to ensure the exploration of all
-energies.  The method introduced in this paper is designed to scale appropriately
-as bin size is changed.
-
-In this section we will introduce four closely related flat histogram
-methods each of which rely on a weight function $w(E)$.  In these
-algorithms, the probability of accepting a move is given by
-\begin{equation}
-	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
-	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
-\end{equation}
-which biases the simulation in favor of energies with low weights.  A set of
-weights that are proportional to the density of states $D(E)$ of the system will
-result in an entirely flat histogram, which means that the weights should
-converge to being proportional to the density of states.  The natural logarithm
-of the weight is typically stored, since the density of states will often vary
-over a few hundred orders of magnitude. In the microcanonical ensemble, the
-entropy is defined as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the
-logarithm of the weight is an approximation of the entropy.
-
-Each approach uses a random walk in energy space to estimate the density of
-states.  The core of these approaches is to update the weights at each step of
-the simulation
-\begin{equation}
-	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
-	+\gamma_t
-\end{equation}
-where $t$ is the number of the current move, $\gamma_t$ is an update factor
-that varies over the simulation, and $E$ is the current energy.  This update
-causes the random walk to avoid energies that have been frequently sampled,
-enabling a rapid exploration of energy space. This approach violates detailed
-balance, due to the acceptance probabilities changing with each move, but the
-severity of this violation decreases as we decrease $\gamma_t$.  The four
-methods differ primarily in how they schedule the decrease of $\gamma_t$.
-
-\subsection{Wang-Landau}
-
-The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
-landau2014guide} begins with $\gamma^{\text{WL}}_{t=0}=1$, and then decreases
-$\gamma^{\text{WL}}$ in discrete stages.  When the energy histogram is
-sufficiently flat, $\gamma^{\text{WL}}$ is decreased by a specified factor of
-$\frac12$.  The flatness is defined by the ratio between the minimum value of
-the histogram and its average value.  When this flatness reaches a specified
-threshold (typically 0.8), the $\gamma^{\text{WL}}$ value is decreased and the
-histogram is reset to zero.  This approach requires that the energy range of
-interest be known in advance, and difficulties can occur with this flatness
-criteria due to the fact that some energies in this energy range might never be
-sampled~\cite{troster2005wang}.  The entire process is repeated until
-$\gamma^{\text{WL}}$ reaches a desired cutoff.
-
-The Wang-Landau approach thus has three parameters that need be
-specified: the factor by which to decrease $\gamma^{\text{WL}}$ when flatness is
-achieved, the flatness criterion, and the cutoff that determines when
-the computation is complete.  In addition, an energy range (or in
-general, a \emph{set} of energies) must be supplied, so that the
-flatness criterion can be defined.
-
-\subsection{$1/t$-Wang-Landau}
-
-The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
-from dropping below $N_S/t$~\cite{belardinelli2008analysis,
-schneider2017convergence}. The method follows the standard WL algorithm with two
-modifications.  Firstly, the histogram is considered flat, and $\gamma_t$ is
-decreased by a factor of two, when every energy state has been visited once,
-i.e. when the WL ``flatness'' becomes nonzero. Secondly, when
-$\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the algorithm switches to use
-$\gamma_t = N_S/t$ for the remainder of the simulation:
-\begin{align}
-  \gamma_t^{1/t\text{-WL}} = \begin{cases}
-     \gamma^{\text{WL}}_t & \gamma^{\text{WL}}_t > \frac{N_S}{t} \\
-     \frac{N_S}{t} & t \ge t_0
- \end{cases}
-\end{align}
-where $t$ is the number of moves, $\gamma^{\text{WL}}_t$ is the Wang-Landau update factor
-at move $t$, and $N_S$ is the number of energy bins.
-
-\subsection{SAMC}
-The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses the lack of
-convergence of Wang-Landau's approach with a simple schedule by which the update
-factor $\gamma_t$ is continuously decreased~\cite{liang2007stochastic,
-werlich2015stochastic, schneider2017convergence}.  The update factor is defined
-in the original implementation~\cite{liang2007stochastic} in terms of an tunable
-parameter $t_0$,
-\begin{align}
-\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}\label{eq:1}
-\end{align}
-where as above $t$ is the number of moves that have been attempted.
-SAMC offers extreme simplicity, combined with is proven convergence.
-Provided the update factor satisfies
-\begin{align}
-\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
-\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
-\end{align}
-where $\zeta > 1$, Liang has proven that the weights converge to the true
-density of states~\cite{liang2006theory, liang2007stochastic,
-liang2009improving}.  In addition, the energy range need not be known \emph{a priori}.
-The time to converge depends only on the choice of parameter $t_0$.
-Unfortunately, there is no prescription for finding an acceptable value for
-$t_0$, and while the algorithm formally converges, for a poor choice of $t_0$
-that convergence can be far too slow to be practical. Liang \emph{et al.} give a
-rule of thumb in which $t_0$ is chosen in the range from $2N_S$ to $100N_S$
-where $N_S$ is the number of energy bins~\cite{liang2007stochastic}. Schneider
-\emph{et al.} found that for the Ising model this heuristic is helpful for small
-spin systems, but that larger systems require an even higher $t_0$
-value~\cite{schneider2017convergence}.  We will describe below one case we
-examined, in which $t_0$ needs to be as much as two orders of magnitude higher
-than the rule of thumb of $100N_S$ in order to converge in $10^{12}$ moves.
-
-Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t^{\text{SA}}$
-by a factor $\gamma_0$~\cite{werlich2015stochastic}.  While this may result in
-an improved rate of convergence, it adds yet another parameter that must be
-empirically determined, and we have not explored this additional degree of
-freedom.
-
-\subsection{SAMC convergence time}\label{sec:samc-convergence}
-A primary difficulty in using the SAMC method lies in identifying an
-appropriate value for $t_0$.  Although SAMC is proven to formally
-converge regardless of the $t_0$ value, a choice that is either too
-high or too low will result in prohibitively slow convergence to the true
-entropy of the system.  It is instructive to consider separately
-values of $t_0$ that are too low or too high.
-
-We can place a rigorous \emph{lower} bound $t_{\min}$ on the number of
-moves required to find the true entropy by considering the total change
-that needs to be made to the entropy:
-\begin{align}
-  \Delta S_{\text{tot}} \equiv \sum_E S(E) - S_{\min}.
-\end{align}
-This connects with $\gamma_t$ because the $\ln w(E)$ (our approximation for entropy)
-has $\gamma_t$ added to it on each move.  Thus a minimum number of moves
-that could possibly result in the true entropy $S(E)$ starting with a flat set of
-weights is determined by the total entropy change required.  We can estimate
-this number of moves required by summing the $\gamma_t$,
-which we can approximate using an integral:
-\begin{align}
-   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t^{\text{SA}} \\
-  % &= \int_0^{t_{\min}} \gamma(t)dt \\
-  &\approx t_0 + \int_{t_0}^{t_{\min}} \frac{t_0}{t}dt
-  \\
-  &= t_0\left(1 + \ln\left(\frac{t_{\min}}{t_0}\right)\right)
-\end{align}
-Solving for ${t_{\min}}$ we find that
-\begin{align}
-  {t_{\min}} &= t_0 e^{\frac{\Delta S_{\text{tot}}}{t_0} - 1}
-\end{align}
-which means that the minimum time to converge grows exponentially
-as $t_0$ is made smaller.  You \emph{seriously} don't want to underestimate
-$t_0$!
-
-One might reasonably choose to err by selecting a large $t_0$. The rate of
-convergence is harder to estimate when $t_0$ is large, but in general
-$\gamma_t^{\text{SA}}$ itself forms a lower bound on the accuracy with which the
-entropy may be known, with an unknown prefactor which is related to the
-coherence time ($t = t_0$) of the Monte Carlo simulation.  Since
-$\gamma_t^{\text{SA}}$ is given by $t_0/t$, the time to converge to a given
-accuracy is increased in proportion to the ratio by which we overestimate $t_0$.
-Thus, while it is exponentially painful to underestimate $t_0$, overestimating
-by several orders of magnitude is also not acceptable.  We should note that
-these extreme limiting cases do not preclude the possibility that there is a
-wide range of $t_0$ values that lead to an acceptable convergence rate.
-
-\section{SAD Algorithm}\label{sec:sad}
-The Stochastic Approximation with Dynamic update factor (SAD) method
-is a variant of the SAMC Algorithm that attempts to dynamically choose
-the modification factor rather than relying on system dependent
-parameters such as $t_0$ or $\gamma_0$.  There is an immediate
-advantage of such an algorithm where parameters are chosen independent
-of system size or type. Each flat-histogram method has unique
-advantages and disadvantages.  Wang-Landau and $1/t$-WL require an
-energy range for initialization.  SAMC removes this energy range
-requirement but requires simulating every possible energy. Our
-proposed method SAD requires the user to input $T_\text{min}$, the
-lowest temperature of interest, which is an immediate disadvantage of
-the method. However, identifying a minimum temperature of interest
-$T_\text{min}$ may be easier for a user than determining in advance
-an energy range of interest or the unphysical parameter $t_0$.
-
-We set the maximum temperature of interest as $T=\infty$ for a couple
-of reasons.  First, when this temperature is reached, the system is
-maximally randomized, so allowing infinite temperature further ensures
-that the system can surmount an energy barrier of any size.  Second,
-the high temperatures tend to be ``easy'' to simulate, and tend to
-converge very quickly.  We note, however, that this is under the
-assumption that infinite temperature corresponds to a finite energy,
-as is the case for configurational sampling.  Were we to include
-kinetic energy in a simulation (as in molecular dynamics), or to
-examine a system such as a high density Lennard-Jones fluid with an
-entropy maximum at an energy that is far from the energy range of
-physical interest, then we would need to apply either a maximum
-temperature or a maximum energy, either of which could slow down the
-rate at which state space is explored.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../square-well/N50-lndos-comparison.pdf}
-  \caption[The entropy of a square-well fluid with 50 atoms and filling fraction
-        0.3 as the green line.]{The entropy of a square-well fluid with 50 atoms and filling fraction
-        0.3 as the green line.  The green hatched area reflects the
-        minimum entropy change needed to converge to the true value.
-        The light blue area is the quadratic approximation
-        for the change in entropy.  The vertical dotted lines represent
-        the energy corresponding to $T=1/3$ and $T=\infty$.}
-  \label{fig:entropy-cartoon}
-\end{figure}
-
-While for SAMC, the update factor is defined in the original
-implementation (see equation~\eqref{eq:1}), for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
-thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
-the SAMC parameter
-$t_0$ should have dimensions of entropy.
-We begin with an estimate of the average value of the entropy (relative
-to the lowest entropy at $T_{\min}$).  If we assume a quadratic
-dependence on energy (see Fig.~\ref{fig:entropy-cartoon}), this is given by
-\begin{align}
-\langle S\rangle \approx \frac13 \frac{E({T=\infty}) - E(T_{\min})}{T_{\min}}
-\end{align}
-We approximate this energy difference by $E_H -E_L$ where $E_H$ and
-$E_L$ are defined below. The entropy numerator of the update factor in
-general should scale with the total number of interesting energy states
-$N_S$, since updates to the weights are distributed between that many
-energy states.  The product $N_S\langle S\rangle$ is the total change
-of entropy required (starting from constant weights) to find the true
-entropy, and puts a lower bound on the convergence time. After
-\emph{long} times, when all the energies have been explored a long time ago, we
-wish for a lower update factor in order to more rapidly refine the
-remaining error in entropy.  We track the time at which we first visited
-each possible energy.  We define $t_L$ to be the last time that we
-encountered an energy that we currently believe is in the energy range
-of interest, so a $t\gg t_L$ we feel confident that we have established
-the true energy range of interest. We gradually transition to a lower
-update factor (but still asymptotically scaling as $\gamma_t \propto
-1/t$ to ensure eventual convergence).  Finally, we wish for an update
-factor that is \emph{never} greater than 1, because a very large update
-factor could introduce very large errors in entropy that may take many iterations
-to remove.  The SAD expression for $\gamma_t$ which incorporates these
-ideas is:
-\begin{align}
-  \gamma_{t}^{\text{SAD}} =
-     \frac{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
-     }{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
-     }
-\end{align}
-where $E_H$ and $E_L$ are the current estimates for the highest and
-lowest energies of interest as defined below.  This factor
-asymptotically has the same $1/t$ behavior as the original SAMC
-algorithm and with the same $N_S$ prefactor used by the $1/t$-WL
-method; however for earlier values of $t$, the update factor drops as
-$1/t^2$ and jumps every time a new energy is determined to be of
-interest.  This behavior allows SAD to dynamically prevent the update
-factor from decreasing too rapidly.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../square-well/gamma-n256.pdf}
-  \caption{The update factor $\gamma_t$ versus iteration for the
-    square-well fluid with 256 atoms, for four different methods: WL, $1/t$-WL, SAMC,
-    and SAD.}
-    \label{fig:gamma-vs-t}
-\end{figure}
-Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for the related
-methods SAD, WL, $1/t$-WL, and SAMC.  For SAMC, $\gamma_t$ remains
-constant before dropping as $1/t$.  WL $\gamma_t$ remains at 1 for
-many iterations, and then decreases very rapidly, with $1/t$-WL
-behaving similarly but decreasing more aggressively before
-transitioning to a more conservative $1/t$ behavior.  The update
-factor for SAD fluctuates dynamically around a value less than 1 for
-early MC moves, and then decreases as approximately $1/t$ while
-continuing to fluctuate as new energies are found to be important.  At
-intermediate times, the SAD $\gamma_t$ decreases as $1/t^2$ before
-asymptoting to $N_S/t$, which is the same as $1/t$-WL.
-
-Since SAD does not explore \emph{all} energy states, it needs to
-determine what energy range corresponds to the temperature range of
-interest defined by $T_{\min}<T<\infty$. The simulation is responsible
-for determining and updating this energy range. Given the true entropy
-$S(E)$, we can define the interesting energy range as
-%\begin{align}
-  $E(T_{\min}) <E< E(T=\infty)$
-%\end{align}
-where $E(T)$ is the energy that maximizes $S-E/T$.  During the course of the
-simulation, this precise energy is challenging to evaluate accurately.
-In order to ensure that we sample this entire energy range adequately,
-we define two energy limits:  a high energy $E_H$ and a low
-energy $E_L$, which define the range over which the energy histogram
-is made flat. At move $t$, $E_H$ and $E_L$ are the greatest and lowest
-energy that prior to that move
-had the highest histogram value (i.e. been visited the most times) at some point
-during the course of the simulation.
-This definition results in a ``ratcheting'' effect, in which $E_H$
-may only increase, while $E_L$ may only decrease over the course of the
-simulation, which results in a conservative estimate of the range of
-energies that need be sampled.
-
-During the simulation when considering a move inside the energy
-range of interest $E_L \le E \le E_H$, the weights are used as in the
-three methods already described.  If $E\ge E_H$, the weight is taken
-to be
-\begin{align}\label{eq:highw}
-  w(E>E_H) &= w(E_H),
-\end{align}
-which corresponds to an infinite temperature.  This choice ensures
-that if the maximum in entropy is at an energy $E_{\max}>E_H$, then
-the energy $E_{\max}$ will eventually have the highest number of
-counts and the ratcheting will result in $E_H\ge E_{\max}$ .  At lower
-energies, Boltzmann weights corresponding to the minimum temperature
-are used:
-\begin{align}\label{eq:loww}
-  w(E<E_L) &= w(E_L)e^{-\frac{E_L-E}{T_{\min}}}.
-\end{align}
-This choice has the result that if the energy $E_{\min}$ at which the free
-energy at $T_{\min}$ is minimized is less than $E_L$, the lower energy
-limit will ratchet down to include $E_{\min}$.
-Each time we change the value of $E_H$ or $E_L$, the weights outside the
-new portion of the interesting energy range are set to the expressions
-in Equations~\ref{eq:highw} and~\ref{eq:loww}.
-
-%~ \paragraph{Energy binning with SAD}
-
-A significant advantage of SAD over SAMC---which the $1/t$-WL and WL
-methods share \emph{after} they have discovered all the energies---is that
-the schedule for $\gamma_t$ automatically responds to the choice of bin
-size.
-SAD should perform similarly over a reasonable range of bin sizes
-because $\gamma_t \propto \nicefrac{N_S}{t}$.  As the number of energy
-states $N_S$ found increases (fine binning), the time spent $t$ in
-each bin will decrease with the effect that the convergence should be
-roughly independent of the bin size chosen.  SAMC could be used with a
-prefactor $\gamma_0$ to aid in a similar
-way~\cite{werlich2015stochastic} but this adds yet another parameter
-for the user to choose.
-
-\section{Results}\label{sec:results}
-
-\subsection{Square-well fluid}
-As our first test case, we consider the square-well fluid i.e. a
-system of particles whose interactions are governed by a square-well
-potential~\cite{singh2003surface, barker2004perturbationSW}.  The
-square-well potential is an ideal test-bed as it is a simple model for
-a liquid, which includes both attractive and repulsive
-interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
-To date, there have not been any published direct convergence comparison tests
-for flat histogram methods applied to the square-well fluid.
-The potential $U(\textbf{r})$ for such a system is given by
-\begin{equation}
- U(\textbf{r})=\begin{cases} \infty &
- \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
- \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
- \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
-\end{equation}
-where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
-reduced range of the potential well, and $\epsilon$ is its depth. This model has
-the further advantage that binning is not required because the energy is
-discrete.
-
-\begin{figure*}
-  \includegraphics[width=0.5\textwidth]{../square-well/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
-\includegraphics[width=0.5\textwidth]{../square-well/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
-  \caption[(a) The average entropy error for each MC method for $N=50$,
-               $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run.]{(a) The average entropy error for each MC method for $N=50$,
-               $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run. The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded area, and
-           (b) the update factor $\gamma_t$ versus iteration number
-               for the same simulations.
-           (c) The average entropy error for each MC method for the
-               same physical system with a smaller displacement distance
-               $\delta_0 = 0.005\sigma$,
-               as a function of number of iterations run, and
-           (d) the update factor $\gamma_t$ versus iteration number
-               for the same simulations.
-  }\label{fig:fast-slow-gamma}
-\end{figure*}
-
-We tested the algorithms on two square-well fluid systems.  The first
-is a smaller simulation with a particle number of 50, a well-width of
-$\lambda = 1.3$, and a volume corresponding to a filling fraction
-(defined as the fraction of volume filled by atoms) of $\eta =
-0.3$. The second system is larger, with a particle number of 256, a
-well-width of $\lambda = 1.5$, and a volume corresponding to a filling
-fraction of $\eta = 0.17$.  For each system we use a reasonable
-root-mean-square displacement distance $\delta_0 = 0.05\sigma$ for
-proposed moves, and for the smaller system we also use an unreasonably
-small displacement distance of $0.005\sigma$. The simulations explore
-the energy space of the systems with minimum reduced temperatures of
-$T_{\text{min}} = 1/3$ for simulations of the smaller system, and
-$T_{\min}=1$ for the larger system.  All simulations lead to the
-minimum important energy $E_{\min}$ and maximum entropy energy
-$E_{\max}$ being calculated (with the exception of the WL methods
-where both of these parameters are needed \emph{a priori}).
-
-The SAMC simulations
-computed the density of states for the entire range of possible
-energies.  The SAD simulations determined the energy range of interest
-dynamically as described above, based on a specified $T_{\min}$.  For
-the WL and $1/t$-WL simulations, we constrained the simulation to
-remain in the energy range corresponding to $T_{\min} < T < \infty$,
-as determined by a previous SAMC simulation.  Thus the WL and $1/t$-WL
-simulations were given extra information that in practice would not be
-available without additional computational effort, and the SAMC simulations
-computed the entropy over the entire range of possible energies, which
-required more effort.
-
-We use the average entropy error versus moves as a metric to compare
-simulation runtimes and overall convergence. The overall accuracy
-is determined by examining the fractional error of a particular method to
-a precise reference system. For each simulation, the reference system
-is chosen to be the final output of a SAMC simulation with a fixed energy range
-corresponding to the temperature range of interest.
-Although SAMC does not require an energy range as an input parameter,
-we find that by limiting the simulation to this energy range, we can
-achieve much faster convergence with a smaller $t_0$.  We compute an average
-of the error by averaging the error in the entropy over the
-interesting energy range, and then averaging \emph{this} error over several
-simulations run with different random number seeds.
-
-\subsubsection{50 square-well atoms}
-For this 50 atom simulation, we chose a minimum reduced
-temperature of $1/3$, which corresponds to an interesting energy range
-from $-248$ to $-120$.  The number of important energy states
-for this system is therefore $N_S = 129$.  The entropy of this system is shown in
-Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
-energy range the entropy differs by 198, corresponding to a ratio of
-$10^{86}$ between the highest and lowest density of states.
-
-In order to explore the effect of simulation details on convergence,
-we consider two values for the displacement distance by which atoms
-are moved during a Monte Carlo step.  We began with a reasonable
-displacement distance of $\delta_0 = 0.05\sigma$, which corresponds to
-an acceptance rate of proposed moves of 38\%.  We further ran
-simulations with a much smaller displacement distance of $\delta_0 =
-0.005\sigma$, which resulted in an acceptance rate of 86\%, which
-converged more slowly.
-
-Figure~\ref{fig:fast-slow-gamma}a shows the average error in the
-entropy as a function of time for this system with the reasonable
-displacement distance of $\delta_0 = 0.05\sigma$.  The solid/dashed lines
-represent the average of the absolute value of the error in the
-entropy averaged over eight simulations using different random number
-seeds.  The range of average errors for each simulation is shown as a
-shaded region around its mean error.  By the time $10^8$ moves have
-been made all but the SAMC simulation with the shortest $t_0$ have
-begun to converge as $1/\sqrt{t}$.  We then see the WL error
-saturate around $10^{10}$ moves.
-
-Figure~\ref{fig:fast-slow-gamma}c shows the average error in
-the entropy as a function of time for this system with the
-unreasonably small displacement distance of $\delta_0 = 0.005\sigma$.
-%
-The smaller translation scale causes all methods to take additional time to
-explore all energies. Based on random walk scaling, the convergence time of an
-ideal method should scale roughly as $\delta_0^{-2}$ in the limit of small
-$\delta_0$, that is, one order of magnitude decrease in the displacement
-distance should result in two order of magnitude increase in convergence time.
-SAMC simulations with a $t_0$ value that rapidly converged for $\delta =
-0.05\sigma$ do not converge at all in $10^{12}$ moves for a translation scale of
-$\delta = 0.005\sigma$. It is also worth noting that for the smaller
-displacement distance, the SAMC rule of thumb of choosing $t_0$ to be
-approximately $100N_S$ is no longer valid. SAD, WL, and $1/t$-WL handle the
-shift in displacement distance and converge roughly as expected.
-
-The methods SAD, WL, and $1/t$-WL compensate for the smaller
-displacement distance by reducing $\gamma_t$ more slowly, as can be seen from
-Figure~\ref{fig:fast-slow-gamma}b and ~\ref{fig:fast-slow-gamma}d.
-The update factors take approximately 10$\times$ longer to reach steady-state
-for the smaller displacement distance. Because of this update behavior,
-these methods are less sensitive to the choice of displacement
-distance than SAMC is.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{../square-well/periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
-  \caption[The average entropy error for each MC method for $N=256$,
-               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
-               as a function of number of iterations run.]{
-  The average entropy error for each MC method for $N=256$,
-               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
-               as a function of number of iterations run.  The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded area.  The update factor for this
-               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
-\end{figure}
-
-\subsubsection{256 square-well atoms}
-Next we introduce a considerably larger simulation containing 256
-atoms which has a maximum entropy about 1500 greater than its minimum.
-This makes exploring the entire range of energies extremely expensive,
-and strongly favors the methods that restrict the energy (or
-temperature) range of interest.  For this simulation, we chose a much
-higher minimum reduced temperature of $1.0$, which corresponds to an
-interesting energy range from $-915$ to $-509$.  The number of
-important energy states for this system is then $N_S = 407$.  The
-minimum entropy over this energy range is just 395 less than the
-maximum, corresponding to a ratio of only $10^{118}$ between the
-highest and lowest density of states.
-
-Figure~\ref{fig:n256} shows the average error in the entropy as a
-function of moves for this system with the reasonable displacement
-distance of $\delta_0 = 0.05\sigma$.  The solid lines represent the
-average of the absolute value of the error in the entropy averaged
-over eight simulations using different random number seeds.  The range
-of average errors for each simulation is shown as a shaded region
-around its mean error.  By the time $10^8$ moves have been made all
-but the SAMC simulation with the shortest $t_0$ shown have begun to
-converge as $1/\sqrt{t}$.  We then see the WL error saturates around
-$10^{10}$ moves.  Once again, the convergence of SAD is essentially
-the same as that of the $1/t$-WL method.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../square-well/lj-cv.pdf}
-  \caption[The heat capacity of the LJ31 cluster after $10^{12}$ moves
-    as calculated with each method.]{The heat capacity of the LJ31 cluster after $10^{12}$ moves
-    as calculated with each method.  Each simulation used a bin size
-    of $\Delta E=0.01\epsilon$.}
-  \label{fig:lj-cv}
-\end{figure}
-
-\subsection{31-atom Lennard-Jones cluster}
-
-We also examined convergence of each algorithm on the 31-atom cluster
-(LJ31), which is the smallest Lennard-Jones cluster that exhibits
-broken ergodicity and a solid-solid phase transition at a low
-temperature~\cite{doye1998thermodynamics,
-  martiniani2014superposition}.  Clusters of Lennard-Jones atoms have
-been frequently used for testing Monte Carlo
-algorithms~\cite{wales1997global, neirotti2000phase,
-  frantsuzov2005size, mandelshtam2006multiple}.  The LJ31 cluster
-is commonly used for testing convergence of algorithms
-because it features a two-funnel energy landscape with a significant
-barrier between the two low-energy states~\cite{calvo2000entropic,
-  calvo2000phase, poulain2006performances,
-  martiniani2014superposition}.  In particular, the LJ31 cluster
-features a low-temperature solid-solid phase transition that is challenging to
-converge~\cite{mandelshtam2006multiple}.  Poulain \emph{et al.}
-tested a number of variants of the WL algorithm on LJ31 and failed to
-converge the spike in heat capacity corresponding to the solid-solid
-phase transition~\cite{poulain2006performances}.
-
-Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster
-over the temperature range from $k_BT=0.01\epsilon$ to
-$k_BT=0.40\epsilon$.  There are two peaks, a large peak around $k_BT
-\approx 0.3\epsilon$ corresponding to melting temperature, and a
-small peak around $k_BT \approx 0.027\epsilon$ for the solid-solid
-transition from a Mackay-to-anti-Mackay (M $\rightarrow$ aM)
-transition~\cite{doye1997thermally, hendy2001molecular,
-  mandelshtam2006multiple}.  We constrain our atoms within a spherical
-box of radius 2.5$\sigma$, as is common in the
-literature~\cite{poulain2006performances, mandelshtam2006multiple}.
-We note that using a larger box (with radius $5\sigma$) has a large
-effect on the melting peak, and a smaller but still significant effect
-on the solid-solid transition.  Since our focus is on the convergence
-rate, we will restrict ourselves to the smaller box size.  In addition, we
-further restricted all our simulations to negative total potential energies.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{../square-well/lj-cv-error.pdf}
-  \caption[The maximum error in the heat capacity of the LJ31 cluster
-    in the temperature range 0.01~$\epsilon\le k_BT \le 0.05\epsilon$]{The maximum error in the heat capacity of the LJ31 cluster
-    in the temperature range 0.01~$\epsilon\le k_BT \le 0.05\epsilon$.
-    Note that the value of $C_V$ ranges from 90$k_B$ to 120$k_B$, so
-    the smallest (maximum) errors seen are around 1\%.  All WL methods
-    shown were run with a histogram bin width of $\Delta
-    E=0.01\epsilon$.  The converged reference heat capacity was
-    computed with a $1/t$-WL simulation with the energy constrained to
-    a range $-133.53\epsilon\le E\le -110\epsilon$.
-  }
-  \label{fig:lj-cv-error}
-\end{figure}
-
-In order to test the low-temperature convergence, we focused on the
-heat capacity in the temperature range $0.01\epsilon \le k_BT \le
-0.05\epsilon$ (see inset in Fig.~\ref{fig:lj-cv}), which contains
-the solid-solid phase transition.
-%
-We compute the heat capacity from the entropy using a standard
-canonical analysis, which gives
-\begin{align}
-  C_V &= 31\times\frac32k_B + \frac1{k_BT^2}\left< (E- \langle
-  E\rangle)^2\right>
-  \label{eq:cv}
-\end{align}
-where the first term represents the kinetic energy contribution, and
-the second term is the potential energy contribution, with canonical
-averages using the microcanonical entropy to compute multiplicities.
-%
-For our reference heat capacity, we ran a $1/t$-WL simulation with the
-energy constrained to a range $-133.53\epsilon\le E\le -110\epsilon$.
-We chose this energy range through experimentation, as we found the
-convergence rate to be very sensitive to the minimum energy chosen.
-
-%% \jpcom{this is a basically a repeat of above...
-%% The heat capacity changes significantly even at these low temperatures
-%% when the size of the cavity is increased.  We use a 2.5 cavity because
-%% ``everyone else'' does~\cite{poulain2006performances}.  For purposes
-%% of testing the convergence of the algorithm, we don't need to solve
-%% for the true heat capacity.  Furthermore, the heat capacity at fixed
-%% volume is not the most meaningful quantity to predict for a cluster.
-%% }
-We ran simulations with several bin widths using each algorithm we studied
-and show results for $\Delta E=0.1\epsilon$ and
-$0.01\epsilon$.
-When examining the converged results, we saw that each bin size yielded a
-measurably different low temperature heat capacity. The largest
-bin size of $0.1\epsilon$ introduced a maximum error in heat capacity
-of more than $20k_B$ compared with the smaller bin size.  We will focus on
-the smaller bin size, and
-used a bin size of $0.01\epsilon$ for our converged reference heat capacity.
-
-Figure~\ref{fig:lj-cv-error} shows the maximum error in heat capacity
-over the temperature range $0.01\epsilon\le k_BT\le 0.05\epsilon$ as a
-function of the number of Monte Carlo moves performed.
-%
-We explored the convergence behavior of each algorithm at two energy
-bin widths ($\Delta E=0.01\epsilon$ and $\Delta E=0.1\epsilon$).  We
-were unable to converge the WL or $1/t$-WL algorithms with the larger
-bin size.
-%
-The rate of convergence of SAD for the heat capacity seems to be roughly
-independent of the bin width.  In contrast (but not shown) the entropy
-itself shows much larger statistical fluctuations with smaller bin
-sizes.  The canonical averaging process in Eq.~\ref{eq:cv} averages
-out those statistical fluctuations, which reduces the penalty of using
-a smaller bin size.
-
-%% Dependence on bin width.  Figure~\ref{fig:lj-cv-error} shows that the
-%% heat capacity converges just as well for bin widths varying over two
-%% orders of magnitude, although for bins as large as $\Delta
-%% E=0.1\epsilon$ the error in the heat capacity is significant, as is
-%% apparent in Fig~\ref{fig:lj-cv}.  The entropy itself converges much
-%% more slowly for a smaller $\Delta E$, but when summing over energies
-%% the statistical fluctuations in the entropy cancel. WL and $1/t$-WL
-%% show similar behavior, so this is not new, but is encouraging.
-
-The convergence behavior of the WL and $1/t$-WL is incredibly
-sensitive to the minimum energy chosen, which complicates the process of
-obtaining a workable result.  The known ground state energy of this
-cluster is $-133.586422\epsilon$~\cite{northby1987structure,
-  wales1997global}.  With a minimum energy of $-133.58\epsilon$, none
-of our WL or $1/t$-WL simulations converged in 10$^{12}$ moves.  With minimum
-energies of $-133.54\epsilon$, $-133.53\epsilon$, and
-$-133.52\epsilon$, the simulations do seem to converge, although the
-lowest energy case was significantly slower.
-This result highlights the advantage of being able to
-specify a temperature range rather than an energy range when computing
-properties as a function of temperature.  It is not only more
-convenient, but also far more efficient due the the fact that only a
-single simulation need be performed.
-
-%% Specifying temperature range rather than energy range is beneficial.
-%% Both WL and $1/t$-WL are incredibly sensitive to the minimum energy
-%% chosen, as seen in Fig.~\ref{fig:lj-cv-error}.  Cite previous work
-%% that showed them not converging because it chose too low a minimum
-%% energy, proving that it is \emph{not} a trivial procedure to determine
-%% an appropriate energy range.  Specifying the temperature range
-%% \emph{does} matter! We had to do a whole bunch of $1/t$-WL
-%% computations in order to determine what would be a sufficient
-%% low-energy cutoff.
-
-Finally, we want to mention our results for SAMC, which are not shown
-in our plots in order to reduce clutter.  With a well-chosen power of ten value for
-$t_0$, SAMC can converge to the correct result, but less rapidly than
-either $1/t$-WL or SAD.  However, with a poor $t_0$
-parameter, the method cannot converge in practice.  In addition, the
-optimal $t_0$ parameter depends on the bin width as well as the energy
-range chosen, further complicating matters.
-
-%% SAMC is a major pain.  The $t_0$ value needs to be adjusted when
-%% $\Delta E$ is changed or when the energy range is changed.  The energy
-%% range has the same downsides as the WL methods, although it manifests
-%% differently, with failure to converge as the problem, rather than
-%% failure for $\gamma_t$ to decrease.
-
-\section{Conclusions}
-We have introduced a new algorithm, a variant of the SAMC method, which
-effectively samples the
-energy space corresponding to a desired range of temperatures for a
-few systems.
-%
-We find that both SAD and $1/t$-WL converge more rapidly than SAMC, and
-unlike WL consistently converge to the correct density of states. SAD requires
-the user to specify a temperature range of interest rather than an
-energy range of interest as $1/t$-WL does.  For use cases in which
-a range of desired temperatures is known, this will make the SAD method
-considerably more convenient.
-
-We find that SAMC converges for a reasonable choice of
-$t_0$ but this parameter can be difficult to tune especially across
-significantly differing systems.  We find that even simple
-changes to the Monte Carlo moves can have a dramatic effect on the
-range of practical $t_0$ values.  Additionally, SAMC does not
-converge as rapidly as either SAD or $1/t$-WL even for the best choice of $t_0$,
-when a relatively small range of energies is required, because it always
-simulates all possible energies.
-
-Finally, we find that for examining a Lennard-Jones cluster the WL and
-$1/t$-WL algorithms---versions of which had previously been found to
-be unsuccessful on this problem~\cite{poulain2006performances}---have
-convergence properties that are highly dependent on the choice of
-energy range to be examined.  This is particularly problematic when
-examining the low-temperature heat capacity, where it is difficult to
-determine the lowest energy that will have a significant impact at the
-temperatures of interest. This is an example of how SAD is more
-powerful when the range of energies is not known in advance.
+\input{../square-well/sq-well-body.tex}

### 2020-05-02 thesis/chapter-3.tex
added in final chapters including author contributions.
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

@@
-\chapter{Flat histogram method comparison on 2D Ising Model
+\chapter{Flat-histogram method comparison on 2D Ising Model
 }
 
-\section{Contribution of Authors}
-Jordan K. Pommerenck, Tanner T. Simpson, Michael A. Perlin, David J. Roundy
-% \affiliation{Department of Physics, Oregon State University,
-%   Corvallis, OR 97331}
-
-\section{Introduction}
-Flat histogram Monte-Carlo simulation algorithms calculate the thermodynamic
-properties of various systems over a range of temperatures.  The first histogram
-method used a single canonical Monte Carlo simulation to predict properties for
-nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
-samples a narrow energy range given limited by a narrow temperature range, it
-proves computationally inefficient at sampling large energy ranges.
-
-Due to a canonical Monte Carlo simulation's inefficient sampling of large
-energies, a variety of ``flat'' (or ``broad'') histogram
-methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-wang2001determining, wang2001efficient} were developed to explore a wider range
-of energies.  In addition to obtaining thermodynamic information for the entire
-energy range for a single simulation, these approaches cannot be easily trapped
-in a local energy minimum like a canonical simulation.
-
-Wang and Landau introduced one of the most widely used flat-histogram
-Monte-Carlo algorithms that accurately determined the density of states (DOS) for
-a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
-its power, the method unfortunately requires a priori knowledge of several
-user-defined parameters. Thus, for any given system under study, the user needs
-to determine the ideal parameters in order to apply the method. The Wang-Landau
-algorithm is also known to violate detailed balance (although only for brief
-time intervals)~\cite{yan2003fast, shell2002generalization}. With the violation
-of detailed balance, convergence of the algorithm is not guaranteed.
-
-Because of the uncertainty of convergence for WL, many studies were undertaken
-to understand how the modification (or update) factor $\gamma$ impacts the
-convergence~\cite{zhou2005understanding,lee2006convergence,
-belardinelli2007wang}. The ideal choice of decreasing the modification factor
-helps to prevent CPU time from being wasted by continuing to perform
-calculations after the error in the density of states is
-saturated~\cite{belardinelli2008analysis}. Modification factors that decrease
-faster than $1/t$ were found to lead to
-non-convergence~\cite{belardinelli2007fast}.  These findings led to the
-formation of the $1/t$-WL algorithm.
-
-Liang independently considered whether WL could be treated as a special case of
-Stochastic Approximation whose convergence could be mathematically
-proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
-al.~\cite{liang2007stochastic} argued that WL can be considered a form of
-Stochastic Approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
-convergence (if certain conditions are met). Despite the added benefit of
-guaranteed convergence, the method still has a system specific user-defined
-variable. Such variables often create difficulty when applying Monte-Carlo
-methods across arbitrary systems.
-
-Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
-related Statistical Temperature Molecular Dynamics (STMD), an adaption
-of the WL method that approximates the entropy as a piecewise linear function,
-which improves convergence for systems with a continuously varying
-energy~\cite{kim2006statistical, kim2007statistical}. STMC applied to WL
-requires a temperature range be specified rather than an energy range.  Kim
-\emph{et al.} extended this work as Replica Exchange Statistical Temperature
-Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping STMC
-simulations to improve convergence~\cite{kim2009replica}. Recently,
-Junghans \emph{et al.} demonstrated a close connection between
-metadynamics, which was introduced by Laio and
-Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods, with STMD
-forging the connection~\cite{junghans2014molecular}.
-
-The SAD (Stochastic Approximation with
-Dynamic $\gamma$) method outlined by Pommerenck
-et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
-that dynamically chooses the modification factor rather than relying on system
-dependent parameters. SAD shares the same convergence properties with SAMC while
-replacing un-physical user-defined parameters with the algorithms dynamic
-choice.
-
-In this work, we apply the family of weight-based flat histogram Monte
-Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
-
-\section{Ising Model}
-The 2D Ising spin-lattice system is widely used as a testbed when
-benchmarking or comparing Monte-Carlo
-methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The 2nd order
-phase transition behavior and the ability to directly calculate the
-exact solution for finite lattices~\cite{beale1996exact} make the
-system sufficiently interesting for such theoretical comparisons. It is
-also important to note that direct comparison of the other methods can
-be made with WL as its original implementation was done on this
-system~\cite{wang2001determining,wang2001efficient}.
-We test the convergence of several flat histogram methods
-on the periodic 2D square lattice ferromagnetic Ising model with nearest
-neighbor interactions~\cite{landau2004new}.
-\begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
-\end{align}
-The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
-for up or down spins respectively. In the absence of a magnetic field ($h =
-0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
-kaufman1949crystal}:
-\begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
-\end{align}
-where the sum is over nearest neighbor spin sites. Beale showed that for finite
-lattices the Density of States could directly be calculated from the partition
-function~\cite{beale1996exact}
-\begin{align}
-Z = \sum_E g(E) e^{-{\beta E}}
-\end{align}
-where $g(E)$ is the multiplicity of the system which is proportional to
-$D(E)$. We can estimate the average deviation from the exact solution
-using the relation~\cite{schneider2017convergence, shakirov2018convergence,
-barash2017control}:
-\begin{align}
-\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
-\end{align}
-As per the implementation by Wang and
-Landau~\cite{wang2001determining}, the total number of available energy
-states are $N-1$.
-
-\section{Flat histogram methods}\label{sec:histogram}
-Flat histogram methods determine the density of states $D(E)$ over a broad range
-of energies by simulating each energy with equivalent accuracy. Flat histogram
-Monte Carlo methods propose randomly chosen ``moves'' which change the state of
-the system and must satisfy detailed balance.  Each algorithm differs in how it
-determines the probability of accepting a move and in what additional statistics
-must be collected in order decide on that probability.
-
-We describe four closely related flat histogram methods which each rely on a
-weight function $w(E)$ to determine $D(E)$.  For these algorithms, the
-probability of accepting a move is given by
-\begin{equation}
-	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
-	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
-\end{equation}
-which biases the simulation in favor of energies with low weights. The result of
-weights $w(E)$ that are proportional to $D(E)$ is an entirely flat histogram. We
-can relate the entropy to the weights in the microcanonical ensemble, since the
-entropy is defined as $S(E) \equiv k_B\ln(D(E)) \approx \ln w(E)$.
-
-Each approach uses a random walk in energy space to estimate $D(E)$.  At the
-core of each of these approaches is to continuously update the weights at each
-step of the simulation
-\begin{equation}
-	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
-	+\gamma_t
-\end{equation}
-where $t$ is number of the current move, $\gamma(t)$ is a move-dependent update
-factor, and $E$ is the current energy.  This update causes the random walk to
-avoid energies that have been frequently sampled, leading to a rapid exploration
-of energy space. The four methods differ primarily in how they schedule the
-decreasing of $\gamma_t$.
-
-The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
-landau2014guide} explores energy space by initially setting $\gamma=1$, and then
-decreases $\gamma$ in discrete stages. An energy range of interest must be
-specified~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
-which can require multiple simulations if unknown.
-
-The number (``counts'') of moves ending at each energy are stored in a
-histogram.  For a sufficiently flat energy histogram (specified by the user to
-typically be 0.8), $\gamma$ is decreased by a specified factor of $\frac12$ and
-the histogram is reset to zero. The entire process is repeated until $\gamma$
-reaches a desired cutoff.
-
-Zhou et. al also showed that the WL algorithm never converges
-exponentially~\cite{zhou2008optimal} and succesfully bounded the statistical
-error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}. He also suggests
-that a modification factor of $1/t$ is optimal for WL.
-
-The $1/t$-WL method is a refinement of the original implementation that
-addresses the error saturation of WL. Belardinelli and Pereyra
-implemented a schedule~\cite{belardinelli2007fast} that enforced that
-if $\gamma < \nicefrac{1}{t}$ then the update factor is set to
-$\nicefrac{1}{t}$ and the histogram is no longer tracked. Using this
-metric, the error saturation is avoided since the correct density of
-states is approached asymptotically as
-$t^{-\frac12}$~\cite{belardinelli2008analysis}.
-Schneider $\text{et. al}$ outline minor refinements to the $1/t$-WL algorithm
-including $\nicefrac{N_S}{t}$ scaling and switching from standard WL to
-$1/t$-WL when the update factor $\gamma <
-\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the original
-$1/t$ implementation~\cite{belardinelli2007fast}, the update factor is
-decreased once all $N_S$ energy states are visited at least once (i.e.
-$H(E) \neq 0$) effectively avoiding the concept of `flatness'.
-
-Another weight-based flat histogram method is
-the Stochastic Approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
-schedule by which the update factor $\gamma$ is continuously
-decreased~\cite{liang2007stochastic, werlich2015stochastic,
-schneider2017convergence}.  The update factor is defined in the
-original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$,
-\begin{align}
-\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
-\end{align}
-where as above $t$ is the number of moves that have been attempted.
-
-SAMC offers extreme simplicity and proven convergence.
-Provided the update factor satisfies
-\begin{align}
-\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
-\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
-\end{align}
-where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
-to converge to the true density of states~\cite{liang2006theory,
-liang2007stochastic}.  Unlike WL methods, the energy range need not be
-known \emph{a priori.}  The convergenc time depends only on the choice of
-parameter $t_0$.
-The parameter $t_0$ can unfortunately be difficult to chose in advance
-for arbitrary systems.
-Liang \emph{et al.} give a rule of thumb in
-which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
-is the number of energy bins~\cite{liang2007stochastic}.  Schneider
-\emph{et al.} found that for the Ising model this heuristic is helpful
-for small spin systems, but that larger systems require an even higher
-$t_0$ value~\cite{schneider2017convergence}. Our work also confirms this.
-
-Pommerenck et al. propose a refinement~\cite{pommerenck2020stochastic} to SAMC
-where the update factor is determined dynamically rather than by the user.
-Stochastic Approximation with a dynamic $\gamma$ (SAD) requires the user to
-provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
-requiring a priori an energy range of interest; however, this may easier to
-identify and is more physical than the SAMC parameter $t_0$. The update factor
-can be written in terms of the current estimates for the highest $E_H$ and
-lowest $E_L$ energies of interest and the last time that an energy in the range
-of interest is encountered $t_L$.
-\begin{align}
-  \gamma_{t}^{\text{SAD}} =
-     \frac{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
-     }{
-       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
-     }
-\end{align}
-SAD only explores the energy range of interest as specified by the minimum
-temperature of interest $T_{\min} < T < \infty$. During the simulation the two
-energies $E_H$ and $E_L$, are refined such that the range of energies are conservatively
-estimated. The weights are calculated for each energy region according to the original
-prescription.
-\begin{enumerate}
-\item {$E < E_L$:} $w(E>E_H) = w(E_H)$
-\item {$E_L < E < E_H$:} moves are handled the same as other weight-based
-methods that are mentioned
-\item {$E > E_H$:} $w(E<E_L) = w(E_L)e^{-\frac{E_L-E}{T_{\min}}}$
-\end{enumerate}
-Each time the simulation changes the value of $E_H$ or $E_L$, the weights
-within the new portion of the interesting energy range are updated.
-
-\section{Results}
-We tested the algorithms on two 2D Ising model systems.  The first is a smaller
-simulation with a lattic size of 32 and the second has a lattice size of 128.
-For each system we use a reasonable root-mean-square displacement distance
-$\delta_0 = 0.05\sigma$ for proposed moves. The simulations explore the energy
-space of the systems with minimum reduced temperatures of $T_{\text{min}} = 1$
-for simulations All simulations lead to the minimum important energy $E_{\min}$
-and maximum entropy energy $E_{\max}$ being calculated (with the exception of
-the WL methods where both of these parameters are needed a priori).
-
-\begin{figure}
-\includegraphics[width=\textwidth]{../ising/gamma-n32.pdf}
-  \caption{
-  The update factor $\gamma_t$ versus the iteration number for the $N=32 \times 32$
-  system.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{../ising/N32-entropy-error-default.pdf}
-  \caption[The average entropy error for each MC method for $N=32 \times 32$,
-               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-               as a function of number of iterations run.]{
-  The average entropy error for each MC method for $N=32 \times 32$,
-               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-               as a function of number of iterations run.  The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded area.}\label{fig:n32}
-\end{figure}
-
-\subsection{The 32 $\times$ 32 Ising model}
-For this simulation, we chose a minimum reduced temperature of $1$, which
-corresponds to an interesting energy range from $-2048$ to $48$.  The number of
-important energy states for this system is $N_S = 529$.
-
-Fig.~\ref{fig:n32} shows the average error in the entropy as a function of time for
-this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed lines
-represent the average of the absolute value of the error in the entropy averaged
-over eight simulations using different random number seeds. The range of average
-errors for each simulation is shown as a shaded region around its mean error. By
-the time $10^8$ moves have been made all but the WL simulation have begun to
-converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^9$ moves.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{../ising/gamma-n128.pdf}
-  \caption{
-  The update factor $\gamma_t$ versus the iteration number for the $N=128 \times 128$
-  system.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{../ising/N128-entropy-error-default.pdf}
-  \caption[The average entropy error for each MC method for $N=128 \times 128$,
-               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-               as a function of number of iterations run.]{
-  The average entropy error for each MC method for $N=128 \times 128$,
-               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
-               as a function of number of iterations run.  The error is
-               averaged over 8 independent simulations, and the best
-               and worst simulations for each method are shown as a
-               semi-transparent shaded ar.}\label{fig:n128}
-\end{figure}
-
-\subsection{128 $\times$ 128 Ising system}
-For this simulation, we chose a minimum reduced temperature of $1$, which
-corresponds to an interesting energy range from $-32768$ to $100$.  The number of
-important energy states for this system is $N_S = 8214$.
-Fig.~\ref{fig:n128} shows the average error in the entropy as a function of time
-for this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed
-lines represent the average of the absolute value of the error in the entropy
-averaged over eight simulations using different random number seeds. The range
-of average errors for each simulation is shown as a shaded region around its
-mean error. By the time $10^{12}$ moves have been made all but the WL and SAD
-simulations have begun to converge as $1/\sqrt{t}$. We then see the WL error
-saturate around $10^{12}$ moves. The error in the SAD method will continue to
-converge faster than $1/\sqrt{t}$ until the update factor approaches that of
-1/t-WL at which point both methods will converge equivalently.
-
-\section{Conclusion}
-We have examined the convergence of SAD versus other widely used weight-based
-histogram methods.  We have shown that Sad samples the energy space
-corresponding to a desired range of temperatures for a variety of system sizes.
-We find that both SAD and 1/t-WL demonstrate excellent and robust convergence.
-They both converge more rapidly than SAMC, and unlike WL do not suffer from
-error saturation. We find that for larger systems SAD reduces the updated factor
-more slowly (and conservatively) than 1/t-WL. This means that SAD will take
-proportionately more moves to converge to the same value as 1/t-WL as system
-size is increased. SAD requires the user to specify a temperature range of
-interest rather than an energy range of interest as 1/t-WL does. For use cases
-in which a range of desired temperatures is known, this will make the SAD method
-considerably more convenient.
+\input{../ising/ising-body.tex}

### 2020-05-02 thesis/thesis.tex
updated the thesis to read in directly from manuscripts; still working on gas-adsorption.

@@
 \usepackage{graphicx}
-\graphicspath{{images/}}
+\graphicspath{{../square-well/}{../ising/}{../gas-adsorption/figs/}}
+
 \usepackage{caption}
-\usepackage{subcaption}
+%\usepackage{subcaption}
 \usepackage{float}
@@
 \usepackage{nicefrac}
+
+\usepackage{xargs}                      % Use more than one optional parameter in a new commands
+\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
+\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
+\usepackage{mdframed}
+\usepackage{braket}
 
@@
 
+% gas adsorption commands
+\usepackage{subfig}
+\usepackage{url}
+%
+\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
+
+\newcommand{\addcite}{{\bf \color{red} []}}
+\newcommand{\xvec}{\mathbf{x}}
+\newcommand{\thetavec}{\boldsymbol \theta}
+\newcommand\V{\Phi}
+\newcommand\Vk{\tilde\Phi(\vec k)}
+\newcommand\volume{V}
+\newcommand\pfull{\ensuremath{p_{\text{full}}}}
+\newcommand\pempty{\ensuremath{p_{\text{empty}}}}
+\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
+\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}
+\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
+\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}
+\newcommand\gst{\ensuremath{\Delta g_{st}}}
+
+\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
+\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }
+
+\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }
+
+\newtheorem{limitation}{Limitation}
+
 % THE BIBLIOGRAPHY NEEDS TO COME BEFORE ALL OF THE CHAPTERS
@@
 % gas-adsorption paper
-\input{chapter-4}
+%\input{chapter-4}
 % general conclusion linking all manuscripts

### 2020-05-04 ising/ising-body.tex
added coupling constant J and made beta/J for figure. Changed acknowledgement.

@@
 We test the convergence of several flat-histogram methods
-on the periodic 2D square lattice ferromagnetic Ising model with nearest
-neighbor interactions~\cite{landau2004new}.
+on the periodic 2D square lattice ferromagnetic Ising model with identical
+nearest neighbor interactions~\cite{landau2004new} ($J_{ij} = J$).
 \begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
+\mathcal{H} = -J \sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
 \end{align}
@@
 \begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
+\mathcal{H} = - J\sum_{\braket{i,j}} \sigma_i \sigma_j
 \end{align}
@@
 
-We thank Klas Markstr\"{o}m for discussions regarding the exact computation of the partition function for the 2D Ising model.
-We also wish to thank Johannes Zierenberg for helpful discussions regarding WL production run simulations and
-Martin Weigel for discussions regarding entropic population annealing.
+We wish to thank Johannes Zierenberg for helpful discussions regarding WL followed by a production run and for insights into properties of a production run at different stages of convergence.

### 2020-05-04 scripts-python/ising-cv-plot.py

@@ -72,7 +72,7 @@ for name in cv_headers:
     colors.plot(1/Temp, cv / N**2, method=name)
 
 colors.legend(loc='best')
-plt.xlabel(r'$k_B\beta$')
+plt.xlabel(r'$\beta / J$')
 plt.ylabel(r'$c_V$ / $ k_B$')
 plt.xlim(0.3,0.6)
 if N == 32:

### 2020-05-04 ising/ising-body.tex
added some text on ising phase transition.

@@
 \end{align}
-% As per the implementation by Wang and
-% Landau~\cite{wang2001determining}, the total number of available energy
-% states are $N-1$.
 Computing the specific heat capacity $c_V$ presents a difficult challenge for
 any Monte Carlo method due to fluctuations in the derivative of the
-internal energy.  Methods that accurately compute $c_V$ also by
-extension accurately compute the internal energy.
+internal energy around the phase transition. The critical temperature $T_c = 2.
+269T$ or $\beta= 0.441J$ comes directly from the Kramers-Wannier
+duality~\cite{bhattacharjee1995fifty} and marks the transition from a
+disordered to ordered magnetic state. Methods that accurately compute $c_V$
+also by extension accurately compute the internal energy.
 

### 2020-05-04 ising/ising-body.tex
update thesis.bib with ising bib and fix all errors in ising paper.

@@
 also by extension accurately compute the internal energy.
-
-\davidsays{talk about phase transition, e.g. at what temperature?}
 
@@
 
-All the methods except for SAMC gave a reasonable approximation for
-the heat capacity peak resulting from the phase transition within 10$^9$ moves.
-Figure~\ref{fig:N32-cv} shows the specific heat capacity vs.
-the reciprocal temperature $\beta$ at $10^{9}$ moves.
-Each method (except for the SAMC methods) is shown using the same single random number seed.
+All the methods, except for SAMC, use the same single random number seed and
+give a reasonable approximation for the heat capacity peak resulting from the
+phase transition within 10$^9$ moves. Figure~\ref{fig:N32-cv} shows the
+specific heat capacity vs. the reciprocal temperature $\beta$ at $10^{9}$ moves.
+A temperature range of $1.5k_B$ to $5k_B$ is chosen to highlight
+the phase transition at the critical temperature $T_c$.
 
@@
 
-\davidsays{Explain the range of temperature compared, and why} Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
+Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
 this system. The solid/dashed lines
@@
 \subsection{128 $\times$ 128 Ising system}
-Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each
-of the flat-histogram methods. All of the update factors initially
-start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$
-throughout the simulation. After about $10^{13}$ moves,
-$\gamma^{\text{SAD}}_t$ proceeds as $1/t$. \davidsays{fixme similarly}
-Both WL and ``production'' WL update factors $\gamma^{\text{WL}}_t$ are shown.
-The production WL begins after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$.
-The update factor for $1/t$-WL decreases similarly to WL before finding all the energy states and switching to $1/t$.
-All of the SAMC update factors equal 1 until the number of moves is equal to $t_0$ at which point they decrease as $1/t$.
-
-Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a function of time for
-this system. The solid/dashed lines
-represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
+For the larger 2D Ising system, the update factors initially
+start at the same $\gamma_{t=0} = 1$; however, all of the methods take longer
+to proceed as $1/t$ (with SAD taking $10^{13}$ moves). Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the 
+flat-histogram methods. 
+We implement two WL simulations followed by a production run each beginning
+after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$ respectively.
+Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a 
+function of time for this system. The solid/dashed lines represent the average
+of the maximum value of the error in the specific heat capacity $c_V$ averaged
 over eight simulations using different random number seeds. The range of maximum
@@
 the time $10^{13}$ moves have been made all but the WL simulation have begun to
-converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$ moves.
+converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$
+moves. The two WL simulations followed by a production run highlight that
+choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
+properties.
 
 \section{Conclusion}
-\davidsays{remove this?}
-The method performs as well as 1/t-WL and SAD.  Unfortunately, the addition of another tunable parameter $\gamma_{t}^{\min}$ makes the method somewhat more difficult to implement in practice.
-
 We find that both SAD, 1/t-WL, and WL followed by a production run
-(with an adequately small $\gamma^{\min}$) demonstrate excellent and robust convergence.
-They all converge more rapidly than SAMC, and unlike pure WL do not suffer from
+(with an adequately small $\gamma^{\min}$) demonstrate excellent and robust
+convergence. They all converge more rapidly than SAMC, and unlike pure WL do
+not suffer from
 error saturation. We find that for larger systems SAD reduces the updated factor
@@
 size is increased. SAD requires the user to specify a temperature range of
-interest rather than an energy range of interest as the WL methods do.
-The 2D Ising model is a system where the energy range is known \emph{a priori.}, which makes a comparison between WL methods and SAD (which requires a minimum temperature) more biased towards WL methods.
-For the majority of cases in which the energy range of the system is not known and where a range of desired temperatures is known, this will make the SAD method
+interest rather than an energy range of interest as the WL methods do. The 2D
+Ising model is a system where the energy range is known \emph{a priori.}, which
+makes a comparison between WL methods and SAD (which requires a minimum
+temperature) more biased towards WL methods. For the majority of cases in which
+the energy range of the system is not known and where a range of desired
+temperatures is known, this will make the SAD method
 considerably more convenient. This work also visually demonstrates
-that WL followed by a production run performs extremely well and is preferable to WL for ensuring both ergodicity and detailed balance.
+that WL followed by a production run performs extremely well and is preferable
+to WL for ensuring both ergodicity and detailed balance.
 

### 2020-05-05 gas-adsorption/gas-ads-appendix.tex
update the paper and conform to all the thesis guidlines.

--- 2020-05-02 gas-adsorption/gas-ads.tex
@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp]{revtex4-1}
-\usepackage{graphicx} % Include figure files
-\graphicspath{{figs/}}
-\usepackage{color}
-\usepackage{amsmath}
-\usepackage{subfig}
-
-\usepackage{nicefrac} % Include for inline fractions
-\usepackage{url}
-
-% \usepackage{cite}
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage{color}  % Coloured text etc.
-%
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
-\usepackage{mdframed}
-
-\newcommand{\addcite}{{\bf \color{red} []}}
-\newcommand{\xvec}{\mathbf{x}}
-\newcommand{\thetavec}{\boldsymbol \theta}
-
-\newcommand\V{\Phi}
-\newcommand\Vk{\tilde\Phi(\vec k)}
-\newcommand\volume{V}
-
-\newcommand{\pink}[1]{{\bf \color{pink} #1}}
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\orange}[1]{{\bf \color{orange} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
-
-%\newcommandx{\davidsays}[1]{\begin{mdframed}[backgroundcolor=blue!24]\textbf{David says:} #1\end{mdframed} }
-%\newcommandx{\davidsays}[2][1=inline]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{David says:} #2} }
-\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
-\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
-\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }
-
-%% \renewcommand\davidsays[1]{}
-%% \renewcommand{\corysays}[2][1=inline]{}
-%TC:macro \corysays [ignore]
-%TC:macro \davidsays [ignore]
-%TC:macro \jordan [ignore]
-
-\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }
-
-\newtheorem{limitation}{Limitation}
-
-\begin{document}
-\title{
-An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}
-% other ideas:
-%Gas delivery via pressure-swing adsorption in nanoporous materials: setting storage targets}
-% A thermodynamic upper bound to gas delivery via pressure-swing adsorption in nanoporous materials:}
-
-\author{Jordan K. Pommerenck}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-\author{Cory M. Simon}
-\affiliation{School of Chemical, Biological, and Environmental Engineering,
-  Corvallis, OR 97331}
-\author{David Roundy}
-\affiliation{Department of Physics, Oregon State University,
-  Corvallis, OR 97331}
-
-\begin{abstract}
-  Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles [].
-One strategy to densify these gases is to pack the fuel tank with a porous material [].
-  Metal-organic frameworks are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases [].
-The US Department of Energy (DOE) has set volumetric deliverable capacity targets [] which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels.
-Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction []; none have established a rigorous upper bound on the deliverable capacity.
-  Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing.
-% "pressure-swing storage doesn't make sense to me.
-storage and delivery  maybe but awk.
-  % I added rigid since that could be important...
-  To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
-  \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
-  Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
-  We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-  % However, this upper bound directly leads to important realizations which should guide future development.
-  However, limitations to the scope of applicability of our upper bound guide future material development efforts.
- % Firstly, one could extract the gas at a higher temperature than that used while filling [].
-  Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
-  Secondly, the fundamental physics of our upper bound do not pertain to
-  % rule out
-  any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
-  % firstly, secondly vs first, second?
-\end{abstract}
-
-\newcommand\pfull{\ensuremath{p_{\text{full}}}}
-\newcommand\pempty{\ensuremath{p_{\text{empty}}}}
-
-\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
-\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}
-
-\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
-\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}
-
-\maketitle
-
-\newcommand\gst{\ensuremath{\Delta g_{st}}}
-
-\paragraph*{to do}
-\begin{enumerate}
-    \item See david/corysays
-    \item make the code presentable
-    \item do the FIXME in code
-\end{enumerate}
-
-\input{gas-ads-body.tex}
-
-\bibliography{bibfile}
-
 \appendix
@@
 So the answer is not so simple.}
-
-%TC:endignore
-
-\end{document}

### 2020-05-05 gas-adsorption/gas-ads.tex
update the paper and conform to all the thesis guidlines.

@@
 
-\appendix
-\clearpage
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}%
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}%
-\renewcommand{\thesection}{SI \Alph{section}}%
-
-\section{An external, spatially uniform potential $\V$ shifts the chemical potential $\mu$ of the gas} \label{sec:V_shifts_chem_pot}
-We show that imposing an external, spatially uniform potential $\V$ to a gas has the effect of shifting the chemical potential $\mu$ of the gas, recovering eqn.~\ref{eq:mof-density}.
-Consider a control volume $\Omega$ with volume $V=|\Omega|$ (large enough to neglect boundary effects) that is endowed with the external, spatially uniform potential $\V$.
-Impose the grand-canonical ensemble, where this control volume can exchange energy and particles with a bath of gas at temperature $T$ and chemical potential $\mu$.
-
-To denote a microstate of this system, let $N$ be the number of gas particles in the control volume and $\mathbf{x}_1,...,\mathbf{x}_N$ be their positions.
-Then, the potential energy $E$ of a microstate of the control volume is:
-\begin{equation}
-    E(\mathbf{x}_1,...,\mathbf{x}_N) = N\V + E_{gg}(\mathbf{x}_1,...,\mathbf{x}_N),
-\end{equation} where $E_{gg}$ is the (unknown and complicated) interatomic potential for gas-gas interactions that governs the (real) gas properties.
-The first term arises from each gas molecule feeling the external potential $\V$, where $\V < 0$ corresponds to attraction.
-
-\corysays{wut about orientations? shld we include those?}
-
-The grand-canonical partition function of the control volume is:
-\begin{multline}
-    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\xvec_1, ..., \xvec_N)} e^{\beta (\mu - \V) N} d\xvec_1 \cdots d\xvec_N.
-    \label{eq:gcpf}
-\end{multline} We recognize this as the grand canonical partition function of the bulk gas in the control volume without the external potential, but shifted in chemical potential:
-\begin{equation}
-    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
-    \label{eq:xi_vs_xi0}
-\end{equation}
-where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the bulk gas in the control volume in the absence of an external potential.
-Importantly, this equivalency is predicated on the potential $\V$ being spatially uniform.
-Therefore, the thermodynamic properties of the gas sitting in the spatially uniform external potential $\V$-- i.e., adsorbed in our idealized substrate-- are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$ ($T$ fixed here).
-As $\V$ becomes more negative, corresponding to a more attractive adsorbent, the thermodynamic properties of the adsorbed gas in our ideal substrate are equivalent to the gas at higher chemical potential.
-
-\section{Proof of extremum}
-To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.
-In this proof we will make use of the Fourier transform of this potential:
-\begin{align}
-    \Vk \equiv \iiint \V(\vec r) e^{-i\vec k\cdot \vec r} d\vec r
-\end{align}
-The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$.
-Therefore, to show that a uniform potential extremizes the deliverable capacity, we must show that the functional derivative of the deliverable capacity with respect to $\Vk$ is zero for \emph{nonzero} values of $\vec k$, i.e.
-\begin{align}
-    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \vec k\ne 0.
-\end{align}
-We note that this functional derivative may be non-zero for $\vec k=0$ because we separately maximize with respect to the particular uniform potential $\V$.
-This means that
-\begin{align}
-    \frac{\delta N_H}{\delta \V(\vec k)} &= -\frac{\delta N_L}{\delta \Vk}
-\end{align}
-where $N_H$ and $N_L$ are the number of particles at the low and high pressure.
-
-Because the chemical potential $\mu$ varies monotonically with $N$ at fixed temperature, we can consider how the chemical potential varies as we change $\Vk$ with the number of molecules held fixed.
-We demonstrate this using the cyclic chain rule, which shows us that
-\begin{align}
-    \left(\frac{\delta N}{\delta \Vk}\right)_{\mu} &=
-    -\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}
-    \left(\frac{\partial N}{\partial \mu}\right)_{\Vk}.
-\end{align}
-Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \V(\vec k)}\right)_{\mu}=0$.
-Thus we consider
-\begin{align}
-    \left(\frac{\delta \mu}{\delta \tilde\V(\vec k)}\right)_N
-    &= \left(\frac{\delta \left(\frac{\partial F}{\partial N}\right)_{\volume}}{\delta \Vk}\right)_N
-    \\
-    &= \left(\frac{\partial \left(\frac{\delta F}{\delta \Vk}\right)_{\volume}}{\partial N}\right)_{\volume}
-    \label{eq:dmudpot}
-\end{align}
-where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reordered the functional and partial derivatives.
-Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ just gives the number density:
-\begin{align}
-    \frac{\delta F}{\delta \Vk} &= \rho(\vec k)
-\end{align}
-The number density is itself spatially uniform for any system that is stable in a fluid state at this density (i.e.
-does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.
-Thus the functional derivative $\frac{\delta F}{\delta \V(\vec r)}$ is actually a uniform function.
-We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
-\begin{align}
-    \frac{\delta \mu}{\delta \Vk} &\propto \delta(\vec k) \\
-    \frac{\delta N}{\delta \Vk} &\propto \delta(\vec k)
-\end{align}
-Thus the functional derivative of both the chemical potential and $N$ with regard to $\V(\vec r)$ are themselves spatially uniform.
-Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e.
-$\vec k=0$), the derivative of $D$ with respect to any change of potential is zero.
-
-This demonstrates that a spatially uniform potential leads to an extremum value of the deliverable capacity.  This proof is insufficient, however, to show that it must be a true maximum.
-
-\section{The real-substrate analogy of $\V$ is $\gst$}
-\label{sec:phi-is-delta-g}
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{g_st_piston_2.png}
-    \caption{Cartoon illustrating the definition of $\gst$ and a hypothetical experiment to measure it with a piston and removable barrier that separates two equal volumes, one with a porous material (bottom) and one with free space (top).
-The initial state (left) is a bulk gas phase at density $\rho$ and temperature $T$.
-An infinitely-thin, removable barrier separates the gas from the porous material (bottom).
-The barrier is then slowly removed, and the piston slowly pushes all gas molecules into the porous material in an isothermal process.
-The final state (right) is an adsorbed gas phase at density $\rho$ and temperature $T$.
-The quantity $\gst$ follows from {\color{red} the heat released?}.}
-    \label{fig:delta-G-cartoon}
-\end{figure}
-
-The parameter describing our idealized substrate is $\V$, the spatially uniform potential felt by a gas molecule adsorbed in the idealized substrate. 
-A natural question is how this potential relates to the properties of real substrates. 
-The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the gas (see Sec.~\ref{sec:V_shifts_chem_pot}).
-Because our ideal substrate shifts the chemical potential of the gas molecules by providing a spatially uniform potential energy field, the entropy of the gas in the ideal substrate is equal to the entropy of the gas in its bulk state at the same density and temperature.
-In contrast, a real substrate provides a non-spatially uniform potential.
-Consequently, the entropy of the gas inside a real substrate is \emph{not} equal to the entropy of the bulk gas at the same density and temperature.
-Therefore, the parameter analogous to $\V$ in a real substrate will involve both energy and entropy.
-The real-substrate analogy of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
-\begin{equation}
-   \gst(\rho, T) \equiv
-   %\frac{G_{\text{ads}}(\rho, T) - G_{\text{gas}}(\rho, T)}{N}\\ 
-    \mu_{\text{ads}}(\rho, T) - \mu_{\text{gas}}(\rho, T).
-  % \gst(\rho, T) &\equiv g_{\text{gas}}(\rho, T) - g_{\text{ads}}(\rho, T)
-  \label{eq:g_st}
-\end{equation}
-The isosteric Gibbs free energy difference $\gst$ is the difference in molar Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and the bulk gas \emph{with the same density of gas molecules}.
-The quantity $\gst$ does \emph{not} correspond to a change in the molar Gibbs free energy as a molecule is adsorbed, which is zero under conditions of coexistence.
-The quantity $\gst$ in a real substrate is a direct analogy to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
-Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to  measure $\gst$ via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
-Note that $\gst$ is a property of both the substrate and the identity of the gas.
-Because real substrates offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho, T)=\V$.
-Consequently, throughout this article, we show $\gst(\rho, T)$ for real substrates at both conditions relevant to gas storage and delivery, $\pfull$ and $\pempty$.
-% In the initial state, the partition isolates the bulk gas exhibiting density $\rho$ and temperature $T$.
-Then, the partition is removed and the piston pushes down on the gas so that the gas slowly adsorbs into the MOF in an isothermal process.
-\red{isothermal rite?} In the final state, the piston has pushed all gas into the MOF, resulting in an adsorbed gas phase with the same density $\rho$ and temperature $T$ as the initial state.
-The isosteric Gibbs free energy of adsorption, $\gst$, then follows from \red{is it the amount of heat released?}.
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} looks too jumbled. two separate figs? plus $G_{st}$ is capitalized here, should be lowercase. where is all this data from}
-
-In practice, we can readily compute $\gst(\rho, T)$ of a real gas/substrate system from (i) the (experimental or simulated) equilibrium adsorption isotherm of the gas in the substrate and (ii) the (experimental or simulated) chemical potential of the bulk gas.
-Consider the real substrate in thermodynamic equilibrium with a bulk gas at fixed temperature $T$ and pressure $p$, and let $\rho=\rho(p, T)$ be the density of gas in the substrate. 
-At coexistence, the chemical potential of the bulk gas is equal to the chemical potential of the adsorbed gas in the substrate.
-Thus, we can use the experimentally known molar Gibbs free energy of the pure, bulk gas system at temperature $T$ and pressure $p$ to determine the molar Gibbs free energy of the adsorbed system: $\mu_{\text{ads}}(\rho, T)=\mu_{\text{gas}}(p, T)$.
-We can then also look up the known chemical potential of the bulk gas at the same density and temperature as in the substrate, $\mu_{\text{gas}}(\rho, T)$.
-Via eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.
-
-An interesting question is how $\gst$ relates to the commonly measured and reported isosteric heat of adsorption $q_{st}$, which is roughly the energy change when a gas molecule is adsorbed~\cite{sircar1999isosteric, tian2017differential}.
-Figure~\ref{fig:qst-vs-delta-G} shows how $q_{st}$ compares to $\gst$ for several prominent adsorbents.
-In every case, $|q_{st}|>|\gst|$ because the gas in the adsorbent always has less entropy than the gas in the bulk at the same density and temperature.
-That is, while adsorption is energetically favored, it is entropically disfavored due to the restrictions imposed on the configuration of the gas molecules via steric interactions with the substrate itself; this counters the energetic attraction.
-
-\corysays{can u mathematically justify why $\gst$ is the relevant parameter through the grand canonical partition function in eqn.~\ref{eq:gcpf}?
-it is not true that $\Xi(\mu, V, T)$ for the real substrate is equal to $\Xi(\mu+\gst, V, T)$}
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} should be split in two. one for methane one for h2.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{qst-vs-delta-G}
-    \caption{Relationship between $\gst$ and the isosteric heat $q_{st}$ for several prominent adsorbents.
-The dots represent the properties of methane adsorption at 298\ K and at 5\ bar and 100\ bar.
-The $\times$'s represent the properties of hydrogen adsorption at 77\ K and at 5~bar and 100~bar.}
-    \label{fig:qst-vs-delta-G}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{methane-298-gst}
-    \caption{The density-dependence of $\gst$ of methane in several adsorbents (298\ K). Note that $\gst$ is monotonic in $\rho$. Data from Refs.~\ref{mason2014evaluating,furukawa2009storage}.
-    }
-    \label{fig:methane-gst}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{four-cases}
-    \caption{Four cases we consider.}
-    \label{fig:delta-gst-maximum}
-\end{figure}
-
-\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
-Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding to a full and empty density.
-The deliverable capacity of the MOF is equal to the difference between the full and empty density.
-Examining Fig.~\ref{fig:methane-gst}, we see that a significant variety of known experimental $\Delta g_\text{st}$ curves are monotonic.
-Furthermore, our qualitative argument in Section~\ref{sec:upper-bound} suggests that this function \emph{should} be monotonically increasing for rigid substrates, as increasing the density of gas forces some of the gas to reside in higher-energy sites.
-If this is always the case, the deliverable capacity will always be bounded by our theoretical upper bound, even for a value of $\V$ that is not optimal.
-
-To show this, we perform a thought experiment illustrated in Fig.~\ref{fig:delta-gst-maximum}.
-We begin with two pairs of boxes containing gas.
-Two of the boxes are filled with a MOF, and the other two have a uniform potential energy $\V$.
-We could imagine the two boxes at $\V$ as being empty but at very low altitude where gravity provides a potential energy difference (relative to the box with the MOF in it) of $\V$ to each gas molecule (perhaps we are on Jupiter?).
-One box of each sort has a density of gas $\rhofull$ corresponding to $\pfull$ in the MOF system, while the other two boxes has gas density $\rhoempty$.
-
-We now consider what happens if we introduce a diffusive connection between two boxes that are initially at the same density, for instance by connecting them with a tube.
-If the chemical potential of the two boxes are equal, their density will be unchanged, otherwise gas will flow into the box with lower chemical potential.
-The difference in chemical potential is equal to the difference
-\begin{align}
-   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
-   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
-   \\
-   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
-\end{align}
-Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at their initial density after they are connected.
-Thus the deliverable capacity of the MOF will be greater than the deliverable capacity of the model with uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e.
-if $\gst(\rho)$ does \emph{not} monotonically increase.
-By the same token, if we consider the case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable capacity than our model, the MOF must have less residual gas, which means that gas must spontaneously flow from the low-density MOF to the box with potential $\V$, which means that $\gst(\rhoempty)>\V$.
-So once again exceeding our upper bound requires a material with a $\gst(\rho)$ that does not increase monotonically.
-
-Taken together, this indicates that not only is our absolute upper bound an upper bound for rigid MOFs, but the green curve labeled $\rho_D(\gst)$ in Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and \ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal $\gst$.
-
-\section{Cryogenic hydrogen storage}\label{sec:cryo-hydrogen}
-\begin{figure}
-    \centering
-    FIGURE HERE A B C D
-    \includegraphics[width=0.95\columnwidth]{hydrogen-77-n-vs-G}
-    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
-    \label{fig:hydrogen-77-D}
-\end{figure}
-
-One approach to increase the deliverable capacity is to reduce the storage temperature.
-This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling point of nitrogen.
-The DOE ULTIMATE target in this case looks far more achievable, and with a much lower $|\gst|$.
-In fact, an empty tank at this temperature can satisfy the DOE 2020 target.
-The DOE ULTIMATE target is 14\% below the upper bound.
-Actual MOFs fall far short of the theoretical maximum.
-\red{the target is not just a storage density, it specifies temps and pressures too...
-is 77 K allowed?} \davidsays{My recollection of my reading of the DOE target was that anything was allowed, but that the target included ``everything'', which would include both the tank and any cryogenic equipment required.
-So the answer is not so simple.}
-
-%TC:endignore
+\input{gas-ads-appendix.tex}
 

### 2020-05-06 ising/ising-body.tex
made a number of ising edits and a thesis edit.

@@
 where the update factor is determined dynamically rather than by the user.
-stochastic approximation with a dynamic $\gamma$ (SAD) requires the user to
-provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
+Stochastic approximation with a dynamic $\gamma$ (SAD) requires the user to
+provide a minimum temperature of interest $T_{\min}$. This is analogous to WL
 requiring \emph{a priori.} an energy range of interest; however, this is almost always easier to
@@
 can be written in terms of the current estimates for the highest $E_H$ and
-lowest $E_L$ energies of interest and the last time that an energy in the range
+lowest $E_L$ interesting energies and the last time that an energy in the range
 of interest is encountered $t_L$.
@@
 temperature of interest $T_{\min} < T < \infty$. During the simulation the two
-energies $E_H$ and $E_L$, are refined such that the range of energies are conservatively
-estimated. The weights are calculated for each energy region according to the original
-prescription.
+energies $E_H$ and $E_L$, are refined such that the range of energies are
+conservatively estimated. The weights are calculated for each energy region according to the original prescription.
 \begin{enumerate}
@@
 \section{Results}
-We tested the algorithms on two different system sizes of the 2D Ising model.  The first is a smaller
-simulation with a lattice size of $N = 32 \times 32$ and the second has a lattice size of $N = 128 \times 128$. The SAD method explores the energy
-space of each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations lead to the minimum important energy $E_{\min}$
-and maximum entropy energy $E_{\max}$ being calculated (with the exception of
-the WL methods where both of these parameters are needed \emph{a priori.}).
+We test the algorithms on two different system sizes of the 2D Ising model.
+The first is a smaller simulation with a lattice size of $N = 32 \times 32$ and
+the second has a lattice size of $N = 128 \times 128$. The SAD method explores 
+the energy space of each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations calculate the minimum important energy $E_{\min}$
+and maximum entropy energy $E_{\max}$ (with the exception of the WL methods
+where both of these parameters are needed \emph{a priori.}).
 
@@
 
-All the methods, except for SAMC, use the same single random number seed and
+All of the methods, except for SAMC, use the same single random number seed and
 give a reasonable approximation for the heat capacity peak resulting from the
-phase transition within 10$^9$ moves. Figure~\ref{fig:N32-cv} shows the
+phase transition after only 10$^9$ moves. Figure~\ref{fig:N32-cv} shows the
 specific heat capacity vs. the reciprocal temperature $\beta$ at $10^{9}$ moves.
@@
 
-Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a function of time for
-this system. The solid/dashed lines
-represent the average of the maximum value of the error in the specific heat capacity $c_V$ averaged
+Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a
+function of time for this system. The solid/dashed lines represent the average
+of the maximum value of the error in the specific heat capacity $c_V$ averaged
 over eight simulations using different random number seeds. The range of maximum
@@
 Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a 
-function of time for this system. The solid/dashed lines represent the average
-of the maximum value of the error in the specific heat capacity $c_V$ averaged
+function of time for this system. The solid/dashed lines represent the maximum
+value of the error in the specific heat capacity $c_V$ averaged
 over eight simulations using different random number seeds. The range of maximum
@@
 that WL followed by a production run performs extremely well and is preferable
-to WL for ensuring both ergodicity and detailed balance.
+to pure WL for ensuring both ergodicity and detailed balance.
 

### 2020-05-06 thesis/chapter-1.tex
working on introduction fitting together.

@@
 
-Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles~\cite{?}.
-One strategy to densify these gases is to pack the fuel tank with a porous material~\cite{?}.
-Metal-organic frameworks (MOFs) are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases~\cite{?}.
-The US Department of Energy (DOE) has set volumetric deliverable capacity targets~\cite{?} which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels.
-Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction~\cite{?}; none have established a rigorous upper bound on the deliverable capacity.
+% This is part of the abstract of our paper.
+Due to their low volumetric energy density at ambient conditions, both hydrogen
+and natural gas are challenging to economically store onboard environmentally
+sustainable vehicles~\cite{?}. One strategy to densify these gases is to pack
+the fuel tank with a porous material~\cite{?}. Metal-organic frameworks (MOFs)
+are tunable, nanoporous materials with large internal surface areas and show
+considerable promise for densifying gases~\cite{?}. The US Department of Energy
+(DOE) has set volumetric deliverable capacity targets~\cite{?} which, if met,
+would help to enable commercial adoption of hydrogen/natural gas as
+transportation fuels. Many have attempted to establish theoretical upper bounds
+on the deliverable capacity via pressure-swing adsorption using simplified
+models of the gas-substrate and gas-gas interaction~\cite{?}; none have
+established a rigorous upper bound on the deliverable capacity.
 
-Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing. 
-To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
-Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
-We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-However, limitations to the scope of applicability of our upper bound guide future material development efforts.
-Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes~\cite{?}.
-Secondly, the fundamental physics of our upper bound do not pertain to
-any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target~\cite{?}.
+% This is part of the abstract(s) for APS Conferences.
+Here, we present a theoretical upper bound on the deliverable capacity of a
+rigid material via an isothermal pressure-swing. To provide an extremum, we
+consider a substrate that provides a spatially uniform potential energy field
+for the gas. Our bound is unique in that it directly relies on experimentally
+measured properties of the (real) bulk gas, without making approximations.
+We conclude that the goals set by the US DOE for natural gas and hydrogen
+storage are theoretically possible, but sufficiently close to the upper bound
+as to be impractical for any real porous material. However, limitations to the
+scope of applicability of our upper bound guide future material development efforts. Firstly, one could heat the adsorbent to drive off residual gas when
+the pressure diminishes~\cite{?}. Secondly, the fundamental physics of our
+upper bound do not pertain to any material that changes its structure in
+response to adsorbed gas, suggesting that flexible materials could still
+satisfy the DOE targets~\cite{?}.
 
-Consider a pressure vessel onboard a vehicle (i.e.
-fuel tank) packed with porous material.
-At the filling stage, the tank is connected and opened to a (pure) gaseous reservoir at pressure $p_\text{full}$ and allowed to equilibrate.
-At this point, the adsorbed gas tank is considered full.
-While driving, gas desorbs from the adsorbent to the engine/fuel cell, driven by a pressure differential.
-The tank is considered depleted/empty when the pressure has dropped to $p_\text{empty}$, the pressure at which the flow rate of gas from the tank to the engine is insufficient.
-
-However, given $p_\text{empty} \neq 0$ (pulling vacuum), residual gas will remain trapped in the adsorbent.
-Therefore, the driving range of the vehicle is primarily determined by the \emph{deliverable capacity}, defined as the density of gas in the tank at $p_\text{full}$ minus the residual gas remaining at $p_\text{empty}$.
-The [isothermal, volumetric] deliverable capacity is an intrinsic property of the nanoporous material and its interaction with the gas.
-
-\jpsays{explain how 2D SAD is needed for this and how we are working toward it}
+% Now add why a multidimensional 2D SAD would help and how we are working toward it.
 
 \section{Summary}
-We have provided the background necessary for understanding the various flat-histogram methods and the
-various applied test systems. In the following chapters, we detail our manuscripts and the method SAD
-applied to the various physical test systems. We test SAD against a variety of other weight-based flat-histogram
-methods on the square-well fluid, Lennard-Jones clusters, and the 2D Ising model. Each of these systems provides
-unique challenges to our newly developed method SAD and convergence robustness is determined best by the application
-to multiple systems.
-
-\jpsays{Tie this in with gas adsorption!}
+We have provided the background necessary for understanding the various
+flat-histogram methods and the various applied test systems. In the following
+chapters, we detail our manuscripts and the method SAD applied to the various
+physical test systems. We test SAD against a variety of other weight-based
+flat-histogram methods on the square-well fluid, a Lennard-Jones cluster, and
+the 2D Ising model. Each of these systems provides unique challenges to our
+newly developed method SAD and convergence robustness is determined best by the
+application to multiple systems.
+% Now tie this in with gas adsorption.

### 2020-05-06 gas-adsorption/gas-ads-body.tex
multiple changes to introduction and added citations to gas-ads abstract.

@@
 % (200 bar for CH$_4$ to achieve 9.2 MJ/L \cite{makal2012methane} and 700 bar for H$_2$ to achieve 4.5-5.3 MJ/L~\cite{unknown}).
-Both approaches require expensive infrastructure at refilling stations and significant energy input; e.g., liquefaction of hydrogen consumes at least 30\% of its energy content \cite{bossel2003energy}.
-Moreover, high-pressure storage tanks are heavy, thick-walled, and non-conformable, while cryogenic storage tanks are bulky, expensive, and afflicted by boil-off losses \cite{hasan2009minimizing}.
+Both approaches require expensive infrastructure at refilling stations and significant energy input; e.g., liquefaction of hydrogen consumes at least 30\% of its energy content~\cite{bossel2003energy}.
+Moreover, high-pressure storage tanks are heavy, thick-walled, and non-conformable, while cryogenic storage tanks are bulky, expensive, and afflicted by boil-off losses~\cite{hasan2009minimizing}.
 % \jordan{
@@
 
-A promising approach to densify natural gas \cite{makal2012methane,mason2014evaluating} and hydrogen \cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is through physical adsorption in nanoporous materials at room temperature~\cite{schoedel2016role}.
+A promising approach to densify natural gas~\cite{makal2012methane,mason2014evaluating} and hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is through physical adsorption in nanoporous materials at room temperature~\cite{schoedel2016role}.
 The internal surfaces of porous materials attract gas molecules through van der Waals, electrostatic, etc.
@@
 For a vehicle employing a porous material to store natural gas or hydrogen, the (volumetric) \emph{deliverable capacity} of the gas is the primary thermodynamic property of the material that determines the driving range~\cite{mason2014evaluating}.
-The adsorbed gas storage tank delivers the gaseous fuel to the engine via an (assumed) isothermal pressure swing \cite{sircar2002pressure}.
+The adsorbed gas storage tank delivers the gaseous fuel to the engine via an (assumed) isothermal pressure swing~\cite{sircar2002pressure}.
 The deliverable capacity (see Fig.~\ref{fig:delcap}) is the density of the gas in the material at the storage pressure $\pfull$ minus the residual gas that remains adsorbed at the lowest pressure $\pempty$ such that sufficient flow is maintained to feed the engine.
 For commercial feasibility, the US Department of Energy (DOE) has set deliverable capacity targets for both adsorbed methane and hydrogen storage and delivery for vehicles.
-For methane, the Advanced Research Projects Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$ per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure swing \cite{simon2015materials}.
-For hydrogen, the DOE set a series of progressive targets at five year intervals, with the ultimate target of 0.05 kg H$_2$/L \cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a minimum of -40 $^\circ$C \cite{allendorf2018assessment}.
-Thus far, despite the emergence of highly tunable materials with enormous surface areas, such as metal-organic frameworks \cite{furukawa2013chemistry}, no porous material has met these deliverable capacity targets.
+For methane, the Advanced Research Projects Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$ per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure swing~\cite{simon2015materials}.
+For hydrogen, the DOE set a series of progressive targets at five year intervals, with the ultimate target of 0.05 kg H$_2$/L~\cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a minimum of -40 $^\circ$C~\cite{allendorf2018assessment}.
+Thus far, despite the emergence of highly tunable materials with enormous surface areas, such as metal-organic frameworks~\cite{furukawa2013chemistry}, no porous material has met these deliverable capacity targets.
 
@@
 
-G\'omez-Gualdr\'on \emph{et al.} \cite{gomez2017impact} introduced a model that accounted for gas-gas interactions via an intermolecular potential and idealized the substrate in two different ways: (1) discrete adsorption sites packed into an FCC lattice and (2) a volume endowed with a spatially uniform, background potential energy field.
+G\'omez-Gualdr\'on \emph{et al.}~\cite{gomez2017impact} introduced a model that accounted for gas-gas interactions via an intermolecular potential and idealized the substrate in two different ways: (1) discrete adsorption sites packed into an FCC lattice and (2) a volume endowed with a spatially uniform, background potential energy field.
 According to molecular simulation of methane adsorption, the ARPA-E deliverable capacity target of 315 L STP/L could be reached in both of these idealized substrates.
@@
 %More works obtained an empirical upper bound on the deliverable capacity by explicitly accounting for the atoms of the material needed to create the potential energy field felt by the gas molecules.
-Simulations of methane adsorption in hundreds of thousands of explicit nanoporous crystal structures---both real and hypothetical---suggested that the ARPA-E deliverable capacity target is infeasible \cite{simon2015materials}; the highest simulated methane deliverable capacity was 196 cm$^3$~STP/cm$^3$.
+Simulations of methane adsorption in hundreds of thousands of explicit nanoporous crystal structures---both real and hypothetical---suggested that the ARPA-E deliverable capacity target is infeasible~\cite{simon2015materials}; the highest simulated methane deliverable capacity was 196 cm$^3$~STP/cm$^3$.
 Confidence in this conclusion rests upon (i) the accuracy of the intermolecular potentials describing the molecular interactions and (ii) the sufficient sampling of material space, i.e.\ that the structures considered are representative of the set of possible materials.
-To further explore material space and address sensitivity to the potentials used, scaling the Lennard-Jones potential well depths of material atoms to model enhanced interactions \cite{gomez2014exploring}, placing Lennard-Jones spheres in a unit cell randomly to form ``pseudo-materials'' \cite{kaija2018high}, and generating fictitious potential energy fields via a generative adversarial network trained on zeolite structures \cite{lee2019predicting} all generated model substrates that all failed to meet the ARPA-E methane deliverable capacity target.
+To further explore material space and address sensitivity to the potentials used, scaling the Lennard-Jones potential well depths of material atoms to model enhanced interactions~\cite{gomez2014exploring}, placing Lennard-Jones spheres in a unit cell randomly to form ``pseudo-materials''~\cite{kaija2018high}, and generating fictitious potential energy fields via a generative adversarial network trained on zeolite structures~\cite{lee2019predicting} all generated model substrates that all failed to meet the ARPA-E methane deliverable capacity target.
 
@@
 The first limitation of our proof is that we restricted ourselves to \emph{isothermal} pressure swing storage.
-By raising the temperature of the adsorbent during gas discharge to drive off residual gas \cite{gomez2014exploring}, the deliverable capacity could be enhanced, albeit at the cost of a more complicated engineering design of the fuel tank and vehicle.
+By raising the temperature of the adsorbent during gas discharge to drive off residual gas~\cite{gomez2014exploring}, the deliverable capacity could be enhanced, albeit at the cost of a more complicated engineering design of the fuel tank and vehicle.
 %While this approach poses engineering challenges, in principle it could allow an arbitrarily large deliverable capacity.
@@
 A flagship example is MOF Co(bdp)~\cite{choi2008broadly}, which possesses a wine-rack-like topology capable of hinge motion.
-At low methane pressure, Co(bdp) adopts a collapsed, nonporous state, but expands to a porous state and fills with gas at higher pressures \cite{mason2015methane}.
+At low methane pressure, Co(bdp) adopts a collapsed, nonporous state, but expands to a porous state and fills with gas at higher pressures~\cite{mason2015methane}.
 This allows Co(bdp) to fully expel its residual gas at low pressures.

### 2020-05-06 gas-adsorption/gas-ads.tex
made a number of ising edits and a thesis edit.
multiple changes to introduction and added citations to gas-ads abstract.

@@
 \begin{abstract}
-  Due to their low volumetric energy density at ambient conditions, both hydrogen and natural gas are challenging to economically store onboard environmentally sustainable vehicles [].
-One strategy to densify these gases is to pack the fuel tank with a porous material [].
-  Metal-organic frameworks are tunable, nanoporous materials with large internal surface areas and show considerable promise for densifying gases [].
-The US Department of Energy (DOE) has set volumetric deliverable capacity targets [] which, if met, would help to enable commercial adoption of hydrogen/natural gas as transportation fuels.
-Many have attempted to establish theoretical upper bounds on the deliverable capacity via pressure-swing adsorption using simplified models of the gas-substrate and gas-gas interaction []; none have established a rigorous upper bound on the deliverable capacity.
-  Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing.
+Due to their low volumetric energy density at ambient conditions, both hydrogen
+and natural gas are challenging to economically store onboard environmentally
+sustainable vehicles~\cite{mason2014evaluating, sircar2002pressure}. One
+strategy to densify these gases is to pack the fuel tank with a porous
+material~\cite{?}. Metal-organic frameworks (MOFs) are tunable, nanoporous
+materials with large internal surface areas and show considerable promise for
+densifying gases~\cite{makal2012methane,mason2014evaluating,
+suh2011hydrogen,garcia2018benchmark, schoedel2016role}. The US Department of
+Energy (DOE) has set volumetric deliverable capacity targets~\cite{simon2015materials, h2targetsDOE} which, if met, would help to enable 
+commercial adoption of hydrogen/natural gas as
+transportation fuels. Many have attempted to establish theoretical upper bounds
+on the deliverable capacity via pressure-swing adsorption using simplified
+models of the gas-substrate and gas-gas interaction~\cite{gomez2014exploring, 
+gomez2017impact, kaija2018high, lee2019predicting}; none have
+established a rigorous upper bound on the deliverable capacity.
+Here, we present a theoretical upper bound on the deliverable capacity of a
+rigid material via an isothermal pressure-swing. To provide an extremum, we
+consider a substrate that provides a spatially uniform potential energy field
+for the gas. Our bound is unique in that it directly relies on experimentally
+measured properties of the (real) bulk gas, without making approximations.
+We conclude that the goals set by the US DOE for natural gas and hydrogen
+storage are theoretically possible, but sufficiently close to the upper bound
+as to be impractical for any real porous material. However, limitations to the
+scope of applicability of our upper bound guide future material development 
+efforts. Firstly, one could heat the adsorbent to drive off residual gas when
+the pressure diminishes~\cite{gomez2014exploring}. Secondly, the fundamental
+physics of our upper bound do not pertain to any material that changes its
+structure in response to adsorbed gas, suggesting that flexible materials could
+still satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly, mason2015methane}.
+
+%   Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing.
 % "pressure-swing storage doesn't make sense to me.
 storage and delivery  maybe but awk.
-  % I added rigid since that could be important...
-  To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
-  \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
-  Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
-  We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-  % However, this upper bound directly leads to important realizations which should guide future development.
-  However, limitations to the scope of applicability of our upper bound guide future material development efforts.
- % Firstly, one could extract the gas at a higher temperature than that used while filling [].
-  Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
-  Secondly, the fundamental physics of our upper bound do not pertain to
-  % rule out
-  any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
-  % firstly, secondly vs first, second?
+%   % I added rigid since that could be important...
+%   To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
+%   \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
+%   Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
+%   We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
+%   % However, this upper bound directly leads to important realizations which should guide future development.
+%   However, limitations to the scope of applicability of our upper bound guide future material development efforts.
+%  % Firstly, one could extract the gas at a higher temperature than that used while filling [].
+%   Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
+%   Secondly, the fundamental physics of our upper bound do not pertain to
+%   % rule out
+%   any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
+%   % firstly, secondly vs first, second?
 \end{abstract}
@@
 
+\appendix
+\clearpage
 \input{gas-ads-appendix.tex}

### 2020-05-06 thesis/chapter-1.tex
multiple changes to introduction and added citations to gas-ads abstract.

@@
 and natural gas are challenging to economically store onboard environmentally
-sustainable vehicles~\cite{?}. One strategy to densify these gases is to pack
-the fuel tank with a porous material~\cite{?}. Metal-organic frameworks (MOFs)
-are tunable, nanoporous materials with large internal surface areas and show
-considerable promise for densifying gases~\cite{?}. The US Department of Energy
-(DOE) has set volumetric deliverable capacity targets~\cite{?} which, if met,
-would help to enable commercial adoption of hydrogen/natural gas as
+sustainable vehicles~\cite{mason2014evaluating, sircar2002pressure}. One
+strategy to densify these gases is to pack the fuel tank with a porous
+material~\cite{?}. Metal-organic frameworks (MOFs) are tunable, nanoporous
+materials with large internal surface areas and show considerable promise for
+densifying gases~\cite{makal2012methane,mason2014evaluating,
+suh2011hydrogen,garcia2018benchmark, schoedel2016role}. The US Department of
+Energy (DOE) has set volumetric deliverable capacity targets~\cite{simon2015materials, h2targetsDOE} which, if met, would help to enable 
+commercial adoption of hydrogen/natural gas as
 transportation fuels. Many have attempted to establish theoretical upper bounds
 on the deliverable capacity via pressure-swing adsorption using simplified
-models of the gas-substrate and gas-gas interaction~\cite{?}; none have
+models of the gas-substrate and gas-gas interaction~\cite{gomez2014exploring, 
+gomez2017impact, kaija2018high, lee2019predicting}; none have
 established a rigorous upper bound on the deliverable capacity.
 
-% This is part of the abstract(s) for APS Conferences.
 Here, we present a theoretical upper bound on the deliverable capacity of a
@@
 as to be impractical for any real porous material. However, limitations to the
-scope of applicability of our upper bound guide future material development efforts. Firstly, one could heat the adsorbent to drive off residual gas when
-the pressure diminishes~\cite{?}. Secondly, the fundamental physics of our
-upper bound do not pertain to any material that changes its structure in
-response to adsorbed gas, suggesting that flexible materials could still
-satisfy the DOE targets~\cite{?}.
+scope of applicability of our upper bound guide future material development 
+efforts. Firstly, one could heat the adsorbent to drive off residual gas when
+the pressure diminishes~\cite{gomez2014exploring}. Secondly, the fundamental
+physics of our upper bound do not pertain to any material that changes its
+structure in response to adsorbed gas, suggesting that flexible materials could
+still satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly, mason2015methane}.
 
 % Now add why a multidimensional 2D SAD would help and how we are working toward it.
+Due to the difficulty in designing the ``ideal'' MOF for gas storage
+applications, computational methods are necessary that can predict the
+associated thermodynamic properties of interest. While SAD is an invaluable tool
+for calculating a single isotherm (at a given density), an even more powerful
+method is necessary for determining the best MOF for gas storage applications.
+A multidimensional 2D SAD would allow researchers to compute all properties of
+interest for all densities and temperatures. Our preliminary research has
+yielded a robust method that has been tested on multiple systems and can be
+extended into a 2D formulation.
 
@@
 % Now tie this in with gas adsorption.
+We are currently working on two extensions which will enable a multidimensional
+SAD. The first is adaptive interpolation histogram binning. This technique will
+decrease the simulation time by orders of magnitude. The second is a version of
+SAD that simulates ``number'' space rather than energy space. With the success
+of these two extensions, we can begin the development of 2D SAD and thus provide
+a powerful way for researchers to work toward enabling environmentally
+sustainable light vehicle storage.

### 2020-05-07 ising/ising.tex
made a number of ising edits and a thesis edit.
went through the abstract and made edits.

@@
 
-\title{Flat-histogram method comparison on 2D Ising model
+\title{Flat-histogram method comparison on the 2D Ising model
 }
@@
 \begin{abstract}
-We examine the convergence of the stochastic approximation with a dynamic
-update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
-stochastic approximation Monte Carlo (SAMC) and Wang-Landau (WL) methods show that SAD performs robustly and without user input knowledge of an energy range.
-We also compare SAD with ``production'' run WL with fixed weights which is the first of its kind recorded.
-Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
+We examine the convergence of stochastic approximation with a dynamic
+update factor (SAD) with other flat-histogram methods applied to the 2D
+Ising model. The comparison with stochastic approximation Monte Carlo (SAMC)
+and Wang-Landau (WL) methods shows that SAD performs robustly and without user
+input knowledge of an energy range. We also compare SAD with WL followed by a
+``production run'' with fixed weights which is the first comparison of its kind
+recorded. We find that WL followed by a production run is preferable to pure WL
+for preserving ergodicity and detailed balance. Since SAD is more powerful in
+the common case in which the range of energies is not known in advance, the 2D
+Ising model presents a unique challenge since the energy range is readily known
+and easily discovered by WL methods.
 \end{abstract}

### 2020-05-07 ising/ising-body.tex
updated the data and corrected the paper.

@@
 samples a narrow energy range, it proves computationally inefficient at sampling large energy ranges.
-
 Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram
@@
 The SAD (stochastic approximation with a
-dynamic $\gamma$) method outlined by Pommerenck
+dynamic $\gamma$) method as outlined by Pommerenck
 et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
-that dynamically chooses the modification factor rather than relying on system
-dependent parameters. SAD shares the same convergence properties with SAMC while
-replacing un-physical user-defined parameters with the algorithms dynamic
-choice.
+that dynamically chooses the modification factor rather than relying on
+un-physical user-defined parameters. SAD shares the same convergence properties
+with SAMC while replacing un-physical user-defined parameters with the
+algorithms dynamic choice.
 
-In this work, we apply the family of weight-based flat-histogram Monte
-Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
+In this work, we compare the convergence properties of five flat-histogram
+methods. We detail how each method is implemented and apply the family of
+weight-based flat-histogram Monte Carlo methods (pure WL, WL followed by a
+"production run", 1/t-WL, SAMC, and SAD) to the 2D Ising model.
 
@@
 \section{Flat-histogram methods}\label{sec:histogram}
-Flat-histogram methods determine the density of states $D(E)$ over a broad range
+Flat-histogram methods compute the density of states $D(E)$ over a broad range
 of energies by simulating each energy with equivalent accuracy. Flat-histogram
@@
 of energy space. Flat-histogram methods differ primarily in how they schedule
-the decrease of $\gamma_t$.  Figure~\ref{fig:N32-gamma} shows ...
+the decrease of $\gamma_t$.  Figure~\ref{fig:N32-gamma} shows several
+flat-histogram methods each decreasing $\gamma_t$ as a function of moves.
+Methods that decrease $\gamma_t$ too rapidly can fail to converge while methods
+that decrease too slowly can take infinitely long to converge to the correct
+$D(E)$.
 

### 2020-05-08 ising/ising-body.tex
made some changes in conclusion.

@@
 \section{Conclusion}
-We find that both SAD, 1/t-WL, and WL followed by a production run
+We find that SAD, 1/t-WL, and WL followed by a production run
 (with an adequately small $\gamma^{\min}$) demonstrate excellent and robust
@@
 not suffer from
-error saturation. We find that for larger systems SAD reduces the updated factor
-more slowly (and conservatively) than 1/t-WL and WL followed by a
+error saturation. We find that for larger Ising systems SAD reduces the update
+factor more slowly (and conservatively) than 1/t-WL and WL followed by a
 production run. This means that SAD will take
 proportionately more moves to converge to the same value as 1/t-WL as system
-size is increased. SAD requires the user to specify a temperature range of
-interest rather than an energy range of interest as the WL methods do. The 2D
+size is increased (WL methods are given the energy range \emph{a priori.} and
+converge independent of system size). SAD requires the user to specify a
+temperature range of interest rather than an energy range of interest as the WL
+methods do. The 2D
 Ising model is a system where the energy range is known \emph{a priori.}, which
@@
 \section{Acknowledgments}
-
 We wish to thank Johannes Zierenberg for helpful discussions regarding WL followed by a production run and for insights into properties of a production run at different stages of convergence.

### 2020-05-08 ising/ising-body.tex
tweak to parameters for SAD

--- 2020-05-07 ising/ising-body.tex
@@
 requiring \emph{a priori.} an energy range of interest; however, this is almost always easier to
-identify and is more physical than the SAMC parameter $t_0$. The update factor
-can be written in terms of the current estimates for the highest $E_H$ and
-lowest $E_L$ interesting energies and the last time that an energy in the range
-of interest is encountered $t_L$.
+identify and is more physical than the SAMC parameter $t_0$. The
+update factor is given by:
 \begin{align}
@@
 \end{align}
+where $E_H$ and $E_L$ are the current estimates for the highest and
+lowest interesting energies and $t_L$ is the last time at which an
+energy in the range of interest is encountered.
+%
 SAD only explores the energy range of interest as specified by the minimum

### 2020-05-08 ising/ising-body.tex
fix some 1/t and edit abstract and stuff

@@
 These studies led to the
-formation of the $1/t$-WL algorithm and also sparked some debate as to whether
-the entire simulation should be split into multiple parts, called ``production run" WL, in order to preserve ergodicity and
+formation of the $1/t$-WL algorithm and have also led some researchers to follow
+WL with a ``production run'' with fixed weights, in order to preserve ergodicity and
 detailed balance~\cite{jayasri2005wang, mukhopadhyay2008monte}.
+%
+To our knowledge, this work is the first test of the convergence
+properties of WL followed by a production run with comparison to other
+methods.
 
@@
 weight-based flat-histogram Monte Carlo methods (pure WL, WL followed by a
-"production run", 1/t-WL, SAMC, and SAD) to the 2D Ising model.
+``production run'', $1/t$-WL, SAMC, and SAD) to the 2D Ising model.
 
@@
 \section{Conclusion}
-We find that SAD, 1/t-WL, and WL followed by a production run
-(with an adequately small $\gamma^{\min}$) demonstrate excellent and robust
-convergence. They all converge more rapidly than SAMC, and unlike pure WL do
-not suffer from
-error saturation. We find that for larger Ising systems SAD reduces the update
-factor more slowly (and conservatively) than 1/t-WL and WL followed by a
-production run. This means that SAD will take
-proportionately more moves to converge to the same value as 1/t-WL as system
-size is increased (WL methods are given the energy range \emph{a priori.} and
-converge independent of system size). SAD requires the user to specify a
-temperature range of interest rather than an energy range of interest as the WL
-methods do. The 2D
-Ising model is a system where the energy range is known \emph{a priori.}, which
-makes a comparison between WL methods and SAD (which requires a minimum
-temperature) more biased towards WL methods. For the majority of cases in which
-the energy range of the system is not known and where a range of desired
-temperatures is known, this will make the SAD method
-considerably more convenient. This work also visually demonstrates
-that WL followed by a production run performs extremely well and is preferable
-to pure WL for ensuring both ergodicity and detailed balance.
+We find that SAD, $1/t$-WL, and WL followed by a production run (with an
+adequately small $\gamma^{\min}$) demonstrate excellent and robust
+convergence. They all converge more rapidly than SAMC, and unlike pure
+WL do not suffer from error saturation. We find that for larger Ising
+systems SAD reduces the update factor more slowly (and conservatively)
+than $1/t$-WL and WL followed by a production run. This means that SAD
+will take proportionately more moves to converge to the same value as
+$1/t$-WL as system size is increased.  However, the WL methods are
+solving a simpler problem because they are given the energy range
+\emph{a priori}, rather than a temperature range of interest such as
+SAD requires.  The 2D Ising model is a system where the energy range
+is known \emph{a priori.}, which makes a comparison between WL methods
+and SAD more biased in favor of WL methods. For the general case in
+which the energy range of the system is not known and where a range of
+desired temperatures is known, this will make the SAD method
+considerably more convenient, and quite possibly more efficient than a
+process involving multiple simulations to determine the energy range
+of interest. This work also dramatically demonstrates that WL followed
+by a production run performs extremely well and is preferable to pure
+WL for ensuring both ergodicity and detailed balance.
 

### 2020-05-08 ising/ising.tex
fix some 1/t and edit abstract and stuff

--- 2020-05-02 ising/ising.tex
@@
 \begin{abstract}
-We examine the convergence of the stochastic approximation with a dynamic
-update factor (SAD) algorithm applied to the 2D Ising model. Comparison with
-stochastic approximation Monte Carlo (SAMC) and Wang-Landau (WL) methods show that SAD performs robustly and without user input knowledge of an energy range.
-We also compare SAD with ``production'' run WL with fixed weights which is the first of its kind recorded.
-Since SAD is more powerful in the common case in which the range of energies is not known in advance, the 2D Ising model presents a unique challenge since the energy range is readily known and easily discovered by WL methods.
+  We compare the convergence of several flat-histogram methods applied
+  to the 2D Ising model, including the recently introduced stochastic
+  approximation with a dynamic update factor (SAD) method.  We compare
+  this method with the Wang-Landau (WL) method, the $1/t$ variant of
+  the WL method, and standard stochastic approximation Monte Carlo
+  (SAMC).  In addition, we consider a procedure WL followed by a
+  ``production run'' with fixed weights that refines the estimation of
+  the entropy.  To our knowledge, this work is the first to test this
+  approach against other methods.
+  %% To our knowledge, the
+  %% convergence rate of this production-run WL approach has never been 
+%% We examine the convergence of stochastic approximation with a dynamic
+%% update factor (SAD) with other flat-histogram methods applied to the 2D
+%% Ising model. The comparison with stochastic approximation Monte Carlo (SAMC)
+%% and Wang-Landau (WL) methods shows that SAD performs robustly and without user
+%% input knowledge of an energy range. We also compare SAD with WL followed by a
+%% ``production run'' with fixed weights which is the first comparison of its kind
+%% recorded.
+We find that WL followed by a production run \emph{does} converge to
+the true density of states, in contrast to pure WL.  Three of the
+methods converge robustly: SAD, $1/t$-WL, and WL followed by a
+production run.  Of these, SAD does not require \emph{a priori}
+knowledge of the energy range.
 \end{abstract}

### 2020-05-08 ising/ising-body.tex
more verbiage on the production run results

@@
 converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$
-moves. The two WL simulations followed by a production run highlight that
-choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
-properties.
+moves.  The WL simulation followed by a production run when
+$\gamma^{\text{WL}}_t$ reached $10^{-4}$ sometimes dramatically failed to
+converge, and with a \davidsays{few} random seeds converged very
+nicely.  This highlights a risk taken when setting $\gamma^{\min}$.
+If the weights are insufficiently converged, the production run will
+fail to explore all energies. \davidsays{Can you identify whether it got
+stuck at high or low energy?}  With the smaller value of
+$\gamma^{\min}=10^{-6}$ consistently and efficiently converged.
+
+%% The two WL simulations followed by a production run highlight that
+%% choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
+%% properties.
 

### 2020-05-08 ising/ising-body.tex
made changes; had to add body to main in order for aps to show pdf?.

--- 2020-05-08 ising/ising-body.tex
@@
 converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$
-moves. The two WL simulations followed by a production run highlight that
-choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
-properties.
+moves.  The WL simulation followed by a production run when
+$\gamma^{\text{WL}}_t$ reached $10^{-4}$ sometimes dramatically failed to
+converge, although five of the eight random seeds converged very
+nicely.  This highlights a risk taken when setting $\gamma^{\min}$.
+If the weights are insufficiently converged, the production run will
+fail to explore all energies, in this case, three of the simulations became ``stuck'' at low energies. With the smaller value of
+$\gamma^{\min}=10^{-6}$, the method consistently and efficiently converged.
+
+%% The two WL simulations followed by a production run highlight that
+%% choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
+%% properties.
 

### 2020-05-08 ising/ising.tex
made changes; had to add body to main in order for aps to show pdf?.

@@
 
-% define colors for comments
-\definecolor{dark-gray}{gray}{0.10}
-\definecolor{light-gray}{gray}{0.70}
-
 \newcommand{\red}[1]{{\bf \color{red} #1}}
@@
 \newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
+
 \begin{document}
-
 \title{Flat-histogram method comparison on the 2D Ising model
@@
 the entropy.  To our knowledge, this work is the first to test this
-  approach against other methods.
-  %% To our knowledge, the
-  %% convergence rate of this production-run WL approach has never been 
-%% We examine the convergence of stochastic approximation with a dynamic
-%% update factor (SAD) with other flat-histogram methods applied to the 2D
-%% Ising model. The comparison with stochastic approximation Monte Carlo (SAMC)
-%% and Wang-Landau (WL) methods shows that SAD performs robustly and without user
-%% input knowledge of an energy range. We also compare SAD with WL followed by a
-%% ``production run'' with fixed weights which is the first comparison of its kind
-%% recorded.
-We find that WL followed by a production run \emph{does} converge to
+approach against other methods. We find that WL followed by a production run \emph{does} converge to
 the true density of states, in contrast to pure WL.  Three of the
@@
 
-\input{ising-body.tex}
+\section{Introduction}
+Flat-histogram Monte Carlo simulation algorithms calculate the thermodynamic
+properties of various systems over a range of temperatures.  The first histogram
+method used a single canonical Monte Carlo simulation to predict properties for
+nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
+samples a narrow energy range, it proves computationally inefficient at sampling large energy ranges.
+Multicanonical methods, introduced by Berg and Neuhaus, enabled flat-histogram
+sampling which improved the exploration of configurational space and allowed the
+simulation to overcome free-energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
+These works led to increase in the development of a variety of ``flat'' (or ``broad'') histogram
+methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
+wang2001determining, wang2001efficient} which could explore a wider range
+of energies.  In addition to obtaining thermodynamic information for the entire
+energy range for a single simulation, these approaches cannot be easily trapped
+in a local energy minimum like a canonical simulation.
+
+Wang and Landau introduced one of the most widely used flat-histogram
+Monte Carlo algorithms that determined the density of states (DOS) for
+a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
+its power, the method unfortunately requires \emph{a priori.} knowledge of several
+user-defined parameters. Thus, for any given system under study, the user needs
+to determine the ideal parameters in order to apply the method. The Wang-Landau
+algorithm is also known to violate detailed balance (although only for brief
+time intervals)~\cite{yan2003fast, shell2002generalization}. With the violation
+of detailed balance, convergence of the algorithm is not guaranteed.
+
+Because of the uncertainty of convergence for WL, many studies have been undertaken
+to understand how the modification (or update) factor $\gamma$ impacts the
+convergence~\cite{zhou2005understanding,lee2006convergence,
+belardinelli2007wang}.
+Belardinelli and Pereyra showed that an update factor that decreases faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang, belardinelli2008analysis, zhou2008optimal}, where $t$ corresponds to the number of moves.
+Schneider \emph{et al.} outline minor refinements algorithm including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.
+These studies led to the
+formation of the $1/t$-WL algorithm and have also led some researchers to follow
+WL with a ``production run'' with fixed weights, in order to preserve ergodicity and
+detailed balance~\cite{jayasri2005wang, mukhopadhyay2008monte}.
+%
+To our knowledge, this work is the first test of the convergence
+properties of WL followed by a production run with comparison to other
+methods.
+
+Liang independently considered whether WL could be treated as a special case of
+stochastic approximation whose convergence could be mathematically
+proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang \emph{et
+al.}~\cite{liang2007stochastic} argued that WL can be considered a form of
+stochastic approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
+convergence (if certain conditions are met). Despite the added benefit of
+guaranteed convergence, the method still has a system specific user-defined
+variable. Such variables often create difficulty when applying Monte Carlo
+methods across arbitrary systems.
+
+Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
+related Statistical Temperature Molecular Dynamics (STMD), an adaption
+of the WL method that approximates the entropy as a piecewise linear function,
+which improves convergence for systems with a continuously varying
+energy~\cite{kim2006statistical, kim2007statistical}. STMC applied to WL
+requires a temperature range be specified rather than an energy range.  Kim
+\emph{et al.} extended this work as Replica Exchange Statistical Temperature
+Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping STMC
+simulations to improve convergence~\cite{kim2009replica}. Recently,
+Junghans \emph{et al.} demonstrated a close connection between
+metadynamics, which was introduced by Laio and
+Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods, with STMD
+forging the connection~\cite{junghans2014molecular}.
+
+The SAD (stochastic approximation with a
+dynamic $\gamma$) method as outlined by Pommerenck
+et.al~\cite{pommerenck2020stochastic} is a special version of the SAMC algorithm
+that dynamically chooses the modification factor rather than relying on
+un-physical user-defined parameters. SAD shares the same convergence properties
+with SAMC while replacing un-physical user-defined parameters with the
+algorithms dynamic choice.
+
+In this work, we compare the convergence properties of five flat-histogram
+methods. We detail how each method is implemented and apply the family of
+weight-based flat-histogram Monte Carlo methods (pure WL, WL followed by a
+``production run'', $1/t$-WL, SAMC, and SAD) to the 2D Ising model.
+
+\section{Ising Model}
+The 2D Ising spin-lattice system is widely used as a testbed when
+benchmarking or comparing Monte Carlo
+methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing, barash2019estimating}. The 2nd order
+phase transition behavior and the ability to directly calculate the
+exact solution for finite lattices~\cite{beale1996exact, haggkvist2004computation} make the
+system sufficiently interesting for such theoretical comparisons. It is
+also important to note that direct comparison of the other methods can
+be made with WL as its original implementation was done on this
+system~\cite{wang2001determining,wang2001efficient}.
+We test the convergence of several flat-histogram methods
+on the periodic 2D square lattice ferromagnetic Ising model with identical
+nearest neighbor interactions~\cite{landau2004new} ($J_{ij} = J$).
+\begin{align}
+\mathcal{H} = -J \sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
+\end{align}
+The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
+for up or down spins respectively. In the absence of a magnetic field ($h =
+0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
+kaufman1949crystal}:
+\begin{align}
+\mathcal{H} = - J\sum_{\braket{i,j}} \sigma_i \sigma_j
+\end{align}
+where the sum is over nearest neighbor spin sites. Beale showed that for finite
+lattices the Density of States could directly be calculated from the partition
+function~\cite{beale1996exact}
+\begin{align}
+Z = \sum_E g(E) e^{-{\beta E}}
+\end{align}
+% where $g(E)$ is the multiplicity of the system which is proportional to
+% $D(E)$. We can estimate the average deviation from the exact solution
+% using the relation~\cite{schneider2017convergence, shakirov2018convergence,
+% barash2017control}:
+% \begin{align}
+% \braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
+% \end{align}
+where $g(E)$ is the multiplicity of the system which is proportional to
+$D(E)$. We can compute the maximum deviation in the canonical specific heat capacity $c_V$ from the exact solution~\cite{schneider2017convergence, shakirov2018convergence,
+barash2017control,barash2017gpu}:
+\begin{align}
+\epsilon \equiv \max_T \left|\left(c_V(T)_{\text{MC}} - c_V(T)_{\text{Beale}}\right)\right|
+\end{align}
+Computing the specific heat capacity $c_V$ presents a difficult challenge for
+any Monte Carlo method due to fluctuations in the derivative of the
+internal energy around the phase transition. The critical temperature $T_c = 2.
+269T$ or $\beta= 0.441J$ comes directly from the Kramers-Wannier
+duality~\cite{bhattacharjee1995fifty} and marks the transition from a
+disordered to ordered magnetic state. Methods that accurately compute $c_V$
+also by extension accurately compute the internal energy.
+
+\section{Flat-histogram methods}\label{sec:histogram}
+Flat-histogram methods compute the density of states $D(E)$ over a broad range
+of energies by simulating each energy with equivalent accuracy. Flat-histogram
+Monte Carlo methods propose randomly chosen ``moves'' which change the state of
+the system and must satisfy detailed balance.  Each algorithm differs in how it
+determines the probability of accepting a move and in what additional statistics
+must be collected in order decide on that probability.
+
+We describe several closely related flat-histogram methods which each rely on a
+weight function $w(E)$ to determine $D(E)$.  For these algorithms, the
+probability of accepting a move is given by
+\begin{equation}
+	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
+	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
+\end{equation}
+which biases the simulation in favor of energies with low weights. The result of
+weights $w(E)$ that are proportional to $D(E)$ is an entirely flat-histogram. We
+can relate the entropy to the weights in the microcanonical ensemble, since the
+entropy is defined as $S(E) \equiv k_B\ln(D(E)) \approx \ln w(E)$.
+
+Flat-histogram methods employ a random walk in energy space to estimate $D(E)$.  Each method operates by continuously updating the weights at each
+step of the simulation
+\begin{equation}
+	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
+	+\gamma_t
+\end{equation}
+where $t$ is number of the current move, $\gamma(t)$ is a move-dependent update
+factor, and $E$ is the current energy.  This update causes the random walk to
+avoid frequent sampling of the same energies, leading to a rapid exploration
+of energy space. Flat-histogram methods differ primarily in how they schedule
+the decrease of $\gamma_t$.  Figure~\ref{fig:N32-gamma} shows several
+flat-histogram methods each decreasing $\gamma_t$ as a function of moves.
+Methods that decrease $\gamma_t$ too rapidly can fail to converge while methods
+that decrease too slowly can take infinitely long to converge to the correct
+$D(E)$.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{gamma-n32.pdf}
+  \caption{
+  The update factor $\gamma_t$ versus the iteration number for the $N=32 \times 32$
+  system.}
+  \label{fig:N32-gamma}
+\end{figure}
+
+The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
+landau2014guide} explores energy space by setting
+$\gamma_{t=0}^{\text{WL}}=1$, and then decreases $\gamma^{\text{WL}}$ in prescribed stages. An energy range of
+interest must be specified~\cite{wang2001efficient, schulz2003avoiding,yan2003fast}, which often requires multiple simulations if unknown.
+%
+The number (``counts'') of moves ending at each energy are stored in a
+histogram.
+For a sufficiently flat energy histogram (typically user-specified to be 0.8), $\gamma^{\text{WL}}$ is decreased by a specified factor of $\frac12$ and the histogram is reset to zero.
+The entire process is repeated until $\gamma^{\text{WL}}$ reaches a desired cutoff.
+
+The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
+from dropping below $N_S/t$~\cite{belardinelli2008analysis,
+schneider2017convergence}. The method follows the standard WL algorithm with two
+modifications.  Firstly, when each energy state has been visited once, the histogram is considered flat and $\gamma_t$ is
+decreased by a factor of two. Secondly, when
+$\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the update factor becomes
+$\gamma_t = N_S/t$ for the remainder of the simulation:
+\begin{align}
+  \gamma_t^{1/t\text{-WL}} = \begin{cases}
+     \gamma^{\text{WL}}_t & \gamma^{\text{WL}}_t > \frac{N_S}{t} \\
+     \frac{N_S}{t} & t \ge t_0
+ \end{cases}
+\end{align}
+where $t$ is the number of moves, $\gamma^{\text{WL}}_t$ is the Wang-Landau update factor
+at move $t$, and $N_S$ is the number of energy bins.
+
+The WL method can be terminated after $\gamma$ reaches a specified
+minimum $\gamma^{\min}$ followed by a \emph{production run} in which the
+weights are held fixed~\cite{gross2018massively}. WL is used to
+generate the weights resulting in a
+flat-histogram~\cite{janke2017generalized}.  The entropy is then
+computed (up to a constant) by adding the logarithm of the production
+histogram to the logarithm of the weights.  The production run thus
+satisfies detailed balance, and will ideally be ergodic; however, the
+convergence of the simulation is still impacted by the choice of the
+minimum $\gamma^{\min}$.
+
+Another weight-based flat-histogram method is
+the stochastic approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
+schedule by which the update factor $\gamma^{\text{SA}}_t$ is continuously
+decreased~\cite{liang2007stochastic, werlich2015stochastic,
+schneider2017convergence}.  The update factor is defined in the
+original implementation~\cite{liang2007stochastic} in terms of an
+arbitrary tunable parameter $t_0$,
+\begin{align}
+\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
+\end{align}
+where as above $t$ is the number of moves that have been attempted.
+
+The implementation of SAMC is extremely simplistic.
+In addition, Liang has proven that the weights converge to the true
+density of states~\cite{liang2006theory, liang2007stochastic,
+liang2009improving} provided the update factor satisfies
+\begin{align}
+\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
+\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
+\end{align}
+where $\zeta > 1$.  Unlike WL methods, the energy range need not be
+known \emph{a priori.} and the convergence time depends only on the choice of
+parameter $t_0$.
+Unfortunately, $t_0$ can be difficult to chose in advance
+for arbitrary systems.
+Liang \emph{et al.} give a rule of thumb in
+which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
+is the number of energy bins~\cite{liang2007stochastic}.  Schneider
+\emph{et al.} found and we confirm that for the Ising model this heuristic is helpful for small spin systems, but that larger systems require an even higher
+$t_0$ value~\cite{schneider2017convergence}.
+
+Pommerenck \emph{et al.} propose a refinement~\cite{pommerenck2020stochastic} to SAMC
+where the update factor is determined dynamically rather than by the user.
+Stochastic approximation with a dynamic $\gamma$ (SAD) requires the user to
+provide a minimum temperature of interest $T_{\min}$. This is analogous to WL
+requiring \emph{a priori.} an energy range of interest; however, this is almost always easier to
+identify and is more physical than the SAMC parameter $t_0$. The
+update factor is given by:
+\begin{align}
+  \gamma_{t}^{\text{SAD}} =
+     \frac{
+       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
+     }{
+       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
+     }
+\end{align}
+where $E_H$ and $E_L$ are the current estimates for the highest and
+lowest interesting energies and $t_L$ is the last time at which an
+energy in the range of interest is encountered.
+%
+SAD only explores the energy range of interest as specified by the minimum
+temperature of interest $T_{\min} < T < \infty$. During the simulation the two
+energies $E_H$ and $E_L$, are refined such that the range of energies are
+conservatively estimated. The weights are calculated for each energy region according to the original prescription.
+\begin{enumerate}
+\item {$E < E_L$:} $w(E>E_H) = w(E_H)$
+\item {$E_L < E < E_H$:} moves are handled the same as other weight-based
+methods that are mentioned
+\item {$E > E_H$:} $w(E<E_L) = w(E_L)e^{-\frac{E_L-E}{T_{\min}}}$
+\end{enumerate}
+Each time the simulation changes the value of $E_H$ or $E_L$, the weights
+within the new portion of the interesting energy range are updated.
+
+\section{Results}
+We test the algorithms on two different system sizes of the 2D Ising model.
+The first is a smaller simulation with a lattice size of $N = 32 \times 32$ and
+the second has a lattice size of $N = 128 \times 128$. The SAD method explores 
+the energy space of each system using a minimum reduced temperature of $T_{\text{min}} = 1$. All simulations calculate the minimum important energy $E_{\min}$
+and maximum entropy energy $E_{\max}$ (with the exception of the WL methods
+where both of these parameters are needed \emph{a priori.}).
+
+% \begin{figure}
+% \includegraphics[width=\columnwidth]{N32-entropy-error-default.pdf}
+%   \caption{
+%   The average entropy error for each MC method for $N=32 \times 32$,
+%                $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+%                as a function of number of iterations run.  The error is
+%                averaged over 8 independent simulations, and the best
+%                and worst simulations for each method are shown as a
+%                semi-transparent shaded area.}\label{fig:n32}
+% \end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{N32-Cv.pdf}
+  \caption{
+    The specific heat capacity versus the reciprocal temperature $\beta$ for the $N=32 \times 32$ system for each histogram method at $10^{9}$ moves.}
+  \label{fig:N32-cv}
+\end{figure}
+
+\subsection{The 32 $\times$ 32 Ising model}
+Figure~\ref{fig:N32-gamma} shows the update factor $\gamma_t$ for each
+of the flat-histogram methods. All of the update factors initially
+start at $\gamma_{t=0} = 1$. SAD dynamically updates $\gamma_t$
+throughout the simulation. After about $10^{10}$ moves,
+$\gamma^{\text{SAD}}_t$ proceeds as $1/t$. We show
+$\gamma^{\text{WL}}_t$ for WL both with and without a production
+run. The WL production run begins after $\gamma^{\text{WL}}_t$ has reached
+$10^{-4}$. The update factor for $1/t$-WL decreases similarly to WL
+before finding all the energy states and switching to $1/t$. All of
+the SAMC update factors equal 1 until the number of moves is equal to
+$t_0$ at which point they decrease as $1/t$.
+
+All of the methods, except for SAMC, use the same single random number seed and
+give a reasonable approximation for the heat capacity peak resulting from the
+phase transition after only 10$^9$ moves. Figure~\ref{fig:N32-cv} shows the
+specific heat capacity vs. the reciprocal temperature $\beta$ at $10^{9}$ moves.
+A temperature range of $1.5k_B$ to $5k_B$ is chosen to highlight
+the phase transition at the critical temperature $T_c$.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{N32-Cv-error-default.pdf}
+  \caption{The maximum error in the specific heat capacity for each method for the $N=32 \times 32$ and $T_{\min} = 1$ as a function of number of iterations run.
+The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
+  \label{fig:N32-cv-error}
+\end{figure}
+
+Figure~\ref{fig:N32-cv-error} shows the maximum error in the heat capacity as a
+function of time for this system. The solid/dashed lines represent the average
+of the maximum value of the error in the specific heat capacity $c_V$ averaged
+over eight simulations using different random number seeds. The range of maximum
+errors for each simulation is shown as a shaded region. By
+the time $10^8$ moves have been made all but the WL simulation have begun to
+converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^9$ moves.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{gamma-n128.pdf}
+  \caption{
+  The update factor $\gamma_t$ versus the iteration number for the $N=128 \times 128$
+  system.}
+  \label{fig:N128-gamma}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{N128-Cv-error-default.pdf}
+  \caption{The maximum error in the specific heat capacity for each method for the $N=128 \times 128$ and $T_{\min} = 1$ as a function of number of iterations run.
+The maximum error is averaged over 8 independent simulations, and the best and worst simulations for each method are shown as a semi-transparent shaded area.}
+  \label{fig:N128-cv-error}
+\end{figure}
+
+% \begin{figure}
+% \includegraphics[width=\columnwidth]{N128-entropy-error-default.pdf}
+%   \caption{
+%   The average entropy error for each MC method for $N=128 \times 128$,
+%                $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+%                as a function of number of iterations run.  The error is
+%                averaged over 8 independent simulations, and the best
+%                and worst simulations for each method are shown as a
+%                semi-transparent shaded ar.}\label{fig:n128}
+% \end{figure}
+
+\subsection{128 $\times$ 128 Ising system}
+For the larger 2D Ising system, the update factors initially
+start at the same $\gamma_{t=0} = 1$; however, all of the methods take longer
+to proceed as $1/t$ (with SAD taking $10^{13}$ moves). Fig.~\ref{fig:N128-gamma} shows the update factor $\gamma_t$ for each of the 
+flat-histogram methods. 
+We implement two WL simulations followed by a production run each beginning
+after $\gamma^{\text{WL}}_t$ has reached $10^{-4}$ and $10^{-6}$ respectively.
+Fig.~\ref{fig:N128-cv-error} shows the maximum error in the heat capacity as a 
+function of time for this system. The solid/dashed lines represent the maximum
+value of the error in the specific heat capacity $c_V$ averaged
+over eight simulations using different random number seeds. The range of maximum
+errors for each simulation is shown as a shaded region. By
+the time $10^{13}$ moves have been made all but the WL simulation have begun to
+converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^{12}$
+moves.  The WL simulation followed by a production run when
+$\gamma^{\text{WL}}_t$ reached $10^{-4}$ sometimes dramatically failed to
+converge, although five of the eight random seeds converged very
+nicely.  This highlights a risk taken when setting $\gamma^{\min}$.
+If the weights are insufficiently converged, the production run will
+fail to explore all energies, in this case, three of the simulations became ``stuck'' at low energies. With the smaller value of
+$\gamma^{\min}=10^{-6}$, the method consistently and efficiently converged.
+
+%% The two WL simulations followed by a production run highlight that
+%% choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
+%% properties.
+
+\section{Conclusion}
+We find that SAD, $1/t$-WL, and WL followed by a production run (with an
+adequately small $\gamma^{\min}$) demonstrate excellent and robust
+convergence. They all converge more rapidly than SAMC, and unlike pure
+WL do not suffer from error saturation. We find that for larger Ising
+systems SAD reduces the update factor more slowly (and conservatively)
+than $1/t$-WL and WL followed by a production run. This means that SAD
+will take proportionately more moves to converge to the same value as
+$1/t$-WL as system size is increased.  However, the WL methods are
+solving a simpler problem because they are given the energy range
+\emph{a priori}, rather than a temperature range of interest such as
+SAD requires.  The 2D Ising model is a system where the energy range
+is known \emph{a priori.}, which makes a comparison between WL methods
+and SAD more biased in favor of WL methods. For the general case in
+which the energy range of the system is not known and where a range of
+desired temperatures is known, this will make the SAD method
+considerably more convenient, and quite possibly more efficient than a
+process involving multiple simulations to determine the energy range
+of interest. This work also dramatically demonstrates that WL followed
+by a production run performs extremely well and is preferable to pure
+WL for ensuring both ergodicity and detailed balance.
+
+\section{Acknowledgments}
+We wish to thank Johannes Zierenberg for helpful discussions regarding WL followed by a production run and for insights into properties of a production run at different stages of convergence.
 

### 2020-05-11 gas-adsorption/gas-ads-appendix.tex
made a number of ising edits and a thesis edit.
compute Vopt and add to text. Fix reference to SI in text.

@@
-\appendix
-\clearpage
 \setcounter{table}{0}
@@
 
-\section{Proof of extremum}
+\section{Proof of extremum}\label{sec:proof-extremum}
 To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.

### 2020-05-11 gas-adsorption/gas-ads-body.tex
compute Vopt and add to text. Fix reference to SI in text.

@@
 %Our theoretical framework thus recovers an upper bound on the deliverable capacity of methane (298\ K, $\pfull=$65\ bar, $\pempty=$5.8\ bar) in porous materials as 374\ L STP/L.
-The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, \red{achieved for $\V_{opt} = ...$}.
+The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, achieved for $\V_{opt} =$ 5.9\ kJ/mol.
 %
@@
 %yields the maximum deliverable capacity over all possible potential energy fields. 
-In the SI (see Section~\ref{sec:fixme} \davidsays{Look up section label}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest.
+In the SI (see Section~\ref{sec:proof-extremum}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest.
 The potential energy field is \emph{static} if it is unaffected by the presence of the gas; consequently, our model does not apply to flexible materials that undergo gas-induced conformation changes \cite{schneemann2014flexible}.

### 2020-05-11 gas-adsorption/gas-ads-body.tex
remove out of date comments and answer comments.

@@
 Moreover, high-pressure storage tanks are heavy, thick-walled, and non-conformable, while cryogenic storage tanks are bulky, expensive, and afflicted by boil-off losses~\cite{hasan2009minimizing}.
-% \jordan{
-% Both forms of densification suffer several drawbacks varying from safety issues,  excessively expensive, boil-off~\cite{hasan2009minimizing}, and extensive refilling infrastructure~\cite{bossel2003energy}.
-% }
-% % Jordan: I commented this out! % Cory: I shorted it. I think it's important.
 
@@
 interactions to achieve a higher adsorbed gas density than the bulk gas at the same temperature and pressure, allowing for room temperature and lower-pressure storage and alleviating many drawbacks of high-pressure storage.
-% This would thereby reduce the cost of infrastructure at refilling stations, allow thinner-walled and therefore cheaper and lighter pressure vessels, reduce the energy requirements for gas densification, and alleviate safety concerns with high pressure storage.
-
-%As most homes in the US are connected to a natural gas pipeline \addcite, adsorbed natural gas fuel tanks could permit at-home fueling.
-
-% Jordan: I commented this out!
-% Cory: I think I agree that this paragraph is not necessary, thanks.
-% Current research is towards synthesizing a nanoporous material that can maximally densify natural gas/hydrogen whilst satisfying cost and stability constraints.
-So far, metal-organic frameworks (MOFs) \cite{furukawa2013chemistry} have demonstrated the highest internal surface areas ($>$7000 m$^2$/g \cite{farha2012metal}) and absorptive capacities for hydrogen \cite{suh2011hydrogen,garcia2018benchmark} and methane \cite{makal2012methane,mason2014evaluating}.
-Importantly, MOFs are highly adjustable materials synthesized by combining organic molecules (struts) and metal clusters (nodes) that
-% % I cut [often] here, since I'm not sure what the exception is and I don't think it's relevant.
-% coordinate to form a three-dimensional, porous framework \cite{furukawa2013chemistry}.
-This modular synthesis enables one to fine-tune the MOF chemistry to arrive at an optimal material for methane or hydrogen storage and delivery \cite{schoedel2016role}.
-See Fig.~\ref{fig:example_MOF} for the crystal structure of a canonical MOF, CuBTC \cite{chui1999chemically}.
 
@@
 Intuitively, the deliverable capacity of a gas in such a homogeneous substrate with the optimal potential energy is an upper bound because, in a realistic material, (i) spatial inhomogeneity of the potential energy results in some points in the control volume offering a suboptimal attraction for the gas and (ii) the atoms required to create the potential field will exclude gas from occupying a fraction of the control volume.
-%\corysays{the suboptimal energy is hard to argue since that would affect entropy too...}
-
-% \corysays{I propose to replace ``homogenous'' with ``spatially uniform'' everywhere instead of using both interchangeably.
-% \davidsays{We replaced ``homogeneous potential'' with spatially uniform, but were less confident replacing ``homogeneous substrate'', which has a slightly different meaning, and might need a slightly different name.
-e.g.
-maybe ``uniform substrate'' or ``model substrate''?}}
-
-%\corysays{what do u think about callin this an ``ideal substrate'' or giving it some name so we can refer to it hereon?
-\davidsays{I think I'd prefer ``homogeneous substrate'' or ``idealized substrate''}}
-
-%\subsection{Finding the density}\label{sec:finding-n}
 
@@
 See Sec.~\ref{sec:V_shifts_chem_pot} for a derivation.
-% where $\rho_g(\mu,T)$ is the density of the pure gas at chemical potential $\mu$, shown for methane in Fig.~\ref{fig:density-vs-mu-ch4}. 
-% cory says: ^ we already defined \rho_g in pargraph above.
 
@@
 
-% Yo David, I changed ur paragraph below. tell me wut u think.
-% To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a \emph{maximum} deliverable capacity over all static potential energy fields, consider the case of a material that offers a non-spatially-uniform potential energy felt by the gas molecules.We wish to show that this material will have a lower deliverable capacity than our ideal substrate.
-At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, some of the gas molecules will be forced into higher energy locations.
-Thus, the mean attraction will be lowest when the tank is ``full,'' and highest when the tank is ``empty.'' This is the opposite of what we would want to achieve a high deliverable capacity.
-Based on this qualitative reasoning, we expect the spatially uniform potential \red{$\V_{opt}$} to give the maximal deliverable capacity, since it will provide the most attraction for the molecules added at high densities.
-
-% \corysays{``To qualitatively argue that the spatially uniform potential provides a \emph{maximum} deliverable capacity'' do we need to qualify that this spatially uniform potential must hv the optimal $\V$?
-obviously a spatially uniform potential can be better than a spatially uniform one with a poor $\V$.
-maybe we should start with our optimal homogenous potential, then say introduce a little blip in it from there, like in calculus of variations.}
-
 \begin{figure}
@@
 
-\corysays{I know it doesn't matter, but should we choose L STP/L or cm3 STP/cm3?
-we use both, some might get confused/distracted?
-\davidsays{We chose to use the units used for each of the targets thinking it would be more familiar and/or easier to compare.
-I don't like either set of units, and if we were to pick ``our'' choice, I'd lean towards something dimensionless with a sane comparison (e.g.
-relative to the gas density at our high pressure)}}
-
-\corysays{need to explain why the MOFs are lines not points... that \gst has a range for MOFs.
-\jpsays{due to a different deliverable capacity at full pressure and empty pressure.
-For hydrogen the SI (by Paula Garcia-Holley) has total uptake plots so we could make a curve even but would have a lot of error due to manual extraction?}}
-
 \begin{figure}
@@
 as written the legend is not correct since those are not the functions being shown.
-in the main text we never discussed what \gst is for the MOF and why the MOFs are lines not points (\gst is a function of density now)}
+in the main text we never discussed what \gst is for the MOF and why the MOFs are lines not points (\gst is a function of density now)
+\jpsays{due to a different deliverable capacity at full pressure and empty pressure.
+For hydrogen the SI (by Paula Garcia-Holley) has total uptake plots so we could make a curve even but would have a lot of error due to manual extraction?}}
 
@@
 By raising the temperature of the adsorbent during gas discharge to drive off residual gas~\cite{gomez2014exploring}, the deliverable capacity could be enhanced, albeit at the cost of a more complicated engineering design of the fuel tank and vehicle.
-%While this approach poses engineering challenges, in principle it could allow an arbitrarily large deliverable capacity.
-
-% \corysays{no, cant lower filling temp. when park it will heat up then go above pressure rating of tank...}
 
@@
 We note that our upper bound can readily be applied to the storage of other gasses of interest, with the proviso that the gas is far from crystallization. Our code is available at \davidsays{XXXXX}.
-
-% \clearpage
-
-% Here we will articulate the assumptions required for our upper bound to hold. Some of these assumptions serve as a guide for chemists who aim to synthesize a material that violates our upper bound.
-
-% Our proof that the spatially uniform potential is an extremum assumes that the fluid is stable at the temperature and densities involved.
-% \begin{limitation}\label{limit:crystallization}
-% Our proof assumes the gas is cannot become spontaneously solid at the temperature of interest.
-% \end{limitation}
-% Our upper bound will \emph{not} be valid for gasses which may crystallize in their pure form.
-This is because a periodic potential can interact with a high-density crystal far more strongly than it interacts with a low-density gas.
-We are not aware of a practical application of this principle, but theoretically it is possible.
-For the application of gas storage for light gasses such as methane or hydrogen at moderately high temperatures (say liquid nitrogen or warmer), the pure gas is sufficiently far from its crystal form that we need not concern ourselves with crystallization.
-We do not see this particular assumption as one that can be worked around.
-
-% There remains the possibility that the spatially uniform potential is a local maximum, but there is another global maximum.
-We do not see how to formally rule out this possibility for an arbitrary gas-gas interaction, but the qualitative arguments above seem sufficient to rule out this possibility for a real gas.
-
-% However, there was a more fundamental assumption, which is that the porous material acts as an effective potential for the gas molecules, and that the gas-gas interaction is unmodified.
-% \begin{limitation}\label{limit:field}
-% Our proof assumes the interaction of the gas with the substrate does not modify the substrate.  Equivalently,  the interaction of substrate with gas operates as a simple potential energy field.
-% \end{limitation}
-% For \emph{most} porous materials this seems likely to be a pretty good approximation.
-However, there are unusual cases where the substrate can in effect provide a very strongly density-dependent interaction.
-This is perhaps easiest to see in the example of \davidsays{look up the collapsing MOFs description from the RMS-MOF paper}, in which the gas can introduce a phase change in the solid, which has the effect of drastically lowering the density of gas in the material.
-Such materials can indeed violate our upper bound.
-\davidsays{Remember to discuss this in the conclusions or perhaps discussion, we can frame these limitations as guides.
-The limitations of our upper bound are the design requirements for a material that will do any better.}
-
-% [impurities?]
-
-% \section{Conclusion}
-% As many have suspected \davidsays{cite previous work}, pressure-swing adsorption in rigid porous solids is not going to enable reaching targets.
-\davidsays{conclude explicitly about each target}  Unlike previous work, our upper bound on deliverable capacity is based on experimentally known properties of the gas of interest rather than simulations.
-
-% The limitations on our proof of an upper bound on the deliverable capacity provide hints as to approaches that could in principle achieve the targets.
-In this paper we have identified three limitations of our upper bound on the deliverable capacity.
-Two of these three limitations suggest engineering strategies that do hold promise of achieving the targets for hydrogen and methane storage.
-Limitation \ref{limit:crystallization}, which states that the gas must not crystallize does not means for improvement, since it is a limit on the gasses to which our bound may be applied, and neither hydrogen nor methane are in any danger of crystallizing.
-
-% Crystal density vs. pelletiziation density.
-
-% A more subtle limitation of our proof is that it assumes that the substrate itself is not modified by the presence of the gas.
-While this assumption is never precisely true, since any adsorbed atom will change the electron distribution of the substrate to some degree, it tends to be sufficiently accurate that most atomistic simulations can get away with assuming that the substrate is unmodified.
-A dramatic exception to that rule is the field of flexible MOFs, in which the presence of the gas can induce a phase change which can \emph{drastically} modify the gas-substrate interaction~\davidsays{cite experiments}\cite{witman2019}.
-
-% wut about mixtures?

### 2020-05-11 gas-adsorption/gas-ads-body.tex
checkout all of the calculations and removed comments that were answered.

@@
 For an adsorbent with at least 84\% void fraction, the ARPA-E target is theoretically possible.
-The experimental deliverable capacities of several adsorbents~\cite{mason2014evaluating} are also shown over a range of $\gst$ (converted from the measured adsorption isotherm as explained in \davidsays{confirm this} Sec.~\ref{sec:phi-is-delta-g}).
+The experimental deliverable capacities of several adsorbents~\cite{mason2014evaluating} are also shown over a range of $\gst$ (converted from the measured adsorption isotherm as explained in Sec.~\ref{sec:phi-is-delta-g}).
 These can be compared with the highest observed deliverable capacities for methane at room temperature in rigid materials, up to 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}.
@@
 
-\jpsays{I edited everything below this comment. The original is commented out}
-
 Storage of hydrogen is considerably more challenging owing to its relatively weak interaction with adsorbents.
-The DOE ULTIMATE deliverable capacity target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C \red{check}.
+The DOE ULTIMATE deliverable capacity target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C.
 Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve for hydrogen storage along with a range of experimental measurements for known MOFs.
 The DOE ULTIMATE deliverable capacity target is theoretically possible, however by such a small margin that we can safely rule out the possibility of reaching this target through storage and delivery of hydrogen in \emph{any} rigid substrate at room temperature.
-Such a material would require a void fraction of at least 94\% \red{check}.
+Such a material would require a void fraction of at least 94\%.
 On top of this, the DOE ULTIMATE target requires an optimal $|\gst|$ of 10~kJ/mol which is far greater than what is found in observed porous materials.
@@
 While these upper bounds do not rule out the discovery of materials that reach current DOE targets, they cast strong doubt on the possibility of achieving these goals when we consider the additional constraints imposed due to steric hindrance between substrate atoms and the adsorbate.
-Our upper bound does indicate that those goals cannot be exceeded by more than 19\% for methane and 6\% for hydrogen \red{check}.
+Our upper bound does indicate that those goals cannot be exceeded by more than 16\% for methane and 6\% for hydrogen.
 Fortunately, there are some limitations to these upper bounds which suggest avenues for future developments which could allow the targets to be met.

### 2020-05-11 gas-adsorption/gas-ads-body.tex
working through the paper and making edits and removing outdated comments.

@@
 \section{Introduction}
-The transportation sector is dominantly powered by petroleum-based fuels \cite{davis2009transportation}.
-In the US, it is responsible for 36\% of energy-related carbon dioxide emissions \cite{useia} and 50,000 premature deaths per year associated with particulate matter and ozone emissions \cite{caiazzo2013air}.
-Natural gas and hydrogen (H$_2$) are alternative transportation fuels that emit fewer/less pollutants 
-than petroleum-based fuels, and thus their widespread adoption could mitigate climate change 
-\cite{mcglade2015geographical} and improve air quality for human health. Further, as the finite global 
-petroleum resources are declining rapidly \cite{sorrell2010global}, the development of technologies for the widespread
-adoption of sustainable transportation fuels, such as hydrogen, is critical.
+The transportation sector is dominantly powered by petroleum-based
+fuels~\cite{davis2009transportation}. In the United Sates, it is responsible
+for 36\% of energy-related carbon dioxide emissions \cite{useia} and 50,000
+premature deaths per year associated with particulate matter and ozone
+emissions~\cite{caiazzo2013air}. Natural gas and hydrogen (H$_2$) are
+alternative transportation fuels that emit fewer/less pollutants than
+petroleum-based fuels, and thus their widespread adoption could mitigate
+climate change~\cite{mcglade2015geographical} and improve air quality for human
+health. Further, as the finite global petroleum resources are declining
+rapidly~\cite{sorrell2010global}, the development of technologies for the
+widespread adoption of sustainable transportation fuels, such as hydrogen, is
+critical.
 
-Natural gas, mostly methane, is considered a transition (to a renewable and clean) fuel because 
-it emits 25\% less carbon dioxide \cite{eia2013much} and less toxic 
-byproducts \cite{wang2000full} upon combustion per unit energy produced compared to gasoline. From an 
-economic standpoint, the supply of natural gas in the US is increasing as a result of hydraulic 
-fracturing and horizontal drilling techniques \cite{usnatgassupply}. A positive environmental outlook 
-for natural gas, however, is predicated on mitigating fugitive emissions (methane is itself a potent 
-greenhouse gas) \cite{alvarez2012greater} and groundwater contamination \cite{osborn2011methane}
-from hydraulic fracturing.
+Natural gas, mostly methane, is considered a transition (to a renewable and
+clean) fuel because it emits 25\% less carbon dioxide~\cite{eia2013much} and
+fewer toxic byproducts~\cite{wang2000full} upon combustion per unit energy
+produced compared to gasoline. From an economic standpoint, the supply of
+natural gas in the United States is increasing as a result of hydraulic
+fracturing and horizontal drilling techniques~\cite{usnatgassupply}. A positive
+environmental outlook for natural gas, however, is predicated on mitigating
+fugitive emissions (methane is itself a potent greenhouse
+gas)~\cite{alvarez2012greater} and groundwater
+contamination~\cite{osborn2011methane} from hydraulic fracturing.
 
-Hydrogen (H$_2$) is the ultimate transportation fuel because it emits only water when it electrochemically reacts with oxygen in a fuel cell to produce electricity and power a vehicle.
-% Moreover, elemental hydrogen is very abundant, though bonded with oxygen (in water) or carbon (in hydrocarbons); H$_2$ does not occur naturally in abundant quantities.
-Currently, hydrogen (H$_2$) is primarily produced by steam reforming of natural gas followed by the
-water-gas shift reaction, which emits carbon dioxide \cite{crabtree2004hydrogen}. Notably, the 
-environmental allure of hydrogen is predicated on its production via a renewable means, e.g.\ splitting
+Hydrogen (H$_2$) is the ultimate transportation fuel because it emits only
+water when it electrochemically reacts with oxygen in a fuel cell to power a
+vehicle. Currently, hydrogen (H$_2$) is primarily produced by steam reforming
+of natural gas followed by the water-gas shift reaction, which emits carbon
+dioxide \cite{crabtree2004hydrogen}. Notably, the environmental allure of
+hydrogen is predicated on its production via a renewable means, e.g.\ splitting
 water using wind-generated electricity.
 
-% \jordan{I got the values from wikipedia but the citation that wikipedia gave didn't link to the correct paper? Article is Energy Density.
-% THIS PAPER says H2 at 700 bar is 4.7 MJ/L: Review on processing of metal–organic framework (MOF) materials towards system integration for hydrogen storage
-% }
 At ambient conditions, both methane

### 2020-05-12 gas-adsorption/gas-ads-body.tex
I have worked through to section II.

@@
 
-At ambient conditions, both methane
-%(0.036 MJ/L)
-and hydrogen gas
-%(0.01 MJ/L)
-possess a very low volumetric energy density compared to (liquid) gasoline.
-%(34.2 MJ/L)
-Consequently, under storage space constraints in passenger vehicles, natural gas and hydrogen must be densified for onboard storage to achieve a reasonable driving range on a ``full'' tank of fuel.
-Traditional densification approaches are liquefaction (cryogenic storage near atmospheric pressure)
+% methane (0.036 MJ/L); hydrogen gas (0.01 MJ/L); liquid gasoline (34.2 MJ/L).
 % (at 1 atm, liquid CH$_4$ at 111~K to achieve 22.2 MJ/L~\cite{makal2012methane}, liquid H$_2$ at 20 K to achieve 8~MJ/L \cite{suh2011hydrogen})
-or compression to high pressures (room temperature storage).
 % (200 bar for CH$_4$ to achieve 9.2 MJ/L \cite{makal2012methane} and 700 bar for H$_2$ to achieve 4.5-5.3 MJ/L~\cite{unknown}).
-Both approaches require expensive infrastructure at refilling stations and significant energy input; e.g., liquefaction of hydrogen consumes at least 30\% of its energy content~\cite{bossel2003energy}.
-Moreover, high-pressure storage tanks are heavy, thick-walled, and non-conformable, while cryogenic storage tanks are bulky, expensive, and afflicted by boil-off losses~\cite{hasan2009minimizing}.
 
-A promising approach to densify natural gas~\cite{makal2012methane,mason2014evaluating} and hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is through physical adsorption in nanoporous materials at room temperature~\cite{schoedel2016role}.
-The internal surfaces of porous materials attract gas molecules through van der Waals, electrostatic, etc.
-interactions to achieve a higher adsorbed gas density than the bulk gas at the same temperature and pressure, allowing for room temperature and lower-pressure storage and alleviating many drawbacks of high-pressure storage.
+At ambient conditions, both methane and hydrogen gas possess a very low
+volumetric energy density compared to gasoline. Consequently, under storage
+space constraints in passenger vehicles, natural gas and hydrogen must be
+densified for onboard storage to achieve a reasonable driving range on a
+``full'' tank of fuel. Traditional densification approaches are liquefaction,
+at cryogenic temperatures and atmospheric pressure, or compression, at room
+temperature and high pressures. Both approaches require expensive
+infrastructure at refilling stations and significant energy input; e.g.,
+liquefaction of hydrogen consumes at least 30\% of its energy
+content~\cite{bossel2003energy}. Moreover, high-pressure storage tanks are
+heavy, thick-walled, and non-conformable, while cryogenic storage tanks are
+bulky, expensive, and afflicted by boil-off losses~\cite{hasan2009minimizing}.
 
-For a vehicle employing a porous material to store natural gas or hydrogen, the (volumetric) \emph{deliverable capacity} of the gas is the primary thermodynamic property of the material that determines the driving range~\cite{mason2014evaluating}.
-The adsorbed gas storage tank delivers the gaseous fuel to the engine via an (assumed) isothermal pressure swing~\cite{sircar2002pressure}.
-The deliverable capacity (see Fig.~\ref{fig:delcap}) is the density of the gas in the material at the storage pressure $\pfull$ minus the residual gas that remains adsorbed at the lowest pressure $\pempty$ such that sufficient flow is maintained to feed the engine.
-For commercial feasibility, the US Department of Energy (DOE) has set deliverable capacity targets for both adsorbed methane and hydrogen storage and delivery for vehicles.
-For methane, the Advanced Research Projects Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$ per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure swing~\cite{simon2015materials}.
-For hydrogen, the DOE set a series of progressive targets at five year intervals, with the ultimate target of 0.05 kg H$_2$/L~\cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a minimum of -40 $^\circ$C~\cite{allendorf2018assessment}.
-Thus far, despite the emergence of highly tunable materials with enormous surface areas, such as metal-organic frameworks~\cite{furukawa2013chemistry}, no porous material has met these deliverable capacity targets.
+A promising approach to densify natural
+gas~\cite{makal2012methane,mason2014evaluating} and
+hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is
+through physical adsorption in nanoporous materials at room
+temperature~\cite{schoedel2016role}. The internal surfaces of porous materials
+attract gas molecules through van der Waals, electrostatic, etc. interactions
+to achieve a higher adsorbed gas density than the bulk gas at the same
+temperature and pressure, allowing for room temperature and lower-pressure
+storage and alleviating many drawbacks of high-pressure storage.
+
+For a vehicle employing a porous material to store natural gas or hydrogen, the
+(volumetric) \emph{deliverable capacity} of the gas is the primary
+thermodynamic property that determines the driving
+range~\cite{mason2014evaluating}. The adsorbed gas storage tank delivers the
+gaseous fuel to the engine via an (assumed) isothermal pressure
+swing~\cite{sircar2002pressure}. The deliverable capacity (see
+Fig.~\ref{fig:delcap}) is the density of the gas in the material at the storage
+pressure $\pfull$ minus the residual gas that remains adsorbed at the lowest
+pressure $\pempty$ such that sufficient flow is maintained to feed the engine.
+For commercial feasibility, the US Department of Energy (DOE) has set
+deliverable capacity targets for both adsorbed methane and hydrogen storage and
+delivery for vehicles. For methane, the Advanced Research Projects
+Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$
+per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure
+swing~\cite{simon2015materials}. For hydrogen, the DOE set a series of
+progressive targets at five year intervals, with the ultimate target of 0.05 kg
+H$_2$/L~\cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a
+minimum of -40 $^\circ$C~\cite{allendorf2018assessment}. Thus far, despite the
+emergence of highly tunable materials with enormous surface areas, such as
+metal-organic frameworks~\cite{furukawa2013chemistry}, no porous material has
+met these deliverable capacity targets.
 
@@
 
-To set realistic performance targets and optimally allocate resources for research efforts, in this work, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a rigid porous material and uses as input the experimentally measured properties of the bulk gas.
-Our extremum is provided by a substrate that offers a spatially uniform potential energy field felt by the gas.
-Applying our framework to methane and hydrogen gas, we find the US DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-Optimistically, new paradigms outside the scope of applicability of our theoretical framework, such as gas-induced structural transitions of the material, hold promise for meeting these targets, as evidenced by flexible MOF Co(bdp) which currently boasts the largest methane deliverable capacity~\cite{mason2015methane}.
-
-% Jordan: I commented this out!
-% To optimally allocate resources for research efforts and set realistic performance targets, it is important to assess the thermodynamic limits of technologies.
-Here, we present a theoretical framework that places an intrinsic upper limit on the deliverable capacity of gas in a nanoporous material and uses as input experimental measurements of the bulk gas.
-Applying our framework to methane and hydrogen gas, we find the DOE deliverable capacity targets for natural gas and hydrogen storage and delivery are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
+To set realistic performance targets and optimally allocate research resources,
+in this work, we present a theoretical framework that places an intrinsic upper
+limit on the deliverable capacity of any gas in a rigid porous material and
+uses as input the experimentally measured properties of the bulk gas. Our
+extremum is provided by a substrate that offers a spatially uniform potential
+energy field felt by the gas. Applying our framework to methane and hydrogen
+gas, we find the US DOE deliverable capacity targets for natural gas and
+hydrogen storage and delivery are theoretically possible, but sufficiently
+close to the upper bound as to be impractical for any real porous material.
+Optimistically, new paradigms outside the scope of applicability of our
+theoretical framework, such as gas-induced structural transitions of the
+material, hold promise for meeting these targets, as evidenced by flexible MOF
+Co(bdp) which currently boasts the largest methane deliverable
+capacity~\cite{mason2015methane}.
 

### 2020-05-12 gas-adsorption/gas-ads-body.tex
I worked through section III.

@@
 \section{Gas storage \& delivery by isothermal, pressure-swing adsorption}
-% Here, we pose the context and question that we address in this work.
-% Consider pure gas storage and delivery by isothermal, pressure-swing adsorption \cite{sircar2002pressure}.
-Consider a pressure vessel onboard a vehicle (i.e.
-fuel tank) packed with porous material.
-At the filling stage, the tank is connected to a (pure) gaseous reservoir at pressure $\pfull$ and allowed to equilibrate.
-At this point, the adsorbed gas tank is considered full.
-While driving, gas desorbs from the adsorbent to the engine/fuel cell, driven by a pressure differential.
-The tank is considered depleted/empty when the pressure has dropped to $\pempty$, the pressure at which the flow rate of gas from the tank to the engine is insufficient.
-%
-However, given $\pempty \neq 0$ (pulling vacuum), residual gas will remain trapped in the adsorbent.
-Therefore, the driving range of the vehicle is primarily determined by the \emph{deliverable capacity}, defined as the density of gas in the tank at $\pfull$ minus density of the the residual gas remaining at $\pempty$.
-The [isothermal, volumetric] deliverable capacity is an intrinsic property of the nanoporous material and its interaction with the gas.
-%Precisely, let $\rho(p; T)$ be the density of gas in the adsorbent when in equilibrium with a bath of pure gas at pressure $p$ and temperature $T$, the \emph{adsorption isotherm}; $\rho(p,T)$ is essentially the equation of state of the gas adsorbed in the porous material.
-The [intrinsic, isothermal] deliverable capacity $\rho_D$ of the gas in the adsorbent is:
-%\begin{equation}
-%    \rho_D(\pfull, \pempty; T) \equiv \rho(\pfull; T) - \rho(\pempty; T).
-%    \label{eq:delcap}
-%\end{equation}
-% The pressures $\pfull$ and $\pempty$ correspond to chemical potentials $\mufull$ and $\muempty$ via an equation of state for the gas.
-% Eqn.~\ref{eq:delcap}
-% This assumes that the tank reaches equilibrium %, neglecting the kinetics of ad/desorption
-% at the end of the filling and depletion stages.
-
-%The assumption of isothermal operation neglects (i) heat released/consumed during ad/desorption, which can diminish the deliverable capacity by raising/lowering the temperature of the adsorbent during filling/discharge \cite{mota1997dynamics,chang1996behavior}, and (ii) an engineering strategy to use waste heat from the engine to heat the adsorbent and drive off residual gas when it nears $\pempty$ \cite{gomez2014exploring}.
-
-%Second, an engineering strategy to increase the deliverable capacity beyond the isothermal deliverable capacity is to reroute waste heat from the engine to the adsorbed gas tank during delivery.
-By raising the temperature of the adsorbent during the discharge stage, we drive off a fraction of the residual gas that would otherwise remain when the tank reaches $\pempty$ .
-In such a case, the [non-isothermal] deliverable capacity of the tank is $\rho(\pfull, T)-\rho(\pempty; T_H)$ where $T_H$ is the temperature to which we heat the adsorbent at discharge.
-This strategy, however, requires a more complicated design.
-
-% \begin{limitation}\label{limit:isothermal}
-% We assume isothermal pressure-swing adsorption.
-It is possible to achieve a greater deliverable capacity by raising the temperature of the adsorbent as the gas is extracted to drive off the residual gas.
-% \end{limitation}
+Consider a pressure vessel onboard a vehicle (i.e. fuel tank) packed with
+porous material. At the filling stage, the tank is connected to a (pure)
+gaseous reservoir at pressure $\pfull$ and allowed to equilibrate. At this
+point, the adsorbed gas tank is considered full. While driving, gas desorbs
+from the adsorbent to the engine/fuel cell, driven by a pressure differential.
+The tank is considered depleted/empty when the pressure has dropped to
+$\pempty$, the pressure at which the flow rate of gas from the tank to the
+engine is insufficient. However, given $\pempty \neq 0$ (pulling vacuum),
+residual gas will remain trapped in the adsorbent. Therefore, the driving range
+of the vehicle is primarily determined by the \emph{deliverable capacity}. The
+isothermal, volumetric deliverable capacity is an intrinsic property of the
+nanoporous material and its interaction with the gas.
 
 \section{Review of previous work}
-There has been considerable work attempting to establish an upper bound on the isothermal deliverable capacity in pressure-swing adsorption.
-% We now review previous work towards placing an upper bound on the isothermal deliverable capacity of hydrogen or methane by pressure-swing adsorption.
-% The first body of work rests on the simplified Langmuir model; the second body of work rests upon imposing energetic models for the gas-substrate and gas-gas interactions.
+There has been considerable work attempting to establish an upper bound on the
+isothermal deliverable capacity in pressure-swing adsorption.
 
-Early work showed that, in the simplified Langmuir model, there exists an optimal free energy of adsorption (which determines the Langmuir constant) that maximizes the deliverable capacity $\rho_D$~\cite{matranga1992storage,bhatia2006optimum,simon2014optimizing}.
-If the gas-substrate interaction is too weak (strong), too little (much) gas adsorbs (is retained) at $\pfull$ ($\pempty$), diminishing $\rho_D$.
-An upper bound on the deliverable capacity of a Langmuir material follows if each adsorption site is endowed with the optimal free energy of adsorption.
-However, remaining is the question of how many adsorption sites per volume a porous material can practically offer, under the constraint that these adsorption sites offer the optimal free energy of adsorption.
-Moreover, gas-gas attractions, neglected in the Langmuir model, could recruit more gas in the material at $\pfull$ than at $\pempty$ and enhance the deliverable capacity~\cite{simon2014optimizing}.
+Early work showed that, in the simplified Langmuir model, there exists an
+optimal free energy of adsorption (which determines the Langmuir constant) that
+maximizes the deliverable capacity
+$\rho_D$~\cite{matranga1992storage,bhatia2006optimum,simon2014optimizing}. If
+the gas-substrate interaction is too weak (strong), too little (much) gas
+adsorbs (is retained) at $\pfull$ ($\pempty$), diminishing $\rho_D$. An upper
+bound on the deliverable capacity of a Langmuir material follows if each
+adsorption site is endowed with the optimal free energy of adsorption. However,
+remaining is the question of how many adsorption sites per volume a porous
+material can practically offer, under the constraint that these adsorption
+sites provide the optimal free energy of adsorption. Moreover, gas-gas
+attractions, neglected in the Langmuir model, could recruit more gas in the
+material at $\pfull$ than at $\pempty$ and enhance the deliverable
+capacity~\cite{simon2014optimizing}.
 
-G\'omez-Gualdr\'on \emph{et al.}~\cite{gomez2017impact} introduced a model that accounted for gas-gas interactions via an intermolecular potential and idealized the substrate in two different ways: (1) discrete adsorption sites packed into an FCC lattice and (2) a volume endowed with a spatially uniform, background potential energy field.
-According to molecular simulation of methane adsorption, the ARPA-E deliverable capacity target of 315 L STP/L could be reached in both of these idealized substrates.
-%The ARPA-E target of 315 L STP/L was reached when adsorption sites were endowed with an optimal heat of adsorption of 11.5 kJ/mol and adsorption sites were $5.3$ \AA~apart.
-In a second model, they endowed a volume with a background potential energy field, then simulated the adsorption of methane in the volume in the grand-canonical ensemble, using a Lennard-Jones potential for methane-methane interactions.
-The maximal deliverable capacity of 357 L STP/L was achieved with a background heat of adsorption of 8.6 kJ/mol.
-However, both models neglected the space occupied by atoms of the porous material that are needed to endow the adsorption sites/volume with the attractive potential energy.
-% To address this, in the third model, they constructed pseudo materials by placing Lennard-Jones spheres in space with tunable well depths.
-When constrained to form a connected network, none of the psuedo-materials reached the ARPA-E deliverable capacity target.
+G\'omez-Gualdr\'on \emph{et al.}~\cite{gomez2017impact} introduced a model that
+accounted for gas-gas interactions via an intermolecular potential and
+idealized the substrate in two different ways: (1) discrete adsorption sites
+packed into an FCC lattice and (2) a volume endowed with a spatially uniform,
+background potential energy field. According to molecular simulation of methane
+adsorption, the ARPA-E deliverable capacity target of 315 L STP/L could be
+reached in both of these idealized substrates. However, both models neglected
+the space occupied by atoms of the porous material that are needed to endow the
+adsorption sites/volume with the attractive potential energy.
 
-A third body of work attempted to account for both gas-gas interactions \emph{and} steric interactions of the gas with the atoms of the porous material.
-% explores the upper limits of the deliverable capacity while accounting for gas-gas interactions \emph{and} the atoms of the porous material that are required to attract the guest; these material atoms exclude the adsorbate from regions of the control volume.
-%More works obtained an empirical upper bound on the deliverable capacity by explicitly accounting for the atoms of the material needed to create the potential energy field felt by the gas molecules.
-Simulations of methane adsorption in hundreds of thousands of explicit nanoporous crystal structures---both real and hypothetical---suggested that the ARPA-E deliverable capacity target is infeasible~\cite{simon2015materials}; the highest simulated methane deliverable capacity was 196 cm$^3$~STP/cm$^3$.
-Confidence in this conclusion rests upon (i) the accuracy of the intermolecular potentials describing the molecular interactions and (ii) the sufficient sampling of material space, i.e.\ that the structures considered are representative of the set of possible materials.
-To further explore material space and address sensitivity to the potentials used, scaling the Lennard-Jones potential well depths of material atoms to model enhanced interactions~\cite{gomez2014exploring}, placing Lennard-Jones spheres in a unit cell randomly to form ``pseudo-materials''~\cite{kaija2018high}, and generating fictitious potential energy fields via a generative adversarial network trained on zeolite structures~\cite{lee2019predicting} all generated model substrates that all failed to meet the ARPA-E methane deliverable capacity target.
+A third body of work attempted to account for both gas-gas interactions
+\emph{and} steric interactions of the gas with the atoms of the porous
+material. Simulations of methane adsorption in hundreds of thousands of
+explicit nanoporous crystal structures---both real and hypothetical---suggested
+that the ARPA-E deliverable capacity target is
+infeasible~\cite{simon2015materials}; the highest simulated methane deliverable
+capacity was 196 cm$^3$~STP/cm$^3$. Confidence in this conclusion rests upon
+(i) the accuracy of the intermolecular potentials describing the molecular
+interactions and (ii) the sufficient sampling of material space, i.e.\ that the
+structures considered are representative of the set of possible materials. To
+further explore material space and address sensitivity to the potentials used,
+scaling the Lennard-Jones potential well depths of material atoms to model
+enhanced interactions~\cite{gomez2014exploring}, placing Lennard-Jones spheres
+in a unit cell randomly to form ``pseudo-materials''~\cite{kaija2018high}, and
+generating fictitious potential energy fields via a generative adversarial
+network trained on zeolite structures~\cite{lee2019predicting} all generated
+model substrates that failed to meet the ARPA-E methane deliverable
+capacity target.
 
@@
 
-In this work, we place a rigorous upper bound on the deliverable capacity of a pure gas in a rigid substrate by endowing a control volume with a spatially uniform background energy field.
-Instead of using a molecular model for the gas~\cite{gomez2017impact}, we use the experimental equation of state to account for gas-gas interactions.
-In addition, we prove using the calculus of variations that this spatially uniform substrate yields a maximal deliverable capacity.
-We use our framework to place an upper bound on the deliverable capacity of methane and hydrogen.
+In this work, we place a rigorous upper bound on the deliverable capacity of a
+pure gas in a rigid substrate by endowing a control volume with a spatially
+uniform background energy field. Instead of using a molecular model for the
+gas~\cite{gomez2017impact}, we use the experimental equation of state to
+account for gas-gas interactions. In addition, we prove using the calculus of
+variations that this spatially uniform substrate yields a maximal deliverable
+capacity. We use our framework to place an upper bound on the deliverable
+capacity of methane and hydrogen gas.
 
@@
 For comparison, we also show the density of methane adsorbed into several porous materials.
-
-% \davidsays{This paragraph could move considerably later. Or cut entirely?}
-% We model the interaction of substrate with gas as a simple Gibbs free energy of adsorption per mole (or atom) \gst.
-We will assume that this Gibbs free energy is independent of the concentration of gas.
-At low concentrations, this assumption is reasonable, because the gas molecules can be assumed to be widely separated and their interactions weak.
-In the limit of high concentrations, we expect that the attraction of the gas to the substrate will be diminished, since the ``good spots'' will already be occupied.
-The choice of a concentration-independent $\gst$ will thus have a tendency to overestimate the deliverable capacity, providing an upper bound (for more on this, see Sec.~\ref{sec:qualitative-max}).
 

### 2020-05-13 gas-adsorption/gas-ads-body.tex
working through section IV.

@@
 
-The thermodynamic properties of a bulk, pure gas are characterized by an equation of state. 
-Of particular interest for gas storage and delivery is the density of the gas as a function of chemical potential $\mu$ and temperature $T$, $\rho_g(\mu,T)$, which is shown for methane gas in Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K.
-For comparison, we also show the density of methane adsorbed into several porous materials.
+The thermodynamic properties of a bulk, pure gas are characterized by an
+equation of state. Of particular interest for gas storage and delivery is the
+density of the gas $\rho_g(\mu,T)$ as a function of chemical potential $\mu$
+and temperature $T$, which is shown for methane gas in
+Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K. For comparison, we also show
+the density of methane adsorbed into several porous materials.
 
-To place an upper bound on the deliverable capacity, consider a substrate whose sole interaction with the gas is to introduce a spatially uniform potential energy $\V$
-for gas molecules in the control volume (where $\V<0$ for an attractive potential).
-Because this interaction is uniform within the substrate, the gas-gas interactions and thus fluid structure in such a substrate are identical to those of the pure gas at the same density and temperature.
-This allows us to obtain the adsorption properties of this model substrate using the experimentally measured properties of the pure gas~\cite{nist}.
-Intuitively, the deliverable capacity of a gas in such a homogeneous substrate with the optimal potential energy is an upper bound because, in a realistic material, (i) spatial inhomogeneity of the potential energy results in some points in the control volume offering a suboptimal attraction for the gas and (ii) the atoms required to create the potential field will exclude gas from occupying a fraction of the control volume.
+To place an upper bound on the deliverable capacity, consider a substrate whose
+sole interaction with the gas is to introduce a spatially uniform potential
+energy $\V$ for gas molecules in the control volume (where $\V<0$ for an
+attractive potential). Because this interaction is uniform within the
+substrate, the gas-gas interactions and thus fluid structure in such a
+substrate are identical to those of the pure gas at the same density and
+temperature. This allows us to obtain the adsorption properties of this model
+substrate using the experimentally measured properties of the pure
+gas~\cite{nist}. Intuitively, the deliverable capacity of a gas in such a
+homogeneous substrate with the optimal potential energy is an upper bound
+because, in a real material, (i) spatial inhomogeneity of the potential
+energy results in some points in the control volume offering a suboptimal
+attraction for the gas and (ii) the atoms required to create the potential
+field exclude gas from occupying a fraction of the control volume.
 
-The spatially uniform potential $\V$ representing gas-substrate interactions behaves as an external potential and effectively shifts the chemical potential (or, equivalently, molar Gibbs free energy) of the gas in the material, just as gravitational potential energy causes the density of air to vary with altitude.
+The spatially uniform potential $\V$ representing gas-substrate interactions
+behaves as an external potential and effectively shifts the chemical potential
+(or, equivalently, molar Gibbs free energy) of the gas in the material, just as
+gravitational potential energy causes the density of air to vary with altitude.
 Consequently, the density of gas in our homogeneous substrate is:
@@
 \end{align}
-where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the pressures $\pfull$ and $\pempty$, respectively.
-Thus, $\rho_D$ for our homogeneous substrate as a function of $\V$ is the difference between two shifted versions of $\rho_g(\mu; T)$.
-Figure~\ref{fig:methane-298-D} shows the two shifted bulk methane density curves and their difference.
-The horizontal axis
-is the difference in molar Gibbs free energy at fixed density and temperature between the gas within the substrate and the pure gas (\gst), which is equal to $\V$ in our model.
-We see a potential $\V$ that maximizes the deliverable capacity of our ideal substrate, balancing the need to maximize the density at $\pfull$ against the need to minimize residual gas retained at $\pempty$.
-%Our theoretical framework thus recovers an upper bound on the deliverable capacity of methane (298\ K, $\pfull=$65\ bar, $\pempty=$5.8\ bar) in porous materials as 374\ L STP/L.
-The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar, $\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L, achieved for $\V_{opt} =$ 5.9\ kJ/mol.
-%
-%As a ballpark, we can expect the attractive potential to be optimal around the maximum slope of $\rho_g(\mu)$.
-This is the point where the steric repulsion between the gas atoms starts leading to saturation of the density.
+where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the
+pressures $\pfull$ and $\pempty$, respectively. Thus, $\rho_D$ for our
+homogeneous substrate as a function of $\V$ is the difference between two
+shifted versions of $\rho_g(\mu; T)$. Figure~\ref{fig:methane-298-D} shows the
+two shifted bulk methane density curves and their difference. The horizontal
+axis is the difference in molar Gibbs free energy at fixed density and
+temperature between the gas within the substrate and the pure gas (\gst), which
+is equal to $\V$ in our model. We see a potential $\V$ that maximizes the
+deliverable capacity of our ideal substrate, balancing the need to maximize the
+density at $\pfull$ against the need to minimize residual gas retained at
+$\pempty$.
 
-An essential question is whether our idealized substrate offering the spatially uniform potential $\V_{opt}$ for the gas places an upper bound on the deliverable capacity in all rigid porous materials.
-%yields the maximum deliverable capacity over all possible potential energy fields. 
-In the SI (see Section~\ref{sec:proof-extremum}), we show that our idealized substrate with spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the deliverable capacity over all possible \emph{static potential energy fields}, provided the gas does not crystallize at the temperature and in the density range of interest.
-The potential energy field is \emph{static} if it is unaffected by the presence of the gas; consequently, our model does not apply to flexible materials that undergo gas-induced conformation changes \cite{schneemann2014flexible}.
-We next argue that this extremum is a maximum by addressing two alternative possibilities: the extremum could turn out to be (i) a saddle point or (ii) a local, rather than global, maximum.
+The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar,
+$\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L,
+achieved for $\V_{opt} =$ 5.9\ kJ/mol.
+
+An essential question is whether our idealized substrate, with spatially
+uniform potential $\V_{opt}$ for the gas, places an upper bound on the
+deliverable capacity in all rigid porous materials. In the SI (see
+Section~\ref{sec:proof-extremum}), we show that our idealized substrate with
+spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the
+deliverable capacity over all possible \emph{static potential energy fields},
+provided the gas does not crystallize at the temperature and in the density
+range of interest. The potential energy field is \emph{static} if it is
+unaffected by the presence of the gas; consequently, our model does not apply
+to flexible materials that undergo gas-induced conformation
+changes~\cite{schneemann2014flexible}. We next argue that this extremum is a
+maximum by addressing two alternative possibilities: the extremum could turn
+out to be (i) a saddle point or (ii) a local, rather than global, maximum.
 

### 2020-05-13 gas-adsorption/gas-ads-body.tex
fixed some typos and removed out of date comments that are wrong.

@@
 
-Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum $\rho_D$ over all \emph{spatially uniform}, static potential energy fields.
-To qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a maximum deliverable capacity over \emph{all} (including non-spatially uniform) static potential energy fields (a broader claim), consider a non-spatially-uniform variation on $\V_{opt}$.
-We will show that the effect of this variation is to reduce the deliverable capacity.
-At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while at higher densities, the additional gas molecules will be forced into higher energy locations.
-Thus, the mean attraction will be lowest at \pfull\  and highest at \pempty.
-This will result in $\gst$ increasing monotonically with with pressure.
-As shown in the Supplementary Information (see Sec.~\ref{sec:monotonic}), this results in the deliverable capacity of a non-uniform potential being lower than the deliverable capacity of a uniform potential with $\V=\gst$ for either $\pfull$ or $\pempty$.
-Thus a spatially uniform potential will give the best deliverable capacity amongst all static potential energy fields that give the same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will give the a greater deliverable than any static non-uniform potential.
+Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum
+$\rho_D$ over all \emph{spatially uniform}, static potential energy fields. To
+qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a
+maximum deliverable capacity over \emph{all} (including non-spatially uniform)
+static potential energy fields (a broader claim), consider a
+non-spatially-uniform variation on $\V_{opt}$. We will show that the effect of
+this variation is to reduce the deliverable capacity. At low adsorbed gas
+densities, the lowest-energy positions will be preferentially occupied, while
+at higher densities, the additional gas molecules will be forced into higher
+energy locations. Thus, the mean attraction will be lowest at \pfull\ and
+highest at \pempty. This will result in $\gst$ increasing monotonically with
+pressure. As shown in the SI (see Sec.~\ref{sec:monotonic}), this results in
+the deliverable capacity of a non-uniform potential being lower than the
+deliverable capacity of a uniform potential with $\V=\gst$ for either $\pfull$
+or $\pempty$. Thus, a spatially uniform potential will give the best
+deliverable capacity amongst all static potential energy fields that give the
+same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will yield a
+greater deliverable capacity over any static non-uniform potential.
 
@@
     \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
-    Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line (not visible at this scale).}
+    Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line.}
     \label{fig:methane-298-D}
@@
 \section{Results}
-Although our theoretical framework allows us to place an upper bound on the deliverable capacity of many different gases in a rigid porous material, we focus on the maximal deliverable capacity of methane and hydrogen gas in our homogeneous substrate due to their application as transportation fuels.
-To characterize the density of the gases, $\rho_g(p, T)$, we use experimental data from NIST~\cite{nist}, which naturally includes quantum effects that particularly affect hydrogen at low temperature~\cite{kumar2006quantum}.
-In the context of storage onboard passenger vehicles, we compare our upper bound with several prominent porous materials using experimental adsorption isotherms from the literature; we also compare with deliverable capacity targets set by the US DOE.
+Although our theoretical framework allows us to place an upper bound on the
+deliverable capacity of many different gases in a rigid porous material, we
+focus on the maximal deliverable capacity of methane and hydrogen gas in our
+homogeneous substrate due to their application as transportation fuels. To
+characterize the density of the gases, $\rho_g(p, T)$, we use experimental data
+from NIST~\cite{nist}, which naturally includes quantum effects that
+particularly affect hydrogen at low temperature~\cite{kumar2006quantum}. In the
+context of storage onboard passenger vehicles, we compare our upper bound with
+several prominent porous materials using experimental adsorption isotherms from
+the literature; we also compare with deliverable capacity targets set by the US
+DOE.
 
-Figure~\ref{fig:methane-298-D} shows the upper bound for methane storage at room temperature~(374~cm$^3$~STP/cm$^3$). In addition to the predicted maximum deliverable capacity as a function of \gst,
-\corysays{subtlety here: isnt ur proof restricted to saying $\V_{opt}$ provides the upper bound?
-does it really prove that, for fixed and nonoptimal $\gst$, a spatially uniform potential is optimal over a non-spatially uniform one?
-I don't think it does...}
-the ARPA-E target of 315~cm$^3$~STP/cm$^3$~\cite{arpaemove} is shown for context.
-For an adsorbent with at least 84\% void fraction, the ARPA-E target is theoretically possible.
-The experimental deliverable capacities of several adsorbents~\cite{mason2014evaluating} are also shown over a range of $\gst$ (converted from the measured adsorption isotherm as explained in Sec.~\ref{sec:phi-is-delta-g}).
-These can be compared with the highest observed deliverable capacities for methane at room temperature in rigid materials, up to 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}.
-% don't think the below is helpful or correct: Gst opt is reached by some... there are void fraction constraints too.
-% To reach the ARPA-E target, an adsorbent material with a higher $|\gst|$ would need to be found.
+Figure~\ref{fig:methane-298-D} shows the upper bound for methane storage at
+room temperature~(374~cm$^3$~STP/cm$^3$). In addition to the predicted maximum
+deliverable capacity as a function of \gst, the ARPA-E target of
+315~cm$^3$~STP/cm$^3$~\cite{arpaemove} is shown for context. For an adsorbent
+with at least 84\% void fraction, the ARPA-E target is theoretically possible.
+The experimental deliverable capacities of several
+adsorbents~\cite{mason2014evaluating} are also shown over a range of $\gst$
+(converted from the measured adsorption isotherm as explained in
+Sec.~\ref{sec:phi-is-delta-g}). These can be compared with the highest observed
+deliverable capacities for methane at room temperature in rigid materials, up
+to 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}.
 
@@
 \end{figure}
-
-\corysays{Figs.~\ref{fig:hydrogen-298-D} ~\ref{fig:methane-298-D} why not just let $\gst$ be negative and have an arrow pointing in direction of more attraction?
-as written the legend is not correct since those are not the functions being shown.
-in the main text we never discussed what \gst is for the MOF and why the MOFs are lines not points (\gst is a function of density now)
-\jpsays{due to a different deliverable capacity at full pressure and empty pressure.
-For hydrogen the SI (by Paula Garcia-Holley) has total uptake plots so we could make a curve even but would have a lot of error due to manual extraction?}}
 

### 2020-05-13 gas-adsorption/gas-ads-body.tex
starting on the conclusion section.

@@
 
-Storage of hydrogen is considerably more challenging owing to its relatively weak interaction with adsorbents.
-The DOE ULTIMATE deliverable capacity target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C.
-Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve for hydrogen storage along with a range of experimental measurements for known MOFs.
-The DOE ULTIMATE deliverable capacity target is theoretically possible, however by such a small margin that we can safely rule out the possibility of reaching this target through storage and delivery of hydrogen in \emph{any} rigid substrate at room temperature.
-Such a material would require a void fraction of at least 94\%.
-On top of this, the DOE ULTIMATE target requires an optimal $|\gst|$ of 10~kJ/mol which is far greater than what is found in observed porous materials.
-This reflects the known fact that hydrogen interacts with substrates far more weakly than methane does.
-One possibility that has been pursued is storage at cryogenic temperatures.
-See Section~\ref{sec:cryo-hydrogen} in the Supplementary Information for the upper bound for storage of hydrogen at 77 K.
+Storage of hydrogen is considerably more challenging owing to its relatively
+weak interaction with adsorbents. The DOE ULTIMATE deliverable capacity
+target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C.
+Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve for
+hydrogen storage along with a range of experimental measurements for known
+MOFs. The DOE ULTIMATE deliverable capacity target is theoretically possible,
+however by such a small margin that we can safely rule out the possibility of
+reaching this target through storage and delivery of hydrogen in \emph{any}
+rigid substrate at room temperature. Such a material would require a void
+fraction of at least 94\%. On top of this, the DOE ULTIMATE target requires an
+optimal $|\gst|$ of 10~kJ/mol which is far greater than what is found in
+observed porous materials. This reflects the known fact that hydrogen interacts
+with substrates far more weakly than methane does. One possibility that has
+been pursued is storage at cryogenic temperatures. See
+Section~\ref{sec:cryo-hydrogen} in the SI for the upper bound for hydrogen
+storage at 77 K.
 
-Why do we not experimentally observe materials approaching the upper bound on the deliverable capacity?
-Foremost, any adsorbent substrate will be composed of atoms, which will exclude the gas from some volume.
-Due to strong short-range interactions, substrate atoms must be uniformly distributed to achieve strong attraction for gas atoms.
-Uniform atom distribution sets a limit on the pore volume.
-% Because strong interactions are short-range, achieving a strong attraction requires that substrate atoms be distributed throughout the volume, which places a limit on the pore volume.
-In contrast, real materials have a spatially non-uniform attraction for gas atoms.
-The result of this is that there are regions of space with non-optimal attraction.
-For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.
-% Finally, real materials must have a spatially non-uniform attraction for gas atoms, resulting in some regions of space where the attraction is not optimal.
-For hydrogen, there is the further issue that there are \emph{no} known physical interactions that are sufficiently strong to give an optimal deliverable capacity at room temperature.
+Why do we not experimentally observe materials approaching the upper bound on
+the deliverable capacity? Foremost, any adsorbent substrate will be composed of
+atoms, which will exclude the gas from some volume. Due to strong short-range
+interactions, substrate atoms must be uniformly distributed to achieve strong
+attraction for gas atoms. Uniform atom distribution sets a limit on the pore
+volume. In contrast, real materials have a spatially non-uniform attraction for
+gas atoms. The result of this is that there are regions of space with
+non-optimal attraction. For hydrogen, there is the further issue that there are
+\emph{no} known physical interactions that are sufficiently strong to give an
+optimal deliverable capacity at room temperature.
 

### 2020-05-13 gas-adsorption/gas-ads-body.tex
fixed the conclusion.

@@
 \section{Conclusion}
-We have established an upper bound on the deliverable capacity via pressure-swing adsorption in rigid porous solids based on experimentally measured properties of pure gases.
-While these upper bounds do not rule out the discovery of materials that reach current DOE targets, they cast strong doubt on the possibility of achieving these goals when we consider the additional constraints imposed due to steric hindrance between substrate atoms and the adsorbate.
-Our upper bound does indicate that those goals cannot be exceeded by more than 16\% for methane and 6\% for hydrogen.
-Fortunately, there are some limitations to these upper bounds which suggest avenues for future developments which could allow the targets to be met.
+We have established an upper bound on the deliverable capacity via
+pressure-swing adsorption in rigid porous solids based on experimentally
+measured properties of pure gases. While these upper bounds do not rule out the
+discovery of materials that reach current DOE targets, they cast strong doubt
+on the possibility of achieving these goals when we consider the additional
+constraints imposed due to steric hindrance between substrate atoms and the
+adsorbate. Our upper bound does indicate that those goals cannot be exceeded by
+more than 16\% for methane and 6\% for hydrogen. Fortunately, there are some
+limitations to these upper bounds which suggest avenues for future developments
+which could allow the targets to be met.
 
-The first limitation of our proof is that we restricted ourselves to \emph{isothermal} pressure swing storage.
-By raising the temperature of the adsorbent during gas discharge to drive off residual gas~\cite{gomez2014exploring}, the deliverable capacity could be enhanced, albeit at the cost of a more complicated engineering design of the fuel tank and vehicle.
+The first limitation of our proof is that we restricted ourselves to
+\emph{isothermal} pressure swing storage. By raising the temperature of the
+adsorbent during gas discharge to drive off residual
+gas~\cite{gomez2014exploring}, the deliverable capacity could be enhanced,
+albeit at the cost of a more complicated engineering design of the fuel tank
+and vehicle.
 
-The second limitation of our proof arises in the assumption of a rigid substrate.
-The rigid substrate acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
-% The second limitation of our proof arises in the assumption of a rigid substrate, which acts as a static potential energy field for gas molecules that is unchanged by the adsorption of gas.
-For most porous materials this is a reasonable approximation, and this assumption is frequently made in both the simulation and theory of porous materials~\cite{witman2017influence}.
-However, there are cases where the substrate can in effect provide a very strongly gas-density-dependent interaction through structural flexibility~\cite{schneemann2014flexible}.
-A flagship example is MOF Co(bdp)~\cite{choi2008broadly}, which possesses a wine-rack-like topology capable of hinge motion.
-At low methane pressure, Co(bdp) adopts a collapsed, nonporous state, but expands to a porous state and fills with gas at higher pressures~\cite{mason2015methane}.
-This allows Co(bdp) to fully expel its residual gas at low pressures.
-Our upper bound does not consider such material effects and thus flexible materials could have significantly higher deliverable capacities.
-% Such materials can indeed violate our upper bound and thus could lead to significantly improved deliverable capacities.
+The second limitation of our proof arises in the assumption of a rigid
+substrate. The rigid substrate acts as a static potential energy field for gas
+molecules that is unchanged by the adsorption of gas. For most porous materials
+this is a reasonable approximation, and this assumption is frequently made in
+both the simulation and theory of porous materials~\cite{witman2017influence}.
+However, there are cases where the substrate can provide a very strong
+gas-density-dependent interaction through structural
+flexibility~\cite{schneemann2014flexible}. A flagship example is MOF
+Co(bdp)~\cite{choi2008broadly}, which possesses a wine-rack-like topology
+capable of hinge motion. At low methane pressure, Co(bdp) adopts a collapsed,
+nonporous state, but expands to a porous state and fills with gas at higher
+pressures~\cite{mason2015methane}. This allows Co(bdp) to fully expel its
+residual gas at low pressures. Our upper bound does not consider such material
+effects, and thus flexible materials could have significantly higher
+deliverable capacities.
 
-We note that our upper bound can readily be applied to the storage of other gasses of interest, with the proviso that the gas is far from crystallization. Our code is available at \davidsays{XXXXX}.
+We note that our upper bound can readily be applied to the storage of other
+gasses of interest, with the proviso that the gas is far from crystallization.
+Our code is available at \davidsays{XXXXX}.

### 2020-05-15 gas-adsorption/gas-ads-appendix.tex
working on appendix SI C.

--- 2020-05-05 gas-adsorption/gas-ads-appendix.tex
@@
 \section{An external, spatially uniform potential $\V$ shifts the chemical potential $\mu$ of the gas} \label{sec:V_shifts_chem_pot}
-We show that imposing an external, spatially uniform potential $\V$ to a gas has the effect of shifting the chemical potential $\mu$ of the gas, recovering eqn.~\ref{eq:mof-density}.
-Consider a control volume $\Omega$ with volume $V=|\Omega|$ (large enough to neglect boundary effects) that is endowed with the external, spatially uniform potential $\V$.
-Impose the grand-canonical ensemble, where this control volume can exchange energy and particles with a bath of gas at temperature $T$ and chemical potential $\mu$.
+We show that imposing an external, spatially uniform potential $\V$ to a gas
+has the effect of shifting the chemical potential $\mu$ of the gas, recovering
+eqn.~\ref{eq:mof-density}. Consider a control volume $\Omega$ with volume
+$V=|\Omega|$ (large enough to neglect boundary effects) that is endowed with
+the external, spatially uniform potential $\V$. Impose the grand-canonical
+ensemble, where this control volume can exchange energy and particles with a
+bath of gas at temperature $T$ and chemical potential $\mu$.
 
-To denote a microstate of this system, let $N$ be the number of gas particles in the control volume and $\mathbf{x}_1,...,\mathbf{x}_N$ be their positions.
+To denote a microstate of this system, let $N$ be the number of gas particles
+in the control volume and $\mathbf{x}_1,...,\mathbf{x}_N$ be their positions.
 Then, the potential energy $E$ of a microstate of the control volume is:
-\begin{equation}
-    E(\mathbf{x}_1,...,\mathbf{x}_N) = N\V + E_{gg}(\mathbf{x}_1,...,\mathbf{x}_N),
-\end{equation} where $E_{gg}$ is the (unknown and complicated) interatomic potential for gas-gas interactions that governs the (real) gas properties.
-The first term arises from each gas molecule feeling the external potential $\V$, where $\V < 0$ corresponds to attraction.
+\begin{equation} E(\mathbf{x}_1,...,\mathbf{x}_N) = N\V +
+E_{gg}(\mathbf{x}_1,...,\mathbf{x}_N), \end{equation} where $E_{gg}$ is the
+(unknown and complicated) interatomic potential for gas-gas interactions that
+governs the (real) gas properties. The first term arises from each gas molecule
+experiencing the external potential $\V$, where $\V < 0$ corresponds to
+attraction.
 
@@
     \label{eq:gcpf}
-\end{multline} We recognize this as the grand canonical partition function of the bulk gas in the control volume without the external potential, but shifted in chemical potential:
+\end{multline}
+We recognize this as the grand canonical partition function of the bulk gas in
+the control volume without the external potential, but shifted in chemical
+potential:
 \begin{equation}
@@
 \end{equation}
-where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the bulk gas in the control volume in the absence of an external potential.
-Importantly, this equivalency is predicated on the potential $\V$ being spatially uniform.
-Therefore, the thermodynamic properties of the gas sitting in the spatially uniform external potential $\V$-- i.e., adsorbed in our idealized substrate-- are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$ ($T$ fixed here).
-As $\V$ becomes more negative, corresponding to a more attractive adsorbent, the thermodynamic properties of the adsorbed gas in our ideal substrate are equivalent to the gas at higher chemical potential.
+where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the bulk
+gas in the control volume in the absence of an external potential. Importantly,
+this equivalency depends on the potential $\V$ being spatially uniform.
+Therefore, the thermodynamic properties of the gas atoms in the spatially
+uniform external potential $\V$-- i.e., adsorbed in our idealized substrate--
+are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$
+(where $T$ is held fixed). As $\V$ becomes more negative, corresponding to a
+more attractive adsorbent, the thermodynamic properties of the adsorbed gas in
+our ideal substrate are equivalent to the gas at higher chemical potential.
 
-\section{Proof of extremum}
-To show that a uniform potential gives the highest deliverable capacity, we consider a potential of interaction between gas and substrate $\V(\vec r)$ that varies with space.
-In this proof we will make use of the Fourier transform of this potential:
+\section{Proof of extremum}\label{sec:proof-extremum}
+To show that a uniform potential gives the highest deliverable capacity, we
+consider a interaction potential between gas and substrate $\V(\vec r)$ that
+varies with space.  In this proof we will make use of the Fourier transform of
+this potential:
 \begin{align}
@@
 \end{align}
-The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$.
-Therefore, to show that a uniform potential extremizes the deliverable capacity, we must show that the functional derivative of the deliverable capacity with respect to $\Vk$ is zero for \emph{nonzero} values of $\vec k$, i.e.
+The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$. Therefore, to show that a uniform potential
+extremizes the deliverable capacity, we must show that the functional
+derivative of the deliverable capacity with respect to $\Vk$ is zero for
+\emph{nonzero} values of $\vec k$, i.e.
 \begin{align}
@@
 \end{align}
-We note that this functional derivative may be non-zero for $\vec k=0$ because we separately maximize with respect to the particular uniform potential $\V$.
+We note that this functional derivative may be non-zero for $\vec k=0$ because
+we separately maximize with respect to the particular uniform potential $\V$.
 This means that
@@
 
-Because the chemical potential $\mu$ varies monotonically with $N$ at fixed temperature, we can consider how the chemical potential varies as we change $\Vk$ with the number of molecules held fixed.
-We demonstrate this using the cyclic chain rule, which shows us that
+Because the chemical potential $\mu$ varies monotonically with $N$ at fixed
+temperature, we can consider how the chemical potential varies as we change
+$\Vk$ with the number of molecules held fixed. We demonstrate this using the
+cyclic chain rule, which shows us that
 \begin{align}
@@
 \end{align}
-Thus the functional derivative of both the chemical potential and $N$ with regard to $\V(\vec r)$ are themselves spatially uniform.
+Thus, the functional derivative of both $\mu$ and $N$ with regard to $\V(\vec r)$ are themselves spatially uniform.
 Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e.
@@
 
-This demonstrates that a spatially uniform potential leads to an extremum value of the deliverable capacity.  This proof is insufficient, however, to show that it must be a true maximum.
+This demonstrates that a spatially uniform potential leads to an extremum value
+of the deliverable capacity. This proof is insufficient, however, to show that
+it must be a true maximum.
 

### 2020-05-15 gas-adsorption/gas-ads-appendix.tex
use bold face for all vectors

@@
 To denote a microstate of this system, let $N$ be the number of gas particles
-in the control volume and $\mathbf{x}_1,...,\mathbf{x}_N$ be their positions.
+in the control volume and $\mathbf{r}_1,...,\mathbf{r}_N$ be their positions.
 Then, the potential energy $E$ of a microstate of the control volume is:
-\begin{equation} E(\mathbf{x}_1,...,\mathbf{x}_N) = N\V +
-E_{gg}(\mathbf{x}_1,...,\mathbf{x}_N), \end{equation} where $E_{gg}$ is the
+\begin{equation} E(\rvec_1,...,\rvec_N) = N\V +
+E_{gg}(\rvec_1,...,\rvec_N), \end{equation} where $E_{gg}$ is the
 (unknown and complicated) interatomic potential for gas-gas interactions that
@@
 \begin{multline}
-    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\xvec_1, ..., \xvec_N)} e^{\beta (\mu - \V) N} d\xvec_1 \cdots d\xvec_N.
+    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\rvec_1, ..., \rvec_N)} e^{\beta (\mu - \V) N} d\rvec_1 \cdots d\rvec_N.
     \label{eq:gcpf}
@@
 To show that a uniform potential gives the highest deliverable capacity, we
-consider a interaction potential between gas and substrate $\V(\vec r)$ that
+consider a interaction potential between gas and substrate $\V(\rvec)$ that
 varies with space.  In this proof we will make use of the Fourier transform of
@@
 \begin{align}
-    \Vk \equiv \iiint \V(\vec r) e^{-i\vec k\cdot \vec r} d\vec r
+    \Vk \equiv \iiint \V(\rvec) e^{-i\kvec\cdot \rvec} d\rvec
 \end{align}
-The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\vec k)\propto\delta(\vec k)$. Therefore, to show that a uniform potential
+The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\kvec)\propto\delta(\kvec)$. Therefore, to show that a uniform potential
 extremizes the deliverable capacity, we must show that the functional
 derivative of the deliverable capacity with respect to $\Vk$ is zero for
-\emph{nonzero} values of $\vec k$, i.e.
+\emph{nonzero} values of $\kvec$, i.e.
 \begin{align}
-    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \vec k\ne 0.
+    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \kvec\ne 0.
 \end{align}
-We note that this functional derivative may be non-zero for $\vec k=0$ because
+We note that this functional derivative may be non-zero for $\kvec=0$ because
 we separately maximize with respect to the particular uniform potential $\V$.
@@
 \begin{align}
-    \frac{\delta N_H}{\delta \V(\vec k)} &= -\frac{\delta N_L}{\delta \Vk}
+    \frac{\delta N_H}{\delta \V(\kvec)} &= -\frac{\delta N_L}{\delta \Vk}
 \end{align}
@@
     \left(\frac{\delta N}{\delta \Vk}\right)_{\mu} &=
-    -\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}
+    -\left(\frac{\delta \mu}{\delta \V(\kvec)}\right)_{N}
     \left(\frac{\partial N}{\partial \mu}\right)_{\Vk}.
 \end{align}
-Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \V(\vec k)}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \V(\vec k)}\right)_{\mu}=0$.
+Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \V(\kvec)}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \V(\kvec)}\right)_{\mu}=0$.
 Thus we consider
 \begin{align}
-    \left(\frac{\delta \mu}{\delta \tilde\V(\vec k)}\right)_N
+    \left(\frac{\delta \mu}{\delta \tilde\V(\kvec)}\right)_N
     &= \left(\frac{\delta \left(\frac{\partial F}{\partial N}\right)_{\volume}}{\delta \Vk}\right)_N
@@
 \begin{align}
-    \frac{\delta F}{\delta \Vk} &= \rho(\vec k)
+    \frac{\delta F}{\delta \Vk} &= \rho(\kvec)
 \end{align}
@@
 does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.
-Thus the functional derivative $\frac{\delta F}{\delta \V(\vec r)}$ is actually a uniform function.
+Thus the functional derivative $\frac{\delta F}{\delta \V(\rvec)}$ is actually a uniform function.
 We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
 \begin{align}
-    \frac{\delta \mu}{\delta \Vk} &\propto \delta(\vec k) \\
-    \frac{\delta N}{\delta \Vk} &\propto \delta(\vec k)
+    \frac{\delta \mu}{\delta \Vk} &\propto \delta(\kvec) \\
+    \frac{\delta N}{\delta \Vk} &\propto \delta(\kvec)
 \end{align}
-Thus, the functional derivative of both $\mu$ and $N$ with regard to $\V(\vec r)$ are themselves spatially uniform.
+Thus, the functional derivative of both $\mu$ and $N$ with regard to $\V(\rvec)$ are themselves spatially uniform.
 Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e.
-$\vec k=0$), the derivative of $D$ with respect to any change of potential is zero.
+$\kvec=0$), the derivative of $D$ with respect to any change of potential is zero.
 

### 2020-05-15 gas-adsorption/gas-ads-appendix.tex
lots of edits during our meeting

@@
 E_{gg}(\rvec_1,...,\rvec_N), \end{equation} where $E_{gg}$ is the
-(unknown and complicated) interatomic potential for gas-gas interactions that
-governs the (real) gas properties. The first term arises from each gas molecule
-experiencing the external potential $\V$, where $\V < 0$ corresponds to
-attraction.
+(unknown and complicated) interatomic potential for gas-gas
+interactions that governs the (real) gas properties. The first term
+arises from each gas molecule experiencing the external potential
+$\V$, where $\V < 0$ corresponds to attraction.  To account
+  for molecular rotational and vibrational degrees of freedom, we
+  could treat the positions $\rvec_i$ as the locations of atoms rather
+  than molecules.  In this case the gas-gas interactions will include
+  both \emph{intra-molecular} interactions and \emph{inter-molecular}
+  interactions.
 
-\corysays{wut about orientations? shld we include those?}
-
-The grand-canonical partition function of the control volume is:
+The grand partition function of the control volume is:
 \begin{multline}
@@
 \end{equation}
-where $\Xi_0(\mu, V, T)$ is the grand-canonical partition function of the bulk
+where $\Xi_0(\mu, V, T)$ is the grand partition function of the bulk
 gas in the control volume in the absence of an external potential. Importantly,
@@
 
-\section{The real-substrate analogy of $\V$ is $\gst$}
+\section{The real-substrate analog of $\V$ is $\gst$}
 \label{sec:phi-is-delta-g}
@@
     \includegraphics[width=\columnwidth]{g_st_piston_2.png}
-    \caption{Cartoon illustrating the definition of $\gst$ and a hypothetical experiment to measure it with a piston and removable barrier that separates two equal volumes, one with a porous material (bottom) and one with free space (top).
-The initial state (left) is a bulk gas phase at density $\rho$ and temperature $T$.
-An infinitely-thin, removable barrier separates the gas from the porous material (bottom).
-The barrier is then slowly removed, and the piston slowly pushes all gas molecules into the porous material in an isothermal process.
-The final state (right) is an adsorbed gas phase at density $\rho$ and temperature $T$.
-The quantity $\gst$ follows from {\color{red} the heat released?}.}
+    \caption{Cartoon illustrating the definition of $\gst$ and a
+      hypothetical experiment to measure it with a piston and
+      removable barrier that separates two equal volumes, one with a
+      porous material (bottom) and one with free space (top). The
+      initial state (left) is a bulk gas phase at density $\rho$ and
+      temperature $T$. An infinitely-thin, removable barrier separates
+      the gas from the porous material (bottom). The barrier is then
+      slowly removed, and the piston slowly pushes all gas molecules
+      into the porous material in an isothermal process. The final
+      state (right) is an adsorbed gas phase at density $\rho$ and
+      temperature $T$. The quantity $\gst$ follows from \davidsays{the heat released?}.}
     \label{fig:delta-G-cartoon}
@@
 Therefore, the parameter analogous to $\V$ in a real substrate will involve both energy and entropy.
-The real-substrate analogy of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
+The real-substrate analog of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
 \begin{equation}
@@
 The quantity $\gst$ does \emph{not} correspond to a change in the molar Gibbs free energy as a molecule is adsorbed, which is zero under conditions of coexistence.
-The quantity $\gst$ in a real substrate is a direct analogy to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
+The quantity $\gst$ in a real substrate is a direct analog to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
 Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to  measure $\gst$ via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
@@
 The isosteric Gibbs free energy of adsorption, $\gst$, then follows from \red{is it the amount of heat released?}.
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} looks too jumbled. two separate figs? plus $G_{st}$ is capitalized here, should be lowercase. where is all this data from}
 
@@
 
-\corysays{can u mathematically justify why $\gst$ is the relevant parameter through the grand canonical partition function in eqn.~\ref{eq:gcpf}?
-it is not true that $\Xi(\mu, V, T)$ for the real substrate is equal to $\Xi(\mu+\gst, V, T)$}
-
-\corysays{Fig.~\ref{fig:qst-vs-delta-G} should be split in two. one for methane one for h2.}
-
 \begin{figure}
@@
     \includegraphics[width=0.95\columnwidth]{qst-vs-delta-G}
-    \caption{Relationship between $\gst$ and the isosteric heat $q_{st}$ for several prominent adsorbents.
-The dots represent the properties of methane adsorption at 298\ K and at 5\ bar and 100\ bar.
-The $\times$'s represent the properties of hydrogen adsorption at 77\ K and at 5~bar and 100~bar.}
+    \caption{Relationship between $\gst$ and the isosteric heat
+      $q_{st}$ for several prominent adsorbents at room
+      temperature. The dots represent the properties of methane
+      adsorption at 5~bar and 65~bar. The $+$'s represent the
+      properties of hydrogen adsorption at 5~bar and
+      100~bar. \davidsays{Add citations.}}
     \label{fig:qst-vs-delta-G}
@@
     \includegraphics[width=0.95\columnwidth]{methane-298-gst}
-    \caption{The density-dependence of $\gst$ of methane in several adsorbents (298\ K). Note that $\gst$ is monotonic in $\rho$. Data from Refs.~\ref{mason2014evaluating,furukawa2009storage}.
+    \caption{The density-dependence of $\gst$ of methane in several
+      adsorbents (298\ K). Note that $\gst$ is monotonic in
+      $\rho$. Data from
+      Refs.~\cite{mason2014evaluating, furukawa2009storage}.
     }
@@
     \includegraphics[width=0.95\columnwidth]{four-cases}
-    \caption{Four cases we consider.}
+    \caption{Four cases we consider. \davidsays{MAKE THIS FIGURE}}
     \label{fig:delta-gst-maximum}
@@
     \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
+Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-77-D}
@@
 Actual MOFs fall far short of the theoretical maximum.
-\red{the target is not just a storage density, it specifies temps and pressures too...
-is 77 K allowed?} \davidsays{My recollection of my reading of the DOE target was that anything was allowed, but that the target included ``everything'', which would include both the tank and any cryogenic equipment required.
-So the answer is not so simple.}

### 2020-05-15 gas-adsorption/gas-ads-body.tex
lots of edits during our meeting

@@
     \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
+Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-298-D}

### 2020-05-15 gas-adsorption/colors.py

@@ -43,6 +43,12 @@ def color(mof):
         return colors[the_order.index(mof)]
     return 'k'
 
+def symbol(gas):
+    if gas == 'methane':
+        return '.'
+    else:
+        return '+' 
+
 latex_mof_names = {
     'Ni2dobdc': 'Ni$_2$(dodbc)',
     'Co2dobdc': 'Co$_2$(dodbc)',

### 2020-05-15 gas-adsorption/gibbs-qst.py

@@ -340,7 +340,7 @@ for mof in colors.order(CH4_298_E_Data): # For each MOF
     qst = CH4_298_E_Data[mof]
     plt.plot([delta_G_lo_p/(kJ/mol), delta_G_hi_p/(kJ/mol)],
              [qst/(kJ/mol), qst/(kJ/mol)],
-             '.-', label=colors.latex_mof(mof), color=colors.color(mof))
+             colors.symbol('methane')+'-', label=colors.latex_mof(mof), color=colors.color(mof))
     mofs_shown.add(colors.latex_mof(mof))
 
 
@@ -378,23 +378,23 @@ nh = 1/Vh
 mu_empty = my_interp(p_empty, ph, mu_h)
 mu_full = my_interp(p_full, pc, mu_c)
 
-for mof in colors.order(H2_77_Data): # For each MOF
-    rho_lo_p = H2_77_Data[mof]['rho_h']*H2_crystal_density[mof]
-    delta_G_lo_p = my_interp(rho_lo_p, nh, mu_h) - mu_empty
-    # convert units: mmol/gram*molecular_weight/crystal_density,
-    rho_hi_p = H2_77_Data[mof]['rho_c']*H2_crystal_density[mof]
-    delta_G_hi_p = my_interp(rho_hi_p, nc, mu_c) - mu_full
+# for mof in colors.order(H2_77_Data): # For each MOF
+#     rho_lo_p = H2_77_Data[mof]['rho_h']*H2_crystal_density[mof]
+#     delta_G_lo_p = my_interp(rho_lo_p, nh, mu_h) - mu_empty
+#     # convert units: mmol/gram*molecular_weight/crystal_density,
+#     rho_hi_p = H2_77_Data[mof]['rho_c']*H2_crystal_density[mof]
+#     delta_G_hi_p = my_interp(rho_hi_p, nc, mu_c) - mu_full
 
-    qst = H2_77_Data[mof]
-    if colors.latex_mof(mof) not in mofs_shown:
-        plt.plot([delta_G_lo_p/(kJ/mol), delta_G_hi_p/(kJ/mol)],
-                 [qst['qh']/(kJ/mol), qst['qc']/(kJ/mol)],
-                 'x-', label=colors.latex_mof(mof), color=colors.color(mof))
-        mofs_shown.add(colors.latex_mof(mof))
-    else:
-        plt.plot([delta_G_lo_p/(kJ/mol), delta_G_hi_p/(kJ/mol)],
[…]

### 2020-05-15 gas-adsorption/isothermal-ads-gas-plot.py

@@ -256,7 +256,7 @@ for mof in colors.order(mof_isotherms): # For each MOF
 
     plt.plot([delta_G_lo_p/(kJ/mol), delta_G_hi_p/(kJ/mol)],
             [(rho_hi_p-rho_lo_p)/density_units, (rho_hi_p-rho_lo_p)/density_units],
-            'x-', label=colors.latex_mof(mof), color=colors.color(mof))
+            colors.symbol(basename)+'-', label=colors.latex_mof(mof), color=colors.color(mof))
 
 stepby = None
 if 20 < ymax < 200:
@@ -294,7 +294,7 @@ for mof in mof_isotherms: # For each MOF
         p_MOF = data[1]*bar
         rho_MOF = data[2]*density_units
         mu_ax.plot(mu_from_p(p_MOF)/(kJ/mol),
-                   rho_MOF/density_units, '.-', color=colors.color(mof), label=colors.latex_mof(mof))
+                   rho_MOF/density_units, colors.symbol(basename)+'-', color=colors.color(mof), label=colors.latex_mof(mof))
     except OSError: # Raise an OSError if file does not exist?
         print(f'Isotherm curves: The file "{fname}" does not existg continuing...')
         continue

### 2020-05-16 gas-adsorption/gas-ads-appendix.tex
fixed the appendix; now to add the figure.

@@
       state (right) is an adsorbed gas phase at density $\rho$ and
-      temperature $T$. The quantity $\gst$ follows from \davidsays{the heat released?}.}
+      temperature $T$.}
     \label{fig:delta-G-cartoon}
@@
 
-The parameter describing our idealized substrate is $\V$, the spatially uniform potential felt by a gas molecule adsorbed in the idealized substrate. 
-A natural question is how this potential relates to the properties of real substrates. 
-The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the gas (see Sec.~\ref{sec:V_shifts_chem_pot}).
-Because our ideal substrate shifts the chemical potential of the gas molecules by providing a spatially uniform potential energy field, the entropy of the gas in the ideal substrate is equal to the entropy of the gas in its bulk state at the same density and temperature.
-In contrast, a real substrate provides a non-spatially uniform potential.
-Consequently, the entropy of the gas inside a real substrate is \emph{not} equal to the entropy of the bulk gas at the same density and temperature.
-Therefore, the parameter analogous to $\V$ in a real substrate will involve both energy and entropy.
-The real-substrate analog of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
+The parameter describing our idealized substrate is $\V$, the spatially uniform
+potential felt by a gas molecule adsorbed in the idealized substrate. A natural
+question is how this potential relates to the properties of real substrates.
+The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the
+gas (see Sec.~\ref{sec:V_shifts_chem_pot}). Because our ideal substrate shifts
+the chemical potential of the gas molecules by providing a spatially uniform
+potential energy field, the entropy of the gas in the ideal substrate is equal
+to the entropy of the gas in its bulk state at the same density and
+temperature. In contrast, a real substrate provides a non-spatially uniform
+potential. Consequently, the entropy of the gas inside a real substrate is
+\emph{not} equal to the entropy of the bulk gas at the same density and
+temperature. Therefore, the parameter analogous to $\V$ in a real substrate
+will involve both energy and entropy. The real-substrate analog of $\V$ is the
+shift of molar Gibbs free energy provided by the substrate, specifically an
+\emph{isosteric} (or constant-density) shift of the Gibbs free energy:
 \begin{equation}
@@
 \end{equation}
-The isosteric Gibbs free energy difference $\gst$ is the difference in molar Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and the bulk gas \emph{with the same density of gas molecules}.
-The quantity $\gst$ does \emph{not} correspond to a change in the molar Gibbs free energy as a molecule is adsorbed, which is zero under conditions of coexistence.
-The quantity $\gst$ in a real substrate is a direct analog to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
-Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to  measure $\gst$ via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
-Note that $\gst$ is a property of both the substrate and the identity of the gas.
-Because real substrates offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho, T)=\V$.
-Consequently, throughout this article, we show $\gst(\rho, T)$ for real substrates at both conditions relevant to gas storage and delivery, $\pfull$ and $\pempty$.
-% In the initial state, the partition isolates the bulk gas exhibiting density $\rho$ and temperature $T$.
-Then, the partition is removed and the piston pushes down on the gas so that the gas slowly adsorbs into the MOF in an isothermal process.
-\red{isothermal rite?} In the final state, the piston has pushed all gas into the MOF, resulting in an adsorbed gas phase with the same density $\rho$ and temperature $T$ as the initial state.
-The isosteric Gibbs free energy of adsorption, $\gst$, then follows from \red{is it the amount of heat released?}.
+The isosteric Gibbs free energy difference $\gst$ is the difference in molar
+Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and
+the bulk gas \emph{with the same density of gas molecules}. The quantity $\gst$
+does \emph{not} correspond to a change in the molar Gibbs free energy as a
+molecule is adsorbed, which is zero under conditions of coexistence. The
+quantity $\gst$ in a real substrate is a direct analog to $\V$ in our ideal
+substrate because it is the chemical potential shift needed to impose on the
+bulk gas in coexistence with the real substrate to achieve the same density as
+in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
+Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to
+measure $\gst$ via a piston with a removable partition that separates a volume
+of free space from the same volume of substrate. Note that $\gst$ is a property
+of both the substrate and the identity of the gas. Because real substrates
+offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function
+of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho,
+T)=\V$. Consequently, throughout this article, we show $\gst(\rho, T)$ for real
+substrates at both conditions relevant to gas storage and delivery, $\pfull$
+and $\pempty$.
 
-In practice, we can readily compute $\gst(\rho, T)$ of a real gas/substrate system from (i) the (experimental or simulated) equilibrium adsorption isotherm of the gas in the substrate and (ii) the (experimental or simulated) chemical potential of the bulk gas.
-Consider the real substrate in thermodynamic equilibrium with a bulk gas at fixed temperature $T$ and pressure $p$, and let $\rho=\rho(p, T)$ be the density of gas in the substrate. 
-At coexistence, the chemical potential of the bulk gas is equal to the chemical potential of the adsorbed gas in the substrate.
-Thus, we can use the experimentally known molar Gibbs free energy of the pure, bulk gas system at temperature $T$ and pressure $p$ to determine the molar Gibbs free energy of the adsorbed system: $\mu_{\text{ads}}(\rho, T)=\mu_{\text{gas}}(p, T)$.
-We can then also look up the known chemical potential of the bulk gas at the same density and temperature as in the substrate, $\mu_{\text{gas}}(\rho, T)$.
-Via eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.
+In practice, we can readily compute $\gst(\rho, T)$ of a real gas/substrate
+system from (i) the (experimental or simulated) equilibrium adsorption isotherm
+of the gas in the substrate and (ii) the (experimental or simulated) chemical
+potential of the bulk gas. Consider the real substrate in thermodynamic
+equilibrium with a bulk gas at fixed temperature $T$ and pressure $p$, and let
+$\rho=\rho(p, T)$ be the density of gas in the substrate. At coexistence, the
+chemical potential of the bulk gas is equal to the chemical potential of the
+adsorbed gas in the substrate. Thus, we can use the experimentally known molar
+Gibbs free energy of the pure, bulk gas system at temperature $T$ and pressure
+$p$ to determine the molar Gibbs free energy of the adsorbed system:
+$\mu_{\text{ads}}(\rho, T)=\mu_{\text{gas}}(p, T)$. We can then also look up
+the known chemical potential of the bulk gas at the same density and
+temperature as in the substrate, $\mu_{\text{gas}}(\rho, T)$. Via
+eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.
 
-An interesting question is how $\gst$ relates to the commonly measured and reported isosteric heat of adsorption $q_{st}$, which is roughly the energy change when a gas molecule is adsorbed~\cite{sircar1999isosteric, tian2017differential}.
-Figure~\ref{fig:qst-vs-delta-G} shows how $q_{st}$ compares to $\gst$ for several prominent adsorbents.
-In every case, $|q_{st}|>|\gst|$ because the gas in the adsorbent always has less entropy than the gas in the bulk at the same density and temperature.
-That is, while adsorption is energetically favored, it is entropically disfavored due to the restrictions imposed on the configuration of the gas molecules via steric interactions with the substrate itself; this counters the energetic attraction.
+An interesting question is how $\gst$ relates to the commonly measured and
+reported isosteric heat of adsorption $q_{st}$, which is roughly the energy
+change when a gas molecule is adsorbed~\cite{sircar1999isosteric,
+tian2017differential}. Figure~\ref{fig:qst-vs-delta-G} shows how $q_{st}$
+compares to $\gst$ for several prominent adsorbents. In every case,
+$|q_{st}|>|\gst|$ because the gas in the adsorbent always has less entropy than
+the gas in the bulk at the same density and temperature. That is, while
+adsorption is energetically favored, it is entropically disfavored due to the
+restrictions imposed on the configuration of the gas molecules via steric
+interactions with the substrate itself; this counters the energetic attraction.
 
@@
       $q_{st}$ for several prominent adsorbents at room
-      temperature. The dots represent the properties of methane
-      adsorption at 5~bar and 65~bar. The $+$'s represent the
+      temperature (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}). The dots represent the properties of methane
+      adsorption at 5.8~bar and 65~bar. The $+$'s represent the
       properties of hydrogen adsorption at 5~bar and
-      100~bar. \davidsays{Add citations.}}
+      100~bar.}
     \label{fig:qst-vs-delta-G}
@@
 \section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
-Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding to a full and empty density.
-The deliverable capacity of the MOF is equal to the difference between the full and empty density.
-Examining Fig.~\ref{fig:methane-gst}, we see that a significant variety of known experimental $\Delta g_\text{st}$ curves are monotonic.
-Furthermore, our qualitative argument in Section~\ref{sec:upper-bound} suggests that this function \emph{should} be monotonically increasing for rigid substrates, as increasing the density of gas forces some of the gas to reside in higher-energy sites.
-If this is always the case, the deliverable capacity will always be bounded by our theoretical upper bound, even for a value of $\V$ that is not optimal.
+Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding
+to a full and empty density. The deliverable capacity of the MOF is equal to
+the difference between the full and empty density. Examining
+Fig.~\ref{fig:methane-gst}, we see that a significant variety of known
+experimental $\Delta g_\text{st}$ curves are monotonic. Furthermore, our
+qualitative argument in Section~\ref{sec:upper-bound} suggests that this
+function \emph{should} monotonically increase for rigid substrates, as
+increasing the density of gas causes some of the gas to reside in higher-energy
+sites. If this is always the case, the deliverable capacity will always be
+bounded by our theoretical upper bound, even for a value of $\V$ that is not
+optimal.
 
-To show this, we perform a thought experiment illustrated in Fig.~\ref{fig:delta-gst-maximum}.
-We begin with two pairs of boxes containing gas.
-Two of the boxes are filled with a MOF, and the other two have a uniform potential energy $\V$.
-We could imagine the two boxes at $\V$ as being empty but at very low altitude where gravity provides a potential energy difference (relative to the box with the MOF in it) of $\V$ to each gas molecule (perhaps we are on Jupiter?).
-One box of each sort has a density of gas $\rhofull$ corresponding to $\pfull$ in the MOF system, while the other two boxes has gas density $\rhoempty$.
+To show this, we perform a thought experiment illustrated in
+Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of boxes containing
+gas. Two of the boxes are filled with a MOF, and the other two have a uniform
+potential energy $\V$. We could imagine the two boxes at $\V$ as being empty
+but at very low altitude where gravity provides a potential energy difference
+(relative to the box with the MOF in it) of $\V$ to each gas molecule (perhaps
+we are on Jupiter?). One box of each sort has a density of gas $\rhofull$
+corresponding to $\pfull$ in the MOF system, while the other two boxes has gas
+density $\rhoempty$.
 
-We now consider what happens if we introduce a diffusive connection between two boxes that are initially at the same density, for instance by connecting them with a tube.
-If the chemical potential of the two boxes are equal, their density will be unchanged, otherwise gas will flow into the box with lower chemical potential.
-The difference in chemical potential is equal to the difference
+We now consider what happens if we introduce a diffusive connection between two
+boxes that are initially at the same density, for instance by connecting them
+with a tube. If the chemical potential of the two boxes are equal, their
+density will be unchanged, otherwise gas will flow into the box with lower
+chemical potential. The difference in chemical potential is equal to the
+difference
 \begin{align}
@@
 \end{align}
-Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at their initial density after they are connected.
-Thus the deliverable capacity of the MOF will be greater than the deliverable capacity of the model with uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e.
-if $\gst(\rho)$ does \emph{not} monotonically increase.
-By the same token, if we consider the case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable capacity than our model, the MOF must have less residual gas, which means that gas must spontaneously flow from the low-density MOF to the box with potential $\V$, which means that $\gst(\rhoempty)>\V$.
-So once again exceeding our upper bound requires a material with a $\gst(\rho)$ that does not increase monotonically.
+Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at
+their initial density after they are connected. Thus, the deliverable capacity
+of the MOF will be greater than the deliverable capacity of the model with
+uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$
+does \emph{not} monotonically increase. By the same token, if we consider the
+case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable
+capacity than our model, the MOF must have less residual gas, which means that
+gas must spontaneously flow from the low-density MOF to the box with potential
+$\V$, which means that $\gst(\rhoempty)>\V$. Once again, exceeding our upper
+bound requires a material with a $\gst(\rho)$ that does not increase
+monotonically.
 
-Taken together, this indicates that not only is our absolute upper bound an upper bound for rigid MOFs, but the green curve labeled $\rho_D(\gst)$ in Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and \ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal $\gst$.
+Taken together, this indicates that not only is our absolute upper bound an
+upper bound for rigid MOFs, but the green curve labeled $\rho_D(\gst)$ in
+Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and
+\ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal
+$\gst$.
 
@@
     \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
+Experimental deliverable capacities for several MOFs (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-77-D}
@@
 
-One approach to increase the deliverable capacity is to reduce the storage temperature.
-This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling point of nitrogen.
-The DOE ULTIMATE target in this case looks far more achievable, and with a much lower $|\gst|$.
-In fact, an empty tank at this temperature can satisfy the DOE 2020 target.
-The DOE ULTIMATE target is 14\% below the upper bound.
-Actual MOFs fall far short of the theoretical maximum.
+One approach to increase the deliverable capacity is to reduce the storage
+temperature. This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows
+the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling
+point of nitrogen. The DOE ULTIMATE target in this case looks far more
+achievable, and with a much lower $|\gst|$. In fact, an empty tank at this
+temperature can satisfy the DOE 2020 target. The DOE ULTIMATE target is 14\%
+below the upper bound. Actual MOFs fall far short of the theoretical maximum.

### 2020-05-16 gas-adsorption/gas-ads-body.tex
fixed the appendix; now to add the figure.

--- 2020-05-13 gas-adsorption/gas-ads-body.tex
@@
     \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
-Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
+Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-298-D}

### 2020-05-19 gas-adsorption/gas-ads-appendix.tex
added new figure with latex in inkscape.
made some final changes to the appendix.

@@
 Therefore, the thermodynamic properties of the gas atoms in the spatially
-uniform external potential $\V$-- i.e., adsorbed in our idealized substrate--
+uniform external potential $\V$ (adsorbed in our idealized substrate)
 are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$
@@
 more attractive adsorbent, the thermodynamic properties of the adsorbed gas in
-our ideal substrate are equivalent to the gas at higher chemical potential.
+our ideal substrate are equivalent to the gas at a higher chemical potential.
 
@@
 To show that a uniform potential gives the highest deliverable capacity, we
-consider a interaction potential between gas and substrate $\V(\rvec)$ that
-varies with space.  In this proof we will make use of the Fourier transform of
+consider an interaction potential between gas and substrate $\V(\rvec)$ that
+varies in space.  In this proof, we make use of the Fourier transform of
 this potential:
@@
 \end{align}
-The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\kvec)\propto\delta(\kvec)$. Therefore, to show that a uniform potential
-extremizes the deliverable capacity, we must show that the functional
+The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\kvec)\propto\delta(\kvec)$. In order for a uniform potential
+to extremize the deliverable capacity, we must show that the functional
 derivative of the deliverable capacity with respect to $\Vk$ is zero for
@@
 where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reordered the functional and partial derivatives.
-Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ just gives the number density:
+Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ yields the number density:
 \begin{align}
@@
 \end{align}
-The number density is itself spatially uniform for any system that is stable in a fluid state at this density (i.e.
+The number density is itself spatially uniform for any stable system in a fluid state at this density (i.e.
 does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.
-Thus the functional derivative $\frac{\delta F}{\delta \V(\rvec)}$ is actually a uniform function.
+Thus, the functional derivative $\frac{\delta F}{\delta \V(\rvec)}$ is actually a uniform function.
 We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
@@
 The isosteric Gibbs free energy difference $\gst$ is the difference in molar
-Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and
+Gibbs free energy (equivalent to chemical potential) between the adsorbed gas system and
 the bulk gas \emph{with the same density of gas molecules}. The quantity $\gst$
@@
     \centering
-    \includegraphics[width=0.95\columnwidth]{four-cases}
-    \caption{Four cases we consider. \davidsays{MAKE THIS FIGURE}}
+    \includegraphics[width=0.95\columnwidth]{four-cases-pro}
+    \caption{Two pairs of boxes containing gas with one of the pair being
+    filled with MOF. One box of each sort has a density of gas $\rhofull$
+    corresponding to $\pfull$ in the MOF system, while the other two boxes has
+    gas density $\rhoempty$}
     \label{fig:delta-gst-maximum}
@@
 gas. Two of the boxes are filled with a MOF, and the other two have a uniform
-potential energy $\V$. We could imagine the two boxes at $\V$ as being empty
+potential energy $\V$. We imagine the two boxes at $\V$ as being empty
 but at very low altitude where gravity provides a potential energy difference

### 2020-05-20 thesis/pretext.tex
updated the data and corrected the paper.
the pretext is correct.

@@
 	\large
-	AN ABSTRACT OF THE THESIS OF
+	AN ABSTRACT OF THE DISSERTATION OF
 \end{center}
@@
 \underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
-presented on April 30, 2020
+presented on June 8, 2020
 \vspace{1.0cm}
@@
 \underline{\hspace{11cm}}
-\vspace{0.5cm}
+\vspace{0.1cm}
 
@@
 $\copyright$ Copyright by Jordan K. Pommerenck \\
-April 30, 2020 \\
-(All Rights Reserved)
+June 8, 2020 \\
+All Rights Reserved
 \end{center}
@@
 \vspace{3.0cm}
-Presented April 30, 2020  \\
+Presented June 8, 2020  \\
 Commencement June 2020
@@
 \underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
-April 30, 2020. \\
+\underline{June 8, 2020.} \\
 

### 2020-05-22 gas-adsorption/gas-ads.tex
remove out of date comments and answer comments.
working through the paper and making edits and removing outdated comments.
use bold face for all vectors
made cory comment fixes for abstract.

@@
 % \usepackage{cite}
-\usepackage{xargs}                      % Use more than one optional parameter in a new commands
-\usepackage{color}  % Coloured text etc.
+\usepackage{xargs}
+% Use more than one optional parameter in a new commands
+\usepackage{color}
 %
@@
 \newcommand{\addcite}{{\bf \color{red} []}}
-\newcommand{\xvec}{\mathbf{x}}
+\newcommand{\rvec}{\mathbf{r}}
+\newcommand{\kvec}{\mathbf{k}}
 \newcommand{\thetavec}{\boldsymbol \theta}
@@
 \newcommand\V{\Phi}
-\newcommand\Vk{\tilde\Phi(\vec k)}
+\newcommand\Vk{\tilde\Phi(\kvec)}
 \newcommand\volume{V}
@@
 
-%\newcommandx{\davidsays}[1]{\begin{mdframed}[backgroundcolor=blue!24]\textbf{David says:} #1\end{mdframed} }
-%\newcommandx{\davidsays}[2][1=inline]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{David says:} #2} }
 \newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
@@
 \newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }
-
-%% \renewcommand\davidsays[1]{}
-%% \renewcommand{\corysays}[2][1=inline]{}
-%TC:macro \corysays [ignore]
-%TC:macro \davidsays [ignore]
-%TC:macro \jordan [ignore]
 
@@
 \title{
-An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}
-% other ideas:
-%Gas delivery via pressure-swing adsorption in nanoporous materials: setting storage targets}
-% A thermodynamic upper bound to gas delivery via pressure-swing adsorption in nanoporous materials:}
+An upper bound to gas storage and delivery via pressure-swing adsorption in
+porous materials}
 
@@
 Due to their low volumetric energy density at ambient conditions, both hydrogen
-and natural gas are challenging to economically store onboard environmentally
-sustainable vehicles~\cite{mason2014evaluating, sircar2002pressure}. One
-strategy to densify these gases is to pack the fuel tank with a porous
-material~\cite{?}. Metal-organic frameworks (MOFs) are tunable, nanoporous
-materials with large internal surface areas and show considerable promise for
-densifying gases~\cite{makal2012methane,mason2014evaluating,
+and natural gas are challenging to economically store onboard vehicles as
+fuels~\cite{mason2014evaluating, sircar2002pressure}. One strategy to densify
+these gases is to pack the fuel tank with a porous
+material~\cite{schoedel2016role}. Metal-organic frameworks (MOFs) are tunable,
+nanoporous materials with large internal surface areas and show considerable
+promise for densifying gases~\cite{makal2012methane,mason2014evaluating,
 suh2011hydrogen,garcia2018benchmark, schoedel2016role}. The US Department of
-Energy (DOE) has set volumetric deliverable capacity targets~\cite{simon2015materials, h2targetsDOE} which, if met, would help to enable 
-commercial adoption of hydrogen/natural gas as
-transportation fuels. Many have attempted to establish theoretical upper bounds
-on the deliverable capacity via pressure-swing adsorption using simplified
-models of the gas-substrate and gas-gas interaction~\cite{gomez2014exploring, 
-gomez2017impact, kaija2018high, lee2019predicting}; none have
-established a rigorous upper bound on the deliverable capacity.
-Here, we present a theoretical upper bound on the deliverable capacity of a
-rigid material via an isothermal pressure-swing. To provide an extremum, we
-consider a substrate that provides a spatially uniform potential energy field
-for the gas. Our bound is unique in that it directly relies on experimentally
-measured properties of the (real) bulk gas, without making approximations.
-We conclude that the goals set by the US DOE for natural gas and hydrogen
-storage are theoretically possible, but sufficiently close to the upper bound
-as to be impractical for any real porous material. However, limitations to the
-scope of applicability of our upper bound guide future material development 
-efforts. Firstly, one could heat the adsorbent to drive off residual gas when
-the pressure diminishes~\cite{gomez2014exploring}. Secondly, the fundamental
-physics of our upper bound do not pertain to any material that changes its
-structure in response to adsorbed gas, suggesting that flexible materials could
-still satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly, mason2015methane}.
+Energy (DOE) has set volumetric deliverable capacity
+targets~\cite{simon2015materials, h2targetsDOE} which, if met, would help
+enable commercial adoption of hydrogen/natural gas as transportation fuels.
+Many have attempted to establish theoretical upper bounds on the deliverable
+capacity via pressure-swing adsorption using simplified models of the
+gas-substrate and gas-gas interaction~\cite{gomez2014exploring,
+gomez2017impact, kaija2018high, lee2019predicting}; none have established a
+rigorous upper bound on the deliverable capacity. Here, we present a
+theoretical upper bound on the deliverable capacity of a rigid material via an
+isothermal pressure-swing. To provide an extremum, we consider a substrate that
+provides a spatially uniform potential energy field for the gas. Our bound is
+unique in that it directly relies on experimentally measured properties of the
+(real) bulk gas, without making approximations. We conclude that the goals set
+by the US DOE for natural gas and hydrogen storage are theoretically possible,
+but sufficiently close to the upper bound as to be impractical for any real
+porous material. However, limitations to the scope of applicability of our
+upper bound guide future material development efforts. Firstly, one could heat
+the adsorbent to drive off residual gas trapped in the
+adsorbent~\cite{gomez2014exploring}. Secondly, the fundamental physics of our
+upper bound do not pertain to any material that changes its structure in
+response to adsorbed gas, suggesting that flexible materials could still
+satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly,
+mason2015methane}.
 
-%   Here, we present a theoretical upper bound on the deliverable capacity of a rigid material via an isothermal pressure-swing.
-% "pressure-swing storage doesn't make sense to me.
-storage and delivery  maybe but awk.
-%   % I added rigid since that could be important...
-%   To provide an extremum, we consider a substrate that provides a spatially uniform potential energy field for the gas. 
-%   \davidsays{Maybe we can add a phrase for an optimal uniform attraction?}
-%   Our bound is unique in that it directly relies on experimentally measured properties of the (real) bulk gas, without making approximations.
-%   We conclude that the goals set by the US DOE for natural gas and hydrogen storage are theoretically possible, but sufficiently close to the upper bound as to be impractical for any real porous material.
-%   % However, this upper bound directly leads to important realizations which should guide future development.
-%   However, limitations to the scope of applicability of our upper bound guide future material development efforts.
-%  % Firstly, one could extract the gas at a higher temperature than that used while filling [].
-%   Firstly, one could heat the adsorbent to drive off residual gas when the pressure diminishes [].
-%   Secondly, the fundamental physics of our upper bound do not pertain to
-%   % rule out
-%   any material that changes its structure in response to adsorbed gas, suggesting that flexible materials could still satisfy the DOE target [].
-%   % firstly, secondly vs first, second?
 \end{abstract}
@@
 
-\paragraph*{to do}
-\begin{enumerate}
-    \item See david/corysays
-    \item make the code presentable
-    \item do the FIXME in code
-\end{enumerate}
-
 \input{gas-ads-body.tex}

### 2020-05-22 gas-adsorption/gas-ads-body.tex
made some more of corys edits and fixed MOF names in ledgend.

@@
 fuels~\cite{davis2009transportation}. In the United Sates, it is responsible
-for 36\% of energy-related carbon dioxide emissions \cite{useia} and 50,000
+for 36\% of energy-related carbon dioxide emissions \cite{useia} and $\sim$50,000
 premature deaths per year associated with particulate matter and ozone
@@
 temperature and high pressures. Both approaches require expensive
-infrastructure at refilling stations and significant energy input; e.g.,
-liquefaction of hydrogen consumes at least 30\% of its energy
+infrastructure at refilling stations and significant energy input; e.g., the
+energy input to liquify hydrogen is $\sim$30\% of its energy
 content~\cite{bossel2003energy}. Moreover, high-pressure storage tanks are
@@
 gas~\cite{makal2012methane,mason2014evaluating} and
-hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage is
-through physical adsorption in nanoporous materials at room
-temperature~\cite{schoedel2016role}. The internal surfaces of porous materials
+hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage at
+room temperature is through physical adsorption in nanoporous
+materials~\cite{schoedel2016role}. The internal surfaces of porous materials
 attract gas molecules through van der Waals, electrostatic, etc. interactions
@@
 temperature and pressure, allowing for room temperature and lower-pressure
-storage and alleviating many drawbacks of high-pressure storage.
+storage and alleviating many drawbacks of high-pressure, low-temperature
+storage.
 
@@
 minimum of -40 $^\circ$C~\cite{allendorf2018assessment}. Thus far, despite the
-emergence of highly tunable materials with enormous surface areas, such as
+emergence of highly tunable materials with large surface areas, such as
 metal-organic frameworks~\cite{furukawa2013chemistry}, no porous material has
@@
 residual gas will remain trapped in the adsorbent. Therefore, the driving range
-of the vehicle is primarily determined by the \emph{deliverable capacity}. The
-isothermal, volumetric deliverable capacity is an intrinsic property of the
-nanoporous material and its interaction with the gas.
+of the vehicle is primarily determined by the \emph{deliverable capacity} of
+the gas in the material (see Fig.~\ref{fig:delcap}): the density of gas at
+$\pfull$ minus that at $\pempty$. The isothermal, volumetric deliverable
+capacity is an intrinsic property of the nanoporous material and its
+interaction with the gas.
 
@@
 the space occupied by atoms of the porous material that are needed to endow the
-adsorption sites/volume with the attractive potential energy.
+adsorption sites/volume with the attractive potential energy and rely on a
+molecular model for methane.
 
-A third body of work attempted to account for both gas-gas interactions
-\emph{and} steric interactions of the gas with the atoms of the porous
-material. Simulations of methane adsorption in hundreds of thousands of
-explicit nanoporous crystal structures---both real and hypothetical---suggested
-that the ARPA-E deliverable capacity target is
-infeasible~\cite{simon2015materials}; the highest simulated methane deliverable
-capacity was 196 cm$^3$~STP/cm$^3$. Confidence in this conclusion rests upon
-(i) the accuracy of the intermolecular potentials describing the molecular
-interactions and (ii) the sufficient sampling of material space, i.e.\ that the
-structures considered are representative of the set of possible materials. To
-further explore material space and address sensitivity to the potentials used,
-scaling the Lennard-Jones potential well depths of material atoms to model
-enhanced interactions~\cite{gomez2014exploring}, placing Lennard-Jones spheres
-in a unit cell randomly to form ``pseudo-materials''~\cite{kaija2018high}, and
-generating fictitious potential energy fields via a generative adversarial
-network trained on zeolite structures~\cite{lee2019predicting} all generated
-model substrates that failed to meet the ARPA-E methane deliverable
-capacity target.
+A third body of work incorporated both gas-gas interactions \emph{and} steric
+interactions of the gas with the atoms of the porous material. Simulations of
+methane adsorption in hundreds of thousands of explicit nanoporous crystal
+structures---both real and hypothetical---suggested that the ARPA-E deliverable
+capacity target is infeasible~\cite{simon2015materials}; the highest simulated
+methane deliverable capacity was 196 cm$^3$~STP/cm$^3$. Confidence in this
+conclusion rests upon (i) the accuracy of the intermolecular potentials
+describing the molecular interactions and (ii) the sufficient sampling of
+material space, i.e.\ that the structures considered are representative of the
+set of possible materials. To further explore material space and address
+sensitivity to the intermolecular potentials, scaling the Lennard-Jones
+potential well depths of material atoms to model enhanced
+interactions~\cite{gomez2014exploring}, placing Lennard-Jones spheres in a unit
+cell randomly to form ``pseudo-materials''~\cite{kaija2018high}, and generating
+fictitious potential energy fields via a generative adversarial network trained
+on zeolite structures~\cite{lee2019predicting} all generated model substrates
+that failed to meet the ARPA-E methane deliverable capacity target.
 
@@
 density of the gas $\rho_g(\mu,T)$ as a function of chemical potential $\mu$
-and temperature $T$, which is shown for methane gas in
+and temperature $T$, is shown for methane gas in
 Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K. For comparison, we also show
@@
 two shifted bulk methane density curves and their difference. The horizontal
-axis is the difference in molar Gibbs free energy at fixed density and
-temperature between the gas within the substrate and the pure gas (\gst), which
+axis is the difference in molar Gibbs free energy, at fixed density and
+temperature, between the gas within the substrate and the pure gas (\gst), which
 is equal to $\V$ in our model. We see a potential $\V$ that maximizes the
@@
 static potential energy fields (a broader claim), consider a
-non-spatially-uniform variation on $\V_{opt}$. We will show that the effect of
-this variation is to reduce the deliverable capacity. At low adsorbed gas
+non-spatially-uniform variation on $\V_{opt}$. The effect of
+this variation is to reduce the deliverable capacity: At low adsorbed gas
 densities, the lowest-energy positions will be preferentially occupied, while
@@
 energy locations. Thus, the mean attraction will be lowest at \pfull\ and
-highest at \pempty. This will result in $\gst$ increasing monotonically with
+highest at \pempty. As a result, $\gst$ increases monotonically with
 pressure. As shown in the SI (see Sec.~\ref{sec:monotonic}), this results in
-the deliverable capacity of a non-uniform potential being lower than the
-deliverable capacity of a uniform potential with $\V=\gst$ for either $\pfull$
-or $\pempty$. Thus, a spatially uniform potential will give the best
+the deliverable capacity of a non-uniform potential with a given $\gst$ at $\pfull$ or $pempty$ being lower than the
+deliverable capacity of a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
 deliverable capacity amongst all static potential energy fields that give the
 same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will yield a
-greater deliverable capacity over any static non-uniform potential.
+greater deliverable capacity over any static, non-uniform potential.
 
@@
     \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
-    Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line.}
+    Experimental deliverable capacities for several porous materials (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line.}
     \label{fig:methane-298-D}
@@
 Although our theoretical framework allows us to place an upper bound on the
-deliverable capacity of many different gases in a rigid porous material, we
+deliverable capacity of many different gases in rigid porous materials, we
 focus on the maximal deliverable capacity of methane and hydrogen gas in our

### 2020-05-22 gas-adsorption/colors.py

@@ -54,8 +54,9 @@ latex_mof_names = {
     'Co2dobdc': 'Co$_2$(dodbc)',
     'Mg2dobdc':'Mg$_2$(dodbc)',
     'Zn2bdc2dabco': 'Zn$_2$(bdc)$_2$(dabco)',
-    'CYCU3Al': 'CYCU$_3$Al',
-    'UiO68Ant': 'UiO68Ant',
+    'CuMOF74': 'Cu-MOF-74',
+    'CYCU3Al': 'CYCU-3-Al',
+    'UiO68Ant': 'UiO-68-Ant',
     'rhtMOF7': 'rht-MOF-7',
 }
 def latex_mof(mof):

### 2020-05-22 gas-adsorption/gas-ads-body.tex
working on appendix right now.

@@
 room temperature~(374~cm$^3$~STP/cm$^3$). In addition to the predicted maximum
-deliverable capacity as a function of \gst, the ARPA-E target of
+deliverable capacity as a function of $\V$, the ARPA-E target of
 315~cm$^3$~STP/cm$^3$~\cite{arpaemove} is shown for context. For an adsorbent
@@
 (converted from the measured adsorption isotherm as explained in
-Sec.~\ref{sec:phi-is-delta-g}). These can be compared with the highest observed
-deliverable capacities for methane at room temperature in rigid materials, up
-to 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}.
+Sec.~\ref{sec:phi-is-delta-g}). For context, the highest observed
+deliverable capacity for methane at room temperature in a rigid material is 208~cm$^3$~STP/cm$^3$~\cite{simon2015materials}.
 
@@
 Storage of hydrogen is considerably more challenging owing to its relatively
-weak interaction with adsorbents. The DOE ULTIMATE deliverable capacity
-target~\cite{DOE} is within 6\% of the upper bound at 25$^\circ$C.
-Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve for
-hydrogen storage along with a range of experimental measurements for known
-MOFs. The DOE ULTIMATE deliverable capacity target is theoretically possible,
-however by such a small margin that we can safely rule out the possibility of
-reaching this target through storage and delivery of hydrogen in \emph{any}
-rigid substrate at room temperature. Such a material would require a void
-fraction of at least 94\%. On top of this, the DOE ULTIMATE target requires an
-optimal $|\gst|$ of 10~kJ/mol which is far greater than what is found in
-observed porous materials. This reflects the known fact that hydrogen interacts
-with substrates far more weakly than methane does. One possibility that has
-been pursued is storage at cryogenic temperatures. See
+weak interaction with adsorbents. For room temperature storage, the DOE
+ULTIMATE deliverable capacity target~\cite{DOE} is within 6\% of the upper
+bound. Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve
+for hydrogen storage along with experimental measurements for known MOFs. The
+DOE ULTIMATE deliverable capacity target is theoretically possible, however by
+such a small margin that we can safely rule out the possibility of reaching
+this target through storage and delivery of hydrogen in \emph{any} rigid
+substrate at room temperature. Such a material would require a void fraction of
+at least 94\%. On top of this, the DOE ULTIMATE target requires an optimal
+$|\gst|$ of 10~kJ/mol which is far greater than what is found in observed
+porous materials. This reflects the known fact that hydrogen interacts with
+substrates far more weakly than methane does. One possibility that has been
+pursued is storage at cryogenic temperatures. See
 Section~\ref{sec:cryo-hydrogen} in the SI for the upper bound for hydrogen
@@
 atoms, which will exclude the gas from some volume. Due to strong short-range
-interactions, substrate atoms must be uniformly distributed to achieve strong
-attraction for gas atoms. Uniform atom distribution sets a limit on the pore
+interactions, substrate atoms must be approximately uniformly distributed to achieve strong
+attraction for gas atoms, imposing a limit on the pore
 volume. In contrast, real materials have a spatially non-uniform attraction for
-gas atoms. The result of this is that there are regions of space with
+gas atoms. As a result, there are regions of space with
 non-optimal attraction. For hydrogen, there is the further issue that there are
-\emph{no} known physical interactions that are sufficiently strong to give an
+no known \emph{physical} interactions that are sufficiently strong to give an
 optimal deliverable capacity at room temperature.
@@
 more than 16\% for methane and 6\% for hydrogen. Fortunately, there are some
-limitations to these upper bounds which suggest avenues for future developments
-which could allow the targets to be met.
+limitations to these upper bounds which suggest avenues for future developments.
 
 The first limitation of our proof is that we restricted ourselves to
-\emph{isothermal} pressure swing storage. By raising the temperature of the
+\emph{isothermal} pressure swing storage (which is consistent with the targets set by DOE). By raising the temperature of the
 adsorbent during gas discharge to drive off residual

### 2020-05-22 gas-adsorption/gas-ads-appendix.tex
made two changes in appendix.
improve a few things in our meeting

@@
 \renewcommand{\thefigure}{S\arabic{figure}}%
-\renewcommand{\thesection}{SI \Alph{section}}%
+\renewcommand{\thesection}{SI~\Alph{section}}%
 
@@
 
-The grand partition function of the control volume is:
+The grand canonical partition function of the control volume is:
 \begin{multline}
@@
 \begin{align}
-    \frac{\delta N_H}{\delta \V(\kvec)} &= -\frac{\delta N_L}{\delta \Vk}
+    \frac{\delta N_H}{\delta \Vk} &= \frac{\delta N_L}{\delta \Vk}
 \end{align}
@@
     \left(\frac{\delta N}{\delta \Vk}\right)_{\mu} &=
-    -\left(\frac{\delta \mu}{\delta \V(\kvec)}\right)_{N}
+    -\left(\frac{\delta \mu}{\delta \Vk}\right)_{N}
     \left(\frac{\partial N}{\partial \mu}\right)_{\Vk}.
 \end{align}
-Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \V(\kvec)}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \V(\kvec)}\right)_{\mu}=0$.
+Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \Vk}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \Vk}\right)_{\mu}=0$.
 Thus we consider
 \begin{align}
-    \left(\frac{\delta \mu}{\delta \tilde\V(\kvec)}\right)_N
+    \left(\frac{\delta \mu}{\delta \Vk}\right)_N
     &= \left(\frac{\delta \left(\frac{\partial F}{\partial N}\right)_{\volume}}{\delta \Vk}\right)_N
@@
 \begin{align}
-    \frac{\delta \mu}{\delta \Vk} &\propto \delta(\kvec) \\
-    \frac{\delta N}{\delta \Vk} &\propto \delta(\kvec)
+    \left(\frac{\delta \mu}{\delta \Vk}\right)_N &\propto \delta(\kvec) \\
+    \left(\frac{\delta N}{\delta \Vk}\right)_\mu &\propto \delta(\kvec)
 \end{align}

### 2020-05-22 gas-adsorption/gas-ads-body.tex
improve a few things in our meeting

@@
 temperature, between the gas within the substrate and the pure gas (\gst), which
-is equal to $\V$ in our model. We see a potential $\V$ that maximizes the
+is equal to $\V$ in our model, and is pressure-dependent in a real
+crystal~(see Sec.~\ref{sec:phi-is-delta-g}). We see a potential $\V$ that maximizes the
 deliverable capacity of our ideal substrate, balancing the need to maximize the

### 2020-05-23 gas-adsorption/gas-ads-appendix.tex
made a number of changes but couldnt get line numbering working with figures and math.

@@
     \includegraphics[width=0.95\columnwidth]{four-cases-pro}
-    \caption{Two pairs of boxes containing gas with one of the pair being
+    \caption{Two pairs of boxes containing gas with one of the pairs
     filled with MOF. One box of each sort has a density of gas $\rhofull$
@@
 \section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
-Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding
+Every MOF has a $\Delta g_\text{st}$ at $\pfull$ and $\pempty$ corresponding
 to a full and empty density. The deliverable capacity of the MOF is equal to
@@
 \ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal
-$\gst$.
+$\gst$ at either $\pfull$ or $\pempty$.
 
@@
     \centering
-    FIGURE HERE A B C D
     \includegraphics[width=0.95\columnwidth]{hydrogen-77-n-vs-G}
-    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$.
+    \caption{Deliverable capacity of hydrogen at 77\ K as a function of the attractive Gibbs free energy $\gst$.
 Experimental deliverable capacities for several MOFs (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}

### 2020-05-25 gas-adsorption/gas-ads.tex
made a number of changes but couldnt get line numbering working with figures and math.
I have added in the hyperlink to github repo.

@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp]{revtex4-1}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,nature]{revtex4-1}
 \usepackage{graphicx} % Include figure files
@@
 
+\usepackage{lineno} % I added this to do line numbering
+
 \usepackage{nicefrac} % Include for inline fractions
 \usepackage{url}
+\usepackage[hypertexnames=false]{hyperref}
+\hypersetup{
+    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
+    pdftoolbar=true,        % show Acrobat’s toolbar?
+    pdfmenubar=true,        % show Acrobat’s menu?
+    pdffitwindow=false,     % window fit to page when opened
+    pdfstartview={FitH},    % fits the width of the page to the window
+    pdftitle={Pommerenck Thesis},    % title
+    pdfauthor={Jordan K. Pommerenck},     % author
+    pdfsubject={Subject},   % subject of the document
+    pdfcreator={Creator},   % creator of the document
+    pdfproducer={Producer}, % producer of the document
+    pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
+    pdfnewwindow=true,      % links in new PDF window
+    colorlinks=true,       % false: boxed links; true: colored links
+    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
+    citecolor=black,        % color of links to bibliography
+    filecolor=magenta,      % color of file links
+    urlcolor=cyan           % color of external links
+}
 

### 2020-05-26 gas-adsorption/gas-ads-appendix.tex
added line numbering successfully.

@@
 Then, the potential energy $E$ of a microstate of the control volume is:
+
 \begin{equation} E(\rvec_1,...,\rvec_N) = N\V +
-E_{gg}(\rvec_1,...,\rvec_N), \end{equation} where $E_{gg}$ is the
-(unknown and complicated) interatomic potential for gas-gas
-interactions that governs the (real) gas properties. The first term
-arises from each gas molecule experiencing the external potential
-$\V$, where $\V < 0$ corresponds to attraction.  To account
-  for molecular rotational and vibrational degrees of freedom, we
-  could treat the positions $\rvec_i$ as the locations of atoms rather
-  than molecules.  In this case the gas-gas interactions will include
-  both \emph{intra-molecular} interactions and \emph{inter-molecular}
-  interactions.
+E_{gg}(\rvec_1,...,\rvec_N),
+\end{equation}
+where $E_{gg}$ is the (unknown and complicated) interatomic potential for
+gas-gas interactions that governs the (real) gas properties. The first term
+arises from each gas molecule experiencing the external potential $\V$, where
+$\V < 0$ corresponds to attraction. To account for molecular rotational and
+vibrational degrees of freedom, we could treat the positions $\rvec_i$ as the
+locations of atoms rather than molecules. In this case the gas-gas interactions
+will include both \emph{intra-molecular} interactions and
+\emph{inter-molecular} interactions.
 
 The grand canonical partition function of the control volume is:
+
 \begin{multline}
@@
 potential:
+
 \begin{equation}
@@
 this potential:
+
 \begin{align}
@@
 \emph{nonzero} values of $\kvec$, i.e.
+
 \begin{align}
@@
 This means that
+
 \begin{align}
@@
 cyclic chain rule, which shows us that
+
 \begin{align}
@@
 Thus we consider
+
 \begin{align}
@@
 Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ yields the number density:
+
 \begin{align}
@@
 We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
+
 \begin{align}
@@
 \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
+
 \begin{equation}
@@
 difference
+
 \begin{align}

### 2020-05-27 dissertation/dissertation.tex
working on a new template for thesis.

\documentclass[double,12pt]{beavtex}
\usepackage{graphicx}
\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command

\renewcommand{\chaptername}{}

\title{Branching Random Walk and Probability Problems
from Physics and Biology}
\author{Torrey Johnson}
\degree{Doctor of Philosophy}
\doctype{Dissertation}
\department{Mathematics}
\depttype{Department}
\depthead{Chair}
\major{Mathematics}
\advisor{Edward C Waymire}
\submitdate{June 7, 2012}
\commencementyear{2013}
\abstract{\input{chapter-5.tex}}
\acknowledgements{\input{chapter-5.tex}}

\begin{document}

\maketitle

\mainmatter

\begingroup
\let\clearpage\relax
\begin{center}
\large
Branching Random Walk and Probability Problems
from Physics and Biology
\normalsize
\end{center}
\chapter{Introduction}
hello this is the chapter
%\input{introduction.tex}
\endgroup

% \chapter{Tree Polymers in the Infinite Volume Limit at Critical Strong Disorder}
% \input{csd.tex}

% \chapter{Simple Branching Random Walk and an Associated Tree Polymer}
% \input{brwtree.tex}

% \chapter{Branching for a Decision: A Site-Selection Model for Honeybee Swarms}
% \input{beetext.tex}

% \chapter{Conclusion}
% \input{conclusion.tex}

% \bibliographystyle{plain}
% \bibliography{bib}

% \appendix
% \chapter{Code for Swarm Simulation}
% \input{appendix.tex}
\end{document}

### 2020-05-27 dissertation/pretext.tex
working on a new template for thesis.

--- 2020-05-20 thesis/pretext.tex
@@
 \underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
-presented on June 8, 2020
+\nobreak presented on \underline{June 8, 2020}
 \vspace{1.0cm}
@@
 \justify{}
-Title: \underline{Advancing renewable gas storage using flat-histogram methods}.
+Title: \underline{Advancing Renewable Gas Storage using Flat-histogram Methods}.
 \vspace{2.0cm}
@@
 \begin{center}
-Advancing renewable gas storage using flat-histogram methods
+Advancing Renewable Gas Storage using Flat-histogram Methods
 
@@
 \begin{flushleft}
-Chair of the Department of Physics
+Head of the Department of Physics
 \vspace{1.0cm}

### 2020-05-27 dissertation/abstract.tex
added a new dissertation tex which conforms to OSU guidlines.

--- 2020-04-30 thesis/pretext.tex
@@
-%\thispagestyle{plain}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	AN ABSTRACT OF THE THESIS OF
-\end{center}
-
-\justify{}
-\underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
-presented on April 30, 2020
-\vspace{1.0cm}
-
-\justify{}
-Title: \underline{Advancing renewable gas storage using flat-histogram methods}.
-\vspace{2.0cm}
-
-\justify{}
-Abstract approved:
-\underline{\hspace{11cm}}
-\vspace{0.5cm}
-
-%\centering{}
-\hspace{7.0cm} David J. Roundy
-\vspace{1.0cm}
-
-\justify{}
-\doublespacing
-This work introduces the novel flat-histogram Monte
-Carlo (MC) method stochastic approximation with a dynamic update factor (SAD) and explores the convergence properties of a
-variety of related weight-based MC methods. The new method is applied
-to a number of physical `test’ systems including the 2D Ising model, a
-square-well fluid, and a Lennard-Jones cluster. A driving motivation for
-developing novel Monte Carlo methods that have physically based tunable
+This work introduces the novel flat-histogram Monte Carlo (MC) method
+stochastic approximation with a dynamic update factor (SAD) and explores the
+convergence properties of a variety of related weight-based MC methods. The new
+method is applied to a number of physical `test’ systems including the 2D Ising
+model, a square-well fluid, and a Lennard-Jones cluster. A driving motivation
+for developing novel Monte Carlo methods that have physically based tunable
 parameters is to provide simulation methods for exploring metal-organic
-frameworks (MOFs). A theoretical framework for gas adsorption and
-delivery in metal-organic frameworks is developed and compared with
-experimental data. SAD is introduced as a powerful method for examining
-thermodynamic properties for MOFs. A preliminary groundwork is laid for
-the future development of a multi-dimensional SAD which would be able
-to compute all properties of interest for any given MOF thereby providing a powerful way for researchers to determine any porous media's suitability for gas storage applications.
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-
-\vspace*{4.0cm}
-\begin{center}
-$\copyright$ Copyright by Jordan K. Pommerenck \\
-April 30, 2020 \\
-(All Rights Reserved)
-\end{center}
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-
-\begin{center}
-Advancing renewable gas storage using flat-histogram methods
-
-\vspace{1.0cm}
-by \\
-Jordan K. Pommerenck \\
-\vspace{3.0cm}
-A DISSERTATION \\
-\vspace{0.5cm}
-submitted to \\
-\vspace{0.5cm}
-Oregon State University \\
-\vspace{3.0cm}
-in partial fulfillment of \\
-the requirements for the \\
-degree of \\
-\vspace{1.0cm}
-Doctor of Philosophy \\
-\vspace{3.0cm}
-Presented April 30, 2020  \\
-Commencement June 2020
-\end{center}
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-\justify{}
-\underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
-April 30, 2020. \\
-
-\justify{}
-\vspace{0.5cm}
-APPROVED: \\
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Major Professor, representing Physics
-\vspace{1.0cm}
-\end{flushleft}
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Chair of the Department of Physics
-\vspace{1.0cm}
-\end{flushleft}
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Dean of the Graduate School
-\vspace{1.0cm}
-\end{flushleft}
-
-\vspace{3.0cm}
-\justify{}
-I understand that my dissertation will become part of the permanent collection
-of Oregon State University libraries. My signature below authorizes release of my dissertation to any reader upon request. \\
-\vspace{0.5cm}
-\begin{flushleft}
-\underline{\hspace{15cm}}
-\end{flushleft}
-\centering{}
-Jordan K. Pommerenck, Author
-
-\newpage{}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	ACKNOWLEDGEMENTS
-\end{center}
-\justify{}
-\doublespacing
-I gratefully acknowledge my graduate advisor David Roundy for his
-constant support and guidance throughout my research. His breadth of knowledge
-and hands-off approach to advising have tremendously helped my development as an independent scientific researcher.
-I express thanks to my entire graduate committee: David McIntyre, Davide Lazzati, Yun-Shik Lee, and graduate council representative Chih-hung (Alex) Chang.
-I would like to acknowledge the many outstanding undergraduate and graduate students that have positively impacted me personally and professionally while at Oregon State University.
-Finally, I would like to thank my parents for their unconditional support throughout my Ph.D.
-research.
-
-\newpage{}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	CONTRIBUTION OF AUTHORS
-\end{center}
-\justify{}
-\doublespacing
-Michael A.
-Perlin worked on much of the early C{}\verb!++!
-codebase.
-This codebase was used primarily in the early study of many of the MC methods before switching to RUST.
-In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
-Tanner T.
-Simpson performed many of the early simulations for the flat-histogram methods using the C{}\verb!++!
-codebase.
-He also developed using Python some of the plotting scripts that allowed the saved C{}\verb!++!
-data to be visualized.
-In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
-Cory M.
-Simon collaborated together with us on the third manuscript \emph{An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}.
-He contributed significantly to the introduction section of the paper and helped in the editing phase throughout preparation.
-In addition, he designed the first figure in the paper and the first figure in the supplemental to the paper.
+frameworks (MOFs). A theoretical framework for gas adsorption and delivery in
+metal-organic frameworks is developed and compared with experimental data. SAD
+is introduced as a powerful method for examining thermodynamic properties for
+MOFs. A preliminary groundwork is laid for the future development of a
+multi-dimensional SAD which would be able to compute all properties of interest
+for any given MOF thereby providing a powerful way for researchers to determine
+any porous media's suitability for gas storage applications.

### 2020-05-27 dissertation/acknowledgements.tex
added a new dissertation tex which conforms to OSU guidlines.

--- 2020-04-30 thesis/pretext.tex
@@
-%\thispagestyle{plain}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	AN ABSTRACT OF THE THESIS OF
-\end{center}
-
-\justify{}
-\underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
-presented on April 30, 2020
-\vspace{1.0cm}
-
-\justify{}
-Title: \underline{Advancing renewable gas storage using flat-histogram methods}.
-\vspace{2.0cm}
-
-\justify{}
-Abstract approved:
-\underline{\hspace{11cm}}
-\vspace{0.5cm}
-
-%\centering{}
-\hspace{7.0cm} David J. Roundy
-\vspace{1.0cm}
-
-\justify{}
-\doublespacing
-This work introduces the novel flat-histogram Monte
-Carlo (MC) method stochastic approximation with a dynamic update factor (SAD) and explores the convergence properties of a
-variety of related weight-based MC methods. The new method is applied
-to a number of physical `test’ systems including the 2D Ising model, a
-square-well fluid, and a Lennard-Jones cluster. A driving motivation for
-developing novel Monte Carlo methods that have physically based tunable
-parameters is to provide simulation methods for exploring metal-organic
-frameworks (MOFs). A theoretical framework for gas adsorption and
-delivery in metal-organic frameworks is developed and compared with
-experimental data. SAD is introduced as a powerful method for examining
-thermodynamic properties for MOFs. A preliminary groundwork is laid for
-the future development of a multi-dimensional SAD which would be able
-to compute all properties of interest for any given MOF thereby providing a powerful way for researchers to determine any porous media's suitability for gas storage applications.
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-
-\vspace*{4.0cm}
-\begin{center}
-$\copyright$ Copyright by Jordan K. Pommerenck \\
-April 30, 2020 \\
-(All Rights Reserved)
-\end{center}
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-
-\begin{center}
-Advancing renewable gas storage using flat-histogram methods
-
-\vspace{1.0cm}
-by \\
-Jordan K. Pommerenck \\
-\vspace{3.0cm}
-A DISSERTATION \\
-\vspace{0.5cm}
-submitted to \\
-\vspace{0.5cm}
-Oregon State University \\
-\vspace{3.0cm}
-in partial fulfillment of \\
-the requirements for the \\
-degree of \\
-\vspace{1.0cm}
-Doctor of Philosophy \\
-\vspace{3.0cm}
-Presented April 30, 2020  \\
-Commencement June 2020
-\end{center}
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-\justify{}
-\underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
-April 30, 2020. \\
-
-\justify{}
-\vspace{0.5cm}
-APPROVED: \\
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Major Professor, representing Physics
-\vspace{1.0cm}
-\end{flushleft}
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Chair of the Department of Physics
-\vspace{1.0cm}
-\end{flushleft}
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Dean of the Graduate School
-\vspace{1.0cm}
-\end{flushleft}
-
-\vspace{3.0cm}
-\justify{}
-I understand that my dissertation will become part of the permanent collection
-of Oregon State University libraries. My signature below authorizes release of my dissertation to any reader upon request. \\
-\vspace{0.5cm}
-\begin{flushleft}
-\underline{\hspace{15cm}}
-\end{flushleft}
-\centering{}
-Jordan K. Pommerenck, Author
-
-\newpage{}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	ACKNOWLEDGEMENTS
-\end{center}
-\justify{}
-\doublespacing
-I gratefully acknowledge my graduate advisor David Roundy for his
-constant support and guidance throughout my research. His breadth of knowledge
-and hands-off approach to advising have tremendously helped my development as an independent scientific researcher.
-I express thanks to my entire graduate committee: David McIntyre, Davide Lazzati, Yun-Shik Lee, and graduate council representative Chih-hung (Alex) Chang.
-I would like to acknowledge the many outstanding undergraduate and graduate students that have positively impacted me personally and professionally while at Oregon State University.
-Finally, I would like to thank my parents for their unconditional support throughout my Ph.D.
-research.
-
-\newpage{}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	CONTRIBUTION OF AUTHORS
-\end{center}
-\justify{}
-\doublespacing
-Michael A.
-Perlin worked on much of the early C{}\verb!++!
-codebase.
-This codebase was used primarily in the early study of many of the MC methods before switching to RUST.
-In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
-Tanner T.
-Simpson performed many of the early simulations for the flat-histogram methods using the C{}\verb!++!
-codebase.
-He also developed using Python some of the plotting scripts that allowed the saved C{}\verb!++!
-data to be visualized.
-In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
-Cory M.
-Simon collaborated together with us on the third manuscript \emph{An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}.
-He contributed significantly to the introduction section of the paper and helped in the editing phase throughout preparation.
-In addition, he designed the first figure in the paper and the first figure in the supplemental to the paper.
+I gratefully acknowledge my graduate advisor David Roundy for his constant
+support and guidance throughout my research. His breadth of knowledge and
+hands-off approach to advising have tremendously helped my development as an
+independent scientific researcher. I express thanks to my entire graduate
+committee: David McIntyre, Davide Lazzati, Yun-Shik Lee, and graduate council
+representative Chih-hung (Alex) Chang. I would like to acknowledge the many
+outstanding undergraduate and graduate students that have positively impacted
+me personally and professionally while at Oregon State University. Finally, I
+would like to thank my parents for their unconditional support throughout my
+Ph.D. research.

### 2020-05-27 dissertation/contributors.tex
added a new dissertation tex which conforms to OSU guidlines.

--- 2020-04-30 thesis/pretext.tex
@@
-%\thispagestyle{plain}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	AN ABSTRACT OF THE THESIS OF
-\end{center}
-
-\justify{}
-\underline{Jordan K. Pommerenck} for the degree of \underline{Doctor of Philosophy} in \underline{Physics}
-presented on April 30, 2020
-\vspace{1.0cm}
-
-\justify{}
-Title: \underline{Advancing renewable gas storage using flat-histogram methods}.
-\vspace{2.0cm}
-
-\justify{}
-Abstract approved:
-\underline{\hspace{11cm}}
-\vspace{0.5cm}
-
-%\centering{}
-\hspace{7.0cm} David J. Roundy
-\vspace{1.0cm}
-
-\justify{}
-\doublespacing
-This work introduces the novel flat-histogram Monte
-Carlo (MC) method stochastic approximation with a dynamic update factor (SAD) and explores the convergence properties of a
-variety of related weight-based MC methods. The new method is applied
-to a number of physical `test’ systems including the 2D Ising model, a
-square-well fluid, and a Lennard-Jones cluster. A driving motivation for
-developing novel Monte Carlo methods that have physically based tunable
-parameters is to provide simulation methods for exploring metal-organic
-frameworks (MOFs). A theoretical framework for gas adsorption and
-delivery in metal-organic frameworks is developed and compared with
-experimental data. SAD is introduced as a powerful method for examining
-thermodynamic properties for MOFs. A preliminary groundwork is laid for
-the future development of a multi-dimensional SAD which would be able
-to compute all properties of interest for any given MOF thereby providing a powerful way for researchers to determine any porous media's suitability for gas storage applications.
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-
-\vspace*{4.0cm}
-\begin{center}
-$\copyright$ Copyright by Jordan K. Pommerenck \\
-April 30, 2020 \\
-(All Rights Reserved)
-\end{center}
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-
-\begin{center}
-Advancing renewable gas storage using flat-histogram methods
-
-\vspace{1.0cm}
-by \\
-Jordan K. Pommerenck \\
-\vspace{3.0cm}
-A DISSERTATION \\
-\vspace{0.5cm}
-submitted to \\
-\vspace{0.5cm}
-Oregon State University \\
-\vspace{3.0cm}
-in partial fulfillment of \\
-the requirements for the \\
-degree of \\
-\vspace{1.0cm}
-Doctor of Philosophy \\
-\vspace{3.0cm}
-Presented April 30, 2020  \\
-Commencement June 2020
-\end{center}
-
-\newpage{}
-\thispagestyle{empty}
-\singlespacing
-\justify{}
-\underline{Doctor of Philosophy} dissertation of \underline{Jordan K. Pommerenck} presented on
-April 30, 2020. \\
-
-\justify{}
-\vspace{0.5cm}
-APPROVED: \\
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Major Professor, representing Physics
-\vspace{1.0cm}
-\end{flushleft}
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Chair of the Department of Physics
-\vspace{1.0cm}
-\end{flushleft}
-
-\justify{}
-\underline{\hspace{15cm}}
-\begin{flushleft}
-Dean of the Graduate School
-\vspace{1.0cm}
-\end{flushleft}
-
-\vspace{3.0cm}
-\justify{}
-I understand that my dissertation will become part of the permanent collection
-of Oregon State University libraries. My signature below authorizes release of my dissertation to any reader upon request. \\
-\vspace{0.5cm}
-\begin{flushleft}
-\underline{\hspace{15cm}}
-\end{flushleft}
-\centering{}
-Jordan K. Pommerenck, Author
-
-\newpage{}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	ACKNOWLEDGEMENTS
-\end{center}
-\justify{}
-\doublespacing
-I gratefully acknowledge my graduate advisor David Roundy for his
-constant support and guidance throughout my research. His breadth of knowledge
-and hands-off approach to advising have tremendously helped my development as an independent scientific researcher.
-I express thanks to my entire graduate committee: David McIntyre, Davide Lazzati, Yun-Shik Lee, and graduate council representative Chih-hung (Alex) Chang.
-I would like to acknowledge the many outstanding undergraduate and graduate students that have positively impacted me personally and professionally while at Oregon State University.
-Finally, I would like to thank my parents for their unconditional support throughout my Ph.D.
-research.
-
-\newpage{}
-\thispagestyle{empty}
-\begin{center}
-	\large
-	CONTRIBUTION OF AUTHORS
-\end{center}
-\justify{}
-\doublespacing
-Michael A.
-Perlin worked on much of the early C{}\verb!++!
-codebase.
-This codebase was used primarily in the early study of many of the MC methods before switching to RUST.
-In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
-Tanner T.
-Simpson performed many of the early simulations for the flat-histogram methods using the C{}\verb!++!
-codebase.
-He also developed using Python some of the plotting scripts that allowed the saved C{}\verb!++!
-data to be visualized.
-In addition, he helped in the editing phase during the preparation of the first manuscript \emph{Stochastic approximation Monte Carlo with a dynamic update factor}.
-Cory M.
-Simon collaborated together with us on the third manuscript \emph{An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials}.
-He contributed significantly to the introduction section of the paper and helped in the editing phase throughout preparation.
-In addition, he designed the first figure in the paper and the first figure in the supplemental to the paper.
+Michael A. Perlin worked on much of the early C{}\verb!++! codebase. This
+codebase was used primarily in the early study of many of the MC methods before
+switching to RUST. In addition, he helped in the editing phase during the
+preparation of the first manuscript \emph{Stochastic approximation Monte Carlo
+with a dynamic update factor}. Tanner T. Simpson performed many of the early
+simulations for the flat-histogram methods using the C{}\verb!++! codebase. He
+also developed using Python some of the plotting scripts that allowed the saved
+C{}\verb!++! data to be visualized. In addition, he helped in the editing phase
+during the preparation of the first manuscript \emph{Stochastic approximation
+Monte Carlo with a dynamic update factor}. Cory M. Simon collaborated together
+with us on the third manuscript \emph{An upper bound to gas delivery via
+pressure-swing adsorption in nanoporous materials}. He contributed
+significantly to the introduction section of the paper and helped in the
+editing phase throughout preparation. In addition, he designed the first figure
+in the paper and the first figure in the supplemental to the paper.

### 2020-05-27 dissertation/dissertation.tex
added a new dissertation tex which conforms to OSU guidlines.

@@
 \documentclass[double,12pt]{beavtex}
+
+\usepackage{fancyhdr}
+% link to section command
+\newcommand{\HR}[1]{\textbf{\hyperref[#1]{\ref{#1}}}}
+
+% for creating an index at end of document
+\usepackage{imakeidx}
+\makeindex[columns=1, title=Alphabetical Index]
+
+% indexing commands
+\newcommand{\bind}[1]{\textbf{#1}\index{#1}}
+\newcommand{\iind}[1]{\textit{#1}\index{#1}}
+\newcommand{\ind}[1]{#1\index{#1}}
+
+\usepackage[hypertexnames=false]{hyperref}
+\hypersetup{
+    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
+    pdftoolbar=true,        % show Acrobat’s toolbar?
+    pdfmenubar=true,        % show Acrobat’s menu?
+    pdffitwindow=false,     % window fit to page when opened
+    pdfstartview={FitH},    % fits the width of the page to the window
+    pdftitle={Pommerenck Thesis},    % title
+    pdfauthor={Jordan K. Pommerenck},     % author
+    pdfsubject={Subject},   % subject of the document
+    pdfcreator={Creator},   % creator of the document
+    pdfproducer={Producer}, % producer of the document
+    pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
+    pdfnewwindow=true,      % links in new PDF window
+    colorlinks=true,       % false: boxed links; true: colored links
+    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
+    citecolor=black,        % color of links to bibliography
+    filecolor=magenta,      % color of file links
+    urlcolor=cyan           % color of external links
+}
+%\usepackage{blindtext}
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage[document]{ragged2e}
+% differential and partials package
+\usepackage[thinc]{esdiff}
+\usepackage{physics}
+\newcommand{\D}[1]{\text{d}#1}
+% the nth superscript package
+\usepackage[super]{nth}
+
+\usepackage{indentfirst}
 \usepackage{graphicx}
-\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
-\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
+\graphicspath{{../square-well/}{../ising/}{../gas-adsorption/figs/}}
+
+\usepackage{caption}
+%\usepackage{subcaption}
+\usepackage{float}
+\usepackage{amsmath,amssymb}
+\usepackage{nicefrac}
+
+% Use more than one optional parameter in a new commands
+\usepackage{xargs}
+\usepackage[pdftex,dvipsnames]{xcolor}
+\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
+\usepackage{mdframed}
+\usepackage{braket}
+
+% gas adsorption commands
+\usepackage{subfig}
+\usepackage{url}
+%
+\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
+
+\newcommand{\addcite}{{\bf \color{red} []}}
+\newcommand{\rvec}{\mathbf{r}}
+\newcommand{\kvec}{\mathbf{k}}
+\newcommand{\thetavec}{\boldsymbol \theta}
+\newcommand\V{\Phi}
+\newcommand\Vk{\tilde\Phi(\kvec)}
+\newcommand\volume{V}
+\newcommand\pfull{\ensuremath{p_{\text{full}}}}
+\newcommand\pempty{\ensuremath{p_{\text{empty}}}}
+\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
+\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}
+\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
+\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}
+\newcommand\gst{\ensuremath{\Delta g_{st}}}
+
+\newcommandx{\jordan}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{\textbf{Jordan says:} #2} }
+\newcommandx{\corysays}[2][1=inline]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Cory says:} #2} }
+
+\newcommandx{\maybecut}[1]{\begin{mdframed}[backgroundcolor=cyan!24]\textbf{Maybe cut:} #1\end{mdframed} }
+
+\newtheorem{limitation}{Limitation}
 
@@
 
-\title{Branching Random Walk and Probability Problems
-from Physics and Biology}
-\author{Torrey Johnson}
+\title{Advancing Renewable Gas Storage using Flat-histogram Methods}
+\author{Jordan K. Pommerenck}
 \degree{Doctor of Philosophy}
 \doctype{Dissertation}
-\department{Mathematics}
+\department{Physics}
 \depttype{Department}
-\depthead{Chair}
-\major{Mathematics}
-\advisor{Edward C Waymire}
-\submitdate{June 7, 2012}
-\commencementyear{2013}
-\abstract{\input{chapter-5.tex}}
-\acknowledgements{\input{chapter-5.tex}}
+\depthead{Head}
+\major{Physics}
+\advisor{David Roundy}
+\submitdate{June 8, 2020}
+\commencementyear{2020}
+\abstract{\input{abstract.tex}}
+\acknowledgements{\input{acknowledgements.tex}}
+\contributors{\input{contributors.tex}}
 
@@
 \large
-Branching Random Walk and Probability Problems
-from Physics and Biology
+Advancing Renewable Gas Storage using Flat-histogram Methods
 \normalsize
@@
 \chapter{Introduction}
-hello this is the chapter
-%\input{introduction.tex}
+\input{introduction.tex}
 \endgroup
 
-% \chapter{Tree Polymers in the Infinite Volume Limit at Critical Strong Disorder}
-% \input{csd.tex}
+\chapter{Stochastic Approximation Monte Carlo with a Dynamic Update
+Factor}
+\vspace*{2.0cm}
 
-% \chapter{Simple Branching Random Walk and an Associated Tree Polymer}
-% \input{brwtree.tex}
+\textbf{Authors:} \\
+Jordan K. Pommerenck, Tanner T. Simpson, Michael A. Perlin,
+and David Roundy
 
-% \chapter{Branching for a Decision: A Site-Selection Model for Honeybee Swarms}
-% \input{beetext.tex}
+\vspace*{4.0cm}
 
-% \chapter{Conclusion}
-% \input{conclusion.tex}
+\textbf{Physical Review E} \\
+\href{https://doi.org/10.1103/PhysRevE.101.013301}{Journal Online Access} \\
+\textbf{101(1)}, 013301 (2020)~\cite{pommerenck2020stochastic}
 
-% \bibliographystyle{plain}
-% \bibliography{bib}
+\clearpage{}
 
-% \appendix
-% \chapter{Code for Swarm Simulation}
-% \input{appendix.tex}
+\input{../square-well/sq-well-body.tex}
+
+\chapter{Flat-histogram method comparison on 2D Ising Model
+}
+\input{../ising/ising-body.tex}
+
+\chapter{An upper bound to gas storage and delivery via pressure-swing
+adsorption in porous materials}
+\input{../gas-adsorption/gas-ads-body.tex}
+
+\chapter{Conclusion}
+\input{conclusion.tex}
+
+\bibliographystyle{plain}
+\bibliography{thesis}
+
+\appendix
+\input{../gas-adsorption/gas-ads-appendix.tex}
 \end{document}

### 2020-05-29 dissertation/introduction.tex
added a new dissertation tex which conforms to OSU guidlines.
made changes to thesis and cover letter.

--- 2020-05-06 thesis/chapter-1.tex
@@
-\chapter{Introduction}
-
 Monte Carlo methods are a powerful simulation tool that allow scientists to explore properties of systems that may otherwise by inaccessible via experimental methods.
@@
 
-In our works, we use the \textbf{Xoroshiro128+} RNG which is named after its logical operations (XOR $\xrightarrow{}$ rotate $\xrightarrow{}$ shift $\xrightarrow{}$ rotate).
+In our work, we use the \textbf{Xoroshiro128+} RNG which is named after its logical operations (XOR $\xrightarrow{}$ rotate $\xrightarrow{}$ shift $\xrightarrow{}$ rotate).
 The RNG has ultra-fast speed 0.72 $\nicefrac{\text{ns}}{\text{64-bits}}$ and incredible statistical quality~\cite{blackman2018scrambled}.
@@
 strategy to densify these gases is to pack the fuel tank with a porous
-material~\cite{?}. Metal-organic frameworks (MOFs) are tunable, nanoporous
+material~\cite{schoedel2016role}. Metal-organic frameworks (MOFs) are tunable, nanoporous
 materials with large internal surface areas and show considerable promise for

### 2020-05-29 gas-adsorption/gas-ads-body.tex
I have added in the hyperlink to github repo.
added line numbering successfully.
made the figure somewhat smaller.

@@
     \subfloat[][]{
-        \includegraphics[width=0.98\columnwidth]{usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
+        \includegraphics[width=0.8\columnwidth]{usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
     }
@@
 Consequently, the density of gas in our homogeneous substrate is:
+
 \begin{align}
@@
 The deliverable capacity of gas in our homogeneous substrate is thus:
+
 \begin{align}
@@
 gasses of interest, with the proviso that the gas is far from crystallization.
-Our code is available at \davidsays{XXXXX}.
+Our code is available at \href{https://github.com/Jordan-Pommerenck/isothermal-gas-adsorption}{https://github.com/Jordan-Pommerenck/isothermal-gas-adsorption}.

### 2020-05-30 dissertation/introduction.tex
made all changes in thesis after meeting and also changed fig in gas adsorption.

@@
-Monte Carlo methods are a powerful simulation tool that allow scientists to explore properties of systems that may otherwise by inaccessible via experimental methods.
-Monte Carlo methods can be used to solve a variety of physical problems yielding insight into statistical averages due to random processes.
-During the last two decades considerable research has been done on developing and refining flat-histogram Monte Carlo methods.
-These Monte Carlo (MC) methods are capable of exploring all of the energy phase space for all given temperatures in a \emph{single} simulation.
-In this work, we first outline the necessary statistical mechanics background.
-We provide a description of MC methods and flat-histogram methods, as well as, our newly developed method SAD (stochastic approximation Monte Carlo with a dynamic update factor).
-We discuss physics rich test systems that can be used to explore the convergence and statistical properties of our newly developed method SAD.
+Monte Carlo methods are a powerful simulation tool that allow scientists to
+explore properties of systems that may otherwise by inaccessible via
+experimental methods. Monte Carlo methods can be used to solve a variety of
+physical problems yielding insight into statistical averages due to random
+processes. During the last two decades considerable research has been done on
+developing and refining flat-histogram Monte Carlo methods. These Monte Carlo
+(MC) methods are capable of exploring all of the energy phase space for all
+given temperatures in a \emph{single} simulation. In this work, we first
+outline the necessary statistical mechanics background. We provide a
+description of MC methods and flat-histogram methods, as well as, our newly
+developed method SAD (stochastic approximation Monte Carlo with a dynamic
+update factor). We discuss physics rich test systems that can be used to
+explore the convergence and statistical properties of our newly developed
+method SAD.
 
@@
     S(E_1,E_2) &= k_B \ln(\Omega(E_1)\Omega_2(E_2)) \\
-    &= k_B \left( \Omega_1(E_1) + \Omega_2(E_2)\right) \\
+    &= k_B \left( \ln(\Omega_1(E_1)) + \ln(\Omega_2(E_2))\right) \\
     &= k_B \left( S_1(E_1) + S_2(E_2)\right)
@@
 \begin{align}
-    T = \left(\diff{S(E)}{E}\right)_V = \frac{k_B}{\Omega(E)}\left(\diff{\Omega(E)}{E}\right)_V
+    \frac{1}{T} = \left(\diff{S(E)}{E}\right)_V = \frac{k_B}{\Omega(E)}\left(\diff{\Omega(E)}{E}\right)_V
 \end{align}
@@
     \D{U} =  T\D{S} - p\D{V} \\
-    T = \left(\diff{U}{S}\right)_V
+    T = \left(\pdv{U}{S}\right)_V
 \end{align}
@@
 \begin{align}
-    \mathcal{P} = e^{\nicefrac{F-E}{k_BT}}
+    \mathcal{P} = e^{\nicefrac{\left(F-E\right)}{k_BT}}
 \end{align}
-where $F\left(N,V,T\right)$ is the free energy of the system $\Gamma$ and will be discussed in detail in section~\HR{potentials}.
+where $F\left(N,V,T\right)$ is the free energy of the system $\Gamma$ and will be discussed in detail in Section~\HR{potentials}.
 
@@
 have units of energy $k_BT$. In this section, we will discuss five common thermodynamic potentials and how they are mathematically related through Maxwell relations.
-While the internal energy $U(E)$ has been briefly mentioned in section~\HR{laws}, we outline in detail the relationship between the state variables.
+While the internal energy $U(E)$ has been briefly mentioned in Section~\HR{laws}, we outline in detail the relationship between the state variables.
 \begin{align}
@@
 \begin{align}\label{maxF}
-    S = \left( \pdv{F}{T}\right )_{V,N}  \quad\quad 
-    p = -\left( \pdv{F}{V}\right )_{T,N}  \quad\quad 
+    S = -\left( \pdv{F}{T}\right )_{V,N}  \quad\quad 
+    p = \left( \pdv{F}{V}\right )_{T,N}  \quad\quad 
     \mu = \left( \pdv{F}{N}\right )_{T,V}  \quad
@@
 \begin{align}
-    \left( {\pdv{G}{T}{p}} \right )_{N} = -\left( \pdv{S}{p}\right )_{T,N} = - \left( \pdv{V}{T}\right )_{p,N}
+    \left( {\pdv{G}{T}{p}} \right )_{N} = -\left( \pdv{S}{p}\right )_{T,N} = \left( \pdv{V}{T}\right )_{p,N}
 \end{align}
@@
 \begin{align}
-    C_V = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{V,N} = \frac{1}{T}\left(\pdv{S}{T}\right )_{V,N}
+    C_V = T\left(\pdv{S}{T}\right )_{V,N}
 \end{align}
@@
 \begin{align}
-    C_p = \left(\pdv{U}{S}\right )_{V,N}\left( \pdv{S}{T} \right )_{p,N} = T\left(\pdv{S}{T}\right )_{p,N}
+    C_p = T\left(\pdv{S}{T}\right )_{p,N}
 \end{align}
@@
 \begin{align}
-    \mathcal{P} = e^{\nicefrac{\Phi + \mu N-E}{k_BT}}
+    \mathcal{P} = e^{\nicefrac{\left(\Phi + \mu N-E\right)}{k_BT}}
 \end{align}

### 2020-05-30 gas-adsorption/gas-ads-body.tex
made all changes in thesis after meeting and also changed fig in gas adsorption.

@@
 
-\begin{figure}
-    \centering
-    \subfloat[][]{
-        \includegraphics[width=0.6\columnwidth]{hkust-1_fancy.png} \label{fig:example_MOF}
-    }
-    \qquad
-    \subfloat[][]{
-        \includegraphics[width=0.8\columnwidth]{usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
-    }
-    \caption{Gas storage and delivery using metal-organic frameworks (MOFs).
+% \begin{figure}
+%     \centering
+%     \subfloat[][]{
+%         \includegraphics[width=0.6\columnwidth]{hkust-1_fancy.png} \label{fig:example_MOF}
+%     }
+%     \qquad
+%     \subfloat[][]{
+%         \includegraphics[width=0.8\columnwidth]{usable_capacity_illustration_toy_ish.pdf} \label{fig:delcap}
+%     }
+%     \caption{Gas storage and delivery using metal-organic frameworks (MOFs).
 (a) the crystal structure of an archetype MOF, CuBTC \cite{chui1999chemically}.
@@
 The deliverable capacity $\rho_D$ is illustrated as the density of gas in the MOF at the storage pressure $\pfull$ minus the density at the discharge pressure $\pempty$.
-    }
-    \label{fig:fig1}
+%     }
+%     \label{fig:fig1}
+% \end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\columnwidth]{Combine_Figure.pdf}
+    \begin{minipage}{0.25\textwidth}
+    \phantomsubcaption{\label{fig:example_MOF}}
+    \end{minipage}
+    \begin{minipage}{0.25\textwidth}
+    \phantomsubcaption{\label{fig:delcap}}
+    \end{minipage}
+
+    \caption{\label{fig:fig1} Gas storage and delivery using metal-organic frameworks (MOFs).
+(\subref{fig:example_MOF}) the crystal structure of an archetype MOF, CuBTC \cite{chui1999chemically}.
+(\subref{fig:delcap}) the methane adsorption isotherm in CuBTC \cite{chui1999chemically} at 298 K (blue) (data from Ref.~\cite{mason2014evaluating}).
+The deliverable capacity $\rho_D$ is illustrated as the density of gas in the MOF at the storage pressure $\pfull$ minus the density at the discharge pressure $\pempty$.}
 \end{figure}

### 2020-06-01 gas-adsorption/gas-ads-appendix.tex
the latest paper changes.

@@
 \begin{align}
-    \frac{\delta D}{\delta \Vk} &= 0, \text{ if } \kvec\ne 0.
+    \frac{\delta D}{\delta \Vk} &= \mathbf{0}, \text{ if } \kvec\ne \mathbf{0}.
 \end{align}
-We note that this functional derivative may be non-zero for $\kvec=0$ because
-we separately maximize with respect to the particular uniform potential $\V$.
-This means that
+We note that this functional derivative may be non-zero for $\kvec=\mathbf{0}$ 
+because we separately maximize with respect to the particular uniform potential 
+$\V$. This means that
 
 \begin{align}
-    \frac{\delta N_H}{\delta \Vk} &= \frac{\delta N_L}{\delta \Vk}
+    \frac{\delta N_\text{full}}{\delta \Vk} &= \frac{\delta N_\text{empty}}{\delta \Vk}
 \end{align}
-where $N_H$ and $N_L$ are the number of particles at the low and high pressure.
+where $N_\text{full}$ and $N_\text{empty}$ are the number of particles at the full and empty pressure, respectively.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{g_st_piston_2.png}
-    \caption{Cartoon illustrating the definition of $\gst$ and a
-      hypothetical experiment to measure it with a piston and
-      removable barrier that separates two equal volumes, one with a
-      porous material (bottom) and one with free space (top). The
-      initial state (left) is a bulk gas phase at density $\rho$ and
-      temperature $T$. An infinitely-thin, removable barrier separates
-      the gas from the porous material (bottom). The barrier is then
-      slowly removed, and the piston slowly pushes all gas molecules
-      into the porous material in an isothermal process. The final
-      state (right) is an adsorbed gas phase at density $\rho$ and
-      temperature $T$.}
+    \includegraphics[width=\columnwidth]{g_st_simple.pdf}
+    \caption{Cartoon illustrating the definition of $\gst$.
+for a given adsorbent and gas, $\gst$=$\gst (\rho, T)$ is defined as the difference in chemical potential between the adsorbed gas system and bulk gas system when both exhibit the same density and temperature.}
     \label{fig:delta-G-cartoon}
@@
     \includegraphics[width=0.95\columnwidth]{four-cases-pro}
-    \caption{Two pairs of boxes containing gas with one of the pairs
-    filled with MOF. One box of each sort has a density of gas $\rhofull$
-    corresponding to $\pfull$ in the MOF system, while the other two boxes has
-    gas density $\rhoempty$}
+    \caption{Two pairs of two connected and equal volumes containing gas.
+In the top pair, the density of gas in each volume is $\rhofull$.
+In the bottom pair, the density of gas in each volume is $\rhoempty$.
+The two volumes on the left are packed with a porous material.
+The two volumes on the right offer a spatially uniform background potential energy $\V$ (our idealized substrate).
+A connection between the two volumes in each pair has a valve to allow flow of gas between them}
     \label{fig:delta-gst-maximum}
@@
 increasing the density of gas causes some of the gas to reside in higher-energy
-sites. If this is always the case, the deliverable capacity will always be
-bounded by our theoretical upper bound, even for a value of $\V$ that is not
-optimal.
+sites.
+If this is always the case, the deliverable capacity of a real material (with a non-spatially uniform potential) with a given $\gst$ at $\pfull$ or $\pempty$ is bounded below by the deliverable capacity in our idealized substrate with $\V=\gst$, even for a value of $\V$ that is not optimal
 
 To show this, we perform a thought experiment illustrated in
-Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of boxes containing
-gas. Two of the boxes are filled with a MOF, and the other two have a uniform
-potential energy $\V$. We imagine the two boxes at $\V$ as being empty
-but at very low altitude where gravity provides a potential energy difference
-(relative to the box with the MOF in it) of $\V$ to each gas molecule (perhaps
-we are on Jupiter?). One box of each sort has a density of gas $\rhofull$
-corresponding to $\pfull$ in the MOF system, while the other two boxes has gas
-density $\rhoempty$.
+Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of volumes containing
+gas. Within each pair, one volume is filled with a porous material, while the
+other offers a uniform potential energy $\V$ (our idealized substrate). In the
+first (second) pair, the density of gas in each volume is $\rhofull$
+($\rhoempty$), corresponding to the density of gas adsorbed in the porous
+material at $\pfull$ ($\pempty$).
 
-We now consider what happens if we introduce a diffusive connection between two
-boxes that are initially at the same density, for instance by connecting them
-with a tube. If the chemical potential of the two boxes are equal, their
-density will be unchanged, otherwise gas will flow into the box with lower
-chemical potential. The difference in chemical potential is equal to the
-difference
+We now consider what happens if we introduce a diffusive connection between a
+volume of MOF and a volume with an idealized substrate that initially contain
+the same density of gas, for instance by connecting them with a tube. If the
+chemical potential of the two volumes are equal, their density will be
+unchanged; otherwise gas will flow into the volume with lower chemical
+potential. The difference in chemical potential is equal to the difference
 
@@
 \end{align}
-Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at
-their initial density after they are connected. Thus, the deliverable capacity
-of the MOF will be greater than the deliverable capacity of the model with
-uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$
-does \emph{not} monotonically increase. By the same token, if we consider the
-case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable
-capacity than our model, the MOF must have less residual gas, which means that
-gas must spontaneously flow from the low-density MOF to the box with potential
-$\V$, which means that $\gst(\rhoempty)>\V$. Once again, exceeding our upper
-bound requires a material with a $\gst(\rho)$ that does not increase
-monotonically.
+where $\mu_{\text{MOF}}(\rho)$ and $\mu_{\V}(\rho)$ are the chemical potentials
+of gas in the porous material and ideal substrate with potential $\V$, respectively. Thus if $\gst(\rhoempty) =
+\V$ then the low-density containers will remain at their initial density after
+they are connected. Thus, the deliverable capacity of the MOF will be greater
+than the deliverable capacity of the ideal substrate with uniform potential $\V$ if and
+only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$ does \emph{not} monotonically
+increase, since the gas in the high-density system will flow from the ideal
+substrate to the MOF. By the same token, if we consider the case where
+$\gst(\rhofull) = \V$, then for the MOF to achieve a greater deliverable
+capacity than the idealized substrate, the MOF must have less residual gas,
+which means that gas must spontaneously flow from the low-density MOF to the
+volume with potential $\V$, which means that $\gst(\rhoempty)>\V$. Once again,
+exceeding our upper bound requires a material with a $\gst(\rho)$ that does not
+increase monotonically.
 
@@
     \caption{Deliverable capacity of hydrogen at 77\ K as a function of the attractive Gibbs free energy $\gst$.
-Experimental deliverable capacities for several MOFs (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
+Experimental deliverable capacities for several porous materials (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-77-D}

### 2020-06-01 gas-adsorption/gas-ads-body.tex
the latest paper changes.

@@
 The transportation sector is dominantly powered by petroleum-based
-fuels~\cite{davis2009transportation}. In the United Sates, it is responsible
+fuels~\cite{davis2009transportation}. In the United States, it is responsible
 for 36\% of energy-related carbon dioxide emissions \cite{useia} and $\sim$50,000
@@
 materials~\cite{schoedel2016role}. The internal surfaces of porous materials
-attract gas molecules through van der Waals, electrostatic, etc. interactions
+attract gas molecules through van der Waals, electrostatic, etc.\ interactions
 to achieve a higher adsorbed gas density than the bulk gas at the same
@@
 hydrogen storage and delivery are theoretically possible, but sufficiently
-close to the upper bound as to be impractical for any real porous material.
-Optimistically, new paradigms outside the scope of applicability of our
-theoretical framework, such as gas-induced structural transitions of the
+close to the upper bound as to be impractical for any real, rigid porous
+material. Optimistically, new paradigms outside the scope of applicability of
+our theoretical framework, such as gas-induced structural transitions of the
 material, hold promise for meeting these targets, as evidenced by flexible MOF
@@
 adsorption, the ARPA-E deliverable capacity target of 315 L STP/L could be
-reached in both of these idealized substrates. However, both models neglected
-the space occupied by atoms of the porous material that are needed to endow the
-adsorption sites/volume with the attractive potential energy and rely on a
-molecular model for methane.
+reached in both of these idealized substrates. However, both models (i)
+neglected the space occupied by atoms of the porous material that are needed to
+endow the adsorption sites/volume with the attractive potential energy and (ii)
+rely on a molecular model for methane.
 
@@
 set of possible materials. To further explore material space and address
-sensitivity to the intermolecular potentials, scaling the Lennard-Jones
+sensitivity to the intermolecular potentials: scaling the Lennard-Jones
 potential well depths of material atoms to model enhanced
@@
   \includegraphics[width=\columnwidth]{methane-298-rho-mu}
-  \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several MOFs at 298\ K as a function of chemical potential (bottom axis) and pressure (top axis).
+  \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several porous materials at 298\ K as a function of chemical potential (bottom axis) and pressure (top axis).
 The bulk methane density is from the National Institute of Standards and Technology (NIST)~\cite{nist}.
-The methane adsorption isotherms in the MOFs are experimental data from Ref.~\cite{mason2014evaluating, furukawa2009storage}.
+The methane adsorption isotherms in the porous materials are experimental data from Ref.~\cite{mason2014evaluating, furukawa2009storage}.
   }
@@
 two shifted bulk methane density curves and their difference. The horizontal
-axis is the difference in molar Gibbs free energy, at fixed density and
-temperature, between the gas within the substrate and the pure gas (\gst), which
-is equal to $\V$ in our model, and is pressure-dependent in a real
-crystal~(see Sec.~\ref{sec:phi-is-delta-g}). We see a potential $\V$ that maximizes the
-deliverable capacity of our ideal substrate, balancing the need to maximize the
-density at $\pfull$ against the need to minimize residual gas retained at
-$\pempty$.
+axis is the difference in molar Gibbs free energy, at fixed adsorbent density
+and temperature, between the gas within the substrate and the pure gas (\gst),
+which is equal to $\V$ in our model, and is pressure-dependent in a real
+crystal~(see Sec.~\ref{sec:phi-is-delta-g}). We see a potential $\V$ that
+maximizes the deliverable capacity of our ideal substrate, balancing the need
+to maximize the density at $\pfull$ against the need to minimize residual gas
+retained at $\pempty$.
 
@@
 uniform potential $\V_{opt}$ for the gas, places an upper bound on the
-deliverable capacity in all rigid porous materials. In the SI (see
-Section~\ref{sec:proof-extremum}), we show that our idealized substrate with
+deliverable capacity in all rigid porous materials. In
+Section~\ref{sec:proof-extremum}, we show that our idealized substrate with
 spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the
@@
 highest at \pempty. As a result, $\gst$ increases monotonically with
-pressure. As shown in the SI (see Sec.~\ref{sec:monotonic}), this results in
-the deliverable capacity of a non-uniform potential with a given $\gst$ at $\pfull$ or $pempty$ being lower than the
+pressure. As shown in Sec.~\ref{sec:monotonic}, this results in
+the deliverable capacity of a non-uniform potential with a given $\gst$ at $\pfull$ or $\pempty$ being lower than the
 deliverable capacity of a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
@@
     \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
-Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
+Experimental deliverable capacities for several porous materials (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-298-D}
@@
 this is a reasonable approximation, and this assumption is frequently made in
-both the simulation and theory of porous materials~\cite{witman2017influence}.
+both the simulation and theory of porous materials~\cite{duren2009using}.
 However, there are cases where the substrate can provide a very strong
@@
 pressures~\cite{mason2015methane}. This allows Co(bdp) to fully expel its
-residual gas at low pressures. Our upper bound does not consider such material
-effects, and thus flexible materials could have significantly higher
-deliverable capacities.
+residual gas at low pressures. Flexible materials could have significantly 
+higher deliverable capacities, as our upper bound does not pertain to them.
 

### 2020-06-01 gas-adsorption/gas-ads.tex
added line numbering successfully.
made all changes in thesis after meeting and also changed fig in gas adsorption.
the latest paper changes.

@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,nature]{revtex4-1}
+\RequirePackage[mathlines]{lineno}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,nature]{revtex4-2}
+\setlength{\columnsep}{1cm}
+\linenumbers
+
 \usepackage{graphicx} % Include figure files
@@
 \usepackage{amsmath}
-\usepackage{subfig}
 
-\usepackage{lineno} % I added this to do line numbering
-
+%\usepackage{lineno} % I added this to do line numbering
+\usepackage{subcaption}
 \usepackage{nicefrac} % Include for inline fractions
@@
 
+\begin{linenomath}
 \begin{abstract}
@@
 rigorous upper bound on the deliverable capacity. Here, we present a
-theoretical upper bound on the deliverable capacity of a rigid material via an
-isothermal pressure-swing. To provide an extremum, we consider a substrate that
-provides a spatially uniform potential energy field for the gas. Our bound is
-unique in that it directly relies on experimentally measured properties of the
-(real) bulk gas, without making approximations. We conclude that the goals set
-by the US DOE for natural gas and hydrogen storage are theoretically possible,
-but sufficiently close to the upper bound as to be impractical for any real
-porous material. However, limitations to the scope of applicability of our
-upper bound guide future material development efforts. Firstly, one could heat
-the adsorbent to drive off residual gas trapped in the
-adsorbent~\cite{gomez2014exploring}. Secondly, the fundamental physics of our
-upper bound do not pertain to any material that changes its structure in
-response to adsorbed gas, suggesting that flexible materials could still
-satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly,
+theoretical upper bound on the deliverable capacity of a gas in a rigid
+material via an isothermal pressure-swing. To provide an extremum, we consider
+a substrate that provides a spatially uniform potential energy field for the
+gas. Our bound is unique in that it directly relies on experimentally measured
+properties of the (real) bulk gas, without making approximations. We conclude
+that the goals set by the US DOE for room-temperature natural gas and hydrogen
+storage are theoretically possible, but sufficiently close to the upper bound
+as to be impractical for any real, rigid porous material. However, limitations
+to the scope of applicability of our upper bound guide fuel tank design and
+future material development. Firstly, one could heat the adsorbent to drive off
+trapped, residual gas in the adsorbent~\cite{gomez2014exploring}. Secondly, the
+physics of our upper bound do not pertain to any material that changes its
+structure in response to adsorbed gas, suggesting that flexible materials could
+still satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly,
 mason2015methane}.
@@
 \end{abstract}
+\end{linenomath}
 

### 2020-06-01 gas-adsorption/gas-ads-appendix.tex
rework appendix D to hopefully fit better with new figure

@@
 increasing the density of gas causes some of the gas to reside in higher-energy
-sites.
-If this is always the case, the deliverable capacity of a real material (with a non-spatially uniform potential) with a given $\gst$ at $\pfull$ or $\pempty$ is bounded below by the deliverable capacity in our idealized substrate with $\V=\gst$, even for a value of $\V$ that is not optimal
+sites. If this is always the case, the deliverable capacity of a real material
+(with a non-spatially uniform potential) with a given $\gst$ at $\pfull$ or
+$\pempty$ is bounded below by the deliverable capacity in our idealized
+substrate with $\V=\gst$, even for a value of $\V$ that is not optimal.
 
 To show this, we perform a thought experiment illustrated in
-Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of volumes containing
-gas. Within each pair, one volume is filled with a porous material, while the
-other offers a uniform potential energy $\V$ (our idealized substrate). In the
-first (second) pair, the density of gas in each volume is $\rhofull$
-($\rhoempty$), corresponding to the density of gas adsorbed in the porous
-material at $\pfull$ ($\pempty$).
+Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of volumes.  Two of
+these volumes contain a real adsorbent material.  These adsorbent volumes are
+connected to bulk gas reservoirs at $\pfull$ and $\pempty$ respectively, and
+therefore have densities $\rhofull$ and $\rhoempty$.  The other two volumes
+volumes have equal density to the two adsorbent volumes ($\rhofull$ and $\rhoempty$), but contain our
+idealized substrate with uniform potential energy $\V$.
 
-We now consider what happens if we introduce a diffusive connection between a
+We now consider what happens if we open a diffusive connection between a
 volume of MOF and a volume with an idealized substrate that initially contain
 the same density of gas, for instance by connecting them with a tube. If the
-chemical potential of the two volumes are equal, their density will be
-unchanged; otherwise gas will flow into the volume with lower chemical
-potential. The difference in chemical potential is equal to the difference
-
+chemical potential in the idealized substrate is higher than in the porous material,
+then gas will flow out of the idealized substrate, lowering its density.  Conversely
+if the chemical potential is lower in the idealized substrate than in the
+volume of porous material, then gas will flow into the idealized substrate, increasing
+its density.
+The difference in chemical potential between those two volumes is
 \begin{align}
@@
 where $\mu_{\text{MOF}}(\rho)$ and $\mu_{\V}(\rho)$ are the chemical potentials
-of gas in the porous material and ideal substrate with potential $\V$, respectively. Thus if $\gst(\rhoempty) =
+of gas in the porous material and ideal substrate with potential $\V$, respectively.
+Thus if $\gst(\rhoempty) =
 \V$ then the low-density containers will remain at their initial density after

### 2020-06-02 gas-adsorption/gas-ads-appendix.tex
made the final changes.

@@
 \end{align}
-where $N_\text{full}$ and $N_\text{empty}$ are the number of particles at the full and empty pressure, respectively.
+where $N_\text{full}$ and $N_\text{empty}$ are the number of particles at the full and empty pressure, $\pfull$ and $\pempty$ respectively.
 
@@
     \caption{Cartoon illustrating the definition of $\gst$.
-for a given adsorbent and gas, $\gst$=$\gst (\rho, T)$ is defined as the difference in chemical potential between the adsorbed gas system and bulk gas system when both exhibit the same density and temperature.}
+For a given adsorbent and gas, $\gst$=$\gst (\rho, T)$ is defined as the difference in chemical potential between the adsorbed gas system and bulk gas system when both exhibit the same density $\rho$ and temperature $T$.}
     \label{fig:delta-G-cartoon}
@@
 \includegraphics[width=0.95\columnwidth]{four-cases-pro}
-\caption{Two pairs of two connected and equal volumes containing gas.
-In the top pair, the density of gas in each volume is $\rhofull$.
-In the bottom pair, the density of gas in each volume is $\rhoempty$.
-The two volumes on the left are packed with a porous material.
-The two volumes on the right offer a spatially uniform background potential energy $\V$ (our idealized substrate).
-A connection between the two volumes in each pair has a valve to allow flow of gas between them}
+\caption{A bulk gas reservoir is connected to a volume filled with adsorbent material (green/blue: material atoms, red: gas particles).
+The adsorbent is also connected, but with a closed valve, to an equal volume containing the ideal substrate (shaded blue background represents the uniform potential energy field $\V$) with an equal density of gas as in the adsorbent.
+The top (bottom) shows a realization of this connected system where the pressure of the bulk gas is $\pfull$ ($pempty$), and thus the density of gas in the adsorbent is $\rhofull$ ($\rhoempty$) by definition.
+The deliverable capacity of the adsorbent is $\rhofull - \rhoempty$.
+We can determine if the deliverable capacity of gas in the ideal substrate is lower or higher than in the adsorbent by considering what happens when we open the valves connecting the two pairs of volumes containing adsorbent and the ideal substrate.}
 \label{fig:delta-gst-maximum}
@@
 \section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
-Every MOF has a $\Delta g_\text{st}$ at $\pfull$ and $\pempty$ corresponding
-to a full and empty density. The deliverable capacity of the MOF is equal to
+Every adsorbent has a $\Delta g_\text{st}$ at $\pfull$ and $\pempty$ corresponding
+to a full and empty density, $\rhofull$ and $\rhoempty$ respectively. The deliverable capacity of the adsorbent is equal to
 the difference between the full and empty density. Examining
@@
 We now consider what happens if we open a diffusive connection between a
-volume of MOF and a volume with an idealized substrate that initially contain
+volume of adsorbent and a volume with an idealized substrate that initially contain
 the same density of gas, for instance by connecting them with a tube. If the
 chemical potential in the idealized substrate is higher than in the porous material,
-then gas will flow out of the idealized substrate, lowering its density.  Conversely
+then gas will flow out of the idealized substrate, lowering its density.  Conversely,
 if the chemical potential is lower in the idealized substrate than in the
@@
 \begin{align}
-   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
+   \gst(\rho) - \V &= \left(\mu_{\text{ads}}(\rho) - \mu_{\text{gas}}(\rho)\right)
    - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
    \\
-   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
+   &= \mu_{\text{ads}}(\rho) - \mu_{\V}(\rho).
 \end{align}
-where $\mu_{\text{MOF}}(\rho)$ and $\mu_{\V}(\rho)$ are the chemical potentials
+where $\mu_{\text{ads}}(\rho)$ and $\mu_{\V}(\rho)$ are the chemical potentials
 of gas in the porous material and ideal substrate with potential $\V$, respectively.
 Thus if $\gst(\rhoempty) =
-\V$ then the low-density containers will remain at their initial density after
-they are connected. Thus, the deliverable capacity of the MOF will be greater
+\V$, the low-density containers will remain at their initial density after
+they are connected. Thus, the deliverable capacity of the adsorbent will be greater
 than the deliverable capacity of the ideal substrate with uniform potential $\V$ if and
 only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$ does \emph{not} monotonically
-increase, since the gas in the high-density system will flow from the ideal
-substrate to the MOF. By the same token, if we consider the case where
-$\gst(\rhofull) = \V$, then for the MOF to achieve a greater deliverable
-capacity than the idealized substrate, the MOF must have less residual gas,
-which means that gas must spontaneously flow from the low-density MOF to the
+increase, since then the gas in the high-density system will flow from the ideal
+substrate to the adsorbent. By the same token, if we consider the case where
+$\gst(\rhofull) = \V$, then for the adsorbent to achieve a greater deliverable
+capacity than the idealized substrate, the adsorbent must have less residual gas,
+which means that gas must spontaneously flow from the low-density adsorbent to the
 volume with potential $\V$, which means that $\gst(\rhoempty)>\V$. Once again,
@@
 Taken together, this indicates that not only is our absolute upper bound an
-upper bound for rigid MOFs, but the green curve labeled $\rho_D(\gst)$ in
+upper bound for rigid adsorbents, but the green curve labeled $\rho_D(\gst)$ in
 Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and
@@
 \section{Cryogenic hydrogen storage}\label{sec:cryo-hydrogen}
+One approach to increase the deliverable capacity is to reduce the storage
+temperature. This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows
+the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling
+point of nitrogen. The DOE ULTIMATE target in this case looks far more
+achievable, and with a much lower $|\gst|$. In fact, an empty tank at this
+temperature can satisfy the DOE 2020 target. The DOE ULTIMATE target is 14\%
+below the upper bound. Actual adsorbents fall far short of the theoretical maximum.
+
 \begin{figure}
@@
 \end{figure}
-
-One approach to increase the deliverable capacity is to reduce the storage
-temperature. This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows
-the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling
-point of nitrogen. The DOE ULTIMATE target in this case looks far more
-achievable, and with a much lower $|\gst|$. In fact, an empty tank at this
-temperature can satisfy the DOE 2020 target. The DOE ULTIMATE target is 14\%
-below the upper bound. Actual MOFs fall far short of the theoretical maximum.

### 2020-06-02 gas-adsorption/gas-ads-body.tex
made the final changes.

@@
 reached in both of these idealized substrates. However, both models (i)
-neglected the space occupied by atoms of the porous material that are needed to
+neglect the space occupied by atoms of the porous material that are needed to
 endow the adsorption sites/volume with the attractive potential energy and (ii)
@@
 two shifted bulk methane density curves and their difference. The horizontal
-axis is the difference in molar Gibbs free energy, at fixed adsorbent density
-and temperature, between the gas within the substrate and the pure gas (\gst),
+axis is the difference in molar Gibbs free energy of the gas, at fixed density (the density in the adsorbent) and temperature, between the gas within the substrate and the pure gas (\gst),
 which is equal to $\V$ in our model, and is pressure-dependent in a real
@@
 bound. Figure~\ref{fig:hydrogen-298-D} shows the theoretical upper bound curve
-for hydrogen storage along with experimental measurements for known MOFs. The
+for hydrogen storage along with experimental measurements for known adsorbents. The
 DOE ULTIMATE deliverable capacity target is theoretically possible, however by
@@
 measured properties of pure gases. While these upper bounds do not rule out the
-discovery of materials that reach current DOE targets, they cast strong doubt
+discovery of materials that reach current DOE targets for room temperature hydrogen and methane storage, they cast strong doubt
 on the possibility of achieving these goals when we consider the additional

### 2020-06-02 gas-adsorption/gas-ads-appendix.tex
final paper changes.

@@
 \end{align}
-where $N_\text{full}$ and $N_\text{empty}$ are the number of particles at the full and empty pressure, $\pfull$ and $\pempty$ respectively.
+where $N_\text{full}$ and $N_\text{empty}$ are the number of particles at the full and empty pressure, $\pfull$ and $\pempty$, respectively.
 
@@
 The adsorbent is also connected, but with a closed valve, to an equal volume containing the ideal substrate (shaded blue background represents the uniform potential energy field $\V$) with an equal density of gas as in the adsorbent.
-The top (bottom) shows a realization of this connected system where the pressure of the bulk gas is $\pfull$ ($pempty$), and thus the density of gas in the adsorbent is $\rhofull$ ($\rhoempty$) by definition.
+The top (bottom) shows a realization of this connected system where the pressure of the bulk gas is $\pfull$ ($\pempty$), and thus the density of gas in the adsorbent is $\rhofull$ ($\rhoempty$) by definition.
 The deliverable capacity of the adsorbent is $\rhofull - \rhoempty$.
@@
 Every adsorbent has a $\Delta g_\text{st}$ at $\pfull$ and $\pempty$ corresponding
-to a full and empty density, $\rhofull$ and $\rhoempty$ respectively. The deliverable capacity of the adsorbent is equal to
+to a full and empty density, $\rhofull$ and $\rhoempty$, respectively. The deliverable capacity of the adsorbent is equal to
 the difference between the full and empty density. Examining
@@
 connected to bulk gas reservoirs at $\pfull$ and $\pempty$ respectively, and
-therefore have densities $\rhofull$ and $\rhoempty$.  The other two volumes
-volumes have equal density to the two adsorbent volumes ($\rhofull$ and $\rhoempty$), but contain our
+by definition contain gas at density $\rhofull$ and $\rhoempty$.  The other two volumes
+volumes contain gas at a density equal to the two adsorbent volumes ($\rhofull$ and $\rhoempty$), but contain our
 idealized substrate with uniform potential energy $\V$.

### 2020-06-02 gas-adsorption/gas-ads.tex
final paper changes.

@@
 \RequirePackage[mathlines]{lineno}
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,nature]{revtex4-2}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,aps,pre,10pt]{revtex4-1}
 \setlength{\columnsep}{1cm}
@@
 \usepackage{url}
-\usepackage[hypertexnames=false]{hyperref}
-\hypersetup{
-    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
-    pdftoolbar=true,        % show Acrobat’s toolbar?
-    pdfmenubar=true,        % show Acrobat’s menu?
-    pdffitwindow=false,     % window fit to page when opened
-    pdfstartview={FitH},    % fits the width of the page to the window
-    pdftitle={Pommerenck Thesis},    % title
-    pdfauthor={Jordan K. Pommerenck},     % author
-    pdfsubject={Subject},   % subject of the document
-    pdfcreator={Creator},   % creator of the document
-    pdfproducer={Producer}, % producer of the document
-    pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
-    pdfnewwindow=true,      % links in new PDF window
-    colorlinks=true,       % false: boxed links; true: colored links
-    linkcolor=black,          % color of internal links (change box color with linkbordercolor)
-    citecolor=black,        % color of links to bibliography
-    filecolor=magenta,      % color of file links
-    urlcolor=cyan           % color of external links
-}
+%\usepackage[hypertexnames=false]{hyperref}
+% \hypersetup{
+%     unicode=false,          % non-Latin characters in Acrobat’s bookmarks
+%     pdftoolbar=true,        % show Acrobat’s toolbar?
+%     pdfmenubar=true,        % show Acrobat’s menu?
+%     pdffitwindow=false,     % window fit to page when opened
+%     pdfstartview={FitH},    % fits the width of the page to the window
+%     pdftitle={Pommerenck Thesis},    % title
+%     pdfauthor={Jordan K. Pommerenck},     % author
+%     pdfsubject={Subject},   % subject of the document
+%     pdfcreator={Creator},   % creator of the document
+%     pdfproducer={Producer}, % producer of the document
+%     pdfkeywords={flat-histogram, Monte Carlo, Thesis}, % list of keywords
+%     pdfnewwindow=true,      % links in new PDF window
+%     colorlinks=true,       % false: boxed links; true: colored links
+%     linkcolor=black,          % color of internal links (change box color with linkbordercolor)
+%     citecolor=black,        % color of links to bibliography
+%     filecolor=magenta,      % color of file links
+%     urlcolor=cyan           % color of external links
+% }
 

### 2020-06-06 dissertation/conclusion.tex
added a new dissertation tex which conforms to OSU guidlines.
fixed the conclusion to specification and greatly increased length.

--- 2020-04-30 thesis/chapter-5.tex
@@
-\chapter{General Conclusion}
+In the first manuscript \emph{Stochastic approximation Monte Carlo with a
+dynamic update factor}, we developed and refined a novel Monte Carlo variant of
+SAMC called SAD. We compared the convergence properties of SAD to a variety of
+other weight-based flat-histogram methods. We also examined the convergence
+behavior applied to the square-well fluid and a Lennard-Jones cluster. We found
+that the method SAD performs robustly on both systems independent of
+scaling. SAD effectively and efficiently samples the entire energy space defined
+by a chosen temperature range.
 
-Hello this is the conclusion.
+In addition, we found that SAD converges far more rapidly than SAMC. The
+convergence of SAD rivals WL methods, which were provided the energy range of
+interest \emph{a priori}. Unlike WL, which can converge to the wrong density of
+states, SAD shares the proven convergence properties of SAMC. We found that,
+while SAMC converges for a reasonable choice of $t_0$, this nonphysical
+parameter can be especially difficult to tune. Even for the best choice of
+$t_0$, SAMC does not converge as rapidly as SAD or the WL methods, when a
+relatively small range of energies is required, because the algorithm simulates
+all possible energies.
+
+We also found that SAD is far more powerful in the common scenario where the
+range of energies is not known in advance. For the 31 atom Lennard-Jones
+cluster system, the WL and $1/t$-WL algorithms showed convergence properties
+that were highly dependent on the choice of energy range examined. This was
+particularly problematic when examining heat capacity at low temperatures,
+where it was difficult to determine the lowest energy that would have a
+significant impact at the temperatures of interest.
+
+In the second manuscript \emph{Flat-histogram method comparison on 2D Ising
+model}, we tested SAD on the 2D Ising model against a number of weight-based
+methods including a ``production'' run WL. We also examined the convergence
+behavior applied to the 2D Ising model. We found that SAD, $1/t$-WL, and WL
+followed by a production run (with an adequately small $\gamma^{\min}$)
+demonstrate excellent and robust convergence.
+
+We found that for larger Ising systems, SAD reduced the update factor more
+slowly (and conservatively) than $1/t$-WL and WL followed by a production run.
+However, the WL methods solved a simpler problem because they were given the
+energy range \emph{a priori}, rather than a temperature range of interest such
+as SAD requires. The 2D Ising model is a system where the energy range is known
+\emph{a priori.}, which made a comparison between WL methods and SAD more
+biased in favor of WL methods. We also showed in this work that WL followed by
+a production run performed extremely well and is preferable to pure WL for
+ensuring both ergodicity and detailed balance.
+
+In the third manuscript \emph{An upper bound to gas delivery via pressure-swing
+adsorption in nanoporous materials}, we developed an upper bound on the
+deliverable capacity via pressure-swing adsorption in rigid porous solids and
+compared the theory with experimental properties of pure gases. Although the
+upper bound does not rule out the discovery of materials that reach current DOE
+targets for room temperature hydrogen and methane storage, it casts strong
+doubt on the possibility of achieving these goals when considering the
+additional constraints imposed due to steric hindrance between substrate atoms
+and the adsorbate.
+
+Additionally, we found that the upper bound helps by directing research toward
+materials with an ideal energy of attraction as computed from the theory. Our
+upper bound can also readily be applied to the storage of other gases of
+interest (beyond hydrogen and methane), with the proviso that the gas is far
+from crystallization.
+
+A driving motivation for the development of the novel Monte Carlo method SAD
+and a theoretical upper bound on the deliverable capacity in rigid porous
+solids is to allow the simulation of key thermodynamic properties of porous
+materials. By computationally determining a materials deliverable capacity,
+researchers save countless development and experimental hours. In particular,
+this body of research heavily contributes to enabling light-vehicle gas storage
+applications. However, the are two important research extensions that remain in
+order to completely develop a method suitable for simulating the thermodynamic
+properties of interest for any porous material. (i) The first is the
+development of SAD with a minimum pressure $p_{\min}$ instead of a $T_{\min}$.
+By varying the number of particles, the simulation can yield thermodynamic
+insights for all densities (which is of necessary for computing the deliverable
+capacity). (ii) The second step that remains is the merger of each SAD method
+such that a multidimensional 2D SAD is formed. Such a Monte Carlo method would
+be an invaluable tool for helping to enable light-vehicle gas storage
+technology.

### 2020-06-06 dissertation/abstract.tex
reformed my abstract in response to comments.

@@
 method is applied to a number of physical `test’ systems including the 2D Ising
-model, a square-well fluid, and a Lennard-Jones cluster. A driving motivation
-for developing novel Monte Carlo methods that have physically based tunable
-parameters is to provide simulation methods for exploring metal-organic
-frameworks (MOFs). A theoretical framework for gas adsorption and delivery in
-metal-organic frameworks is developed and compared with experimental data. SAD
-is introduced as a powerful method for examining thermodynamic properties for
-MOFs. A preliminary groundwork is laid for the future development of a
-multi-dimensional SAD which would be able to compute all properties of interest
-for any given MOF thereby providing a powerful way for researchers to determine
-any porous media's suitability for gas storage applications.
+model, a square-well fluid, and a 31 atom Lennard-Jones cluster. We find that
+SAD performs robustly on all of the systems. Also, SAD efficiently samples the
+entire energy space defined by a chosen temperature range (rather than using
+unphysical parameters). Additionally, we develop a theoretical upper bound for
+gas adsorption and delivery in porous materials and compare the upper bound with
+experimental data.
+A driving motivation for developing the novel Monte Carlo method SAD is to
+provide a simulation method for computing the thermodynamic properties of
+porous materials. By computationally determining a materials deliverable
+capacity, researchers save countless development and experimental hours. In
+particular, this body of research heavily contributes to enabling light-vehicle
+gas storage applications.

### 2020-06-06 dissertation/introduction.tex
finished all of the corrections and edits.

@@
-Monte Carlo methods are a powerful simulation tool that allow scientists to
+Monte Carlo (MC) methods are a powerful simulation tool that allow scientists to
 explore properties of systems that may otherwise by inaccessible via
@@
 developing and refining flat-histogram Monte Carlo methods. These Monte Carlo
-(MC) methods are capable of exploring all of the energy phase space for all
+methods are capable of exploring all of the energy phase space for all
 given temperatures in a \emph{single} simulation. In this work, we first
@@
 
-A driving motivation for developing novel Monte Carlo methods that have physically based tunable parameters is to provide simulation methods for exploring metal-organic frameworks (MOFs).
-In this work, we examine MOFs and
-their application to gas storage and delivery. We formulate a theoretical upper
-bound on the maximum amount of gas that can be stored and released at different
-pressures.
-We propose SAD as a powerful method for examining thermodynamic properties for MOFs.
-We lay a groundwork for 2D SAD which would be able to compute isotherms for any given MOF thereby providing a powerful way for researchers to determine a MOF's suitability for gas storage applications.
+A driving motivation for developing novel Monte Carlo methods that have
+physically based tunable parameters is to provide simulation methods for
+exploring porous materials. In this work, we examine the theoretical upper
+bound on the storage capacity of porous materials in the context of
+light-vehicle fuel storage. We formulate a theoretical upper bound on the
+maximum amount of gas that can be stored and released at different pressures.
+We propose SAD as a potentially powerful method for examining thermodynamic
+properties for porous materials. This body of research serves as a groundwork
+for multidimensional SAD which would be able to compute isotherms for any given
+porous material thereby providing a powerful way for researchers to determine
+the suitability for gas storage applications.
 
-In the following chapters, we follow the manuscript outline with each chapter corresponding to our published or submitted research on the SAD method applied to various physical systems and our work on establishing a theoretical upper bound on gas storage and delivery for MOFs.
+In the following chapters, we follow the manuscript outline with each chapter
+corresponding to our published or submitted research on the SAD method applied
+to various physical systems, and our work on establishing a theoretical upper
+bound on gas storage and delivery for MOFs.
 
 \section{Micro-Canonical Ensemble}\label{micro}
-A statistical ensemble is defined as a probability distribution $\mathcal{P}\left(\Omega\right)$ for the state of the system.
-The micro-canonical\index{ensemble!micro-canonical} is a representation for a system $\Gamma$ that is determined by a fixed number\index{extensive!number $N$} of particles $N$ (and fixed components $\sum N_i$ for a multi-component system), a fixed volume\index{extensive!volume $V$} $V$, and a fixed energy\index{extensive!energy $E$} $E$.
-All possible states for the system $\Gamma\left(N,V,E\right)$ have the same number of particles, energy, and volume due to the system being in isolation from any external environment.
-The micro-canonical ensemble\index{ensemble!micro-canonical} is for this reason also known as the $NVE$ ensemble as the micro-states depend on the \ind{extensive} macroscopic details.
+A statistical ensemble is defined as a probability distribution
+$\mathcal{P}\left(\Omega\right)$ for the state of the system. The
+micro-canonical\index{ensemble!micro-canonical} is a representation for a
+system $\Gamma$ that is determined by a fixed number\index{extensive!number
+$N$} of particles $N$ (and fixed components $\sum N_i$ for a multi-component
+system), a fixed volume\index{extensive!volume $V$} $V$, and a fixed
+energy\index{extensive!energy $E$} $E$. All possible states for the system
+$\Gamma\left(N,V,E\right)$ have the same number of particles, energy, and
+volume due to the system being in isolation from any external environment. The
+micro-canonical ensemble\index{ensemble!micro-canonical} is for this reason
+also known as the $NVE$ ensemble as the micro-states depend on the
+\ind{extensive} macroscopic details.
 
@@
 
-\subsection{Definition of Entropy $S(E)$}\label{entropy}
-The entropy $S(E)$ can be defined for any given system $\Gamma$ with discrete energy states using one of the most important formulae in physics:
+\subsection{The Boltzmann Entropy $S(E)$}\label{entropy}
+The Boltzmann entropy $S(E)$ can be defined for any given system $\Gamma$ with discrete energy states by relating it to the microstates:
 \begin{align}\label{sdos}
@@
     &= k_B \left( \ln(\Omega_1(E_1)) + \ln(\Omega_2(E_2))\right) \\
-    &= k_B \left( S_1(E_1) + S_2(E_2)\right)
+    &= \left( S_1(E_1) + S_2(E_2)\right)
 \end{align}
@@
 
-\subsection{The Laws of Thermodynamics and $S(E)$}\label{laws}
+\subsection{Thermodynamics in the Micro-Canonical Ensemble}\label{laws}
+The statistical entropy defined in Eq.~\HR{sdos} is consistent with the \nth{1} and \nth{2} laws of
+thermodynamics for any system with uniformly distributed microstates.
 The \nth{1} law of thermodynamics states that for an isolated system $\Gamma$, the total internal energy is equal to the sum of the heat $Q$ and work $W$ done.
@@
 \end{align}
-The relationship between heat and entropy for a system is invaluable for the \nth{1} law to relate the conjugate variables T, $S(E)$, and $U(E)$.
+The relationship between heat and entropy for a system is useful for the \nth{1} law to relate the conjugate variables T, $S(E)$, and $U(E)$.
 
 \section{Canonical Ensemble}
-The Canonical ensemble is a representation for a system $\Gamma$ that is determined by a fixed number of particles $N$ (and fixed components $\sum N_i$ for a multi-component system), a fixed volume $V$, and a fixed temperature $T$.
-All possible states for the system $\Gamma\left(N,V,T\right)$ have the same number of particles, energy, and temperature due to the system being in thermal equilibrium with a heat bath.
-The canonical ensemble is for this reason also known as the $NVT$ ensemble.
+The Canonical ensemble is a representation for a system $\Gamma$ that is
+determined by a fixed number of particles $N$ (and fixed components $\sum N_i$
+for a multi-component system), a fixed volume $V$, and a fixed temperature $T$.
+All possible states for the system $\Gamma\left(N,V,T\right)$ have the same
+number of particles, energy, and temperature due to the system being in thermal
+equilibrium with a heat bath. The canonical ensemble is for this reason also
+known as the $NVT$ ensemble.
 
-The probability $\mathcal{P}$ of selecting a given micro-state at random is
+The probability $\mathcal{P}$ of selecting a given micro-state $i$ at random is
 \begin{align}
-    \mathcal{P} = e^{\nicefrac{\left(F-E\right)}{k_BT}}
+    \mathcal{P_i} = \frac{e^{\nicefrac{\left(-E_i\right)}{k_BT}}}{Z}
 \end{align}
-where $F\left(N,V,T\right)$ is the free energy of the system $\Gamma$ and will be discussed in detail in Section~\HR{potentials}.
+where $Z$ is the canonical partition function and will be discussed in detail in Section~\HR{betaZ}.
 
-\subsection{Boltzmann Factor and Partition Function}
-In statistical thermodynamics $\beta$ is defined as the coldness or the reciprocal of the temperature.
+\subsection{The Canonical Partition Function}\label{betaZ}
+In statistical thermodynamics $\beta$ is defined as the coldness or the reciprocal of the canonical temperature.
 \begin{align}
-    \beta = \frac{1}{k_BT} = \frac{1}{k_B}
+    \beta = \frac{1}{k_BT}
 \end{align}
@@
 \end{align}
+The partition function depends both on the microstate energies as well as the
+temperature $T$. Because the microstate energies are influenced by microscopic
+properties, $Z$ has important statistical meaning.
 
-\subsection{Thermodynamic Potentials}\label{potentials}
-Thermodynamic potentials describe a given thermodynamic state of a given system and
-have units of energy $k_BT$. In this section, we will discuss five common thermodynamic potentials and how they are mathematically related through Maxwell relations.
-While the internal energy $U(E)$ has been briefly mentioned in Section~\HR{laws}, we outline in detail the relationship between the state variables.
+\section{Thermodynamic Potentials}\label{potentials}
+Thermodynamic potentials describe a given thermodynamic state of a given system
+and have units of energy $k_BT$. In this section, we will discuss the
+thermodynamic potentials ($U$, $F$, $H$, and $G$) and how they are mathematically related through
+Maxwell relations. While the internal energy $U(E)$ has been briefly mentioned
+in Section~\HR{laws}, we outline in detail the relationship between the state
+variables.
 \begin{align}
@@
 \subsection{Maxwell Relations}\label{maxwell}
-Maxwell's relations are second derivative equalities that relate important thermodynamic variables.
-Maxwell relations are an extension of the Schwarz theorem which states that the order of differentiation is irrelevant.
+Maxwell's relations are second derivative equalities that relate thermodynamic variables.
+Maxwell relations are formalism of the Schwarz theorem which states that the order of differentiation is irrelevant.
 \begin{align}
@@
 The Maxwell relations from Section~\HR{maxwell} can be written in terms of thermodynamic properties of the material.
-The isochoric heat capacity $C_V$ of a given material is defined as the amount of heat necessary to raise the temperature by a given amount.
+The isochoric heat capacity $C_V$ of a given material is defined as the amount of heat necessary to raise the temperature of the material.
 \begin{align}
@@
 \end{align}
-Similarly, the isochoric coefficient of thermal expansion can be written as the following:
+Similarly, the isochoric coefficient of thermal expansion can be written as
 \begin{align}
-    \alpha_V = \frac{1}{p}\left( \pdv{p}{T} \right )_{V,N}
+    \alpha_V = \frac{1}{p}\left( \pdv{p}{T} \right )_{V,N}.
 \end{align}
-The compressibility determines to what extent a given material will change volume based on the pressure applied to the system.  The isothermal compressibility can be written in terms of the following:
+The compressibility determines to what extent a given material will change volume based on the pressure applied to the system.  The isothermal compressibility can be written:
 \begin{align}
-    \kappa_T = -\frac{1}{V}\left( \pdv{V}{p} \right )_{T,N}
+    \kappa_T = -\frac{1}{V}\left( \pdv{V}{p} \right )_{T,N}.
 \end{align}
-Similarly, the isentropic compressibility can be written as the following:
+Similarly, the isentropic compressibility can be written as
 \begin{align}
-    \kappa_S = -\frac{1}{V}\left( \pdv{V}{p} \right )_{S,N}
+    \kappa_S = -\frac{1}{V}\left( \pdv{V}{p} \right )_{S,N}.
 \end{align}
-A major driving force behind introducing the concepts of thermal expansion and compressibility is to right complex derivatives in terms of these parameters.
+A major driving force behind introducing the concepts of thermal expansion and compressibility is the ability to write complex derivatives in terms of these parameters.
 The internal energy derivative and entropy derivative can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
@@
 The grand canonical ensemble is a representation for a system $\Gamma$ that is determined by a fixed chemical potential $\mu$, a fixed volume $V$, and a fixed temperature $T$.
-All possible states for the system $\Gamma\left(\mu,V,T\right)$ have the same chemical potential, volume, and temperature due to the system being in thermal and chemical equilibrium with a given resevoir.
+All possible states for the system $\Gamma\left(\mu,V,T\right)$ have the same chemical potential, volume, and temperature due to the system being in thermal and chemical equilibrium with a given reservoir.
 The grand canonical ensemble is for this reason also known as the $\mu VT$ ensemble.
@@
 Analogous to potential energy, higher chemical potential drives particles to lower potential (thereby usually changing the particle number).
-The chemical potential $\mu_i$ of species $i$ can be represented using a variety of free energies (all of which are equivalent but not necessarily useful for any given experiment).
+The chemical potential $\mu_i$ of species $i$ can be represented using a variety of free energies (all of which are equivalent).
 \begin{align}
@@
 \section{Monte Carlo Simulation}\label{monte}
-Experiment and simulation blend together to help form a more complete understanding of our knowledge of how the universe works.
-Computational `experiments' such as Monte Carlo methods aid in this understanding by allowing scientists to explore system configurations that are not easily accessible to traditional experiment.
+Experiment and simulation are both necessary to help form a more complete
+understanding of our knowledge of how the universe works. Computational
+`experiments' such as Monte Carlo methods aid in this understanding by allowing
+scientists to explore system configurations that are not easily accessible to
+traditional experiment.
 
-In the field of statistical physics, Monte Carlo methods are used to determine equilibrium 
-properties (thermal averages) of multi-particle systems.  A number of Monte Carlo simulations 
-have been developed over the last century. Some of these methods are completely unique while 
-others are more sophisticated variants of an original method. All of the methods share a number 
-of similarities including picking random moves (we will discuss this process in Section~\HR{rng}) and determining the acceptance or rejection of a 
-move based on a probability distribution. They differ primarily in how they explore phase space 
-and thus compute thermodynamic properties such as the density of states $\Omega(E)$. The accuracy of any MC simulation depends on the details of how the system is `explored'. To reduce the error 
-in simulation one can in principle simply continue to allow the simulation more time to explore 
-phase space.
+In the field of statistical physics, Monte Carlo methods are used to determine
+equilibrium properties (thermal averages) of multi-particle systems. A number
+of Monte Carlo simulations have been developed over the last century. Some of
+these methods are completely unique while others are more sophisticated
+variants of an original method. All of the methods share a number of
+similarities including picking random moves (we will discuss this process in
+Section~\HR{rng}) and determining the acceptance or rejection of a move based
+on a probability distribution. They differ primarily in how they explore phase
+space and thus compute thermodynamic properties such as the density of states
+$\Omega(E)$. The accuracy of any MC simulation depends on the details of how
+the system is `explored'. To reduce the error in simulation one can in
+principle simply continue to allow the simulation more time to explore phase
+space.
 
-The motivation for new research into developing MC methods stems from the desire to solve ever-increasingly complex physical problems in a faster simulation time.
-Many of the originally developed MC methods are unable to effectively explore the complex phase space of real (and even many test) systems.
-Flat-histogram methods (introduced in Section~\HR{fhm}) are able to explore all energies and temperatures in a `single' simulation.
-In addition, they are able to explore complex energy landscapes without getting `stuck' (either never converging or converging at such a slow rate as to not practically converge).
-Many of these flat-histogram methods also have proven convergence on the statistical error thus guaranteeing convergence.
-In the following sections, we will discuss a number of MC methods including importance sampling canonical methods, and flat-histogram methods leading up to our method SAD introduced in this work discussed in Section~\HR{fhm}.
+The motivation for new research into developing MC methods stems from the
+desire to solve ever-increasingly complex physical problems in a faster
+simulation time. Many of the originally developed MC methods are unable to
+effectively explore the complex phase space of real (and even many test)
+systems. Flat-histogram methods (introduced in Section~\HR{fhm}) are able to
+explore all energies and temperatures in a `single' simulation. In addition,
+they are able to explore complex energy landscapes without getting `stuck'
+(either never converging or converging at such a slow rate as to not
+practically converge). Many of these flat-histogram methods also have proven
+convergence on the statistical error thus guaranteeing convergence. In the
+following sections, we will discuss a number of MC methods including importance
+sampling and flat-histogram methods introduced in Section~\HR{cmc} and
+~\HR{fhm}.
 
@@
 
-Since Monte Carlo moves are inherently stochastic in nature, it is imperative that any useful random number generator be fast and ideally non-deterministic.
-In practice, pseudo-random number generators (PRNGs) are used in order to generate the fast random streams although they are deterministic.
-In order for a PRNG to be acceptable for use in Monte Carlo simulation it must meet certain criteria~\cite{sawilowsky2003you}.
-First, the PRNG must have a long `period' before repeating values.
-Poor PRNGs (that is ones with short periods) can lead to systematic error in Monte Carlo simulations~\cite{ferrenberg1992monte, plascak2002cluster, ossola2004systematic}.
-Second, the PRNG must actually be able to pass certain statistical tests such as implemented in the TestU01 library~\cite{l2007testu01}.
+Since Monte Carlo moves are inherently stochastic in nature, it is imperative
+that any useful random number generator be fast and ideally non-deterministic.
+In practice, pseudo-random number generators (PRNGs) are used in order to
+generate the fast random streams although they are deterministic. In order for
+a PRNG to be acceptable for use in Monte Carlo simulation it must meet certain
+criteria~\cite{sawilowsky2003you}. First, the PRNG must have a long `period'
+before repeating values. Poor PRNGs (that is ones with short periods) can lead
+to systematic error in Monte Carlo simulations~\cite{ferrenberg1992monte,
+plascak2002cluster, ossola2004systematic}. Second, the PRNG must actually be
+able to pass certain statistical tests such as implemented in the TestU01
+library~\cite{l2007testu01}.
 
-There are a variety of PRNGs each with benefits and few without some form of drawback.
-We detail some of the well known PRNGs here.
-Linear congruential generators (LCGs) have a very fast time performance and are straightforward to program (in terms of size and complexity).
-Unfortunately, LCGs have multiple issues regarding the generated random number statistical quality~\cite{coddington1994analysis, barry1996recommendations}.
-While the period of the PRNG can be made acceptable by using 64-bit integers, the statistical correlations remain.
-Other PRNGs are the belong to the shift-register family (Xorshift RNGs).
-Like the LCG family, the shift-register RNGs fail many statistical tests~\cite{panneton2005xorshift} and are therefore often considered unacceptable for use in Monte Carlo simulations.
+There are a variety of PRNGs each with benefits and few without some form of
+drawback. We detail some of the well known PRNGs here. Linear congruential
+generators (LCGs) have a very fast time performance and are straightforward to
+program (in terms of size and complexity). Unfortunately, LCGs have multiple
+issues regarding the generated random number statistical
+quality~\cite{coddington1994analysis, barry1996recommendations}. While the
+period of the PRNG can be made acceptable by using 64-bit integers, the
+statistical correlations remain. Other PRNGs belong to the
+shift-register family (Xorshift RNGs). Like the LCG family, the shift-register
+RNGs fail many statistical tests~\cite{panneton2005xorshift} and are therefore
+often considered unacceptable for use in Monte Carlo simulations.
 
 In our work, we use the \textbf{Xoroshiro128+} RNG which is named after its logical operations (XOR $\xrightarrow{}$ rotate $\xrightarrow{}$ shift $\xrightarrow{}$ rotate).
-The RNG has ultra-fast speed 0.72 $\nicefrac{\text{ns}}{\text{64-bits}}$ and incredible statistical quality~\cite{blackman2018scrambled}.
+The RNG has ultra-fast speed 0.72 $\nicefrac{\text{ns}}{\text{64-bits}}$ and excellent statistical quality~\cite{blackman2018scrambled}.
 
@@
 In this section, we introduce canonical Monte Carlo and how it relates to flat-histogram methods introduced in Section~\HR{fhm}.
-The first importance sampling method developed was the famous Metropolis method~\cite{metropolis1953equation}.
+The first importance sampling method developed was the Metropolis method~\cite{metropolis1953equation}.
 The method works by choosing an initial state for the system $\Gamma$.
@@
 \subsection{Flat-Histogram Methods}\label{fhm}
-The goal of flat-histogram methods (also called \emph{broad-histogram}
-or \emph{multicanonical} methods) is to simulate each energy with
-similar accuracy.  In this way, they can accurately determine the density of states
-over a broad range of energies and thus determine the thermodynamic
-quantities such as heat capacity or internal energy
-over a broad range of temperatures.
-Properties that require more information, such as a spatial
-correlation function or a response function, can still be computed
-for any temperature, provided statistics are collected for each
-individual energy, which can then be reweighted for any
+The goal of flat-histogram methods (also called \emph{broad-histogram} or
+\emph{multicanonical} methods) is to sample each microstate with equal
+accuracy. In this way, the method can accurately determine the density of
+states over a broad range of energies and thus determine the thermodynamic
+quantities such as heat capacity or internal energy over a broad range of
+temperatures, where canonical MC cannot. Properties that require more
+information, such as a spatial correlation function or a response function, can
+still be computed for any temperature, provided statistics are collected for
+each individual energy, which can then be reweighted for any
 temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
-errington2003direct}.
-All the flat-histogram Monte Carlo methods begin with randomly chosen
-`moves' which change the state of the system and must satisfy
+errington2003direct}. All the flat-histogram Monte Carlo methods begin with
+randomly chosen `moves' which change the state of the system and must satisfy
 detailed balance. All of the tested algorithms differ in how they determine the
-probability of accepting a move and in what additional statistics must
-be collected in order to decide on that probability.
-Flat-histogram methods calculate the density of states $D(E)$ for a discrete set
-of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
+probability of accepting a move and in what additional statistics must be
+collected in order to decide on that probability. Flat-histogram methods
+calculate the density of states $D(E)$ for a discrete set of
+energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
 trebst2004optimizing}. Therefore, energy binning becomes an important
@@
 balance, due to the acceptance probabilities changing with each move, but the
-severity of this violation decreases as we decrease $\gamma_t$.  The four
+severity of this violation decreases as we decrease $\gamma_t$.  Flat-histogram
 methods differ primarily in how they schedule the decrease of $\gamma_t$.
 
-In this work, we develop a new method called Stochastic Approximation with Dynamic update factor (SAD) method
-which is a variant of the SAMC Algorithm that attempts to dynamically choose
-the modification factor rather than relying on system dependent
-parameters such as $t_0$ or $\gamma_0$.  There is an immediate
-advantage of such an algorithm where parameters are chosen independent
-of system size or type. Our
-proposed method SAD requires the user to input $T_\text{min}$, the
-lowest temperature of interest, which is an immediate disadvantage of
-the method. However, identifying a minimum temperature of interest
-$T_\text{min}$ may be easier for a user than determining in advance
-an energy range of interest or the unphysical parameter $t_0$ such as other
-flat-histogram methods must do.
+In this work, we develop a new method called Stochastic Approximation with
+Dynamic update factor (SAD) method which is a variant of the SAMC Algorithm
+that attempts to dynamically choose the modification factor rather than relying
+on system dependent parameters. There is an immediate advantage of such an
+algorithm where parameters are chosen independent of system size or type. Our
+proposed method SAD requires the user to input $T_\text{min}$, the lowest
+temperature of interest, which is an immediate disadvantage of the method.
+However, identifying a minimum temperature of interest $T_\text{min}$ may be
+easier for a user than determining in advance an energy range of interest or
+the unphysical parameter $t_0$ such as other flat-histogram methods must do.
 
 \section{Test System Details}
-We introduce a number of systems that are suitable for benchmarking the convergence properties of various
-flat-histogram MC methods. Testing on a variety of systems is necessary for ensuring the robustness of a
-given MC method. The square-well fluid is chosen as a test-bed due to the fact that convergence properties
-of flat-histogram weight-based MC methods have never been studied on this system. Secondly, the MC method
-will have to simulate the potentially complex liquid-vapor phase coexistence of the system. The 2D Ising
-model is a simple system with a second order phase transition. In addition, the majority of MC methods have
-been tested on the Ising model so the speed and accuracy with which a given MC method can converge can often
-be directly compared with the literature. A Lennard-Jones cluster is a heavily studied system that can be
-simple or incredibly complex for a MC method to accurately solve. In order to solve for solid-solid transitions
-for certain cluster configurations, it is necessary to explore incredibly low reduced temperatures. This can
-make finding the global minima configuration extremely difficult for MC methods in general.
+We introduce a number of systems that are suitable for benchmarking the
+convergence properties of various flat-histogram MC methods. Testing on a
+variety of systems is necessary for ensuring the robustness of a given MC
+method. (i) The square-well fluid is chosen as a test-bed due to the fact that
+convergence properties of flat-histogram weight-based MC methods have never
+been studied on this system. Secondly, the MC method will have to simulate the
+potentially complex liquid-vapor phase coexistence of the system. (ii) The 2D Ising
+model is a simple system with a second order phase transition. In addition, the
+majority of MC methods have been tested on the Ising model so the speed and
+accuracy with which a given MC method can converge can often be directly
+compared with the literature. (iii) Lennard-Jones clusters are a heavily studied
+system that can be simple or incredibly complex for a MC method to accurately
+solve. In order to solve for solid-solid transitions for certain cluster
+configurations, it is necessary to explore incredibly low reduced temperatures.
+This can make finding the global minima configuration extremely difficult for
+MC methods in general.
 
@@
 \begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
+\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i \sigma_i
 \end{align}
@@
 \end{align}
-A number of challenges exit when trying to accurately simulate the density of states of a cluster system.
+A number of challenges exist when trying to accurately simulate the density of states of a cluster system.
 Narrow potential energy funnels~\cite{mandelshtam2006multiple, wales1997global} and low temperature solid-solid
@@
 (which could lead to boil-off of the particles). A `constraining' sphere with radius $R_c$ is used to prevent
-this but it is usually empirically determined further complicating the simulation process.
+this but it is usually empirically determined, further complicating the simulation process.
 
@@
 
-% This is part of the abstract of our paper.
 Due to their low volumetric energy density at ambient conditions, both hydrogen
@@
 
-Here, we present a theoretical upper bound on the deliverable capacity of a
+We present a theoretical upper bound on the deliverable capacity of a
 rigid material via an isothermal pressure-swing. To provide an extremum, we
@@
 method is necessary for determining the best MOF for gas storage applications.
-A multidimensional 2D SAD would allow researchers to compute all properties of
-interest for all densities and temperatures. Our preliminary research has
+A multidimensional SAD would allow researchers to compute all properties of
+interest for all densities and temperatures. Our research has
 yielded a robust method that has been tested on multiple systems and can be
-extended into a 2D formulation.
+extended into a higher dimension formulation.
 
@@
 application to multiple systems.
-% Now tie this in with gas adsorption.
+
 We are currently working on two extensions which will enable a multidimensional
-SAD. The first is adaptive interpolation histogram binning. This technique will
-decrease the simulation time by orders of magnitude. The second is a version of
-SAD that simulates ``number'' space rather than energy space. With the success
-of these two extensions, we can begin the development of 2D SAD and thus provide
-a powerful way for researchers to work toward enabling environmentally
-sustainable light vehicle storage.
+SAD. The first is (i) adaptive interpolation histogram binning. This technique
+will decrease the simulation time by orders of magnitude. The second is (ii) a
+version of SAD that simulates ``number'' space rather than energy space. These
+two extensions will enable all densities to be extracted and should decrease
+simulation time. The development of 2D SAD will provide a powerful way for
+researchers to work toward enabling environmentally sustainable light vehicle
+storage.

### 2020-06-08 dissertation/introduction.tex
made all committee changes; now working on putting appendix into chapter.

@@
 All possible states for the system $\Gamma\left(N,V,T\right)$ have the same
-number of particles, energy, and temperature due to the system being in thermal
+number of particles, volume, and temperature due to the system being in thermal
 equilibrium with a heat bath. The canonical ensemble is for this reason also
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figs/well-potential.pdf}
+    \caption{The square-well and the Lennard-Jones potentials are graphically shown vs. the radial distance from the particle.
+    }
+    \label{fig:well-potential}
+  \end{figure}
+
 \subsection{Square-Well Fluid}
 The square-well fluid is a system of particles whose interactions are governed by a square-well
-potential~\cite{singh2003surface, barker2004perturbationSW}.  The square-well potential is an ideal test-bed
+potential~\cite{singh2003surface, barker2004perturbationSW} as shown in Fig.~\ref{fig:well-potential}.  The square-well potential is an ideal test-bed
 for Monte Carlo methods as it is a simple model for a liquid, which includes both attractive and repulsive
@@
 for up or down spins respectively. In the absence of a magnetic field ($h =
-0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
+0$), we can write the Hamiltonian as follows~\cite{onsager1944crystal,
 kaufman1949crystal}:
@@
 methods~\cite{doye1998thermodynamics, poulain2006performances, martiniani2014superposition}. The potential
-energy for interacting atomic particles can be written as follows:
+energy for interacting atomic particles can be written as follows and can visually be seen in Fig.~\cite{fig:well-potential}:
 \begin{align}

### 2020-06-09 gas-adsorption/gas-ads-complete.tex
final dissertation edits.

--- 2020-06-02 gas-adsorption/gas-ads-body.tex
@@
 
+% MERGE APPENDIX IN HERE !!! %
+\subsection{An external, spatially uniform potential $\V$ shifts the chemical potential $\mu$ of the gas} \label{sec:V_shifts_chem_pot}
+We show that imposing an external, spatially uniform potential $\V$ to a gas
+has the effect of shifting the chemical potential $\mu$ of the gas. Consider a control volume $\Omega$ with volume
+$V=|\Omega|$ (large enough to neglect boundary effects) that is endowed with
+the external, spatially uniform potential $\V$. Impose the grand-canonical
+ensemble, where this control volume can exchange energy and particles with a
+bath of gas at temperature $T$ and chemical potential $\mu$.
+
+To denote a microstate of this system, let $N$ be the number of gas particles
+in the control volume and $\mathbf{r}_1,...,\mathbf{r}_N$ be their positions.
+Then, the potential energy $E$ of a microstate of the control volume is:
+
+\begin{equation} E(\rvec_1,...,\rvec_N) = N\V +
+E_{gg}(\rvec_1,...,\rvec_N),
+\end{equation}
+where $E_{gg}$ is the (unknown and complicated) interatomic potential for
+gas-gas interactions that governs the (real) gas properties. The first term
+arises from each gas molecule experiencing the external potential $\V$, where
+$\V < 0$ corresponds to attraction. To account for molecular rotational and
+vibrational degrees of freedom, we could treat the positions $\rvec_i$ as the
+locations of atoms rather than molecules. In this case the gas-gas interactions
+will include both \emph{intra-molecular} interactions and
+\emph{inter-molecular} interactions.
+
+The grand canonical partition function of the control volume is:
+
+\begin{multline}
+    \Xi(\mu, V, T)= \\ \displaystyle \sum_{N=0}^\infty \frac{1}{\Lambda^{3N}N!} \int_{\Omega} \cdots \int_{\Omega} e^{-\beta E_{gg}(\rvec_1, ..., \rvec_N)} e^{\beta (\mu - \V) N} d\rvec_1 \cdots d\rvec_N.
+    \label{eq:gcpf}
+\end{multline}
+We recognize this as the grand canonical partition function of the bulk gas in
+the control volume without the external potential, but shifted in chemical
+potential:
+
+\begin{equation}
+    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
+    \label{eq:xi_vs_xi0}
+\end{equation}
+where $\Xi_0(\mu, V, T)$ is the grand partition function of the bulk
+gas in the control volume in the absence of an external potential. Importantly,
+this equivalency depends on the potential $\V$ being spatially uniform.
+Therefore, the thermodynamic properties of the gas atoms in the spatially
+uniform external potential $\V$ (adsorbed in our idealized substrate)
+are equivalent to the properties of the bulk gas at chemical potential $\mu-\V$
+(where $T$ is held fixed). As $\V$ becomes more negative, corresponding to a
+more attractive adsorbent, the thermodynamic properties of the adsorbed gas in
+our ideal substrate are equivalent to the gas at a higher chemical potential.
+
+\subsection{Proof of extremum}\label{sec:proof-extremum}
+To show that a uniform potential gives the highest deliverable capacity, we
+consider an interaction potential between gas and substrate $\V(\rvec)$ that
+varies in space.  In this proof, we make use of the Fourier transform of
+this potential:
+\begin{align}
+    \Vk \equiv \iiint \V(\rvec) e^{-i\kvec\cdot \rvec} d\rvec
+\end{align}
+The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\kvec)\propto\delta(\kvec)$. In order for a uniform potential
+to extremize the deliverable capacity, we must show that the functional
+derivative of the deliverable capacity with respect to $\Vk$ is zero for
+\emph{nonzero} values of $\kvec$, i.e.
+\begin{align}
+    \frac{\delta D}{\delta \Vk} &= \mathbf{0}, \text{ if } \kvec\ne \mathbf{0}.
+\end{align}
+We note that this functional derivative may be non-zero for $\kvec=\mathbf{0}$
+because we separately maximize with respect to the particular uniform potential
+$\V$. This means that
+\begin{align}
+    \frac{\delta N_\text{full}}{\delta \Vk} &= \frac{\delta N_\text{empty}}{\delta \Vk}
+\end{align}
+where $N_\text{full}$ and $N_\text{empty}$ are the number of particles at the full and empty pressure, $\pfull$ and $\pempty$, respectively.
+
+Because the chemical potential $\mu$ varies monotonically with $N$ at fixed
+temperature, we can consider how the chemical potential varies as we change
+$\Vk$ with the number of molecules held fixed. We demonstrate this using the
+cyclic chain rule, which shows us that
+\begin{align}
+    \left(\frac{\delta N}{\delta \Vk}\right)_{\mu} &=
+    -\left(\frac{\delta \mu}{\delta \Vk}\right)_{N}
+    \left(\frac{\partial N}{\partial \mu}\right)_{\Vk}.
+\end{align}
+Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \Vk}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \Vk}\right)_{\mu}=0$.
+Thus we consider
+\begin{align}
+    \left(\frac{\delta \mu}{\delta \Vk}\right)_N
+    &= \left(\frac{\delta \left(\frac{\partial F}{\partial N}\right)_{\volume}}{\delta \Vk}\right)_N
+    \\
+    &= \left(\frac{\partial \left(\frac{\delta F}{\delta \Vk}\right)_{\volume}}{\partial N}\right)_{\volume}
+    \label{eq:dmudpot}
+\end{align}
+where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reordered the functional and partial derivatives.
+Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ yields the number density:
+\begin{align}
+    \frac{\delta F}{\delta \Vk} &= \rho(\kvec)
+\end{align}
+The number density is itself spatially uniform for any stable system in a fluid state at this density (i.e.
+does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.
+Thus, the functional derivative $\frac{\delta F}{\delta \V(\rvec)}$ is actually a uniform function.
+We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
+\begin{align}
+    \left(\frac{\delta \mu}{\delta \Vk}\right)_N &\propto \delta(\kvec) \\
+    \left(\frac{\delta N}{\delta \Vk}\right)_\mu &\propto \delta(\kvec)
+\end{align}
+Thus, the functional derivative of both $\mu$ and $N$ with regard to $\V(\rvec)$ are themselves spatially uniform.
+Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e.
+$\kvec=0$), the derivative of $D$ with respect to any change of potential is zero.
+
+This demonstrates that a spatially uniform potential leads to an extremum value
+of the deliverable capacity. This proof is insufficient, however, to show that
+it must be a true maximum.
+
+\subsection{The real-substrate analog of $\V$ is $\gst$}
+\label{sec:phi-is-delta-g}
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{g_st_simple.pdf}
+    \caption{Cartoon illustrating the definition of $\gst$.
+For a given adsorbent and gas, $\gst$=$\gst (\rho, T)$ is defined as the difference in chemical potential between the adsorbed gas system and bulk gas system when both exhibit the same density $\rho$ and temperature $T$.}
+    \label{fig:delta-G-cartoon}
+\end{figure}
+
+The parameter describing our idealized substrate is $\V$, the spatially uniform
+potential felt by a gas molecule adsorbed in the idealized substrate. A natural
+question is how this potential relates to the properties of real substrates.
+The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the
+gas (see Sec.~\ref{sec:V_shifts_chem_pot}). Because our ideal substrate shifts
+the chemical potential of the gas molecules by providing a spatially uniform
+potential energy field, the entropy of the gas in the ideal substrate is equal
+to the entropy of the gas in its bulk state at the same density and
+temperature. In contrast, a real substrate provides a non-spatially uniform
+potential. Consequently, the entropy of the gas inside a real substrate is
+\emph{not} equal to the entropy of the bulk gas at the same density and
+temperature. Therefore, the parameter analogous to $\V$ in a real substrate
+will involve both energy and entropy. The real-substrate analog of $\V$ is the
+shift of molar Gibbs free energy provided by the substrate, specifically an
+\emph{isosteric} (or constant-density) shift of the Gibbs free energy:
+
+\begin{equation}
+   \gst(\rho, T) \equiv
+   %\frac{G_{\text{ads}}(\rho, T) - G_{\text{gas}}(\rho, T)}{N}\\
+    \mu_{\text{ads}}(\rho, T) - \mu_{\text{gas}}(\rho, T).
+  % \gst(\rho, T) &\equiv g_{\text{gas}}(\rho, T) - g_{\text{ads}}(\rho, T)
+  \label{eq:g_st}
+\end{equation}
+The isosteric Gibbs free energy difference $\gst$ is the difference in molar
+Gibbs free energy (equivalent to chemical potential) between the adsorbed gas system and
+the bulk gas \emph{with the same density of gas molecules}. The quantity $\gst$
+does \emph{not} correspond to a change in the molar Gibbs free energy as a
+molecule is adsorbed, which is zero under conditions of coexistence. The
+quantity $\gst$ in a real substrate is a direct analog to $\V$ in our ideal
+substrate because it is the chemical potential shift needed to impose on the
+bulk gas in coexistence with the real substrate to achieve the same density as
+in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
+Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to
+measure $\gst$ via a piston with a removable partition that separates a volume
+of free space from the same volume of substrate. Note that $\gst$ is a property
+of both the substrate and the identity of the gas. Because real substrates
+offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function
+of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho,
+T)=\V$. Consequently, throughout this article, we show $\gst(\rho, T)$ for real
+substrates at both conditions relevant to gas storage and delivery, $\pfull$
+and $\pempty$.
+
+In practice, we can readily compute $\gst(\rho, T)$ of a real gas/substrate
+system from (i) the (experimental or simulated) equilibrium adsorption isotherm
+of the gas in the substrate and (ii) the (experimental or simulated) chemical
+potential of the bulk gas. Consider the real substrate in thermodynamic
+equilibrium with a bulk gas at fixed temperature $T$ and pressure $p$, and let
+$\rho=\rho(p, T)$ be the density of gas in the substrate. At coexistence, the
+chemical potential of the bulk gas is equal to the chemical potential of the
+adsorbed gas in the substrate. Thus, we can use the experimentally known molar
+Gibbs free energy of the pure, bulk gas system at temperature $T$ and pressure
+$p$ to determine the molar Gibbs free energy of the adsorbed system:
+$\mu_{\text{ads}}(\rho, T)=\mu_{\text{gas}}(p, T)$. We can then also look up
+the known chemical potential of the bulk gas at the same density and
+temperature as in the substrate, $\mu_{\text{gas}}(\rho, T)$. Via
+eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.
+
+An interesting question is how $\gst$ relates to the commonly measured and
+reported isosteric heat of adsorption $q_{st}$, which is roughly the energy
+change when a gas molecule is adsorbed~\cite{sircar1999isosteric,
+tian2017differential}. Figure~\ref{fig:qst-vs-delta-G} shows how $q_{st}$
+compares to $\gst$ for several prominent adsorbents. In every case,
+$|q_{st}|>|\gst|$ because the gas in the adsorbent always has less entropy than
+the gas in the bulk at the same density and temperature. That is, while
+adsorption is energetically favored, it is entropically disfavored due to the
+restrictions imposed on the configuration of the gas molecules via steric
+interactions with the substrate itself; this counters the energetic attraction.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{qst-vs-delta-G}
+    \caption{Relationship between $\gst$ and the isosteric heat
+      $q_{st}$ for several prominent adsorbents at room
+      temperature (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}). The dots represent the properties of methane
+      adsorption at 5.8~bar and 65~bar. The $+$'s represent the
+      properties of hydrogen adsorption at 5~bar and
+      100~bar.}
+    \label{fig:qst-vs-delta-G}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{methane-298-gst}
+    \caption{The density-dependence of $\gst$ of methane in several
+      adsorbents (298\ K). Note that $\gst$ is monotonic in
+      $\rho$. Data from
+      Refs.~\cite{mason2014evaluating, furukawa2009storage}.
+    }
+    \label{fig:methane-gst}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.95\columnwidth]{four-cases-pro}
+\caption{A bulk gas reservoir is connected to a volume filled with adsorbent material (green/blue: material atoms, red: gas particles).
+The adsorbent is also connected, but with a closed valve, to an equal volume containing the ideal substrate (shaded blue background represents the uniform potential energy field $\V$) with an equal density of gas as in the adsorbent.
+The top (bottom) shows a realization of this connected system where the pressure of the bulk gas is $\pfull$ ($\pempty$), and thus the density of gas in the adsorbent is $\rhofull$ ($\rhoempty$) by definition.
+The deliverable capacity of the adsorbent is $\rhofull - \rhoempty$.
+We can determine if the deliverable capacity of gas in the ideal substrate is lower or higher than in the adsorbent by considering what happens when we open the valves connecting the two pairs of volumes containing adsorbent and the ideal substrate.}
+\label{fig:delta-gst-maximum}
+\end{figure}
+
+\subsection{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
+Every adsorbent has a $\Delta g_\text{st}$ at $\pfull$ and $\pempty$ corresponding
+to a full and empty density, $\rhofull$ and $\rhoempty$, respectively. The deliverable capacity of the adsorbent is equal to
+the difference between the full and empty density. Examining
+Fig.~\ref{fig:methane-gst}, we see that a significant variety of known
+experimental $\Delta g_\text{st}$ curves are monotonic. Furthermore, our
+qualitative argument in Section~\ref{sec:upper-bound} suggests that this
+function \emph{should} monotonically increase for rigid substrates, as
+increasing the density of gas causes some of the gas to reside in higher-energy
+sites. If this is always the case, the deliverable capacity of a real material
+(with a non-spatially uniform potential) with a given $\gst$ at $\pfull$ or
+$\pempty$ is bounded below by the deliverable capacity in our idealized
+substrate with $\V=\gst$, even for a value of $\V$ that is not optimal.
+
+To show this, we perform a thought experiment illustrated in
+Fig.~\ref{fig:delta-gst-maximum}. We begin with two pairs of volumes.  Two of
+these volumes contain a real adsorbent material.  These adsorbent volumes are
+connected to bulk gas reservoirs at $\pfull$ and $\pempty$ respectively, and
+by definition contain gas at density $\rhofull$ and $\rhoempty$.  The other two volumes
+volumes contain gas at a density equal to the two adsorbent volumes ($\rhofull$ and $\rhoempty$), but contain our
+idealized substrate with uniform potential energy $\V$.
+
+We now consider what happens if we open a diffusive connection between a
+volume of adsorbent and a volume with an idealized substrate that initially contain
+the same density of gas, for instance by connecting them with a tube. If the
+chemical potential in the idealized substrate is higher than in the porous material,
+then gas will flow out of the idealized substrate, lowering its density.  Conversely,
+if the chemical potential is lower in the idealized substrate than in the
+volume of porous material, then gas will flow into the idealized substrate, increasing
+its density.
+The difference in chemical potential between those two volumes is
+\begin{align}
+   \gst(\rho) - \V &= \left(\mu_{\text{ads}}(\rho) - \mu_{\text{gas}}(\rho)\right)
+   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
+   \\
+   &= \mu_{\text{ads}}(\rho) - \mu_{\V}(\rho).
+\end{align}
+where $\mu_{\text{ads}}(\rho)$ and $\mu_{\V}(\rho)$ are the chemical potentials
+of gas in the porous material and ideal substrate with potential $\V$, respectively.
+Thus if $\gst(\rhoempty) =
+\V$, the low-density containers will remain at their initial density after
+they are connected. Thus, the deliverable capacity of the adsorbent will be greater
+than the deliverable capacity of the ideal substrate with uniform potential $\V$ if and
+only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$ does \emph{not} monotonically
+increase, since then the gas in the high-density system will flow from the ideal
+substrate to the adsorbent. By the same token, if we consider the case where
+$\gst(\rhofull) = \V$, then for the adsorbent to achieve a greater deliverable
+capacity than the idealized substrate, the adsorbent must have less residual gas,
+which means that gas must spontaneously flow from the low-density adsorbent to the
+volume with potential $\V$, which means that $\gst(\rhoempty)>\V$. Once again,
+exceeding our upper bound requires a material with a $\gst(\rho)$ that does not
+increase monotonically.
+
+Taken together, this indicates that not only is our absolute upper bound an
+upper bound for rigid adsorbents, but the green curve labeled $\rho_D(\gst)$ in
+Figs.~\ref{fig:methane-298-D}, \ref{fig:hydrogen-298-D}, and
+\ref{fig:hydrogen-77-D} is an upper bound for materials with a non-optimal
+$\gst$ at either $\pfull$ or $\pempty$.
+
+% End here %
+
 To place an upper bound on the deliverable capacity, consider a substrate whose
@@
 Consequently, the density of gas in our homogeneous substrate is:
-
 \begin{align}
@@
 The deliverable capacity of gas in our homogeneous substrate is thus:
-
 \begin{align}
@@
 porous materials. This reflects the known fact that hydrogen interacts with
-substrates far more weakly than methane does. One possibility that has been
-pursued is storage at cryogenic temperatures. See
-Section~\ref{sec:cryo-hydrogen} in the SI for the upper bound for hydrogen
-storage at 77 K.
+substrates far more weakly than methane does. 
+One approach to increase the deliverable capacity is to reduce the storage
+temperature. This is illustrated in Fig.~\ref{fig:hydrogen-77-D}, which shows
+the upper bound to the deliverable capacity of hydrogen at 77\ K, the boiling
+point of nitrogen. The DOE ULTIMATE target in this case looks far more
+achievable, and with a much lower $|\gst|$. In fact, an empty tank at this
+temperature can satisfy the DOE 2020 target. The DOE ULTIMATE target is 14\%
+below the upper bound. Actual adsorbents fall far short of the theoretical maximum.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{hydrogen-77-n-vs-G}
+    \caption{Deliverable capacity of hydrogen at 77\ K as a function of the attractive Gibbs free energy $\gst$.
+Experimental deliverable capacities for several porous materials (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
+    \label{fig:hydrogen-77-D}
+\end{figure}
 

### 2020-06-12 gas-adsorption/gas-ads-body.tex
added citation to bib file and found reasonable place to put it in paper.

@@
 metal-organic frameworks~\cite{furukawa2013chemistry}, no porous material has
-met these deliverable capacity targets.
+met these deliverable capacity targets~\cite{firlej2013understanding}.
 

### 2020-06-18 ising/referee-response.md
add our referee response in ising paper.

Dear editor,

We appreciate the referees' careful reading of our paper, and feel that their
feedback has led to a significant improvement. We have addressed all of the
first referee's comments by added two paragraphs to the introduction and a
section in the paper to 'Address the general problem of all "flat-histogram"
methods'. We also strengthen the abstract to address the concerns raised by the
second referee. We point out that: (i) Our paper is the first of its kind (to
the best of our knowledge) to compare ``production run'' WL with other
flat-histogram methods. (ii) It is important to test SAD's convergence
properties on the 2D Ising where the range of energies is easily known (thus
biasing in favor of WL methods). ...

Response to first referee
# -----------------------------------------------------------------------------#

The first referee points out that 'Simulation set-up and analysis protocols are
sound and give confidence in the quality of the presented data and
conclusions.'

The referee also asks that three additions be made to the paper.

(i) It would be 'worthwhile to work out the [slower convergence for SAD due to
the energy range being an a priori parameter in WL/MUCA] in the paper, and, if
possible, quantify the loss in efficiency coming from energy range
determination.'

We have added a paragraph? subsection in the SAD/results portion of the paper that
address this.

(ii) Address the general problem of all "flat histogram" methods that the
performance of these methods is due to hidden barriers almost always lower than
what one would expect from a random walk in energy (in most cases still
exponentially growing). This has been discussed by Nadler et al (PRE 75 (2007)
026109) who also discus a protocol for optimizing the ensemble which minimizes
the residual exponential slowing down. The authors should point out if and how
SAD can be adapted to such "non-flat-histogram" methods.

We have added a paragraph that 'Address the general problem of all "flat
histogram" methods' in the introduction.

(iii) Add a few short remarks on comparison with replica exchange methods would
be helpful.

We have added a 'few short remarks' on comparison with replica exchange in the
introduction.

Response to second referee
# -----------------------------------------------------------------------------#

We thank the second referee for bringing to our attention two points so that we
can better address them in our paper and make them completely clear.

(i) The work is almost a repetition of what the authors did when they
introduced the algorithm.

Our paper is the first of its kind (to the best of our knowledge) to compare
``production run'' WL with other flat-histogram methods. The first (not the second) referee
notes rather strongly that this is the only way that WL should be done;
however, no research work has ever compared this with other flat-histogram
methods and certainly not with the completeness shown here. Also, the very fact that
it is compared with pure WL is proof for the first referee's statement that
"this is how WL should be done" since it performs superior to pure WL.

(ii) It is not clear[ly] understood why the current calculations were not
included in the original article. Although the results presented by the
[authors] are original, they are only a simple extension of the first work, and
do not meet the standards of PRE readers.

The second referee an important question here to which there are two answers.
First, we only recently became aware of ``production run'' WL and wanted to
rapidly compare this with other histogram methods due to the lack of literature
detailing its performance vs other flat-histogram methods. Second, SAD was
initially tested on methods where it could benefit from knowing the temperature
range of interest in advance. It is critical to test it's convergence
properties on a different type of system where the range of energies is easily
known (thus biasing in favor of WL methods).

In response to the second referee's comments, we have better highlighted these
important points in the abstract.

### 2020-06-18 ising/referee-response.md
ideas for referee response

@@
 
+## Ideas
+
+- Measure time until we last find a new important energy
+- Find the fraction of moves (in the end) that are outside the range of interesting energies.
+- Discuss in results or discussion
+
 (ii) Address the general problem of all "flat histogram" methods that the
@@
 exponentially growing). This has been discussed by Nadler et al (PRE 75 (2007)
-026109) who also discus a protocol for optimizing the ensemble which minimizes
+1)      who also discus a protocol for optimizing the ensemble which minimizes
 the residual exponential slowing down. The authors should point out if and how
@@
 
+## Ideas
+
+- Read this paper (probably both of us) and write it up.  We need to describe the paper and discuss how this relates.
+- My guess is that SAD is like any of the other methods for non-flat-histogram approaches.
+- (Optimized ensemble *follows* a flat histogram method.)
+
 (iii) Add a few short remarks on comparison with replica exchange methods would
@@
 introduction.
+
+## This is a bit more work
+
+- Our method *should* be amenable to replica exchange parallelization, just as WL is.
 
@@
 
+## Notes
+
+- We think it is important to bridge the disconnect between *users* of WL and developers of improved-flat histogram methods.
+- Maybe two paragraphs? THIS IS REALLY IMPORTANT (IN THE LETTER)
+
 (ii) It is not clear[ly] understood why the current calculations were not
@@
 
-The second referee an important question here to which there are two answers.
+The second referee asks an important question here to which there are two answers.
 First, we only recently became aware of ``production run'' WL and wanted to
@@
 
+## Notes
+
+- Two paragraphs for two answers?
+- We only became aware of "production run" WL as a result of the response to our
+  first paper.  As mentioned (hinted at) in the acknowledgement, XXX emailed us after
+  reading our paper and suggested that we test ...
+- 2nd: Ising is different.  Range of energies is known.  We omitted it in ther first
+  paper due to size and time.,  These computations are lengthy
+  and slow.
+
 In response to the second referee's comments, we have better highlighted these

### 2020-06-20 ising/referee-response.md
finished the referee response. Now making paper changes.

--- 2020-06-18 ising/referee-response.md
@@
 
-(i) It would be 'worthwhile to work out the [slower convergence for SAD due to
+(i) "It would be worthwhile to work out the [slower convergence for SAD due to
 the energy range being an a priori parameter in WL/MUCA] in the paper, and, if
 possible, quantify the loss in efficiency coming from energy range
-determination.'
+determination."
 
-We have added a paragraph? subsection in the SAD/results portion of the paper that
-address this.
+We have added a paragraph in the discussion section that explains the loss in
+efficiency SAD suffers by not specifying an energy range a priori. We are able
+to quantify this by identifying the last time SAD finds a new important energy
+and then finding the fraction of moves (in the end) that are outside the range
+of interesting energies.
 
-(ii) Address the general problem of all "flat histogram" methods that the
+(ii) "Address the general problem of all "flat histogram" methods that the
 performance of these methods is due to hidden barriers almost always lower than
@@
 exponentially growing). This has been discussed by Nadler et al (PRE 75 (2007)
-026109) who also discus a protocol for optimizing the ensemble which minimizes
-the residual exponential slowing down. The authors should point out if and how
-SAD can be adapted to such "non-flat-histogram" methods.
+1) who also discus a protocol for optimizing the ensemble which minimizes the
+residual exponential slowing down. The authors should point out if and how SAD
+can be adapted to such "non-flat-histogram" methods."
 
 We have added a paragraph that 'Address the general problem of all "flat
-histogram" methods' in the introduction.
+histogram" methods' in the introduction. We discuss how the performance is
+impacted due to hidden barriers. We also add that WL has already been adapted
+to replica-exchange and point out that this should be possible for SAD as well.
 
@@
 We have added a 'few short remarks' on comparison with replica exchange in the
-introduction.
+introduction. WL has been formulated as replica-exchange WLMC whereby it adopts
+the concept of conformational swapping directly from parallel tempering. Each replica
+is allowed to traverse the entire energy range via overlapping energy windows.
+
+While we have performed limited comparison with replica exchange methods (in
+our previous work we found the heat capacity to be consistent with computations
+performed with replica exchange), SAD could be extended in much the same way
+that WL was. In fact, this would make an interesting comparison with replica
+exchange statistical temperature Monte Carlo (as they both specify a
+temperature range) and should be studied on a temperature range system in a
+future work.
 
@@
 
-(i) The work is almost a repetition of what the authors did when they
-introduced the algorithm.
+(i) "The work is almost a repetition of what the authors did when they
+introduced the algorithm."
 
 Our paper is the first of its kind (to the best of our knowledge) to compare
-``production run'' WL with other flat-histogram methods. The first (not the second) referee
-notes rather strongly that this is the only way that WL should be done;
-however, no research work has ever compared this with other flat-histogram
-methods and certainly not with the completeness shown here. Also, the very fact that
-it is compared with pure WL is proof for the first referee's statement that
-"this is how WL should be done" since it performs superior to pure WL.
+``production run'' WL with other flat-histogram methods. The first (not the
+second) referee notes rather strongly that this is the only way that WL should
+be done; however, no research work has ever compared this with other
+flat-histogram methods and certainly not with the completeness shown here.
+Also, the very fact that it is compared with pure WL is proof for the first
+referee's statement that "this is how WL should be done" since it performs
+superior to pure WL.
 
-(ii) It is not clear[ly] understood why the current calculations were not
+We believe that this paper helps by bridging the disconnect between *users* of
+WL and developers of improved flat-histogram methods. The majority of
+developers use pure WL when comparing multiple flat-histogram methods since
+there is no direct comparison of WL followed by a production run with methods
+such as 1/t-WL or SAMC.
+
+(ii) "It is not clear[ly] understood why the current calculations were not
 included in the original article. Although the results presented by the
 [authors] are original, they are only a simple extension of the first work, and
-do not meet the standards of PRE readers.
+do not meet the standards of PRE readers."
 
-The second referee an important question here to which there are two answers.
-First, we only recently became aware of ``production run'' WL and wanted to
-rapidly compare this with other histogram methods due to the lack of literature
-detailing its performance vs other flat-histogram methods. Second, SAD was
-initially tested on methods where it could benefit from knowing the temperature
-range of interest in advance. It is critical to test it's convergence
-properties on a different type of system where the range of energies is easily
-known (thus biasing in favor of WL methods).
+The second referee asks an important question here to which there are two
+answers. We only became aware of "production run" WL as a result of the
+response to our first paper. As mentioned (hinted at) in the acknowledgement,
+Johannes Zierenberg emailed us after reading our paper and suggested that we
+test WL followed by a production run with a number of flat-histogram methods.
+Due to the lack of literature detailing its performance vs other flat-histogram
+methods, the 2D Ising model seems the logical starting point for such a
+comparison.
+
+Second, SAD was initially tested on methods where it could benefit from knowing
+the temperature range of interest in advance. It is critical to test it's
+convergence properties on a different type of system where the range of
+energies is easily known (thus biasing in favor of WL methods). Also in the
+case of the 128 X 128 2D Ising system, the computation time is quite lengthy
+(more than 6 months for some calculations) representing a superior and robust
+comparison result.
 
 In response to the second referee's comments, we have better highlighted these
-important points in the abstract.
+important points in the abstract and conclusion.

### 2020-06-20 ising/ising-body.tex
made some of the changes. Needs some work.

@@
 
+Nadler \emph{et al.} and Trebst \emph{et al.} systematically examined optimized
+ensembles due to performance issues flat-histogram methods faced when
+confronted with hidden energy barriers~\cite{nadler2007generalized,
+nadler2007dynamics, trebst2004optimizing}. Trebst \emph{et al.} used their
+optimized ensemble to systematically set the temperature range(s) and
+implemented an optimized replica-exchange~\cite{trebst2006optimized}.
+Replica-exchange Wang-Landau has been developed with great
+success~\cite{perera2016magnetic} and adapting flat-histogram methods to use
+such an optimized ensemble is doable
+
+Last time for 32 X 32 is TL 2047388. Do I look at histogram to figure this out???
+
+But optimized ensemble requires the energy range???
+
 Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
-related Statistical Temperature Molecular Dynamics (STMD), an adaption
-of the WL method that approximates the entropy as a piecewise linear function,
-which improves convergence for systems with a continuously varying
+related Statistical Temperature Molecular Dynamics (STMD), an adaption of the
+WL method that approximates the entropy as a piecewise linear function, which
+improves convergence for systems with a continuously varying
 energy~\cite{kim2006statistical, kim2007statistical}. STMC applied to WL
@@
 Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping STMC
-simulations to improve convergence~\cite{kim2009replica}. Recently,
-Junghans \emph{et al.} demonstrated a close connection between
-metadynamics, which was introduced by Laio and
-Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods, with STMD
-forging the connection~\cite{junghans2014molecular}.
+simulations to improve convergence~\cite{kim2009replica}. Recently, Junghans
+\emph{et al.} demonstrated a close connection between metadynamics, which was
+introduced by Laio and Parinello~\cite{laio2002escaping}, and WL-based Monte
+Carlo methods, with STMD forging the connection~\cite{junghans2014molecular}.
 
@@
 by a production run performs extremely well and is preferable to pure
-WL for ensuring both ergodicity and detailed balance.
+WL for ensuring both ergodicity and detailed balance. While this has long
+been thought to be the case and that a multi-canonical run where WL is used to
+determine the parameters is the only way WL should be used, this research
+represents the first conclusive comparison among all of these
+flat-histogram methods.
 

### 2020-06-20 ising/ising.tex
fix ising paper so that it updates from ising-body.tex.
made some of the changes. Needs some work.

--- 2020-05-08 ising/ising.tex
@@
 \begin{abstract}
-  We compare the convergence of several flat-histogram methods applied
-  to the 2D Ising model, including the recently introduced stochastic
-  approximation with a dynamic update factor (SAD) method.  We compare
-  this method with the Wang-Landau (WL) method, the $1/t$ variant of
-  the WL method, and standard stochastic approximation Monte Carlo
-  (SAMC).  In addition, we consider a procedure WL followed by a
-  ``production run'' with fixed weights that refines the estimation of
-  the entropy.  To our knowledge, this work is the first to test this
-  approach against other methods.
-  %% To our knowledge, the
-  %% convergence rate of this production-run WL approach has never been 
-%% We examine the convergence of stochastic approximation with a dynamic
-%% update factor (SAD) with other flat-histogram methods applied to the 2D
-%% Ising model. The comparison with stochastic approximation Monte Carlo (SAMC)
-%% and Wang-Landau (WL) methods shows that SAD performs robustly and without user
-%% input knowledge of an energy range. We also compare SAD with WL followed by a
-%% ``production run'' with fixed weights which is the first comparison of its kind
-%% recorded.
-We find that WL followed by a production run \emph{does} converge to
-the true density of states, in contrast to pure WL.  Three of the
-methods converge robustly: SAD, $1/t$-WL, and WL followed by a
-production run.  Of these, SAD does not require \emph{a priori}
-knowledge of the energy range.
+
+We compare the convergence of several flat-histogram methods applied to the 2D
+Ising model, including the recently introduced stochastic approximation with a
+dynamic update factor (SAD) method. We compare this method with the Wang-Landau
+(WL) method, the $1/t$ variant of the WL method, and standard stochastic
+approximation Monte Carlo (SAMC). In addition, we consider a procedure WL
+followed by a ``production run'' with fixed weights that refines the estimation
+of the entropy. To our knowledge, this work is the first to test this approach
+against other methods. We find that WL followed by a production run \emph{does}
+converge to the true density of states, in contrast to pure WL. Three of the
+methods converge robustly: SAD, $1/t$-WL, and WL followed by a production run.
+Of these, SAD does not require \emph{a priori} knowledge of the energy range.
+This work also shows that WL followed by a production run performs superior to
+other forms of WL while ensuring both ergodicity and detailed balance.
+
 \end{abstract}

### 2020-06-22 ising/referee-response.md
a few suggestions on ising edit

@@
 second referee. We point out that: (i) Our paper is the first of its kind (to
-the best of our knowledge) to compare ``production run'' WL with other
+the best of our knowledge) to compare ``production run'' WL with other recent
 flat-histogram methods. (ii) It is important to test SAD's convergence

### 2020-06-22 ising/ising-body.tex
a few suggestions on ising edit

@@
 
+\davidsays{let's keep replica exchange and optimized ensemble in separate paragraphs}
+
 Nadler \emph{et al.} and Trebst \emph{et al.} systematically examined optimized
@@
 optimized ensemble to systematically set the temperature range(s) and
-implemented an optimized replica-exchange~\cite{trebst2006optimized}.
-Replica-exchange Wang-Landau has been developed with great
-success~\cite{perera2016magnetic} and adapting flat-histogram methods to use
-such an optimized ensemble is doable
+implemented an optimized replica-exchange~\cite{trebst2006optimized}
+\davidsays{looks like this is not a flat-histogram approach, scrap this reference}.
 
@@
 But optimized ensemble requires the energy range???
+
+Replica exchange Wang-Landau (REWL) was introduced \davidsays{Need to discuss this, and each of these algorithms
+are sufficiently similar that this approach should apply to them.}\cite{vogel2013rewl}.
+Replica-exchange Wang-Landau has been developed with great
+success~\cite{perera2016magnetic} and adapting flat-histogram methods to use
+such an optimized ensemble is doable
 

### 2020-06-23 ising/ising-body.tex
made minor changes; working on other paper now.

--- 2020-06-20 ising/ising-body.tex
@@
 
+% optimized ensemble paragraph
 Nadler \emph{et al.} and Trebst \emph{et al.} systematically examined optimized
@@
 nadler2007dynamics, trebst2004optimizing}. Trebst \emph{et al.} used their
-optimized ensemble to systematically set the temperature range(s) and
-implemented an optimized replica-exchange~\cite{trebst2006optimized}.
+optimized ensemble to systematically set the temperature range(s).
+
+% replica exchange paragraph
+Replica exchange Wang-Landau (REWL) was introduced~\cite{vogel2013rewl}.
 Replica-exchange Wang-Landau has been developed with great
@@
 
-Last time for 32 X 32 is TL 2047388. Do I look at histogram to figure this out???
-
+% This goes in discussion section
+Last time for 32 X 32 is TL 2047388. Do I look at histogram to figure this out??
 But optimized ensemble requires the energy range???
+Also, SAD doesn't appear to be impacted by the larger energy range? histogram counts outside the interesting energy range are negligible compared with count in the range of interest.
 

### 2020-06-25 gas-adsorption/rsc-gas-ads-complete.tex
add gas ads rcs format for paper.

--- 2020-06-09 gas-adsorption/gas-ads-complete.tex
@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%This is the LaTeX ARTICLE template for RSC journals
+%Copyright The Royal Society of Chemistry 2016
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[twoside,twocolumn,9pt]{article}
+\usepackage{extsizes}
+\usepackage[super,sort&compress,comma]{natbib} 
+\usepackage[version=3]{mhchem}
+\usepackage[left=1.5cm, right=1.5cm, top=1.785cm, bottom=2.0cm]{geometry}
+\usepackage{balance}
+\usepackage{times,mathptmx}
+\usepackage{sectsty}
+\usepackage{graphicx} 
+\usepackage{lastpage}
+\usepackage[format=plain,justification=justified,singlelinecheck=false,font={stretch=1.125,small,sf},labelfont=bf,labelsep=space]{caption}
+\usepackage{float}
+\usepackage{fancyhdr}
+\usepackage{fnpos}
+\usepackage[english]{babel}
+\addto{\captionsenglish}{%
+  \renewcommand{\refname}{Notes and references}
+}
+\usepackage{array}
+\usepackage{droidsans}
+\usepackage{charter}
+\usepackage[T1]{fontenc}
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{setspace}
+\usepackage[compact]{titlesec}
+\usepackage{hyperref}
+%%%Please don't disable any packages in the preamble, as this may cause the template to display incorrectly.%%%
+
+\graphicspath{{figs/}}
+
+\usepackage{subcaption}
+\usepackage{nicefrac} % Include for inline fractions
+\usepackage{url}
+
+% \usepackage{cite}
+\usepackage{xargs}
+% Use more than one optional parameter in a new commands
+\usepackage{color}
+%
+\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
+\usepackage{mdframed}
+
+\newcommand{\addcite}{{\bf \color{red} []}}
+\newcommand{\rvec}{\mathbf{r}}
+\newcommand{\kvec}{\mathbf{k}}
+\newcommand{\thetavec}{\boldsymbol \theta}
+
+\newcommand\V{\Phi}
+\newcommand\Vk{\tilde\Phi(\kvec)}
+\newcommand\volume{V}
+
+\newcommand\pfull{\ensuremath{p_{\text{full}}}}
+\newcommand\pempty{\ensuremath{p_{\text{empty}}}}
+
+\newcommand\mufull{\ensuremath{\mu_{\text{full}}}}
+\newcommand\muempty{\ensuremath{\mu_{\text{empty}}}}
+
+\newcommand\rhofull{\ensuremath{\rho_{\text{full}}}}
+\newcommand\rhoempty{\ensuremath{\rho_{\text{empty}}}}
+
+\newcommand\gst{\ensuremath{\Delta g_{st}}}
+
+\newcommand{\red}[1]{{\bf \color{red} #1}}
+\newcommand{\blue}[1]{{\bf \color{blue} #1}}
+
+\newcommand{\davidsays}[1]{{\color{red} [\blue{D:} \emph{#1}]}}
+\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
+
+\newtheorem{limitation}{Limitation}
+
+\definecolor{cream}{RGB}{222,217,201}
+
+\begin{document}
+
+\pagestyle{fancy}
+\thispagestyle{plain}
+\fancypagestyle{plain}{
+
+%%%HEADER%%%
+\fancyhead[C]{\includegraphics[width=18.5cm]{head_foot/header_bar}}
+\fancyhead[L]{\hspace{0cm}\vspace{1.5cm}\includegraphics[height=30pt]{head_foot/journal_name}}
+\fancyhead[R]{\hspace{0cm}\vspace{1.7cm}\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}
+\renewcommand{\headrulewidth}{0pt}
+}
+%%%END OF HEADER%%%
+
+%%%PAGE SETUP - Please do not change any commands within this section%%%
+\makeFNbottom
+\makeatletter
+\renewcommand\LARGE{\@setfontsize\LARGE{15pt}{17}}
+\renewcommand\Large{\@setfontsize\Large{12pt}{14}}
+\renewcommand\large{\@setfontsize\large{10pt}{12}}
+\renewcommand\footnotesize{\@setfontsize\footnotesize{7pt}{10}}
+\makeatother
+
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\renewcommand\footnoterule{\vspace*{1pt}% 
+\color{cream}\hrule width 3.5in height 0.4pt \color{black}\vspace*{5pt}} 
+\setcounter{secnumdepth}{5}
+
+\makeatletter 
+\renewcommand\@biblabel[1]{#1}            
+\renewcommand\@makefntext[1]% 
+{\noindent\makebox[0pt][r]{\@thefnmark\,}#1}
+\makeatother 
+\renewcommand{\figurename}{\small{Fig.}~}
+\sectionfont{\sffamily\Large}
+\subsectionfont{\normalsize}
+\subsubsectionfont{\bf}
+\setstretch{1.125} %In particular, please do not alter this line.
+\setlength{\skip\footins}{0.8cm}
+\setlength{\footnotesep}{0.25cm}
+\setlength{\jot}{10pt}
+\titlespacing*{\section}{0pt}{4pt}{4pt}
+\titlespacing*{\subsection}{0pt}{15pt}{1pt}
+%%%END OF PAGE SETUP%%%
+
+%%%FOOTER%%%
+\fancyfoot{}
+\fancyfoot[LO,RE]{\vspace{-7.1pt}\includegraphics[height=9pt]{head_foot/LF}}
+\fancyfoot[CO]{\vspace{-7.1pt}\hspace{13.2cm}\includegraphics{head_foot/RF}}
+\fancyfoot[CE]{\vspace{-7.2pt}\hspace{-14.2cm}\includegraphics{head_foot/RF}}
+\fancyfoot[RO]{\footnotesize{\sffamily{1--\pageref{LastPage} ~\textbar  \hspace{2pt}\thepage}}}
+\fancyfoot[LE]{\footnotesize{\sffamily{\thepage~\textbar\hspace{3.45cm} 1--\pageref{LastPage}}}}
+\fancyhead{}
+\renewcommand{\headrulewidth}{0pt} 
+\renewcommand{\footrulewidth}{0pt}
+\setlength{\arrayrulewidth}{1pt}
+\setlength{\columnsep}{6.5mm}
+\setlength\bibsep{1pt}
+%%%END OF FOOTER%%%
+
+%%%FIGURE SETUP - please do not change any commands within this section%%%
+\makeatletter 
+\newlength{\figrulesep} 
+\setlength{\figrulesep}{0.5\textfloatsep}
+
+\newcommand{\topfigrule}{\vspace*{-1pt}% 
+\noindent{\color{cream}\rule[-\figrulesep]{\columnwidth}{1.5pt}} }
+
+\newcommand{\botfigrule}{\vspace*{-2pt}% 
+\noindent{\color{cream}\rule[\figrulesep]{\columnwidth}{1.5pt}} }
+
+\newcommand{\dblfigrule}{\vspace*{-1pt}% 
+\noindent{\color{cream}\rule[-\figrulesep]{\textwidth}{1.5pt}} }
+
+\makeatother
+%%%END OF FIGURE SETUP%%%
+
+%%%TITLE, AUTHORS AND ABSTRACT%%%
+\twocolumn[
+  \begin{@twocolumnfalse}
+\vspace{3cm}
+\sffamily
+\begin{tabular}{m{4.5cm} p{13.5cm} }
+
+\includegraphics{head_foot/DOI} & \noindent\LARGE{\textbf{An upper bound to gas storage and delivery via pressure-swing adsorption in
+porous materials$^\dag$}} \\%Article title goes here instead of the text "This is the title"
+\vspace{0.3cm} & \vspace{0.3cm} \\
+
+ & \noindent\large{Jordan K. Pommerenck,$^{\ast}$\textit{$^{a}$} Cory M. Simon,\textit{$^{b\ddag}$} and David Roundy\textit{$^{a}$}} \\%Author names go here instead of "Full name", etc.
+
+\includegraphics{head_foot/dates} & \noindent\normalsize{The abstract should be a single paragraph which summarises the content of the article.
+Any references in the abstract should be written out in full \textit{e.g.}\ [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].} \\%The abstrast goes here instead of the text "The abstract should be..."
+
+\end{tabular}
+
+ \end{@twocolumnfalse} \vspace{0.6cm}
+
+  ]
+%%%END OF TITLE, AUTHORS AND ABSTRACT%%%
+
+%%%FONT SETUP - please do not change any commands within this section
+\renewcommand*\rmdefault{bch}\normalfont\upshape
+\rmfamily
+\section*{}
+\vspace{-1cm}
+
+%%%FOOTNOTES%%%
+
+\footnotetext{\textit{$^{a}$~Address, Address, Town, Country. Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
+\footnotetext{\textit{$^{b}$~Address, Address, Town, Country. }}
+
+%Please use \dag to cite the ESI in the main text of the article.
+%If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
+\footnotetext{\dag~Electronic Supplementary Information (ESI) available: [details of any supplementary information available should be included here]. See DOI: 00.0000/00000000.}
+%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required
+
+\footnotetext{\ddag~Additional footnotes to the title and authors can be included \textit{e.g.}\ `Present address:' or `These authors contributed equally to this work' as above using the symbols: \ddag, \textsection, and \P.
+Please place the appropriate symbol next to the author's name and include a \texttt{\textbackslash footnotetext} entry in the the correct place in the list.}
+
+%%%END OF FOOTNOTES%%%
+
+%%%MAIN TEXT%%%%
 \section{Introduction}
@@
 
-\section{Conclusion}
+\section{Conclusions}
 We have established an upper bound on the deliverable capacity via
@@
 Our code is available at \href{https://github.com/Jordan-Pommerenck/isothermal-gas-adsorption}{https://github.com/Jordan-Pommerenck/isothermal-gas-adsorption}.
+
+\section*{Conflicts of interest}
+In accordance with our policy on \href{http://www.rsc.org/journals-books-databases/journal-authors-reviewers/author-responsibilities/#code-of-conduct}{Conflicts of interest} please ensure that a conflicts of interest statement is included in your manuscript here.
+Please note that this statement is required for all submitted manuscripts.
+If no conflicts exist, please state that ``There are no conflicts to declare''.
+
+\section*{Acknowledgements}
+The Acknowledgements come at the end of an article after Conflicts of interest and before the Notes and references.
+
+%%%END OF MAIN TEXT%%%
+
+%The \balance command can be used to balance the columns on the final page if desired. It should be placed anywhere within the first column of the last page.
+
+\balance
+
+%If notes are included in your references you can change the title from 'References' to 'Notes and references' using the following command:
+%\renewcommand\refname{Notes and references}
+
+%%%REFERENCES%%%
+\bibliography{bibfile} %You need to replace "rsc" on this line with the name of your .bib file
+\bibliographystyle{rsc} %the RSC's .bst file
+
+\end{document}

### 2020-06-25 ising/ising-body.tex
edit of the discussion

@@
 such an optimized ensemble is doable
-
-% This goes in discussion section
-Last time for 32 X 32 is TL 2047388. Do I look at histogram to figure this out??
-But optimized ensemble requires the energy range???
-Also, SAD doesn't appear to be impacted by the larger energy range? histogram counts outside the interesting energy range are negligible compared with count in the range of interest.
 
@@
 
+Figure~\ref{fig:N128-cv-error} shows that the SAD algorithm converges
+signiticantly more slowly that the converging WL methods.
+%We interpret this as due to the
+%fact that the SAD simulation must take the time to discover the energy range
+%of interest dynamically.
+The SAD algorithm on average takes around
+\davidsays{$10^9$} moves to identify as important all \davidsays{8192}
+negative energy states, which is around \davidsays{twice as long} as
+the number of moves that
+the WL methods require in order to explore all the states in this
+energy range.  The main difference between the convergence of these
+methods is that the WL-based methods decrease $\gamma_t$ far more rapidly,
+which leads to more rapid convergence.  The Wang-Landau approaches can
+get away with this because the range of energies of interest is given
+as an input rather than an output, allowing a more aggressive schedule
+of reduction of $\gamma_t$.  This aggressive behavior is precisely what
+requires that Wang-Landau methods be followed by some correction stage
+(either $1/t$ or a production run) in order to correct residual errors.
+
 %% The two WL simulations followed by a production run highlight that

### 2020-06-29 ising/referee-response.md
update the referee response to be consistent with our paper edits.

@@
 flat-histogram methods. (ii) It is important to test SAD's convergence
-properties on the 2D Ising where the range of energies is easily known (thus
-biasing in favor of WL methods). ...
+properties on the 2D Ising where the range of energies is easily known.
 
@@
 and then finding the fraction of moves (in the end) that are outside the range
-of interesting energies.
+of interesting energies. We find that SAD histogram counts outside the interesting
+energy range are negligible. We mention this in the discussion paragraph.
 
@@
 convergence properties on a different type of system where the range of
-energies is easily known (thus biasing in favor of WL methods). Also in the
-case of the 128 X 128 2D Ising system, the computation time is quite lengthy
-(more than 6 months for some calculations) representing a superior and robust
-comparison result.
+energies is easily known (we find that this does not measurably bias in favor
+of WL methods). Also in the case of the 128 X 128 2D Ising system, the
+computation time is quite lengthy (more than 6 months for some calculations)
+representing a superior and robust comparison result.
 
 In response to the second referee's comments, we have better highlighted these
-important points in the abstract and conclusion.
+important points in the abstract, discussion paragraph, and conclusion.

### 2020-06-29 ising/ising-body.tex
made ising paper edits and added a few more citations on replica exchange.

@@
 % optimized ensemble paragraph
-Nadler \emph{et al.} and Trebst \emph{et al.} systematically examined optimized
-ensembles due to performance issues flat-histogram methods faced when
-confronted with hidden energy barriers~\cite{nadler2007generalized,
-nadler2007dynamics, trebst2004optimizing}. Trebst \emph{et al.} used their
-optimized ensemble to systematically set the temperature range(s).
+A problem that flat-histogram methods face is that the convergence rate is
+impacted by energy barriers and bottleneck which make traversing the phase
+space difficult. Nadler \emph{et al.} and Trebst \emph{et al.} systematically
+examined optimized ensembles due to performance issues flat-histogram methods
+faced when confronted with hidden energy barriers~\cite{nadler2007generalized,
+nadler2007dynamics, trebst2004optimizing}. The goal of these methods is to make
+the statistical errors uniform such that low energy errors do not dominate over
+high energy errors.
 
 % replica exchange paragraph
-Replica exchange Wang-Landau (REWL) was introduced~\cite{vogel2013rewl}.
-Replica-exchange Wang-Landau has been developed with great
-success~\cite{perera2016magnetic} and adapting flat-histogram methods to use
-such an optimized ensemble is doable
+
+Parallel tempering (replica exchange) is a powerful method for overcoming
+energy barriers~\cite{geyer1995annealing, hukushima1996exchange,
+hansmann1997parallel}. The method works by breaking the simulation up into many
+`replicas' and simulating in parallel. At certain times, neighboring replicas
+are allowed to exchange which enables the phase space to be more evenly
+explored.
+
+Flat-histogram methods, such as Replica exchange Wang-Landau
+(REWL)~\cite{vogel2013rewl,perera2016magnetic}, have been proposed that combine
+histogram and replica exchange methods. This demonstrates that flat-histogram
+methods can be extended to non-flat-histogram methods.
 
@@
 Figure~\ref{fig:N128-cv-error} shows that the SAD algorithm converges
-signiticantly more slowly that the converging WL methods.
-%We interpret this as due to the
-%fact that the SAD simulation must take the time to discover the energy range
-%of interest dynamically.
+significantly more slowly than the converging WL methods.
 The SAD algorithm on average takes around
-\davidsays{$10^9$} moves to identify as important all \davidsays{8192}
-negative energy states, which is around \davidsays{twice as long} as
+$10^9$ moves to identify as important all 8192
+negative energy states, which is around twice as long as
 the number of moves that
@@
 
-%% The two WL simulations followed by a production run highlight that
-%% choosing too large a $\gamma^{\text{WL}}_t$ will result in poor convergence
-%% properties.
-
 \section{Conclusion}
@@
 will take proportionately more moves to converge to the same value as
-$1/t$-WL as system size is increased.  However, the WL methods are
-solving a simpler problem because they are given the energy range
+$1/t$-WL as system size is increased.  While the WL methods are
+are given the energy range
 \emph{a priori}, rather than a temperature range of interest such as
-SAD requires.  The 2D Ising model is a system where the energy range
-is known \emph{a priori.}, which makes a comparison between WL methods
-and SAD more biased in favor of WL methods. For the general case in
+SAD requires, we find that the SAD histogram counts for energies outside the range of interest are negligible. For the general case in
 which the energy range of the system is not known and where a range of
-desired temperatures is known, this will make the SAD method
+desired temperatures is known, the SAD method is
 considerably more convenient, and quite possibly more efficient than a
-process involving multiple simulations to determine the energy range
+process involving multiple simulations to determine an energy range
 of interest. This work also dramatically demonstrates that WL followed
@@
 determine the parameters is the only way WL should be used, this research
-represents the first conclusive comparison among all of these
+represents the first detailed comparison among all of these
 flat-histogram methods.

### 2020-07-02 ising/ising-body.tex
edit intro paragraphs

@@
 % optimized ensemble paragraph
-A problem that flat-histogram methods face is that the convergence rate is
-impacted by energy barriers and bottleneck which make traversing the phase
+Another challenge that flat-histogram methods face is that the convergence rate is
+impacted by energy barriers and bottleneck which can make traversing the phase
 space difficult. Nadler \emph{et al.} and Trebst \emph{et al.} systematically
-examined optimized ensembles due to performance issues flat-histogram methods
-faced when confronted with hidden energy barriers~\cite{nadler2007generalized,
-nadler2007dynamics, trebst2004optimizing}. The goal of these methods is to make
-the statistical errors uniform such that low energy errors do not dominate over
-high energy errors.
+examined optimized ensembles to address performance issues flat-histogram methods
+face when confronted with hidden energy barriers~\cite{nadler2007generalized,
+nadler2007dynamics, trebst2004optimizing}.
+The approach involves using an ensemble that does \emph{not} result in a
+flat histogram, but instead optimizes the rate of diffusion between low
+energy and high energy states.
+The goal of these methods is to make the statistical errors uniform.
+%
+Optimized ensemble methods typically begin by using a flat-histogram method
+to get a first approximation for the weights, which means that any of the
+methods
+tested in this work could be used as a starting point for an optimized
+ensemble simulation.
 
@@
 
-Parallel tempering (replica exchange) is a powerful method for overcoming
-energy barriers~\cite{geyer1995annealing, hukushima1996exchange,
-hansmann1997parallel}. The method works by breaking the simulation up into many
-`replicas' and simulating in parallel. At certain times, neighboring replicas
-are allowed to exchange which enables the phase space to be more evenly
-explored.
-
-Flat-histogram methods, such as Replica exchange Wang-Landau
-(REWL)~\cite{vogel2013rewl,perera2016magnetic}, have been proposed that combine
-histogram and replica exchange methods. This demonstrates that flat-histogram
-methods can be extended to non-flat-histogram methods.
+An approach to parallelizing flat-histogram Monte Carlo methods is
+the replica-exchange approach, which was pioneered by parallel tempering
+algorithms~\cite{geyer1995annealing, hukushima1996exchange,
+hansmann1997parallel}.  Vogel \emph{et al.} adapted this approach to develop
+the Replica Exchange Wang-Landau (REWL)~\cite{vogel2013rewl} approach.  This
+approach for parallelization is sufficiently simple and general that it could
+equally be applied to any of the methods explored in this paper.
 

### 2020-07-02 ising/referee-response.md
refine referee response

--- 2020-06-20 ising/referee-response.md
@@
 
-We appreciate the referees' careful reading of our paper, and feel that their
-feedback has led to a significant improvement. We have addressed all of the
-first referee's comments by added two paragraphs to the introduction and a
-section in the paper to 'Address the general problem of all "flat-histogram"
-methods'. We also strengthen the abstract to address the concerns raised by the
-second referee. We point out that: (i) Our paper is the first of its kind (to
-the best of our knowledge) to compare ``production run'' WL with other
-flat-histogram methods. (ii) It is important to test SAD's convergence
-properties on the 2D Ising where the range of energies is easily known (thus
-biasing in favor of WL methods). ...
+We appreciate the referees' careful reading of our paper, and feel
+that their feedback has led to a significant improvement. We have
+addressed all of the first referee's comments by adding two paragraphs
+to the introduction, and a section in the paper to 'Address the
+general problem of all "flat-histogram" methods'. We also strengthen
+the abstract to address the concerns raised by the second referee. We
+point out that:
+
+(1) It is relevant to test SAD's convergence properties on the 2D
+    Ising where the range of energies is easily known.  This system in
+    particular is a kind of worst-case scenario for SAD, and our
+    previous paper focused on systems where SAD was more likely to be
+    applied in practice.
+
+(2) Our paper is the first of its kind (to the best of our knowledge)
+    to compare ``production run'' WL with other recent flat-histogram
+    methods.  This is particularly relevant, as it is common in this
+    field to discount WL as "not converging to the true answer" (as we
+    havepreviously done ourselves, to our chagrin), without
+    recognizing that when it is used in practice it is followed by a
+    "production run", which has the effect of causing the the
+    simulation to converge to the true answer.
 
 Response to first referee
-# -----------------------------------------------------------------------------#
+# --------------------------------------------------------------------#
 
-The first referee points out that 'Simulation set-up and analysis protocols are
-sound and give confidence in the quality of the presented data and
-conclusions.'
+The first referee points out that 'Simulation set-up and analysis
+protocols are sound and give confidence in the quality of the
+presented data and conclusions.'
 
@@
 
-(i) "It would be worthwhile to work out the [slower convergence for SAD due to
-the energy range being an a priori parameter in WL/MUCA] in the paper, and, if
-possible, quantify the loss in efficiency coming from energy range
-determination."
+(i) "It would be worthwhile to work out this point in the paper
+[slower convergence for SAD due to the energy range being an a priori
+parameter in WL/MUCA], and, if possible, quantify the loss in
+efficiency coming from energy range determination."
 
-We have added a paragraph in the discussion section that explains the loss in
-efficiency SAD suffers by not specifying an energy range a priori. We are able
-to quantify this by identifying the last time SAD finds a new important energy
-and then finding the fraction of moves (in the end) that are outside the range
-of interesting energies.
+We have added a paragraph in the results section that explains the
+loss in efficiency SAD suffers by not specifying an energy range a
+priori. We explain that the fraction of moves (in the end) that are
+outside the range of interesting energies is negligible, and the time
+spent before all interesting energies have been identified is not
+*much* longer than the time WL takes to find all energies, but that
+SAD is more conservative than WL after this point in decreasing
+gamma.  We point out that standard WL requires a production run (or
+the 1/t variant) in order to prevent *overly* agresssively decreasing
+gamma, so SAD and WL are erring in opposite directions.
 
-(ii) "Address the general problem of all "flat histogram" methods that the
-performance of these methods is due to hidden barriers almost always lower than
-what one would expect from a random walk in energy (in most cases still
-exponentially growing). This has been discussed by Nadler et al (PRE 75 (2007)
-1) who also discus a protocol for optimizing the ensemble which minimizes the
-residual exponential slowing down. The authors should point out if and how SAD
-can be adapted to such "non-flat-histogram" methods."
+(ii) "Address the general problem of all "flat histogram" methods that
+the performance of these methods is due to hidden barriers almost
+always lower than what one would expect from a random walk in energy
+(in most cases still exponentially growing). This has been discussed
+by Nadler et al (PRE 75 (2007) 1) who also discus a protocol for
+optimizing the ensemble which minimizes the residual exponential
+slowing down. The authors should point out if and how SAD can be
+adapted to such "non-flat-histogram" methods."
 
-We have added a paragraph that 'Address the general problem of all "flat
-histogram" methods' in the introduction. We discuss how the performance is
-impacted due to hidden barriers. We also add that WL has already been adapted
-to replica-exchange and point out that this should be possible for SAD as well.
+We have added a paragraph that addresses the optimized ensemble
+approach.  Our understanding is that any of these flat-histogram
+methods could be followed by the optimized ensemble method, which
+requires as a starting point a set of weights that allow efficient
+sampling of the energy range of interest.
 
-(iii) Add a few short remarks on comparison with replica exchange methods would
-be helpful.
+(iii) Add a few short remarks on comparison with replica exchange
+methods would be helpful.
 
-We have added a 'few short remarks' on comparison with replica exchange in the
-introduction. WL has been formulated as replica-exchange WLMC whereby it adopts
-the concept of conformational swapping directly from parallel tempering. Each replica
-is allowed to traverse the entire energy range via overlapping energy windows.
-
-While we have performed limited comparison with replica exchange methods (in
-our previous work we found the heat capacity to be consistent with computations
-performed with replica exchange), SAD could be extended in much the same way
-that WL was. In fact, this would make an interesting comparison with replica
-exchange statistical temperature Monte Carlo (as they both specify a
-temperature range) and should be studied on a temperature range system in a
-future work.
+We added a paragraph very briefly describing how replica exchange has
+been adapted to WL, and make the point that the same adaption could
+apply to any of these very similar flat-histogram methods.
 
 Response to second referee
-# -----------------------------------------------------------------------------#
+# ------------------------------------------------------------------#
 
-We thank the second referee for bringing to our attention two points so that we
-can better address them in our paper and make them completely clear.
+We thank the second referee for bringing to our attention two points
+that we can better clarify in our paper.
 
@@
 
-Our paper is the first of its kind (to the best of our knowledge) to compare
-``production run'' WL with other flat-histogram methods. The first (not the
-second) referee notes rather strongly that this is the only way that WL should
-be done; however, no research work has ever compared this with other
-flat-histogram methods and certainly not with the completeness shown here.
-Also, the very fact that it is compared with pure WL is proof for the first
-referee's statement that "this is how WL should be done" since it performs
-superior to pure WL.
+Our paper is the first of its kind (to the best of our knowledge) to
+compare ``production run'' WL with other flat-histogram methods. The
+first (not the second) referee notes rather strongly that this is the
+only way that WL should be done; however, no research work has ever
+compared this with other flat-histogram methods and certainly not with
+the completeness shown here.  This contribution alone makes our paper
+valuable, even were SAD not included in the mix.
 
-We believe that this paper helps by bridging the disconnect between *users* of
-WL and developers of improved flat-histogram methods. The majority of
-developers use pure WL when comparing multiple flat-histogram methods since
-there is no direct comparison of WL followed by a production run with methods
-such as 1/t-WL or SAMC.
+We believe that this paper helps by bridging the disconnect between
+*users* of WL and developers of improved flat-histogram methods. The
+majority of developers use pure WL when comparing multiple
+flat-histogram methods since there is no direct comparison of WL
+followed by a production run with methods such as 1/t-WL or SAMC.
 
-(ii) "It is not clear[ly] understood why the current calculations were not
-included in the original article. Although the results presented by the
-[authors] are original, they are only a simple extension of the first work, and
-do not meet the standards of PRE readers."
+(ii) "It is not clear[ly] understood why the current calculations were
+not included in the original article. Although the results presented
+by the [authors] are original, they are only a simple extension of the
+first work, and do not meet the standards of PRE readers."
 
-The second referee asks an important question here to which there are two
-answers. We only became aware of "production run" WL as a result of the
-response to our first paper. As mentioned (hinted at) in the acknowledgement,
-Johannes Zierenberg emailed us after reading our paper and suggested that we
-test WL followed by a production run with a number of flat-histogram methods.
-Due to the lack of literature detailing its performance vs other flat-histogram
-methods, the 2D Ising model seems the logical starting point for such a
-comparison.
+The second referee asks an important question here to which there are
+two answers. We only became aware of "production run" WL as a result
+of the response to our first paper. As mentioned (hinted at) in the
+acknowledgement, Johannes Zierenberg emailed us after reading our
+paper and suggested that we test WL followed by a production run with
+a number of flat-histogram methods.  Due to the lack of literature
+detailing its performance vs other flat-histogram methods, the 2D
+Ising model seems the logical starting point for such a comparison.
 
-Second, SAD was initially tested on methods where it could benefit from knowing
-the temperature range of interest in advance. It is critical to test it's
-convergence properties on a different type of system where the range of
-energies is easily known (thus biasing in favor of WL methods). Also in the
-case of the 128 X 128 2D Ising system, the computation time is quite lengthy
-(more than 6 months for some calculations) representing a superior and robust
-comparison result.
+Second, SAD was initially tested on methods where it could benefit
+from knowing the temperature range of interest in advance.  It made
+sense in the first paper to focus on systems in which the energy range
+of interest was *not* easily known as these systems are both more
+common and are the kind of systems in which we expect SAD to be
+applied.  Nevertheless, it is still valuable to test SAD's convergence
+properties on a system where the range of energies is easily known.
 
-In response to the second referee's comments, we have better highlighted these
-important points in the abstract and conclusion.
+We also point out that these simulations are time-consuming.  Some of
+those in the current paper ran for more than 6 months.  It is not
+reasonable to expect us to complete all possible simulations before
+publication of a paper.
+
+In response to the second referee's comments, we have better
+highlighted these two important points in the abstract, discussion
+paragraph, and conclusion.

### 2020-07-08 gas-adsorption/rsc-gas-ads-complete.tex
fix rsc paper style to prevent overflow.

@@
 \usepackage{balance}
-\usepackage{times,mathptmx}
+\usepackage{mathptmx}
 \usepackage{sectsty}
@@
 %%%Please don't disable any packages in the preamble, as this may cause the template to display incorrectly.%%%
+
+%\usepackage{epstopdf}%This line makes .eps figures into .pdf - please comment out if not required.
+\definecolor{cream}{RGB}{222,217,201}
 
@@
 %%%HEADER%%%
-\fancyhead[C]{\includegraphics[width=18.5cm]{head_foot/header_bar}}
-\fancyhead[L]{\hspace{0cm}\vspace{1.5cm}\includegraphics[height=30pt]{head_foot/journal_name}}
-\fancyhead[R]{\hspace{0cm}\vspace{1.7cm}\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}
+% \fancyhead[C]{\includegraphics[width=18.5cm]{head_foot/header_bar}}
+% \fancyhead[L]{\hspace{0cm}\vspace{1.5cm}\includegraphics[height=30pt]{head_foot/journal_name}}
+% \fancyhead[R]{\hspace{0cm}\vspace{1.7cm}\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}
 \renewcommand{\headrulewidth}{0pt}
@@
   \begin{@twocolumnfalse}
-\vspace{3cm}
+{\includegraphics[height=30pt]{head_foot/journal_name}\hfill\raisebox{0pt}[0pt][0pt]{\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}\\[1ex]
+\includegraphics[width=18.5cm]{head_foot/header_bar}}\par
+\vspace{1em}
 \sffamily

### 2020-07-08 gas-adsorption/rsc-gas-ads-complete.tex
add appendix; still need to change citations to long form.

@@
 
-\includegraphics{head_foot/dates} & \noindent\normalsize{The abstract should be a single paragraph which summarises the content of the article.
-Any references in the abstract should be written out in full \textit{e.g.}\ [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].} \\%The abstrast goes here instead of the text "The abstract should be..."
+\includegraphics{head_foot/dates} & \noindent\normalsize{Due to their low volumetric energy density at ambient conditions, both hydrogen
+and natural gas are challenging to economically store onboard vehicles as
+fuels~\cite{mason2014evaluating, sircar2002pressure}. One strategy to densify
+these gases is to pack the fuel tank with a porous
+material~\cite{schoedel2016role}. Metal-organic frameworks (MOFs) are tunable,
+nanoporous materials with large internal surface areas and show considerable
+promise for densifying gases~\cite{makal2012methane,mason2014evaluating,
+suh2011hydrogen,garcia2018benchmark, schoedel2016role}. The US Department of
+Energy (DOE) has set volumetric deliverable capacity
+targets~\cite{simon2015materials, h2targetsDOE} which, if met, would help
+enable commercial adoption of hydrogen/natural gas as transportation fuels.
+Many have attempted to establish theoretical upper bounds on the deliverable
+capacity via pressure-swing adsorption using simplified models of the
+gas-substrate and gas-gas interaction~\cite{gomez2014exploring,
+gomez2017impact, kaija2018high, lee2019predicting}; none have established a
+rigorous upper bound on the deliverable capacity. Here, we present a
+theoretical upper bound on the deliverable capacity of a gas in a rigid
+material via an isothermal pressure-swing. To provide an extremum, we consider
+a substrate that provides a spatially uniform potential energy field for the
+gas. Our bound is unique in that it directly relies on experimentally measured
+properties of the (real) bulk gas, without making approximations. We conclude
+that the goals set by the US DOE for room-temperature natural gas and hydrogen
+storage are theoretically possible, but sufficiently close to the upper bound
+as to be impractical for any real, rigid porous material. However, limitations
+to the scope of applicability of our upper bound guide fuel tank design and
+future material development. Firstly, one could heat the adsorbent to drive off
+trapped, residual gas in the adsorbent~\cite{gomez2014exploring}. Secondly, the
+physics of our upper bound do not pertain to any material that changes its
+structure in response to adsorbed gas, suggesting that flexible materials could
+still satisfy the DOE targets~\cite{schneemann2014flexible, choi2008broadly,
+mason2015methane}}.
+
+%Any references in the abstract should be written out in full \textit{e.g.}\ [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].} \\%The abstrast goes here instead of the text "The abstract should be..."
 

### 2020-07-09 gas-adsorption/rsc-gas-ads-complete.tex
reorder sections a bit, and move figures around a bit

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\columnwidth]{Combine_Figure.pdf}
+    \begin{minipage}{0.25\textwidth}
+    \phantomsubcaption{\label{fig:example_MOF}}
+    \end{minipage}
+    \begin{minipage}{0.25\textwidth}
+    \phantomsubcaption{\label{fig:delcap}}
+    \end{minipage}
+
+    \caption{\label{fig:fig1} Gas storage and delivery using metal-organic frameworks (MOFs).
+(\subref{fig:example_MOF}) the crystal structure of an archetype MOF, CuBTC \cite{chui1999chemically}.
+(\subref{fig:delcap}) the methane adsorption isotherm in CuBTC \cite{chui1999chemically} at 298 K (blue) (data from Ref.~\cite{mason2014evaluating}).
+The deliverable capacity $\rho_D$ is illustrated as the density of gas in the MOF at the storage pressure $\pfull$ minus the density at the discharge pressure $\pempty$.}
+\end{figure}
+
 For a vehicle employing a porous material to store natural gas or hydrogen, the
@@
 % \end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\columnwidth]{Combine_Figure.pdf}
-    \begin{minipage}{0.25\textwidth}
-    \phantomsubcaption{\label{fig:example_MOF}}
-    \end{minipage}
-    \begin{minipage}{0.25\textwidth}
-    \phantomsubcaption{\label{fig:delcap}}
-    \end{minipage}
-
-    \caption{\label{fig:fig1} Gas storage and delivery using metal-organic frameworks (MOFs).
-(\subref{fig:example_MOF}) the crystal structure of an archetype MOF, CuBTC \cite{chui1999chemically}.
-(\subref{fig:delcap}) the methane adsorption isotherm in CuBTC \cite{chui1999chemically} at 298 K (blue) (data from Ref.~\cite{mason2014evaluating}).
-The deliverable capacity $\rho_D$ is illustrated as the density of gas in the MOF at the storage pressure $\pfull$ minus the density at the discharge pressure $\pempty$.}
-\end{figure}
 
@@
 
-\section{An upper bound on the deliverable capacity of a pure gas in a rigid porous material}\label{sec:upper-bound}
+\section{Methods}
+\subsection{An upper bound on the deliverable capacity}\label{sec:upper-bound}
+We now develop a thermodynamic framework that places an upper bound on the deliverable capacity of any pure gas in a rigid porous material.
+
+The thermodynamic properties of a bulk, pure gas are characterized by an
+equation of state. Of particular interest for gas storage and delivery is the
+density of the gas $\rho_g(\mu,T)$ as a function of chemical potential $\mu$
+and temperature $T$, is shown for methane gas in
+Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K. For comparison, we also show
+the density of methane adsorbed into several porous materials.
+
 \begin{figure}
@@
 
-We now develop a thermodynamic framework that places an upper bound on the deliverable capacity of any pure gas in a rigid porous material.
+  To place an upper bound on the deliverable capacity, consider a substrate whose
+sole interaction with the gas is to introduce a spatially uniform potential
+energy $\V$ for gas molecules in the control volume (where $\V<0$ for an
+attractive potential). Because this interaction is uniform within the
+substrate, the gas-gas interactions and thus fluid structure in such a
+substrate are identical to those of the pure gas at the same density and
+temperature. This allows us to obtain the adsorption properties of this model
+substrate using the experimentally measured properties of the pure
+gas~\cite{nist}. Intuitively, the deliverable capacity of a gas in such a
+homogeneous substrate with the optimal potential energy is an upper bound
+because, in a real material, (i) spatial inhomogeneity of the potential
+energy results in some points in the control volume offering a suboptimal
+attraction for the gas and (ii) the atoms required to create the potential
+field exclude gas from occupying a fraction of the control volume.
 
-The thermodynamic properties of a bulk, pure gas are characterized by an
-equation of state. Of particular interest for gas storage and delivery is the
-density of the gas $\rho_g(\mu,T)$ as a function of chemical potential $\mu$
-and temperature $T$, is shown for methane gas in
-Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K. For comparison, we also show
-the density of methane adsorbed into several porous materials.
+The spatially uniform potential $\V$ representing gas-substrate interactions
+behaves as an external potential and effectively shifts the chemical potential
+(or, equivalently, molar Gibbs free energy) of the gas in the material, just as
+gravitational potential energy causes the density of air to vary with altitude.
+Consequently, the density of gas in our homogeneous substrate is:
+\begin{align}
+    \rho(\mu,T) &= \rho_g(\mu - \V,T). \label{eq:mof-density}
+\end{align}
+See Sec.~\ref{sec:V_shifts_chem_pot} for a derivation.
+
+The deliverable capacity of gas in our homogeneous substrate is thus:
+\begin{align}
+    \rho_D(\V,T) &= \rho(\mufull,T) - \rho(\muempty,T),
+    \label{eq:DofPhi}
+    \\
+    &= \rho_g(\mufull-\V,T) - \rho_g(\muempty-\V,T),
+\end{align}
+where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the
+pressures $\pfull$ and $\pempty$, respectively. Thus, $\rho_D$ for our
+homogeneous substrate as a function of $\V$ is the difference between two
+shifted versions of $\rho_g(\mu; T)$. Figure~\ref{fig:methane-298-D} shows the
+two shifted bulk methane density curves and their difference. The horizontal
+axis is the difference in molar Gibbs free energy of the gas, at fixed density (the density in the adsorbent) and temperature, between the gas within the substrate and the pure gas (\gst),
+which is equal to $\V$ in our model, and is pressure-dependent in a real
+crystal~(see Sec.~\ref{sec:phi-is-delta-g}). We see a potential $\V$ that
+maximizes the deliverable capacity of our ideal substrate, balancing the need
+to maximize the density at $\pfull$ against the need to minimize residual gas
+retained at $\pempty$.
+
+The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar,
+$\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L,
+achieved for $\V_{opt} =$ 5.9\ kJ/mol.
+
+An essential question is whether our idealized substrate, with spatially
+uniform potential $\V_{opt}$ for the gas, places an upper bound on the
+deliverable capacity in all rigid porous materials. In
+Section~\ref{sec:proof-extremum}, we show that our idealized substrate with
+spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the
+deliverable capacity over all possible \emph{static potential energy fields},
+provided the gas does not crystallize at the temperature and in the density
+range of interest. The potential energy field is \emph{static} if it is
+unaffected by the presence of the gas; consequently, our model does not apply
+to flexible materials that undergo gas-induced conformation
+changes~\cite{schneemann2014flexible}. We next argue that this extremum is a
+maximum by addressing two alternative possibilities: the extremum could turn
+out to be (i) a saddle point or (ii) a local, rather than global, maximum.
+
+Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum
+$\rho_D$ over all \emph{spatially uniform}, static potential energy fields. To
+qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a
+maximum deliverable capacity over \emph{all} (including non-spatially uniform)
+static potential energy fields (a broader claim), consider a
+non-spatially-uniform variation on $\V_{opt}$. The effect of
+this variation is to reduce the deliverable capacity: At low adsorbed gas
+densities, the lowest-energy positions will be preferentially occupied, while
+at higher densities, the additional gas molecules will be forced into higher
+energy locations. Thus, the mean attraction will be lowest at \pfull\ and
+highest at \pempty. As a result, $\gst$ increases monotonically with
+pressure. As shown in Sec.~\ref{sec:monotonic}, this results in
+the deliverable capacity of a non-uniform potential with a given $\gst$ at $\pfull$ or $\pempty$ being lower than the
+deliverable capacity of a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
+deliverable capacity amongst all static potential energy fields that give the
+same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will yield a
+greater deliverable capacity over any static, non-uniform potential.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{methane-298-n-vs-G}
+    \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
+    Experimental deliverable capacities for several porous materials (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line.}
+    \label{fig:methane-298-D}
+\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.95\columnwidth]{four-cases-pro}
-\caption{A bulk gas reservoir is connected to a volume filled with adsorbent material (green/blue: material atoms, red: gas particles).
-The adsorbent is also connected, but with a closed valve, to an equal volume containing the ideal substrate (shaded blue background represents the uniform potential energy field $\V$) with an equal density of gas as in the adsorbent.
-The top (bottom) shows a realization of this connected system where the pressure of the bulk gas is $\pfull$ ($\pempty$), and thus the density of gas in the adsorbent is $\rhofull$ ($\rhoempty$) by definition.
-The deliverable capacity of the adsorbent is $\rhofull - \rhoempty$.
-We can determine if the deliverable capacity of gas in the ideal substrate is lower or higher than in the adsorbent by considering what happens when we open the valves connecting the two pairs of volumes containing adsorbent and the ideal substrate.}
-\label{fig:delta-gst-maximum}
-\end{figure}
-
 \subsection{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
@@
 substrate with $\V=\gst$, even for a value of $\V$ that is not optimal.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{four-cases-pro}
+    \caption{A bulk gas reservoir is connected to a volume filled with adsorbent material (green/blue: material atoms, red: gas particles).
+The adsorbent is also connected, but with a closed valve, to an equal volume containing the ideal substrate (shaded blue background represents the uniform potential energy field $\V$) with an equal density of gas as in the adsorbent.
+The top (bottom) shows a realization of this connected system where the pressure of the bulk gas is $\pfull$ ($\pempty$), and thus the density of gas in the adsorbent is $\rhofull$ ($\rhoempty$) by definition.
+The deliverable capacity of the adsorbent is $\rhofull - \rhoempty$.
+We can determine if the deliverable capacity of gas in the ideal substrate is lower or higher than in the adsorbent by considering what happens when we open the valves connecting the two pairs of volumes containing adsorbent and the ideal substrate.}
+    \label{fig:delta-gst-maximum}
+\end{figure}
 
@@
 % End here %
-
-To place an upper bound on the deliverable capacity, consider a substrate whose
-sole interaction with the gas is to introduce a spatially uniform potential
-energy $\V$ for gas molecules in the control volume (where $\V<0$ for an
-attractive potential). Because this interaction is uniform within the
-substrate, the gas-gas interactions and thus fluid structure in such a
-substrate are identical to those of the pure gas at the same density and
-temperature. This allows us to obtain the adsorption properties of this model
-substrate using the experimentally measured properties of the pure
-gas~\cite{nist}. Intuitively, the deliverable capacity of a gas in such a
-homogeneous substrate with the optimal potential energy is an upper bound
-because, in a real material, (i) spatial inhomogeneity of the potential
-energy results in some points in the control volume offering a suboptimal
-attraction for the gas and (ii) the atoms required to create the potential
-field exclude gas from occupying a fraction of the control volume.
-
-The spatially uniform potential $\V$ representing gas-substrate interactions
-behaves as an external potential and effectively shifts the chemical potential
-(or, equivalently, molar Gibbs free energy) of the gas in the material, just as
-gravitational potential energy causes the density of air to vary with altitude.
-Consequently, the density of gas in our homogeneous substrate is:
-\begin{align}
-    \rho(\mu,T) &= \rho_g(\mu - \V,T). \label{eq:mof-density}
-\end{align}
-See Sec.~\ref{sec:V_shifts_chem_pot} for a derivation.
-
-The deliverable capacity of gas in our homogeneous substrate is thus:
-\begin{align}
-    \rho_D(\V,T) &= \rho(\mufull,T) - \rho(\muempty,T),
-    \label{eq:DofPhi}
-    \\
-    &= \rho_g(\mufull-\V,T) - \rho_g(\muempty-\V,T),
-\end{align}
-where $\mufull$ and $\muempty$ are the chemical potentials corresponding to the
-pressures $\pfull$ and $\pempty$, respectively. Thus, $\rho_D$ for our
-homogeneous substrate as a function of $\V$ is the difference between two
-shifted versions of $\rho_g(\mu; T)$. Figure~\ref{fig:methane-298-D} shows the
-two shifted bulk methane density curves and their difference. The horizontal
-axis is the difference in molar Gibbs free energy of the gas, at fixed density (the density in the adsorbent) and temperature, between the gas within the substrate and the pure gas (\gst),
-which is equal to $\V$ in our model, and is pressure-dependent in a real
-crystal~(see Sec.~\ref{sec:phi-is-delta-g}). We see a potential $\V$ that
-maximizes the deliverable capacity of our ideal substrate, balancing the need
-to maximize the density at $\pfull$ against the need to minimize residual gas
-retained at $\pempty$.
-
-The optimal deliverable capacity of methane (298\ K, $\pfull=$ 65\ bar,
-$\pempty=$ 5.8\ bar) in our ideal, spatially uniform substrate is 374\ L STP/L,
-achieved for $\V_{opt} =$ 5.9\ kJ/mol.
-
-An essential question is whether our idealized substrate, with spatially
-uniform potential $\V_{opt}$ for the gas, places an upper bound on the
-deliverable capacity in all rigid porous materials. In
-Section~\ref{sec:proof-extremum}, we show that our idealized substrate with
-spatially uniform potential $\V_{opt}$ yields an \emph{extremum} of the
-deliverable capacity over all possible \emph{static potential energy fields},
-provided the gas does not crystallize at the temperature and in the density
-range of interest. The potential energy field is \emph{static} if it is
-unaffected by the presence of the gas; consequently, our model does not apply
-to flexible materials that undergo gas-induced conformation
-changes~\cite{schneemann2014flexible}. We next argue that this extremum is a
-maximum by addressing two alternative possibilities: the extremum could turn
-out to be (i) a saddle point or (ii) a local, rather than global, maximum.
-
-Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum
-$\rho_D$ over all \emph{spatially uniform}, static potential energy fields. To
-qualitatively argue that the spatially uniform potential $\V_{opt}$ provides a
-maximum deliverable capacity over \emph{all} (including non-spatially uniform)
-static potential energy fields (a broader claim), consider a
-non-spatially-uniform variation on $\V_{opt}$. The effect of
-this variation is to reduce the deliverable capacity: At low adsorbed gas
-densities, the lowest-energy positions will be preferentially occupied, while
-at higher densities, the additional gas molecules will be forced into higher
-energy locations. Thus, the mean attraction will be lowest at \pfull\ and
-highest at \pempty. As a result, $\gst$ increases monotonically with
-pressure. As shown in Sec.~\ref{sec:monotonic}, this results in
-the deliverable capacity of a non-uniform potential with a given $\gst$ at $\pfull$ or $\pempty$ being lower than the
-deliverable capacity of a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
-deliverable capacity amongst all static potential energy fields that give the
-same $\gst$, and the \emph{optimal} uniform potential $\V_{opt}$ will yield a
-greater deliverable capacity over any static, non-uniform potential.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\columnwidth]{methane-298-n-vs-G}
-    \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
-    Experimental deliverable capacities for several porous materials (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line.}
-    \label{fig:methane-298-D}
-\end{figure}
 

### 2020-07-09 gas-adsorption/rsc-gas-ads-complete.tex
more edits, including abstract rewrite

--- 2020-07-08 gas-adsorption/rsc-gas-ads-complete.tex
@@
 \includegraphics{head_foot/DOI} & \noindent\LARGE{\textbf{An upper bound to gas storage and delivery via pressure-swing adsorption in
-porous materials$^\dag$}} \\%Article title goes here instead of the text "This is the title"
+porous materials}} \\%Article title goes here instead of the text "This is the title"
 \vspace{0.3cm} & \vspace{0.3cm} \\
 
- & \noindent\large{Jordan K. Pommerenck,$^{\ast}$\textit{$^{a}$} Cory M. Simon,\textit{$^{b\ddag}$} and David Roundy\textit{$^{a}$}} \\%Author names go here instead of "Full name", etc.
+ & \noindent\large{Jordan K. Pommerenck,$^{\ast}$\textit{$^{a}$} Cory M. Simon,\textit{$^{b}$} and David Roundy\textit{$^{a}$}} \\%Author names go here instead of "Full name", etc.
 
-\includegraphics{head_foot/dates} & \noindent\normalsize{The abstract should be a single paragraph which summarises the content of the article.
-Any references in the abstract should be written out in full \textit{e.g.}\ [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].} \\%The abstrast goes here instead of the text "The abstract should be..."
+\includegraphics{head_foot/dates} & \noindent\normalsize{
+    Both hydrogen
+and natural gas are challenging to economically store onboard vehicles as
+fuels, due to their low volumetric energy density at ambient conditions.
+One strategy to densify
+these gases is to pack the fuel tank with a porous
+adsorbent material.
+% 
+The US Department of
+Energy (DOE) has set volumetric deliverable capacity
+targets which, if met, would help
+enable commercial adoption of hydrogen/natural gas as transportation fuels.
+%
+%Many have attempted to establish theoretical upper bounds on the deliverable
+%capacity via pressure-swing adsorption using simplified models of the
+%gas-substrate and gas-gas interaction~\cite{gomez2014exploring,
+%gomez2017impact, kaija2018high, lee2019predicting}; none have established a
+%rigorous upper bound on the deliverable capacity.
+%
+Here, we present a
+theoretical upper bound on the deliverable capacity of a gas in a rigid
+porous
+material via an isothermal pressure swing.
+%
+To provide an extremum, we consider
+a substrate that provides a spatially uniform potential energy field for the
+gas. Our bound is unique in that it relies directly on experimentally measured
+properties of the pure gas.
+%
+We conclude
+that the goals set by the DOE for room-temperature natural gas and hydrogen
+storage are just barely theoretically possible.
+The goals are likely to be impossible for any real,
+rigid porous material because of steric
+repulsion, which must reduce the capacity below our upper
+bound. Limitations
+to the scope of applicability of our upper bound may guide fuel tank design and
+future material development. Firstly, one could avoid using an \emph{isothermal}
+pressure swing by heating the adsorbent to drive off
+trapped, residual gas. Secondly,
+our upper bound assumes the material does not change its
+structure in response to adsorbed gas, suggesting that flexible materials could
+still satisfy the DOE targets.}
+
+%Any references in the abstract should be written out in full \textit{e.g.}\ [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].} \\%The abstrast goes here instead of the text "The abstract should be..."
 
@@
 %%%FOOTNOTES%%%
-
-\footnotetext{\textit{$^{a}$~Address, Address, Town, Country. Fax: XX XXXX XXXX; Tel: XX XXXX XXXX; E-mail: xxxx@aaa.bbb.ccc}}
-\footnotetext{\textit{$^{b}$~Address, Address, Town, Country. }}
+\footnotetext{\textit{$^{*}$~E-mail: pommerjo@oregonstate.edu}}
+\footnotetext{\textit{$^{a}$~Department of Physics, Oregon State University, Corvallis, OR 97331, USA.}}
+\footnotetext{\textit{$^{b}$~Department of Chemical, Biological, and Environmental Engineering, Oregon State University, Corvallis, OR 97331, USA. }}
 
@@
 %If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
-\footnotetext{\dag~Electronic Supplementary Information (ESI) available: [details of any supplementary information available should be included here]. See DOI: 00.0000/00000000.}
+
+%\footnotetext{\dag~Electronic Supplementary Information (ESI) available: [details of any supplementary information available should be included here]. See DOI: 00.0000/00000000.}
+
 %additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required
 
-\footnotetext{\ddag~Additional footnotes to the title and authors can be included \textit{e.g.}\ `Present address:' or `These authors contributed equally to this work' as above using the symbols: \ddag, \textsection, and \P.
+%\footnotetext{\ddag~Additional footnotes to the title and authors can be included \textit{e.g.}\ `Present address:' or `These authors contributed equally to this work' as above using the symbols: \ddag, \textsection, and \P.
 Please place the appropriate symbol next to the author's name and include a \texttt{\textbackslash footnotetext} entry in the the correct place in the list.}
@@
 \section*{Conflicts of interest}
-In accordance with our policy on \href{http://www.rsc.org/journals-books-databases/journal-authors-reviewers/author-responsibilities/#code-of-conduct}{Conflicts of interest} please ensure that a conflicts of interest statement is included in your manuscript here.
-Please note that this statement is required for all submitted manuscripts.
-If no conflicts exist, please state that ``There are no conflicts to declare''.
+There are no conflicts to declare.
 
-\section*{Acknowledgements}
-The Acknowledgements come at the end of an article after Conflicts of interest and before the Notes and references.
+%\section*{Acknowledgements}
+%The Acknowledgements come at the end of an article after Conflicts of interest and before the Notes and references.
 

### 2020-07-10 gas-adsorption/rsc-gas-ads-complete.tex
fix affiliation
shout out to cofs

@@
 \footnotetext{\textit{$^{a}$~Department of Physics, Oregon State University, Corvallis, OR 97331, USA.}}
-\footnotetext{\textit{$^{b}$~Department of Chemical, Biological, and Environmental Engineering, Oregon State University, Corvallis, OR 97331, USA. }}
+\footnotetext{\textit{$^{b}$~School of Chemical, Biological, and Environmental Engineering, Oregon State University, Corvallis, OR 97331, USA. }}
 
@@
 emergence of highly tunable materials with large surface areas, such as
-metal-organic frameworks~\cite{furukawa2013chemistry}, no porous material has
-met these deliverable capacity targets.
+metal-organic frameworks~\cite{furukawa2013chemistry} and covalent organic 
+frameworks~\cite{}, no porous material has met these deliverable capacity targets.
 

### 2020-07-10 gas-adsorption/rsc-gas-ads-complete.tex
reorder intro sentences

@@
 water when it electrochemically reacts with oxygen in a fuel cell to power a
-vehicle. Currently, hydrogen (H$_2$) is primarily produced by steam reforming
+vehicle. 
+Notably, the environmental allure of hydrogen is predicated on its production 
+via a renewable means, e.g.\ splitting water using wind-generated electricity;
+currently, hydrogen (H$_2$) is primarily produced by steam reforming
 of natural gas followed by the water-gas shift reaction, which emits carbon
-dioxide \cite{crabtree2004hydrogen}. Notably, the environmental allure of
-hydrogen is predicated on its production via a renewable means, e.g.\ splitting
-water using wind-generated electricity.
+dioxide \cite{crabtree2004hydrogen}.
 

### 2020-07-10 gas-adsorption/rsc-gas-ads-complete.tex
made \V_opt paragraph have two separate ideas, instead of intertwining them which was confusing

@@
 temperature and pressure, allowing for room temperature and lower-pressure
-storage and alleviating many drawbacks of high-pressure, low-temperature
+storage and alleviating many drawbacks of high-pressure or low-temperature
 storage.
@@
 
-In this work, we place a rigorous upper bound on the deliverable capacity of a
+In this work, we place a rigorous upper bound on the isothermal deliverable capacity of a
 pure gas in a rigid substrate by endowing a control volume with a spatially
@@
 static potential energy fields (a broader claim), consider a
-non-spatially-uniform variation on $\V_{opt}$. The effect of
-this variation is to reduce the deliverable capacity: At low adsorbed gas
-densities, the lowest-energy positions will be preferentially occupied, while
+non-spatially-uniform variation on $\V_{opt}$. 
+At low adsorbed gas densities, the lowest-energy positions will be preferentially occupied, while
 at higher densities, the additional gas molecules will be forced into higher
 energy locations. Thus, the mean attraction will be lowest at \pfull\ and
-highest at \pempty. As a result, $\gst$ increases monotonically with
+highest at \pempty. 
+Therefore, the effect of a non-spatially uniform variation on $\V_{opt}$ is to reduce the deliverable capacity.
+By the same argument, $\gst$ increases monotonically with
 pressure. As shown in Sec.~\ref{sec:monotonic}, this results in
-the deliverable capacity of a non-uniform potential with a given $\gst$ at $\pfull$ or $\pempty$ being lower than the
-deliverable capacity of a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
+the deliverable capacity provided by a non-uniform potential with a given $\gst$ at $\pfull$ or $\pempty$ being lower than the
+deliverable capacity provided by a uniform potential with $\V=\gst$. Thus, a spatially uniform potential gives the maximal
 deliverable capacity amongst all static potential energy fields that give the
@@
 The grand canonical partition function of the control volume is:
-
 \begin{multline}
@@
 potential:
-
 \begin{equation}

### 2020-07-10 gas-adsorption/rsc-gas-ads-complete.tex
random changes

@@
 To show that a uniform potential gives the highest deliverable capacity, we
-consider an interaction potential between gas and substrate $\V(\rvec)$ that
+consider an interaction potential between gas and substrate offering a potential energy field $\V(\rvec)$ that
 varies in space.  In this proof, we make use of the Fourier transform of
@@
 \end{align}
-The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\kvec)\propto\delta(\kvec)$. In order for a uniform potential
+The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\kvec)\propto\delta(\kvec)$. For a uniform potential
 to extremize the deliverable capacity, we must show that the functional
@@
 \begin{align}
-    \frac{\delta D}{\delta \Vk} &= \mathbf{0}, \text{ if } \kvec\ne \mathbf{0}.
+    \frac{\delta \rho_D}{\delta \Vk} &= \mathbf{0}, \text{ if } \kvec\ne \mathbf{0}.
 \end{align}
@@
 \end{align}
-where $N_\text{full}$ and $N_\text{empty}$ are the number of particles at the full and empty pressure, $\pfull$ and $\pempty$, respectively.
+where $N_\text{full}$ and $N_\text{empty}$ are the number of particles in the substrate at the full and empty pressure, $\pfull$ and $\pempty$, respectively.
 
@@
     \\
-    &= \left(\frac{\partial \left(\frac{\delta F}{\delta \Vk}\right)_{\volume}}{\partial N}\right)_{\volume}
+    &= \left(\frac{\partial \left(\frac{\delta F}{\delta \Vk}\right)_{N}}{\partial N}\right)_{\volume}
     \label{eq:dmudpot}
@@
 Thus, the functional derivative of both $\mu$ and $N$ with regard to $\V(\rvec)$ are themselves spatially uniform.
-Since we already maximize $D$ with respect to the spatially uniform component of the potential (i.e.
-$\kvec=0$), the derivative of $D$ with respect to any change of potential is zero.
+Since we already maximize $\rho_D$ with respect to the spatially uniform component of the potential (i.e.
+$\kvec=0$), the derivative of $\rho_D$ with respect to any change of potential is zero.
 
@@
 substrate at room temperature. Such a material would require a void fraction of
-at least 94\%. On top of this, the DOE ULTIMATE target requires an optimal
+at least 94\%. On top of this, meeting the DOE ULTIMATE target requires an optimal
 $|\gst|$ of 10~kJ/mol which is far greater than what is found in observed
@@
 Why do we not experimentally observe materials approaching the upper bound on
-the deliverable capacity? Foremost, any adsorbent substrate will be composed of
+the deliverable capacity? First and foremost, any adsorbent substrate will be composed of
 atoms, which will exclude the gas from some volume. Due to strong short-range
@@
 attraction for gas atoms, imposing a limit on the pore
-volume. In contrast, real materials have a spatially non-uniform attraction for
+volume. Second, in contrast to the homogeneous substrate, real materials have a spatially non-uniform attraction for
 gas atoms. As a result, there are regions of space with

### 2020-07-11 gas-adsorption/rsc-gas-ads-complete.tex
added refs
fix the citations for fig 8 and 9 regarding cof-102.

@@
 
+Finally, empirical models that statistically relate e.g., surface area to 
+hydrogen adsorption \cite{firlej2013understanding,goldsmith2013theoretical}
+have been used to determine what properties a material must possess to reach
+a given storage target.
+
 % Brown and Freeman \cite{bhown2011analysis}
@@
     \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
-Experimental deliverable capacities for several porous materials (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
+Experimental deliverable capacities for several porous materials (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-298-D}
@@
     \caption{Deliverable capacity of hydrogen at 77\ K as a function of the attractive Gibbs free energy $\gst$.
-Experimental deliverable capacities for several porous materials (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
+Experimental deliverable capacities for several porous materials (data from Ref.~\cite{garcia2018benchmark, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
     \label{fig:hydrogen-77-D}

### 2020-07-14 gas-adsorption/rsc-gas-ads-complete.tex
minor revisions to proof text

@@
 \subsection{Proof of extremum}\label{sec:proof-extremum}
+
 To show that a uniform potential gives the highest deliverable capacity, we
-consider an interaction potential between gas and substrate offering a potential energy field $\V(\rvec)$ that
-varies in space.  In this proof, we make use of the Fourier transform of
-this potential:
+consider an interaction potential between gas and substrate offering a
+potential energy field $\V(\rvec)$ that varies in space. We will show that the
+uniform potential has the highest deliverable capacify of the set of potentials
+having the same mean value, and to find the maximum deliverable capacity one
+maximizes over this mean value. In this proof, we make use of the Fourier
+transform of this potential:
 \begin{align}
@@
 \end{align}
-The Fourier transform of a uniform potential is a Dirac delta function $\tilde{\V}(\kvec)\propto\delta(\kvec)$. For a uniform potential
-to extremize the deliverable capacity, we must show that the functional
-derivative of the deliverable capacity with respect to $\Vk$ is zero for
-\emph{nonzero} values of $\kvec$, i.e.
+The Fourier transform of a uniform potential is a Dirac delta function
+$\tilde{\V}(\kvec)\propto\delta(\kvec)$. For a uniform potential to extremize
+the deliverable capacity, we must show that the functional derivative of the
+deliverable capacity with respect to $\Vk$ is zero for \emph{nonzero} values of
+$\kvec$, i.e.
 \begin{align}
@@
 We note that this functional derivative may be non-zero for $\kvec=\mathbf{0}$
-because we separately maximize with respect to the particular uniform potential
-$\V$. This means that
+because we separately maximize with respect to the mean value of the uniform
+potential $\V$. This means that
 \begin{align}
@@
 \end{align}
-where $N_\text{full}$ and $N_\text{empty}$ are the number of particles in the substrate at the full and empty pressure, $\pfull$ and $\pempty$, respectively.
+where $N_\text{full}$ and $N_\text{empty}$ are the number of particles in the
+substrate at the full and empty pressure, $\pfull$ and $\pempty$, respectively.
 
@@
 \end{align}
-Since changing the chemical potential changes the number of molecules in the general case, if we can show that $\left(\frac{\delta \mu}{\delta \Vk}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta \Vk}\right)_{\mu}=0$.
-Thus we consider
+Since changing the chemical potential changes the number of molecules in the
+general case, if we can show that $\left(\frac{\delta \mu}{\delta
+\Vk}\right)_{N}=0$ then we will have shown that $\left(\frac{\delta N}{\delta
+\Vk}\right)_{\mu}=0$. Thus we consider
 \begin{align}
@@
 \end{align}
-where we have made use of the derivative relationship between $\mu$ and the Helmholtz free energy $F$, and have then reordered the functional and partial derivatives.
-Let us consider the interior derivative first.  The derivative of the Helmholtz free energy with respect to the external potential $\Vk$ yields the number density:
+where we have made use of the derivative relationship between $\mu$ and the
+Helmholtz free energy $F$, and have then reordered the functional and partial
+derivatives. Let us consider the interior derivative first. The derivative of
+the Helmholtz free energy with respect to the external potential $\Vk$ yields
+the number density:
 \begin{align}
@@
 \end{align}
-The number density is itself spatially uniform for any stable system in a fluid state at this density (i.e.
-does not spontaneously crystallize), and thus has a Fourier transform that is proportional to a Dirac $\delta$-function.
-Thus, the functional derivative $\frac{\delta F}{\delta \V(\rvec)}$ is actually a uniform function.
-We can insert this expression into Eq.~\ref{eq:dmudpot} to find that
+The number density given a uniform potential is itself spatially uniform for
+any stable system in a fluid state at this density (i.e. does not spontaneously
+crystallize), and thus has a Fourier transform that is proportional to a Dirac
+$\delta$-function. Thus, the functional derivative $\frac{\delta F}{\delta
+\V(\rvec)}$ is actually a uniform function. We can insert this expression into
+Eq.~\ref{eq:dmudpot} to find that
 \begin{align}
@@
 \end{align}
-Thus, the functional derivative of both $\mu$ and $N$ with regard to $\V(\rvec)$ are themselves spatially uniform.
-Since we already maximize $\rho_D$ with respect to the spatially uniform component of the potential (i.e.
-$\kvec=0$), the derivative of $\rho_D$ with respect to any change of potential is zero.
+Thus, the functional derivative of both $\mu$ and $N$ with regard to
+$\V(\rvec)$ are themselves spatially uniform. Since we already maximize
+$\rho_D$ with respect to the spatially uniform component of the potential (i.e.
+$\kvec=0$), the derivative of $\rho_D$ with respect to any change of potential
+is zero.
 

### 2020-08-24 ising/ising-body.tex
added the necessary citation.

@@
 belardinelli2007wang}.
-Belardinelli and Pereyra showed that an update factor that decreases faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang, belardinelli2008analysis, zhou2008optimal}, where $t$ corresponds to the number of moves.
+Belardinelli and Pereyra showed that an update factor that decreases faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007fast, belardinelli2007wang, belardinelli2008analysis, zhou2008optimal}, where $t$ corresponds to the number of moves.
 Schneider \emph{et al.} outline minor refinements algorithm including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.

### 2020-09-03 gas-adsorption/colors.py
changed formatting of MOF names.

@@ -69,6 +69,8 @@ latex_mof_names = {
     'CYCU3Al': 'CYCU-3-Al',
     'UiO68Ant': 'UiO-68-Ant',
     'rhtMOF7': 'rht-MOF-7',
+    'HKUST1-mono': 'HKUST-1-mono',
+    'NJUBai43': 'NJU-Bai-43',
 }
 def latex_mof(mof):
     if mof in latex_mof_names:

### 2020-09-03 gas-adsorption/rsc-gas-ads-complete.tex
add the paper citations for the data we just added to the paper.

@@
 The bulk methane density is from the National Institute of Standards and Technology (NIST)~\cite{nist}.
-The methane adsorption isotherms in the porous materials are experimental data from Ref.~\cite{mason2014evaluating, furukawa2009storage}.
+The methane adsorption isotherms in the porous materials are experimental data from Ref.~\cite{mason2014evaluating, furukawa2009storage, tian2018sol, gandara2014high, zhang2017fine}.
     }
@@
     \caption{The deliverable capacity of methane as a function of the attractive Gibbs free energy $|\gst|$.
-    Experimental deliverable capacities for several porous materials (data from Ref.~\cite{mason2014evaluating, furukawa2009storage}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line.}
+    Experimental deliverable capacities for several porous materials (data from Ref.~\cite{mason2014evaluating, furukawa2009storage, tian2018sol, gandara2014high, zhang2017fine}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as dots connected by a line.}
     \label{fig:methane-298-D}

### 2021-02-25 gas-adsorption/referee-response.md
start a response to referees

## Response to referees

# Referee 1

Referee 1 states that "Phenomenological predictions without any relation to real
materials are not very helpful, if not useless."  The referee fails to grasp
that our theory *does* take into account the gas, and places a fundamental limit
on the capacity of an entire class of materials (inflexible porous materials).
This is far from useless.  The referee suggests as superior an approach
involving screening hundreds of thousands of actual MOFs.  We will point out
that such an approach cannot possibly determine that there isn't a better MOF
out there, while our approach can do so.

TODO: We should find a place to edit the paper to make this more clear, and
state here that we did so.

TODO: Maybe include some more discussion of pore volume, just to illustrate that
our method does give information related to things like pore volume?

The other suggestion of Referee 1 is that we clearly state the capacity that we report, as discussed by Parilla et al.

TODO: We must do so.

# Referee 3

Referee 3 states that the conclusions are convincing, and makes two suggestions
for improving the paper:

1. The referee suggests that we attempt to quantify the possible effect of a
   flexible MOF on our upper bound.

TODO: Add a bit of discussion of how and why our model does not apply to
flexible MOFs, and why we cannot place *any* upper bound on a flexible MOF using
our approach.

2. The referee asks that we provide some guidelines in terms of working
   conditions to reach the DOE target.  They further suggest looking at a
   temperature difference to see how that could impact the deliverable capacity.

TODO: Add *some* sort of extra discussion on the topic.  Our model cannot
predict a temperature+pressure swing cycle because that would require
information about the entropy of adsorption at the two temperatures, which is
rather trickier, I think.

### 2021-02-25 gas-adsorption/referee-response.md
sketch out a note to editor
add a bit more response

@@
 ## Response to referees
+
+Dear editor,
+
+The responses of referees 1 and 3 could hardly be more divergent.  We FIXME
+SHOULD ADD A SUCCINCT ARGUMENT HERE WHY REFEREE 3 IS BETTER.
 
@@
 
-TODO: Maybe include some more discussion of pore volume, just to illustrate that
-our method does give information related to things like pore volume?
+The referee gives a litany of structural properties that we do not address in
+our model: "pore volume, pore size, specific surface area, crystal density,
+skeletal density, packing density, different adsorption sites etc."
+
+    We note that some of these (e.g. crystal density, skeletal density)
+    manifestly have no impact on the property we are studying, which is
+    volumetric storage capacity.  Others of course do impact the volumetric
+    storage capacity of a material, but do not relate to an upper bound
+    specifically.  TODO: We do add a discussion of the pore volume fraction
+    and how it can constrained by our upper bound, to make more clear the
+    range of applications of our bound.  TODO: We can put a lower bound on
+    pore volume fraction contingent on the maximum density having a given value.
+    It will take some explanation, but I think is worthwhile.
 

### 2021-03-05 gas-adsorption/rsc-gas-ads-complete.tex
start flexible thingy

@@
 potential energy field $\V(\rvec)$ that varies in space. We will show that the
-uniform potential has the highest deliverable capacify of the set of potentials
+uniform potential has the highest deliverable capacity of the set of potentials
 having the same mean value, and to find the maximum deliverable capacity one
@@
 
+\section{Flexible two-state model}
+
+\begin{align}
+    |n(p_{hi})\V| &> |\Delta F| \\
+    |n(p_{lo})\V| &< |\Delta F|
+\end{align}
+This means that for a given $\Delta F$ we can put an upper bound on the
+deliverable capacity.
+
 \section{Conclusions}

### 2021-03-06 gas-adsorption/isothermal-ads-gas-plot.py
create new flexible MOF figure

@@ -68,9 +68,9 @@ gram = numericalunits.g
 kg = numericalunits.kg
 L = numericalunits.L
 Joule = numericalunits.J
-mol = numericalunits.mol
+mol = 6.02214e23
 cm = numericalunits.cm
-mmol = numericalunits.mmol
+mmol = 1e-3*mol
 
 mg = numericalunits.mg    # milligram
 mL = numericalunits.mL    # milliliter
@@ -349,6 +349,31 @@ plt.ylabel('$\Delta g_{st}$ (kJ/mol)')
 plt.legend(loc='lower right')
 plt.tight_layout()
 plt.savefig('figs/' + basename + '-' + temperature + '-gst.pdf', transparent=True)
+
+plt.figure('capacity vs Delta F', figsize=(5,3.75))
+delta_F_flexible = np.linspace(0, 1e23*kJ/mol/cm**3, 100)
+capacities = np.zeros_like(delta_F_flexible)
+
+def n_from_mu_with_nans(particular_mu):
+    return np.interp(particular_mu, mu, n, left=np.nan, right=np.nan)
+
+mu_lo = mu_from_p(p_empty)
+n_lo = n_from_mu_with_nans(mu_lo + Gads)
+n_hi = n_from_mu_with_nans(mu_from_p(p_full) + Gads)
+print('n_lo', n_lo/density_units)
+for i in range(len(capacities)):
+    to_minimize = abs(n_lo*Gads - delta_F_flexible[i])
+    to_minimize[np.isnan(to_minimize)] = 1e200
+    which = np.argmin(to_minimize)
+    capacities[i] = n_hi[which]
+
+plt.plot(delta_F_flexible/(kJ/mol/cm**3), capacities/density_units, '.-', label='good stuff')
+plt.xlabel(r'$\Delta F$ (kJ/mol/cm$^3$)')
+plt.ylabel(r'$\rho_D$ (%s) two-phase assumption' % density_unit_name)
+
+plt.legend()
+plt.tight_layout()
+plt.savefig('figs/' + basename + '-' + temperature + '-flexible.pdf', transparent=True)
 
 if 'noshow' not in sys.argv:
     plt.show()

### 2021-03-09 gas-adsorption/rsc-gas-ads-complete.tex
create new flexible MOF figure
start on flexible two-state model

@@
 \begin{align}
-    |n(p_{hi})\V| &> |\Delta F| \\
-    |n(p_{lo})\V| &< |\Delta F|
+    V_{\text{open}}|n(p_{hi})\V| &> |\Delta F| \\
+    V_{\text{open}}|n(p_{lo})\V| &< |\Delta F|
 \end{align}
@@
 deliverable capacity.
+
+\davidsays{This means we should pick the $\V$ to be big enough to make us just
+on the boundary of not collapsing.  This means that:}
+\begin{align}
+    n(\mu_{lo} + \V)\V = \Delta F
+\end{align}
+This is a nonlinear equation that we want to solve.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{methane-298-flexible}
+    \caption{Deliverable capacity of methane in a flexible substrate at 298\ K as a function of the free energy difference 
+      between open and closed phases. The experimental deliverable capacity for Co(bdp) is shown for
+       comparison~\cite{fixme}, along with the free energy difference between open and closed conformations~\cite{fixthis}.}
+    \label{fig:methane-flexible}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{hydrogen-298-flexible}
+    \caption{Deliverable capacity of hydrogen in a flexible substrate at 298\ K as a function of the free energy difference 
+      between open and closed phases.}
+    \label{fig:hydrogen-flexible}
+\end{figure}
 

### 2021-03-09 gas-adsorption/isothermal-ads-gas-plot.py

@@ -70,11 +70,13 @@ L = numericalunits.L
 Joule = numericalunits.J
 mol = 6.02214e23
 cm = numericalunits.cm
+angstrom = 1e-8*cm
 mmol = 1e-3*mol
 
 mg = numericalunits.mg    # milligram
 mL = numericalunits.mL    # milliliter
 kJ = numericalunits.kJ     # kiloJoule
+J = numericalunits.J
 
 # --- Command line arguments --- #
 
@@ -351,7 +353,16 @@ plt.tight_layout()
 plt.savefig('figs/' + basename + '-' + temperature + '-gst.pdf', transparent=True)
 
 plt.figure('capacity vs Delta F', figsize=(5,3.75))
-delta_F_flexible = np.linspace(0, 1e23*kJ/mol/cm**3, 100)
+
+# delta_F_units = kJ/cm**3
+# delta_F_unit_name = 'kJ/cm$^3$'
+delta_F_units = J/cm**3
+delta_F_unit_name = 'J/cm$^3$'
+# delta_F_units = kJ/angstrom**3
+# delta_F_unit_name = f'kJ/$\AA^3$'
+
+
+delta_F_flexible = np.linspace(0, 0.04*kJ/cm**3, 300)
 capacities = np.zeros_like(delta_F_flexible)
 
 def n_from_mu_with_nans(particular_mu):
@@ -362,14 +373,49 @@ n_lo = n_from_mu_with_nans(mu_lo + Gads)
 n_hi = n_from_mu_with_nans(mu_from_p(p_full) + Gads)
 print('n_lo', n_lo/density_units)
 for i in range(len(capacities)):
-    to_minimize = abs(n_lo*Gads - delta_F_flexible[i])
-    to_minimize[np.isnan(to_minimize)] = 1e200
-    which = np.argmin(to_minimize)
-    capacities[i] = n_hi[which]
+    # to_minimize = abs(n_lo*Gads - delta_F_flexible[i])
+    # to_minimize[np.isnan(to_minimize)] = 1e200
+    # which = np.argmin(to_minimize)
+    # capacities[i] = n_hi[which]
+    capacities[i] = np.interp(delta_F_flexible[i], n_lo*Gads, n_hi)
 
-plt.plot(delta_F_flexible/(kJ/mol/cm**3), capacities/density_units, '.-', label='good stuff')
-plt.xlabel(r'$\Delta F$ (kJ/mol/cm$^3$)')
[…]

### 2021-03-12 gas-adsorption/referee-response.md
add referee response about the flexible model

@@
 
-TODO: Add a bit of discussion of how and why our model does not apply to
-flexible MOFs, and why we cannot place *any* upper bound on a flexible MOF using
-our approach.
+We have added a section "Flexible two-phase model" discussing flexible MOFs. We
+cannot place a formal upper bound on a flexible MOF in general, but we can put
+an upper bound on the capacity of a two-phase MOF such as Co(bdp), if we
+constrain the free energy difference between the two phases. This suggests that
+this type of a system will require a more stable closed phase in order to reach
+the DOE targets.
 
-2. The referee asks that we provide some guidelines in terms of working
+1. The referee asks that we provide some guidelines in terms of working
    conditions to reach the DOE target.  They further suggest looking at a

### 2021-03-12 gas-adsorption/rsc-gas-ads-complete.tex
add referee response about the flexible model

@@
 \usepackage{mdframed}
+
+\usepackage[color]{changebar}
+\cbcolor{red}
 
@@
 
-\section{Flexible two-state model}
-
-\begin{align}
-    V_{\text{open}}|n(p_{hi})\V| &> |\Delta F| \\
-    V_{\text{open}}|n(p_{lo})\V| &< |\Delta F|
-\end{align}
-This means that for a given $\Delta F$ we can put an upper bound on the
-deliverable capacity.
-
-\davidsays{This means we should pick the $\V$ to be big enough to make us just
-on the boundary of not collapsing.  This means that:}
-\begin{align}
-    n(\mu_{lo} + \V)\V = \Delta F
-\end{align}
-This is a nonlinear equation that we want to solve.
-
+\cbstart
+\section{Flexible two-phase model}
 \begin{figure}
@@
 \end{figure}
+Although our limit does not directly apply to materials that change conformation in
+response to gas adsorption, we can make an estimate of the benefits to be gained
+materials such as \davidsays{FIXME AND CITE} which undergo a phase change in the
+presence of gas.  We can model such a material as a two-phase system with one ``open''
+phase and a ``closed'' phase in which all gas is expelled.  We assume there is a
+free energy difference $\Delta G$ between the two phases in the absence of gas, with
+the closed phase more stable.  In addition, we assume a $\V$ as above \davidsays{REFERENCE EQUATION?}
+for the adsorption of gas into the open phase.  In order for the system to benefit
+from the two phases, the open phase must be preferred at $\pfull$, with the closed phase
+preferred at $\pempty$.  This requires that
+\begin{align}
+    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> |\Delta G| \\
+    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< |\Delta G|
+\end{align}
+where $\rho_{\text{open}}(p)$ is the density of gas in the open phase of the material at
+the given pressure.  This model has no upper bound on the deliverable capacity, since
+$\V$ and $\Delta G$ can together be increased without bound.  We can, however, place a
+bound for a fixed $\Delta G$, since if the gas is too strongly attracted to the open phase
+the material will remain open at $\pempty$.  This upper bound is found by identifying the
+attractive interaction $\V$ that solves the equation
+\begin{align}
+    \rho(\muempty{} + \V)\V = \Delta G/V_{\text{open}}
+\end{align}
+for $\V$, which gives the largest attraction consistent with the material closing
+at the empty pressure.
+Figure~\ref{fig:methane-flexible} illustrates
+how this upper bound depends on $\Delta G$ for methane.  For comparison, we include the free energy
+difference for Co(bdp)~\davidsays{CITE} and its measured deliverable capacity.  While this
+upper bound is not rigorous, it suggests that a two-phase flexible material for methane storage
+would benefit from significantly greater stability of the closed phase---provided the open phase
+had correspondingly greater attraction for the gas molecules.  Figure~\ref{fig:hydrogen-flexible}
+shows the corresponding result for hydrogen storage, indicating that for a flexible material to
+reach the DOE ULTIMATE deliverable capacity would require far greater stability for the closed
+phase, as well as an unprecedented level of attractive interaction for hydrogen.
 
@@
 \end{figure}
+
+\cbend
 

### 2021-03-12 gas-adsorption/isothermal-ads-gas-plot.py

@@ -394,17 +394,21 @@ delta_F_cobdp = 3.3*kJ/mol/volume_cobdp
 
 methane_codbp_capacity = 196 # Mason et al. Nature "Methane storage in flexible ..."
 
-plt.plot([delta_F_cobdp/delta_F_units, delta_F_cobdp/delta_F_units],
-         [0, capacities.max()/density_units], color='xkcd:green', label='Co(BDP)')
+experiment_symbol = ''
 if basename == 'methane':
+    experiment_symbol = colors.symbol(basename)
     plt.plot([delta_F_cobdp/delta_F_units],
-             [methane_codbp_capacity], colors.symbol(basename), color='xkcd:green', markersize=12)
+             [methane_codbp_capacity], experiment_symbol, color='xkcd:green', markersize=12)
+plt.plot([delta_F_cobdp/delta_F_units, delta_F_cobdp/delta_F_units],
+         [-capacities.max()/density_units, 1.1*capacities.max()/density_units],
+         experiment_symbol+'-', color='xkcd:green', label='Co(bdp)',
+         markersize=12, linewidth=0.5)
 
 # delta_F_MIL_53 = 2.5*kJ/mol/(1012*angstrom**3)
 # plt.plot([delta_F_MIL_53/delta_F_units, delta_F_MIL_53/delta_F_units],
 #          [0, capacities.max()/density_units], label='MIL-53(Cr)')
 
-plt.plot(delta_F_flexible/delta_F_units, capacities/density_units, '-', label='good stuff')
+plt.plot(delta_F_flexible/delta_F_units, capacities/density_units, '-', label='Two-phase upper bound')
 
 # this paper gives the \Delta F of Co(BDP) at 77 K as 3.3 kJ/mol, of
 # Cu(4,4′-bipy)(dhbc)2 at 298 K as ~ 4 kJ/mol (see table 1) and for MIL-53 at 304
@@ -412,7 +416,7 @@ plt.plot(delta_F_flexible/delta_F_units, capacities/density_units, '-', label='g
 # and they treat the MOF as a two-state system, which is kind of true. so yeah
 # these are directly appropriate for comparing to our model.
 
-plt.xlabel(rf'$\Delta F/V_{{open}}$ ({delta_F_unit_name})')
+plt.xlabel(rf'$\Delta G/V_{{open}}$ ({delta_F_unit_name})')
[…]

### 2021-03-23 gas-adsorption/rsc-gas-ads-complete.tex
add cartoon from Cory
unique => only bound possible. dont need to justify novelty
improve intro

@@
 a substrate that provides a spatially uniform potential energy field for the
-gas. Our bound is unique in that it relies directly on experimentally measured
+gas. Our bound relies directly on experimentally measured
 properties of the pure gas.
@@
 We conclude
-that the goals set by the DOE for room-temperature natural gas and hydrogen
+that the deliverable capacity targets set by the DOE for room-temperature natural gas and hydrogen
 storage are just barely theoretically possible.
-The goals are likely to be impossible for any real,
+The targets are likely to be impossible for any real,
 rigid porous material because of steric
-repulsion, which must reduce the capacity below our upper
+repulsion, which reduces the deliverable capacity below our upper
 bound. Limitations
@@
 The transportation sector is dominantly powered by petroleum-based
-fuels~\cite{davis2009transportation}. In the United States, it is responsible
+fuels~\cite{davis2009transportation}. 
+Consequently, in the United States, it is responsible
 for 36\% of energy-related carbon dioxide emissions \cite{useia} and $\sim$50,000
 premature deaths per year associated with particulate matter and ozone
-emissions~\cite{caiazzo2013air}. Natural gas and hydrogen (H$_2$) are
-alternative transportation fuels that emit fewer/less pollutants than
-petroleum-based fuels, and thus their widespread adoption could mitigate
-climate change~\cite{mcglade2015geographical} and improve air quality for human
-health. Further, as the finite global petroleum resources are declining
-rapidly~\cite{sorrell2010global}, the development of technologies for the
-widespread adoption of sustainable transportation fuels, such as hydrogen, is
+emissions~\cite{caiazzo2013air}. 
+Natural gas and hydrogen (H$_2$) are
+alternative transportation fuels that, if adopted widely,
+could mitigate climate change~\cite{mcglade2015geographical} and improve air quality for human
+health. 
+More, the finite global petroleum resources are declining
+rapidly~\cite{sorrell2010global}.
+The development of technologies for the
+widespread adoption of sustainable transportation fuels, such as hydrogen, is therefore
 critical.
@@
 At ambient conditions, both methane and hydrogen gas possess a very low
-volumetric energy density compared to gasoline. Consequently, under storage
+volumetric energy density compared to (liquid) gasoline. Consequently, under storage
 space constraints in passenger vehicles, natural gas and hydrogen must be
@@
 energy input to liquify hydrogen is $\sim$30\% of its energy
-content~\cite{bossel2003energy}. Moreover, high-pressure storage tanks are
+content~\cite{bossel2003energy}. Also, high-pressure storage tanks are
 heavy, thick-walled, and non-conformable, while cryogenic storage tanks are
@@
 to achieve a higher adsorbed gas density than the bulk gas at the same
-temperature and pressure, allowing for room temperature and lower-pressure
-storage and alleviating many drawbacks of high-pressure or low-temperature
+temperature and pressure.
+Nanoporous materials could thus allow for room temperature and lower-pressure
+storage of natural gas and hydrogen and alleviate many drawbacks of high-pressure or low-temperature
 storage.
@@
 thermodynamic property that determines the driving
-range~\cite{mason2014evaluating}. The adsorbed gas storage tank delivers the
+range on a given ``full'' tank of fuel~\cite{mason2014evaluating}. 
+The adsorbed gas storage tank delivers the
 gaseous fuel to the engine via an (assumed) isothermal pressure
@@
 Agency--Energy (ARPA-E) set a deliverable capacity target of 315\ L STP CH$_4$
-per L of adsorbent at 298 K using a 65\ bar to 5.8\ bar pressure
+per L of adsorbent at 298 K using a 65\,bar to 5.8\,bar pressure
 swing~\cite{simon2015materials}. For hydrogen, the DOE set a series of
-progressive targets at five year intervals, with the ultimate target of 0.05 kg
+progressive targets at five year intervals, with the ultimate target of 0.05\,kg
 H$_2$/L~\cite{h2targetsDOE} using a 100\ bar to 5\ bar pressure swing at a
@@
 metal-organic frameworks~\cite{furukawa2013chemistry} and covalent organic 
-frameworks~\cite{}, no porous material has met these deliverable capacity targets.
+frameworks~\cite{diercks2017atom}, no porous material has met these deliverable capacity targets.
 
@@
 
-To set realistic performance targets and optimally allocate research resources,
+To set realistic performance targets and optimally allocate resources to research efforts,
 in this work, we present a theoretical framework that places an intrinsic upper
-limit on the deliverable capacity of any gas in a rigid porous material and
+limit on the deliverable capacity of a pure gas in a rigid porous material and
 uses as input the experimentally measured properties of the bulk gas. Our
@@
 \cbstart
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\columnwidth]{figs/two-phase}
+    \caption{Cartoon of our model.}
+    \label{fig:flexible-cartoon}
+\end{figure}
 \section{Flexible two-phase model}

### 2021-03-23 gas-adsorption/rsc-gas-ads-complete.tex
use g for gas instead of 0 + minor stuff

@@
 residual gas will remain trapped in the adsorbent. Therefore, the driving range
-of the vehicle is primarily determined by the \emph{deliverable capacity} of
+of the vehicle is primarily determined by the \emph{deliverable capacity} $\rho_D$ of
 the gas in the material (see Fig.~\ref{fig:delcap}): the density of gas at
@@
 optimal free energy of adsorption (which determines the Langmuir constant) that
-maximizes the deliverable capacity
-$\rho_D$~\cite{matranga1992storage,bhatia2006optimum,simon2014optimizing}. If
+maximizes the deliverable capacity~\cite{matranga1992storage,bhatia2006optimum,simon2014optimizing}. If
 the gas-substrate interaction is too weak (strong), too little (much) gas
-adsorbs (is retained) at $\pfull$ ($\pempty$), diminishing $\rho_D$. An upper
+adsorbs (is retained) at $\pfull$ ($\pempty$), diminishing the deliverable capacity. An upper
 bound on the deliverable capacity of a Langmuir material follows if each
@@
 accounted for gas-gas interactions via an intermolecular potential and
-idealized the substrate in two different ways: (1) discrete adsorption sites
+idealized the substrate in two different ways---as a: (1) set of discrete adsorption sites
 packed into an FCC lattice and (2) a volume endowed with a spatially uniform,
@@
 material space, i.e.\ that the structures considered are representative of the
-set of possible materials. To further explore material space and address
+set of possible materials \cite{Moosavi2020}. To further explore material space and address
 sensitivity to the intermolecular potentials: scaling the Lennard-Jones
@@
 density of the gas $\rho_g(\mu,T)$ as a function of chemical potential $\mu$
-and temperature $T$, is shown for methane gas in
+and temperature $T$. We show $\rho_g(\mu, T)$ for methane gas in
 Fig.~\ref{fig:density-vs-mu-ch4} at $T=298$\ K. For comparison, we also show
@@
     \includegraphics[width=\columnwidth]{methane-298-rho-mu}
-    \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several porous materials at 298\ K as a function of chemical potential (bottom axis) and pressure (top axis).
+    \caption{The density of bulk methane, the ideal gas, and adsorbed gas in several porous materials at 298\,K as a function of chemical potential (bottom axis) and pressure (top axis).
 The bulk methane density is from the National Institute of Standards and Technology (NIST)~\cite{nist}.
@@
 attraction for the gas and (ii) the atoms required to create the potential
-field exclude gas from occupying a fraction of the control volume.
+field exclude gas from occupying a fraction of the control volume and thus reduce utilized space.
 
@@
 axis is the difference in molar Gibbs free energy of the gas, at fixed density (the density in the adsorbent) and temperature, between the gas within the substrate and the pure gas (\gst),
-which is equal to $\V$ in our model, and is pressure-dependent in a real
+which is equal to $\V$ in our model, but is pressure-dependent in a real
 crystal~(see Sec.~\ref{sec:phi-is-delta-g}). We see a potential $\V$ that
@@
 \begin{equation}
-    \Xi(\mu, V, T)=\Xi_0(\mu - \V, V, T)
+    \Xi(\mu, V, T)=\Xi_g(\mu - \V, V, T)
     \label{eq:xi_vs_xi0}
 \end{equation}
-where $\Xi_0(\mu, V, T)$ is the grand partition function of the bulk
+where $\Xi_g(\mu, V, T)$ is the grand partition function of the bulk
 gas in the control volume in the absence of an external potential. Importantly,

### 2021-04-21 gas-adsorption/rsc-gas-ads-complete.tex
small changes

@@
 Gibbs free energy (equivalent to chemical potential) between the adsorbed gas system and
-the bulk gas \emph{with the same density of gas molecules}. The quantity $\gst$
+the bulk gas \emph{with the same density of gas molecules} and at the same temperature. The quantity $\gst$
 does \emph{not} correspond to a change in the molar Gibbs free energy as a
@@
 in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
-Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to
-measure $\gst$ via a piston with a removable partition that separates a volume
-of free space from the same volume of substrate. Note that $\gst$ is a property
+Figure~\ref{fig:delta-G-cartoon} illustrates $\gst$.
+% David:  no longer illustrates the expt.
+A hypothetical experiment to measure $\gst$ is via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
+Initialize the experiment with all gas in the free space and an empty substrate.
+Then, slowly (i) remove the partition and (ii) use the piston to push the gas into the substrate and eliminate the free space volume, in an isothermal process.
+$\gst$ is the ... \davidsays{wut? energy it takes to do this or?}
+% David: what is GST
+Note that $\gst$ is a property
 of both the substrate and the identity of the gas. Because real substrates
@@
 eqn.~\ref{eq:g_st}, $\gst$ follows from subtracting the two quantities.
+% David: is it fishy to say \mu_gas(p) then \mu_gas(\rho)? seems sloppy notation or is it ok?
 
@@
 atoms, which will exclude the gas from some volume. Due to strong short-range
-interactions, substrate atoms must be approximately uniformly distributed to achieve strong
+interactions, from a longer length-scale perspective, substrate atoms must be approximately uniformly distributed to achieve strong
 attraction for gas atoms, imposing a limit on the pore

### 2021-04-21 gas-adsorption/rsc-gas-ads-complete.tex
some edits on flexible MOF section

@@
     \includegraphics[width=0.95\columnwidth]{figs/two-phase}
-    \caption{Cartoon of our model.}
+    \caption{Cartoon of a two-phase, collapsing material.
+The material can exist in two phases: an open phase in which gas molecules can adsorb and a closed phase that is collapsed and cannot fit any gas molecules.
+Take the closed phase as more stable, so that the material can collapse and expels residual gas.
+    The pressure of gas at which the structural transition occurs depends on the gas-host interactions and relative stability of the phases.
+We wish for the structural transition to occur between $\pfull$ and $\pempty$.
+    }
     \label{fig:flexible-cartoon}
 \end{figure}
+
 \section{Flexible two-phase model}
@@
     \includegraphics[width=0.95\columnwidth]{methane-298-flexible}
-    \caption{Deliverable capacity of methane in a flexible substrate at 298\ K as a function of the free energy difference 
+    \caption{Maximal possible deliverable capacity of methane in a two-phase, collapsing substrate at 298\ K as a function of the free energy difference 
       between open and closed phases. The experimental deliverable capacity for Co(bdp) is shown for
@@
 \end{figure}
-Although our limit does not directly apply to materials that change conformation in
-response to gas adsorption, we can make an estimate of the benefits to be gained
-materials such as \davidsays{FIXME AND CITE} which undergo a phase change in the
-presence of gas.  We can model such a material as a two-phase system with one ``open''
-phase and a ``closed'' phase in which all gas is expelled.  We assume there is a
-free energy difference $\Delta G$ between the two phases in the absence of gas, with
-the closed phase more stable.  In addition, we assume a $\V$ as above \davidsays{REFERENCE EQUATION?}
-for the adsorption of gas into the open phase.  In order for the system to benefit
-from the two phases, the open phase must be preferred at $\pfull$, with the closed phase
-preferred at $\pempty$.  This requires that
-\begin{align}
-    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> |\Delta G| \\
-    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< |\Delta G|
-\end{align}
-where $\rho_{\text{open}}(p)$ is the density of gas in the open phase of the material at
-the given pressure.  This model has no upper bound on the deliverable capacity, since
-$\V$ and $\Delta G$ can together be increased without bound.  We can, however, place a
-bound for a fixed $\Delta G$, since if the gas is too strongly attracted to the open phase
-the material will remain open at $\pempty$.  This upper bound is found by identifying the
-attractive interaction $\V$ that solves the equation
-\begin{align}
-    \rho(\muempty{} + \V)\V = \Delta G/V_{\text{open}}
-\end{align}
-for $\V$, which gives the largest attraction consistent with the material closing
-at the empty pressure.
+Although our deliverable capacity limit does not directly apply to materials that change conformation in
+response to gas adsorption, we can estimate enhancements in the deliverable capacity provided by
+such materials. 
+For example, consider a material such as Co(bdp) \cite{mason2015methane} that 
+can, actuated by the presence of gas, switch between two structural phases: 
+an ``open'' phase where gas molecules can fit in the pore space and a 
+collapsed, ``closed'' phase in which no gas can fit inside.
+Let $\Delta G=G_{\text{open}}-G_{\text{closed}}$ be the free energy difference between the two phases in the absence of gas.
+% David: introduce units of delta G here.
+Suppose the closed phase is more stable ($\implies \Delta G >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
+In addition, we assume the open phase hosts a uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb, so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.
+For an enhancement of the deliverable capacity by this two-phase, flexible material, 
+the open phase must be preferred at $\pfull$, with the closed phase preferred at $\pempty$. 
+This requires that, neglecting gas-gas interactions:
+% David: doesn't this also neglect entropy? wut about the *free energy* of the gas?
+% David: I suggest \Delta G_{host} to be clear?
+\begin{align}
+    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta G \\
+    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta G
+\end{align}
+% David: I put \V<0, think we should say the sign of \Delta G to be clear too.
+% David: specify units on \delta G so clear where V comes from; might be better to absorb V into delta G?
+where $\rho_{\text{open}}(p)$ is the density of gas in the open phase of the material at pressure $p$. 
+I.e., (1, to be open at $\pfull$) the gas-host attractions at $\pfull$ must overcome the free energy penalty to open the structure and 
+(2, to be closed at $\pempty$) the free energy penalty to open the structure must be larger than the (weak) gas-host attractions at $\pempty$.
+This approximate model of the two-phase material has no upper bound on the deliverable capacity, since
+$\V$ and $\Delta G$ can together be increased without bound.
+We can, however, place a bound for a fixed $\Delta G$, since if the gas is too strongly attracted to the open phase,
+the material will remain open at $\pempty$ and our rigid upper bound applies.
+% David: does our rigid upper bound apply then?
+The upper bound for the two-phase material is found by identifying the
+attractive interaction $\V$ that balances the gas-host attractions in the idealized-open-substrate at $\pempty$ with the free energy penalty to open:
+\begin{align}
+    V_{\text{open}} \rho_g(\muempty{} - \V_{opt}, T)\V_{opt} = \Delta G.
+\end{align}
+% David: left is negative. what is the right? negative as well? how did u define it?
+This gives the largest possible attraction $|\V|$ consistent with the material closing
+at $\pempty$ to expel all residual gas. 
+The deliverable capacity is then the density of gas in the open phase at $\pfull$, $\rho_g(\mufull{} - \V_{opt}, T)$.
+% David: changed it to \V_opt, or not necessary?
+
 Figure~\ref{fig:methane-flexible} illustrates
-how this upper bound depends on $\Delta G$ for methane.  For comparison, we include the free energy
-difference for Co(bdp)~\davidsays{CITE} and its measured deliverable capacity.  While this
-upper bound is not rigorous, it suggests that a two-phase flexible material for methane storage
+how the upper bound on the deliverable capacity in a two-phase, collapsing material depends on $\Delta G$ for methane. 
+For comparison, we include the free energy
+difference for Co(bdp)~\davidsays{CITE} and its measured deliverable capacity.  
+While this upper bound is not rigorous, it suggests that a two-phase flexible material for methane storage
 would benefit from significantly greater stability of the closed phase---provided the open phase
 had correspondingly greater attraction for the gas molecules.  Figure~\ref{fig:hydrogen-flexible}
-shows the corresponding result for hydrogen storage, indicating that for a flexible material to
-reach the DOE ULTIMATE deliverable capacity would require far greater stability for the closed
+shows the corresponding result for hydrogen storage, indicating that, for a flexible material to
+reach the DOE ULTIMATE deliverable capacity, it would require far greater stability for the closed
 phase, as well as an unprecedented level of attractive interaction for hydrogen.

### 2021-04-22 gas-adsorption/rsc-gas-ads-complete.tex
change davidsays to corysays for clarity
change Fig. 4 and add new text (in davidsays) about measuring gst

@@
 \newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
+\newcommand{\corysays}[1]{{\color{red} [\blue{Cory:} \emph{#1}]}}
 
@@
 Then, slowly (i) remove the partition and (ii) use the piston to push the gas into the substrate and eliminate the free space volume, in an isothermal process.
-$\gst$ is the ... \davidsays{wut? energy it takes to do this or?}
+$\gst$ is the ... \corysays{wut? energy it takes to do this or?}
+%
+\davidsays{Okay, here is an improved experiment to measure $\gst$.  Two containers of
+equal volume are connected by a hose (with neglible volume).  One container holds
+the substrate, and the other is empty (apart from gas).    The altitude of one container
+is adjusted until the number of gas molecules is equal in each container.  The value
+of $\gst$ is the gravitational potential energy difference of gas molecules in the two
+boxes.}
 % David: what is GST

### 2021-06-16 gas-adsorption/referee-response.md
small changes in response to referees

@@
 
-The responses of referees 1 and 3 could hardly be more divergent.  We FIXME
-SHOULD ADD A SUCCINCT ARGUMENT HERE WHY REFEREE 3 IS BETTER.
+The responses of referees 1 and 3 could hardly be more divergent.
+Our work is impactful precisely because it can tell researchers when to stop looking...
+or perhaps more relevantly it can tell funding agencies when to stop funding research in this direction.
+This is threatening, and Referee 1 doesn't seem interested in knowing whether we should stop looking at porous materials for natural gas and hydrogen storage.
 
@@
 
-TODO: We must do so.
+   The deliverable capacity is defined in the paper as "the density of the gas
+   in the material at the storage pressure p_full minus the residual gas that
+   remains adsorbed at the lowest pressure p_empty such that sufficient flow is
+   maintained to feed the engine."  We believe this is clear. The various
+   capacities of Parilla et al. are geared towards experimental measurements and
+   account for experimental uncertainty in the volume of material, an issue
+   which is not present in a theoretical paper.
+We note that the paper Referee 1 brought up as an example does not any more clearly define "usable capacity" as they refer to what we call "deliverable capacity".
+They do cite Parilla et al., but only in the context of describing their experimental equipment.
 

### 2021-06-16 gas-adsorption/rsc-gas-ads-complete.tex
nicer name for beginning and ending edit section
add label to changebar, and use macros that maybe mean more than cbstart and cbend?
small changes in response to referees

@@
 \cbcolor{red}
+\newcommand\startreferees[1]{\cbstart \noindent\textit{\color{red}%
+[\textbf{To referees:} #1]}
+
+}
+\newcommand\donereferees{\cbend}
 
@@
 
+\startreferees{Deliverable capacity defined below}
 For a vehicle employing a porous material to store natural gas or hydrogen, the
@@
 frameworks~\cite{diercks2017atom}, no porous material has met these deliverable capacity targets.
+
+\donereferees
 
@@
 
-\cbstart
+\startreferees{Flexible MOFs section added below}
 \begin{figure}
@@
 
-\cbend
+\donereferees
 

### 2021-06-17 gas-adsorption/rsc-gas-ads-complete.tex
switch to Helmholtz free energy for flexible mof

@@
 collapsed, ``closed'' phase in which no gas can fit inside.
-Let $\Delta G=G_{\text{open}}-G_{\text{closed}}$ be the free energy difference between the two phases in the absence of gas.
+Let $\Delta F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free
+energy difference between the two phases in the absence of gas.
 % David: introduce units of delta G here.
-Suppose the closed phase is more stable ($\implies \Delta G >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
+Suppose the closed phase is more stable ($\implies \Delta F >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
 In addition, we assume the open phase hosts a uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb, so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.
@@
 \begin{align}
-    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta G \\
-    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta G
+    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta F -pV_{\text{open}} \\
+    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta F -pV_{\text{closed}}
 \end{align}
@@
 % David: specify units on \delta G so clear where V comes from; might be better to absorb V into delta G?
-where $\rho_{\text{open}}(p)$ is the density of gas in the open phase of the material at pressure $p$. 
+where $\rho_{\text{open}}(p)$ is the density of gas
+in the open phase of the material at pressure $p$. We will neglect the $pV$
+terms in the Gibbs free energy, under the assumption that the mechanical
+work associated with the transition is neglible when compared with the
+adsorption energies involved.
 I.e., (1, to be open at $\pfull$) the gas-host attractions at $\pfull$ must overcome the free energy penalty to open the structure and 
@@
 This approximate model of the two-phase material has no upper bound on the deliverable capacity, since
-$\V$ and $\Delta G$ can together be increased without bound.
-We can, however, place a bound for a fixed $\Delta G$, since if the gas is too strongly attracted to the open phase,
+$\V$ and $\Delta F$ can together be increased without bound.
+We can, however, place a bound for a fixed $\Delta F$, since if the gas is too strongly attracted to the open phase,
 the material will remain open at $\pempty$ and our rigid upper bound applies.
@@
 \begin{align}
-    V_{\text{open}} \rho_g(\muempty{} - \V_{opt}, T)\V_{opt} = \Delta G.
+    V_{\text{open}} \rho_g(\muempty{} - \V_{opt}, T)\V_{opt} = \Delta F.
 \end{align}
@@
 Figure~\ref{fig:methane-flexible} illustrates
-how the upper bound on the deliverable capacity in a two-phase, collapsing material depends on $\Delta G$ for methane. 
+how the upper bound on the deliverable capacity of methane in a two-phase,
+collapsing material depends on $\Delta F$. 
 For comparison, we include the free energy

### 2021-06-28 gas-adsorption/referee-response.md
a couple of edits to the response

@@
     storage capacity of a material, but do not relate to an upper bound
-    specifically.  TODO: We do add a discussion of the pore volume fraction
-    and how it can constrained by our upper bound, to make more clear the
-    range of applications of our bound.  TODO: We can put a lower bound on
-    pore volume fraction contingent on the maximum density having a given value.
-    It will take some explanation, but I think is worthwhile.
+    specifically.
+
+    As discussed in the second and third paragraphs of the results section, our results do put a lower bound of 85% on the void fraction for any porous material that will meet the ARPA-E target for methane volumetric deliverable capacity, and a lower bound of 94% for a material reaching the DOE ULTIMATE target for hydrogen at 298K.
+These results *can* guide exploration, if we assume that reaching those targets is the goal.
 

### 2021-07-06 gas-adsorption/referee-response.md
added note about temperature-pressure swings, and a davidsays that I wonder if we *can* do something here

@@
 
-TODO: Add *some* sort of extra discussion on the topic.  Our model cannot
-predict a temperature+pressure swing cycle because that would require
-information about the entropy of adsorption at the two temperatures, which is
-rather trickier, I think.
+We added a sentence mentioning that we cannot apply our upper bound in the case
+where the temperature is changed, because a uniform potential may no longer lead
+to the optimal deliverable capacity.

### 2021-07-06 gas-adsorption/rsc-gas-ads-complete.tex
added note about temperature-pressure swings, and a davidsays that I wonder if we *can* do something here

@@
 
+\startreferees{Added last sentence to clarify why we cannot place a bound
+  on the deliverable capacity when the temperature is changed.}
+
 The first limitation of our proof is that we restricted ourselves to
@@
 and vehicle.
+%
+We cannot place an upper bound on such a process using our approach because with
+changing temperature a uniform potential may not give the maximum deliverable
+capacity.  \davidsays{As far as I can intuit, \emph{normally} the entropy
+increases with temperature, and if that is the case then a uniform potential
+will give the optimum deliverable capacity.  If we can prove that for a subset
+of substrate interactions (e.g. when the substrate acts as an external potential
+felt by the gas), the optimal potential is uniform then we could apply our model
+to temperature-pressure swings.}
+
+\donereferees
 

### 2021-07-06 gas-adsorption/rsc-gas-ads-complete.tex
fix a davidsays

@@
 in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
-Figure~\ref{fig:delta-G-cartoon} illustrates $\gst$.
+% Figure~\ref{fig:delta-G-cartoon} illustrates  $\gst$.
 % David:  no longer illustrates the expt.
-A hypothetical experiment to measure $\gst$ is via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
-Initialize the experiment with all gas in the free space and an empty substrate.
-Then, slowly (i) remove the partition and (ii) use the piston to push the gas into the substrate and eliminate the free space volume, in an isothermal process.
-$\gst$ is the ... \corysays{wut? energy it takes to do this or?}
+% A hypothetical experiment to measure $\gst$ is via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
+% Initialize the experiment with all gas in the free space and an empty substrate.
+% Then, slowly (i) remove the partition and (ii) use the piston to push the gas into the substrate and eliminate the free space volume, in an isothermal process.
+% $\gst$ is the ... \corysays{wut? energy it takes to do this or?}
 %
-\davidsays{Okay, here is an improved experiment to measure $\gst$.  Two containers of
-equal volume are connected by a hose (with neglible volume).  One container holds
-the substrate, and the other is empty (apart from gas).    The altitude of one container
-is adjusted until the number of gas molecules is equal in each container.  The value
-of $\gst$ is the gravitational potential energy difference of gas molecules in the two
-boxes.}
+% \davidsays{Okay, here is an improved experiment to measure $\gst$.  Two containers of
+% equal volume are connected by a hose (with neglible volume).  One container holds
+% the substrate, and the other is empty (apart from gas).    The altitude of one container
+% is adjusted until the number of gas molecules is equal in each container.  The value
+% of $\gst$ is the gravitational potential energy difference of gas molecules in the two
+% boxes.}
+%
+Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to
+measure $\gst$.  Two containers of equal volume are connected by a hose.  One
+container holds the substrate, and the other is empty (apart from gas).    The
+altitude of one container is adjusted until the number of gas molecules is equal
+in each container.  The value of $\gst$ is the gravitational potential energy
+difference of gas molecules in the two boxes.
+
 % David: what is GST

### 2021-07-07 gas-adsorption/referee-response.md
more edits

@@
 This is far from useless.  The referee suggests as superior an approach
-involving screening hundreds of thousands of actual MOFs.  We will point out
-that such an approach cannot possibly determine that there isn't a better MOF
-out there, while our approach can do so.
+involving screening hundreds of thousands of actual or theoretical MOFs.
 
-TODO: We should find a place to edit the paper to make this more clear, and
-state here that we did so.
+    Such a study is indeed valuable for guiding the development of improved MOFs,
+    but is entirely unable to place a rigorous upper bound on the attainable
+    capacity, since it is always possible that there is a better structure out
+    there. We have edited the paragraph on page 3 of the paper that discusses
+    this sort of study to make this more clear (and added a reference the paper
+    by Ahmed et al.).
 

### 2021-07-07 gas-adsorption/rsc-gas-ads-complete.tex
more edits

@@
 
-Finally, empirical models that statistically relate e.g., surface area to 
-hydrogen adsorption \cite{firlej2013understanding,goldsmith2013theoretical}
-have been used to determine what properties a material must possess to reach
-a given storage target.
+\startreferees{Added discussion of screening studies}
+Finally, a number of empirical studies have been performed that computationally
+simulate a adsorption in a vast array of possible porous
+structures~\cite{firlej2013understanding, goldsmith2013theoretical,
+ahmed2019exceptional}.  Such studies can guide the development of new materials
+by allowing researchers to identify commonalities among materials that have
+desirable properties.  However, such an approach can only place a \emph{lower
+bound} on the highest attainable deliverable capacity.
+
+\donereferees
 

### 2021-07-07 gas-adsorption/referee-response.md
reformatting of referee response
add explanation of why non-uniform potential could be good if temperature changes

@@
 
-Referee 1 states that "Phenomenological predictions without any relation to real
-materials are not very helpful, if not useless."  The referee fails to grasp
-that our theory *does* take into account the gas, and places a fundamental limit
-on the capacity of an entire class of materials (inflexible porous materials).
-This is far from useless.  The referee suggests as superior an approach
-involving screening hundreds of thousands of actual or theoretical MOFs.
+1. Referee 1 states that "Phenomenological predictions without any relation to
+   real materials are not very helpful, if not useless."  The referee fails to
+   grasp that our theory *does* take into account the gas, and places a
+   fundamental limit on the capacity of an entire class of materials (inflexible
+   porous materials). This is far from useless.  The referee suggests as
+   superior an approach involving screening hundreds of thousands of actual or
+   theoretical MOFs.
 
@@
 capacity, since it is always possible that there is a better structure out
-    there. We have edited the paragraph on page 3 of the paper that discusses
-    this sort of study to make this more clear (and added a reference the paper
-    by Ahmed et al.).
+there. We have edited the paragraph on page 3 of the paper that discusses this
+sort of study to make this more clear (and added a reference the paper by Ahmed
+et al.).
 
-The referee gives a litany of structural properties that we do not address in
+2. The referee gives a litany of structural properties that we do not address in
    our model: "pore volume, pore size, specific surface area, crystal density,
@@
 
-    We note that some of these (e.g. crystal density, skeletal density)
-    manifestly have no impact on the property we are studying, which is
-    volumetric storage capacity.  Others of course do impact the volumetric
-    storage capacity of a material, but do not relate to an upper bound
-    specifically.
+We note that some of these (e.g. crystal density, skeletal density) manifestly
+have no impact on the property we are studying, which is volumetric storage
+capacity.  Others of course do impact the volumetric storage capacity of a
+material, but do not relate to an upper bound specifically.
 
@@
 
-The other suggestion of Referee 1 is that we clearly state the capacity that we report, as discussed by Parilla et al.
+3. The final suggestion of Referee 1 is that we clearly state the capacity that  
+   we report, as discussed by Parilla et al.
 
-   The deliverable capacity is defined in the paper as "the density of the gas
-   in the material at the storage pressure p_full minus the residual gas that
-   remains adsorbed at the lowest pressure p_empty such that sufficient flow is
-   maintained to feed the engine."  We believe this is clear. The various
-   capacities of Parilla et al. are geared towards experimental measurements and
-   account for experimental uncertainty in the volume of material, an issue
-   which is not present in a theoretical paper.
-We note that the paper Referee 1 brought up as an example does not any more clearly define "usable capacity" as they refer to what we call "deliverable capacity".
-They do cite Parilla et al., but only in the context of describing their experimental equipment.
+The deliverable capacity is defined in the paper as "the density of the gas in
+the material at the storage pressure p_full minus the residual gas that remains
+adsorbed at the lowest pressure p_empty such that sufficient flow is maintained
+to feed the engine."  We believe this is clear. The various capacities of
+Parilla et al. are geared towards experimental measurements and account for
+experimental uncertainty in the volume of material, an issue which is not
+present in a theoretical paper.  We note that the paper Referee 1 brought up as
+an example does not any more clearly define "usable capacity" as they refer to
+what we call "deliverable capacity".  They do cite Parilla et al., but only in
+the context of describing their experimental equipment.
 
@@
 
-1. The referee suggests that we attempt to quantify the possible effect of a
+4. The referee suggests that we attempt to quantify the possible effect of a
    flexible MOF on our upper bound.
@@
 
-1. The referee asks that we provide some guidelines in terms of working
+5. The referee asks that we provide some guidelines in terms of working
    conditions to reach the DOE target.  They further suggest looking at a
@@
 
-We added a sentence mentioning that we cannot apply our upper bound in the case
+We added a short explanation of why we cannot apply our upper bound in the case
 where the temperature is changed, because a uniform potential may no longer lead
 to the optimal deliverable capacity.
+Specifically, a non-uniform potential will decrease the entropy of the adsorbed gas, and that decrease will make it easier to desorb the gas by raising the temperature.

### 2021-07-07 gas-adsorption/rsc-gas-ads-complete.tex
add explanation of why non-uniform potential could be good if temperature changes

@@
 
-\startreferees{Added last sentence to clarify why we cannot place a bound
-  on the deliverable capacity when the temperature is changed.}
+\startreferees{Added an explanation of why we cannot place a bound on the
+  deliverable capacity when the temperature is changed.}
 
@@
 changing temperature a uniform potential may not give the maximum deliverable
-capacity.  \davidsays{As far as I can intuit, \emph{normally} the entropy
-increases with temperature, and if that is the case then a uniform potential
-will give the optimum deliverable capacity.  If we can prove that for a subset
-of substrate interactions (e.g. when the substrate acts as an external potential
-felt by the gas), the optimal potential is uniform then we could apply our model
-to temperature-pressure swings.}
+capacity.  A nonuniform potential will lower the entropy of the adsorbed gas,
+which will make the gas easier to release by raising the temperature.
 

### 2021-07-07 gas-adsorption/referee-response.md
add paragraph explaining about markings in the paper for changed sections.

@@
 
-The responses of referees 1 and 3 could hardly be more divergent.
-Our work is impactful precisely because it can tell researchers when to stop looking...
-or perhaps more relevantly it can tell funding agencies when to stop funding research in this direction.
-This is threatening, and Referee 1 doesn't seem interested in knowing whether we should stop looking at porous materials for natural gas and hydrogen storage.
+Thank you so much for being so flexible with extensions to the deadline for our
+paper revision.  We have made a few major changes in response in particular to
+suggestions by referee 3 (see below).  The revised PDF has substantially
+modified sections noted with a red bar in the margin, and a short red note
+explaining what the change was.
+
+The responses of referees 1 and 3 could hardly be more divergent.  Our work is
+impactful precisely because it can tell researchers when to stop looking... or
+perhaps more relevantly it can tell funding agencies when to stop funding
+research in this direction.  This is threatening, and Referee 1 doesn't seem
+interested in knowing whether we should stop looking at porous materials for
+natural gas and hydrogen storage.
 

### 2021-07-08 gas-adsorption/rsc-gas-ads-complete.tex
edits from jordan

@@
 Finally, a number of empirical studies have been performed that computationally
-simulate a adsorption in a vast array of possible porous
+simulate an adsorption in a vast array of possible porous
 structures~\cite{firlej2013understanding, goldsmith2013theoretical,
@@
     \caption{Cartoon of a two-phase, collapsing material.
-The material can exist in two phases: an open phase in which gas molecules can adsorb and a closed phase that is collapsed and cannot fit any gas molecules.
-Take the closed phase as more stable, so that the material can collapse and expels residual gas.
+The material can exist in two phases: an open phase in which gas molecules can adsorb, and a closed phase that is collapsed and cannot fit any gas molecules.
+Take the closed phase as more stable, so that the material can collapse and expel residual gas.
     The pressure of gas at which the structural transition occurs depends on the gas-host interactions and relative stability of the phases.
@@
       between open and closed phases. The experimental deliverable capacity for Co(bdp) is shown for
-       comparison~\cite{fixme}, along with the free energy difference between open and closed conformations~\cite{fixthis}.}
+       comparison~\cite{mason2015methane}, along with the free energy difference between open and closed conformations~\cite{mason2015methane, choi2008broadly}.}
     \label{fig:methane-flexible}
@@
 For comparison, we include the free energy
-difference for Co(bdp)~\davidsays{CITE} and its measured deliverable capacity.  
+difference for Co(bdp)~\cite{choi2008broadly and its measured deliverable capacity.  
 While this upper bound is not rigorous, it suggests that a two-phase flexible material for methane storage
@@
 gasses of interest, with the proviso that the gas is far from crystallization.
-Our code is available at \href{https://github.com/Jordan-Pommerenck/isothermal-gas-adsorption}{https://github.com/Jordan-Pommerenck/isothermal-gas-adsorption}.
+Our code is available at \href{https://https://github.com/droundy/thesis-pommerenck/tree/master/gas-adsorption}{https://https://github.com/droundy/thesis-pommerenck/tree/master/gas-adsorption}.
 

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
fix to latex
edit caption

--- 2021-07-07 gas-adsorption/rsc-gas-ads-complete.tex
@@
 \cbcolor{red}
-\newcommand\startreferees[1]{\cbstart \noindent\textit{\color{red}%
-[\textbf{To referees:} #1]}
-
-}
-\newcommand\donereferees{\cbend}
 
@@
 \newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommand{\corysays}[1]{{\color{red} [\blue{Cory:} \emph{#1}]}}
 
@@
 
-\startreferees{Deliverable capacity defined below}
 For a vehicle employing a porous material to store natural gas or hydrogen, the
@@
 
-\donereferees
-
 % \begin{figure}
@@
 
-\startreferees{Added discussion of screening studies}
-Finally, a number of empirical studies have been performed that computationally
-simulate a adsorption in a vast array of possible porous
-structures~\cite{firlej2013understanding, goldsmith2013theoretical,
-ahmed2019exceptional}.  Such studies can guide the development of new materials
-by allowing researchers to identify commonalities among materials that have
-desirable properties.  However, such an approach can only place a \emph{lower
-bound} on the highest attainable deliverable capacity.
-
-\donereferees
+Finally, empirical models that statistically relate e.g., surface area to 
+hydrogen adsorption \cite{firlej2013understanding,goldsmith2013theoretical}
+have been used to determine what properties a material must possess to reach
+a given storage target.
 
@@
     \includegraphics[width=\columnwidth]{g_st_simple.pdf}
-    \caption{Cartoon illustrating the definition of $\gst$.
-For a given adsorbent and gas, $\gst$=$\gst (\rho, T)$ is defined as the difference in chemical potential between the adsorbed gas system and bulk gas system when both exhibit the same density $\rho$ and temperature $T$.}
+    \caption{A illustrative thought experiment to measure $\gst$ by adjusting the altitute difference between two connected volumes, one free space and another containing a porous material.
+    $\gst$ is the difference in the gravitational potential energy when the density of gas in each volume is equal.
+    % For a given adsorbent and gas, $\gst$=$\gst (\rho, T)$ is defined as the difference in chemical potential between the adsorbed gas system and bulk gas system when both exhibit the same density $\rho$ and temperature $T$.
+    }
     \label{fig:delta-G-cartoon}
@@
 in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
-% Figure~\ref{fig:delta-G-cartoon} illustrates  $\gst$.
+Figure~\ref{fig:delta-G-cartoon} illustrates $\gst$.
 % David:  no longer illustrates the expt.
-% A hypothetical experiment to measure $\gst$ is via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
-% Initialize the experiment with all gas in the free space and an empty substrate.
-% Then, slowly (i) remove the partition and (ii) use the piston to push the gas into the substrate and eliminate the free space volume, in an isothermal process.
-% $\gst$ is the ... \corysays{wut? energy it takes to do this or?}
-%
-% \davidsays{Okay, here is an improved experiment to measure $\gst$.  Two containers of
-% equal volume are connected by a hose (with neglible volume).  One container holds
-% the substrate, and the other is empty (apart from gas).    The altitude of one container
-% is adjusted until the number of gas molecules is equal in each container.  The value
-% of $\gst$ is the gravitational potential energy difference of gas molecules in the two
-% boxes.}
-%
-Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to
-measure $\gst$.  Two containers of equal volume are connected by a hose.  One
-container holds the substrate, and the other is empty (apart from gas).    The
-altitude of one container is adjusted until the number of gas molecules is equal
-in each container.  The value of $\gst$ is the gravitational potential energy
-difference of gas molecules in the two boxes.
-
+A hypothetical experiment to measure $\gst$ is via a piston with a removable partition that separates a volume of free space from the same volume of substrate.
+Initialize the experiment with all gas in the free space and an empty substrate.
+Then, slowly (i) remove the partition and (ii) use the piston to push the gas into the substrate and eliminate the free space volume, in an isothermal process.
+$\gst$ is the ... \davidsays{wut? energy it takes to do this or?}
 % David: what is GST
@@
 
-\startreferees{Flexible MOFs section added below}
+\cbstart
 \begin{figure}
@@
 collapsed, ``closed'' phase in which no gas can fit inside.
-Let $\Delta F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free
-energy difference between the two phases in the absence of gas.
+Let $\Delta G=G_{\text{open}}-G_{\text{closed}}$ be the free energy difference between the two phases in the absence of gas.
 % David: introduce units of delta G here.
-Suppose the closed phase is more stable ($\implies \Delta F >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
+Suppose the closed phase is more stable ($\implies \Delta G >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
 In addition, we assume the open phase hosts a uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb, so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.
@@
 \begin{align}
-    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta F -pV_{\text{open}} \\
-    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta F -pV_{\text{closed}}
+    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta G \\
+    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta G
 \end{align}
@@
 % David: specify units on \delta G so clear where V comes from; might be better to absorb V into delta G?
-where $\rho_{\text{open}}(p)$ is the density of gas
-in the open phase of the material at pressure $p$. We will neglect the $pV$
-terms in the Gibbs free energy, under the assumption that the mechanical
-work associated with the transition is neglible when compared with the
-adsorption energies involved.
+where $\rho_{\text{open}}(p)$ is the density of gas in the open phase of the material at pressure $p$. 
 I.e., (1, to be open at $\pfull$) the gas-host attractions at $\pfull$ must overcome the free energy penalty to open the structure and 
@@
 This approximate model of the two-phase material has no upper bound on the deliverable capacity, since
-$\V$ and $\Delta F$ can together be increased without bound.
-We can, however, place a bound for a fixed $\Delta F$, since if the gas is too strongly attracted to the open phase,
+$\V$ and $\Delta G$ can together be increased without bound.
+We can, however, place a bound for a fixed $\Delta G$, since if the gas is too strongly attracted to the open phase,
 the material will remain open at $\pempty$ and our rigid upper bound applies.
@@
 \begin{align}
-    V_{\text{open}} \rho_g(\muempty{} - \V_{opt}, T)\V_{opt} = \Delta F.
+    V_{\text{open}} \rho_g(\muempty{} - \V_{opt}, T)\V_{opt} = \Delta G.
 \end{align}
@@
 Figure~\ref{fig:methane-flexible} illustrates
-how the upper bound on the deliverable capacity of methane in a two-phase,
-collapsing material depends on $\Delta F$. 
+how the upper bound on the deliverable capacity in a two-phase, collapsing material depends on $\Delta G$ for methane. 
 For comparison, we include the free energy
@@
 
-\donereferees
+\cbend
 
@@
 
-\startreferees{Added an explanation of why we cannot place a bound on the
-  deliverable capacity when the temperature is changed.}
-
 The first limitation of our proof is that we restricted ourselves to
@@
 and vehicle.
-%
-We cannot place an upper bound on such a process using our approach because with
-changing temperature a uniform potential may not give the maximum deliverable
-capacity.  A nonuniform potential will lower the entropy of the adsorbed gas,
-which will make the gas easier to release by raising the temperature.
-
-\donereferees
 

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex

--- 2021-07-08 gas-adsorption/rsc-gas-ads-complete.tex
@@
 For comparison, we include the free energy
-difference for Co(bdp)~\cite{choi2008broadly and its measured deliverable capacity.  
+difference for Co(bdp)~\cite{choi2008broadly} and its measured deliverable capacity.  
 While this upper bound is not rigorous, it suggests that a two-phase flexible material for methane storage

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
osmotic pot

@@
 For example, consider a material such as Co(bdp) \cite{mason2015methane} that 
-can, actuated by the presence of gas, switch between two structural phases: 
+can, actuated by the presence of gas, switch between two distinct structural phases: 
 an ``open'' phase where gas molecules can fit in the pore space and a 
 collapsed, ``closed'' phase in which no gas can fit inside.
-Let $\Delta F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free
-energy difference between the two phases in the absence of gas.
-% David: introduce units of delta G here.
+
+Let $\Delta F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free energy difference per mole of adsorbent between the two phases in the absence of gas.
 Suppose the closed phase is more stable ($\implies \Delta F >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
 In addition, we assume the open phase hosts a uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb, so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.
+In terms of $\rho_g(\mu, T)$, then, the grand potential $\Omega$ in our idealized substrate is:
+\begin{equation}
+    \Omega(T, \mu) = - \int_{-\infty}^\mu \rho_g(\mu-\V)d\mu.
+\end{equation}
+See Coudert et al. \cite{coudert2008thermodynamics}.
+However, the \emph{osmotic potential}.
+% David: introduce units of delta G here.
+
 For an enhancement of the deliverable capacity by this two-phase, flexible material, 
@@
 \begin{align}
-    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta F -pV_{\text{open}} \\
-    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta F -pV_{\text{closed}}
+    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta F \\
+    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta F
 \end{align}
@@
 where $\rho_{\text{open}}(p)$ is the density of gas
-in the open phase of the material at pressure $p$. We will neglect the $pV$
-terms in the Gibbs free energy, under the assumption that the mechanical
-work associated with the transition is neglible when compared with the
-adsorption energies involved.
+in the open phase of the material at pressure $p$ and $V_{\text{open}}$ is the molar volume of the open phase of the adsorbent, and we neglect the $pV$
+terms in the Gibbs free energy.
+%under the assumption that the mechanical
+%work associated with the transition is neglible when compared with the
+%adsorption energies involved.
 I.e., (1, to be open at $\pfull$) the gas-host attractions at $\pfull$ must overcome the free energy penalty to open the structure and 

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
fix typo

@@
     \includegraphics[width=\columnwidth]{g_st_simple.pdf}
-    \caption{A illustrative thought experiment to measure $\gst$ by adjusting the altitute difference between two connected volumes, one free space and another containing a porous material.
+    \caption{An illustrative thought experiment to measure $\gst$ by adjusting the altitute difference between two connected volumes, one free space and another containing a porous material.
     $\gst$ is the difference in the gravitational potential energy when the density of gas in each volume is equal.

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
edit empirical study paragraph. and fix reference

@@
 material space, i.e.\ that the structures considered are representative of the
-set of possible materials \cite{Moosavi2020}. To further explore material space and address
+set of possible materials \cite{Moosavi2020understanding}. To further explore material space and address
 sensitivity to the intermolecular potentials: scaling the Lennard-Jones
@@
 \startreferees{Added discussion of screening studies}
-Finally, a number of empirical studies have been performed that computationally
-simulate an adsorption in a vast array of possible porous
-structures~\cite{firlej2013understanding, goldsmith2013theoretical,
-ahmed2019exceptional}.  Such studies can guide the development of new materials
+Finally, in a number of studies, molecular simulations of gas adsorption in
+a large pool of both existing and predicted nanoporous
+structures empirically shed light on performance limitations \cite{firlej2013understanding, goldsmith2013theoretical,
+ahmed2019exceptional,simon2015materials}.
+Such studies can guide the development of new materials
 by allowing researchers to identify commonalities among materials that have
-desirable properties.  However, such an approach can only place a \emph{lower
-bound} on the highest attainable deliverable capacity.
-
+desirable properties.  
+However, the largest observed simulated deliverable capacity among the pool of candidate materials in such studies
+is only a \emph{lower bound} on the highest attainable deliverable capacity that is is predicted upon the accuracy of the molecular models.
 \donereferees

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
we dont look at any 'ol condition
mason is all we need to cite for that; both numbers from there

@@
 capacity. We use our framework to place an upper bound on the deliverable
-capacity of methane and hydrogen gas.
+capacity of methane and hydrogen gas at conditions relevant to vehicular storage.
 
@@
       between open and closed phases. The experimental deliverable capacity for Co(bdp) is shown for
-       comparison~\cite{mason2015methane}, along with the free energy difference between open and closed conformations~\cite{mason2015methane, choi2008broadly}.}
+       comparison, along with the free energy difference between open and closed conformations~\cite{mason2015methane}.}
     \label{fig:methane-flexible}
@@
     \includegraphics[width=0.95\columnwidth]{hydrogen-298-flexible}
-    \caption{Deliverable capacity of hydrogen in a flexible substrate at 298\ K as a function of the free energy difference 
+    \caption{Deliverable capacity of hydrogen in a flexible, two-phase substrate at 298\ K as a function of the free energy difference 
       between open and closed phases.}

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
in discussion, note how we did address flexible models

@@
 %
-We cannot place an upper bound on such a process using our approach because with
-changing temperature a uniform potential may not give the maximum deliverable
+We cannot place an upper bound on such a process using our approach because, with
+changing temperature, a uniform potential may not give the maximum deliverable
 capacity.  A nonuniform potential will lower the entropy of the adsorbed gas,
@@
 residual gas at low pressures. Flexible materials could have significantly 
-higher deliverable capacities, as our upper bound does not pertain to them.
+higher deliverable capacities, as we have shown for two-phase structures 
+that collapse and expel all residual gas at low pressures.
 

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
edits

@@
 \end{figure}
-Although our deliverable capacity limit does not directly apply to materials that change conformation in
-response to gas adsorption, we can estimate enhancements in the deliverable capacity provided by
-such materials. 
-For example, consider a material such as Co(bdp) \cite{mason2015methane} that 
-can, actuated by the presence of gas, switch between two distinct structural phases: 
+While our deliverable capacity limit does not apply to materials that change conformation in
+response to gas adsorption, here we adapt our theoretical framework to pertain to bistable adsorbents modeled after Co(bdp) \cite{mason2015methane}.
+Actuated by the presence of gas, Co(bdp) can switch between two distinct structural phases:
 an ``open'' phase where gas molecules can fit in the pore space and a 
@@
 
+We follow Coudert et al. \cite{coudert2008thermodynamics} in developing a thermodynamic model for a two-phase adsorbent. 
 Let $\Delta F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free energy difference per mole of adsorbent between the two phases in the absence of gas.
 Suppose the closed phase is more stable ($\implies \Delta F >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
-In addition, we assume the open phase hosts a uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb, so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.
+In addition, assume the open phase hosts a uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb, so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.
 In terms of $\rho_g(\mu, T)$, then, the grand potential $\Omega$ in our idealized substrate is:
@@
 \end{equation}
-See Coudert et al. \cite{coudert2008thermodynamics}.
-However, the \emph{osmotic potential}.
+However, the \emph{osmotic potential} 
 % David: introduce units of delta G here.

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
fix plot, and some edits on the paper

@@
 \end{figure}
-While our deliverable capacity limit does not apply to materials that change conformation in
-response to gas adsorption, here we adapt our theoretical framework to pertain to bistable adsorbents modeled after Co(bdp) \cite{mason2015methane}.
-Actuated by the presence of gas, Co(bdp) can switch between two distinct structural phases:
-an ``open'' phase where gas molecules can fit in the pore space and a 
-collapsed, ``closed'' phase in which no gas can fit inside.
-
-We follow Coudert et al. \cite{coudert2008thermodynamics} in developing a thermodynamic model for a two-phase adsorbent. 
-Let $\Delta F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free energy difference per mole of adsorbent between the two phases in the absence of gas.
-Suppose the closed phase is more stable ($\implies \Delta F >0$), so that upon a reduction in the pressure of the gas, the material will collapse and expel all residual gas to enhance the deliverable capacity compared to rigid materials that will retain some gas at $\pempty$.
-In addition, assume the open phase hosts a uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb, so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.
-In terms of $\rho_g(\mu, T)$, then, the grand potential $\Omega$ in our idealized substrate is:
+While our deliverable capacity limit does not apply to materials that change
+conformation in response to gas adsorption, here we adapt our theoretical
+framework to pertain to bistable adsorbents modeled after Co(bdp)
+\cite{mason2015methane}. Actuated by the presence of gas, Co(bdp) can switch
+between two distinct structural phases: an ``open'' phase where gas molecules
+can fit in the pore space and a collapsed, ``closed'' phase in which no gas can
+fit inside.
+
+We follow Coudert et al. \cite{coudert2008thermodynamics} in developing a
+thermodynamic model for a two-phase adsorbent. Let $\Delta
+F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free energy difference per
+mole of adsorbent between the two phases in the absence of gas. Suppose the
+closed phase is more stable ($\implies \Delta F >0$), so that upon a reduction
+in the pressure of the gas, the material will collapse and expel all residual
+gas to enhance the deliverable capacity compared to rigid materials that will
+retain some gas at $\pempty$. In addition, assume the open phase hosts a
+uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb,
+so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.  This
+approximate model of the two-phase material has no upper bound on the
+deliverable capacity, since $\V$ and $\Delta F$ can together be increased
+without bound.  We can, however, place a bound for a fixed $\Delta F$, since if
+the gas is too strongly attracted to the open phase, the material will remain
+open at $\pempty$ and our rigid upper bound applies.
+
+The upper bound for the two-phase material is found by identifying the
+attractive interaction $\V$ that balances the gas-host attractions in the
+idealized-open-substrate at $\pempty$ with the free energy penalty to open.
+
+In
+terms of $\rho_g(\mu, T)$, then, the grand potential $\Omega$ in our idealized
+substrate is:
 \begin{equation}
@@
 \end{equation}
-However, the \emph{osmotic potential} 
+However, the \emph{osmotic potential} governs which phase the material adopts.
+The material changes phases when osmotic potentials of the two phases are equal.
+% At the pressure at which they switch when the osmotic potentials in the two phases
+% are equal.
+The upper bound for the two-phase material is found by identifying the largest
+possible attraction $|\V|$ consistent with the material transitioning between
+$\pempty$ and $\pfull$ to expel all residual gas.
+%
+This means the material should change phases at pressure $\pempty$, which gives
+\begin{align}
+    % F_{closed} + 0 &= F_{open} + \Omega(T, \mu) \\
+    \frac{\Delta F}{V_{\text{open}}} &= \int_{-\infty}^{\muempty} \rho_g(\mu-\V_{opt})d\mu.
+\end{align}
+where we follow Coudert \emph{et al.}\cite{coudert2008thermodynamics} and
+neglect the $pV$ terms, and $V_{\text{open}}$ is the molar volume of the open
+phase.
+
 % David: introduce units of delta G here.
@@
 %adsorption energies involved.
-I.e., (1, to be open at $\pfull$) the gas-host attractions at $\pfull$ must overcome the free energy penalty to open the structure and 
-(2, to be closed at $\pempty$) the free energy penalty to open the structure must be larger than the (weak) gas-host attractions at $\pempty$.
-This approximate model of the two-phase material has no upper bound on the deliverable capacity, since
-$\V$ and $\Delta F$ can together be increased without bound.
-We can, however, place a bound for a fixed $\Delta F$, since if the gas is too strongly attracted to the open phase,
-the material will remain open at $\pempty$ and our rigid upper bound applies.
+I.e., (1, to be open at $\pfull$) the gas-host attractions at $\pfull$ must
+overcome the free energy penalty to open the structure and (2, to be closed at
+$\pempty$) the free energy penalty to open the structure must be larger than the
+(weak) gas-host attractions at $\pempty$. This approximate model of the
+two-phase material has no upper bound on the deliverable capacity, since $\V$
+and $\Delta F$ can together be increased without bound. We can, however, place a
+bound for a fixed $\Delta F$, since if the gas is too strongly attracted to the
+open phase, the material will remain open at $\pempty$ and our rigid upper bound
+applies.
 % David: does our rigid upper bound apply then?
 The upper bound for the two-phase material is found by identifying the
-attractive interaction $\V$ that balances the gas-host attractions in the idealized-open-substrate at $\pempty$ with the free energy penalty to open:
+attractive interaction $\V$ that balances the gas-host attractions in the
+idealized-open-substrate at $\pempty$ with the free energy penalty to open:
 \begin{align}

### 2021-07-09 gas-adsorption/isothermal-ads-gas-plot.py

@@ -352,35 +352,29 @@ plt.legend(loc='lower right')
 plt.tight_layout()
 plt.savefig('figs/' + basename + '-' + temperature + '-gst.pdf', transparent=True)
 
-plt.figure('capacity vs Delta F', figsize=(5,3.75))
+fig, ax = plt.subplots(2, 1, sharex=True, figsize=(5,3.75),
+                       gridspec_kw={'height_ratios': [0.8,2]})
 
-# delta_F_units = kJ/cm**3
-# delta_F_unit_name = 'kJ/cm$^3$'
 delta_F_units = J/cm**3
 delta_F_unit_name = 'J/cm$^3$'
-# delta_F_units = kJ/angstrom**3
-# delta_F_unit_name = f'kJ/$\AA^3$'
 
+max_delta_F = 0.03*kJ/cm**3
 
-delta_F_flexible = np.linspace(0, 0.04*kJ/cm**3, 300)
-capacities = np.zeros_like(delta_F_flexible)
+capacities = np.zeros_like(Gads)
 
 def n_from_mu_with_nans(particular_mu):
     return np.interp(particular_mu, mu, n, left=np.nan, right=np.nan)
 
-mu_lo = mu_from_p(p_empty)
-n_lo = n_from_mu_with_nans(mu_lo + Gads)
-n_hi = n_from_mu_with_nans(mu_from_p(p_full) + Gads)
-print('n_lo', n_lo/density_units)
-for i in range(len(capacities)):
-    # to_minimize = abs(n_lo*Gads - delta_F_flexible[i])
-    # to_minimize[np.isnan(to_minimize)] = 1e200
-    # which = np.argmin(to_minimize)
-    # capacities[i] = n_hi[which]
-    capacities[i] = np.interp(delta_F_flexible[i], n_lo*Gads, n_hi)
+mu_values = np.linspace(mu[0], mu_empty, 100000)
+dmu = mu_values[1] - mu_values[0]
+grand_free_energy_full = np.zeros_like(Gads)
+for i in range(len(Gads)):
+    grand_integrand = n_from_mu_with_nans(mu_values + Gads[i])
+    grand_free_energy_full[i] = grand_integrand.sum()*dmu
+    capacities[i] = n_from_mu_with_nans(mu_full + Gads[i])
 
 for n_goal, label, color, style in gas.n_goals:
-    plt.text(delta_F_flexible[len(delta_F_flexible)//4]/delta_F_units,
[…]

### 2021-07-09 gas-adsorption/rsc-gas-ads-complete.tex
k

@@
 We follow Coudert et al. \cite{coudert2008thermodynamics} in developing a
-thermodynamic model for a two-phase adsorbent. Let $\Delta
-F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free energy difference per
+thermodynamic model for a two-phase, collapsing adsorbent like Co(bdp). 
+Let $\Delta F=F_{\text{open}}-F_{\text{closed}}$ be the Helmholtz free energy difference per
 mole of adsorbent between the two phases in the absence of gas. Suppose the
 closed phase is more stable ($\implies \Delta F >0$), so that upon a reduction
-in the pressure of the gas, the material will collapse and expel all residual
+in the pressure of the gas, the material can collapse and expel all residual
 gas to enhance the deliverable capacity compared to rigid materials that will
@@
 uniform, attractive potential energy field $\V<0$ for gas molecules to adsorb,
-so that eqn.~\ref{eq:xi_vs_xi0} holds for the open phase of the material.  This
-approximate model of the two-phase material has no upper bound on the
-deliverable capacity, since $\V$ and $\Delta F$ can together be increased
-without bound.  We can, however, place a bound for a fixed $\Delta F$, since if
+so that eqn.~\ref{eq:xi_vs_xi0} holds for the material when in the open phase.
+
+This model of a two-phase, collapsing material has no upper bound on the
+deliverable capacity, since $|\V|$ and $\Delta F$ can together be increased
+without bound.
+We can, however, place a bound on the deliverable capacity for a fixed $\Delta F$, since if
 the gas is too strongly attracted to the open phase, the material will remain
-open at $\pempty$ and our rigid upper bound applies.
-
-The upper bound for the two-phase material is found by identifying the
+open at $\pempty$, and, then, our rigid upper bound applies.
+We place an upper bound on the deliverable capacity in the two-phase material by identifying the
 attractive interaction $\V$ that balances the gas-host attractions in the
@@
 
-In
-terms of $\rho_g(\mu, T)$, then, the grand potential $\Omega$ in our idealized
-substrate is:
-\begin{equation}
-    \Omega(T, \mu) = - \int_{-\infty}^\mu \rho_g(\mu-\V)d\mu.
-\end{equation}
-However, the \emph{osmotic potential} governs which phase the material adopts.
-The material changes phases when osmotic potentials of the two phases are equal.
+For an enhancement of the deliverable capacity by this two-phase, flexible material, 
+the open phase must be preferred at $\pfull$, with the closed phase preferred at $\pempty$. 
 % At the pressure at which they switch when the osmotic potentials in the two phases
@@
 $\pempty$ and $\pfull$ to expel all residual gas.
+The material changes phases when the osmotic potentials of the open and closed systems 
+with and without gas, respectively, are equal \cite{coudert2008thermodynamics} (neglecting any hystersis).
+Thus, the optimal background potential $\V_{opt}$ equates the grand potential of the gas in the open phase 
+with the free energy difference between the two phases at $\pempty$:
+%In terms of $\rho_g(\mu, T)$, the grand potential $\Omega$ in our idealized
+%substrate is:
+%\begin{equation}
+%    \Omega(T, \mu) = - \int_{-\infty}^\mu \rho_g(\mu-\V)d\mu.
+%\end{equation}
 %
-This means the material should change phases at pressure $\pempty$, which gives
+% This means the material should change phases at pressure $\pempty$, which gives
 \begin{align}
     % F_{closed} + 0 &= F_{open} + \Omega(T, \mu) \\
-    \frac{\Delta F}{V_{\text{open}}} &= \int_{-\infty}^{\muempty} \rho_g(\mu-\V_{opt})d\mu.
+    \frac{\Delta F}{V_{\text{open}}} &= \int_{-\infty}^{\muempty} \rho_g(\mu-\V_{opt})d\mu, \label{eq:v_opt_bistable}
 \end{align}
@@
 phase.
-
-% David: introduce units of delta G here.
-
-For an enhancement of the deliverable capacity by this two-phase, flexible material, 
-the open phase must be preferred at $\pfull$, with the closed phase preferred at $\pempty$. 
-This requires that, neglecting gas-gas interactions:
-% David: doesn't this also neglect entropy? wut about the *free energy* of the gas?
-% David: I suggest \Delta G_{host} to be clear?
-\begin{align}
-    V_{\text{open}}|\rho_{\text{open}}(\pfull)\V| &> \Delta F \\
-    V_{\text{open}}|\rho_{\text{open}}(\pempty)\V| &< \Delta F
-\end{align}
-% David: I put \V<0, think we should say the sign of \Delta G to be clear too.
-% David: specify units on \delta G so clear where V comes from; might be better to absorb V into delta G?
-where $\rho_{\text{open}}(p)$ is the density of gas
-in the open phase of the material at pressure $p$ and $V_{\text{open}}$ is the molar volume of the open phase of the adsorbent, and we neglect the $pV$
-terms in the Gibbs free energy.
-%under the assumption that the mechanical
-%work associated with the transition is neglible when compared with the
-%adsorption energies involved.
-I.e., (1, to be open at $\pfull$) the gas-host attractions at $\pfull$ must
-overcome the free energy penalty to open the structure and (2, to be closed at
-$\pempty$) the free energy penalty to open the structure must be larger than the
-(weak) gas-host attractions at $\pempty$. This approximate model of the
-two-phase material has no upper bound on the deliverable capacity, since $\V$
-and $\Delta F$ can together be increased without bound. We can, however, place a
-bound for a fixed $\Delta F$, since if the gas is too strongly attracted to the
-open phase, the material will remain open at $\pempty$ and our rigid upper bound
-applies.
-% David: does our rigid upper bound apply then?
-The upper bound for the two-phase material is found by identifying the
-attractive interaction $\V$ that balances the gas-host attractions in the
-idealized-open-substrate at $\pempty$ with the free energy penalty to open:
-\begin{align}
-    V_{\text{open}} \rho_g(\muempty{} - \V_{opt}, T)\V_{opt} = \Delta F.
-\end{align}
-% David: left is negative. what is the right? negative as well? how did u define it?
-This gives the largest possible attraction $|\V|$ consistent with the material closing
-at $\pempty$ to expel all residual gas. 
-The deliverable capacity is then the density of gas in the open phase at $\pfull$, $\rho_g(\mufull{} - \V_{opt}, T)$.
+I.e., $|\V_{opt}|$ gives the maximal gas-open-host attractions under the constraint that the material collapses and expels all residual gas at $\pempty$; if the attraction were greater, the material would refuse to collapse and expel its gas at $\pfull$.
+The deliverable capacity is then the density of gas in the open phase at $\pfull$, $\rho_g(\mufull{} - \V_{opt}, T)$ with $\V_{opt}$ satisfying eqn.~\ref{eq:v_opt_bistable}.
 % David: changed it to \V_opt, or not necessary?
@@
 how the upper bound on the deliverable capacity of methane in a two-phase,
-collapsing material depends on $\Delta F$. 
+collapsing material depends on $\Delta F$-- and the $\V_{opt}$ that maximizes the deliverable capacity. 
 For comparison, we include the free energy
 difference for Co(bdp)~\cite{choi2008broadly} and its measured deliverable capacity.
-While this upper bound is not rigorous, it suggests that a two-phase flexible material for methane storage
-would benefit from significantly greater stability of the closed phase---provided the open phase
-had correspondingly greater attraction for the gas molecules.  Figure~\ref{fig:hydrogen-flexible}
-shows the corresponding result for hydrogen storage, indicating that, for a flexible material to
+Indeed, the deliverable capacity of a two-phase, collapsing material can exceed that of a rigid material if 
+(i) the closed phase is sufficiently stable relative to the open phase
+and
+(ii) the gas has a correspondingly high affinity for the open phase.
+Figure~\ref{fig:hydrogen-flexible} shows the corresponding result for hydrogen storage, indicating that, for a flexible material to
 reach the DOE ULTIMATE deliverable capacity, it would require far greater stability for the closed

### 2021-07-09 gas-adsorption/isothermal-ads-gas-plot.py
add label for normal upper bound

@@ -374,7 +374,7 @@ for i in range(len(Gads)):
     capacities[i] = n_from_mu_with_nans(mu_full + Gads[i])
 
 for n_goal, label, color, style in gas.n_goals:
-    plt.text(max_delta_F/delta_F_units/4,
+    plt.text(max_delta_F/delta_F_units/40,
              n_goal/density_units, label, color=color,
              verticalalignment = 'center_baseline',
              horizontalalignment = 'left',
@@ -382,6 +382,14 @@ for n_goal, label, color, style in gas.n_goals:
     plt.axhline(n_goal/density_units, color=color, linestyle=style, linewidth=0.5)
 
 plt.axhline(ymax, color='g', linestyle='--', linewidth=0.5)
+upper_bound_x_position = 11.5
+if basename == 'methane':
+    upper_bound_x_position = 6
+plt.text(upper_bound_x_position,
+            ymax, 'UPPER BOUND NO PHASE CHANGE', color='g',
+            verticalalignment = 'center_baseline',
+            horizontalalignment = 'left',
+            bbox=dict(facecolor='white', edgecolor='white', alpha=1, pad=0))
 
 if basename == 'methane':
     volume_cobdp = 773.58*angstrom**3

### 2021-09-04 gas-adsorption/rsc-gas-ads-complete.tex
last edits?
structure prop rel.

@@
 
-\startreferees{Added discussion of screening studies}
 Finally, in a number of studies, molecular simulations of gas adsorption in
@@
 is only a \emph{lower bound} on the highest attainable deliverable capacity that is is predicted upon the accuracy of the molecular models.
-\donereferees
 
@@
 
-Fig.~\ref{fig:methane-298-D} clearly shows that $\V_{opt}$ provides a maximum
+Fig.~\ref{fig:methane-298-D} shows that $\V_{opt}$ provides a maximum
 $\rho_D$ over all \emph{spatially uniform}, static potential energy fields. To
@@
 which will make the gas easier to release by raising the temperature.
+
+\startreferees{Explain how the point of our paper is not to find structure-property relationships.}
+
+Several studies elucidated relationships between structural features of nanoporous materials---such as void fraction, pore size, and surface area---and their deliverable capacity of various gases \cite{gomez2014exploring,ahmed2019exceptional,simon2015materials,moghadam2018computer,Tong2018}.
+Such structure-property relationships serve as useful heuristics for designing new materials with high deliverable capacity.
+However, the intention of our paper is not to discover structure-property relationships.
+Rather, we place our upper bound on the deliverable capacity of [rigid] materials that holds irrespective of how the structural features of the material are tuned.
 

### 2021-09-04 gas-adsorption/rsc-gas-ads-complete.tex
added discussion abt grav vs vol del cap

@@
 frameworks~\cite{diercks2017atom}, no porous material has met these deliverable capacity targets.
-
-\donereferees
 
@@
 
-\startreferees{Flexible MOFs section added below}
 \begin{figure}
@@
 
-\donereferees
-
 \section{Conclusions}
@@
 which will make the gas easier to release by raising the temperature.
-
-\startreferees{Explain how the point of our paper is not to find structure-property relationships.}
-
-Several studies elucidated relationships between structural features of nanoporous materials---such as void fraction, pore size, and surface area---and their deliverable capacity of various gases \cite{gomez2014exploring,ahmed2019exceptional,simon2015materials,moghadam2018computer,Tong2018}.
-Such structure-property relationships serve as useful heuristics for designing new materials with high deliverable capacity.
-However, the intention of our paper is not to discover structure-property relationships.
-Rather, we place our upper bound on the deliverable capacity of [rigid] materials that holds irrespective of how the structural features of the material are tuned.
-
-\donereferees
 
@@
 
+\startreferees{Explain how the point of our paper is not to find structure-property relationships.}
+Several studies elucidated relationships between structural features of nanoporous materials---such as void fraction, pore size, and surface area---and their deliverable capacity of various gases \cite{gomez2014exploring,ahmed2019exceptional,simon2015materials,moghadam2018computer,Tong2018}.
+Such structure-property relationships serve as useful heuristics for designing new materials with high deliverable capacity.
+However, the intention of our paper is not to discover structure-property relationships.
+Rather, we place our upper bound on the deliverable capacity of [rigid] materials that holds irrespective of how the structural features of the material are tuned.
+
+\donereferees
+
+\startreferees{Discuss gravimetric vs volumetric deliverable capacity}
+
+For vehicular gas storage, gravimetric deliverable capacity (amount of gas per mass of material) may be considered as a material performance metric to avoid a heavy adsorbed-gas fuel tank.
+However, the volumetric deliverable capacity (amount of gas per volume of material) dominantly determines the driving range on a full fuel tank of a given volume \cite{mason2014evaluating}.
+That said, several articles have explored the tradeoff between gravimetric and volumetric deliverable capacity, particularly for hydrogen storage \cite{gomez2017understanding,chen2020balancing,goldsmith2013theoretical,ahmed2021predicting}.
+
+\donereferees
+
 We note that our upper bound can readily be applied to the storage of other

### 2021-09-27 gas-adsorption/gas-ads-revtex.tex
convert paper to revtex

--- 2021-09-04 gas-adsorption/rsc-gas-ads-complete.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%This is the LaTeX ARTICLE template for RSC journals
-%Copyright The Royal Society of Chemistry 2016
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[twoside,twocolumn,9pt]{article}
-\usepackage{extsizes}
-\usepackage[super,sort&compress,comma]{natbib} 
-\usepackage[version=3]{mhchem}
-\usepackage[left=1.5cm, right=1.5cm, top=1.785cm, bottom=2.0cm]{geometry}
-\usepackage{balance}
-\usepackage{mathptmx}
-\usepackage{sectsty}
+\documentclass[pre,twocolumn]{revtex4-2}
+
 \usepackage{graphicx} 
-\usepackage{lastpage}
-\usepackage[format=plain,justification=justified,singlelinecheck=false,font={stretch=1.125,small,sf},labelfont=bf,labelsep=space]{caption}
-\usepackage{float}
-\usepackage{fancyhdr}
-\usepackage{fnpos}
-\usepackage[english]{babel}
-\addto{\captionsenglish}{%
-  \renewcommand{\refname}{Notes and references}
-}
 \usepackage{array}
-\usepackage{droidsans}
-\usepackage{charter}
-\usepackage[T1]{fontenc}
 \usepackage[usenames,dvipsnames]{xcolor}
-\usepackage{setspace}
-\usepackage[compact]{titlesec}
-\usepackage{hyperref}
-%%%Please don't disable any packages in the preamble, as this may cause the template to display incorrectly.%%%
 
@@
 \usepackage{url}
+\usepackage{amsmath}
 
@@
 %
-\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
 \usepackage{mdframed}
@@
 
-\pagestyle{fancy}
-\thispagestyle{plain}
-\fancypagestyle{plain}{
-
-%%%HEADER%%%
-% \fancyhead[C]{\includegraphics[width=18.5cm]{head_foot/header_bar}}
-% \fancyhead[L]{\hspace{0cm}\vspace{1.5cm}\includegraphics[height=30pt]{head_foot/journal_name}}
-% \fancyhead[R]{\hspace{0cm}\vspace{1.7cm}\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}
-\renewcommand{\headrulewidth}{0pt}
-}
-%%%END OF HEADER%%%
-
-%%%PAGE SETUP - Please do not change any commands within this section%%%
-\makeFNbottom
-\makeatletter
-\renewcommand\LARGE{\@setfontsize\LARGE{15pt}{17}}
-\renewcommand\Large{\@setfontsize\Large{12pt}{14}}
-\renewcommand\large{\@setfontsize\large{10pt}{12}}
-\renewcommand\footnotesize{\@setfontsize\footnotesize{7pt}{10}}
-\makeatother
-
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
-\renewcommand\footnoterule{\vspace*{1pt}% 
-\color{cream}\hrule width 3.5in height 0.4pt \color{black}\vspace*{5pt}} 
-\setcounter{secnumdepth}{5}
-
-\makeatletter 
-\renewcommand\@biblabel[1]{#1}            
-\renewcommand\@makefntext[1]% 
-{\noindent\makebox[0pt][r]{\@thefnmark\,}#1}
-\makeatother 
-\renewcommand{\figurename}{\small{Fig.}~}
-\sectionfont{\sffamily\Large}
-\subsectionfont{\normalsize}
-\subsubsectionfont{\bf}
-\setstretch{1.125} %In particular, please do not alter this line.
-\setlength{\skip\footins}{0.8cm}
-\setlength{\footnotesep}{0.25cm}
-\setlength{\jot}{10pt}
-\titlespacing*{\section}{0pt}{4pt}{4pt}
-\titlespacing*{\subsection}{0pt}{15pt}{1pt}
-%%%END OF PAGE SETUP%%%
-
-%%%FOOTER%%%
-\fancyfoot{}
-\fancyfoot[LO,RE]{\vspace{-7.1pt}\includegraphics[height=9pt]{head_foot/LF}}
-\fancyfoot[CO]{\vspace{-7.1pt}\hspace{13.2cm}\includegraphics{head_foot/RF}}
-\fancyfoot[CE]{\vspace{-7.2pt}\hspace{-14.2cm}\includegraphics{head_foot/RF}}
-\fancyfoot[RO]{\footnotesize{\sffamily{1--\pageref{LastPage} ~\textbar  \hspace{2pt}\thepage}}}
-\fancyfoot[LE]{\footnotesize{\sffamily{\thepage~\textbar\hspace{3.45cm} 1--\pageref{LastPage}}}}
-\fancyhead{}
-\renewcommand{\headrulewidth}{0pt} 
-\renewcommand{\footrulewidth}{0pt}
-\setlength{\arrayrulewidth}{1pt}
-\setlength{\columnsep}{6.5mm}
-\setlength\bibsep{1pt}
-%%%END OF FOOTER%%%
-
-%%%FIGURE SETUP - please do not change any commands within this section%%%
-\makeatletter 
-\newlength{\figrulesep} 
-\setlength{\figrulesep}{0.5\textfloatsep}
-
-\newcommand{\topfigrule}{\vspace*{-1pt}% 
-\noindent{\color{cream}\rule[-\figrulesep]{\columnwidth}{1.5pt}} }
-
-\newcommand{\botfigrule}{\vspace*{-2pt}% 
-\noindent{\color{cream}\rule[\figrulesep]{\columnwidth}{1.5pt}} }
-
-\newcommand{\dblfigrule}{\vspace*{-1pt}% 
-\noindent{\color{cream}\rule[-\figrulesep]{\textwidth}{1.5pt}} }
-
-\makeatother
-%%%END OF FIGURE SETUP%%%
-
-%%%TITLE, AUTHORS AND ABSTRACT%%%
-\twocolumn[
-  \begin{@twocolumnfalse}
-{\includegraphics[height=30pt]{head_foot/journal_name}\hfill\raisebox{0pt}[0pt][0pt]{\includegraphics[height=55pt]{head_foot/RSC_LOGO_CMYK}}\\[1ex]
-\includegraphics[width=18.5cm]{head_foot/header_bar}}\par
-\vspace{1em}
-\sffamily
-\begin{tabular}{m{4.5cm} p{13.5cm} }
-
-\includegraphics{head_foot/DOI} & \noindent\LARGE{\textbf{An upper bound to gas storage and delivery via pressure-swing adsorption in
-porous materials}} \\%Article title goes here instead of the text "This is the title"
-\vspace{0.3cm} & \vspace{0.3cm} \\
-
- & \noindent\large{Jordan K. Pommerenck,$^{\ast}$\textit{$^{a}$} Cory M. Simon,\textit{$^{b}$} and David Roundy\textit{$^{a}$}} \\%Author names go here instead of "Full name", etc.
-
-\includegraphics{head_foot/dates} & \noindent\normalsize{
+\author{Jordan Pommerenck}
+\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\author{Cory Simon}
+\affiliation{Department of Chemical Engineering, Oregon State University, Corvallis, OR 97331}
+\author{David Roundy}
+\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+
+%  & \noindent\large{Jordan K. Pommerenck,$^{\ast}$\textit{$^{a}$} Cory M. Simon,\textit{$^{b}$} and David Roundy\textit{$^{a}$}} \\%Author names go here instead of "Full name", etc.
+\begin{abstract}
     Both hydrogen
@@
     structure in response to adsorbed gas, suggesting that flexible materials could
-still satisfy the DOE targets.}
-
-%Any references in the abstract should be written out in full \textit{e.g.}\ [Surname \textit{et al., Journal Title}, 2000, \textbf{35}, 3523].} \\%The abstrast goes here instead of the text "The abstract should be..."
-
-\end{tabular}
-
- \end{@twocolumnfalse} \vspace{0.6cm}
-
-  ]
-%%%END OF TITLE, AUTHORS AND ABSTRACT%%%
-
-%%%FONT SETUP - please do not change any commands within this section
-\renewcommand*\rmdefault{bch}\normalfont\upshape
-\rmfamily
-\section*{}
-\vspace{-1cm}
+    still satisfy the DOE targets.    
+\end{abstract}
+
+\maketitle
 
 %%%FOOTNOTES%%%
-\footnotetext{\textit{$^{*}$~E-mail: pommerjo@oregonstate.edu}}
-\footnotetext{\textit{$^{a}$~Department of Physics, Oregon State University, Corvallis, OR 97331, USA.}}
-\footnotetext{\textit{$^{b}$~School of Chemical, Biological, and Environmental Engineering, Oregon State University, Corvallis, OR 97331, USA. }}
-
-%Please use \dag to cite the ESI in the main text of the article.
-%If you article does not have ESI please remove the the \dag symbol from the title and the footnotetext below.
-
-%\footnotetext{\dag~Electronic Supplementary Information (ESI) available: [details of any supplementary information available should be included here]. See DOI: 00.0000/00000000.}
-
-%additional addresses can be cited as above using the lower-case letters, c, d, e... If all authors are from the same address, no letter is required
-
-%\footnotetext{\ddag~Additional footnotes to the title and authors can be included \textit{e.g.}\ `Present address:' or `These authors contributed equally to this work' as above using the symbols: \ddag, \textsection, and \P.
-Please place the appropriate symbol next to the author's name and include a \texttt{\textbackslash footnotetext} entry in the the correct place in the list.}
-
-%%%END OF FOOTNOTES%%%
-
-%%%MAIN TEXT%%%%
+% \footnotetext{\textit{$^{*}$~E-mail: pommerjo@oregonstate.edu}}
+% \footnotetext{\textit{$^{a}$~Department of Physics, Oregon State University, Corvallis, OR 97331, USA.}}
+% \footnotetext{\textit{$^{b}$~School of Chemical, Biological, and Environmental Engineering, Oregon State University, Corvallis, OR 97331, USA. }}
+
 \section{Introduction}
@@
 
-%%%END OF MAIN TEXT%%%
-
-%The \balance command can be used to balance the columns on the final page if desired. It should be placed anywhere within the first column of the last page.
-
-\balance
-
-%If notes are included in your references you can change the title from 'References' to 'Notes and references' using the following command:
-%\renewcommand\refname{Notes and references}
-
 %%%REFERENCES%%%
 \bibliography{bibfile} %You need to replace "rsc" on this line with the name of your .bib file
-\bibliographystyle{rsc} %the RSC's .bst file
 

### 2021-10-01 gas-adsorption/gas-ads-revtex.tex
comment out comments to referees

@@
 
-\startreferees{Explain how the point of our paper is not to find structure-property relationships.}
-Several studies elucidated relationships between structural features of nanoporous materials---such as void fraction, pore size, and surface area---and their deliverable capacity of various gases \cite{gomez2014exploring,ahmed2019exceptional,simon2015materials,moghadam2018computer,Tong2018}.
-Such structure-property relationships serve as useful heuristics for designing new materials with high deliverable capacity.
-However, the intention of our paper is not to discover structure-property relationships.
-Rather, we place our upper bound on the deliverable capacity of [rigid] materials that holds irrespective of how the structural features of the material are tuned.
-
-\donereferees
-
-\startreferees{Discuss gravimetric vs volumetric deliverable capacity}
-
-For vehicular gas storage, gravimetric deliverable capacity (amount of gas per mass of material) may be considered as a material performance metric to avoid a heavy adsorbed-gas fuel tank.
-However, the volumetric deliverable capacity (amount of gas per volume of material) dominantly determines the driving range on a full fuel tank of a given volume \cite{mason2014evaluating}.
-That said, several articles have explored the tradeoff between gravimetric and volumetric deliverable capacity, particularly for hydrogen storage \cite{gomez2017understanding,chen2020balancing,goldsmith2013theoretical,ahmed2021predicting}.
-
-\donereferees
+% \startreferees{Explain how the point of our paper is not to find structure-property relationships.}
+Several studies elucidated relationships between structural features of
+nanoporous materials---such as void fraction, pore size, and surface area---and
+their deliverable capacity of various gases
+\cite{gomez2014exploring,ahmed2019exceptional,simon2015materials,moghadam2018computer,Tong2018}.
+Such structure-property relationships serve as useful heuristics for designing
+new materials with high deliverable capacity. However, the intention of our
+paper is not to discover structure-property relationships. Rather, we place our
+upper bound on the deliverable capacity of [rigid] materials that holds
+irrespective of how the structural features of the material are tuned.
+
+% \donereferees
+
+% \startreferees{Discuss gravimetric vs volumetric deliverable capacity}
+
+For vehicular gas storage, gravimetric deliverable capacity (amount of gas per
+mass of material) may be considered as a material performance metric to avoid a
+heavy adsorbed-gas fuel tank. However, the volumetric deliverable capacity
+(amount of gas per volume of material) dominantly determines the driving range
+on a full fuel tank of a given volume \cite{mason2014evaluating}. That said,
+several articles have explored the tradeoff between gravimetric and volumetric
+deliverable capacity, particularly for hydrogen storage
+\cite{gomez2017understanding,chen2020balancing,goldsmith2013theoretical,ahmed2021predicting}.
+
+% \donereferees
 

### 2022-03-10 gas-adsorption/gas-ads-prresearch.tex
add upload to arxiv
start working on referee response

--- 2021-10-01 gas-adsorption/gas-ads-revtex.tex
@@
-\documentclass[pre,twocolumn]{revtex4-2}
+\documentclass[aps,prl,twocolumn]{revtex4-2}
 
@@
 \affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+
+\title{An upper bound to gas storage and delivery via pressure-swing adsorption in
+porous materials}
 
@@
 gas~\cite{makal2012methane,mason2014evaluating} and
-hydrogen~\cite{suh2011hydrogen,garcia2018benchmark} for vehicular storage at
+hydrogen~\cite{suh2011hydrogen,garcia2018benchmark, record-high-hydrogen-storage-2018} for vehicular storage at
 room temperature is through physical adsorption in nanoporous
@@
     \includegraphics[width=0.95\columnwidth]{hydrogen-298-n-vs-G}
-    \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.
-Experimental deliverable capacities for several porous materials (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
+    \caption{Deliverable capacity of hydrogen at room temperature as a function
+    of the attractive Gibbs free energy $|\gst|$.  Experimental deliverable
+    capacities for several porous materials (data from
+    Ref.~\cite{garcia2018benchmark, record-high-hydrogen-storage-2018}) are shown along with the experimental
+    values for $\gst$ at the empty and full pressures shown as $+$'s connected
+    by a line.}
     \label{fig:hydrogen-298-D}
@@
 
-\section*{Conflicts of interest}
-There are no conflicts to declare.
-
 %\section*{Acknowledgements}

### 2022-03-10 gas-adsorption/colors.py

@@ -16,6 +16,7 @@ print(kelly_colors)
 colors = kelly_colors # list(prop_cycle.by_key()['color'])
 
 the_order = [
+    'Ni2m-dobdc',
     'MOF519',
     'NJUBai43',
     'NU1000',
@@ -61,6 +62,7 @@ def symbol(gas):
         return '+' 
 
 latex_mof_names = {
+    'Ni2m-dobdc': 'Ni$_2$($m$-dodbc)',
     'Ni2dobdc': 'Ni$_2$(dodbc)',
     'Co2dobdc': 'Co$_2$(dodbc)',
     'Mg2dobdc':'Mg$_2$(dodbc)',

### 2022-03-10 gas-adsorption/data_scripts/hydrogen.py

@@ -80,6 +80,10 @@ H2_298_Data = {
         'rho_full': 5.29*mmol/gram,
         'rho_empty': 0.31*mmol/gram,
     },
+    'Ni2m-dobdc': {
+        'rho_full': 11.9*gram/molecular_weight/L,
+        'rho_empty': 0.9*gram/molecular_weight/L,
+    },
 }
 
 H2_77_Data = { # from Table S4. In the supplemental.

### 2022-03-10 gas-adsorption/isothermal-ads-gas-cartoon.py

@@ -193,6 +193,9 @@ crystal_density = {
     'NU1102':           0.403*gram/cm**3,
     'NU1103':           0.298*gram/cm**3,
     'COF102':           0.41*gram/cm**3,
+    # WARNING, the following is false, and is intended to compensate for rho
+    # being given in mmol/L not mmol/g
+    'Ni2m-dobdc':       1,
 }
 
 mof_isotherms = gas.isotherm_experiments(T[0], 5.8, 65)
@@ -209,9 +212,10 @@ for mof in colors.order(mof_isotherms): # For each MOF
         rho_hi_p = mof_isotherms[mof]['rho_full']*crystal_density[mof]
     delta_G_hi_p = np.interp(rho_hi_p, n, mu) - mu_full
 
-    plt.plot([delta_G_lo_p/(kJ/mol), delta_G_hi_p/(kJ/mol)],
-            [(rho_hi_p-rho_lo_p)/density_units, (rho_hi_p-rho_lo_p)/density_units],
-            colors.symbol(basename)+'-', label=colors.latex_mof(mof), color=colors.color(mof))
+    if mof in mof_isotherms:
+        plt.plot([delta_G_lo_p/(kJ/mol), delta_G_hi_p/(kJ/mol)],
+                [(rho_hi_p-rho_lo_p)/density_units, (rho_hi_p-rho_lo_p)/density_units],
+                colors.symbol(basename)+'-', label=colors.latex_mof(mof), color=colors.color(mof))
 
 stepby = None
 if 20 < ymax < 200:

### 2022-03-10 gas-adsorption/isothermal-ads-gas-plot.py

@@ -240,6 +240,9 @@ crystal_density = {
     'NU1102':           0.403*gram/cm**3,
     'NU1103':           0.298*gram/cm**3,
     'COF102':           0.41*gram/cm**3,
+    # WARNING, the following is false, and is intended to compensate for rho
+    # being given in mmol/L not mmol/g
+    'Ni2m-dobdc':       1,
 }
 
 mof_isotherms = gas.isotherm_experiments(T[0], float(sys.argv[3]), float(sys.argv[4]))
